### 2022-04-29 cmsb22.tex
LNCS template with authors

\documentclass[runningheads]{llncs}
\usepackage[T1]{fontenc}
% \usepackage{graphicx}
\usepackage{hyperref}
\usepackage{color}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Minimal trap-spaces of Logical models are maximal siphons of their Petri-net encoding}
\titlerunning{Minimal trap-spaces as maximal siphons}
\author{Van-Giang Trinh\inst{1}\orcidID{0000-1111-2222-3333} \and \\
Sylvain Soliman\inst{2}\orcidID{0000-0001-5525-7418}}
% \authorrunning{F. Author et al.}

\institute{
Aix-Marseille Université, Marseille, France\\
\email{???@lis-lab.fr}\\
\url{???} \and
Lifeware team, Inria Saclay center, Palaiseau, France\\
\email{Sylvain.Soliman@inria.fr}\\
\url{http://lifeware.inria.fr/~soliman/}}

\maketitle

\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{Logical models \and Boolean models \and Trap-spaces \and Attractor computation \and Petri-nets \and Siphons.}
\end{abstract}
%
%
%
\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
% \includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.

\subsubsection{Acknowledgments} Please place your acknowledgments at
the end of the paper, preceded by an unnumbered run-in heading (i.e.
3rd-level heading).

\bibliographystyle{splncs04}
\bibliography{cmsb.bib}

\end{document}

### 2022-04-29 cmsb22.tex
Preliminary abstract, and article structure

@@
   \email{Sylvain.Soliman@inria.fr}\\
-\url{http://lifeware.inria.fr/~soliman/}}
+  \url{http://lifeware.inria.fr/~soliman/}
+}
 
@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
+  % The abstract should briefly summarize the contents of the paper in
+  % 150--250 words.
 
-\keywords{Logical models \and Boolean models \and Trap-spaces \and Attractor computation \and Petri-nets \and Siphons.}
+  Discrete modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
+  This is even more true for very big models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
+  The analysis of such models is mostly based on attractor computation, since those correspond roughly to \emph{phenotypes}, and the recent use of trap-spaces made a real breakthrough in that domain allowing to consider medium-sized models that used to be out of reach.
+  However, with the continuing increase in model-size, the state of the art computation of minimal trap-spaces based on prime-implicants shows its limits as there can be very many such implicants.
+  In this article we present an alternative method to compute minimal trap-spaces, and hence complex attractors, that relies on a completely different technique, namely the enumeration of maximal siphons in the Petri-net encoding of the original Logical model.
+  We then demonstrate its efficiency and compare it to implicant-based methods on a few relevant big Boolean models.
+
+\keywords{Logical models \and Boolean models \and Trap-spaces \and Attractor computation \and Petri-nets \and Siphons}
 \end{abstract}
-%
-%
-%
-\section{First Section}
-\subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is
-not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
 
-Subsequent paragraphs, however, are indented.
+\section{Introduction}
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
+Discrete modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
+This is even more true for very big models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
+The analysis of such models is mostly based on attractor computation, since those correspond roughly to \emph{phenotypes}, and the recent use of trap-spaces made a real breakthrough in that domain allowing to consider medium-sized models that used to be out of reach.
+However, with the continuing increase in model-size, the state of the art computation of minimal trap-spaces based on prime-implicants shows its limits as there can be very many such implicants.
+In this article we present an alternative method to compute minimal trap-spaces, and hence complex attractors, that relies on a completely different technique, namely the enumeration of maximal siphons in the Petri-net encoding of the original Logical model.
+We then demonstrate its efficiency and compare it to implicant-based methods on a few relevant big Boolean models.
 
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
+\section{Minimal trap-spaces as maximal siphons}
+\subsection{Preliminaries}
+\subsubsection{Traps-spaces}
+\subsubsection{Petri-net encoding of Logical models}
+\subsubsection{Siphons}
 
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
+\subsection{New method}
 
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
+\section{Evaluation}
+\subsection{Benchmark vs. implicant-based methods}
+\subsection{Detailed biological example}
 
-\begin{figure}
-% \includegraphics[width=\textwidth]{fig1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig1}
-\end{figure}
+\section{Conclusion}
 
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
+% \begin{table}
+% \caption{Table captions should be placed above the
+% tables.}\label{tab1}
+% \begin{tabular}{|l|l|l|}
+% \hline
+% Heading level &  Example & Font size and style\\
+% \hline
+% Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
+% 1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
+% 2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
+% 3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
+% 4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
+% \hline
+% \end{tabular}
+% \end{table}
+
+% \noindent Displayed equations are centered and set on a separate
+% line.
+% \begin{equation}
+% x + y = z
+% \end{equation}
+% Please try to avoid rasterized images for line-art diagrams and
+% schemas. Whenever possible, use vector graphics instead (see
+% Fig.~\ref{fig1}).
+
+% \begin{figure}
+% % \includegraphics[width=\textwidth]{fig1.eps}
+% \caption{A figure caption is always placed below the illustration.
+% Please note that short captions are centered, while long ones are
+% justified by the macro package automatically.} \label{fig1}
+% \end{figure}
+
+% \begin{theorem}
+% This is a sample theorem. The run-in heading is set in bold, while
+% the following text appears in italics. Definitions, lemmas,
+% propositions, and corollaries are styled the same way.
+% \end{theorem}
+% %
+% % the environments 'definition', 'lemma', 'proposition', 'corollary',
+% % 'remark', and 'example' are defined in the LLNCS documentclass as well.
+% %
+% \begin{proof}
+% Proofs, examples, and remarks have the initial word in italics,
+% while the following text appears in normal font.
+% \end{proof}
+% For citations of references, we prefer the use of square brackets
+% and consecutive numbers. Citations using labels or the author/year
+% convention are also acceptable. The following bibliography provides
+% a sample reference list with entries for journal
+% articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
+% book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
+% and a homepage~\cite{ref_url1}. Multiple citations are grouped
+% \cite{ref_article1,ref_lncs1,ref_book1},
+% \cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
 

### 2022-04-29 cmsb22.tex
Header
Some more stuff

@@
 \documentclass[runningheads]{llncs}
-\usepackage[T1]{fontenc}
+\usepackage[T5]{fontenc}
 % \usepackage{graphicx}
@@
 \titlerunning{Minimal trap-spaces as maximal siphons}
-\author{Van-Giang Trinh\inst{1}\orcidID{0000-1111-2222-3333} \and \\
+\author{Van-Giang Trịnh\inst{1}\orcidID{0000-0001-6581-998X} \and \\
   Sylvain Soliman\inst{2}\orcidID{0000-0001-5525-7418}}
@@
   Aix-Marseille Université, Marseille, France\\
-  \email{???@lis-lab.fr}\\
-  \url{???} \and
+  \email{Trinh.Van-Giang@lis-lab.fr}
+  \and
   Lifeware team, Inria Saclay center, Palaiseau, France\\
-  \email{Sylvain.Soliman@inria.fr}\\
-  \url{http://lifeware.inria.fr/~soliman/}
+  \email{Sylvain.Soliman@inria.fr}%\\
+  %\url{http://lifeware.inria.fr/~soliman/}
 }
@@
 \subsubsection{Petri-net encoding of Logical models}
+
+Original encoding~\cite{chaouiya2011petri} vs. more recent one~\cite{chatain2014characterization}
+
 \subsubsection{Siphons}
@@
 \subsection{New method}
+
+SBML-Qual~\cite{chaouiya2013sbml}/BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} to PNML\footnote{\url{https://www.pnml.org/}} works via GINsim/BioLQM~\cite{chaouiya2012logical} with
+\begin{verbatim}
+\verb|java -jar GINsim.jar -lqm <input.sbml/input.bnet> <output.pnml>|
+\end{verbatim}
 
@@
 \section{Conclusion}
+
+\subsubsection{Acknowledgments}
+Coffee is good, you should drink some!
+
+\bibliographystyle{splncs04}
+\bibliography{cmsb22.bib}
+
+\end{document}
 
@@
 % \end{proof}
-% For citations of references, we prefer the use of square brackets
-% and consecutive numbers. Citations using labels or the author/year
-% convention are also acceptable. The following bibliography provides
-% a sample reference list with entries for journal
-% articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
-% book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-% and a homepage~\cite{ref_url1}. Multiple citations are grouped
-% \cite{ref_article1,ref_lncs1,ref_book1},
-% \cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-
-\subsubsection{Acknowledgments} Please place your acknowledgments at
-the end of the paper, preceded by an unnumbered run-in heading (i.e.
-3rd-level heading).
-
-\bibliographystyle{splncs04}
-\bibliography{cmsb.bib}
-
-\end{document}

### 2022-04-29 cmsb22.tex
A little something everywhere

@@
 \subsubsection{Traps-spaces}
+
+\cite{klarner2015computing,klarner2017pyboolnet,cifuentes2020control}
+
 \subsubsection{Petri-net encoding of Logical models}
 
-Original encoding~\cite{chaouiya2011petri} vs. more recent one~\cite{chatain2014characterization}
+Original encoding~\cite{chaouiya2004qualitative,chaouiya2011petri} vs. more recent one~\cite{chatain2014characterization}.
+All identical on Boolean models?
 
@@
 
-SBML-Qual~\cite{chaouiya2013sbml}/BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} to PNML\footnote{\url{https://www.pnml.org/}} works via GINsim/BioLQM~\cite{chaouiya2012logical} with
+SBML-Qual~\cite{chaouiya2013sbml}/BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} to PNML\footnote{\url{https://www.pnml.org/}} works via GINsim~\cite{chaouiya2012logical}/BioLQM\footnote{http://www.colomoto.org/biolqm/} with
 \begin{verbatim}
-\verb|java -jar GINsim.jar -lqm <input.sbml/input.bnet> <output.pnml>|
+java -jar GINsim.jar -lqm <input.sbml/input.bnet> <output.pnml>
 \end{verbatim}
@@
 \subsection{Benchmark vs. implicant-based methods}
+
+\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}
+
 \subsection{Detailed biological example}
+
+TBD
 

### 2022-04-30 trappist.py
Preliminary PNML to NetworkX parser…

@@ -1,0 +1,103 @@
+"""Compute minimal trap-spaces of a Petri-net encoded Boolean model.
+
+Copyright (C) 2022 Sylvain.Soliman@inria.fr and giang.trinh91@gmail.com
+
+This program is free software: you can redistribute it and/or modify
+it under the terms of the GNU General Public License as published by
+the Free Software Foundation, either version 3 of the License, or
+(at your option) any later version.
+
+This program is distributed in the hope that it will be useful,
+but WITHOUT ANY WARRANTY; without even the implied warranty of
+MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License
+along with this program.  If not, see <https://www.gnu.org/licenses/>.
+"""
+
+import argparse
+import subprocess
+import sys
+import tempfile
+import xml.etree.ElementTree as etree
+from typing import IO
+
+import networkx as nx
+
+
+version = "0.1.0"
+
+
+def read_pnml(fileobj: IO) -> nx.DiGraph:
+    """Parse the given file."""
+    root = etree.parse(fileobj).getroot()
+    if root.tag != "pnml":
+        raise ValueError("Currently limited to parsing PNML files")
+    net = nx.DiGraph()
+
+    for place in root.findall("./net/place"):
+        net.add_node(place.get("id"), name=place.find("./name/value").text)
+        # print(place.get("id"), place.find("./name/value").text)
+
+    for transition in root.findall("./net/transition"):
+        net.add_node(transition.get("id"))
+
+    for arc in root.findall("./net/arc"):
+        net.add_edge(arc.get("source"), arc.get("target"))
+
+    print(net.number_of_nodes(), net.size())
+    return net
+
+
+def write_asp(petri_net: nx.DiGraph, asp_file: IO):
+    """Write the ASP program for the maximal conflict-free siphons of petri_net."""
+    pass
+
+
+def solve_asp(asp_filename: str):
+    """Run an ASP solver on program asp_file and get the solutions."""
+    result = subprocess.run(
+        [
+            "echo",
+            "clingo",
+            "--dom-pref=32",
+            "--heu=domain",
+            "--dom-mod=7",
+            asp_filename,
+            "0",
+        ],
+        capture_output=True,
+    )
+    print(result.stdout)
+
+
+def main():
[…]

### 2022-04-30 trappist.py
Generating an ASP file

@@ -37,38 +37,49 @@ def read_pnml(fileobj: IO) -> nx.DiGraph:
     net = nx.DiGraph()
 
     for place in root.findall("./net/place"):
-        net.add_node(place.get("id"), name=place.find("./name/value").text)
+        net.add_node(
+            place.get("id"), kind="place", name=place.find("./name/value").text
+        )
         # print(place.get("id"), place.find("./name/value").text)
 
     for transition in root.findall("./net/transition"):
-        net.add_node(transition.get("id"))
+        net.add_node(transition.get("id"), kind="transition")
 
     for arc in root.findall("./net/arc"):
         net.add_edge(arc.get("source"), arc.get("target"))
 
-    print(net.number_of_nodes(), net.size())
     return net
 
 
 def write_asp(petri_net: nx.DiGraph, asp_file: IO):
     """Write the ASP program for the maximal conflict-free siphons of petri_net."""
-    pass
+    for node, kind in petri_net.nodes(data="kind"):
+        if kind == "place":
+            print("{", node, "}", file=asp_file)
+            if not node.startswith("-"):
+                print(f":- {node}, -{node}.", file=asp_file)    # conflict-freeness
+        else:   # it's a transition, apply siphon (if one succ is true, one pred must be true)
+            preds = list(petri_net.predecessors(node))
+            or_preds = "; ".join(preds)
+            for succ in petri_net.successors(node):
+                if succ not in preds:   # optimize obvious tautologies
+                    print(f"{or_preds} :- {succ}", file=asp_file)
 
 
 def solve_asp(asp_filename: str):
     """Run an ASP solver on program asp_file and get the solutions."""
-    result = subprocess.run(
-        [
-            "echo",
-            "clingo",
-            "--dom-pref=32",
-            "--heu=domain",
-            "--dom-mod=7",
-            asp_filename,
-            "0",
-        ],
-        capture_output=True,
-    )
+    # result = subprocess.run(
+    #     [
+    #         "clingo",
+    #         "--dom-pref=32",
+    #         "--heu=domain",
+    #         "--dom-mod=7",
+    #         asp_filename,
+    #         "0",
+    #     ],
+    #     capture_output=True,
+    # )
+    result = subprocess.run(["cat", asp_filename])
     print(result.stdout)
 
 
@@ -94,7 +105,7 @@ def main():
 
     petri_net = read_pnml(args.infile)
 
[…]

### 2022-04-30 trappist.py
Working ASP program

@@ -17,6 +17,7 @@ along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 
 import argparse
+import os
 import subprocess
 import sys
 import tempfile
@@ -40,7 +41,6 @@ def read_pnml(fileobj: IO) -> nx.DiGraph:
         net.add_node(
             place.get("id"), kind="place", name=place.find("./name/value").text
         )
-        # print(place.get("id"), place.find("./name/value").text)
 
     for transition in root.findall("./net/transition"):
         net.add_node(transition.get("id"), kind="transition")
@@ -51,36 +51,51 @@ def read_pnml(fileobj: IO) -> nx.DiGraph:
     return net
 
 
+def pnml_to_asp(name: str) -> str:
+    if name.startswith("-"):
+        return "n" + name[1:]
+    return "p" + name
+
+
 def write_asp(petri_net: nx.DiGraph, asp_file: IO):
     """Write the ASP program for the maximal conflict-free siphons of petri_net."""
     for node, kind in petri_net.nodes(data="kind"):
         if kind == "place":
-            print("{", node, "}", file=asp_file)
+            print("{", pnml_to_asp(node), "}.", file=asp_file, sep="")
             if not node.startswith("-"):
-                print(f":- {node}, -{node}.", file=asp_file)    # conflict-freeness
-        else:   # it's a transition, apply siphon (if one succ is true, one pred must be true)
+                print(
+                    f":- {pnml_to_asp(node)}, {pnml_to_asp('-' + node)}.", file=asp_file
+                )  # conflict-freeness
+        else:  # it's a transition, apply siphon (if one succ is true, one pred must be true)
             preds = list(petri_net.predecessors(node))
-            or_preds = "; ".join(preds)
+            or_preds = "; ".join(map(pnml_to_asp, preds))
             for succ in petri_net.successors(node):
-                if succ not in preds:   # optimize obvious tautologies
-                    print(f"{or_preds} :- {succ}", file=asp_file)
+                if succ not in preds:  # optimize obvious tautologies
+                    print(f"{or_preds} :- {pnml_to_asp(succ)}.", file=asp_file)
 
 
-def solve_asp(asp_filename: str):
+def solve_asp(asp_filename: str) -> str:
     """Run an ASP solver on program asp_file and get the solutions."""
-    # result = subprocess.run(
-    #     [
[…]

### 2022-05-01 trappist.py
Parsing clingo JSON output

@@ -17,14 +17,15 @@ along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 
 import argparse
+import json
 import os
 import subprocess
 import sys
 import tempfile
 import xml.etree.ElementTree as etree
-from typing import IO
+from typing import IO, List
 
-import networkx as nx
+import networkx as nx   # TODO maybe replace with lists/dicts
 
 
 version = "0.1.0"
@@ -80,22 +81,47 @@ def solve_asp(asp_filename: str) -> str:
         [
             "clingo",
             "0",
-            "--heuristic=Domain",
+            "--heuristic=Domain",   # maximal w.r.t. inclusion
             "--enum-mod=domRec",
             "--dom-mod=3",
+            "--outf=2",             # json output
             asp_filename,
         ],
         capture_output=True,
         text=True,
+        # check=True,
     )
-    # print(result.stderr)
+    if result.returncode != 30:     # https://www.mat.unical.it/aspcomp2013/files/aspoutput.txt
+        print(result.stderr)        # 30: SAT, all enumerated, optima found
 
     return result.stdout
 
 
-def write_solutions(asp_output: str):
+def solution_to_bool(places: List[str], sol: List[str]) -> List[str]:
+    """Convert a list of present places in sol, to a tri-valued vector."""
+    return([place_in_sol(sol, p) for p in places])
+
+
+def place_in_sol(sol: List[str], place: str) -> str:
+    """Return 0/1/- if place is absent, present or does not appear in sol."""
+    if "p" + place in sol:
+        return "1"
+    if "n" + place in sol:
+        return "0"
+    return "-"
+
+
+def write_solutions(asp_output: str, petri_net: nx.DiGraph):
     """Display the ASP output back as trap-spaces."""
-    print(asp_output)
+    places = []
+    for node, kind in petri_net.nodes(data="kind"):
+        if kind == "place" and not node.startswith("-"):
+            places.append(node)
+    print(" ".join(places))
+    solutions = json.loads(asp_output)
+    for sol in solutions["Call"][0]["Witnesses"]:
+        print(" ".join(solution_to_bool(places, sol["Value"])))
+    print("Total time:", solutions["Time"]["Total"], "s")
 
 
 def main():
@@ -125,7 +151,7 @@ def main():
         write_asp(petri_net, asp_file)
     solutions = solve_asp(tmpname)
     os.unlink(tmpname)
-    write_solutions(solutions)
+    write_solutions(solutions, petri_net)
 
 
 if __name__ == "__main__":

### 2022-05-01 trappist.py
Cleaner error handling

@@ -91,8 +91,10 @@ def solve_asp(asp_filename: str) -> str:
         text=True,
         # check=True,
     )
-    if result.returncode != 30:     # https://www.mat.unical.it/aspcomp2013/files/aspoutput.txt
-        print(result.stderr)        # 30: SAT, all enumerated, optima found
+
+    # https://www.mat.unical.it/aspcomp2013/files/aspoutput.txt
+    if result.returncode != 30:     # 30: SAT, all enumerated, optima found
+        result.check_returncode()   # will raise CalledProcessError
 
     return result.stdout
 

### 2022-05-01 cmsb22.tex
Comments, title
Getting some text in

@@
 \usepackage{hyperref}
-\usepackage{color}
+\usepackage{xcolor}
 \renewcommand\UrlFont{\color{blue}\rmfamily}
+
+\newcommand{\vangiang}[1]{\textcolor{magenta}{#1}}
+\newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
 
@@
 
+\sylvain{We could go for a completely different title, e.g. focused on performance much more, like ``Scaling-up attractor computation for Logical models''}
+
+\cite{glass1973logical,thomas1973boolean,thomas1990biological}
+
 Discrete modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
@@
 
+Purely boolean, asynchronous.
+
 \cite{klarner2015computing,klarner2017pyboolnet,cifuentes2020control}
@@
 \subsubsection{Petri-net encoding of Logical models}
+\label{sec:encoding}
 
-Original encoding~\cite{chaouiya2004qualitative,chaouiya2011petri} vs. more recent one~\cite{chatain2014characterization}.
-All identical on Boolean models?
+The link between Logical models \emph{à la} Thomas and Petri-nets was originally established in~\cite{chaouiya2004qualitative} in order to make available formal methods like model-checking for the analysis of such systems.
+The basic encoding only holds for purely Boolean models but was later extended to multi-valued models in two ways, either in \cite{chaouiya2011petri} with non 1-safe Petri-nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.
+
+Since our study is focused on Boolean models, we will only briefly recall the original encoding here.
+Its basis is that every \emph{gene} of the original model is represented by two separate places, corresponding to its two states, active, and inactive.
+Each conjunct of the logical function that activate the \emph{gene} will lead to a transition, consuming the inactive place, producing the active place, and with all other literals both consumed and produced.
+And conversely for the inactivation.
+
+Note that given a Logical model in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., one of the packages of SBML v3~\cite{keating2020sbml}, one can easily obtain its Petri-net encoding in the PNML\footnote{\url{https://www.pnml.org/}} standard using the BioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
+This piece of software extracted from GINsim~\cite{chaouiya2012logical} and part of the CoLoMoTo\footnote{\url{http://colomoto.org/}} software suite allows for easy conversion between standard formats.
+It also accepts many other common formats for Logical models, notably the \verb|.bnet| files of the  BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} tools.
+
+Conversion is executed as follows:
+
+\noindent{\small \verb|java -jar GINsim.jar -lqm <input.{sbml,bnet,zginml,...}> <output.pnml>|}
 
@@
 
+Siphons are a classical property of place-transition nets~\cite{peterson81petri}.
+Note however that the use of siphons for the analysis of biological models, though it is not new~\cite{angeli2007petri,angeli2011persistence,degrand2020graphical}, has been mostly relevant to the ODE-based continuous semantics of Chemical Reaction Networks.
+
+We recall here the basic definition:
+
+\begin{definition}
+
+  A siphon is a set of places such that for each edge from a transition to any place of the siphon, there is an edge from a place of the siphon to that transition.
+
+\end{definition}
+
+Intuitively a siphon is a set of places that once empty remains empty, i.e.,
+a set of species that cannot be produced again once they have been completely
+consumed.
+
 \subsection{New method}
-
-SBML-Qual~\cite{chaouiya2013sbml}/BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} to PNML\footnote{\url{https://www.pnml.org/}} works via GINsim~\cite{chaouiya2012logical}/BioLQM\footnote{http://www.colomoto.org/biolqm/} with
-\begin{verbatim}
-java -jar GINsim.jar -lqm <input.sbml/input.bnet> <output.pnml>
-\end{verbatim}
 
@@
 \subsection{Benchmark vs. implicant-based methods}
+
+Using BioLQM~\ref{sec:encoding}
 

### 2022-05-01 cmsb22.tex
More contents, especially intro, updated structure

--- 2022-04-29 cmsb22.tex
@@
 \usepackage{hyperref}
-\usepackage{color}
+\usepackage{xcolor}
 \renewcommand\UrlFont{\color{blue}\rmfamily}
+
+\newcommand{\vangiang}[1]{\textcolor{magenta}{#1}}
+\newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
+
+\newcommand{\bbbb}{{\rm I\!B}}
 
@@
 %
-\title{Minimal trap-spaces of Logical models are maximal siphons of their Petri-net encoding}
+\title{Minimal trap-spaces of Logical models are maximal siphons of their Petri net encoding}
 \titlerunning{Minimal trap-spaces as maximal siphons}
@@
 
-  Discrete modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
+  Boolean modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
   This is even more true for very big models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
@@
   However, with the continuing increase in model-size, the state of the art computation of minimal trap-spaces based on prime-implicants shows its limits as there can be very many such implicants.
-  In this article we present an alternative method to compute minimal trap-spaces, and hence complex attractors, that relies on a completely different technique, namely the enumeration of maximal siphons in the Petri-net encoding of the original Logical model.
+  In this article we present an alternative method to compute minimal trap-spaces, and hence complex attractors, that relies on a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original Logical model.
   We then demonstrate its efficiency and compare it to implicant-based methods on a few relevant big Boolean models.
 
-\keywords{Logical models \and Boolean models \and Trap-spaces \and Attractor computation \and Petri-nets \and Siphons}
+\keywords{Logical models \and Boolean models \and Trap-spaces \and Attractor computation \and Petri nets \and Siphons}
 \end{abstract}
@@
 
-Discrete modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
-This is even more true for very big models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
-The analysis of such models is mostly based on attractor computation, since those correspond roughly to \emph{phenotypes}, and the recent use of trap-spaces made a real breakthrough in that domain allowing to consider medium-sized models that used to be out of reach.
-However, with the continuing increase in model-size, the state of the art computation of minimal trap-spaces based on prime-implicants shows its limits as there can be very many such implicants.
-In this article we present an alternative method to compute minimal trap-spaces, and hence complex attractors, that relies on a completely different technique, namely the enumeration of maximal siphons in the Petri-net encoding of the original Logical model.
+\sylvain{We could go for a completely different title, e.g., focused on performance much more, like ``Scaling-up attractor computation for Logical models''}
+
+From the observation that the transcriptional regulation behaved in a sigmoid step-like way, came the original idea to represent models of gene regulation as discrete event systems.
+Those Gene Regulation Networks (GRN) use thresholds or equivalently logical functions to represent the different regulations~\cite{glass1973logical,thomas1973boolean,thomas1990biological,thomas1991regulatory}.
+
+Boolean modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}.
+This is even more true for very big models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas~\cite{aghamiri2020automated}.
+The analysis of such models is mostly based on attractor computation, since those correspond roughly to \emph{phenotypes}, and the recent use of trap-spaces~\cite{klarner2015computing} made a real breakthrough in that domain allowing to consider medium-sized models that used to be out of reach.
+However, with the continuing increase in model-size, the state of the art computation of minimal trap-spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of such implicants.
+
+Petri nets were introduced in the 60’s as a simple formalism for describing and analyzing information processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri}.
+The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, was introduced quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
+Siphons are such a concept but they have not been used a lot for the analysis of biochemical systems even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{nabli2016enumerating}.
+
+In this article we present an alternative method to compute minimal trap-spaces, and hence complex attractors, that replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original Logical model.
 We then demonstrate its efficiency and compare it to implicant-based methods on a few relevant big Boolean models.
 
+\section{Preliminaries}
+
+TBD.
+Purely Boolean for trap-spaces, asynchronous for encoding.
+
+\subsection{Traps-spaces}
+
+We recall here some definitions from~\cite{klarner2015computing} for the introduction of \emph{trap-spaces}.
+Minimal trap-spaces prove to be a very good approximation of the attractors of a Boolean model and have become the \emph{de facto} standard way to analyze models of a few tens of \emph{genes}~\cite{klarner2017pyboolnet,cifuentes2020control}.
+
+Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S} =\bbbb^{n}\) with \(\bbbb = \{0, 1\}\).
+
+A \emph{partial state} \(P\) of \(\mathcal{M}\) is a subset of \(\mathcal{S}\) where for some \(U\subset V\) the values are fixed by a mapping \(p: U\mapsto\bbbb\) and the other variables (\(V\setminus U\)) are \emph{free}.
+
+\begin{definition}
+
+  A \emph{trap space} of \(\mathcal{M}\) is a partial state \(P\) that is also a \emph{trap set}:
+
+  if \(P'\) is the image of \(P\) by the state transition function, then \(P'\subset P\).
+
+\end{definition}
+
+Here and in the following, the notion of \emph{minimality/maximality} will always be w.r.t. to subset inclusion.
+
+\subsection{Petri net encoding of Logical models}
+\label{sec:encoding}
+
+\begin{definition}
+
+  A \emph{Petri net graph} is a weighted bipartite directed graph \((P, T, W)\),
+  where \(P\) is a non-empty finite set of vertices called \emph{places},
+  \(T\) is a non-empty finite set of vertices called \emph{transitions},
+  \(P \cap T = \emptyset\),
+  and \(W : (P \times T) \cup (T \times P) \mapsto \bbbn \) is a weight function attached to the arcs.
+
+\end{definition}
+
+We will write \(pred(x)\) (resp.\ \(succ(x)\)) to represent the set of vertices that have an (non-zero weighted) arc leading to (resp.\ coming from) \(x\).
+
+The link between Logical models \emph{à la} Thomas and Petri nets was originally established in~\cite{chaouiya2004qualitative} in order to make available formal methods like model-checking for the analysis of such systems.
+The basic encoding into 1-safe (i.e., never more than one token in each place) nets only holds for purely Boolean models but was later extended to multi-valued models in two ways, either in \cite{chaouiya2011petri} with non 1-safe Petri nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.
+
+Since our study is focused on Boolean models, we briefly recall the original encoding here.
+Its basis is that every \emph{gene} of the original model is represented by two separate places, corresponding to its two states, active, and inactive.
+Each conjunct of the logical function that activate the \emph{gene} will lead to a transition, consuming the inactive place, producing the active place, and with all other literals both consumed and produced.
+And conversely for the inactivation.
+
+The main property of this encoding is that it is completely faithful w.r.t.\ the asynchronous semantics.
+Asynchronous transitions of the original model correspond one-to-one with firings of transitions in the Petri net.
+
+Note that given a Logical model in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., one of the packages of SBML v3~\cite{keating2020sbml}, one can easily obtain its Petri net encoding in the PNML\footnote{\url{https://www.pnml.org/}} standard using the BioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
+This piece of software extracted from GINsim~\cite{chaouiya2012logical} and part of the CoLoMoTo\footnote{\url{http://colomoto.org/}} software suite allows for easy conversion between standard formats.
+It also accepts many other common formats for Logical models, notably the \verb|.bnet| files of the  BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} tools.
+
+Conversion is executed as follows:
+
+\noindent{\small \verb|java -jar GINsim.jar -lqm <input.{sbml,bnet,zginml,...}> <output.pnml>|}
+
+\subsection{Siphons}
+
+Siphons are a classical property of place-transition nets~\cite{peterson1981petri}.
+Note however that the use of siphons for the analysis of biological models, though it is not new, has been mostly relevant to the ODE-based continuous semantics of Chemical Reaction Networks~\cite{angeli2007petri,angeli2011persistence,degrand2020graphical}.
+
+We recall here the basic definition establishing that to produce something in a siphon you must consume something from the siphon:
+
+\begin{definition}
+
+  A \emph{siphon} of a Petri net \((P, T, W)\) is a set of places \(S\) such that:
+  \[\forall t\in T, S\cap succ(t)\not =\emptyset\Rightarrow S\cap pred(t)\not =\emptyset\]
+
+\end{definition}
+
+Intuitively a siphon is a set of places that once empty remains empty.
+
+We add a definition related to any set of places of a Petri net encoding a Logical model, and notably a siphon of such a net.
+
+\begin{definition}
+
+  A set of places of Petri net \(P\) encoding logical model \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{gene} of \(\mathcal{M}\).
+
+\end{definition}
+
 \section{Minimal trap-spaces as maximal siphons}
-\subsection{Preliminaries}
-\subsubsection{Traps-spaces}
-\subsubsection{Petri-net encoding of Logical models}
 
-Original encoding~\cite{chaouiya2011petri} vs. more recent one~\cite{chatain2014characterization}
+\begin{theorem}
 
-\subsubsection{Siphons}
+  \(S\) is a minimal trap-space of model \(\mathcal{M}\) iff the places corresponding to the fixed values in \(S\) (active or inactive) are a maximal conflict-free siphon of \((P, T, W)\), the Petri net encoding of \(S\).
 
-\subsection{New method}
+\end{theorem}
 
-SBML-Qual~\cite{chaouiya2013sbml}/BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} to PNML\footnote{\url{https://www.pnml.org/}} works via GINsim/BioLQM~\cite{chaouiya2012logical} with
-\begin{verbatim}
-\verb|java -jar GINsim.jar -lqm <input.sbml/input.bnet> <output.pnml>|
-\end{verbatim}
+Intuitively, a siphon is a set of places that once empty remains so.
+If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the empty places remain empty).
+This is precisely the definition of a trap-space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap-space.
+
+\begin{proof}
+
+  TBD\qed
+
+\end{proof}
+
+\section{Implementation}
+
+ASP
 
 \section{Evaluation}
-\subsection{Benchmark vs. implicant-based methods}
+\subsection{Benchmark vs.\ implicant-based methods}
+
+Using BioLQM~\ref{sec:encoding}
+
+\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}
+
 \subsection{Detailed biological example}
 
+TBD
+
 \section{Conclusion}
+
+Future work: comparison with SAT/MaxSAT based approaches~\cite{nabli2016enumerating}
 
@@
 % \end{figure}
-
-% \begin{theorem}
-% This is a sample theorem. The run-in heading is set in bold, while
-% the following text appears in italics. Definitions, lemmas,
-% propositions, and corollaries are styled the same way.
-% \end{theorem}
-% %
-% % the environments 'definition', 'lemma', 'proposition', 'corollary',
-% % 'remark', and 'example' are defined in the LLNCS documentclass as well.
-% %
-% \begin{proof}
-% Proofs, examples, and remarks have the initial word in italics,
-% while the following text appears in normal font.
-% \end{proof}

### 2022-05-01 trappist.py
flake8 and black

@@ -25,7 +25,7 @@ import tempfile
 import xml.etree.ElementTree as etree
 from typing import IO, List
 
-import networkx as nx   # TODO maybe replace with lists/dicts
+import networkx as nx  # TODO maybe replace with lists/dicts
 
 
 version = "0.1.0"
@@ -53,6 +53,8 @@ def read_pnml(fileobj: IO) -> nx.DiGraph:
 
 
 def pnml_to_asp(name: str) -> str:
+    """Convert a PNML id to an ASP variable."""
+    # TODO handle non-accetable chars
     if name.startswith("-"):
         return "n" + name[1:]
     return "p" + name
@@ -81,10 +83,10 @@ def solve_asp(asp_filename: str) -> str:
         [
             "clingo",
             "0",
-            "--heuristic=Domain",   # maximal w.r.t. inclusion
+            "--heuristic=Domain",  # maximal w.r.t. inclusion
             "--enum-mod=domRec",
             "--dom-mod=3",
-            "--outf=2",             # json output
+            "--outf=2",  # json output
             asp_filename,
         ],
         capture_output=True,
@@ -93,15 +95,15 @@ def solve_asp(asp_filename: str) -> str:
     )
 
     # https://www.mat.unical.it/aspcomp2013/files/aspoutput.txt
-    if result.returncode != 30:     # 30: SAT, all enumerated, optima found
-        result.check_returncode()   # will raise CalledProcessError
+    if result.returncode != 30:  # 30: SAT, all enumerated, optima found
+        result.check_returncode()  # will raise CalledProcessError
 
     return result.stdout
 
 
 def solution_to_bool(places: List[str], sol: List[str]) -> List[str]:
     """Convert a list of present places in sol, to a tri-valued vector."""
-    return([place_in_sol(sol, p) for p in places])
+    return [place_in_sol(sol, p) for p in places]
 
 
 def place_in_sol(sol: List[str], place: str) -> str:

### 2022-05-02 trappist.py
Place in siphon means opposite value is true

@@ -107,11 +107,14 @@ def solution_to_bool(places: List[str], sol: List[str]) -> List[str]:
 
 
 def place_in_sol(sol: List[str], place: str) -> str:
-    """Return 0/1/- if place is absent, present or does not appear in sol."""
+    """Return 0/1/- if place is absent, present or does not appear in sol.
+
+    Remember that being in the siphon means staying empty, so the opposite value is the one fixed.
+    """
     if "p" + place in sol:
-        return "1"
-    if "n" + place in sol:
         return "0"
+    if "n" + place in sol:
+        return "1"
     return "-"
 
 

### 2022-05-02 cmsb22.tex

@@
 \usepackage{xcolor}
+
 \renewcommand\UrlFont{\color{blue}\rmfamily}
@@
 \titlerunning{Minimal trap-spaces as maximal siphons}
-\author{Van-Giang Trịnh\inst{1}\orcidID{0000-0001-6581-998X} \and \\
+\author{Van-Giang Trinh\inst{1}\orcidID{0000-0001-6581-998X} \and \\
   Sylvain Soliman\inst{2}\orcidID{0000-0001-5525-7418}}
@@
 
+%%% Subspace %%%
+A subspace \(m\) of \(S_{\mathcal{M}}\) is characterized by its fixed variables (denoted by \(D_m\)) and free variables.
+This subspace can be specified by an assignment \(m : D_m \mapsto \bbbb\) where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
+The remaining variables of \(\mathcal{M}\), \(V \setminus D_m\), are said to be free, i.e., they can receive any Boolean value.
+We specify subspaces like states but use in addition the symbol \(\star\) to indicate that a variable is free.
+In addition, a subspace \(m\) also refers to the set of states \(\mathcal{S}_{\mathcal{M}}[m] := \{s \in \mathcal{S}_{\mathcal{M}}\;|\;\forall x_i \in D_m : s_i = m(x_i)\}\).
+For example, \(m = \star\star1\) means that \(D_m = \{x_3\}\), \(m(x_3) = 1\), and refers to the set of states \(\{001, 011, 101, 111\}\).
+Let \(\mathcal{S}_{\mathcal{M}}^{\star}\) denote the set of all possible subspaces of \(\mathcal{M}\).
+Note that \(\left|\mathcal{S}_{\mathcal{M}}^{\star}\right| = 3^n\) and \(S_{\mathcal{M}} \subset \mathcal{S}_{\mathcal{M}}^{\star}\)~\cite{klarner2015computing}.
+
+%%% Minimal trap space and example %%%
+A \emph{trap space} is defined as a subspace that is also a trap set.
+Then, we define a partial order \(<\) on \(\mathcal{S}_{\mathcal{M}}^{\star}\) as: \(m < m'\) if and only if \(\mathcal{S}_{\mathcal{M}}[m] \subseteq \mathcal{S}_{\mathcal{M}}[m']\) and \(\mathcal{S}_{\mathcal{M}}[m] \neq \mathcal{S}_{\mathcal{M}}[m']\).
+Consequently, a trap space \(m\) is minimal if and only if there is no trap space \(m' \in \mathcal{S}_{\mathcal{M}}^{\star}\) such that \(m' < m\).
+For example, the BN shown in Example~\ref{example:BN} has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
+Since \(m_1 < m_2\), \(m_1\) is a minimal trap space of the BN.
+
 \subsection{Petri net encoding of Logical models}
@@
 
-\section{Minimal trap-spaces as maximal siphons}
+\section{Minimal trap spaces as maximal conflict-free siphons}
+
+\begin{definition}
+
+  Let \(m\) be a subspace of Boolean model \(\mathcal{M}\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) such that:
+  \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S\]
+
+\end{definition}
 
@@
 
-  \(S\) is a minimal trap-space of model \(\mathcal{M}\) iff the places corresponding to the fixed values in \(S\) (active or inactive) are a maximal conflict-free siphon of \((P, T, W)\), the Petri net encoding of \(S\).
+  Let \(\mathcal{M}\) be a Boolean model and \(\mathcal{P}\) be its Petri net encoding.
+A subspace \(m\) is a trap space of \(\mathcal{M}\) if and only if its mirror \(S\) is a conflict-free siphon of \(\mathcal{P}\).
+
+\end{theorem}
+
+\begin{proof}
+
+  TBD\qed
+
+\end{proof}
+
+\begin{theorem}
+\label{theo:min_ts_max_sp}
+  %\(S\) is a minimal trap-space of Boolean model \(\mathcal{M}\) if and only if the places corresponding to the fixed values in \(S\) (active or inactive) are a maximal conflict-free siphon of \((P, T, W)\), the Petri net encoding of \(S\).
+
+  Let \(\mathcal{M}\) be a Boolean model and \(\mathcal{P}\) be its Petri net encoding.
+A subspace \(m\) is a minimal trap space of \(\mathcal{M}\) if and only if its mirror \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\).
 
@@
 
-\section{Implementation}
+By Theorem~\ref{theo:min_ts_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean model to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
+It is noted that there are no existing methods specifically designed for computing maximal conflict-free siphons (even maximal siphons) of a Petri net.
+The reason is that researchers mainly focus on minimal siphons~\cite{DBLP:journals/isci/LiuB16}.
+Hence, we here propose a new method based on Answer Set Programming (ASP)~\cite{DBLP:journals/aicom/GebserKKOSS11} for computing  maximal conflict-free siphons of a Petri net.
+The details of the proposed method shall be shown in the next section.
+
+\section{Answer set programming-based method}
 

### 2022-05-02 cmsb22.tex
Add the .bib file: use bibitems obtained from dblp, make preservation for some special words, add some new bibitems

@@
 
-Petri nets were introduced in the 60’s as a simple formalism for describing and analyzing information processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri}.
+Petri nets were introduced in the 60’s as a simple formalism for describing and analyzing information processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
 The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, was introduced quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.

### 2022-05-02 cmsb22.tex
Add new content to the Evaluation section

@@
 \section{Evaluation}
+To test the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet}.
+We used a set of real-world Boolean models lying in various scales obtained from the literature.
+These models are complex (i.e., having high average in-degree) and most of them have never been fully analyzed.
+We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
+To solve the ASP problems, we used the ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} with the configuration as same as that used in PyBoolNet~\cite{klarner2015computing}.
+Specifically, ...
+Finally, we set a time limit as one hour for each model.
+
 \subsection{Benchmark vs.\ implicant-based methods}
@@
 \url{http://colomoto.org/biolqm/doc/tools-trapspace.html}
+
+Table~\ref{tab:result_real} shows the experimental results on real-world models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
+
+\begin{table}[!ht]
+\centering
+\caption{Timing comparisons between the proposed method and PyBoolNet on the selected real-world models.
+Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively.
+Column "\(|M|\)" denotes the number of minimal trap spaces.
+The computational time is in seconds.
+"DNF" means that the method did not finish the computation within one hour.}
+\label{tab:result_real}
+\begin{tabular}{|c|r|r|r|r|r|r|r|r|r|}
+\cline{7-10}
+\multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{2}{c|}{PyBoolNet} & \multicolumn{2}{c|}{Proposed method} \\ \hline
+model & $n$ & $e$ & $s$ & $|M|$ & $|F|$ & t1 (s) & t2 (s) & t1 (s) & t2 (s) \\ \hline
+CASCADE-3 & 183 & - & 0 & 1 & 0 & - & - & 1.65 & - \\ \hline
+CHO-2016 & 200 & - & 13 & 1024 & 0 & - & - & 2.10 & - \\ \hline
+CACC & 70 & 153 & 1 & 10 & 2 & - & - & 0.18 & - \\ \hline
+THIL\_17 & 92 & - & 16 & 5724 & 5724 & - & - & 0.90 & - \\ \hline
+ERBB & 247 & 1100 & 22 & 1456394+ & 0 & DNF & DNF & DNF & - \\ \hline
+ER-STRESS & 182 & 266 & 75 & 25600 & 25600 & - & - & 4.54 & - \\ \hline
+AD & 1169 & - & 469 & 49152 & 49152 & - & - & 66.17 & - \\ \hline
+Cholocystokinin & 404 & - & 95 & 705024 & 705024 & - & - & 1080.35 & - \\ \hline
+FIBROBLASTS & 139 & 557 & 9 & 375850 & 32768 & - & - & 331.87 & - \\ \hline
+TcellCheckPoint & 218 & - & 34 & 15360 & 10240 & - & - & 4.33 & - \\ \hline
+HIV-1 & 138 & 368 & 14 & 1408 & 1408 & DNF & DNF & 1.44 & - \\ \hline
+HMOX1 & 145 & 228 & 56 & 5120 & 5120 & - & - & 0.99 & - \\ \hline
+Human network & 1953 & - & 675 & 2048 & 2048 & DNF & DNF & 1770.98 & - \\ \hline
+IL\_6 Signalling & 86 & 149 & 15 & 512 & 448 & - & - & 0.32 & - \\ \hline
+IMMUNE-SYS & 164 & 506 & 13 & 36852 & 27871 & DNF & DNF & 14.19 & - \\ \hline
+IBD & 47 & 287 & 0 & 1 & 0 & DNF & DNF & 296.23 & - \\ \hline
+Influenza A & 131 & 302 & 11 & 2560 & 2560 & - & - & 1.07 & - \\ \hline
+INTERFERON-1 & 121 & 190 & 55 & 512 & 512 & - & - & 0.43 & - \\ \hline
+KORKUT-2015 & 99 & - & 12 & 2060 & 1384 & DNF & DNF & 2.86 & - \\ \hline
+KY-PA & 150 & 304 & 72 & 512 & 512 & - & - & 3.79 & - \\ \hline
+RA Apoptosis & 431 & - & 59 & 640 & 0 & DNF & DNF & 0.70 & - \\ \hline
+CELL CYCLE 2020 & 121 & 493 & - & 78 & 78 & - & - & 3.72 & - \\ \hline
+S1\_Table & 1659 & - & 522 & 1024 & 1024 & DNF & DNF & 1511.07 & - \\ \hline
+Macrophage\_Act & 321 & 533 & 19 & 512 & 384 & - & - & 1.58 & - \\ \hline
+SIPC & 133 & 431 & 0 & 812 & 746 & - & - & 1.40 & - \\ \hline
+SPBP & 66 & 139 & 13 & 512 & 256 & - & - & 0.17 & - \\ \hline
+SN5 & 2746 & - & 835 & 2048 & 2048 & DNF & DNF & 3094.04 & - \\ \hline
+TCR-TLR5-SIG & 130 & 273 & 2 & 48 & 48 & - & - & 0.50 & - \\ \hline
+TLGLSurvival & 61 & 193 & 7 & 318 & 172 & - & - & 0.24 & - \\ \hline
+TUREI-2016 & 4691 & - & 1257 & - & - & DNF & DNF & DNF & DNF \\ \hline
+VIRUS-RE-CYC & 154 & 268 & 25 & 1024 & 1024 & - & - & 0.67 & - \\ \hline
+\end{tabular}
+\end{table}
 

### 2022-05-02 cmsb22.tex
Mostly superficial stuff…

--- 2022-05-02 cmsb22.tex
@@
 \documentclass[runningheads]{llncs}
-\usepackage[T5]{fontenc}
+\usepackage[T1]{fontenc}
 % \usepackage{graphicx}
@@
 \usepackage{xcolor}
+\usepackage{booktabs}
 
@@
 
-Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S} =\bbbb^{n}\) with \(\bbbb = \{0, 1\}\).
-
-A \emph{partial state} \(P\) of \(\mathcal{M}\) is a subset of \(\mathcal{S}\) where for some \(U\subset V\) the values are fixed by a mapping \(p: U\mapsto\bbbb\) and the other variables (\(V\setminus U\)) are \emph{free}.
-
-\begin{definition}
-
-  A \emph{trap space} of \(\mathcal{M}\) is a partial state \(P\) that is also a \emph{trap set}:
-
-  if \(P'\) is the image of \(P\) by the state transition function, then \(P'\subset P\).
-
-\end{definition}
-
-Here and in the following, the notion of \emph{minimality/maximality} will always be w.r.t. to subset inclusion.
+Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\bbbb^{n}\) with \(\bbbb = \{0, 1\}\).
 
 %%% Subspace %%%
-A subspace \(m\) of \(S_{\mathcal{M}}\) is characterized by its fixed variables (denoted by \(D_m\)) and free variables.
-This subspace can be specified by an assignment \(m : D_m \mapsto \bbbb\) where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
-The remaining variables of \(\mathcal{M}\), \(V \setminus D_m\), are said to be free, i.e., they can receive any Boolean value.
-We specify subspaces like states but use in addition the symbol \(\star\) to indicate that a variable is free.
-In addition, a subspace \(m\) also refers to the set of states \(\mathcal{S}_{\mathcal{M}}[m] := \{s \in \mathcal{S}_{\mathcal{M}}\;|\;\forall x_i \in D_m : s_i = m(x_i)\}\).
-For example, \(m = \star\star1\) means that \(D_m = \{x_3\}\), \(m(x_3) = 1\), and refers to the set of states \(\{001, 011, 101, 111\}\).
+A \emph{subspace} \(m\) of \(\mathcal{S}_{\mathcal{M}}\) is characterized by its fixed variables (denoted by \(D_m\)) and free variables.
+This subspace can be specified by an assignment \(D_m \mapsto \bbbb\), also denoted by \(m\) by abuse of notation, and where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
+The remaining variables of \(\mathcal{M}\), \(V \setminus D_m\), are said to be \emph{free}, i.e., they can receive any Boolean value.
+We write subspaces like states but use in addition the symbol \(\star\) to indicate that a variable is free.
+A subspace \(m\) thus corresponds to the set of states \(\mathcal{S}_{\mathcal{M}}[m] := \{s \in \mathcal{S}_{\mathcal{M}}\;|\;\forall x_i \in D_m : s_i = m(x_i)\}\).
+For example, \(m = \star\star1\) means that \(D_m = \{x_3\}\), \(m(x_3) = 1\), and corresponds to the set of states \(\{001, 011, 101, 111\}\).
 Let \(\mathcal{S}_{\mathcal{M}}^{\star}\) denote the set of all possible subspaces of \(\mathcal{M}\).
@@
 Consequently, a trap space \(m\) is minimal if and only if there is no trap space \(m' \in \mathcal{S}_{\mathcal{M}}^{\star}\) such that \(m' < m\).
-For example, the BN shown in Example~\ref{example:BN} has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
-Since \(m_1 < m_2\), \(m_1\) is a minimal trap space of the BN.
+For example, the BN shown in Example~\ref{example:BN} has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\). Since \(m_1 < m_2\), \(m_1\) is a minimal trap space of the BN.
+
+\sylvain{missing example?}
 
@@
 Since our study is focused on Boolean models, we briefly recall the original encoding here.
-Its basis is that every \emph{gene} of the original model is represented by two separate places, corresponding to its two states, active, and inactive.
+Its basis is that every \emph{gene} of the original model is represented by two separate places (\(p\) and \(\overline{p}\)), corresponding to its two states, active, and inactive.
 Each conjunct of the logical function that activate the \emph{gene} will lead to a transition, consuming the inactive place, producing the active place, and with all other literals both consumed and produced.
@@
 
+Intuitively, a siphon is a set of places that once empty remains so.
+If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the empty places remain empty).
+This is precisely the definition of a trap-space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap-space.
+
 \begin{proof}
@@
 
-Intuitively, a siphon is a set of places that once empty remains so.
-If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the empty places remain empty).
-This is precisely the definition of a trap-space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap-space.
-
-\begin{proof}
-
-  TBD\qed
-
-\end{proof}
-
 By Theorem~\ref{theo:min_ts_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean model to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
@@
 \section{Evaluation}
+To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet}.
+We used a set of real-world Boolean models lying in various scales obtained from the literature.
+These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
+We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
+To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
+Specifically, ...
+Finally, we set a time limit as one hour for each model.
+
+\sylvain{we need a source for the models as much as possible, or to refer to some place where people can access them to reproduce the benchmark.}
+
 \subsection{Benchmark vs.\ implicant-based methods}
 
-Using BioLQM~\ref{sec:encoding}
+Table~\ref{tab:result_real} shows the experimental results on real-world models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
 
-\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}
+\begin{table}[!ht]
+  \caption{Timing comparisons between the proposed method and PyBoolNet on the selected real-world models.
+    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively.
+    Column "\(|M|\)" denotes the number of minimal trap spaces.
+    The computation times are all in seconds.
+    "DNF" means that the method did not finish the computation within one hour.}
+  \label{tab:result_real}
+  \begin{tabular}{lrrrrrrrrr}
+    \toprule
+    & & & & & & \multicolumn{2}{c}{PyBoolNet} & \multicolumn{2}{c}{Proposed method}\\
+    \cmidrule(lr){7-8}\cmidrule(lr){9-10}
+    model & $n$ & $e$ & $s$ & $|M|$ & $|F|$ & t1 (s) & t2 (s) & t1 (s) & t2 (s) \\ \midrule
+    CASCADE-3 & 183 & - & 0 & 1 & 0 & - & - & 1.65 & - \\
+    CHO-2016 & 200 & - & 13 & 1024 & 0 & - & - & 2.10 & - \\
+    CACC & 70 & 153 & 1 & 10 & 2 & - & - & 0.18 & - \\
+    THIL\_17 & 92 & - & 16 & 5724 & 5724 & - & - & 0.90 & - \\
+    ERBB & 247 & 1100 & 22 & 1456394+ & 0 & DNF & DNF & DNF & - \\
+    ER-STRESS & 182 & 266 & 75 & 25600 & 25600 & - & - & 4.54 & - \\
+    AD & 1169 & - & 469 & 49152 & 49152 & - & - & 66.17 & - \\
+    Cholocystokinin & 404 & - & 95 & 705024 & 705024 & - & - & 1080.35 & - \\
+    FIBROBLASTS & 139 & 557 & 9 & 375850 & 32768 & - & - & 331.87 & - \\
+    TcellCheckPoint & 218 & - & 34 & 15360 & 10240 & - & - & 4.33 & - \\
+    HIV-1 & 138 & 368 & 14 & 1408 & 1408 & DNF & DNF & 1.44 & - \\
+    HMOX1 & 145 & 228 & 56 & 5120 & 5120 & - & - & 0.99 & - \\
+    Human network & 1953 & - & 675 & 2048 & 2048 & DNF & DNF & 1770.98 & - \\
+    IL\_6 Signalling & 86 & 149 & 15 & 512 & 448 & - & - & 0.32 & - \\
+    IMMUNE-SYS & 164 & 506 & 13 & 36852 & 27871 & DNF & DNF & 14.19 & - \\
+    IBD & 47 & 287 & 0 & 1 & 0 & DNF & DNF & 296.23 & - \\
+    Influenza A & 131 & 302 & 11 & 2560 & 2560 & - & - & 1.07 & - \\
+    INTERFERON-1 & 121 & 190 & 55 & 512 & 512 & - & - & 0.43 & - \\
+    KORKUT-2015 & 99 & - & 12 & 2060 & 1384 & DNF & DNF & 2.86 & - \\
+    KY-PA & 150 & 304 & 72 & 512 & 512 & - & - & 3.79 & - \\
+    RA Apoptosis & 431 & - & 59 & 640 & 0 & DNF & DNF & 0.70 & - \\
+    CELL CYCLE 2020 & 121 & 493 & - & 78 & 78 & - & - & 3.72 & - \\
+    S1\_Table & 1659 & - & 522 & 1024 & 1024 & DNF & DNF & 1511.07 & - \\
+    Macrophage\_Act & 321 & 533 & 19 & 512 & 384 & - & - & 1.58 & - \\
+    SIPC & 133 & 431 & 0 & 812 & 746 & - & - & 1.40 & - \\
+    SPBP & 66 & 139 & 13 & 512 & 256 & - & - & 0.17 & - \\
+    SN5 & 2746 & - & 835 & 2048 & 2048 & DNF & DNF & 3094.04 & - \\
+    TCR-TLR5-SIG & 130 & 273 & 2 & 48 & 48 & - & - & 0.50 & - \\
+    TLGLSurvival & 61 & 193 & 7 & 318 & 172 & - & - & 0.24 & - \\
+    TUREI-2016 & 4691 & - & 1257 & - & - & DNF & DNF & DNF & DNF \\
+    VIRUS-RE-CYC & 154 & 268 & 25 & 1024 & 1024 & - & - & 0.67 & - \\ \bottomrule
+  \end{tabular}
+
+\sylvain{I changed the ``look'' of the table using booktabs, see \url{https://nhigham.com/2019/11/19/better-latex-tables-with-booktabs/} for instance.}
+
+\sylvain{what is $|F|$? what are t1 and t2? why are there no times for BoolNet?}
+\end{table}
 
@@
 
+Also to a BDD-based approach using BioLQM is proposed at \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} but because of the BDD conversion, some models are not even loaded.
+This is also an issue for using BioLQM as a translator/preprocessor and since we do not use BDDs we might rely on writing our own parser for SBML-Qual and such formats.
+
 \subsubsection{Acknowledgments}
@@
 \end{document}
-
-% \begin{table}
-% \caption{Table captions should be placed above the
-% tables.}\label{tab1}
-% \begin{tabular}{|l|l|l|}
-% \hline
-% Heading level &  Example & Font size and style\\
-% \hline
-% Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-% 1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-% 2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-% 3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-% 4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-% \hline
-% \end{tabular}
-% \end{table}
 

### 2022-05-02 cmsb22.tex
Refer to some places (not fully provided now) where people can access the real-world models to reproduce the benchmark. We can also put all the model files into the Git repo.

@@
     model & $n$ & $e$ & $s$ & $|M|$ & $|F|$ & t1 (s) & t2 (s) & t1 (s) & t2 (s) \\ \midrule
-    CASCADE-3 & 183 & - & 0 & 1 & 0 & - & - & 1.65 & - \\
+    CASCADE-3~\cite{Tsirvouli2020} & 183 & - & 0 & 1 & 0 & - & - & 1.65 & - \\
     CHO-2016 & 200 & - & 13 & 1024 & 0 & - & - & 2.10 & - \\
-    CACC & 70 & 153 & 1 & 10 & 2 & - & - & 0.18 & - \\
+    CACC~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 153 & 1 & 10 & 2 & - & - & 0.18 & - \\
     THIL\_17 & 92 & - & 16 & 5724 & 5724 & - & - & 0.90 & - \\
@@
     ER-STRESS & 182 & 266 & 75 & 25600 & 25600 & - & - & 4.54 & - \\
-    AD & 1169 & - & 469 & 49152 & 49152 & - & - & 66.17 & - \\
-    Cholocystokinin & 404 & - & 95 & 705024 & 705024 & - & - & 1080.35 & - \\
-    FIBROBLASTS & 139 & 557 & 9 & 375850 & 32768 & - & - & 331.87 & - \\
+    AD~\cite{aghamiri2020automated} & 1169 & - & 469 & 49152 & 49152 & - & - & 66.17 & - \\
+    Cholocystokinin~\cite{aghamiri2020automated} & 404 & - & 95 & 705024 & 705024 & - & - & 1080.35 & - \\
+    FIBROBLASTS~\cite{helikar2008emergent} & 139 & 557 & 9 & 375850 & 32768 & - & - & 331.87 & - \\
     TcellCheckPoint & 218 & - & 34 & 15360 & 10240 & - & - & 4.33 & - \\
-    HIV-1 & 138 & 368 & 14 & 1408 & 1408 & DNF & DNF & 1.44 & - \\
-    HMOX1 & 145 & 228 & 56 & 5120 & 5120 & - & - & 0.99 & - \\
+    HIV-1~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 368 & 14 & 1408 & 1408 & DNF & DNF & 1.44 & - \\
+    HMOX1~\cite{ostaszewski2021covid19} & 145 & 228 & 56 & 5120 & 5120 & - & - & 0.99 & - \\
     Human network & 1953 & - & 675 & 2048 & 2048 & DNF & DNF & 1770.98 & - \\
-    IL\_6 Signalling & 86 & 149 & 15 & 512 & 448 & - & - & 0.32 & - \\
+    IL\_6 Signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 149 & 15 & 512 & 448 & - & - & 0.32 & - \\
     IMMUNE-SYS & 164 & 506 & 13 & 36852 & 27871 & DNF & DNF & 14.19 & - \\
-    IBD & 47 & 287 & 0 & 1 & 0 & DNF & DNF & 296.23 & - \\
-    Influenza A & 131 & 302 & 11 & 2560 & 2560 & - & - & 1.07 & - \\
-    INTERFERON-1 & 121 & 190 & 55 & 512 & 512 & - & - & 0.43 & - \\
+    IBD~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 287 & 0 & 1 & 0 & DNF & DNF & 296.23 & - \\
+    Influenza A~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 302 & 11 & 2560 & 2560 & - & - & 1.07 & - \\
+    INTERFERON-1~\cite{ostaszewski2021covid19} & 121 & 190 & 55 & 512 & 512 & - & - & 0.43 & - \\
     KORKUT-2015 & 99 & - & 12 & 2060 & 1384 & DNF & DNF & 2.86 & - \\
     KY-PA & 150 & 304 & 72 & 512 & 512 & - & - & 3.79 & - \\
-    RA Apoptosis & 431 & - & 59 & 640 & 0 & DNF & DNF & 0.70 & - \\
+    RA Apoptosis~\cite{aghamiri2020automated} & 431 & - & 59 & 640 & 0 & DNF & DNF & 0.70 & - \\
     CELL CYCLE 2020 & 121 & 493 & - & 78 & 78 & - & - & 3.72 & - \\
     S1\_Table & 1659 & - & 522 & 1024 & 1024 & DNF & DNF & 1511.07 & - \\
-    Macrophage\_Act & 321 & 533 & 19 & 512 & 384 & - & - & 1.58 & - \\
-    SIPC & 133 & 431 & 0 & 812 & 746 & - & - & 1.40 & - \\
+    Macrophage\_Act~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 533 & 19 & 512 & 384 & - & - & 1.58 & - \\
+    SIPC~\cite{Montagud2021} & 133 & 431 & 0 & 812 & 746 & - & - & 1.40 & - \\
     SPBP & 66 & 139 & 13 & 512 & 256 & - & - & 0.17 & - \\
-    SN5 & 2746 & - & 835 & 2048 & 2048 & DNF & DNF & 3094.04 & - \\
+    SN5~\cite{kim2013rmod} & 2746 & - & 835 & 2048 & 2048 & DNF & DNF & 3094.04 & - \\
     TCR-TLR5-SIG & 130 & 273 & 2 & 48 & 48 & - & - & 0.50 & - \\
-    TLGLSurvival & 61 & 193 & 7 & 318 & 172 & - & - & 0.24 & - \\
+    TLGLSurvival~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 193 & 7 & 318 & 172 & - & - & 0.24 & - \\
     TUREI-2016 & 4691 & - & 1257 & - & - & DNF & DNF & DNF & DNF \\

### 2022-05-02 trappist.py
Cleaner code, and tentative benchmarking notebook
not very convincing on PyBoolNet models

@@ -23,7 +23,7 @@ import subprocess
 import sys
 import tempfile
 import xml.etree.ElementTree as etree
-from typing import IO, List
+from typing import Generator, IO, List, Optional
 
 import networkx as nx  # TODO maybe replace with lists/dicts
 
@@ -40,7 +40,7 @@ def read_pnml(fileobj: IO) -> nx.DiGraph:
 
     for place in root.findall("./net/place"):
         net.add_node(
-            place.get("id"), kind="place", name=place.find("./name/value").text
+            place.get("id"), kind="place"  # , name=place.find("./name/value").text
         )
 
     for transition in root.findall("./net/transition"):
@@ -79,6 +79,7 @@ def write_asp(petri_net: nx.DiGraph, asp_file: IO):
 
 def solve_asp(asp_filename: str) -> str:
     """Run an ASP solver on program asp_file and get the solutions."""
+    # FIXME we need a limit on number of solutions
     result = subprocess.run(
         [
             "clingo",
@@ -118,17 +119,45 @@ def place_in_sol(sol: List[str], place: str) -> str:
     return "-"
 
 
-def write_solutions(asp_output: str, petri_net: nx.DiGraph):
+def get_solutions(
+    asp_output: str, petri_net: nx.DiGraph, display: bool
+) -> Optional[Generator[List[str], None, None]]:
     """Display the ASP output back as trap-spaces."""
     places = []
     for node, kind in petri_net.nodes(data="kind"):
         if kind == "place" and not node.startswith("-"):
             places.append(node)
-    print(" ".join(places))
+    if display:
+        print(" ".join(places))
     solutions = json.loads(asp_output)
-    for sol in solutions["Call"][0]["Witnesses"]:
-        print(" ".join(solution_to_bool(places, sol["Value"])))
-    print("Total time:", solutions["Time"]["Total"], "s")
+    if display:
+        print(
+            "\n".join(
+                " ".join(solution_to_bool(places, sol["Value"]))
+                for sol in solutions["Call"][0]["Witnesses"]
+            )
+        )
+        print("Total time:", solutions["Time"]["Total"], "s")
+        return None
+    else:
+        return (
+            solution_to_bool(places, sol["Value"])
+            for sol in solutions["Call"][0]["Witnesses"]
+        )
+
+
+def compute_trap_spaces(
[…]

### 2022-05-03 cmsb22.tex
Update some bibitems. Adjust the Evaluation section: write the introduction graph, adjust the result table

--- 2022-05-02 cmsb22.tex
@@
 To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
-Specifically, ...
+Specifically, we used the configuration \texttt{--dom-pref=32, --heu=domain, --dom-mod=7} (subset maximality).
+We ran all the benchmarks on a virtual machine whose environment is CPU: Intel(R) Core(TM) i7-3630QM 2.40GHz x 4, Memory: 8 GB, Ubuntu 18.04.2 64 bit.
 Finally, we set a time limit as one hour for each model.
@@
 \sylvain{we need a source for the models as much as possible, or to refer to some place where people can access them to reproduce the benchmark.}
+\vangiang{Yes. I have added reference sources for the models but some models are still missing now.}
 
@@
     Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively.
-    Column "\(|M|\)" denotes the number of minimal trap spaces.
-    The computation times are all in seconds.
+For each method, Column "\(|M|\)" denotes the number of minimal trap spaces and Column "time (s)" denotes the computation time in seconds.
     "DNF" means that the method did not finish the computation within one hour.}
+  \centering
   \label{tab:result_real}
-  \begin{tabular}{lrrrrrrrrr}
+  \begin{tabular}{lrrrrrrr}
     \toprule
-    & & & & & & \multicolumn{2}{c}{PyBoolNet} & \multicolumn{2}{c}{Proposed method}\\
-    \cmidrule(lr){7-8}\cmidrule(lr){9-10}
-    model & $n$ & $e$ & $s$ & $|M|$ & $|F|$ & t1 (s) & t2 (s) & t1 (s) & t2 (s) \\ \midrule
-    CASCADE-3 & 183 & - & 0 & 1 & 0 & - & - & 1.65 & - \\
-    CHO-2016 & 200 & - & 13 & 1024 & 0 & - & - & 2.10 & - \\
-    CACC & 70 & 153 & 1 & 10 & 2 & - & - & 0.18 & - \\
-    THIL\_17 & 92 & - & 16 & 5724 & 5724 & - & - & 0.90 & - \\
-    ERBB & 247 & 1100 & 22 & 1456394+ & 0 & DNF & DNF & DNF & - \\
-    ER-STRESS & 182 & 266 & 75 & 25600 & 25600 & - & - & 4.54 & - \\
-    AD & 1169 & - & 469 & 49152 & 49152 & - & - & 66.17 & - \\
-    Cholocystokinin & 404 & - & 95 & 705024 & 705024 & - & - & 1080.35 & - \\
-    FIBROBLASTS & 139 & 557 & 9 & 375850 & 32768 & - & - & 331.87 & - \\
-    TcellCheckPoint & 218 & - & 34 & 15360 & 10240 & - & - & 4.33 & - \\
-    HIV-1 & 138 & 368 & 14 & 1408 & 1408 & DNF & DNF & 1.44 & - \\
-    HMOX1 & 145 & 228 & 56 & 5120 & 5120 & - & - & 0.99 & - \\
-    Human network & 1953 & - & 675 & 2048 & 2048 & DNF & DNF & 1770.98 & - \\
-    IL\_6 Signalling & 86 & 149 & 15 & 512 & 448 & - & - & 0.32 & - \\
-    IMMUNE-SYS & 164 & 506 & 13 & 36852 & 27871 & DNF & DNF & 14.19 & - \\
-    IBD & 47 & 287 & 0 & 1 & 0 & DNF & DNF & 296.23 & - \\
-    Influenza A & 131 & 302 & 11 & 2560 & 2560 & - & - & 1.07 & - \\
-    INTERFERON-1 & 121 & 190 & 55 & 512 & 512 & - & - & 0.43 & - \\
-    KORKUT-2015 & 99 & - & 12 & 2060 & 1384 & DNF & DNF & 2.86 & - \\
-    KY-PA & 150 & 304 & 72 & 512 & 512 & - & - & 3.79 & - \\
-    RA Apoptosis & 431 & - & 59 & 640 & 0 & DNF & DNF & 0.70 & - \\
-    CELL CYCLE 2020 & 121 & 493 & - & 78 & 78 & - & - & 3.72 & - \\
-    S1\_Table & 1659 & - & 522 & 1024 & 1024 & DNF & DNF & 1511.07 & - \\
-    Macrophage\_Act & 321 & 533 & 19 & 512 & 384 & - & - & 1.58 & - \\
-    SIPC & 133 & 431 & 0 & 812 & 746 & - & - & 1.40 & - \\
-    SPBP & 66 & 139 & 13 & 512 & 256 & - & - & 0.17 & - \\
-    SN5 & 2746 & - & 835 & 2048 & 2048 & DNF & DNF & 3094.04 & - \\
-    TCR-TLR5-SIG & 130 & 273 & 2 & 48 & 48 & - & - & 0.50 & - \\
-    TLGLSurvival & 61 & 193 & 7 & 318 & 172 & - & - & 0.24 & - \\
-    TUREI-2016 & 4691 & - & 1257 & - & - & DNF & DNF & DNF & DNF \\
-    VIRUS-RE-CYC & 154 & 268 & 25 & 1024 & 1024 & - & - & 0.67 & - \\ \bottomrule
+    & & & & \multicolumn{2}{c}{PyBoolNet} & \multicolumn{2}{c}{Proposed method}\\
+    \cmidrule(lr){5-6}\cmidrule(lr){7-8}
+    model & $n$ & $e$ & $s$ & $|M|$ & time (s) & $|M|$ & time (s) \\ \midrule
+    cascade-3~\cite{Tsirvouli2020} & 183 & - & 0 & - & - & 1 & 1.65 \\
+    CHO-2016 & 200 & - & 13 & - & - & 1024 & 2.10 \\
+    colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 153 & 1 & - & - & 10 & 0.18 \\
+    THIL\_17 & 92 & - & 16 & - & - & 5724 & 0.90 \\
+    ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 1100 & 22 & DNF & DNF & DNF & DNF \\
+    er-stress & 182 & 266 & 75 & - & - & 25600 & 4.54 \\
+    Alzheimer~\cite{aghamiri2020automated} & 1169 & - & 469 & - & - & 49152 & 66.17 \\
+    cholocystokinin~\cite{aghamiri2020automated} & 404 & - & 95 & - & - & 705024 & 1080.35 \\
+    signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 557 & 9 & - & - & 375850 & 331.87 \\
+    T-cell-check-point & 218 & - & 34 & - & - & 15360 & 4.33 \\
+    HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 368 & 14 & DNF & DNF & 1408 & 1.44 \\
+    HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 228 & 56 & - & - & 5120 & 0.99 \\
+    human-network & 1953 & - & 675 & DNF & DNF & 2048 & 1770.98 \\
+    IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 149 & 15 & - & - & 512 & 0.32 \\
+    immune-system & 164 & 506 & 13 & DNF & DNF & 36852 & 14.19 \\
+    inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 287 & 0 & DNF & DNF & 1 & 296.23 \\
+    influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 302 & 11 & - & - & 2560 & 1.07 \\
+    interferon-1~\cite{ostaszewski2021covid19} & 121 & 190 & 55 & - & - & 512 & 0.43 \\
+    KORKUT-2015 & 99 & - & 12 & DNF & DNF & 2060 & 2.86 \\
+    KY-PA & 150 & 304 & 72 & - & - & 512 & 3.79 \\
+    RA-apoptosis~\cite{aghamiri2020automated} & 431 & - & 59 & DNF & DNF & 640 & 0.70 \\
+    CELL CYCLE 2020 & 121 & 493 & - & - & - & 78 & 3.72 \\
+    S1\_Table & 1659 & - & 522 & DNF & DNF & 1024 & 1511.07 \\
+    signaling-in-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 533 & 19 & - & - & 512 & 1.58 \\
+    SIPC~\cite{Montagud2021} & 133 & 431 & 0 & - & - & 812 & 1.40 \\
+    SPBP & 66 & 139 & 13 & - & - & 512 & 0.17 \\
+    SN5~\cite{kim2013rmod} & 2746 & - & 835 & DNF & DNF & 2048 & 3094.04 \\
+    TCR-TLR5-signaling-2018 & 130 & 273 & 2 & - & - & 48 & 0.50 \\
+    T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 193 & 7 & - & - & 318 & 0.24 \\
+    TUREI-2016 & 4691 & - & 1257 & DNF & DNF & DNF & DNF \\
+    VIRUS-RE-CYC & 154 & 268 & 25 & - & - & 1024 & 0.67 \\ \bottomrule
   \end{tabular}
@@
 \sylvain{I changed the ``look'' of the table using booktabs, see \url{https://nhigham.com/2019/11/19/better-latex-tables-with-booktabs/} for instance.}
+\vangiang{Actually, booktabs looks better. I have never used it before. Thank you for sharing.}
 
 \sylvain{what is $|F|$? what are t1 and t2? why are there no times for BoolNet?}
+\vangiang{I need to open an excel file containing the experimental results but I am unable to access this file now.}
+
 \end{table}

### 2022-05-03 cmsb22.tex
Update some definitions and proofs in Section "Minimal trap spaces as maximal conflict-free siphons".

@@
   A \emph{siphon} of a Petri net \((P, T, W)\) is a set of places \(S\) such that:
-  \[\forall t\in T, S\cap succ(t)\not =\emptyset\Rightarrow S\cap pred(t)\not =\emptyset\]
+  \[\forall t\in T, S\cap succ(t)\not =\emptyset\Rightarrow S\cap pred(t)\not =\emptyset.\] \(\emptyset\) is trivially a siphon.
 
@@
 
-We add a definition related to any set of places of a Petri net encoding a Logical model, and notably a siphon of such a net.
+\section{Minimal trap spaces as maximal conflict-free siphons}
+First, we add a definition related to any set of places of a Petri net encoding a Logical model, and notably a siphon of such a net.
 
@@
 
-\section{Minimal trap spaces as maximal conflict-free siphons}
+Intuitively, a siphon is a set of places that once empty remains so.
+If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the empty places remain empty).
+This is precisely the definition of a trap-space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap-space.
+Hereafter, we formally prove that a maximal conflict-free siphon is equivalent to a minimal trap space.
 
@@
   Let \(m\) be a subspace of Boolean model \(\mathcal{M}\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) such that:
-  \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S\]
+  \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S.\] If \(D_m = \emptyset\), \(S\) is conventionally \(\emptyset\).
 
@@
 \begin{theorem}
+\label{theo:ts_2_sp}
 
@@
 
-Intuitively, a siphon is a set of places that once empty remains so.
-If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the empty places remain empty).
-This is precisely the definition of a trap-space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap-space.
-
 \begin{proof}
 
-  TBD\qed
+  First, we show that if \(m\) is a trap space of \(\mathcal{M}\), then \(S\) is a conflict-free siphon of \(\mathcal{P}\) (*).
+
+  Second, we show that if \(S\) is a conflict-free siphon of \(\mathcal{P}\), then \(m\) is a trap space of \(\mathcal{M}\) (**).
+
+  From (*) and (**), we can conclude the proof.\qed
 
@@
 \begin{theorem}
-\label{theo:min_ts_max_sp}
+\label{theo:min_ts_2_max_sp}
   %\(S\) is a minimal trap-space of Boolean model \(\mathcal{M}\) if and only if the places corresponding to the fixed values in \(S\) (active or inactive) are a maximal conflict-free siphon of \((P, T, W)\), the Petri net encoding of \(S\).
@@
 
-By Theorem~\ref{theo:min_ts_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean model to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
+\begin{proof}
+
+  First, we show that if \(m\) is a minimal trap space of \(\mathcal{M}\), then \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\) (*).
+
+  Second, we show that if \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\), then \(m\) is a minimal trap space of \(\mathcal{M}\) (**).
+
+  From (*) and (**), we can conclude the proof.\qed
+
+\end{proof}
+
+By Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean model to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
 It is noted that there are no existing methods specifically designed for computing maximal conflict-free siphons (even maximal siphons) of a Petri net.

### 2022-05-04 cmsb22.tex
Add the full list of models with reference papers

@@
     model & $n$ & $e$ & $s$ & $|M|$ & time (s) & $|M|$ & time (s) \\ \midrule
-    cascade-3~\cite{Tsirvouli2020} & 183 & - & 0 & - & - & 1 & 1.65 \\
-    CHO-2016 & 200 & - & 13 & - & - & 1024 & 2.10 \\
+    cascade-3~\cite{Tsirvouli2020} & 183 & 603 & 0 & - & - & 1 & 1.65 \\
+    CHO-2016~\cite{lee2019signal} & 200 & 696 & 13 & - & - & 1024 & 2.10 \\
     colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 153 & 1 & - & - & 10 & 0.18 \\
-    THIL\_17 & 92 & - & 16 & - & - & 5724 & 0.90 \\
+    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & - & 16 & - & - & 5724 & 0.90 \\
     ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 1100 & 22 & DNF & DNF & DNF & DNF \\
-    er-stress & 182 & 266 & 75 & - & - & 25600 & 4.54 \\
+    er-stress~\cite{ostaszewski2021covid19} & 182 & 266 & 75 & - & - & 25600 & 4.54 \\
     Alzheimer~\cite{aghamiri2020automated} & 1169 & - & 469 & - & - & 49152 & 66.17 \\
@@
     signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 557 & 9 & - & - & 375850 & 331.87 \\
-    T-cell-check-point & 218 & - & 34 & - & - & 15360 & 4.33 \\
+    T-cell-check-point~\cite{hernandez2020computational} & 218 & - & 34 & - & - & 15360 & 4.33 \\
     HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 368 & 14 & DNF & DNF & 1408 & 1.44 \\
     HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 228 & 56 & - & - & 5120 & 0.99 \\
-    human-network & 1953 & - & 675 & DNF & DNF & 2048 & 1770.98 \\
+    human-network~\cite{kim2011reduction} & 1953 & - & 675 & DNF & DNF & 2048 & 1770.98 \\
     IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 149 & 15 & - & - & 512 & 0.32 \\
-    immune-system & 164 & 506 & 13 & DNF & DNF & 36852 & 14.19 \\
+    immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 506 & 13 & DNF & DNF & 36852 & 14.19 \\
     inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 287 & 0 & DNF & DNF & 1 & 296.23 \\
@@
     interferon-1~\cite{ostaszewski2021covid19} & 121 & 190 & 55 & - & - & 512 & 0.43 \\
-    KORKUT-2015 & 99 & - & 12 & DNF & DNF & 2060 & 2.86 \\
-    KY-PA & 150 & 304 & 72 & - & - & 512 & 3.79 \\
+    Korkut-2015~\cite{lee2019signal} & 99 & - & 12 & DNF & DNF & 2060 & 2.86 \\
+    kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 304 & 72 & - & - & 512 & 3.79 \\
     RA-apoptosis~\cite{aghamiri2020automated} & 431 & - & 59 & DNF & DNF & 640 & 0.70 \\
-    CELL CYCLE 2020 & 121 & 493 & - & - & - & 78 & 3.72 \\
-    S1\_Table & 1659 & - & 522 & DNF & DNF & 1024 & 1511.07 \\
-    signaling-in-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 533 & 19 & - & - & 512 & 1.58 \\
-    SIPC~\cite{Montagud2021} & 133 & 431 & 0 & - & - & 812 & 1.40 \\
-    SPBP & 66 & 139 & 13 & - & - & 512 & 0.17 \\
-    SN5~\cite{kim2013rmod} & 2746 & - & 835 & DNF & DNF & 2048 & 3094.04 \\
-    TCR-TLR5-signaling-2018 & 130 & 273 & 2 & - & - & 48 & 0.50 \\
+    adhesion-CIP-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 493 & - & - & - & 78 & 3.72 \\
+    KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & - & 522 & DNF & DNF & 1024 & 1511.07 \\
+    signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 533 & 19 & - & - & 512 & 1.58 \\
+    signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 431 & 0 & - & - & 812 & 1.40 \\
+    butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 139 & 13 & - & - & 512 & 0.17 \\
+    SN-5~\cite{kim2013rmod} & 2746 & - & 835 & DNF & DNF & 2048 & 3094.04 \\
+    TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 273 & 2 & - & - & 48 & 0.50 \\
     T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 193 & 7 & - & - & 318 & 0.24 \\
-    TUREI-2016 & 4691 & - & 1257 & DNF & DNF & DNF & DNF \\
-    VIRUS-RE-CYC & 154 & 268 & 25 & - & - & 1024 & 0.67 \\ \bottomrule
+    turei-2016~\cite{lee2019signal} & 4691 & - & 1257 & DNF & DNF & DNF & DNF \\
+    virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 268 & 25 & - & - & 1024 & 0.67 \\ \bottomrule
   \end{tabular}

### 2022-05-04 cmsb22.tex

--- 2022-05-03 cmsb22.tex
@@
     model & $n$ & $e$ & $s$ & $|M|$ & time (s) & $|M|$ & time (s) \\ \midrule
-    cascade-3~\cite{Tsirvouli2020} & 183 & - & 0 & - & - & 1 & 1.65 \\
-    CHO-2016 & 200 & - & 13 & - & - & 1024 & 2.10 \\
+    inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 287 & 0 & DNF & DNF & 1 & 296.23 \\
+    T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 193 & 7 & - & - & 318 & 0.24 \\
+    butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 139 & 13 & - & - & 512 & 0.17 \\
     colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 153 & 1 & - & - & 10 & 0.18 \\
-    THIL\_17 & 92 & - & 16 & - & - & 5724 & 0.90 \\
+    IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 149 & 15 & - & - & 512 & 0.32 \\
+    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & - & 16 & - & - & 5724 & 0.90 \\
+    Korkut-2015~\cite{lee2019signal} & 99 & - & 12 & DNF & DNF & 2060 & 2.86 \\ \midrule
+
+    adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 493 & - & - & - & 78 & 3.72 \\
+    interferon-1~\cite{ostaszewski2021covid19} & 121 & 190 & 55 & - & - & 512 & 0.43 \\
+    TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 273 & 2 & - & - & 48 & 0.50 \\
+    influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 302 & 11 & - & - & 2560 & 1.07 \\
+    signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 431 & 0 & - & - & 812 & 1.40 \\
+    HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 368 & 14 & DNF & DNF & 1408 & 1.44 \\
+    signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 557 & 9 & - & - & 375850 & 331.87 \\
+    HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 228 & 56 & - & - & 5120 & 0.99 \\
+    kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 304 & 72 & - & - & 512 & 3.79 \\
+    virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 268 & 25 & - & - & 1024 & 0.67 \\
+    immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 506 & 13 & DNF & DNF & 36852 & 14.19 \\
+    er-stress~\cite{ostaszewski2021covid19} & 182 & 266 & 75 & - & - & 25600 & 4.54 \\
+    cascade-3~\cite{Tsirvouli2020} & 183 & 603 & 0 & - & - & 1 & 1.65 \\
+    CHO-2016~\cite{lee2019signal} & 200 & 696 & 13 & - & - & 1024 & 2.10 \\ \midrule
+
+    T-cell-check-point~\cite{hernandez2020computational} & 218 & - & 34 & - & - & 15360 & 4.33 \\
     ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 1100 & 22 & DNF & DNF & DNF & DNF \\
-    er-stress & 182 & 266 & 75 & - & - & 25600 & 4.54 \\
+    signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 533 & 19 & - & - & 512 & 1.58 \\
+    cholocystokinin~\cite{aghamiri2020automated} & 404 & - & 95 & - & - & 705024 & 1080.35 \\
+    RA-apoptosis~\cite{aghamiri2020automated} & 431 & - & 59 & DNF & DNF & 640 & 0.70 \\ \midrule
+
     Alzheimer~\cite{aghamiri2020automated} & 1169 & - & 469 & - & - & 49152 & 66.17 \\
-    cholocystokinin~\cite{aghamiri2020automated} & 404 & - & 95 & - & - & 705024 & 1080.35 \\
-    signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 557 & 9 & - & - & 375850 & 331.87 \\
-    T-cell-check-point & 218 & - & 34 & - & - & 15360 & 4.33 \\
-    HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 368 & 14 & DNF & DNF & 1408 & 1.44 \\
-    HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 228 & 56 & - & - & 5120 & 0.99 \\
-    human-network & 1953 & - & 675 & DNF & DNF & 2048 & 1770.98 \\
-    IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 149 & 15 & - & - & 512 & 0.32 \\
-    immune-system & 164 & 506 & 13 & DNF & DNF & 36852 & 14.19 \\
-    inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 287 & 0 & DNF & DNF & 1 & 296.23 \\
-    influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 302 & 11 & - & - & 2560 & 1.07 \\
-    interferon-1~\cite{ostaszewski2021covid19} & 121 & 190 & 55 & - & - & 512 & 0.43 \\
-    KORKUT-2015 & 99 & - & 12 & DNF & DNF & 2060 & 2.86 \\
-    KY-PA & 150 & 304 & 72 & - & - & 512 & 3.79 \\
-    RA-apoptosis~\cite{aghamiri2020automated} & 431 & - & 59 & DNF & DNF & 640 & 0.70 \\
-    CELL CYCLE 2020 & 121 & 493 & - & - & - & 78 & 3.72 \\
-    S1\_Table & 1659 & - & 522 & DNF & DNF & 1024 & 1511.07 \\
-    signaling-in-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 533 & 19 & - & - & 512 & 1.58 \\
-    SIPC~\cite{Montagud2021} & 133 & 431 & 0 & - & - & 812 & 1.40 \\
-    SPBP & 66 & 139 & 13 & - & - & 512 & 0.17 \\
-    SN5~\cite{kim2013rmod} & 2746 & - & 835 & DNF & DNF & 2048 & 3094.04 \\
-    TCR-TLR5-signaling-2018 & 130 & 273 & 2 & - & - & 48 & 0.50 \\
-    T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 193 & 7 & - & - & 318 & 0.24 \\
-    TUREI-2016 & 4691 & - & 1257 & DNF & DNF & DNF & DNF \\
-    VIRUS-RE-CYC & 154 & 268 & 25 & - & - & 1024 & 0.67 \\ \bottomrule
+    KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & - & 522 & DNF & DNF & 1024 & 1511.07 \\
+    human-network~\cite{kim2011reduction} & 1953 & - & 675 & DNF & DNF & 2048 & 1770.98 \\ \midrule
+
+    SN-5~\cite{kim2013rmod} & 2746 & - & 835 & DNF & DNF & 2048 & 3094.04 \\
+    turei-2016~\cite{lee2019signal} & 4691 & - & 1257 & DNF & DNF & DNF & DNF \\
+
+    \bottomrule
   \end{tabular}
 
-\sylvain{I changed the ``look'' of the table using booktabs, see \url{https://nhigham.com/2019/11/19/better-latex-tables-with-booktabs/} for instance.}
-\vangiang{Actually, booktabs looks better. I have never used it before. Thank you for sharing.}
+% \sylvain{I changed the ``look'' of the table using booktabs, see \url{https://nhigham.com/2019/11/19/better-latex-tables-with-booktabs/} for instance.}
+% \vangiang{Actually, booktabs looks better. I have never used it before. Thank you for sharing.}
 
-\sylvain{what is $|F|$? what are t1 and t2? why are there no times for BoolNet?}
-\vangiang{I need to open an excel file containing the experimental results but I am unable to access this file now.}
+% \sylvain{what is $|F|$? what are t1 and t2? why are there no times for BoolNet?}
+% \vangiang{I need to open an excel file containing the experimental results but I am unable to access this file now.}
 

### 2022-05-04 cmsb22.tex
Complete the proof of Theorem 2

@@
 \institute{
-  Aix-Marseille Université, Marseille, France\\
+  LIS, Aix-Marseille Université, Marseille, France\\
   \email{Trinh.Van-Giang@lis-lab.fr}
@@
   A set of places of Petri net \(P\) encoding logical model \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{gene} of \(\mathcal{M}\).
+Then, a conflict-free siphon \(S\) is said to be \emph{maximal} if and only if there is no other conflict-free siphon \(S'\) such that \(S \subset S'\).
 
@@
 
-  Let \(m\) be a subspace of Boolean model \(\mathcal{M}\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) such that:
+  Let \(m\) be a subspace of Boolean model \(\mathcal{M}\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) if and only if
   \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S.\] If \(D_m = \emptyset\), \(S\) is conventionally \(\emptyset\).
@@
   First, we show that if \(m\) is a minimal trap space of \(\mathcal{M}\), then \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\) (*).
+Since \(m\) is a trap space of \(\mathcal{M}\), \(S\) is a conflict-free siphon of \(\mathcal{P}\) by Theorem~\ref{theo:ts_2_sp}.
+Assume that \(S\) is not maximal.
+Then, there is another conflict-free siphon \(S'\) such that \(S \subset S'\).
+By Theorem~\ref{theo:ts_2_sp}, there is a trap space \(m'\) corresponding to \(S'\).
+Following the definition of a mirror, \(\mathcal{S}_{\mathcal{M}}[m'] \subset \mathcal{S}_{\mathcal{M}}[m]\), thus \(m' < m\).
+This is a contradiction because \(m\) is a minimal trap space.
+Hence, \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\).
 
   Second, we show that if \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\), then \(m\) is a minimal trap space of \(\mathcal{M}\) (**).
+Since \(S\) is a conflict-free siphon of \(\mathcal{P}\), \(m\) is a trap space of \(\mathcal{M}\) by Theorem~\ref{theo:ts_2_sp}.
+Assume that \(m\) is not minimal.
+Then, there is another trap space \(m'\) such that \(m' < m\).
+In other words, \(\mathcal{S}_{\mathcal{M}}[m'] \subset \mathcal{S}_{\mathcal{M}}[m]\).
+Let \(S'\) be the mirror of \(m'\).
+\(S'\) is a conflict-free siphon by Theorem~\ref{theo:ts_2_sp}.
+Following the definition of a mirror, \(S \subset S'\), which is a contradiction because \(S\) is a maximal conflict-free siphon.
+Hence, \(m\) is a minimal trap space of \(\mathcal{M}\).
 

### 2022-05-04 cmsb22.tex
Complete the proof of Theorem 1, update the definition of a mirror

@@
 
-  Let \(m\) be a subspace of Boolean model \(\mathcal{M}\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) if and only if
-  \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S.\] If \(D_m = \emptyset\), \(S\) is conventionally \(\emptyset\).
+  Let \(m\) be a subspace of Boolean model \(\mathcal{M}\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) such that:
+  \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S\] and \[\forall v \not \in D_m, p_v \not \in S, \overline{p}_v \not \in S.\] If \(D_m = \emptyset\), \(S\) is conventionally \(\emptyset\).
 
@@
   First, we show that if \(m\) is a trap space of \(\mathcal{M}\), then \(S\) is a conflict-free siphon of \(\mathcal{P}\) (*).
+If \(D_m = \emptyset\), then \(S = \emptyset\) is trivially a conflict-free siphon of \(\mathcal{P}\).
+Thus, we consider the case that \(D_m \neq \emptyset\) (resp.
+\(S \neq \emptyset\)).
+Assume that \(S\) is not a siphon of \(\mathcal{P}\).
+Then, there is a transition \(t \in T\) such that \(S\cap succ(t)\not =\emptyset\) but \(S\cap pred(t)\not =\emptyset\).
+In other words, there is a place \(p \in S\) such that \(p \in succ(t)\) but \(p \not \in pred(t)\).
+Let \(v\) be the corresponding node in \(\mathcal{M}\) of \(p\).
+By the characterization of the encoding~\cite{chatain2014characterization}, there is a directional arc from \(t\) to \(p\) and a directional arc from the complementary place of \(p\) to \(t\).
+Without loss of generality, we assume that \(p = p_v\), then there is a directional arc from \(t\) to \(p_v\) and a directional arc from \(\overline{p}_v\) to \(t\).
+In addition, there is also no arc or a bi-directional arc between \(t\) and another place rather than \(p_v\) and \(\overline{p}_v\).
+Thus, there is no connecting arc between \(t\) and any place in \(S \setminus \{p_v\}\) because \(S\cap pred(t)\not =\emptyset\).
+In \(\mathcal{S}_{\mathcal{M}}[m]\), a node in \(V \setminus D_m\) can receive any Boolean value.
+Hence, there is a state \(s \in \mathcal{S}_{\mathcal{M}}[m]\) such that \(m_s(p') = 1, \forall p' \in pred(t) \setminus \{\overline{p_v}\}\) where \(m_s\) is the corresponding marking in \(\mathcal{P}\) of \(s\).
+We also have \(m_s(p_v) = 0\), leading to \(m_s(\overline{p_v}) = 1\) by the characterization of the encoding~\cite{chatain2014characterization}.
+Now, \(t\) is enabled at marking \(m_s\).
+Its firing leads to a new marking \(m'_s\) such that \(m'_s(p_v) = 1\) and \(m'_s(\overline{p}_v) = 0\).
+Let \(s'\) be the corresponding state in \(\mathcal{M}\) of \(m'_s\).
+Since \(m\) is a trap space of \(\mathcal{M}\), \(s' \in \mathcal{S}_{\mathcal{M}}[m]\).
+Then, \(s'(v) = m(v)\), leading to \(m'_s(p_v) = 0\), which is a contradiction.
+Hence, \(S\) is a siphon of \(\mathcal{P}\).
+By the definition of a mirror, \(S\) is also a conflict-free one.
 
   Second, we show that if \(S\) is a conflict-free siphon of \(\mathcal{P}\), then \(m\) is a trap space of \(\mathcal{M}\) (**).
+By the definition of a mirror, \(m\) is a subspace of \(\mathcal{M}\).
+Let \(s\) be an arbitrary state in \(\mathcal{S}_{\mathcal{M}}[m]\) and \(m_s\) be its corresponding marking in \(\mathcal{P}\).
+By the characterization of the encoding~\cite{chatain2014characterization}, \(m_s(p) = 0, \forall p \in S\).
+In any marking \(m'_s\) reachable from \(m_s\) regardless of the update semantics of \(\mathcal{P}\), we have \(m'_s(p) = 0, \forall p \in S\) by the characterization of a siphon~\cite{DBLP:journals/isci/LiuB16}.
+Equivalently, in any state \(s'\) reachable from \(s\) regardless of the update semantics of \(\mathcal{M}\), we have \(s'(v) = s(v) = m(v), \forall v \in D_m\).
+Then, \(s' \in \mathcal{S}_{\mathcal{M}}[m]\).
+By the definition of a trap space and the arbitrariness of \(s\), \(m\) is a trap space of \(\mathcal{M}\).
 

### 2022-05-04 cmsb22.tex
Update the result table, update Section "Answer set programming-based method", update Section "Preliminaries"

@@
 
-\subsection{Traps-spaces}
+\subsection{Traps spaces}
 
-We recall here some definitions from~\cite{klarner2015computing} for the introduction of \emph{trap-spaces}.
+We recall here some definitions from~\cite{klarner2015computing} for the introduction of \emph{trap spaces}.
 Minimal trap-spaces prove to be a very good approximation of the attractors of a Boolean model and have become the \emph{de facto} standard way to analyze models of a few tens of \emph{genes}~\cite{klarner2017pyboolnet,cifuentes2020control}.
 
+%%% Trap set %%%
 Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\bbbb^{n}\) with \(\bbbb = \{0, 1\}\).
+A non-empty set \(T \subseteq S_{\mathcal{M}}\) is a \emph{trap set} of \(\mathcal{M}\) if and only if for every \(x \in T\) and \(y \in S_{\mathcal{M}}\) with \(y\) is reachable from \(x\) it holds that \(y \in T\).
 
@@
 
-\sylvain{missing example?}
+%%% Example %%%
+\begin{example}
+We give a BN \(\mathcal{N} = (V, F)\), where \(V = \{x_1, x_2\}\) and \(F = \{f_1, f_2\}\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
+Herein, \(\land\), \(\lor\), and \(\neg\) denote the CONJUNCTION, DISJUNCTION, and NEGATION logical operators, respectively.
+\label{example:BN}
+\end{example}
+
+%\sylvain{missing example?}
 
@@
 
-  A \emph{Petri net graph} is a weighted bipartite directed graph \((P, T, W)\),
+  A \emph{Petri net} is a weighted bipartite directed graph \((P, T, W)\),
   where \(P\) is a non-empty finite set of vertices called \emph{places},
@@
 
-We will write \(pred(x)\) (resp.\ \(succ(x)\)) to represent the set of vertices that have an (non-zero weighted) arc leading to (resp.\ coming from) \(x\).
+We shall write \(pred(x)\) (resp.\ \(succ(x)\)) to represent the set of vertices that have an (non-zero weighted) arc leading to (resp.\ coming from) \(x\).
 
@@
 Since our study is focused on Boolean models, we briefly recall the original encoding here.
-Its basis is that every \emph{gene} of the original model is represented by two separate places (\(p\) and \(\overline{p}\)), corresponding to its two states, active, and inactive.
-Each conjunct of the logical function that activate the \emph{gene} will lead to a transition, consuming the inactive place, producing the active place, and with all other literals both consumed and produced.
+Its basis is that every node (\emph{gene}) \(v\) of the original model is represented by two separate places (\(p_v\) and \(\overline{p}_v\)), corresponding to its two states, active, and inactive, respectively.
+Each conjunct of the logical function that activate the \emph{gene} will lead to a transition \(t\), consuming the inactive place (i.e., a directional arc from \(\overline{p}_v\) to \(t\)), producing the active place (i.e., a directional arc from \(t\) to \(p_v\)), and with all other literals both consumed and produced (i.e., a bi-directional arc).
 And conversely for the inactivation. 
+Let \(s\) be a state of the Boolean model and \(m_s\) be its the corresponding marking in the encoded Petri net.
+It holds that \(\forall v \in V\), \(s(v) = 0\) if and only if \(m_s(\overline{p}_v) = 1\) and \(s(v) = 1\) if and only if \(m_s(p_v) = 1\).
 
@@
 Let \(v\) be the corresponding node in \(\mathcal{M}\) of \(p\).
-By the characterization of the encoding~\cite{chatain2014characterization}, there is a directional arc from \(t\) to \(p\) and a directional arc from the complementary place of \(p\) to \(t\).
+By the characterization of the encoding~\cite{chaouiya2004qualitative}, there is a directional arc from \(t\) to \(p\) and a directional arc from the complementary place of \(p\) to \(t\).
 Without loss of generality, we assume that \(p = p_v\), then there is a directional arc from \(t\) to \(p_v\) and a directional arc from \(\overline{p}_v\) to \(t\).
@@
 Hence, there is a state \(s \in \mathcal{S}_{\mathcal{M}}[m]\) such that \(m_s(p') = 1, \forall p' \in pred(t) \setminus \{\overline{p_v}\}\) where \(m_s\) is the corresponding marking in \(\mathcal{P}\) of \(s\).
-We also have \(m_s(p_v) = 0\), leading to \(m_s(\overline{p_v}) = 1\) by the characterization of the encoding~\cite{chatain2014characterization}.
+We also have \(m_s(p_v) = 0\), leading to \(m_s(\overline{p_v}) = 1\) by the characterization of the encoding~\cite{chaouiya2004qualitative}.
 Now, \(t\) is enabled at marking \(m_s\).
@@
 Let \(s\) be an arbitrary state in \(\mathcal{S}_{\mathcal{M}}[m]\) and \(m_s\) be its corresponding marking in \(\mathcal{P}\).
-By the characterization of the encoding~\cite{chatain2014characterization}, \(m_s(p) = 0, \forall p \in S\).
+By the characterization of the encoding~\cite{chaouiya2004qualitative}, \(m_s(p) = 0, \forall p \in S\).
 In any marking \(m'_s\) reachable from \(m_s\) regardless of the update semantics of \(\mathcal{P}\), we have \(m'_s(p) = 0, \forall p \in S\) by the characterization of a siphon~\cite{DBLP:journals/isci/LiuB16}.
@@
 \section{Answer set programming-based method}
+First, we show the characterization of a conflict-free siphon of the encoded Petri net \(\mathcal{P}\). ...
 
-ASP
+Then, we translate the above characterization into the ASP \(\mathcal{A}\) as follows.
+We use atom \texttt{p(v, 1)} (resp.
+\texttt{p(v, 0)}) to denote place \(p_v\) (resp.
+\(\overline{p_v}\)), \(\forall v \in V\).
+The constraint (C2) is translated to the rule
+\[\verb|:- p(v, 1), p(v, 0).|\]
+for each \(v \in V\). ...
+
+Naturally, a stable model \(A\) of \(\mathcal{A}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a minimal trap space \(m\) of \(\mathcal{M}\).
+The conversion from \(A\) to \(m\) is straightforward.
+If \(\texttt{p(v, 1)} \in A\) then \(v \in D_m\) and \(m(v) = 0\).
+Conversely, if \(\texttt{p(v, 0)} \in A\) then \(v \in D_m\) and \(m(v) = 1\).
+Otherwise, \(v \not \in D_m\).
+Computing multiple solutions is built into ASP solvers and the solving collection POTASSCO~\cite{DBLP:journals/aicom/GebserKKOSS11} also features the option to find set-inclusion maximal solutions with respect to the predicates that are true.
+By using this built-in option, we can compute all the stable models of \(\mathcal{A}\) in one execution.
 
@@
 We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
+It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models. 
 To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
@@
 
-\sylvain{we need a source for the models as much as possible, or to refer to some place where people can access them to reproduce the benchmark.}
-\vangiang{Yes. I have added reference sources for the models but some models are still missing now.}
+% \sylvain{we need a source for the models as much as possible, or to refer to some place where people can access them to reproduce the benchmark.}
+% \vangiang{Yes. I have added reference sources for the models but some models are still missing now.}
 
@@
 Table~\ref{tab:result_real} shows the experimental results on real-world models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
+
+The first observation is that the proposed method much outperforms PyBoolNet in computational time. More specifically, ...
+\vangiang{In the journal version, we can measure the memory consumption of the two methods}
+
+The second observation is that ...
 
@@
     IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 149 & 15 & - & - & 512 & 0.32 \\
-    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & - & 16 & - & - & 5724 & 0.90 \\
+    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 246 & 16 & - & - & 5724 & 0.90 \\
     Korkut-2015~\cite{lee2019signal} & 99 & - & 12 & DNF & DNF & 2060 & 2.86 \\ \midrule
 
-    adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 493 & - & - & - & 78 & 3.72 \\
+    adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 493 & 4 & - & - & 78 & 3.72 \\
     interferon-1~\cite{ostaszewski2021covid19} & 121 & 190 & 55 & - & - & 512 & 0.43 \\

### 2022-05-04 trappist.py
Some kind of benchmark…
we do have some issues with BioLQM parsing

@@ -77,13 +77,13 @@ def write_asp(petri_net: nx.DiGraph, asp_file: IO):
                     print(f"{or_preds} :- {pnml_to_asp(succ)}.", file=asp_file)
 
 
-def solve_asp(asp_filename: str) -> str:
+def solve_asp(asp_filename: str, max_output: int) -> str:
     """Run an ASP solver on program asp_file and get the solutions."""
     # FIXME we need a limit on number of solutions
     result = subprocess.run(
         [
             "clingo",
-            "0",
+            str(max_output),
             "--heuristic=Domain",  # maximal w.r.t. inclusion
             "--enum-mod=domRec",
             "--dom-mod=3",
@@ -96,7 +96,8 @@ def solve_asp(asp_filename: str) -> str:
     )
 
     # https://www.mat.unical.it/aspcomp2013/files/aspoutput.txt
-    if result.returncode != 30:  # 30: SAT, all enumerated, optima found
+    # 30: SAT, all enumerated, optima found, 10 stopped by max
+    if result.returncode != 30 and result.returncode != 10:
         result.check_returncode()  # will raise CalledProcessError
 
     return result.stdout
@@ -147,7 +148,7 @@ def get_solutions(
 
 
 def compute_trap_spaces(
-    infile: IO, display: bool = False
+    infile: IO, display: bool = False, max_output: int = 0
 ) -> Optional[Generator[List[str], None, None]]:
     """Do the minimal trap-space computation on input file infile."""
     petri_net = read_pnml(infile)
@@ -155,7 +156,7 @@ def compute_trap_spaces(
     (_, tmpname) = tempfile.mkstemp(suffix=".lp", text=True)
     with open(tmpname, "wt") as asp_file:
         write_asp(petri_net, asp_file)
-    solutions = solve_asp(tmpname)
+    solutions = solve_asp(tmpname, max_output)
     os.unlink(tmpname)
     return get_solutions(solutions, petri_net, display)
 
@@ -172,6 +173,13 @@ def main():
         version="%(prog)s v{version}".format(version=version),
     )
     parser.add_argument(
+        "-m",
+        "--max-output",
+        type=int,
+        default=0,
+        help="Maximum number of solutions (0 for all).",
+    )
+    parser.add_argument(
         "infile",
         type=argparse.FileType("r", encoding="utf-8"),
         nargs="?",
@@ -180,7 +188,7 @@ def main():
     )
     args = parser.parse_args()
 
[…]

### 2022-05-04 trappist.py
I cannot write ^^'

@@ -174,7 +174,7 @@ def main():
     )
     parser.add_argument(
         "-m",
-        "--max-output",
+        "--max",
         type=int,
         default=0,
         help="Maximum number of solutions (0 for all).",
@@ -188,7 +188,7 @@ def main():
     )
     args = parser.parse_args()
 
-    compute_trap_spaces(args.infile, display=True, max_output=args.max - output)
+    compute_trap_spaces(args.infile, display=True, max_output=args.max)
 
 
 if __name__ == "__main__":

### 2022-05-04 cmsb22.tex
Preliminary motivating example

@@
 
+\section{Motivating example}
+
+For a few years now the second author has been collaborating with biologists who have been building huge detailed and annotated maps and now wish to analyze the dynamics of the corresponding models.
+One of the main maps studied this way is a map representing knowledge about the Rheumatoïd Arthritis~\cite{singh2018computational}, which was the main motivation for the development of a tool to automatically transform it into an executable Boolean model~\cite{aghamiri2020automated}.
+In the supplementary material of the paper, an excerpt of the map, focused around the apoptosis (cell death) module is transformed into a model of \emph{reasonable} size, namely 181 Boolean variables (model \verb|F5_RA_apoptosis_executable_module.sbml|).
+The study of such model, though, is a big hurdle.
+Indeed, as stated in the article about another model of the same size:
+\emph{``The size of the CaSQ-inferred MAPK model (181 nodes) made the calculation of stable states a non-realistic endeavour.''}
+
+In practice, even if there are many attractors in such a model, to validate it having only a sample of those was deemed useful.
+However, using a state-of-the-art tool like PyBoolNet~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
+And hence, it is not possible to extract any (complex) attractor at all.
+This is also true for the Alzheimer model also mentioned in that same article and originally from~\cite{ogishima2016alzpathway}, but actually not for the MAPK model for which the first trap spaces can be obtained in reasonable time.
+
+Using the method presented above, it is possible to convert the model to PNML in about 1s and to obtain the first 1000 minimal trap-spaces (including ones that contain more than one state) in a few milliseconds.
+Unfortunately since this was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap-spaces.
+
 \section{Evaluation}
@@
 
-\subsection{Detailed biological example}
-
-TBD
-
 \section{Conclusion}

### 2022-05-05 cmsb22.tex
Update Section "Preliminaries", finishi Section "Answer set programming-based method"

@@
 We recall here some definitions from~\cite{klarner2015computing} for the introduction of \emph{trap spaces}.
-Minimal trap-spaces prove to be a very good approximation of the attractors of a Boolean model and have become the \emph{de facto} standard way to analyze models of a few tens of \emph{genes}~\cite{klarner2017pyboolnet,cifuentes2020control}.
+Minimal trap spaces prove to be a very good approximation of the attractors of a Boolean model under asynchronous updating schemes and have become the \emph{de facto} standard way to analyze models of a few tens of \emph{genes}~\cite{klarner2017pyboolnet,cifuentes2020control}.
 
@@
 Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\bbbb^{n}\) with \(\bbbb = \{0, 1\}\).
+A state \(s \in \bbbb^{n}\) is a mapping \(s : V \mapsto \bbbb\) that assigns either 0 (inactive) or 1 (active) to each node.
 A non-empty set \(T \subseteq S_{\mathcal{M}}\) is a \emph{trap set} of \(\mathcal{M}\) if and only if for every \(x \in T\) and \(y \in S_{\mathcal{M}}\) with \(y\) is reachable from \(x\) it holds that \(y \in T\).
@@
 %%% Subspace %%%
-A \emph{subspace} \(m\) of \(\mathcal{S}_{\mathcal{M}}\) is characterized by its fixed variables (denoted by \(D_m\)) and free variables.
+A \emph{subspace} \(m\) of \(\mathcal{S}_{\mathcal{M}}\) is characterized by its fixed nodes (denoted by \(D_m\)) and free nodes.
 This subspace can be specified by an assignment \(D_m \mapsto \bbbb\), also denoted by \(m\) by abuse of notation, and where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
-The remaining variables of \(\mathcal{M}\), \(V \setminus D_m\), are said to be \emph{free}, i.e., they can receive any Boolean value.
-We write subspaces like states but use in addition the symbol \(\star\) to indicate that a variable is free.
+The remaining nodes of \(\mathcal{M}\), \(V \setminus D_m\), are said to be \emph{free}, i.e., they can receive any Boolean value.
+We write subspaces like states but use in addition the symbol \(\star\) to indicate that a node is free.
 A subspace \(m\) thus corresponds to the set of states \(\mathcal{S}_{\mathcal{M}}[m] := \{s \in \mathcal{S}_{\mathcal{M}}\;|\;\forall x_i \in D_m : s_i = m(x_i)\}\).
@@
 Consequently, a trap space \(m\) is minimal if and only if there is no trap space \(m' \in \mathcal{S}_{\mathcal{M}}^{\star}\) such that \(m' < m\).
-For example, the BN shown in Example~\ref{example:BN} has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\). Since \(m_1 < m_2\), \(m_1\) is a minimal trap space of the BN.
+For example, the Boolean model shown in Example~\ref{example:BN} has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
+Since \(m_1 < m_2\), \(m_1\) is a minimal trap space of the Boolean model.
 
@@
 \begin{example}
-We give a BN \(\mathcal{N} = (V, F)\), where \(V = \{x_1, x_2\}\) and \(F = \{f_1, f_2\}\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
+We give a Boolean model \(\mathcal{M} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
 Herein, \(\land\), \(\lor\), and \(\neg\) denote the CONJUNCTION, DISJUNCTION, and NEGATION logical operators, respectively.
@@
 \end{definition}
-
+A \emph{marking} for a Petri net is a mapping \(m : P \mapsto \bbbn\) that assigns a number of tokens to each place.
+A place \(p\) is marked by a marking \(m\) if and only if \(m(p) > 0\).
 We shall write \(pred(x)\) (resp.\ \(succ(x)\)) to represent the set of vertices that have an (non-zero weighted) arc leading to (resp.\ coming from) \(x\).
@@
 Since our study is focused on Boolean models, we briefly recall the original encoding here.
-Its basis is that every node (\emph{gene}) \(v\) of the original model is represented by two separate places (\(p_v\) and \(\overline{p}_v\)), corresponding to its two states, active, and inactive, respectively.
-Each conjunct of the logical function that activate the \emph{gene} will lead to a transition \(t\), consuming the inactive place (i.e., a directional arc from \(\overline{p}_v\) to \(t\)), producing the active place (i.e., a directional arc from \(t\) to \(p_v\)), and with all other literals both consumed and produced (i.e., a bi-directional arc).
+Its basis is that every node (\emph{gene}) \(v\) of the original model \(\mathcal{M} = (V, F)\) is represented by two separate places (\(p_v\) and \(\overline{p}_v\)), corresponding to its two states, active, and inactive, respectively.
+Each conjunct of the logical function that activates the \emph{gene} will lead to a transition \(t\), consuming the inactive place (i.e., a directional arc from \(\overline{p}_v\) to \(t\)), producing the active place (i.e., a directional arc from \(t\) to \(p_v\)), and with all other literals both consumed and produced (i.e., a bi-directional arc).
 And conversely for the inactivation. 
@@
 It holds that \(\forall v \in V\), \(s(v) = 0\) if and only if \(m_s(\overline{p}_v) = 1\) and \(s(v) = 1\) if and only if \(m_s(p_v) = 1\).
+Note also that at any marking \(m\) of the Petri net encoding a Boolean model, it always holds that \(m(p_v) + m(\overline{p}_v) = 1\).
 
@@
 
-Siphons are a classical property of place-transition nets~\cite{peterson1981petri}.
+Siphons are a static and classical property of Petri nets~\cite{peterson1981petri}.
 Note however that the use of siphons for the analysis of biological models, though it is not new, has been mostly relevant to the ODE-based continuous semantics of Chemical Reaction Networks~\cite{angeli2007petri,angeli2011persistence,degrand2020graphical}.
@@
 
-  Let \(m\) be a subspace of Boolean model \(\mathcal{M}\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) such that:
-  \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S\] and \[\forall v \not \in D_m, p_v \not \in S, \overline{p}_v \not \in S.\] If \(D_m = \emptyset\), \(S\) is conventionally \(\emptyset\).
+  Let \(m\) be a subspace of Boolean model \(\mathcal{M} = (V, F)\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) such that:
+  \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S\] and \[\forall v \in V \setminus D_m, p_v \not \in S, \overline{p}_v \not \in S.\]
 
@@
 
-  Let \(\mathcal{M}\) be a Boolean model and \(\mathcal{P}\) be its Petri net encoding.
+  Let \(\mathcal{M} = (V, F)\) be a Boolean model and \(\mathcal{P}\) be its Petri net encoding.
 A subspace \(m\) is a trap space of \(\mathcal{M}\) if and only if its mirror \(S\) is a conflict-free siphon of \(\mathcal{P}\).
@@
 By the characterization of the encoding~\cite{chaouiya2004qualitative}, \(m_s(p) = 0, \forall p \in S\).
-In any marking \(m'_s\) reachable from \(m_s\) regardless of the update semantics of \(\mathcal{P}\), we have \(m'_s(p) = 0, \forall p \in S\) by the characterization of a siphon~\cite{DBLP:journals/isci/LiuB16}.
+In any marking \(m'_s\) reachable from \(m_s\) regardless of the update semantics of \(\mathcal{P}\), we have \(m'_s(p) = 0, \forall p \in S\) by the dynamical property on markings of a siphon~\cite{DBLP:journals/isci/LiuB16}.
 Equivalently, in any state \(s'\) reachable from \(s\) regardless of the update semantics of \(\mathcal{M}\), we have \(s'(v) = s(v) = m(v), \forall v \in D_m\).
@@
 \section{Answer set programming-based method}
-First, we show the characterization of a conflict-free siphon of the encoded Petri net \(\mathcal{P}\). ...
+First, we show the characterization of all conflict-free siphons of the encoded Petri net \(\mathcal{P} = (P, T, W)\).
+Suppose that \(S\) is a generic siphon of \(\mathcal{P}\).
+If a place \(p\) should belong to \(S\), then by definition all the transitions in \(pred(p)\) must belong to \(succ(S)\).
+Note that \(succ(S) = \bigcup_{p \in S}succ(p)\).
+A transition \(t\) belongs to \(post(S)\) if and only if there is at least one place \(p'\) in \(S\) such that \(p' \in pred(t)\).
+Hence, for each transition \(t \in pred(p)\), we can state that
+\begin{equation}
+\label{eq:siphon}
+p \in S \Rightarrow \bigvee_{p' \in pred(t)}p' \in S.
+\end{equation}The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net~\cite{nabli2016enumerating}.
+To make \(S\) to be a conflict-free siphon, we need to add the following rule for each node \(v \in V\) to the system.
+\begin{equation}
+\label{eq:conflict}
+p_v \in S \Rightarrow \overline{p_v} \not \in S \wedge \overline{p_v} \in S \Rightarrow p_v \not \in S
+\end{equation}By definition, the final system fully characterizes all conflict-free siphons of the encoded Petri net.
 
 Then, we translate the above characterization into the ASP \(\mathcal{A}\) as follows.
-We use atom \texttt{p(v, 1)} (resp.
-\texttt{p(v, 0)}) to denote place \(p_v\) (resp.
+We use atom \verb|p(v, 1)| (resp.
+\verb|p(v, 0)|) to denote place \(p_v\) (resp.
 \(\overline{p_v}\)), \(\forall v \in V\).
-The constraint (C2) is translated to the rule
+For each pair \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\), we translate the rule (\ref{eq:siphon}) into the ASP rule
+\[
+\verb|p(v_1, x_1); ... ; p(v_k, x_k) :- p(v, x).|
+\] where \verb|p(v, x)| is the atom representing place \(p\) and \(\{\verb|p(v_1, x_1)|, ..., \verb|p(v_k, x_k)|\}\) is the set of atoms representing places in \(pred(t)\).
+The rule (\ref{eq:conflict}) is translated into the ASP rule
 \[\verb|:- p(v, 1), p(v, 0).|\]
-for each \(v \in V\). ...
+for each \(v \in V\).
+This rule guarantees that \verb|p(v, 1)| and \verb|p(v, 0)| never belong to the same answer set of \(\mathcal{A}\), representing the conflict-freeness.
+Note that the number of atoms of \(\mathcal{A}\) is only \(2n\), which is theoretically advanced to the ASP encoding shown in~\cite{klarner2015computing} where the number of atoms of the encoded ASP equals to the number of prime-implicants of the Boolean model.
 
-Naturally, a stable model \(A\) of \(\mathcal{A}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a minimal trap space \(m\) of \(\mathcal{M}\).
+Naturally, a solution (simply an answer set) \(A\) of \(\mathcal{A}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a minimal trap space \(m\) of \(\mathcal{M}\).
 The conversion from \(A\) to \(m\) is straightforward.
-If \(\texttt{p(v, 1)} \in A\) then \(v \in D_m\) and \(m(v) = 0\).
-Conversely, if \(\texttt{p(v, 0)} \in A\) then \(v \in D_m\) and \(m(v) = 1\).
+If \(\verb|p(v, 1)| \in A\) then \(v \in D_m\) and \(m(v) = 0\).
+Conversely, if \(\verb|p(v, 0)| \in A\) then \(v \in D_m\) and \(m(v) = 1\).
 Otherwise, \(v \not \in D_m\).
-Computing multiple solutions is built into ASP solvers and the solving collection POTASSCO~\cite{DBLP:journals/aicom/GebserKKOSS11} also features the option to find set-inclusion maximal solutions with respect to the predicates that are true.
-By using this built-in option, we can compute all the stable models of \(\mathcal{A}\) in one execution.
+Computing multiple answer sets is built into ASP solvers and the solving collection POTASSCO~\cite{DBLP:journals/aicom/GebserKKOSS11} also features the option to find set-inclusion maximal answer sets with respect to the predicates that are true.
+By using this built-in option, we can compute all the answer sets of \(\mathcal{A}\) in one execution.
 
@@
 To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet}.
-We used a set of real-world Boolean models lying in various scales obtained from the literature.
+We used a set of real-world Boolean models lying in various scales collected from various bibliographic sources.
 These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.

### 2022-05-05 cmsb22.tex

@@
 \author{Van-Giang Trinh\inst{1}\orcidID{0000-0001-6581-998X} \and \\
-  Sylvain Soliman\inst{2}\orcidID{0000-0001-5525-7418}}
-% \authorrunning{F. Author et al.}
+  Sylvain Soliman\inst{2}\orcidID{0000-0001-5525-7418} \and \\
+  Kunihiko Hiraishi\inst{3}\orcidID{0000-0003-1750-1891} \and \\ Belaid Benhamou\inst{1}
+  }
+
+\authorrunning{Trinh et al.}
+
+\vangiang{I added two other authors.
+Prof.
+Belaid Benhamou is currently my postdoctoral supervisor and he will also involve in the topic on siphons.
+Prof.
+Kunihiko Hiraishi was my doctoral supervisor.
+I obtained the method for trap space computation when I was a researcher in his lab and he also contributes to this method.
+Of course, they will involve in revising the papers.
+Is this OK for you?}
 
@@
   LIS, Aix-Marseille Université, Marseille, France\\
-  \email{Trinh.Van-Giang@lis-lab.fr}
+  \email{\{Trinh.Van-Giang, belaid.benhamou\}@lis-lab.fr}
   \and
@@
   %\url{http://lifeware.inria.fr/~soliman/}
+  \and
+  School of Information Science, Japan Advanced Institute of Science and Technology, Japan\\
+  \email{hira@jaist.ac.jp}
 }

### 2022-05-05 cmsb22.tex
Quick pass, title change, much more tomorrow

@@
 \begin{document}
-%
-\title{Minimal trap-spaces of Logical models are maximal siphons of their Petri net encoding}
-\titlerunning{Minimal trap-spaces as maximal siphons}
+
+% \title{Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding}
+% \titlerunning{Minimal trap spaces as maximal siphons}
+\title{Scaling-up attractor computation for Boolean models}
+
 \author{Van-Giang Trinh\inst{1}\orcidID{0000-0001-6581-998X} \and \\
@@
 
-\vangiang{I added two other authors.
-Prof.
-Belaid Benhamou is currently my postdoctoral supervisor and he will also involve in the topic on siphons.
-Prof.
-Kunihiko Hiraishi was my doctoral supervisor.
-I obtained the method for trap space computation when I was a researcher in his lab and he also contributes to this method.
-Of course, they will involve in revising the papers.
-Is this OK for you?}
-
 \institute{
@@
   Lifeware team, Inria Saclay center, Palaiseau, France\\
-  \email{Sylvain.Soliman@inria.fr}%\\
-  %\url{http://lifeware.inria.fr/~soliman/}
+  \email{Sylvain.Soliman@inria.fr}
   \and
@@
   This is even more true for very big models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
-  The analysis of such models is mostly based on attractor computation, since those correspond roughly to \emph{phenotypes}, and the recent use of trap-spaces made a real breakthrough in that domain allowing to consider medium-sized models that used to be out of reach.
-  However, with the continuing increase in model-size, the state of the art computation of minimal trap-spaces based on prime-implicants shows its limits as there can be very many such implicants.
-  In this article we present an alternative method to compute minimal trap-spaces, and hence complex attractors, that relies on a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original Logical model.
+  The analysis of such models is mostly based on attractor computation, since those correspond roughly to \emph{phenotypes}, and the recent use of trap spaces made a real breakthrough in that domain allowing to consider medium-sized models that used to be out of reach.
+  However, with the continuing increase in model-size, the state of the art computation of minimal trap spaces based on prime-implicants shows its limits as there can be very many such implicants.
+  In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, that relies on a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original Logical model.
   We then demonstrate its efficiency and compare it to implicant-based methods on a few relevant big Boolean models.
 
-\keywords{Logical models \and Boolean models \and Trap-spaces \and Attractor computation \and Petri nets \and Siphons}
+\keywords{Logical models \and Boolean models \and Trap spaces \and Attractor computation \and Petri nets \and Siphons}
 \end{abstract}
@@
 \section{Introduction}
-
-\sylvain{We could go for a completely different title, e.g., focused on performance much more, like ``Scaling-up attractor computation for Logical models''}
 
@@
 This is even more true for very big models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas~\cite{aghamiri2020automated}.
-The analysis of such models is mostly based on attractor computation, since those correspond roughly to \emph{phenotypes}, and the recent use of trap-spaces~\cite{klarner2015computing} made a real breakthrough in that domain allowing to consider medium-sized models that used to be out of reach.
-However, with the continuing increase in model-size, the state of the art computation of minimal trap-spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of such implicants.
+The analysis of such models is mostly based on attractor computation, since those correspond roughly to \emph{phenotypes}, and the recent use of trap spaces~\cite{klarner2015computing} made a real breakthrough in that domain allowing to consider medium-sized models that used to be out of reach.
+However, with the continuing increase in model-size, the state of the art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of such implicants.
 
@@
 
-In this article we present an alternative method to compute minimal trap-spaces, and hence complex attractors, that replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original Logical model.
+In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, that replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original Logical model.
 We then demonstrate its efficiency and compare it to implicant-based methods on a few relevant big Boolean models.
@@
 
-TBD.
-Purely Boolean for trap-spaces, asynchronous for encoding.
+We will briefly recall here some preliminaries on Boolean models related to trap spaces and Petri nets.
+To stay in line with the usual considerations we will always assume purely Boolean (and not multi-valued) logical models, and an asynchronous update.
+In the case of multi-level models, an encoding into a Boolean model is always possible (and once again available through BioLQM).
+As far as different update strategies are concerned, we think that even if they might require small technical adaptations, they would not bring much change to the main results of this article.
 
@@
 \end{example}
-
-%\sylvain{missing example?}
 
@@
 
+Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining the Disjunctive Normal Forms (DNF) of each Boolean function.
+Though this might be quite computationally intensive it is important to remark first that contrary to the Prime implicants case, there is no need to find \emph{minimal} DNF.
+Second, in practice this transformation seems to behave quite well on even quite big models as will be shown in the Section~\ref{sec:eval} on evaluation.
+Currently, relying on BioLQM implies going through a BDD encoding of the formulae, which might not be optimal.
+Hence we are still considering a direct encoding that would skip the BDD construction.
+
 \subsection{Siphons}
@@
 If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the empty places remain empty).
-This is precisely the definition of a trap-space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap-space.
+This is precisely the definition of a trap space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap space.
 Hereafter, we formally prove that a maximal conflict-free siphon is equivalent to a minimal trap space.
@@
 \label{theo:min_ts_2_max_sp}
-  %\(S\) is a minimal trap-space of Boolean model \(\mathcal{M}\) if and only if the places corresponding to the fixed values in \(S\) (active or inactive) are a maximal conflict-free siphon of \((P, T, W)\), the Petri net encoding of \(S\).
+  %\(S\) is a minimal trap space of Boolean model \(\mathcal{M}\) if and only if the places corresponding to the fixed values in \(S\) (active or inactive) are a maximal conflict-free siphon of \((P, T, W)\), the Petri net encoding of \(S\).
 
@@
 
-Using the method presented above, it is possible to convert the model to PNML in about 1s and to obtain the first 1000 minimal trap-spaces (including ones that contain more than one state) in a few milliseconds.
-Unfortunately since this was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap-spaces.
+Using the method presented above, it is possible to convert the model to PNML in about 1s and to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) in a few milliseconds.
+Unfortunately since this was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.
 
 \section{Evaluation}
+\label{sec:eval}
 To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet}.
@@
 Finally, we set a time limit as one hour for each model.
-
-% \sylvain{we need a source for the models as much as possible, or to refer to some place where people can access them to reproduce the benchmark.}
-% \vangiang{Yes. I have added reference sources for the models but some models are still missing now.}
 
@@
 
-% \sylvain{I changed the ``look'' of the table using booktabs, see \url{https://nhigham.com/2019/11/19/better-latex-tables-with-booktabs/} for instance.}
-% \vangiang{Actually, booktabs looks better. I have never used it before. Thank you for sharing.}
-
-% \sylvain{what is $|F|$? what are t1 and t2? why are there no times for BoolNet?}
-% \vangiang{I need to open an excel file containing the experimental results but I am unable to access this file now.}
-
 \end{table}

### 2022-05-06 cmsb22.tex
Many small things, plus conclusion

@@
 We will briefly recall here some preliminaries on Boolean models related to trap spaces and Petri nets.
-To stay in line with the usual considerations we will always assume purely Boolean (and not multi-valued) logical models, and an asynchronous update.
-In the case of multi-level models, an encoding into a Boolean model is always possible (and once again available through BioLQM).
+To stay in line with the usual considerations we will always assume purely Boolean logical models, and an asynchronous update.
+In the case of multi-level models, an encoding into a Boolean model is always possible.
 As far as different update strategies are concerned, we think that even if they might require small technical adaptations, they would not bring much change to the main results of this article.
@@
 A \emph{subspace} \(m\) of \(\mathcal{S}_{\mathcal{M}}\) is characterized by its fixed nodes (denoted by \(D_m\)) and free nodes.
-This subspace can be specified by an assignment \(D_m \mapsto \bbbb\), also denoted by \(m\) by abuse of notation, and where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
+This subspace can be specified by an assignment \(D_m \mapsto \bbbb\) where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
 The remaining nodes of \(\mathcal{M}\), \(V \setminus D_m\), are said to be \emph{free}, i.e., they can receive any Boolean value.
@@
 Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining the Disjunctive Normal Forms (DNF) of each Boolean function.
-Though this might be quite computationally intensive it is important to remark first that contrary to the Prime implicants case, there is no need to find \emph{minimal} DNF.
+Though this might be quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
 Second, in practice this transformation seems to behave quite well on even quite big models as will be shown in the Section~\ref{sec:eval} on evaluation.
@@
 
-We recall here the basic definition establishing that to produce something in a siphon you must consume something from the siphon:
+We recall here the basic definition establishing that to produce something in a siphon you must consume something from the siphon.
+This corresponds to the idea that a siphon is a set of places that once empty remains empty.
 
@@
   A \emph{siphon} of a Petri net \((P, T, W)\) is a set of places \(S\) such that:
-  \[\forall t\in T, S\cap succ(t)\not =\emptyset\Rightarrow S\cap pred(t)\not =\emptyset.\] \(\emptyset\) is trivially a siphon.
+  \[\forall t\in T, S\cap succ(t)\not =\emptyset\Rightarrow S\cap pred(t)\not =\emptyset.\]
 
@@
 
-Intuitively a siphon is a set of places that once empty remains empty.
+Note that \(\emptyset\) is trivially a siphon.
 
@@
 Hence, we here propose a new method based on Answer Set Programming (ASP)~\cite{DBLP:journals/aicom/GebserKKOSS11} for computing maximal conflict-free siphons of a Petri net.
-The details of the proposed method shall be shown in the next section.
+The details of the proposed method shall be given in the next section.
 
@@
 Note that \(succ(S) = \bigcup_{p \in S}succ(p)\).
-A transition \(t\) belongs to \(post(S)\) if and only if there is at least one place \(p'\) in \(S\) such that \(p' \in pred(t)\).
+A transition \(t\) belongs to \(succ(S)\) if and only if there is at least one place \(p'\) in \(S\) such that \(p' \in pred(t)\).
 Hence, for each transition \(t \in pred(p)\), we can state that
@@
 
+\sylvain{What do you mean by \emph{advanced}? It is less since there might be an exponential number of prime implicants?}
+
 Naturally, a solution (simply an answer set) \(A\) of \(\mathcal{A}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a minimal trap space \(m\) of \(\mathcal{M}\).
@@
 
-In practice, even if there are many attractors in such a model, to validate it having only a sample of those was deemed useful.
+In practice, even if there is a huge number of attractors in such a model, obtaining a sample of those can reveal very useful to invalidate the model and lead to further refinement.
 However, using a state-of-the-art tool like PyBoolNet~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
@@
 This is also true for the Alzheimer model also mentioned in that same article and originally from~\cite{ogishima2016alzpathway}, but actually not for the MAPK model for which the first trap spaces can be obtained in reasonable time.
+The current practice usually revolves then around fixing some inputs to plausible values and reducing the model accordingly.
+While this approach makes sense, it relies on potentially arbitrary decisions, and \emph{hides away} critical modelling choices that were actually not part of the original Boolean model.
 
@@
 
-\subsection{Benchmark vs.\ implicant-based methods}
+% \subsection{Benchmark vs.\ implicant-based methods}
 
@@
 
-Future work: comparison with SAT/MaxSAT based approaches~\cite{nabli2016enumerating}
+In this article we proposed a new method for the computation of minimal trap-spaces of Boolean models, based on maximal siphons.
+This method is evaluated on big models from the litterature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
+We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ONMKPO21molecular}.
 
-Also to a BDD-based approach using BioLQM is proposed at \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} but because of the BDD conversion, some models are not even loaded.
-This is also an issue for using BioLQM as a translator/preprocessor and since we do not use BDDs we might rely on writing our own parser for SBML-Qual and such formats.
+Though we benchmarked this new approach against the state-of-the-art tool PyBoolNet, there are many more evaluations that we plan to do in the future.
+First, the BioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
+Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
 
-\subsubsection{Acknowledgments}
-Coffee is good, you should drink some!
+Moreover, there are other methods for the maximal siphon computation, like SAT/MaxSAT approaches~\cite{nabli2016enumerating} that could replace our ASP program if they outperform it.
+
+However, the current method already appears to perform very well even on the biggest models we have considered.
+We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.
+
+% \subsubsection{Acknowledgments}
 
@@
 \end{document}
-
-% \noindent Displayed equations are centered and set on a separate
-% line.
-% \begin{equation}
-% x + y = z
-% \end{equation}
-% Please try to avoid rasterized images for line-art diagrams and
-% schemas. Whenever possible, use vector graphics instead (see
-% Fig.~\ref{fig1}).
-
-% \begin{figure}
-% % \includegraphics[width=\textwidth]{fig1.eps}
-% \caption{A figure caption is always placed below the illustration.
-% Please note that short captions are centered, while long ones are
-% justified by the macro package automatically.} \label{fig1}
-% \end{figure}

### 2022-05-06 cmsb22.tex
Add some small comments

--- 2022-05-05 cmsb22.tex
@@
 \title{Scaling-up attractor computation for Boolean models}
+\vangiang{I like the first title :).
+Since we only conduct experiments on real-world Boolean models of not too large size, it seems to make less sense to use "scaling-up" (this should be for the journal version).}
 
@@
 As far as different update strategies are concerned, we think that even if they might require small technical adaptations, they would not bring much change to the main results of this article.
+\vangiang{Minimal trap spaces are a static property of a BN regardless of its update semantics.
+All the results presented in this paper still valid for any update semantics.
+Hence, we do not need to restrict to an asynchronous update.}
 
@@
 
-Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining the Disjunctive Normal Forms (DNF) of each Boolean function.
+Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining the Disjunctive Normal Forms (DNF) of each Boolean function.\vangiang{I think it is better to put a reference here for this argument.}
 Though this might be quite computationally intensive it is important to remark first that contrary to the Prime implicants case, there is no need to find \emph{minimal} DNF.

### 2022-05-06 cmsb22.tex
Biblio and table position

--- 2022-05-06 cmsb22.tex
@@
 \label{sec:eval}
-To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet}.
-We used a set of real-world Boolean models lying in various scales collected from various bibliographic sources.
-These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
-We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
-It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models. 
-To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
-Specifically, we used the configuration \texttt{--dom-pref=32, --heu=domain, --dom-mod=7} (subset maximality).
-We ran all the benchmarks on a virtual machine whose environment is CPU: Intel(R) Core(TM) i7-3630QM 2.40GHz x 4, Memory: 8 GB, Ubuntu 18.04.2 64 bit.
-Finally, we set a time limit as one hour for each model.
 
-% \subsection{Benchmark vs.\ implicant-based methods}
-
-Table~\ref{tab:result_real} shows the experimental results on real-world models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
-
-The first observation is that the proposed method much outperforms PyBoolNet in computational time. More specifically, ...
-\vangiang{In the journal version, we can measure the memory consumption of the two methods}
-
-The second observation is that ...
-
-\begin{table}[!ht]
+\begin{table}[!htb]
   \caption{Timing comparisons between the proposed method and PyBoolNet on the selected real-world models.
@@
 
+To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet}.
+We used a set of real-world Boolean models lying in various scales collected from various bibliographic sources.
+These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
+We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
+It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
+To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
+Specifically, we used the configuration \texttt{--dom-pref=32, --heu=domain, --dom-mod=7} (subset maximality).
+We ran all the benchmarks on a virtual machine whose environment is CPU: Intel(R) Core(TM) i7-3630QM 2.40GHz x 4, Memory: 8 GB, Ubuntu 18.04.2 64 bit.
+Finally, we set a time limit as one hour for each model.
+
+% \subsection{Benchmark vs.\ implicant-based methods}
+
+Table~\ref{tab:result_real} shows the experimental results on real-world models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
+
+The first observation is that the proposed method much outperforms PyBoolNet in computational time. More specifically, ...
+\vangiang{In the journal version, we can measure the memory consumption of the two methods}
+
+The second observation is that ...
+
 \section{Conclusion}
@@
 This method is evaluated on big models from the litterature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
-We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ONMKPO21molecular}.
+We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
 

### 2022-05-06 cmsb22.tex

@@
 
-Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining the Disjunctive Normal Forms (DNF) of each Boolean function.
-Though this might be quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
 Second, in practice this transformation seems to behave quite well on even quite big models as will be shown in the Section~\ref{sec:eval} on evaluation.
@@
 Hence we are still considering a direct encoding that would skip the BDD construction.
+
+\vangiang{I think it is better to put a reference here for this argument.}
 

### 2022-05-06 cmsb22.tex
Answers and cleanup… will work on the abstract now

--- 2022-05-06 cmsb22.tex
@@
 
-% \title{Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding}
-% \titlerunning{Minimal trap spaces as maximal siphons}
-\title{Scaling-up attractor computation for Boolean models}
+\title{Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding}
+\titlerunning{Minimal trap spaces as maximal siphons}
+% \title{Scaling-up attractor computation for Boolean models}
+% \vangiang{I like the first title :).
+Since we only conduct experiments on real-world Boolean models of not too large size, it seems to make less sense to use "scaling-up" (this should be for the journal version).}
 
@@
 We will briefly recall here some preliminaries on Boolean models related to trap spaces and Petri nets.
-To stay in line with the usual considerations we will always assume purely Boolean logical models, and an asynchronous update.
-In the case of multi-level models, an encoding into a Boolean model is always possible.
-As far as different update strategies are concerned, we think that even if they might require small technical adaptations, they would not bring much change to the main results of this article.
+In the case of multi-level Logical models, an encoding into a Boolean model is always possible.
 
@@
 
-Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining the Disjunctive Normal Forms (DNF) of each Boolean function.
-Though this might be quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
-Second, in practice this transformation seems to behave quite well on even quite big models as will be shown in the Section~\ref{sec:eval} on evaluation.
-Currently, relying on BioLQM implies going through a BDD encoding of the formulae, which might not be optimal.
-Hence we are still considering a direct encoding that would skip the BDD construction.
+Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining conditions for the activation/inactivation of the states.
+In~\cite{chaouiya2004qualitative} this took the form of the whole truth table of the Boolean functions, but as shown in Appendix 1 of~\cite{chatain2014characterization} computing Disjunctive Normal Forms (DNF) of each Boolean function is enough.
+Though this might appear quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
+Second, in practice this transformation, here using BDDs in BioLQM, seems to behave quite well on even quite big models as will be shown in the Section~\ref{sec:eval} on evaluation.
 
@@
 
-\sylvain{What do you mean by \emph{advanced}? It is less since there might be an exponential number of prime implicants?}
+\sylvain{What do you mean by \emph{advanced}? It is less since there might be an exponential number of prime-implicants?}
 
@@
 Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
+The tradeoff between a small Petri net, and hence small ASP, and the time it takes to compute it has to be evaluated in depth.
 

### 2022-05-06 cmsb22.tex
Abstract

@@
 
-  Boolean modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
-  This is even more true for very big models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
-  The analysis of such models is mostly based on attractor computation, since those correspond roughly to \emph{phenotypes}, and the recent use of trap spaces made a real breakthrough in that domain allowing to consider medium-sized models that used to be out of reach.
-  However, with the continuing increase in model-size, the state of the art computation of minimal trap spaces based on prime-implicants shows its limits as there can be very many such implicants.
-  In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, that relies on a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original Logical model.
-  We then demonstrate its efficiency and compare it to implicant-based methods on a few relevant big Boolean models.
+  Boolean modelling of gene regulation but also post-transcri\-ptomic models has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
+  This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
+  Besides simulation, the analysis of such models is mostly based on attractor computation, since those correspond roughly to observable biological \emph{phenotypes}.
+The recent use of trap spaces made a real breakthrough in that field allowing to consider medium-sized models that used to be out of reach.
+  However, with the continuing increase in model-size, the state of the art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of implicants.
+
+  In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, of a Boolean model.
+It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
+  After some technical preliminaries, we expose the concrete need for such a method and detail its implementation using Answer Set Programming.
+  We then demonstrate its efficiency and compare it to implicant-based methods on some large Boolean models from the litterature.
 
@@
 
-Boolean modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}.
-This is even more true for very big models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas~\cite{aghamiri2020automated}.
-The analysis of such models is mostly based on attractor computation, since those correspond roughly to \emph{phenotypes}, and the recent use of trap spaces~\cite{klarner2015computing} made a real breakthrough in that domain allowing to consider medium-sized models that used to be out of reach.
-However, with the continuing increase in model-size, the state of the art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of such implicants.
+Boolean modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}, even for modelling post-transcriptional mechanisms.
+This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas~\cite{aghamiri2020automated}.
+Besides simulation, the analysis of such models is mostly based on attractor computation, since those correspond roughly to observable biological \emph{phenotypes}.
+The recent use of trap spaces~\cite{klarner2015computing} made a real breakthrough in that field allowing to consider medium-sized models that used to be out of reach.
+However, with the continuing increase in model-size, the state of the art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of implicants.
 
@@
 
-In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, that replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original Logical model.
-We then demonstrate its efficiency and compare it to implicant-based methods on a few relevant big Boolean models.
+In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, of a Boolean model.
+It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
+After some technical preliminaries, we expose the concrete need for such a method and detail its implementation using Answer Set Programming (ASP).
+We then demonstrate its efficiency and compare it to implicant-based methods on some large Boolean models from the litterature.
+
+All models used for evaluation and the implementation of the presented method are available at\dots.
+
+\sylvain{TBD}
 
@@
 
+\sylvain{I might add a small figure for the example later\dots}
+
 \subsection{Petri net encoding of Logical models}
@@
 
-Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining conditions for the activation/inactivation of the states.
+Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining conditions for the activation and inactivation of the states.
 In~\cite{chaouiya2004qualitative} this took the form of the whole truth table of the Boolean functions, but as shown in Appendix 1 of~\cite{chatain2014characterization} computing Disjunctive Normal Forms (DNF) of each Boolean function is enough.
@@
 \verb|p(v_1, x_1); ... ; p(v_k, x_k) :- p(v, x).|
-\] where \verb|p(v, x)| is the atom representing place \(p\) and \(\{\verb|p(v_1, x_1)|, ..., \verb|p(v_k, x_k)|\}\) is the set of atoms representing places in \(pred(t)\).
+\]
+where \verb|p(v, x)| is the atom representing place \(p\) and \{\verb|p(v_1, x_1)|, \dots,\linebreak \verb|p(v_k, x_k)|\} is the set of atoms representing places in \(pred(t)\).
 The rule (\ref{eq:conflict}) is translated into the ASP rule
@@
 One of the main maps studied this way is a map representing knowledge about the Rheumatoïd Arthritis~\cite{singh2018computational}, which was the main motivation for the development of a tool to automatically transform it into an executable Boolean model~\cite{aghamiri2020automated}.
-In the supplementary material of the paper, an excerpt of the map, focused around the apoptosis (cell death) module is transformed into a model of \emph{reasonable} size, namely 181 Boolean variables (model \verb|F5_RA_apoptosis_executable_module.sbml|).
+In the supplementary material of the paper, an excerpt of the map, focused around the apoptosis (cell death) module is transformed into a model of \emph{reasonable} size, namely 181 Boolean variables (model \verb|F5_RA_apoptosis_executable_module.sbml| of supplementary material S3, and model ``RA-apoptosis'' of Section~\ref{sec:eval}).
 The study of such model, though, is a big hurdle.
@@
 
-Using the method presented above, it is possible to convert the model to PNML in about 1s and to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) in a few milliseconds.
+Using the method presented above, it is possible to convert the model to PNML in about 1s and to obtain the all 640 minimal trap spaces (including ones that contain more than one state) in a few milliseconds.
 Unfortunately since this was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.
@@
 
+\sylvain{the \texttt{dom-pref} thing is deprecated since 2014 but still used in PyBoolNet, I think we might give the more modern version I used in the Python script, but we'll see}
 % \subsection{Benchmark vs.\ implicant-based methods}

### 2022-05-06 cmsb22.tex
Answer question about "advanced", make some small changes in Section "Conclusion"

--- 2022-05-06 cmsb22.tex
@@
   LIS, Aix-Marseille Université, Marseille, France\\
-  \email{\{Trinh.Van-Giang, belaid.benhamou\}@lis-lab.fr}
+  \email{\{trinh.van-giang, belaid.benhamou\}@lis-lab.fr}
   \and
@@
 This rule guarantees that \verb|p(v, 1)| and \verb|p(v, 0)| never belong to the same answer set of \(\mathcal{A}\), representing the conflict-freeness.
-Note that the number of atoms of \(\mathcal{A}\) is only \(2n\), which is theoretically advanced to the ASP encoding shown in~\cite{klarner2015computing} where the number of atoms of the encoded ASP equals to the number of prime-implicants of the Boolean model.
+Note that the number of atoms of \(\mathcal{A}\) is only \(2n\), which is more theoretically advanced than the ASP encoding shown in~\cite{klarner2015computing} where the number of atoms of the encoded ASP equals to the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
 
-\sylvain{What do you mean by \emph{advanced}? It is less since there might be an exponential number of prime-implicants?}
+% \sylvain{What do you mean by \emph{advanced}? It is less since there might be an exponential number of prime-implicants?}
+% \vangiang{Yes. Since there might be an exponential number of prime-implicants, our new ASP encoding is more advanced than the ASP encoding of PyBoolNet.}
 
@@
 
-In this article we proposed a new method for the computation of minimal trap-spaces of Boolean models, based on maximal siphons.
-This method is evaluated on big models from the litterature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
+In this article we proposed a new method for the computation of minimal trap-spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
+This method is evaluated on big models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
 We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
@@
 Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
-The tradeoff between a small Petri net, and hence small ASP, and the time it takes to compute it has to be evaluated in depth.
+The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it has to be evaluated in depth.
 

### 2022-05-06 cmsb22.tex
Details

@@
 This rule guarantees that \verb|p(v, 1)| and \verb|p(v, 0)| never belong to the same answer set of \(\mathcal{A}\), representing the conflict-freeness.
-Note that the number of atoms of \(\mathcal{A}\) is only \(2n\), which is more theoretically advanced than the ASP encoding shown in~\cite{klarner2015computing} where the number of atoms of the encoded ASP equals to the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
-
-% \sylvain{What do you mean by \emph{advanced}? It is less since there might be an exponential number of prime-implicants?}
-% \vangiang{Yes. Since there might be an exponential number of prime-implicants, our new ASP encoding is more advanced than the ASP encoding of PyBoolNet.}
+Note that the number of atoms of \(\mathcal{A}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
 
@@
 In this article we proposed a new method for the computation of minimal trap-spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
-This method is evaluated on big models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
+This method is evaluated on large models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
 We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.

### 2022-05-06 cmsb22.tex

@@
 The recent use of trap spaces made a real breakthrough in that field allowing to consider medium-sized models that used to be out of reach.
-  However, with the continuing increase in model-size, the state of the art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of implicants.
+  However, with the continuing increase in model-size, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of implicants.
 
@@
   After some technical preliminaries, we expose the concrete need for such a method and detail its implementation using Answer Set Programming.
-  We then demonstrate its efficiency and compare it to implicant-based methods on some large Boolean models from the litterature.
+  We then demonstrate its efficiency and compare it to implicant-based methods on some large Boolean models from the literature.
 

### 2022-05-06 cmsb22.tex
Motivating example

--- 2022-05-06 cmsb22.tex
@@
 Though this might appear quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
-Second, in practice this transformation, here using BDDs in BioLQM, seems to behave quite well on even quite big models as will be shown in the Section~\ref{sec:eval} on evaluation.
+In practice this transformation, here using BDDs in BioLQM, seems to behave quite well on even quite big models as will be shown in the Section~\ref{sec:eval} on evaluation.
 
@@
 
-For a few years now the second author has been collaborating with biologists who have been building huge detailed and annotated maps and now wish to analyze the dynamics of the corresponding models.
+For a few years now the second author has been collaborating with biologists who have been building very large detailed and annotated maps and now wish to analyze the dynamics of the corresponding models.
 One of the main maps studied this way is a map representing knowledge about the Rheumatoïd Arthritis~\cite{singh2018computational}, which was the main motivation for the development of a tool to automatically transform it into an executable Boolean model~\cite{aghamiri2020automated}.
@@
 In practice, even if there is a huge number of attractors in such a model, obtaining a sample of those can reveal very useful to invalidate the model and lead to further refinement.
+In particular, it provides a feature-rich alternative to random simulations for this type of very non-deterministic model.
+Being able to detect that there are inconsistencies with published experimental data in some of the, say first 100 attractors, can lead to a much quicker Systems Biology loop: model, invalidate, refine.
+
 However, using a state-of-the-art tool like PyBoolNet~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
 And hence, it is not possible to extract any (complex) attractor at all.
-This is also true for the Alzheimer model also mentioned in that same article and originally from~\cite{ogishima2016alzpathway}, but actually not for the MAPK model for which the first trap spaces can be obtained in reasonable time.
+This is also true for the Alzheimer model also mentioned in that same article and originally from~\cite{ogishima2016alzpathway} (\verb|F4| file in the original supplementary material, and ``Alzheimer'' in Table~\ref{tab:result_real}), but actually not for the MAPK model for which the first trap spaces can be obtained in reasonable time.
 The current practice usually revolves then around fixing some inputs to plausible values and reducing the model accordingly.
@@
 Unfortunately since this was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.
+
+\sylvain{It seems that the RA\_apoptosis model is only 180 nodes, where does the n = 431 come from?}
 

### 2022-05-06 cmsb22.tex
Answer questions

@@
 \usepackage{booktabs}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Drawing
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{tikz}
+\usepackage{amsfonts}
+
+\usetikzlibrary{positioning,automata} % drawing automa
+\usetikzlibrary{calc,fit,tikzmark}
+
+% for captions in figures
+\usepackage{subcaption}
 
@@
 
+\begin{figure}[!ht]
+\centering
+\begin{subfigure}[b]{0.5\textwidth}
+\centering
+\begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]
+   \node[circle,draw] (0) [] {00};
+   \node[circle,draw] (1) [above=of 0] {01};
+   \node[circle,draw] (2) [right=of 0] {10};
+   \node[circle,draw] (3) [above=of 2] {11};
+
+   \draw[->] (0) edge [bend left] (1);
+   \draw[->] (0) edge [bend right] (2);
+
+   \draw[->] (1) edge [bend left] (0);
+   \draw[->] (1) edge [loop right] (1);
+
+   \draw[->] (2) edge [bend right] (0);
+   \draw[->] (2) edge [loop right] (2);
+
+   \draw[->] (3) edge [loop right] (3);
+\end{tikzpicture}
+\caption{}
+\end{subfigure}%
+\begin{subfigure}[b]{0.5\textwidth}
+\centering
+\caption{}
+\end{subfigure}%
+\caption{(a) STG under the fully asynchronous update of the Boolean model. (b) Petri net encoding of the Boolean model.}
+\label{fig:stg_and_PN}
+\end{figure}
+
 \sylvain{I might add a small figure for the example later\dots}
+\vangiang{In this small figure, we can show the STG under the fully asynchronous update and the Petri net encoding of the Boolean model.}
 
@@
 \sylvain{It seems that the RA\_apoptosis model is only 180 nodes, where does the n = 431 come from?}
+\vangiang{n = 431 comes from Table 1 of~\cite{aghamiri2020automated}. I have checked this model. It has only 180 nodes. I will check others models come from~\cite{aghamiri2020automated}.}
 
@@
 \sylvain{the \texttt{dom-pref} thing is deprecated since 2014 but still used in PyBoolNet, I think we might give the more modern version I used in the Python script, but we'll see}
+\vangiang{I also found that you did not use the \texttt{dom-pref} thing in the Python script. I think we should use the same configuration as that used in PyBoolNet.}
+
 % \subsection{Benchmark vs.\ implicant-based methods}

### 2022-05-06 cmsb22.tex
Removed useless stuff

@@
 \begin{abstract}
-  % The abstract should briefly summarize the contents of the paper in
-  % 150--250 words.
 
@@
 
-\sylvain{I might add a small figure for the example later\dots}
 \vangiang{In this small figure, we can show the STG under the fully asynchronous update and the Petri net encoding of the Boolean model.}
@@
 
-\sylvain{It seems that the RA\_apoptosis model is only 180 nodes, where does the n = 431 come from?}
 \vangiang{n = 431 comes from Table 1 of~\cite{aghamiri2020automated}. I have checked this model. It has only 180 nodes. I will check others models come from~\cite{aghamiri2020automated}.}
+\sylvain{that's actually another model (called RA-map or something in the supplementary material), I think the other models should be ok.}
 
@@
 \vangiang{I also found that you did not use the \texttt{dom-pref} thing in the Python script. I think we should use the same configuration as that used in PyBoolNet.}
-
-% \subsection{Benchmark vs.\ implicant-based methods}
+\sylvain{as said above, this is deprecated, and has been since August 2014 \url{https://github.com/potassco/clasp/blob/master/CHANGES}, so I'd advise against using it, it might break any-time.}
 

### 2022-05-06 cmsb22.tex
Add responses, add a small change in Section "Conclusion"

@@
 \sylvain{that's actually another model (called RA-map or something in the supplementary material), I think the other models should be ok.}
+\vangiang{Oh. Thank you for confirming that.}
 
@@
     immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 506 & 13 & DNF & DNF & 36852 & 14.19 \\
+    RA-apoptosis~\cite{aghamiri2020automated} & 181 & - & 59 & DNF & DNF & 640 & 0.70 \\
     er-stress~\cite{ostaszewski2021covid19} & 182 & 266 & 75 & - & - & 25600 & 4.54 \\
@@
     signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 533 & 19 & - & - & 512 & 1.58 \\
-    cholocystokinin~\cite{aghamiri2020automated} & 404 & - & 95 & - & - & 705024 & 1080.35 \\
-    RA-apoptosis~\cite{aghamiri2020automated} & 431 & - & 59 & DNF & DNF & 640 & 0.70 \\ \midrule
+    cholocystokinin~\cite{aghamiri2020automated} & 404 & - & 95 & - & - & 705024 & 1080.35 \\ \midrule
 
@@
 \sylvain{as said above, this is deprecated, and has been since August 2014 \url{https://github.com/potassco/clasp/blob/master/CHANGES}, so I'd advise against using it, it might break any-time.}
+\vangiang{I see. Let's use the more modern version.}
 
@@
 
-Moreover, there are other methods for the maximal siphon computation, like SAT/MaxSAT approaches~\cite{nabli2016enumerating} that could replace our ASP program if they outperform it.
+Moreover, there are possibly other methods for the maximal siphon computation, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}.
+We also plan to investigate them in the future.
+They could replace our ASP program if they outperform it.
 

### 2022-05-08 trappist.py
Add a time-limit

@@ -77,7 +77,7 @@ def write_asp(petri_net: nx.DiGraph, asp_file: IO):
                     print(f"{or_preds} :- {pnml_to_asp(succ)}.", file=asp_file)
 
 
-def solve_asp(asp_filename: str, max_output: int) -> str:
+def solve_asp(asp_filename: str, max_output: int, time_limit: int) -> str:
     """Run an ASP solver on program asp_file and get the solutions."""
     # FIXME we need a limit on number of solutions
     result = subprocess.run(
@@ -88,6 +88,7 @@ def solve_asp(asp_filename: str, max_output: int) -> str:
             "--enum-mod=domRec",
             "--dom-mod=3",
             "--outf=2",  # json output
+            f"--time=limit={time_limit}",
             asp_filename,
         ],
         capture_output=True,
@@ -148,7 +149,7 @@ def get_solutions(
 
 
 def compute_trap_spaces(
-    infile: IO, display: bool = False, max_output: int = 0
+    infile: IO, display: bool = False, max_output: int = 0, time_limit: int = 0
 ) -> Optional[Generator[List[str], None, None]]:
     """Do the minimal trap-space computation on input file infile."""
     petri_net = read_pnml(infile)
@@ -156,7 +157,7 @@ def compute_trap_spaces(
     (_, tmpname) = tempfile.mkstemp(suffix=".lp", text=True)
     with open(tmpname, "wt") as asp_file:
         write_asp(petri_net, asp_file)
-    solutions = solve_asp(tmpname, max_output)
+    solutions = solve_asp(tmpname, max_output, time_limit)
     os.unlink(tmpname)
     return get_solutions(solutions, petri_net, display)
 
@@ -180,6 +181,13 @@ def main():
         help="Maximum number of solutions (0 for all).",
     )
     parser.add_argument(
+        "-t",
+        "--time",
+        type=int,
+        default=0,
+        help="Maximum number of seconds for search (0 for no-limit).",
+    )
+    parser.add_argument(
         "infile",
         type=argparse.FileType("r", encoding="utf-8"),
         nargs="?",
@@ -188,7 +196,9 @@ def main():
     )
     args = parser.parse_args()
 
-    compute_trap_spaces(args.infile, display=True, max_output=args.max)
+    compute_trap_spaces(
+        args.infile, display=True, max_output=args.max, time_limit=args.time
+    )
 
 
 if __name__ == "__main__":

### 2022-05-09 cmsb22.tex
Add concrete examples

@@
 After some technical preliminaries, we expose the concrete need for such a method and detail its implementation using Answer Set Programming (ASP).
-We then demonstrate its efficiency and compare it to implicant-based methods on some large Boolean models from the litterature.
+We then demonstrate its efficiency and compare it to implicant-based methods on some large Boolean models from the literature.
 
@@
 We will briefly recall here some preliminaries on Boolean models related to trap spaces and Petri nets.
-In the case of multi-level Logical models, an encoding into a Boolean model is always possible.
+In the case of multi-level Logical models, an encoding into a Boolean model is always possible~\cite{Didier2011}.
 
@@
 We recall here some definitions from~\cite{klarner2015computing} for the introduction of \emph{trap spaces}.
-Minimal trap spaces prove to be a very good approximation of the attractors of a Boolean model under asynchronous updating schemes and have become the \emph{de facto} standard way to analyze models of a few tens of \emph{genes}~\cite{klarner2017pyboolnet,cifuentes2020control}.
+Minimal trap spaces prove to be a very good approximation of the attractors of a Boolean model under asynchronous update schemes and have become the \emph{de facto} standard way to analyze models of a few tens of \emph{genes}~\cite{klarner2017pyboolnet,cifuentes2020control}.
+
+%%% Boolean model %%%
+Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\bbbb^{n}\) with \(\bbbb = \{0, 1\}\).
+A state \(s \in \bbbb^{n}\) is a mapping \(s : V \mapsto \bbbb\) that assigns either 0 (inactive) or 1 (active) to each node.
+At each time step \(t\), node \(x_i\) can update its state by
+\[x_i(t + 1) = f_i(x(t)).\] An update scheme of a Boolean model specifies the way that the nodes of the model update their states through time evolution~\cite{thomas1991regulatory}.
+There are two main types of update schemes: synchronous and asynchronous.
 
 %%% Trap set %%%
-Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\bbbb^{n}\) with \(\bbbb = \{0, 1\}\).
-A state \(s \in \bbbb^{n}\) is a mapping \(s : V \mapsto \bbbb\) that assigns either 0 (inactive) or 1 (active) to each node.
 A non-empty set \(T \subseteq S_{\mathcal{M}}\) is a \emph{trap set} of \(\mathcal{M}\) if and only if for every \(x \in T\) and \(y \in S_{\mathcal{M}}\) with \(y\) is reachable from \(x\) it holds that \(y \in T\).
@@
 A \emph{trap space} is defined as a subspace that is also a trap set.
+It is noted that trap spaces of a Boolean model are independent of the update scheme of this model~\cite{klarner2015computing}.
 Then, we define a partial order \(<\) on \(\mathcal{S}_{\mathcal{M}}^{\star}\) as: \(m < m'\) if and only if \(\mathcal{S}_{\mathcal{M}}[m] \subseteq \mathcal{S}_{\mathcal{M}}[m']\) and \(\mathcal{S}_{\mathcal{M}}[m] \neq \mathcal{S}_{\mathcal{M}}[m']\).
 Consequently, a trap space \(m\) is minimal if and only if there is no trap space \(m' \in \mathcal{S}_{\mathcal{M}}^{\star}\) such that \(m' < m\).
-For example, the Boolean model shown in Example~\ref{example:BN} has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
+
+%%% Example %%%
+For example, let us consider the Boolean model shown in Example~\ref{example:BN}.
+Figure~\ref{fig:stg_and_PN}(a) shows the dynamics of this model under the fully asynchronous update (i.e., only one node is nondeterministically selected in order to be updated at each time step).
+The model has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
 Since \(m_1 < m_2\), \(m_1\) is a minimal trap space of the Boolean model.
 
-%%% Example %%%
 \begin{example}
@@
 \centering
-\begin{subfigure}[b]{0.5\textwidth}
+\begin{subfigure}[b]{0.4\textwidth}
 \centering
 \begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]
-   \node[circle,draw] (0) [] {00};
-   \node[circle,draw] (1) [above=of 0] {01};
-   \node[circle,draw] (2) [right=of 0] {10};
-   \node[circle,draw] (3) [above=of 2] {11};
+   \node[] (0) [] {00};
+   \node[] (1) [above=of 0] {01};
+   \node[] (2) [right=of 0] {10};
+   \node[] (3) [above=of 2] {11};
 
@@
 \end{subfigure}%
-\begin{subfigure}[b]{0.5\textwidth}
+\begin{subfigure}[b]{0.6\textwidth}
 \centering
+\begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]\footnotesize
+  \node[circle,draw,label=left:$p_{x_1}$] (x1) {};
+  \node[circle,draw,label=left:$\overline{p_{x_1}}$] (nx1) [below=of x1, yshift=-0.5cm] {};
+
+  \node[circle,draw,label=right:$p_{x_2}$] (x2) [right=of x1, xshift=2.0cm] {};
+  \node[circle,draw,label=right:$\overline{p_{x_2}}$] (nx2) [below=of x2, yshift=-0.5cm] {};
+
+  \node[rectangle,draw,label=above:$t^1_{x_1}$] (t1x1) [right=of x1] {};
+  \node[rectangle,draw,label=above:$t^2_{x_1}$] (t2x1) [right=of nx1] {};
+
+  \node[rectangle,draw,label=above:$t^1_{x_2}$] (t1x2) [left=of x2] {};
+  \node[rectangle,draw,label=above:$t^2_{x_2}$] (t2x2) [left=of nx2] {};
+
+  \draw[->] (x1) -- (t1x1);
+  \draw[->] (t1x1) -- (nx1);
+  \draw[<->] (t1x1) -- (nx2);
+
+  \draw[->] (nx1) -- (t2x1);
+  \draw[->] (t2x1) -- (x1);
+  \draw[<->] (t2x1) edge [bend right] (nx2);
+
+  \draw[->] (x2) -- (t2x2);
+  \draw[->] (t2x2) -- (nx2);
+  \draw[<->] (t2x2) edge [bend left] (nx1);
+
+  \draw[->] (nx2) -- (t1x2);
+  \draw[->] (t1x2) -- (x2);
+  \draw[<->] (t1x2) -- (nx1);
+\end{tikzpicture}
 \caption{}
 \end{subfigure}%
-\caption{(a) STG under the fully asynchronous update of the Boolean model. (b) Petri net encoding of the Boolean model.}
+\caption{(a) Dynamics under the fully asynchronous update of the Boolean model. (b) Petri net encoding of the Boolean model. Circles denotes places, whereas rectangles denotes transitions.}
 \label{fig:stg_and_PN}
@@
 
-\vangiang{In this small figure, we can show the STG under the fully asynchronous update and the Petri net encoding of the Boolean model.}
+% \vangiang{In this small figure, we can show the STG under the fully asynchronous update and the Petri net encoding of the Boolean model.}
 
@@
 
-The main property of this encoding is that it is completely faithful w.r.t.\ the asynchronous semantics.
-Asynchronous transitions of the original model correspond one-to-one with firings of transitions in the Petri net.
+% The main property of this encoding is that it is completely faithful w.r.t.\ the asynchronous semantics.
+% Asynchronous transitions of the original model correspond one-to-one with firings of transitions in the Petri net.
+
+The main property of this encoding is that it is completely faithful with respect to the update scheme of the original Boolean model.
+For each node \(v\) of \(\mathcal{M}\), only transitions corresponding to \(v\) can change the current marking of \(p_v\) or \(\overline{p_v}\).
+In addition, at any marking at most one of such transitions is enabled because \(m(p_v) + m(\overline{p}_v) = 1\) holds.
+Hence, for any update scheme in \(\mathcal{M}\), we have a corresponding firing scheme in \(\mathcal{P}\), which preserves the equivalence between the dynamics of \(\mathcal{M}\) and \(\mathcal{P}\)~\cite{DBLP:journals/nc/ChatainHKPT20}.
+
+For illustration, let us reconsider the Boolean model shown in Example~\ref{example:BN}.
+Figure~\ref{fig:stg_and_PN}(b) shows the Petri net encoding of this Boolean model.
+Place \(p_{x_1}\) (resp. \(\overline{p_{x_1}}\)) in \(\mathcal{P}\) represents the activation (resp. the inactivation) of node \(x_1\) in \(\mathcal{M}\).
+Marking \(\{p_{x_1}, \overline{p_{x_2}}\}\) in \(\mathcal{P}\) represents state 10 in \(\mathcal{M}\). 
+Transitions \(t^{1}_{x_1}\) and \(t^{2}_{x_1}\) represent the update of node \(x_1\).
+Of course, at any marking \(t^{1}_{x_1}\) and \(t^{2}_{x_1}\) cannot be both enabled.
+Then, the fully asynchronous update scheme in \(\mathcal{M}\) corresponds to the classical firing scheme in \(\mathcal{P}\) where only one of the enabled transitions at a marking will be fired~\cite{Murata1989}.
 
@@
 It also accepts many other common formats for Logical models, notably the \verb|.bnet| files of the  BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} tools.
-
-Conversion is executed as follows:
+The conversion is executed as follows:
 
@@
 This is precisely the definition of a trap space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap space.
+For example, the Boolean model given in Example~\ref{example:BN} has two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
+The Petri net encoding of this Boolean model has five generic siphons, \(S_1 = \emptyset\), \(S_2 = \{p_{x_1}, \overline{p_{x_1}}\}\), \(S_3 = \{p_{x_2}, \overline{p_{x_2}}\}\), \(S_4 = \{\overline{p_{x_1}}, \overline{p_{x_2}}\}\), and \(S_5 = \{p_{x_1}, \overline{p_{x_1}}, p_{x_2}, \overline{p_{x_2}}\}\).
+However, only \(S_1\) and \(S_4\) are conflict-free siphons and correspond to \(m_2\) and \(m_1\), respectively.
+Since \(S_1 \subset S_4\), \(S_4\) is a maximal siphon corresponding to the minimal trap space \(m_1\).
 Hereafter, we formally prove that a maximal conflict-free siphon is equivalent to a minimal trap space.
@@
 Assume that \(S\) is not a siphon of \(\mathcal{P}\).
-Then, there is a transition \(t \in T\) such that \(S\cap succ(t)\not =\emptyset\) but \(S\cap pred(t)\not =\emptyset\).
+Then, there is a transition \(t \in T\) such that \(S\cap succ(t)\not =\emptyset\) but \(S\cap pred(t)=\emptyset\).
 In other words, there is a place \(p \in S\) such that \(p \in succ(t)\) but \(p \not \in pred(t)\).

### 2022-05-09 trappist.py
Towards benchmarking in a CoLoMoTo notebook

@@ -88,7 +88,7 @@ def solve_asp(asp_filename: str, max_output: int, time_limit: int) -> str:
             "--enum-mod=domRec",
             "--dom-mod=3",
             "--outf=2",  # json output
-            f"--time=limit={time_limit}",
+            f"--time-limit={time_limit}",
             asp_filename,
         ],
         capture_output=True,
@@ -99,6 +99,7 @@ def solve_asp(asp_filename: str, max_output: int, time_limit: int) -> str:
     # https://www.mat.unical.it/aspcomp2013/files/aspoutput.txt
     # 30: SAT, all enumerated, optima found, 10 stopped by max
     if result.returncode != 30 and result.returncode != 10:
+        print(f"Return code from clingo: {result.returncode}")
         result.check_returncode()  # will raise CalledProcessError
 
     return result.stdout

### 2022-05-09 cmsb22.tex
Since we have amsfonts…

@@
 \usepackage[T1]{fontenc}
-% \usepackage{graphicx}
 \usepackage{hyperref}
@@
 \usepackage{booktabs}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+% hyperref for LNCS
 \renewcommand\UrlFont{\color{blue}\rmfamily}
@@
 \newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
-
-\newcommand{\bbbb}{{\rm I\!B}}
 
@@
 %%% Boolean model %%%
-Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\bbbb^{n}\) with \(\bbbb = \{0, 1\}\).
-A state \(s \in \bbbb^{n}\) is a mapping \(s : V \mapsto \bbbb\) that assigns either 0 (inactive) or 1 (active) to each node.
+Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\mathbb{B}^{n}\) with \(\mathbb{B} = \{0, 1\}\).
+A state \(s \in \mathbb{B}^{n}\) is a mapping \(s : V \mapsto \mathbb{B}\) that assigns either 0 (inactive) or 1 (active) to each node.
 At each time step \(t\), node \(x_i\) can update its state by
@@
 A \emph{subspace} \(m\) of \(\mathcal{S}_{\mathcal{M}}\) is characterized by its fixed nodes (denoted by \(D_m\)) and free nodes.
-This subspace can be specified by an assignment \(D_m \mapsto \bbbb\) where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
+This subspace can be specified by an assignment \(D_m \mapsto \mathbb{B}\) where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
 The remaining nodes of \(\mathcal{M}\), \(V \setminus D_m\), are said to be \emph{free}, i.e., they can receive any Boolean value.
@@
   \(P \cap T = \emptyset\),
-  and \(W : (P \times T) \cup (T \times P) \mapsto \bbbn \) is a weight function attached to the arcs.
+  and \(W : (P \times T) \cup (T \times P) \mapsto \mathbb{N} \) is a weight function attached to the arcs.
 
 \end{definition}
-A \emph{marking} for a Petri net is a mapping \(m : P \mapsto \bbbn\) that assigns a number of tokens to each place.
+A \emph{marking} for a Petri net is a mapping \(m : P \mapsto \mathbb{N}\) that assigns a number of tokens to each place.
 A place \(p\) is marked by a marking \(m\) if and only if \(m(p) > 0\).

### 2022-05-09 cmsb22.tex

@@
 By the characterization of the encoding~\cite{chaouiya2004qualitative}, \(m_s(p) = 0, \forall p \in S\).
-In any marking \(m'_s\) reachable from \(m_s\) regardless of the update semantics of \(\mathcal{P}\), we have \(m'_s(p) = 0, \forall p \in S\) by the dynamical property on markings of a siphon~\cite{DBLP:journals/isci/LiuB16}.
-Equivalently, in any state \(s'\) reachable from \(s\) regardless of the update semantics of \(\mathcal{M}\), we have \(s'(v) = s(v) = m(v), \forall v \in D_m\).
+In any marking \(m'_s\) reachable from \(m_s\) regardless of the firing scheme of \(\mathcal{P}\), we have \(m'_s(p) = 0, \forall p \in S\) by the dynamical property on markings of a siphon~\cite{DBLP:journals/isci/LiuB16}.
+Equivalently, in any state \(s'\) reachable from \(s\) regardless of the update scheme of \(\mathcal{M}\), we have \(s'(v) = s(v) = m(v), \forall v \in D_m\).
 Then, \(s' \in \mathcal{S}_{\mathcal{M}}[m]\).
@@
 It is noted that there are no existing methods specifically designed for computing maximal conflict-free siphons (even maximal siphons) of a Petri net.
-The reason is that researchers mainly focus on minimal siphons~\cite{DBLP:journals/isci/LiuB16}.
+The reason might be that researchers mainly focus on minimal siphons~\cite{DBLP:journals/isci/LiuB16}.
 Hence, we here propose a new method based on Answer Set Programming (ASP)~\cite{DBLP:journals/aicom/GebserKKOSS11} for computing maximal conflict-free siphons of a Petri net.
@@
 \end{equation}The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net~\cite{nabli2016enumerating}.
-To make \(S\) to be a conflict-free siphon, we need to add the following rule for each node \(v \in V\) to the system.
+To make \(S\) to be a conflict-free siphon, we need to add to the system the rule
 \begin{equation}
@@
 p_v \in S \Rightarrow \overline{p_v} \not \in S \wedge \overline{p_v} \in S \Rightarrow p_v \not \in S
-\end{equation}By definition, the final system fully characterizes all conflict-free siphons of the encoded Petri net.
+\end{equation}for each node \(v \in V\).
+By definition, the final system fully characterizes all conflict-free siphons of the encoded Petri net.
 
 Then, we translate the above characterization into the ASP \(\mathcal{A}\) as follows.
-We use atom \verb|p(v, 1)| (resp.
+We introduce atom \verb|p(v, 1)| (resp.
 \verb|p(v, 0)|) to denote place \(p_v\) (resp.
@@
 \[\verb|:- p(v, 1), p(v, 0).|\]
-for each \(v \in V\).
-This rule guarantees that \verb|p(v, 1)| and \verb|p(v, 0)| never belong to the same answer set of \(\mathcal{A}\), representing the conflict-freeness.
+for each \(v \in V\). This ASP rule guarantees that two places representing the same node in \(\mathcal{M}\) never belong to the same siphon of \(\mathcal{P}\), representing the conflict-freeness.
+Naturally, a Herbrand model (see, e.g,~\cite{DBLP:journals/aicom/GebserKKOSS11}) of \(\mathcal{A}\) is equivalent to a conflict-free siphon of \(\mathcal{P}\).
+To guarantee that a Herbrand model is also a stable model (an answer set), we need to add to \(\mathcal{A}\) the choice rule
+\[
+\verb|{p(v, 1) ; p(v, 0)}.|
+\] for each \(v \in V\).
 Note that the number of atoms of \(\mathcal{A}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
 
-Naturally, a solution (simply an answer set) \(A\) of \(\mathcal{A}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a minimal trap space \(m\) of \(\mathcal{M}\).
+% For illustration, we consider the Petri net encoding \(\mathcal{P}\) shown in Figure~\ref{fig:stg_and_PN}(b).
+% If \(p_{x_1}\) is in a siphon \(S\), then \(\overline{p_{x_1}}\) or \(\overline{p_{x_2}}\) should be in \(S\).
+% This rule is translated into the ASP rule
+% \[
+% \verb|p(x_1, 0) ; p(x_2, 0) :- p(x_1, 1).|
+% \] To forbid that \(S\) contains both \(p_{x_1}\) and \(\overline{p_{x_1}}\), we have the ASP rule
+% \[
+% \verb|:- p(x_1, 1) , p(x_2, 0).|
+% \] Analogously for other places in \(\mathcal{P}\), we obtain the final ASP as follows:
+
+Now, a solution (simply an answer set) \(A\) of \(\mathcal{A}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a minimal trap space \(m\) of \(\mathcal{M}\).
 The conversion from \(A\) to \(m\) is straightforward.
@@
 
-\vangiang{n = 431 comes from Table 1 of~\cite{aghamiri2020automated}. I have checked this model. It has only 180 nodes. I will check others models come from~\cite{aghamiri2020automated}.}
-\sylvain{that's actually another model (called RA-map or something in the supplementary material), I think the other models should be ok.}
-\vangiang{Oh. Thank you for confirming that.}
+% \vangiang{n = 431 comes from Table 1 of~\cite{aghamiri2020automated}. I have checked this model. It has only 180 nodes. I will check others models come from~\cite{aghamiri2020automated}.}
+% \sylvain{that's actually another model (called RA-map or something in the supplementary material), I think the other models should be ok.}
+% \vangiang{Oh. Thank you for confirming that.}
 
@@
 
-\sylvain{the \texttt{dom-pref} thing is deprecated since 2014 but still used in PyBoolNet, I think we might give the more modern version I used in the Python script, but we'll see}
-\vangiang{I also found that you did not use the \texttt{dom-pref} thing in the Python script. I think we should use the same configuration as that used in PyBoolNet.}
-\sylvain{as said above, this is deprecated, and has been since August 2014 \url{https://github.com/potassco/clasp/blob/master/CHANGES}, so I'd advise against using it, it might break any-time.}
-\vangiang{I see. Let's use the more modern version.}
+% \sylvain{the \texttt{dom-pref} thing is deprecated since 2014 but still used in PyBoolNet, I think we might give the more modern version I used in the Python script, but we'll see}
+% \vangiang{I also found that you did not use the \texttt{dom-pref} thing in the Python script. I think we should use the same configuration as that used in PyBoolNet.}
+% \sylvain{as said above, this is deprecated, and has been since August 2014 \url{https://github.com/potassco/clasp/blob/master/CHANGES}, so I'd advise against using it, it might break any-time.}
+% \vangiang{I see. Let's use the more modern version.}
 
@@
 The first observation is that the proposed method much outperforms PyBoolNet in computational time. More specifically, ...
-\vangiang{In the journal version, we can measure the memory consumption of the two methods}
+% \vangiang{In the journal version, we can measure the memory consumption of the two methods}
 

### 2022-05-09 cmsb22.tex
Updates from Overleaf

@@
 \usepackage[T1]{fontenc}
+% \usepackage{graphicx}
 \usepackage{hyperref}
@@
 \usepackage{booktabs}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-% hyperref for LNCS
 \renewcommand\UrlFont{\color{blue}\rmfamily}
@@
 \newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
+
+\newcommand{\bbbb}{{\rm I\!B}}
 
@@
 %%% Boolean model %%%
-Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\mathbb{B}^{n}\) with \(\mathbb{B} = \{0, 1\}\).
-A state \(s \in \mathbb{B}^{n}\) is a mapping \(s : V \mapsto \mathbb{B}\) that assigns either 0 (inactive) or 1 (active) to each node.
+Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\bbbb^{n}\) with \(\bbbb = \{0, 1\}\).
+A state \(s \in \bbbb^{n}\) is a mapping \(s : V \mapsto \bbbb\) that assigns either 0 (inactive) or 1 (active) to each node.
 At each time step \(t\), node \(x_i\) can update its state by
@@
 A \emph{subspace} \(m\) of \(\mathcal{S}_{\mathcal{M}}\) is characterized by its fixed nodes (denoted by \(D_m\)) and free nodes.
-This subspace can be specified by an assignment \(D_m \mapsto \mathbb{B}\) where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
+This subspace can be specified by an assignment \(D_m \mapsto \bbbb\) where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
 The remaining nodes of \(\mathcal{M}\), \(V \setminus D_m\), are said to be \emph{free}, i.e., they can receive any Boolean value.
@@
   \(P \cap T = \emptyset\),
-  and \(W : (P \times T) \cup (T \times P) \mapsto \mathbb{N} \) is a weight function attached to the arcs.
+  and \(W : (P \times T) \cup (T \times P) \mapsto \bbbn \) is a weight function attached to the arcs.
 
 \end{definition}
-A \emph{marking} for a Petri net is a mapping \(m : P \mapsto \mathbb{N}\) that assigns a number of tokens to each place.
+A \emph{marking} for a Petri net is a mapping \(m : P \mapsto \bbbn\) that assigns a number of tokens to each place.
 A place \(p\) is marked by a marking \(m\) if and only if \(m(p) > 0\).

### 2022-05-09 cmsb22.tex
Some data on PyBoolNet repo

--- 2022-05-09 cmsb22.tex
@@
 
+\appendix
+
+\section{Temporary}
+
+\begin{table}[!htb]
+  \caption{Timing comparisons between the proposed method and PyBoolNet on the PyBoolNet's repository.
+    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively.
+Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap-spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
+  \centering
+  \label{tab:pyboolnet_repo}
+  \begin{tabular}{lrrrrrr}
+    \toprule
+    model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
+    arellano\_rootstem &&&& 4 & 0.04 & \textbf{0.31}\\
+    calzone\_cellfate &&&& 27 & 0.03 & 0.08\\
+    dahlhaus\_neuroplastoma &&&& 32 & 0.04 & 0.11\\
+    davidich\_yeast &&&& 12 & 0.03 & 0.08\\
+    dinwoodie\_life &&&& 7 & 0.03 & 0.08\\
+    dinwoodie\_stomatal &&&& 1 & 0.04 & 0.05\\
+    faure\_cellcycle &&&& 2 & 0.04 & 0.04\\
+    grieco\_mapk &&&& 18 & 0.04 & 0.08\\
+    irons\_yeast &&&& 1 & 0.05 & 0.04\\
+    jaoude\_thdiff &&&& 1000+ & \textbf{1.73} & 0.40\\
+    klamt\_tcr &&&& 8 & 0.06 & 0.04\\
+    krumsiek\_myeloid &&&& 6 & 0.03 & 0.03\\
+    multivalued &&&& 4 & 0.02 & 0.02\\
+    n12c5 &&&& 5 & \textbf{34.37} & 0.06\\
+    n3s1c1a &&&& 2 & 0.02 & 0.03\\
+    n3s1c1b &&&& 2 & 0.02 & 0.03\\
+    n5s3 &&&& 3 & 0.03 & 0.03\\
+    n6s1c2 &&&& 3 & 0.03 & 0.04\\
+    n7s3 &&&& 3 & 0.02 & 0.02\\
+    raf &&&& 2 & 0.03 & 0.02\\
+    randomnet\_n15k3 &&&& 3 & 0.04 & 0.05\\
+    randomnet\_n7k3 &&&& 10 & 0.03 & \textbf{0.19}\\
+    remy\_tumorigenesis &&&& 25 & \textbf{1.79} & 0.13\\
+    saadatpour\_guardcell &&&& 1 & 0.02 & 0.03\\
+    selvaggio\_emt &&&& 1000+ & \textbf{1.03} & 0.23\\
+    tournier\_apoptosis &&&& 3 & 0.05 & 0.06\\
+    xiao\_wnt5a &&&& 4 & 0.06 & 0.03\\
+    zhang\_tlgl &&&& 156 & 0.26 & 0.17\\
+    zhang\_tlgl\_v2 &&&& 258 & 0.19 & 0.17\\
+    \bottomrule
+  \end{tabular}
+
+  \sylvain{I'm not sure that \(e\) and \(s\) are very relevant, moreover they are not readily available\dots}
+\end{table}
+
+As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}, the results are comparable with all minimal trap spaces found very fast.
+Note however that on some very small models, our proposed method was sometimes slower than PyBoolNet, but still way under one second.
+Moreover, we believe that the result on the Arellano model is caused by the cold start of the JVM for BioLQM.
+On the contrary, on every model that was a bit challenging for ByPoolNet, the new method was far more efficient with speedups between one and two orders of magnitude.
+
 \end{document}

### 2022-05-09 cmsb22.tex
Model size

@@
     model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
-    arellano\_rootstem &&&& 4 & 0.04 & \textbf{0.31}\\
-    calzone\_cellfate &&&& 27 & 0.03 & 0.08\\
-    dahlhaus\_neuroplastoma &&&& 32 & 0.04 & 0.11\\
-    davidich\_yeast &&&& 12 & 0.03 & 0.08\\
-    dinwoodie\_life &&&& 7 & 0.03 & 0.08\\
-    dinwoodie\_stomatal &&&& 1 & 0.04 & 0.05\\
-    faure\_cellcycle &&&& 2 & 0.04 & 0.04\\
-    grieco\_mapk &&&& 18 & 0.04 & 0.08\\
-    irons\_yeast &&&& 1 & 0.05 & 0.04\\
-    jaoude\_thdiff &&&& 1000+ & \textbf{1.73} & 0.40\\
-    klamt\_tcr &&&& 8 & 0.06 & 0.04\\
-    krumsiek\_myeloid &&&& 6 & 0.03 & 0.03\\
-    multivalued &&&& 4 & 0.02 & 0.02\\
-    n12c5 &&&& 5 & \textbf{34.37} & 0.06\\
-    n3s1c1a &&&& 2 & 0.02 & 0.03\\
-    n3s1c1b &&&& 2 & 0.02 & 0.03\\
-    n5s3 &&&& 3 & 0.03 & 0.03\\
-    n6s1c2 &&&& 3 & 0.03 & 0.04\\
-    n7s3 &&&& 3 & 0.02 & 0.02\\
-    raf &&&& 2 & 0.03 & 0.02\\
-    randomnet\_n15k3 &&&& 3 & 0.04 & 0.05\\
-    randomnet\_n7k3 &&&& 10 & 0.03 & \textbf{0.19}\\
-    remy\_tumorigenesis &&&& 25 & \textbf{1.79} & 0.13\\
-    saadatpour\_guardcell &&&& 1 & 0.02 & 0.03\\
-    selvaggio\_emt &&&& 1000+ & \textbf{1.03} & 0.23\\
-    tournier\_apoptosis &&&& 3 & 0.05 & 0.06\\
-    xiao\_wnt5a &&&& 4 & 0.06 & 0.03\\
-    zhang\_tlgl &&&& 156 & 0.26 & 0.17\\
-    zhang\_tlgl\_v2 &&&& 258 & 0.19 & 0.17\\
+    arellano\_rootstem & 9 &&& 4 & 0.04 & \textbf{0.31}\\
+    calzone\_cellfate & 28 &&& 27 & 0.03 & 0.08\\
+    dahlhaus\_neuroplastoma & 23 &&& 32 & 0.04 & 0.11\\
+    davidich\_yeast & 10 &&& 12 & 0.03 & 0.08\\
+    dinwoodie\_life & 15 &&& 7 & 0.03 & 0.08\\
+    dinwoodie\_stomatal & 13 &&& 1 & 0.04 & 0.05\\
+    faure\_cellcycle & 10 &&& 2 & 0.04 & 0.04\\
+    grieco\_mapk & 53 &&& 18 & 0.04 & 0.08\\
+    irons\_yeast & 18 &&& 1 & 0.05 & 0.04\\
+    jaoude\_thdiff & 103 &&& 1000+ & \textbf{1.73} & 0.40\\
+    klamt\_tcr & 40 &&& 8 & 0.06 & 0.04\\
+    krumsiek\_myeloid & 11 &&& 6 & 0.03 & 0.03\\
+    multivalued & 13 &&& 4 & 0.02 & 0.02\\
+    n12c5 & 11 &&& 5 & \textbf{34.37} & 0.06\\
+    n3s1c1a & 2 &&& 2 & 0.02 & 0.03\\
+    n3s1c1b & 2 &&& 2 & 0.02 & 0.03\\
+    n5s3 & 4 &&& 3 & 0.03 & 0.03\\
+    n6s1c2 & 5 &&& 3 & 0.03 & 0.04\\
+    n7s3 & 6 &&& 3 & 0.02 & 0.02\\
+    raf & 3 &&& 2 & 0.03 & 0.02\\
+    randomnet\_n15k3 & 15 &&& 3 & 0.04 & 0.05\\
+    randomnet\_n7k3 & 7 &&& 10 & 0.03 & \textbf{0.19}\\
+    remy\_tumorigenesis & 34 &&& 25 & \textbf{1.79} & 0.13\\
+    saadatpour\_guardcell & 13 &&& 1 & 0.02 & 0.03\\
+    selvaggio\_emt & 56 &&& 1000+ & \textbf{1.03} & 0.23\\
+    tournier\_apoptosis & 12 &&& 3 & 0.05 & 0.06\\
+    xiao\_wnt5a & 7 &&& 4 & 0.06 & 0.03\\
+    zhang\_tlgl & 60 &&& 156 & 0.26 & 0.17\\
+    zhang\_tlgl\_v2 & 60 &&& 258 & 0.19 & 0.17\\
     \bottomrule

### 2022-05-09 cmsb22.tex
And more data

@@
 
-In this article we proposed a new method for the computation of minimal trap-spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
+In this article we proposed a new method for the computation of minimal trap spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
 This method is evaluated on large models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
@@
 
-\section{Temporary}
+\section{PyBoolNet repo}
 
@@
 Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap-spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
+    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
   \centering
@@
 
+\section{Selected models}
+
+\begin{table}[!htb]
+  \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
+    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively.
+Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
+  \centering
+  \label{tab:selected_redone}
+  \begin{tabular}{lrrrrrr}
+    \toprule
+    model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
+    CASCADE3 & 183 &&& 1 & 118.54 & 1.03\\
+    KYNURENINE-PATHWAY & 150 &&& 1 & 102.45 & 1.09\\
+    SN5 & 2746 &&& ??? & DNF & DNF\\
+    ERBB-RECEPTOR-SIGNALING & 247 &&& 1000+ & DNF & 3.56\\
+    Executable\_file\_for\_cholocystokinin\_model & 404 &&& 1000+ & 1.46 & 1.79\\
+    korkut\_2015a & 99 &&& 1000+ & DNF & 0.56\\
+    SIGNALING-PATHWAY-FOR-BUTANOL-PRODUCTION & 66 &&& 1 & 0.24 & 0.16\\
+    Corral\_ThIL17diff\_15jan2021 & 92 &&& 1000+ & DNF & 0.32\\
+    HIV-1 & 138 &&& 6 & DNF & 0.33\\
+    Executable\_file\_for\_CaSQ\_derived\_mast\_cell\_activation\_model & 80 &&& 1000+ & ??? & 0.17\\
+    Colitis\_associated\_colon\_cancer & 70 &&& 5 & 0.22 & 0.07\\
+    INTERFERON-1 & 121 &&& 2 & 13.21 & 0.12\\
+    TLGLSurvival & 61 &&& 3 & 0.73 & 0.16\\
+    cho\_2016\_all\_pos & 200 &&& 1000+ & DNF & 0.64\\
+    turei\_2016 & 4691 &&& ??? & DNF & DNF\\
+    Human\_network & 1953 &&& 1000+ & DNF & 42.58\\
+    Executable\_file\_for\_CaSQ\_derived\_MAPK\_model & 182 &&& 1000+ & ??? & 0.32\\
+    TCR-TLR5-SIGNALING-2018 & 130 &&& 48 & 1.52 & 0.17\\
+    Hernandez\_TcellCheckPoints\_13april2020 & 218 &&& 1000+ & 53.51 & 0.55\\
+    IMMUNE-SYSTEM & 164 &&& 1000+ & DNF & 0.92\\
+    HMOX1-PATHWAY & 145 &&& 8 & 3.72 & 0.11\\
+    Signaling\_in\_Macrophage\_Activation & 321 &&& 1 & 13.25 & 0.31\\
+    FIBROBLASTS & 139 &&& 1000+ & DNF & 0.61\\
+    S1\_Table & 1659 &&& 1000+ & DNF & 58.98\\
+    Regan2020\_Adhesion\_CIP\_Migration\_CellCycle\_Apoptosis & 121 &&& 78 & 53.30 & 0.53\\
+    SIGNALING-IN-PROSTATE-CANCER & 133 &&& 1000+ & DNF & 0.46\\
+    InflammatoryBowelDisease & 47 &&& 1 & DNF & 33.02\\
+    Influenza\_A\_Virus\_Replication\_Cycle & 131 &&& 17 & 51.17 & 0.13\\
+    IL\_6\_Signalling & 86 &&& 1 & 0.49 & 0.07\\
+    VIRUS-REPLICATION-CYCLE & 154 &&& 2 & DNF & 0.19\\
+    Executable\_file\_for\_Alzheimer\_model & 1169 &&& 1000+ & DNF & 1.53\\
+    ER-STRESS & 182 &&& 32 & 31.99 & 0.21\\
+    RA\_apoptosis\_executable\_module & 180 &&& 1000+ & DNF & 0.23\\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\sylvain{Some models seem to have a different number of trap spaces here than in the other table (like RA apoptosis), for others quite large time differences (ERBB, inflammatory bowel)\dots}
 \end{document}

### 2022-05-09 trappist.py
Useless commentaries removed

@@ -79,7 +79,6 @@ def write_asp(petri_net: nx.DiGraph, asp_file: IO):
 
 def solve_asp(asp_filename: str, max_output: int, time_limit: int) -> str:
     """Run an ASP solver on program asp_file and get the solutions."""
-    # FIXME we need a limit on number of solutions
     result = subprocess.run(
         [
             "clingo",
@@ -93,7 +92,6 @@ def solve_asp(asp_filename: str, max_output: int, time_limit: int) -> str:
         ],
         capture_output=True,
         text=True,
-        # check=True,
     )
 
     # https://www.mat.unical.it/aspcomp2013/files/aspoutput.txt

### 2022-05-09 cmsb22.tex
Add two missing results

--- 2022-05-09 cmsb22.tex
@@
 
+\section{Selected models}
+
+\begin{table}[!htb]
+  \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
+    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively.
+Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
+  \centering
+  \label{tab:selected_redone}
+  \begin{tabular}{lrrrrrr}
+    \toprule
+    model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
+    CASCADE3 & 183 &&& 1 & 118.54 & 1.03\\
+    KYNURENINE-PATHWAY & 150 &&& 1 & 102.45 & 1.09\\
+    SN5 & 2746 &&& ??? & DNF & DNF\\
+    ERBB-RECEPTOR-SIGNALING & 247 &&& 1000+ & DNF & 3.56\\
+    Executable\_file\_for\_cholocystokinin\_model & 404 &&& 1000+ & 1.46 & 1.79\\
+    korkut\_2015a & 99 &&& 1000+ & DNF & 0.56\\
+    SIGNALING-PATHWAY-FOR-BUTANOL-PRODUCTION & 66 &&& 1 & 0.24 & 0.16\\
+    Corral\_ThIL17diff\_15jan2021 & 92 &&& 1000+ & DNF & 0.32\\
+    HIV-1 & 138 &&& 6 & DNF & 0.33\\
+    Executable\_file\_for\_CaSQ\_derived\_mast\_cell\_activation\_model & 80 &&& 1000+ & 0.89 & 0.17\\
+    Colitis\_associated\_colon\_cancer & 70 &&& 5 & 0.22 & 0.07\\
+    INTERFERON-1 & 121 &&& 2 & 13.21 & 0.12\\
+    TLGLSurvival & 61 &&& 3 & 0.73 & 0.16\\
+    cho\_2016\_all\_pos & 200 &&& 1000+ & DNF & 0.64\\
+    turei\_2016 & 4691 &&& ??? & DNF & DNF\\
+    Human\_network & 1953 &&& 1000+ & DNF & 42.58\\
+    Executable\_file\_for\_CaSQ\_derived\_MAPK\_model & 182 &&& 1000+ & DNF & 0.32\\
+    TCR-TLR5-SIGNALING-2018 & 130 &&& 48 & 1.52 & 0.17\\
+    Hernandez\_TcellCheckPoints\_13april2020 & 218 &&& 1000+ & 53.51 & 0.55\\
+    IMMUNE-SYSTEM & 164 &&& 1000+ & DNF & 0.92\\
+    HMOX1-PATHWAY & 145 &&& 8 & 3.72 & 0.11\\
+    Signaling\_in\_Macrophage\_Activation & 321 &&& 1 & 13.25 & 0.31\\
+    FIBROBLASTS & 139 &&& 1000+ & DNF & 0.61\\
+    S1\_Table & 1659 &&& 1000+ & DNF & 58.98\\
+    Regan2020\_Adhesion\_CIP\_Migration\_CellCycle\_Apoptosis & 121 &&& 78 & 53.30 & 0.53\\
+    SIGNALING-IN-PROSTATE-CANCER & 133 &&& 1000+ & DNF & 0.46\\
+    InflammatoryBowelDisease & 47 &&& 1 & DNF & 33.02\\
+    Influenza\_A\_Virus\_Replication\_Cycle & 131 &&& 17 & 51.17 & 0.13\\
+    IL\_6\_Signalling & 86 &&& 1 & 0.49 & 0.07\\
+    VIRUS-REPLICATION-CYCLE & 154 &&& 2 & DNF & 0.19\\
+    Executable\_file\_for\_Alzheimer\_model & 1169 &&& 1000+ & DNF & 1.53\\
+    ER-STRESS & 182 &&& 32 & 31.99 & 0.21\\
+    RA\_apoptosis\_executable\_module & 180 &&& 1000+ & DNF & 0.23\\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\sylvain{Some models seem to have a different number of trap spaces here than in the other table (like RA apoptosis), for others quite large time differences (ERBB, inflammatory bowel)\dots}
 \end{document}

### 2022-05-09 cmsb22.tex
Answer questions

@@
 \sylvain{Some models seem to have a different number of trap spaces here than in the other table (like RA apoptosis), for others quite large time differences (ERBB, inflammatory bowel)\dots}
+\vangiang{Regarding the time differences, the reason might be that my implementation uses an ASP encoding based on choice rules instead of disjunctive rules or constraint rules.}
+\vangiang{Regarding the differences in the numbers of minimal trap spaces, I have also observed this phenomenon.
+For example, for the Colitis\_associated\_colon\_cancer, the number of minimal trap spaces should be 10 (I am sure that this number is correct).
+For the TLGLSurvival model, the number of minimal trap spaces should be 318.
+Since PyBoolNet and Trappist return the same results for these models, I do not think that there is a problem with Trappist.
+Please check the model files.}
 \end{document}

### 2022-05-09 cmsb22.tex
Comments

@@
 All models used for evaluation and the implementation of the presented method are available at\dots.
-
-\sylvain{TBD}
 
@@
 
-% \vangiang{n = 431 comes from Table 1 of~\cite{aghamiri2020automated}. I have checked this model. It has only 180 nodes. I will check others models come from~\cite{aghamiri2020automated}.}
-% \sylvain{that's actually another model (called RA-map or something in the supplementary material), I think the other models should be ok.}
-% \vangiang{Oh. Thank you for confirming that.}
-
 \section{Evaluation}
@@
 To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
-Specifically, we used the configuration \texttt{--dom-pref=32, --heu=domain, --dom-mod=7} (subset maximality).
+Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heu=domain --dom-mod=7} used by PyBoolNet).
 We ran all the benchmarks on a virtual machine whose environment is CPU: Intel(R) Core(TM) i7-3630QM 2.40GHz x 4, Memory: 8 GB, Ubuntu 18.04.2 64 bit.
 Finally, we set a time limit as one hour for each model.
-
-% \sylvain{the \texttt{dom-pref} thing is deprecated since 2014 but still used in PyBoolNet, I think we might give the more modern version I used in the Python script, but we'll see}
-% \vangiang{I also found that you did not use the \texttt{dom-pref} thing in the Python script. I think we should use the same configuration as that used in PyBoolNet.}
-% \sylvain{as said above, this is deprecated, and has been since August 2014 \url{https://github.com/potassco/clasp/blob/master/CHANGES}, so I'd advise against using it, it might break any-time.}
-% \vangiang{I see. Let's use the more modern version.}
 
@@
   \end{tabular}
+\end{table}
 
   \sylvain{I'm not sure that \(e\) and \(s\) are very relevant, moreover they are not readily available\dots}
-\end{table}
 
@@
 \vangiang{Regarding the time differences, the reason might be that my implementation uses an ASP encoding based on choice rules instead of disjunctive rules or constraint rules.}
+\sylvain{Maybe\dots I don't know ASP well enough to be sure. The fact that you go through a predicate \texttt{p} which adds one level of indirection might also have some effect, I'm not sure.}
+
 \vangiang{Regarding the differences in the numbers of minimal trap spaces, I have also observed this phenomenon.
@@
 Please check the model files.}
+\sylvain{There are not among the files I had to change for BioLQM, so I'm not sure what I can check.
+For TLGL it should be quite obvious since 3 and 318 are pretty far away, but\dots I don't see any thing obviously wrong in the files.
+The 3 minimal trap spaces found are 3 stable-states, all 0 or with a single 1.
+They are the same also found using the BDD of BioLQM, so\dots}
 \end{document}

### 2022-05-10 cmsb22.tex
Small changes in Section "Preliminaries"

@@
 %%% Boolean model %%%
-Given a Boolean model \(\mathcal{M} = (V, F)\) with variables \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\mathbb{B}^{n}\) with \(\mathbb{B} = \{0, 1\}\).
+Given a Boolean model \(\mathcal{M} = (V, F)\) with nodes \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\mathbb{B}^{n}\) with \(\mathbb{B} = \{0, 1\}\).
 A state \(s \in \mathbb{B}^{n}\) is a mapping \(s : V \mapsto \mathbb{B}\) that assigns either 0 (inactive) or 1 (active) to each node.
-At each time step \(t\), node \(x_i\) can update its state by
-\[x_i(t + 1) = f_i(x(t)).\] An update scheme of a Boolean model specifies the way that the nodes of the model update their states through time evolution~\cite{thomas1991regulatory}.
+At each time step \(t\), node \(v_i\) can update its state by
+\[v_i(t + 1) = f_i(v(t)),\] where \(v(t)\) is the state of \(\mathcal{M}\) at time \(t\) and \(v_i(t + 1)\) the state of node \(v_i\) at time \(t + 1\).
+An update scheme of a Boolean model specifies the way that the nodes of the model update their states through time evolution~\cite{thomas1991regulatory}.
 There are two main types of update schemes: synchronous and asynchronous.
@@
 We write subspaces like states but use in addition the symbol \(\star\) to indicate that a node is free.
-A subspace \(m\) thus corresponds to the set of states \(\mathcal{S}_{\mathcal{M}}[m] := \{s \in \mathcal{S}_{\mathcal{M}}\;|\;\forall x_i \in D_m : s_i = m(x_i)\}\).
-For example, \(m = \star\star1\) means that \(D_m = \{x_3\}\), \(m(x_3) = 1\), and corresponds to the set of states \(\{001, 011, 101, 111\}\).
+A subspace \(m\) thus corresponds to the set of states \(\mathcal{S}_{\mathcal{M}}[m] := \{s \in \mathcal{S}_{\mathcal{M}}\;|\;\forall v \in D_m : s(v) = m(v)\}\).
+For example, \(m = \star\star1\) means that \(D_m = \{v_3\}\), \(m(v_3) = 1\), and corresponds to the set of states \(\{001, 011, 101, 111\}\).
 Let \(\mathcal{S}_{\mathcal{M}}^{\star}\) denote the set of all possible subspaces of \(\mathcal{M}\).

### 2022-05-10 cmsb22.tex
Updated bench with scripted models

@@
 Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
+    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes.}
   \centering
@@
     model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
-    CASCADE3 & 183 &&& 1 & 118.54 & 1.03\\
-    KYNURENINE-PATHWAY & 150 &&& 1 & 102.45 & 1.09\\
+    CASCADE3 & 183 &&& 1 & 82.86 & 12.83\\
+    KYNURENINE-PATHWAY & 150 &&& 1000+ & 74.71 & 4.22\\
     SN5 & 2746 &&& ??? & DNF & DNF\\
-    ERBB-RECEPTOR-SIGNALING & 247 &&& 1000+ & DNF & 3.56\\
-    Executable\_file\_for\_cholocystokinin\_model & 404 &&& 1000+ & 1.46 & 1.79\\
-    korkut\_2015a & 99 &&& 1000+ & DNF & 0.56\\
-    SIGNALING-PATHWAY-FOR-BUTANOL-PRODUCTION & 66 &&& 1 & 0.24 & 0.16\\
-    Corral\_ThIL17diff\_15jan2021 & 92 &&& 1000+ & DNF & 0.32\\
-    HIV-1 & 138 &&& 6 & DNF & 0.33\\
-    Executable\_file\_for\_CaSQ\_derived\_mast\_cell\_activation\_model & 80 &&& 1000+ & 0.89 & 0.17\\
-    Colitis\_associated\_colon\_cancer & 70 &&& 5 & 0.22 & 0.07\\
-    INTERFERON-1 & 121 &&& 2 & 13.21 & 0.12\\
-    TLGLSurvival & 61 &&& 3 & 0.73 & 0.16\\
-    cho\_2016\_all\_pos & 200 &&& 1000+ & DNF & 0.64\\
+    ERBB-RECEPTOR-SIGNALING & 247 &&& 1000+ & DNF & 9.27\\
+    Executable\_file\_for\_cholocystokinin\_model & 404 &&& 1000+ & 1.55 & 3.57\\
+    korkut\_2015a & 99 &&& 1000+ & DNF & 3.06\\
+    SIGNALING-PATHWAY-FOR-BUTANOL-PRODUCTION & 66 &&& 1000+ & 0.43 & 0.66\\
+    Corral\_ThIL17diff\_15jan2021 & 92 &&& 1000+ & DNF & 0.91\\
+    HIV-1 & 138 &&& 1000+ & DNF & 3.25\\
+    Executable\_file\_for\_CaSQ\_derived\_mast\_cell\_activation\_model & 80 &&& 1000+ & 0.49 & 0.56\\
+    Colitis\_associated\_colon\_cancer & 70 &&& 10 & 0.15 & 0.22\\
+    INTERFERON-1 & 121 &&& 1000+ & 7.69 & 1.05\\
+    TLGLSurvival & 61 &&& 318 & 0.59 & 0.48\\
+    cho\_2016\_all\_pos & 200 &&& 1000+ & DNF & 3.00\\
     turei\_2016 & 4691 &&& ??? & DNF & DNF\\
-    Human\_network & 1953 &&& 1000+ & DNF & 42.58\\
-    Executable\_file\_for\_CaSQ\_derived\_MAPK\_model & 182 &&& 1000+ & DNF & 0.32\\
-    TCR-TLR5-SIGNALING-2018 & 130 &&& 48 & 1.52 & 0.17\\
-    Hernandez\_TcellCheckPoints\_13april2020 & 218 &&& 1000+ & 53.51 & 0.55\\
-    IMMUNE-SYSTEM & 164 &&& 1000+ & DNF & 0.92\\
-    HMOX1-PATHWAY & 145 &&& 8 & 3.72 & 0.11\\
-    Signaling\_in\_Macrophage\_Activation & 321 &&& 1 & 13.25 & 0.31\\
-    FIBROBLASTS & 139 &&& 1000+ & DNF & 0.61\\
-    S1\_Table & 1659 &&& 1000+ & DNF & 58.98\\
-    Regan2020\_Adhesion\_CIP\_Migration\_CellCycle\_Apoptosis & 121 &&& 78 & 53.30 & 0.53\\
-    SIGNALING-IN-PROSTATE-CANCER & 133 &&& 1000+ & DNF & 0.46\\
-    InflammatoryBowelDisease & 47 &&& 1 & DNF & 33.02\\
-    Influenza\_A\_Virus\_Replication\_Cycle & 131 &&& 17 & 51.17 & 0.13\\
-    IL\_6\_Signalling & 86 &&& 1 & 0.49 & 0.07\\
-    VIRUS-REPLICATION-CYCLE & 154 &&& 2 & DNF & 0.19\\
-    Executable\_file\_for\_Alzheimer\_model & 1169 &&& 1000+ & DNF & 1.53\\
-    ER-STRESS & 182 &&& 32 & 31.99 & 0.21\\
-    RA\_apoptosis\_executable\_module & 180 &&& 1000+ & DNF & 0.23\\
+    Human\_network & 1953 &&& 1000+ & DNF & DNF\\
+    Executable\_file\_for\_CaSQ\_derived\_MAPK\_model & 182 &&& 1000+ & 75.32 & 2.26\\
+    TCR-TLR5-SIGNALING-2018 & 130 &&& 48 & 0.93 & 0.53\\
+    Hernandez\_TcellCheckPoints\_13april2020 & 218 &&& 1000+ & 36.21 & 2.27\\
+    IMMUNE-SYSTEM & 164 &&& 1000+ & DNF & 2.78\\
+    HMOX1-PATHWAY & 145 &&& 1000+ & 3.89 & 0.95\\
+    Signaling\_in\_Macrophage\_Activation & 321 &&& 1000+ & 11.51 & 3.27\\
+    FIBROBLASTS & 139 &&& 1000+ & DNF & 1.91\\
+    S1\_Table & 1659 &&& 1000+ & DNF & DNF\\
+    Regan2020\_Adhesion\_CIP\_Migration\_CellCycle\_Apoptosis & 121 &&& 78 & 40.59 & 2.15\\
+    SIGNALING-IN-PROSTATE-CANCER & 133 &&& 1000+ & DNF & 1.70\\
+    InflammatoryBowelDisease & 47 &&& 1 & DNF & 118.78\\
+    Influenza\_A\_Virus\_Replication\_Cycle & 131 &&& 1000+ & 47.69 & 0.87\\
+    IL\_6\_Signalling & 86 &&& 1000+ & 1.10 & 0.46\\
+    VIRUS-REPLICATION-CYCLE & 154 &&& 1000+ & DNF & 1.16\\
+    Executable\_file\_for\_Alzheimer\_model & 1169 &&& 1000+ & DNF & 6.15\\
+    ER-STRESS & 182 &&& 1000+ & 16.15 & 0.93\\
+    RA\_apoptosis\_executable\_module & 180 &&& 1000+ & DNF & 0.96\\
     \bottomrule
@@
 
-\sylvain{Some models seem to have a different number of trap spaces here than in the other table (like RA apoptosis), for others quite large time differences (ERBB, inflammatory bowel)\dots}
 \vangiang{Regarding the time differences, the reason might be that my implementation uses an ASP encoding based on choice rules instead of disjunctive rules or constraint rules.}
@@
 
-\vangiang{Regarding the differences in the numbers of minimal trap spaces, I have also observed this phenomenon.
-For example, for the Colitis\_associated\_colon\_cancer, the number of minimal trap spaces should be 10 (I am sure that this number is correct).
-For the TLGLSurvival model, the number of minimal trap spaces should be 318.
-Since PyBoolNet and Trappist return the same results for these models, I do not think that there is a problem with Trappist.
-Please check the model files.}
-\sylvain{There are not among the files I had to change for BioLQM, so I'm not sure what I can check.
-For TLGL it should be quite obvious since 3 and 318 are pretty far away, but\dots I don't see any thing obviously wrong in the files.
-The 3 minimal trap spaces found are 3 stable-states, all 0 or with a single 1.
-They are the same also found using the BDD of BioLQM, so\dots}
 \end{document}

### 2022-05-10 cmsb22.tex
Answer two questions

@@
   \sylvain{I'm not sure that \(e\) and \(s\) are very relevant, moreover they are not readily available\dots}
+  \vangiang{Indeed, they are not readily available.
+But if we use more meaningful words, the width of the table may exceed the width of the page.
+Thus, I use \(e\) and \(s\) along with the explanations shown in the table's caption.
+Do you have any other option?}
 
@@
 \sylvain{Maybe\dots I don't know ASP well enough to be sure. The fact that you go through a predicate \texttt{p} which adds one level of indirection might also have some effect, I'm not sure.}
+\vangiang{Oh. I think that this may be a reason. However, I think we should still use a predicate \texttt{p(v, x)} in the paper because of the readability.}
 

### 2022-05-10 cmsb22.tex
Original models and BioLQM transformation to bnet

@@
   \caption{Timing comparisons between the proposed method and PyBoolNet on the PyBoolNet's repository.
-    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively.
-Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+    Column "\(n\)" denotes the numbers of nodes of each model. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
     "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
@@
     \toprule
-    model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
-    arellano\_rootstem & 9 &&& 4 & 0.04 & \textbf{0.31}\\
-    calzone\_cellfate & 28 &&& 27 & 0.03 & 0.08\\
-    dahlhaus\_neuroplastoma & 23 &&& 32 & 0.04 & 0.11\\
-    davidich\_yeast & 10 &&& 12 & 0.03 & 0.08\\
-    dinwoodie\_life & 15 &&& 7 & 0.03 & 0.08\\
-    dinwoodie\_stomatal & 13 &&& 1 & 0.04 & 0.05\\
-    faure\_cellcycle & 10 &&& 2 & 0.04 & 0.04\\
-    grieco\_mapk & 53 &&& 18 & 0.04 & 0.08\\
-    irons\_yeast & 18 &&& 1 & 0.05 & 0.04\\
-    jaoude\_thdiff & 103 &&& 1000+ & \textbf{1.73} & 0.40\\
-    klamt\_tcr & 40 &&& 8 & 0.06 & 0.04\\
-    krumsiek\_myeloid & 11 &&& 6 & 0.03 & 0.03\\
-    multivalued & 13 &&& 4 & 0.02 & 0.02\\
-    n12c5 & 11 &&& 5 & \textbf{34.37} & 0.06\\
-    n3s1c1a & 2 &&& 2 & 0.02 & 0.03\\
-    n3s1c1b & 2 &&& 2 & 0.02 & 0.03\\
-    n5s3 & 4 &&& 3 & 0.03 & 0.03\\
-    n6s1c2 & 5 &&& 3 & 0.03 & 0.04\\
-    n7s3 & 6 &&& 3 & 0.02 & 0.02\\
-    raf & 3 &&& 2 & 0.03 & 0.02\\
-    randomnet\_n15k3 & 15 &&& 3 & 0.04 & 0.05\\
-    randomnet\_n7k3 & 7 &&& 10 & 0.03 & \textbf{0.19}\\
-    remy\_tumorigenesis & 34 &&& 25 & \textbf{1.79} & 0.13\\
-    saadatpour\_guardcell & 13 &&& 1 & 0.02 & 0.03\\
-    selvaggio\_emt & 56 &&& 1000+ & \textbf{1.03} & 0.23\\
-    tournier\_apoptosis & 12 &&& 3 & 0.05 & 0.06\\
-    xiao\_wnt5a & 7 &&& 4 & 0.06 & 0.03\\
-    zhang\_tlgl & 60 &&& 156 & 0.26 & 0.17\\
-    zhang\_tlgl\_v2 & 60 &&& 258 & 0.19 & 0.17\\
+    model & $n$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
+    arellano\_rootstem & 9 & 4 & 0.04 & \textbf{0.31}\\
+    calzone\_cellfate & 28 & 27 & 0.03 & 0.08\\
+    dahlhaus\_neuroplastoma & 23 & 32 & 0.04 & 0.11\\
+    davidich\_yeast & 10 & 12 & 0.03 & 0.08\\
+    dinwoodie\_life & 15 & 7 & 0.03 & 0.08\\
+    dinwoodie\_stomatal & 13 & 1 & 0.04 & 0.05\\
+    faure\_cellcycle & 10 & 2 & 0.04 & 0.04\\
+    grieco\_mapk & 53 & 18 & 0.04 & 0.08\\
+    irons\_yeast & 18 & 1 & 0.05 & 0.04\\
+    jaoude\_thdiff & 103 & 1000+ & \textbf{1.73} & 0.40\\
+    klamt\_tcr & 40 & 8 & 0.06 & 0.04\\
+    krumsiek\_myeloid & 11 & 6 & 0.03 & 0.03\\
+    multivalued & 13 & 4 & 0.02 & 0.02\\
+    n12c5 & 11 & 5 & \textbf{34.37} & 0.06\\
+    n3s1c1a & 2 & 2 & 0.02 & 0.03\\
+    n3s1c1b & 2 & 2 & 0.02 & 0.03\\
+    n5s3 & 4 & 3 & 0.03 & 0.03\\
+    n6s1c2 & 5 & 3 & 0.03 & 0.04\\
+    n7s3 & 6 & 3 & 0.02 & 0.02\\
+    raf & 3 & 2 & 0.03 & 0.02\\
+    randomnet\_n15k3 & 15 & 3 & 0.04 & 0.05\\
+    randomnet\_n7k3 & 7 & 10 & 0.03 & \textbf{0.19}\\
+    remy\_tumorigenesis & 34 & 25 & \textbf{1.79} & 0.13\\
+    saadatpour\_guardcell & 13 & 1 & 0.02 & 0.03\\
+    selvaggio\_emt & 56 & 1000+ & \textbf{1.03} & 0.23\\
+    tournier\_apoptosis & 12 & 3 & 0.05 & 0.06\\
+    xiao\_wnt5a & 7 & 4 & 0.06 & 0.03\\
+    zhang\_tlgl & 60 & 156 & 0.26 & 0.17\\
+    zhang\_tlgl\_v2 & 60 & 258 & 0.19 & 0.17\\
     \bottomrule
@@
 Do you have any other option?}
+  \sylvain{just remove them since they are not relevant}
 
@@
   \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
-    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively.
-Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+    Columns "\(n\)" denotes the numbers of nodes of each model. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
     "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes.}
@@
     \toprule
-    model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
-    CASCADE3 & 183 &&& 1 & 82.86 & 12.83\\
-    KYNURENINE-PATHWAY & 150 &&& 1000+ & 74.71 & 4.22\\
-    SN5 & 2746 &&& ??? & DNF & DNF\\
-    ERBB-RECEPTOR-SIGNALING & 247 &&& 1000+ & DNF & 9.27\\
-    Executable\_file\_for\_cholocystokinin\_model & 404 &&& 1000+ & 1.55 & 3.57\\
-    korkut\_2015a & 99 &&& 1000+ & DNF & 3.06\\
-    SIGNALING-PATHWAY-FOR-BUTANOL-PRODUCTION & 66 &&& 1000+ & 0.43 & 0.66\\
-    Corral\_ThIL17diff\_15jan2021 & 92 &&& 1000+ & DNF & 0.91\\
-    HIV-1 & 138 &&& 1000+ & DNF & 3.25\\
-    Executable\_file\_for\_CaSQ\_derived\_mast\_cell\_activation\_model & 80 &&& 1000+ & 0.49 & 0.56\\
-    Colitis\_associated\_colon\_cancer & 70 &&& 10 & 0.15 & 0.22\\
-    INTERFERON-1 & 121 &&& 1000+ & 7.69 & 1.05\\
-    TLGLSurvival & 61 &&& 318 & 0.59 & 0.48\\
-    cho\_2016\_all\_pos & 200 &&& 1000+ & DNF & 3.00\\
-    turei\_2016 & 4691 &&& ??? & DNF & DNF\\
-    Human\_network & 1953 &&& 1000+ & DNF & DNF\\
-    Executable\_file\_for\_CaSQ\_derived\_MAPK\_model & 182 &&& 1000+ & 75.32 & 2.26\\
-    TCR-TLR5-SIGNALING-2018 & 130 &&& 48 & 0.93 & 0.53\\
-    Hernandez\_TcellCheckPoints\_13april2020 & 218 &&& 1000+ & 36.21 & 2.27\\
-    IMMUNE-SYSTEM & 164 &&& 1000+ & DNF & 2.78\\
-    HMOX1-PATHWAY & 145 &&& 1000+ & 3.89 & 0.95\\
-    Signaling\_in\_Macrophage\_Activation & 321 &&& 1000+ & 11.51 & 3.27\\
-    FIBROBLASTS & 139 &&& 1000+ & DNF & 1.91\\
-    S1\_Table & 1659 &&& 1000+ & DNF & DNF\\
-    Regan2020\_Adhesion\_CIP\_Migration\_CellCycle\_Apoptosis & 121 &&& 78 & 40.59 & 2.15\\
-    SIGNALING-IN-PROSTATE-CANCER & 133 &&& 1000+ & DNF & 1.70\\
-    InflammatoryBowelDisease & 47 &&& 1 & DNF & 118.78\\
-    Influenza\_A\_Virus\_Replication\_Cycle & 131 &&& 1000+ & 47.69 & 0.87\\
-    IL\_6\_Signalling & 86 &&& 1000+ & 1.10 & 0.46\\
-    VIRUS-REPLICATION-CYCLE & 154 &&& 1000+ & DNF & 1.16\\
-    Executable\_file\_for\_Alzheimer\_model & 1169 &&& 1000+ & DNF & 6.15\\
-    ER-STRESS & 182 &&& 1000+ & 16.15 & 0.93\\
-    RA\_apoptosis\_executable\_module & 180 &&& 1000+ & DNF & 0.96\\
+    model & $n$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
+    CASCADE3 & 183 & 1 & 82.86 & 12.83\\
+    KYNURENINE-PATHWAY & 150 & 1000+ & 74.71 & 4.22\\
+    SN5 & 2746 & ??? & DNF & DNF\\
+    ERBB-RECEPTOR-SIGNALING & 247 & 1000+ & DNF & 9.27\\
+    Executable\_file\_for\_cholocystokinin\_model & 404 & 1000+ & 1.55 & 3.57\\
+    korkut\_2015a & 99 & 1000+ & DNF & 3.06\\
+    SIGNALING-PATHWAY-FOR-BUTANOL-PRODUCTION & 66 & 1000+ & 0.43 & 0.66\\
+    Corral\_ThIL17diff\_15jan2021 & 92 & 1000+ & DNF & 0.91\\
+    HIV-1 & 138 & 1000+ & DNF & 3.25\\
+    Executable\_file\_for\_CaSQ\_derived\_mast\_cell\_activation\_model & 80 & 1000+ & 0.49 & 0.56\\
+    Colitis\_associated\_colon\_cancer & 70 & 10 & 0.15 & 0.22\\
+    INTERFERON-1 & 121 & 1000+ & 7.69 & 1.05\\
+    TLGLSurvival & 61 & 318 & 0.59 & 0.48\\
+    cho\_2016\_all\_pos & 200 & 1000+ & DNF & 3.00\\
+    turei\_2016 & 4691 & ??? & DNF & DNF\\
+    Human\_network & 1953 & 1000+ & DNF & DNF\\
+    Executable\_file\_for\_CaSQ\_derived\_MAPK\_model & 182 & 1000+ & 75.32 & 2.26\\
+    TCR-TLR5-SIGNALING-2018 & 130 & 48 & 0.93 & 0.53\\
+    Hernandez\_TcellCheckPoints\_13april2020 & 218 & 1000+ & 36.21 & 2.27\\
+    IMMUNE-SYSTEM & 164 & 1000+ & DNF & 2.78\\
+    HMOX1-PATHWAY & 145 & 1000+ & 3.89 & 0.95\\
+    Signaling\_in\_Macrophage\_Activation & 321 & 1000+ & 11.51 & 3.27\\
+    FIBROBLASTS & 139 & 1000+ & DNF & 1.91\\
+    S1\_Table & 1659 & 1000+ & DNF & DNF\\
+    Regan2020\_Adhesion\_CIP\_Migration\_CellCycle\_Apoptosis & 121 & 78 & 40.59 & 2.15\\
+    SIGNALING-IN-PROSTATE-CANCER & 133 & 1000+ & DNF & 1.70\\
+    InflammatoryBowelDisease & 47 & 1 & DNF & 118.78\\
+    Influenza\_A\_Virus\_Replication\_Cycle & 131 & 1000+ & 47.69 & 0.87\\
+    IL\_6\_Signalling & 86 & 1000+ & 1.10 & 0.46\\
+    VIRUS-REPLICATION-CYCLE & 154 & 1000+ & DNF & 1.16\\
+    Executable\_file\_for\_Alzheimer\_model & 1169 & 1000+ & DNF & 6.15\\
+    ER-STRESS & 182 & 1000+ & 16.15 & 0.93\\
+    RA\_apoptosis\_executable\_module & 180 & 1000+ & DNF & 0.96\\
     \bottomrule

### 2022-05-10 cmsb22.tex
Details
Updated .bnet files when available

@@
   \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
-    Columns "\(n\)" denotes the numbers of nodes of each model. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+    Column "\(n\)" denotes the numbers of nodes of each model. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
     "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes.}
@@
     ERBB-RECEPTOR-SIGNALING & 247 & 1000+ & DNF & 9.27\\
-    Executable\_file\_for\_cholocystokinin\_model & 404 & 1000+ & 1.55 & 3.57\\
+    Executable\_file\_for\_cholocystokinin\_model & 383 & 1000+ & 1.55 & 3.57\\
     korkut\_2015a & 99 & 1000+ & DNF & 3.06\\
@@
     HIV-1 & 138 & 1000+ & DNF & 3.25\\
-    Executable\_file\_for\_CaSQ\_derived\_mast\_cell\_activation\_model & 80 & 1000+ & 0.49 & 0.56\\
+    Executable\_file\_for\_CaSQ\_derived\_mast\_cell\_activation\_model & 73 & 1000+ & 0.49 & 0.56\\
     Colitis\_associated\_colon\_cancer & 70 & 10 & 0.15 & 0.22\\
@@
     Human\_network & 1953 & 1000+ & DNF & DNF\\
-    Executable\_file\_for\_CaSQ\_derived\_MAPK\_model & 182 & 1000+ & 75.32 & 2.26\\
+    Executable\_file\_for\_CaSQ\_derived\_MAPK\_model & 181 & 1000+ & 75.32 & 2.26\\
     TCR-TLR5-SIGNALING-2018 & 130 & 48 & 0.93 & 0.53\\
@@
     VIRUS-REPLICATION-CYCLE & 154 & 1000+ & DNF & 1.16\\
-    Executable\_file\_for\_Alzheimer\_model & 1169 & 1000+ & DNF & 6.15\\
+    Executable\_file\_for\_Alzheimer\_model & 762 & 1000+ & DNF & 6.15\\
     ER-STRESS & 182 & 1000+ & 16.15 & 0.93\\

### 2022-05-10 cmsb22.tex
Updates from Overleaf

--- 2022-05-10 cmsb22.tex
@@
 Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
+    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. A number in bold indicates a ratio greater than three.}
   \centering
@@
     model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
-    arellano\_rootstem & 9 &&& 4 & 0.04 & \textbf{0.31}\\
-    calzone\_cellfate & 28 &&& 27 & 0.03 & 0.08\\
-    dahlhaus\_neuroplastoma & 23 &&& 32 & 0.04 & 0.11\\
-    davidich\_yeast & 10 &&& 12 & 0.03 & 0.08\\
-    dinwoodie\_life & 15 &&& 7 & 0.03 & 0.08\\
-    dinwoodie\_stomatal & 13 &&& 1 & 0.04 & 0.05\\
-    faure\_cellcycle & 10 &&& 2 & 0.04 & 0.04\\
-    grieco\_mapk & 53 &&& 18 & 0.04 & 0.08\\
-    irons\_yeast & 18 &&& 1 & 0.05 & 0.04\\
-    jaoude\_thdiff & 103 &&& 1000+ & \textbf{1.73} & 0.40\\
-    klamt\_tcr & 40 &&& 8 & 0.06 & 0.04\\
-    krumsiek\_myeloid & 11 &&& 6 & 0.03 & 0.03\\
-    multivalued & 13 &&& 4 & 0.02 & 0.02\\
-    n12c5 & 11 &&& 5 & \textbf{34.37} & 0.06\\
-    n3s1c1a & 2 &&& 2 & 0.02 & 0.03\\
-    n3s1c1b & 2 &&& 2 & 0.02 & 0.03\\
-    n5s3 & 4 &&& 3 & 0.03 & 0.03\\
-    n6s1c2 & 5 &&& 3 & 0.03 & 0.04\\
-    n7s3 & 6 &&& 3 & 0.02 & 0.02\\
-    raf & 3 &&& 2 & 0.03 & 0.02\\
-    randomnet\_n15k3 & 15 &&& 3 & 0.04 & 0.05\\
-    randomnet\_n7k3 & 7 &&& 10 & 0.03 & \textbf{0.19}\\
-    remy\_tumorigenesis & 34 &&& 25 & \textbf{1.79} & 0.13\\
-    saadatpour\_guardcell & 13 &&& 1 & 0.02 & 0.03\\
-    selvaggio\_emt & 56 &&& 1000+ & \textbf{1.03} & 0.23\\
-    tournier\_apoptosis & 12 &&& 3 & 0.05 & 0.06\\
-    xiao\_wnt5a & 7 &&& 4 & 0.06 & 0.03\\
-    zhang\_tlgl & 60 &&& 156 & 0.26 & 0.17\\
-    zhang\_tlgl\_v2 & 60 &&& 258 & 0.19 & 0.17\\
+    arellano\_rootstem & 9 && 1 & 4 & 0.04 & \textbf{0.31}\\
+    calzone\_cellfate & 28 && 3 & 27 & 0.03 & 0.08\\
+    dahlhaus\_neuroplastoma & 23 && 5 & 32 & 0.04 & 0.11\\
+    davidich\_yeast & 10 && 0 & 12 & 0.03 & 0.08\\
+    dinwoodie\_life & 15 && 0 & 7 & 0.03 & 0.08\\
+    dinwoodie\_stomatal & 13 && 0 & 1 & 0.04 & 0.05\\
+    faure\_cellcycle & 10 && 1 & 2 & 0.04 & 0.04\\
+    grieco\_mapk & 53 && 4 & 18 & 0.04 & 0.08\\
+    irons\_yeast & 18 && 0 & 1 & 0.05 & 0.04\\
+    jaoude\_thdiff & 103 && 21 & 1000+ & \textbf{1.73} & 0.40\\
+    klamt\_tcr & 40 && 3 & 8 & 0.06 & 0.04\\
+    krumsiek\_myeloid & 11 && 0 & 6 & 0.03 & 0.03\\
+    multivalued & 13 && 1 & 4 & 0.02 & 0.02\\
+    n12c5 & 11 && 0 & 5 & \textbf{34.37} & 0.06\\
+    n3s1c1a & 2 && 0 & 2 & 0.02 & 0.03\\
+    n3s1c1b & 2 && 0 & 2 & 0.02 & 0.03\\
+    n5s3 & 4 && 0 & 3 & 0.03 & 0.03\\
+    n6s1c2 & 5 && 0 & 3 & 0.03 & 0.04\\
+    n7s3 & 6 && 0 & 3 & 0.02 & 0.02\\
+    raf & 3 && 0 & 2 & 0.03 & 0.02\\
+    randomnet\_n15k3 & 15 && 0 & 3 & 0.04 & 0.05\\
+    randomnet\_n7k3 & 7 && 0 & 10 & 0.03 & \textbf{0.19}\\
+    remy\_tumorigenesis & 34 && 4 & 25 & \textbf{1.79} & 0.13\\
+    saadatpour\_guardcell & 13 && 0 & 1 & 0.02 & 0.03\\
+    selvaggio\_emt & 56 && 10 & 1000+ & \textbf{1.03} & 0.23\\
+    tournier\_apoptosis & 12 && 1 & 3 & 0.05 & 0.06\\
+    xiao\_wnt5a & 7 && 1 & 4 & 0.06 & 0.03\\
+    zhang\_tlgl & 60 && 6 & 156 & 0.26 & 0.17\\
+    zhang\_tlgl\_v2 & 60 && 6 & 258 & 0.19 & 0.17\\
     \bottomrule

### 2022-05-10 cmsb22.tex
Some small corrections

--- 2022-05-10 cmsb22.tex
@@
 The recent use of trap spaces~\cite{klarner2015computing} made a real breakthrough in that field allowing to consider medium-sized models that used to be out of reach.
-However, with the continuing increase in model-size, the state of the art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of implicants.
+However, with the continuing increase in model-size, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of implicants.
 
-Petri nets were introduced in the 60’s as a simple formalism for describing and analyzing information processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
+Petri nets were introduced in the 60’s as simple formalism for describing and analyzing information processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
 The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, was introduced quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
@@
 It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
-After some technical preliminaries, we expose the concrete need for such a method and detail its implementation using Answer Set Programming (ASP).
-We then demonstrate its efficiency and compare it to implicant-based methods on some large Boolean models from the literature.
+After some technical preliminaries, we expose the concrete need for such a method and detail its implementation using Answer Set Programming.
+We then demonstrate its efficiency and compare it to implicant-based methods on several large Boolean models from the literature.
 
@@
     Column "\(n\)" denotes the numbers of nodes of each model. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
+    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. A number in bold indicates a ratio greater than three.}
   \centering

### 2022-05-10 cmsb22.tex
Paper updated w/ latest results

@@
   \caption{Timing comparisons between the proposed method and PyBoolNet on the PyBoolNet's repository.
-    Column "\(n\)" denotes the numbers of nodes of each model. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. A number in bold indicates a ratio greater than three.}
+    Column \(n\) denotes the numbers of nodes of each model. Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+    ``DNF'' means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. A number in bold indicates a ratio greater than three.}
   \centering
@@
   \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
-    Column "\(n\)" denotes the numbers of nodes of each model. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes.}
+    Columns \(n\) and \(s\) denote the numbers of nodes (resp.\ input nodes) of each model.
+Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+    ``DNF'' means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes.}
   \centering
@@
     \toprule
-    model & $n$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
-    CASCADE3 & 183 & 1 & 82.86 & 12.83\\
-    KYNURENINE-PATHWAY & 150 & 1000+ & 74.71 & 4.22\\
-    SN5 & 2746 & ??? & DNF & DNF\\
-    ERBB-RECEPTOR-SIGNALING & 247 & 1000+ & DNF & 9.27\\
-    Executable\_file\_for\_cholocystokinin\_model & 383 & 1000+ & 1.55 & 3.57\\
-    korkut\_2015a & 99 & 1000+ & DNF & 3.06\\
-    SIGNALING-PATHWAY-FOR-BUTANOL-PRODUCTION & 66 & 1000+ & 0.43 & 0.66\\
-    Corral\_ThIL17diff\_15jan2021 & 92 & 1000+ & DNF & 0.91\\
-    HIV-1 & 138 & 1000+ & DNF & 3.25\\
-    Executable\_file\_for\_CaSQ\_derived\_mast\_cell\_activation\_model & 73 & 1000+ & 0.49 & 0.56\\
-    Colitis\_associated\_colon\_cancer & 70 & 10 & 0.15 & 0.22\\
-    INTERFERON-1 & 121 & 1000+ & 7.69 & 1.05\\
-    TLGLSurvival & 61 & 318 & 0.59 & 0.48\\
-    cho\_2016\_all\_pos & 200 & 1000+ & DNF & 3.00\\
-    turei\_2016 & 4691 & ??? & DNF & DNF\\
-    Human\_network & 1953 & 1000+ & DNF & DNF\\
-    Executable\_file\_for\_CaSQ\_derived\_MAPK\_model & 181 & 1000+ & 75.32 & 2.26\\
-    TCR-TLR5-SIGNALING-2018 & 130 & 48 & 0.93 & 0.53\\
-    Hernandez\_TcellCheckPoints\_13april2020 & 218 & 1000+ & 36.21 & 2.27\\
-    IMMUNE-SYSTEM & 164 & 1000+ & DNF & 2.78\\
-    HMOX1-PATHWAY & 145 & 1000+ & 3.89 & 0.95\\
-    Signaling\_in\_Macrophage\_Activation & 321 & 1000+ & 11.51 & 3.27\\
-    FIBROBLASTS & 139 & 1000+ & DNF & 1.91\\
-    S1\_Table & 1659 & 1000+ & DNF & DNF\\
-    Regan2020\_Adhesion\_CIP\_Migration\_CellCycle\_Apoptosis & 121 & 78 & 40.59 & 2.15\\
-    SIGNALING-IN-PROSTATE-CANCER & 133 & 1000+ & DNF & 1.70\\
-    InflammatoryBowelDisease & 47 & 1 & DNF & 118.78\\
-    Influenza\_A\_Virus\_Replication\_Cycle & 131 & 1000+ & 47.69 & 0.87\\
-    IL\_6\_Signalling & 86 & 1000+ & 1.10 & 0.46\\
-    VIRUS-REPLICATION-CYCLE & 154 & 1000+ & DNF & 1.16\\
-    Executable\_file\_for\_Alzheimer\_model & 762 & 1000+ & DNF & 6.15\\
-    ER-STRESS & 182 & 1000+ & 16.15 & 0.93\\
-    RA\_apoptosis\_executable\_module & 180 & 1000+ & DNF & 0.96\\
+    model & \(n\) & \(s\) & \(|M|\) & PyBoolNet (s) & Trappist (s) \\
+    \midrule
+    InflammatoryBowelDisease & 47 & 0 & 1 & DNF & 21.90\\
+    TLGLSurvival & 61 & 7 & 318 & 0.73 & 0.14\\
+    SIGNALING-PATHWAY-FOR-BUTANOL-PRODUCTION & 66 & 13 & 1000+ & 0.59 & 0.13\\
+    Colitis\_associated\_colon\_cancer & 70 & 1 & 10 & 0.16 & 0.08\\
+    mast\_cell\_activation\_model\_BCC & 73 & 19 & 1000+ & 0.64 & 0.15\\
+    IL\_6\_Signalling & 86 & 15 & 1000+ & 3.30 & 0.10\\
+    Corral\_ThIL17diff\_15jan2021 & 92 & 16 & 1000+ & DNF & 0.38\\
+    korkut\_2015a & 99 & 12 & 1000+ & DNF & 0.61\\
+    INTERFERON-1 & 121 & 55 & 1000+ & 8.36 & 0.19\\
+    Regan2020\_Adhesion\_CIP\_Migration\_CellCycle\_Apoptosis & 121 & 4 & 78 & 40.81 & 0.33\\
+    TCR-TLR5-SIGNALING-2018 & 130 & 5 & 48 & 0.88 & 0.12\\
+    Influenza\_A\_Virus\_Replication\_Cycle & 131 & 11 & 1000+ & 63.62 & 0.21\\
+    SIGNALING-IN-PROSTATE-CANCER & 133 & 11 & 1000+ & DNF & 0.32\\
+    HIV-1 & 138 & 14 & 1000+ & DNF & 0.87\\
+    FIBROBLASTS & 139 & 9 & 1000+ & DNF & 0.56\\
+    HMOX1-PATHWAY & 145 & 56 & 1000+ & 5.16 & 0.18\\
+    Kynurenine\_Pathway\_stable & 150 & 72 & 1000+ & DNF & 0.74\\
+    VIRUS-REPLICATION-CYCLE & 154 & 25 & 1000+ & DNF & 0.19\\
+    IMMUNE-SYSTEM & 164 & 13 & 1000+ & DNF & 0.67\\
+    RA\_apoptosis\_executable\_module\_BCC & 180 & 59 & 1000+ & DNF & 0.16\\
+    MAPK\_model\_BCC & 181 & 37 & 1000+ & 68.62 & 0.64\\
+    ER-STRESS & 182 & 75 & 1000+ & 28.65 & 0.21\\
+    CASCADE3 & 183 & 1 & 0 & 92.75 &  0.65\\
+    cho\_2016\_all\_pos & 200 & 13 & 1000+ & DNF & 1.01\\
+    Hernandez\_TcellCheckPoints\_13april2020 & 218 & 14 & 1000+ & 37.40 & 0.35\\
+    ERBB-RECEPTOR-SIGNALING & 247 & 22 & 1000+ & DNF & 1.07\\
+    Signaling\_in\_Macrophage\_Activation & 321 & 19 & 1000+ & 15.82 & 0.63\\
+    cholocystokinin\_model\_BCC & 383 & 74 & 1000+ & 1.22 & 0.73\\
+    Alzheimer\_model\_BCC & 762 & 237 & 1000+ & DNF & 1.03\\
+    S1\_Table & 1659 & 521 & 1000+ & DNF & 44.85\\
+    Human\_network & 1953 & 669 & 1000+ & DNF & 42.97\\
+    SN5 & 2746 & 829 & ??? & DNF & DNF\\
+    turei\_2016 & 4691 & 1257 & ??? & DNF & DNF\\
     \bottomrule
@@
 
-\vangiang{Regarding the time differences, the reason might be that my implementation uses an ASP encoding based on choice rules instead of disjunctive rules or constraint rules.}
-\sylvain{Maybe\dots I don't know ASP well enough to be sure. The fact that you go through a predicate \texttt{p} which adds one level of indirection might also have some effect, I'm not sure.}
+% s obtained by: grep -c '^\([a-zA-Z0-9_-]*\),\s*\1\s*\(#.*\)\?$' models/*.bnet
+
 \vangiang{Oh. I think that this may be a reason. However, I think we should still use a predicate \texttt{p(v, x)} in the paper because of the readability.}
+\sylvain{To be honest, I don't think it is more readable\dots}
 

### 2022-05-10 cmsb22.tex
Updates from Overleaf
yalafi (languagetool)

@@
 
-Petri nets were introduced in the 60’s as simple formalism for describing and analyzing information processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
+Petri nets were introduced in the 60s as a simple formalism for describing and analyzing information processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
 The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, was introduced quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
-Siphons are such a concept but they have not been used a lot for the analysis of biochemical systems even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{nabli2016enumerating}.
+Siphons are such a concept, but they have not been used a lot for the analysis of biochemical systems even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{nabli2016enumerating}.
 
@@
 A place \(p\) is marked by a marking \(m\) if and only if \(m(p) > 0\).
-We shall write \(pred(x)\) (resp.\ \(succ(x)\)) to represent the set of vertices that have an (non-zero weighted) arc leading to (resp.\ coming from) \(x\).
+We shall write \(pred(x)\) (resp.\ \(succ(x)\)) to represent the set of vertices that have a (non-zero weighted) arc leading to (resp.\ coming from) \(x\).
 
 The link between Logical models \emph{à la} Thomas and Petri nets was originally established in~\cite{chaouiya2004qualitative} in order to make available formal methods like model-checking for the analysis of such systems.
-The basic encoding into 1-safe (i.e., never more than one token in each place) nets only holds for purely Boolean models but was later extended to multi-valued models in two ways, either in \cite{chaouiya2011petri} with non 1-safe Petri nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.
+The basic encoding into 1-safe (i.e., never more than one token in each place) nets only holds for purely Boolean models but was later extended to multivalued models in two ways, either in~\cite{chaouiya2011petri} with non 1-safe Petri nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.
 
@@
 Its basis is that every node (\emph{gene}) \(v\) of the original model \(\mathcal{M} = (V, F)\) is represented by two separate places (\(p_v\) and \(\overline{p}_v\)), corresponding to its two states, active, and inactive, respectively.
-Each conjunct of the logical function that activates the \emph{gene} will lead to a transition \(t\), consuming the inactive place (i.e., a directional arc from \(\overline{p}_v\) to \(t\)), producing the active place (i.e., a directional arc from \(t\) to \(p_v\)), and with all other literals both consumed and produced (i.e., a bi-directional arc).
+Each conjunct of the logical function that activates the \emph{gene} will lead to a transition \(t\), consuming the inactive place (i.e., a directional arc from \(\overline{p}_v\) to \(t\)), producing the active place (i.e., a directional arc from \(t\) to \(p_v\)), and with all other literals both consumed and produced (i.e., a bidirectional arc).
 And conversely for the inactivation. 
-Let \(s\) be a state of the Boolean model and \(m_s\) be its the corresponding marking in the encoded Petri net.
+Let \(s\) be a state of the Boolean model and \(m_s\) be its corresponding marking in the encoded Petri net.
 It holds that \(\forall v \in V\), \(s(v) = 0\) if and only if \(m_s(\overline{p}_v) = 1\) and \(s(v) = 1\) if and only if \(m_s(p_v) = 1\).
@@
 Without loss of generality, we assume that \(p = p_v\), then there is a directional arc from \(t\) to \(p_v\) and a directional arc from \(\overline{p}_v\) to \(t\).
-In addition, there is also no arc or a bi-directional arc between \(t\) and another place rather than \(p_v\) and \(\overline{p}_v\).
+In addition, there is also no arc or a bidirectional arc between \(t\) and another place rather than \(p_v\) and \(\overline{p}_v\).
 Thus, there is no connecting arc between \(t\) and any place in \(S \setminus \{p_v\}\) because \(S\cap pred(t)\not =\emptyset\).
@@
 p \in S \Rightarrow \bigvee_{p' \in pred(t)}p' \in S.
-\end{equation}The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net~\cite{nabli2016enumerating}.
+\end{equation}
+The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net~\cite{nabli2016enumerating}.
 To make \(S\) to be a conflict-free siphon, we need to add to the system the rule
@@
 \verb|{p(v, 1) ; p(v, 0)}.|
-\] for each \(v \in V\).
+\]
+for each \(v \in V\).
 Note that the number of atoms of \(\mathcal{A}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
@@
 In particular, it provides a feature-rich alternative to random simulations for this type of very non-deterministic model.
-Being able to detect that there are inconsistencies with published experimental data in some of the, say first 100 attractors, can lead to a much quicker Systems Biology loop: model, invalidate, refine.
+Being able to detect that there are inconsistencies with published experimental data in some of the first 1000 attractors, for instance, can lead to a much quicker Systems Biology loop: model, invalidate, refine.
 
@@
 Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heu=domain --dom-mod=7} used by PyBoolNet).
-We ran all the benchmarks on a virtual machine whose environment is CPU: Intel(R) Core(TM) i7-3630QM 2.40GHz x 4, Memory: 8 GB, Ubuntu 18.04.2 64 bit.
+We ran all the benchmarks on a virtual machine whose environment is CPU: Intel® Core™ i7-3630QM 2.40GHz x 4, Memory: 8 GB, Ubuntu 18.04.2 64 bit.
 Finally, we set a time limit as one hour for each model.

### 2022-05-10 cmsb22.tex
Update model names

--- 2022-05-10 cmsb22.tex
@@
     \midrule
-    InflammatoryBowelDisease & 47 & 0 & 1 & DNF & 21.90\\
-    TLGLSurvival & 61 & 7 & 318 & 0.73 & 0.14\\
-    SIGNALING-PATHWAY-FOR-BUTANOL-PRODUCTION & 66 & 13 & 1000+ & 0.59 & 0.13\\
-    Colitis\_associated\_colon\_cancer & 70 & 1 & 10 & 0.16 & 0.08\\
-    mast\_cell\_activation\_model\_BCC & 73 & 19 & 1000+ & 0.64 & 0.15\\
-    IL\_6\_Signalling & 86 & 15 & 1000+ & 3.30 & 0.10\\
-    Corral\_ThIL17diff\_15jan2021 & 92 & 16 & 1000+ & DNF & 0.38\\
-    korkut\_2015a & 99 & 12 & 1000+ & DNF & 0.61\\
-    INTERFERON-1 & 121 & 55 & 1000+ & 8.36 & 0.19\\
-    Regan2020\_Adhesion\_CIP\_Migration\_CellCycle\_Apoptosis & 121 & 4 & 78 & 40.81 & 0.33\\
-    TCR-TLR5-SIGNALING-2018 & 130 & 5 & 48 & 0.88 & 0.12\\
-    Influenza\_A\_Virus\_Replication\_Cycle & 131 & 11 & 1000+ & 63.62 & 0.21\\
-    SIGNALING-IN-PROSTATE-CANCER & 133 & 11 & 1000+ & DNF & 0.32\\
-    HIV-1 & 138 & 14 & 1000+ & DNF & 0.87\\
-    FIBROBLASTS & 139 & 9 & 1000+ & DNF & 0.56\\
-    HMOX1-PATHWAY & 145 & 56 & 1000+ & 5.16 & 0.18\\
-    Kynurenine\_Pathway\_stable & 150 & 72 & 1000+ & DNF & 0.74\\
-    VIRUS-REPLICATION-CYCLE & 154 & 25 & 1000+ & DNF & 0.19\\
-    IMMUNE-SYSTEM & 164 & 13 & 1000+ & DNF & 0.67\\
-    RA\_apoptosis\_executable\_module\_BCC & 180 & 59 & 1000+ & DNF & 0.16\\
-    MAPK\_model\_BCC & 181 & 37 & 1000+ & 68.62 & 0.64\\
-    ER-STRESS & 182 & 75 & 1000+ & 28.65 & 0.21\\
-    CASCADE3 & 183 & 1 & 0 & 92.75 &  0.65\\
-    cho\_2016\_all\_pos & 200 & 13 & 1000+ & DNF & 1.01\\
-    Hernandez\_TcellCheckPoints\_13april2020 & 218 & 14 & 1000+ & 37.40 & 0.35\\
-    ERBB-RECEPTOR-SIGNALING & 247 & 22 & 1000+ & DNF & 1.07\\
-    Signaling\_in\_Macrophage\_Activation & 321 & 19 & 1000+ & 15.82 & 0.63\\
-    cholocystokinin\_model\_BCC & 383 & 74 & 1000+ & 1.22 & 0.73\\
-    Alzheimer\_model\_BCC & 762 & 237 & 1000+ & DNF & 1.03\\
-    S1\_Table & 1659 & 521 & 1000+ & DNF & 44.85\\
-    Human\_network & 1953 & 669 & 1000+ & DNF & 42.97\\
-    SN5 & 2746 & 829 & ??? & DNF & DNF\\
-    turei\_2016 & 4691 & 1257 & ??? & DNF & DNF\\
+    inflammatory-bowel-disease & 47 & 0 & 1 & DNF & 21.90\\
+    T-LGL-survival-network-2008 & 61 & 7 & 318 & 0.73 & 0.14\\
+    butanol-production & 66 & 13 & 1000+ & 0.59 & 0.13\\
+    colitis-associated-colon-cancer & 70 & 1 & 10 & 0.16 & 0.08\\
+    mast-cell-activation & 73 & 19 & 1000+ & 0.64 & 0.15\\
+    IL-6-signalling & 86 & 15 & 1000+ & 3.30 & 0.10\\
+    Corral-ThIL-17-diff & 92 & 16 & 1000+ & DNF & 0.38\\
+    Korkut-2015 & 99 & 12 & 1000+ & DNF & 0.61\\
+    adhesion-cip-migration-cell-cycle & 121 & 4 & 78 & 40.81 & 0.33\\
+    interferon-1 & 121 & 55 & 1000+ & 8.36 & 0.19\\
+    TCR-TLR5-signaling-2018 & 130 & 5 & 48 & 0.88 & 0.12\\
+    influenza-virus-replication-cycle & 131 & 11 & 1000+ & 63.62 & 0.21\\
+    signaling-in-prostate-cancer & 133 & 11 & 1000+ & DNF & 0.32\\
+    HIV-1-with-T-cell-signaling & 138 & 14 & 1000+ & DNF & 0.87\\
+    signal-transduction-in-fibroblasts & 139 & 9 & 1000+ & DNF & 0.56\\
+    HMOX-1-pathway & 145 & 56 & 1000+ & 5.16 & 0.18\\
+    kynurenine-pathway & 150 & 72 & 1000+ & DNF & 0.74\\
+    virus-replication-cycle & 154 & 25 & 1000+ & DNF & 0.19\\
+    immune-system & 164 & 13 & 1000+ & DNF & 0.67\\
+    RA-apoptosis & 180 & 59 & 1000+ & DNF & 0.16\\
+    MAPK & 181 & 37 & 1000+ & 68.62 & 0.64\\
+    er-stress & 182 & 75 & 1000+ & 28.65 & 0.21\\
+    cascade-3 & 183 & 0 & 1 & 92.75 &  0.65\\
+    CHO-2016 & 200 & 13 & 1000+ & DNF & 1.01\\
+    T-cell-check-point & 218 & 14 & 1000+ & 37.40 & 0.35\\
+    ErbB-receptor-signaling & 247 & 22 & 1000+ & DNF & 1.07\\
+    signaling-macrophage-activation & 321 & 19 & 1000+ & 15.82 & 0.63\\
+    cholocystokinin & 383 & 74 & 1000+ & 1.22 & 0.73\\
+    Alzheimer & 762 & 237 & 1000+ & DNF & 1.03\\
+    KEGG-network & 1659 & 521 & 1000+ & DNF & 44.85\\
+    human-network & 1953 & 669 & 1000+ & DNF & 42.97\\
+    SN-5 & 2746 & 829 & ??? & DNF & DNF\\
+    turei-2016 & 4691 & 1257 & ??? & DNF & DNF\\
     \bottomrule

### 2022-05-10 cmsb22.tex
Update the ASP encoding

@@
 
-Then, we translate the above characterization into the ASP \(\mathcal{A}\) as follows.
-We introduce atom \verb|p(v, 1)| (resp.
-\verb|p(v, 0)|) to denote place \(p_v\) (resp.
+Then, we translate the above characterization into the ASP \(\mathcal{L}\) as follows.
+We introduce atom \verb|p-v| (resp.
+\verb|n-v|) to denote place \(p_v\) (resp.
 \(\overline{p_v}\)), \(\forall v \in V\).
+The set of all atoms in \(\mathcal{L}\) is given as \(\mathcal{A} = \bigcup_{v \in V}\{\verb|p-v|, \verb|n-v|\}\).
 For each pair \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\), we translate the rule (\ref{eq:siphon}) into the ASP rule
 \[
-\verb|p(v_1, x_1); ... ; p(v_k, x_k) :- p(v, x).|
+\verb|a_1; ... ; a_k :- a.|
 \]
-where \verb|p(v, x)| is the atom representing place \(p\) and \{\verb|p(v_1, x_1)|, \dots,\linebreak \verb|p(v_k, x_k)|\} is the set of atoms representing places in \(pred(t)\).
+where \(\verb|a| \in \mathcal{A}\) is the atom representing place \(p\) and \(\{\verb|a_1|, \dots,\verb|a_k|\} \subseteq \mathcal{A}\) is the set of atoms representing places in \(pred(t)\).
 The rule (\ref{eq:conflict}) is translated into the ASP rule
-\[\verb|:- p(v, 1), p(v, 0).|\]
+\[\verb|:- p-v, n-v.|\]
 for each \(v \in V\). This ASP rule guarantees that two places representing the same node in \(\mathcal{M}\) never belong to the same siphon of \(\mathcal{P}\), representing the conflict-freeness.
-Naturally, a Herbrand model (see, e.g,~\cite{DBLP:journals/aicom/GebserKKOSS11}) of \(\mathcal{A}\) is equivalent to a conflict-free siphon of \(\mathcal{P}\).
-To guarantee that a Herbrand model is also a stable model (an answer set), we need to add to \(\mathcal{A}\) the choice rule
+Naturally, a Herbrand model (see, e.g,~\cite{DBLP:journals/aicom/GebserKKOSS11}) of \(\mathcal{L}\) is equivalent to a conflict-free siphon of \(\mathcal{P}\).
+To guarantee that a Herbrand model is also a stable model (an answer set), we need to add to \(\mathcal{L}\) the two choice rules
 \[
-\verb|{p(v, 1) ; p(v, 0)}.|
+\verb|{p-v}. {n-v}.|
 \]
 for each \(v \in V\).
-Note that the number of atoms of \(\mathcal{A}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
+Note that the number of atoms of \(\mathcal{L}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
 
@@
 
-Now, a solution (simply an answer set) \(A\) of \(\mathcal{A}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a minimal trap space \(m\) of \(\mathcal{M}\).
+Now, a solution (simply an answer set) \(A \subseteq \mathcal{A}\) of \(\mathcal{L}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a trap space \(m\) of \(\mathcal{M}\).
 The conversion from \(A\) to \(m\) is straightforward.
-If \(\verb|p(v, 1)| \in A\) then \(v \in D_m\) and \(m(v) = 0\).
-Conversely, if \(\verb|p(v, 0)| \in A\) then \(v \in D_m\) and \(m(v) = 1\).
+If \(\verb|p-v| \in A\) then \(v \in D_m\) and \(m(v) = 0\).
+Conversely, if \(\verb|n-v| \in A\) then \(v \in D_m\) and \(m(v) = 1\).
 Otherwise, \(v \not \in D_m\).
-Computing multiple answer sets is built into ASP solvers and the solving collection POTASSCO~\cite{DBLP:journals/aicom/GebserKKOSS11} also features the option to find set-inclusion maximal answer sets with respect to the predicates that are true.
-By using this built-in option, we can compute all the answer sets of \(\mathcal{A}\) in one execution.
+Computing multiple answer sets is built into ASP solvers and the solving collection POTASSCO~\cite{DBLP:journals/aicom/GebserKKOSS11} also features the option to find set-inclusion maximal answer sets with respect to the set of atoms.
+Naturally, a set-inclusion maximal answer set of \(\mathcal{L}\) is equivalent to a maximal conflict-free siphon of \(\mathcal{P}\), thus a minimal trap space of \(\mathcal{M}\).
+By using this built-in option, we can compute all the set-inclusion maximal answer sets of \(\mathcal{L}\) (resp.
+all the minimal trap spaces of \(\mathcal{M}\)) in one execution.
 

### 2022-05-11 cmsb22.tex
New bench, not so sure…

@@
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-    ``DNF'' means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes.}
+    ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of 2 minutes.}
   \centering
@@
     \toprule
-    model & \(n\) & \(s\) & \(|M|\) & PyBoolNet (s) & Trappist (s) \\
+    & & & & PyBoolNet & \multicolumn{2}{c}{Trappist} \\
+    \cmidrule(lr){6-7}
+    model & \(n\) & \(s\) & \(|M|\) & first 1000 & first 1000 & all \\
     \midrule
-    inflammatory-bowel-disease & 47 & 0 & 1 & DNF & 21.90\\
-    T-LGL-survival-network-2008 & 61 & 7 & 318 & 0.73 & 0.14\\
-    butanol-production & 66 & 13 & 1000+ & 0.59 & 0.13\\
-    colitis-associated-colon-cancer & 70 & 1 & 10 & 0.16 & 0.08\\
-    mast-cell-activation & 73 & 19 & 1000+ & 0.64 & 0.15\\
-    IL-6-signalling & 86 & 15 & 1000+ & 3.30 & 0.10\\
-    Corral-ThIL-17-diff & 92 & 16 & 1000+ & DNF & 0.38\\
-    Korkut-2015 & 99 & 12 & 1000+ & DNF & 0.61\\
-    adhesion-cip-migration-cell-cycle & 121 & 4 & 78 & 40.81 & 0.33\\
-    interferon-1 & 121 & 55 & 1000+ & 8.36 & 0.19\\
-    TCR-TLR5-signaling-2018 & 130 & 5 & 48 & 0.88 & 0.12\\
-    influenza-virus-replication-cycle & 131 & 11 & 1000+ & 63.62 & 0.21\\
-    signaling-in-prostate-cancer & 133 & 11 & 1000+ & DNF & 0.32\\
-    HIV-1-with-T-cell-signaling & 138 & 14 & 1000+ & DNF & 0.87\\
-    signal-transduction-in-fibroblasts & 139 & 9 & 1000+ & DNF & 0.56\\
-    HMOX-1-pathway & 145 & 56 & 1000+ & 5.16 & 0.18\\
-    kynurenine-pathway & 150 & 72 & 1000+ & DNF & 0.74\\
-    virus-replication-cycle & 154 & 25 & 1000+ & DNF & 0.19\\
-    immune-system & 164 & 13 & 1000+ & DNF & 0.67\\
-    RA-apoptosis & 180 & 59 & 1000+ & DNF & 0.16\\
-    MAPK & 181 & 37 & 1000+ & 68.62 & 0.64\\
-    er-stress & 182 & 75 & 1000+ & 28.65 & 0.21\\
-    cascade-3 & 183 & 0 & 1 & 92.75 &  0.65\\
-    CHO-2016 & 200 & 13 & 1000+ & DNF & 1.01\\
-    T-cell-check-point & 218 & 14 & 1000+ & 37.40 & 0.35\\
-    ErbB-receptor-signaling & 247 & 22 & 1000+ & DNF & 1.07\\
-    signaling-macrophage-activation & 321 & 19 & 1000+ & 15.82 & 0.63\\
-    cholocystokinin & 383 & 74 & 1000+ & 1.22 & 0.73\\
-    Alzheimer & 762 & 237 & 1000+ & DNF & 1.03\\
-    KEGG-network & 1659 & 521 & 1000+ & DNF & 44.85\\
-    human-network & 1953 & 669 & 1000+ & DNF & 42.97\\
-    SN-5 & 2746 & 829 & ??? & DNF & DNF\\
-    turei-2016 & 4691 & 1257 & ??? & DNF & DNF\\
+    inflammatory-bowel-disease & 47 & 0 & 1 & DNF & 21.90 & 25.75\\
+    T-LGL-survival-network-2008 & 61 & 7 & 318 & 0.73 & 0.14 & 0.14\\
+    butanol-production & 66 & 13 & 8192 & 0.59 & 0.13 & 0.64\\
+    colitis-associated-colon-cancer & 70 & 1 & 10 & 0.16 & 0.08 & 0.06\\
+    mast-cell-activation & 73 & 19 & 1000+ & 0.64 & 0.15 & DNF\\
+    IL-6-signalling & 86 & 15 & 32768 & 3.30 & 0.10 & 5.06\\
+    Corral-ThIL-17-diff & 92 & 16 & 1000+ & DNF & 0.38 & DNF\\
+    Korkut-2015 & 99 & 12 & 18556 & DNF & 0.61 & 5.01\\
+    adhesion-cip-migration-cell-cycle & 121 & 4 & 78 & 40.81 & 0.33 & 0.45\\
+    interferon-1 & 121 & 55 & 1000+ & 8.36 & 0.19 & DNF\\
+    TCR-TLR5-signaling-2018 & 130 & 5 & 48 & 0.88 & 0.12 & 0.11\\
+    influenza-virus-replication-cycle & 131 & 11 & 10128 & 63.62 & 0.21 & 1.95\\
+    signaling-in-prostate-cancer & 133 & 11 & 2760 & DNF & 0.32 & 0.72\\
+    HIV-1-with-T-cell-signaling & 138 & 14 & 39424 & DNF & 0.87 & 7.72\\
+    signal-transduction-in-fibroblasts & 139 & 9 & 1000+ & DNF & 0.56 & DNF\\
+    HMOX-1-pathway & 145 & 56 & 1000+ & 5.16 & 0.18 & DNF\\
+    kynurenine-pathway & 150 & 72 & 1000+ & DNF & 0.74 & DNF\\
+    virus-replication-cycle & 154 & 25 & 1000+ & DNF & 0.19 & DNF\\
+    immune-system & 164 & 13 & 1000+ & DNF & 0.67 & DNF\\
+    RA-apoptosis & 180 & 59 & 1000+ & DNF & 0.16 & DNF\\
+    MAPK & 181 & 37 & 1000+ & 68.62 & 0.64 & DNF\\
+    er-stress & 182 & 75 & 1000+ & 28.65 & 0.21 & DNF\\
+    cascade-3 & 183 & 0 & 1 & 92.75 &  0.65 & 0.83\\
+    CHO-2016 & 200 & 13 & 13312 & DNF & 1.01 & 1.99\\
+    T-cell-check-point & 218 & 14 & 1000+ & 37.40 & 0.35 & DNF\\
+    ErbB-receptor-signaling & 247 & 22 & 1000+ & DNF & 1.07 & DNF\\
+    signaling-macrophage-activation & 321 & 19 & 1000+ & 15.82 & 0.63 & DNF\\
+    cholocystokinin & 383 & 74 & 1000+ & 1.22 & 0.73 & DNF\\
+    Alzheimer & 762 & 237 & 1000+ & DNF & 1.03 & DNF\\
+    KEGG-network & 1659 & 521 & 1000+ & DNF & 44.85 & DNF\\
+    human-network & 1953 & 669 & 1000+ & DNF & 42.97 & DNF\\
+    SN-5 & 2746 & 829 & ??? & DNF & DNF & DNF\\
+    turei-2016 & 4691 & 1257 & ??? & DNF & DNF & DNF\\
     \bottomrule
@@
 
-\vangiang{Oh. I think that this may be a reason. However, I think we should still use a predicate \texttt{p(v, x)} in the paper because of the readability.}
-\sylvain{To be honest, I don't think it is more readable\dots}
-
 \end{document}

### 2022-05-11 cmsb22.tex
A new option for benchmarking selected models

@@
 
-  \sylvain{I'm not sure that \(e\) and \(s\) are very relevant, moreover they are not readily available\dots}
-  \vangiang{Indeed, they are not readily available.
+%   \sylvain{I'm not sure that \(e\) and \(s\) are very relevant, moreover they are not readily available\dots}
+%   \vangiang{Indeed, they are not readily available.
 But if we use more meaningful words, the width of the table may exceed the width of the page.
@@
 Do you have any other option?}
-  \sylvain{just remove them since they are not relevant}
+%   \sylvain{just remove them since they are not relevant}
 
@@
 
+\begin{table}[!htb]
+  \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
+    Columns \(n\) and \(s\) denote the numbers of nodes (resp.\ input nodes) of each model.
+For each method, Column \(|M|\) denotes the number of minimal trap spaces and the computation time is given in seconds.
+    ``DNF'' means that the method did not finish the computation within the timeout of 2 minutes.}
+  \centering
+  \label{tab:selected_new_option}
+  \begin{tabular}{lrrrrrr}
+    \toprule
+    & & & \multicolumn{2}{c}{PyBoolNet} & \multicolumn{2}{c}{Trappist} \\
+    \cmidrule(lr){4-5} \cmidrule(lr){6-7}
+    model & \(n\) & \(s\) & \(|M|\) & time (s) & \(|M|\) & time (s) \\
+    \midrule
+    inflammatory-bowel-disease & 47 & 0 & 0+ & DNF & 1 & 25.75 \\
+    T-LGL-survival-network-2008 & 61 & 7 & 318 & 0.73 & 318 & 0.14 \\
+    ErbB-receptor-signaling & 247 & 22 & 0+ & DNF & 1456394+ & DNF \\
+    SN-5 & 2746 & 829 & 0+ & DNF & 0+ & DNF \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\vangiang{What do you think about the following option?
+We do not set a limit for the number of minimal trap space but a limit of 2 minutes for the computation time.
+There will be 4 cases (see Table~\ref{tab:selected_new_option}).
+Case 1: both methods finished (e.g., the T-LGL-survival-network-2008 model).
+In this case, we compare the computation time between two methods.
+Case 2: PyBoolNet finished and Trappist did not finish.
+In this case, PyBoolNet is of course better than Trappist (but no model for this case).
+Case 3: PyBoolNet did not finish and Trappist finished (e.g., the inflammatory-bowel-disease).
+In this case, Trappist is of course better than PyBoolNet.
+Moreover, 0+ indicates that PyBoolNet even failed at the computation of prime-implicants.
+Case 4: both method did not finish (e.g., the ErbB-receptor-signaling model).
+In this case, we compare the numbers of minimal trap spaces obtained by the two methods within the timeout (more is better).
+Moreover, 0+ also indicates that Trappist even failed at the conversion to PNML models.}
+
 % s obtained by: grep -c '^\([a-zA-Z0-9_-]*\),\s*\1\s*\(#.*\)\?$' models/*.bnet

### 2022-05-11 cmsb22.tex
Come back to the original option

@@
 
+\begin{table}[!htb]
+  \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
+    Columns \(n\) and \(s\) denote the numbers of nodes (resp.\ input nodes) of each model.
+For each method, Column \(|M|\) denotes the number of minimal trap spaces and the computation time is given in seconds.
+    ``DNF'' means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes.}
+  \centering
+  \label{tab:selected_original_option}
+  \begin{tabular}{lrrrrrr}
+    \toprule
+    & & & \multicolumn{2}{c}{PyBoolNet} & \multicolumn{2}{c}{Trappist} \\
+    \cmidrule(lr){4-5} \cmidrule(lr){6-7}
+    model & \(n\) & \(s\) & \(|M|\) & time (s) & \(|M|\) & time (s) \\
+    \midrule
+    inflammatory-bowel-disease & 47 & 0 & 0+ & DNF & 1 & 21.90 \\
+    T-LGL-survival-network-2008 & 61 & 7 & 318 & 0.73 & 318 & 0.14 \\
+    SN-5 & 2746 & 829 & 0+ & DNF & 0+ & DNF \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\vangiang{Original option: We set a limit of 1000 for the number of minimal trap spaces and a limit of 2 minutes for the computation time.
+There will be 4 cases (see Table~\ref{tab:selected_original_option}).
+Case 1: both methods finished (e.g., the T-LGL-survival-network-2008 model).
+In this case, we simply compare the computation time between two methods.
+Case 2: PyBoolNet finished and Trappist did not finish.
+In this case, PyBoolNet is of course better than Trappist (but no model for this case).
+Case 3: PyBoolNet did not finish and Trappist finished (e.g., the inflammatory-bowel-disease).
+In this case, Trappist is of course better than PyBoolNet.
+Moreover, 0+ indicates that PyBoolNet even failed at the computation of prime-implicants.
+Case 4: both methods did not finish.
+In this case, we consider the following subcases.
+Case 4-1: PyBoolNet is 0+ and Trappist is 0+ (e.g., the SN-5 model), i.e., PyBoolNet failed at the computation of prime-implicants and Trappist failed at the conversion to PNML models.
+Case 4-2: PyBoolNet is 0+ and Trappist is 1000+ (no model for this case), i.e., Trappist is better than PyBoolNet.
+Case 4-3: PyBoolNet is 1000+ and Trappist is 0+ (no model for this subcase), i.e., PyBoolNet is better than Trappist.
+Case 4-4: PyBoolNet is 1000+ and Trappist is 1000+ (no model for this case), i.e., we cannot compare between PyBoolNet and Trappist.}
+
 % s obtained by: grep -c '^\([a-zA-Z0-9_-]*\),\s*\1\s*\(#.*\)\?$' models/*.bnet

### 2022-05-11 cmsb22.tex
Add a line…
Integrated evaluation into main text

@@
 
-\begin{table}[!htb]
-  \caption{Timing comparisons between the proposed method and PyBoolNet on the selected real-world models.
-    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively.
-For each method, Column "\(|M|\)" denotes the number of minimal trap spaces and Column "time (s)" denotes the computation time in seconds.
-    "DNF" means that the method did not finish the computation within one hour.}
-  \centering
-  \label{tab:result_real}
-  \begin{tabular}{lrrrrrrr}
-    \toprule
-    & & & & \multicolumn{2}{c}{PyBoolNet} & \multicolumn{2}{c}{Proposed method}\\
-    \cmidrule(lr){5-6}\cmidrule(lr){7-8}
-    model & $n$ & $e$ & $s$ & $|M|$ & time (s) & $|M|$ & time (s) \\ \midrule
-    inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 287 & 0 & DNF & DNF & 1 & 296.23 \\
-    T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 193 & 7 & - & - & 318 & 0.24 \\
-    butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 139 & 13 & - & - & 512 & 0.17 \\
-    colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 153 & 1 & - & - & 10 & 0.18 \\
-    IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 149 & 15 & - & - & 512 & 0.32 \\
-    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 246 & 16 & - & - & 5724 & 0.90 \\
-    Korkut-2015~\cite{lee2019signal} & 99 & - & 12 & DNF & DNF & 2060 & 2.86 \\ \midrule
+To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet} on both its own repository of models and on large models from the literature.
 
-    adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 493 & 4 & - & - & 78 & 3.72 \\
-    interferon-1~\cite{ostaszewski2021covid19} & 121 & 190 & 55 & - & - & 512 & 0.43 \\
-    TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 273 & 2 & - & - & 48 & 0.50 \\
-    influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 302 & 11 & - & - & 2560 & 1.07 \\
-    signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 431 & 0 & - & - & 812 & 1.40 \\
-    HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 368 & 14 & DNF & DNF & 1408 & 1.44 \\
-    signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 557 & 9 & - & - & 375850 & 331.87 \\
-    HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 228 & 56 & - & - & 5120 & 0.99 \\
-    kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 304 & 72 & - & - & 512 & 3.79 \\
-    virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 268 & 25 & - & - & 1024 & 0.67 \\
-    immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 506 & 13 & DNF & DNF & 36852 & 14.19 \\
-    RA-apoptosis~\cite{aghamiri2020automated} & 181 & - & 59 & DNF & DNF & 640 & 0.70 \\
-    er-stress~\cite{ostaszewski2021covid19} & 182 & 266 & 75 & - & - & 25600 & 4.54 \\
-    cascade-3~\cite{Tsirvouli2020} & 183 & 603 & 0 & - & - & 1 & 1.65 \\
-    CHO-2016~\cite{lee2019signal} & 200 & 696 & 13 & - & - & 1024 & 2.10 \\ \midrule
-
-    T-cell-check-point~\cite{hernandez2020computational} & 218 & - & 34 & - & - & 15360 & 4.33 \\
-    ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 1100 & 22 & DNF & DNF & DNF & DNF \\
-    signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 533 & 19 & - & - & 512 & 1.58 \\
-    cholocystokinin~\cite{aghamiri2020automated} & 404 & - & 95 & - & - & 705024 & 1080.35 \\ \midrule
-
-    Alzheimer~\cite{aghamiri2020automated} & 1169 & - & 469 & - & - & 49152 & 66.17 \\
-    KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & - & 522 & DNF & DNF & 1024 & 1511.07 \\
-    human-network~\cite{kim2011reduction} & 1953 & - & 675 & DNF & DNF & 2048 & 1770.98 \\ \midrule
-
-    SN-5~\cite{kim2013rmod} & 2746 & - & 835 & DNF & DNF & 2048 & 3094.04 \\
-    turei-2016~\cite{lee2019signal} & 4691 & - & 1257 & DNF & DNF & DNF & DNF \\
-
-    \bottomrule
-  \end{tabular}
-
-\end{table}
-
-To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet}.
-We used a set of real-world Boolean models lying in various scales collected from various bibliographic sources.
-These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
-We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
-It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
 To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
 Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heu=domain --dom-mod=7} used by PyBoolNet).
-We ran all the benchmarks on a virtual machine whose environment is CPU: Intel® Core™ i7-3630QM 2.40GHz x 4, Memory: 8 GB, Ubuntu 18.04.2 64 bit.
-Finally, we set a time limit as one hour for each model.
+We ran all the benchmarks on an apple laptop whose environment is CPU: Intel\textregistered\ Core\texttrademark\ i7 1.20GHz x 4, 16 GB DDR4 RAM, MacOS 12.3.1.
+Finally, we set a time limit of two minutes for each model.
 
-Table~\ref{tab:result_real} shows the experimental results on real-world models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
-
-The first observation is that the proposed method much outperforms PyBoolNet in computational time. More specifically, ...
-% \vangiang{In the journal version, we can measure the memory consumption of the two methods}
-
-The second observation is that ...
-
-\section{Conclusion}
-
-In this article we proposed a new method for the computation of minimal trap spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
-This method is evaluated on large models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
-We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
-
-Though we benchmarked this new approach against the state-of-the-art tool PyBoolNet, there are many more evaluations that we plan to do in the future.
-First, the BioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
-Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
-The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it has to be evaluated in depth.
-
-Moreover, there are possibly other methods for the maximal siphon computation, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}.
-We also plan to investigate them in the future.
-They could replace our ASP program if they outperform it.
-
-However, the current method already appears to perform very well even on the biggest models we have considered.
-We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.
-
-% \subsubsection{Acknowledgments}
-
-\bibliographystyle{splncs04}
-\bibliography{cmsb22.bib}
-
-\appendix
-
-\section{PyBoolNet repo}
+\subsection{PyBoolNet repo}
 
@@
 
-%   \sylvain{I'm not sure that \(e\) and \(s\) are very relevant, moreover they are not readily available\dots}
-%   \vangiang{Indeed, they are not readily available.
-But if we use more meaningful words, the width of the table may exceed the width of the page.
-Thus, I use \(e\) and \(s\) along with the explanations shown in the table's caption.
-Do you have any other option?}
-%   \sylvain{just remove them since they are not relevant}
-
 As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}, the results are comparable with all minimal trap spaces found very fast.
@@
 
-\section{Selected models}
+\subsection{Selected models}
+
+We used a set of real-world Boolean models lying in various scales collected from various bibliographic sources.
+These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
+We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
+It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
+Table~\ref{tab:result_real} shows the experimental results on real-world models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
 
@@
   \centering
-  \label{tab:selected_redone}
+  \label{tab:result_real}
   \begin{tabular}{lrrrrrr}
@@
     \cmidrule(lr){6-7}
-    model & \(n\) & \(s\) & \(|M|\) & first 1000 & first 1000 & all \\
+    model & \(n\) & \(s\) & \(|M|\) & 1000 & 1000 & all \\
     \midrule
-    inflammatory-bowel-disease & 47 & 0 & 1 & DNF & 21.90 & 25.75\\
-    T-LGL-survival-network-2008 & 61 & 7 & 318 & 0.73 & 0.14 & 0.14\\
-    butanol-production & 66 & 13 & 8192 & 0.59 & 0.13 & 0.64\\
-    colitis-associated-colon-cancer & 70 & 1 & 10 & 0.16 & 0.08 & 0.06\\
-    mast-cell-activation & 73 & 19 & 1000+ & 0.64 & 0.15 & DNF\\
-    IL-6-signalling & 86 & 15 & 32768 & 3.30 & 0.10 & 5.06\\
-    Corral-ThIL-17-diff & 92 & 16 & 1000+ & DNF & 0.38 & DNF\\
-    Korkut-2015 & 99 & 12 & 18556 & DNF & 0.61 & 5.01\\
-    adhesion-cip-migration-cell-cycle & 121 & 4 & 78 & 40.81 & 0.33 & 0.45\\
-    interferon-1 & 121 & 55 & 1000+ & 8.36 & 0.19 & DNF\\
-    TCR-TLR5-signaling-2018 & 130 & 5 & 48 & 0.88 & 0.12 & 0.11\\
-    influenza-virus-replication-cycle & 131 & 11 & 10128 & 63.62 & 0.21 & 1.95\\
-    signaling-in-prostate-cancer & 133 & 11 & 2760 & DNF & 0.32 & 0.72\\
-    HIV-1-with-T-cell-signaling & 138 & 14 & 39424 & DNF & 0.87 & 7.72\\
-    signal-transduction-in-fibroblasts & 139 & 9 & 1000+ & DNF & 0.56 & DNF\\
-    HMOX-1-pathway & 145 & 56 & 1000+ & 5.16 & 0.18 & DNF\\
-    kynurenine-pathway & 150 & 72 & 1000+ & DNF & 0.74 & DNF\\
-    virus-replication-cycle & 154 & 25 & 1000+ & DNF & 0.19 & DNF\\
-    immune-system & 164 & 13 & 1000+ & DNF & 0.67 & DNF\\
-    RA-apoptosis & 180 & 59 & 1000+ & DNF & 0.16 & DNF\\
-    MAPK & 181 & 37 & 1000+ & 68.62 & 0.64 & DNF\\
-    er-stress & 182 & 75 & 1000+ & 28.65 & 0.21 & DNF\\
-    cascade-3 & 183 & 0 & 1 & 92.75 &  0.65 & 0.83\\
-    CHO-2016 & 200 & 13 & 13312 & DNF & 1.01 & 1.99\\
-    T-cell-check-point & 218 & 14 & 1000+ & 37.40 & 0.35 & DNF\\
-    ErbB-receptor-signaling & 247 & 22 & 1000+ & DNF & 1.07 & DNF\\
-    signaling-macrophage-activation & 321 & 19 & 1000+ & 15.82 & 0.63 & DNF\\
-    cholocystokinin & 383 & 74 & 1000+ & 1.22 & 0.73 & DNF\\
-    Alzheimer & 762 & 237 & 1000+ & DNF & 1.03 & DNF\\
-    KEGG-network & 1659 & 521 & 1000+ & DNF & 44.85 & DNF\\
-    human-network & 1953 & 669 & 1000+ & DNF & 42.97 & DNF\\
-    SN-5 & 2746 & 829 & ??? & DNF & DNF & DNF\\
-    turei-2016 & 4691 & 1257 & ??? & DNF & DNF & DNF\\
+    inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & 21.90 & 25.75\\
+    T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.73 & 0.14 & 0.14\\
+    butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.59 & 0.13 & 0.64\\
+    colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.16 & 0.08 & 0.06\\
+    mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & 1000+ & 0.64 & 0.15 & DNF\\
+    IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 3.30 & 0.10 & 5.06\\
+    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & 1000+ & DNF & 0.38 & DNF\\
+    Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 0.61 & 5.01\\ \midrule
+    adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 4 & 78 & 40.81 & 0.33 & 0.45\\
+    interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & 1000+ & 8.36 & 0.19 & DNF\\
+    TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 0.88 & 0.12 & 0.11\\
+    influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 63.62 & 0.21 & 1.95\\
+    signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & 0.32 & 0.72\\
+    HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & 0.87 & 7.72\\
+    signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 9 & 1000+ & DNF & 0.56 & DNF\\
+    HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & 1000+ & 5.16 & 0.18 & DNF\\
+    kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & 1000+ & DNF & 0.74 & DNF\\
+    virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & 1000+ & DNF & 0.19 & DNF\\
+    immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & 1000+ & DNF & 0.67 & DNF\\
+    RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & 1000+ & DNF & 0.16 & DNF\\
+    MAPK~\cite{aghamiri2020automated} & 181 & 37 & 1000+ & 68.62 & 0.64 & DNF\\
+    er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & 1000+ & 28.65 & 0.21 & DNF\\
+    cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 92.75 &  0.65 & 0.83\\
+    CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & 1.01 & 1.99\\
+    T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & 1000+ & 37.40 & 0.35 & DNF\\
+    ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & 1000+ & DNF & 1.07 & DNF\\
+    signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & 1000+ & 15.82 & 0.63 & DNF\\
+    cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & 1000+ & 1.22 & 0.73 & DNF\\
+    Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & 1000+ & DNF & 1.03 & DNF\\
+    KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & 1000+ & DNF & 44.85 & DNF\\
+    human-network~\cite{kim2011reduction} & 1953 & 669 & 1000+ & DNF & 42.97 & DNF\\
+    SN-5~\cite{kim2013rmod} & 2746 & 829 & ??? & DNF & DNF & DNF\\
+    turei-2016~\cite{lee2019signal} & 4691 & 1257 & ??? & DNF & DNF & DNF\\
     \bottomrule
@@
 \end{table}
-
 % s obtained by: grep -c '^\([a-zA-Z0-9_-]*\),\s*\1\s*\(#.*\)\?$' models/*.bnet
 
+Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-invariants.
+Hence, not even a single minimal trap-space was output by the method.
+
+The first observation is that the proposed method much outperforms PyBoolNet in computational time. More specifically, ...
+% \vangiang{In the journal version, we can measure the memory consumption of the two methods}
+
+The second observation is that ...
+
+\section{Conclusion}
+
+In this article we proposed a new method for the computation of minimal trap spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
+This method is evaluated on large models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
+We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
+
+Though we benchmarked this new approach against the state-of-the-art tool PyBoolNet, there are many more evaluations that we plan to do in the future.
+First, the BioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
+Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
+The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it has to be evaluated in depth.
+
+Moreover, there are possibly other methods for the maximal siphon computation, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}.
+We also plan to investigate them in the future.
+They could replace our ASP program if they outperform it.
+
+However, the current method already appears to perform very well even on the biggest models we have considered.
+We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.
+
+% \subsubsection{Acknowledgments}
+
+\bibliographystyle{splncs04}
+\bibliography{cmsb22.bib}
+
 \end{document}

### 2022-05-11 cmsb22.tex
Organization

@@
 To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
-Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heu=domain --dom-mod=7} used by PyBoolNet).
+Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by PyBoolNet).
 We ran all the benchmarks on an apple laptop whose environment is CPU: Intel\textregistered\ Core\texttrademark\ i7 1.20GHz x 4, 16 GB DDR4 RAM, MacOS 12.3.1.
 Finally, we set a time limit of two minutes for each model.
+
+All the models and a CoLoMoTo notebook realizing the benchmarks can be found at \url{https://github.com/soli/trap-spaces-as-siphons}.
+These can be run on a Docker image in the cloud by clicking the Binder button.
 

### 2022-05-11 cmsb22.tex
Tables

@@
   \caption{Timing comparisons between the proposed method and PyBoolNet on the PyBoolNet's repository.
-    Column \(n\) denotes the numbers of nodes of each model. Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-    ``DNF'' means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. A number in bold indicates a ratio greater than three.}
+    Column \(n\) denotes the numbers of nodes of each model, column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+    A number in bold indicates a ratio greater than three.}
   \centering
@@
     \toprule
-    model & $n$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
+    model & $n$ & $|M|$ & PyBoolNet & Trappist \\ \midrule
     arellano\_rootstem & 9 & 4 & 0.04 & \textbf{0.31}\\
@@
     irons\_yeast & 18 & 1 & 0.05 & 0.04\\
-    jaoude\_thdiff & 103 & 1000+ & \textbf{1.73} & 0.40\\
+    jaoude\_thdiff & 103 & >1000 & \textbf{1.73} & 0.40\\
     klamt\_tcr & 40 & 8 & 0.06 & 0.04\\
@@
     saadatpour\_guardcell & 13 & 1 & 0.02 & 0.03\\
-    selvaggio\_emt & 56 & 1000+ & \textbf{1.03} & 0.23\\
+    selvaggio\_emt & 56 & >1000 & \textbf{1.03} & 0.23\\
     tournier\_apoptosis & 12 & 3 & 0.05 & 0.06\\
@@
 
+\setlength{\tabcolsep}{4pt}
 \begin{table}[!htb]
@@
     colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.16 & 0.08 & 0.06\\
-    mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & 1000+ & 0.64 & 0.15 & DNF\\
+    mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & >1000 & 0.64 & 0.15 & DNF\\
     IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 3.30 & 0.10 & 5.06\\
-    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & 1000+ & DNF & 0.38 & DNF\\
+    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & >1000 & DNF & 0.38 & DNF\\
     Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 0.61 & 5.01\\ \midrule
     adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 4 & 78 & 40.81 & 0.33 & 0.45\\
-    interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & 1000+ & 8.36 & 0.19 & DNF\\
+    interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & >1000 & 8.36 & 0.19 & DNF\\
     TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 0.88 & 0.12 & 0.11\\
@@
     HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & 0.87 & 7.72\\
-    signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 9 & 1000+ & DNF & 0.56 & DNF\\
-    HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & 1000+ & 5.16 & 0.18 & DNF\\
-    kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & 1000+ & DNF & 0.74 & DNF\\
-    virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & 1000+ & DNF & 0.19 & DNF\\
-    immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & 1000+ & DNF & 0.67 & DNF\\
-    RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & 1000+ & DNF & 0.16 & DNF\\
-    MAPK~\cite{aghamiri2020automated} & 181 & 37 & 1000+ & 68.62 & 0.64 & DNF\\
-    er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & 1000+ & 28.65 & 0.21 & DNF\\
+    signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 9 & >1000 & DNF & 0.56 & DNF\\
+    HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & >1000 & 5.16 & 0.18 & DNF\\
+    kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & >1000 & DNF & 0.74 & DNF\\
+    virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & >1000 & DNF & 0.19 & DNF\\
+    immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & >1000 & DNF & 0.67 & DNF\\
+    RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & >1000 & DNF & 0.16 & DNF\\
+    MAPK~\cite{aghamiri2020automated} & 181 & 37 & >1000 & 68.62 & 0.64 & DNF\\
+    er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & >1000 & 28.65 & 0.21 & DNF\\
     cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 92.75 &  0.65 & 0.83\\
     CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & 1.01 & 1.99\\
-    T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & 1000+ & 37.40 & 0.35 & DNF\\
-    ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & 1000+ & DNF & 1.07 & DNF\\
-    signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & 1000+ & 15.82 & 0.63 & DNF\\
-    cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & 1000+ & 1.22 & 0.73 & DNF\\
-    Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & 1000+ & DNF & 1.03 & DNF\\
-    KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & 1000+ & DNF & 44.85 & DNF\\
-    human-network~\cite{kim2011reduction} & 1953 & 669 & 1000+ & DNF & 42.97 & DNF\\
+    T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & >1000 & 37.40 & 0.35 & DNF\\
+    ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & >1000 & DNF & 1.07 & DNF\\
+    signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & >1000 & 15.82 & 0.63 & DNF\\
+    cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & >1000 & 1.22 & 0.73 & DNF\\
+    Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & >1000 & DNF & 1.03 & DNF\\ \midrule
+    KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & >1000 & DNF & 44.85 & DNF\\
+    human-network~\cite{kim2011reduction} & 1953 & 669 & >1000 & DNF & 42.97 & DNF\\
     SN-5~\cite{kim2013rmod} & 2746 & 829 & ??? & DNF & DNF & DNF\\

### 2022-05-11 cmsb22.tex
Updates from Overleaf

--- 2022-05-11 cmsb22.tex
@@
 
-To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet} on both its own repository of models and on large models from the literature.
+To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet} on both its own repository of models and large models from the literature.
 
@@
 
-\subsection{PyBoolNet repo}
+\subsection{PyBoolNet repository}
 
@@
 Moreover, we believe that the result on the Arellano model is caused by the cold start of the JVM for BioLQM.
-On the contrary, on every model that was a bit challenging for ByPoolNet, the new method was far more efficient with speedups between one and two orders of magnitude.
+On the contrary, on every model that was a bit challenging for PyPoolNet, the new method was far more efficient with speedups between one and two orders of magnitude.
 

### 2022-05-11 cmsb22.tex
Small corrections

@@
 
-Using the method presented above, it is possible to convert the model to PNML in about 1s and to obtain the all 640 minimal trap spaces (including ones that contain more than one state) in a few milliseconds.
+Using the method presented above, it is possible to convert the model to PNML in about one second and to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) in a few milliseconds.
 Unfortunately since this was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.
@@
 Moreover, we believe that the result on the Arellano model is caused by the cold start of the JVM for BioLQM.
-On the contrary, on every model that was a bit challenging for PyPoolNet, the new method was far more efficient with speedups between one and two orders of magnitude.
+On the contrary, on every model that was a bit challenging for PyBoolNet, the new method was far more efficient with speedups between one and two orders of magnitude.
 
@@
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-    ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of 2 minutes.}
+    ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of two minutes.}
   \centering
@@
 
-Moreover, there are possibly other methods for the maximal siphon computation, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}.
-We also plan to investigate them in the future.
+In addition, there are possibly other methods for computing maximal conflict-free siphons in Petri nets, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}.
+Although these approaches do not directly support the maximal conflict-free siphon computation now, we plan to investigate them in the future.
 They could replace our ASP program if they outperform it.

### 2022-05-12 cmsb22.tex
Some updates

@@
 
-To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet} on both its own repository of models and large models from the literature.
+To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
 
-To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
+To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet}.
 Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by PyBoolNet).
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons between the proposed method and PyBoolNet on the PyBoolNet's repository.
-    Column \(n\) denotes the numbers of nodes of each model, column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+  \caption{Timing comparisons between the proposed method (i.e., Trappist) and PyBoolNet on the PyBoolNet's repository.
+    Column \(n\) denotes the number of nodes of each model, Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
     A number in bold indicates a ratio greater than three.}
@@
 Note however that on some very small models, our proposed method was sometimes slower than PyBoolNet, but still way under one second.
-Moreover, we believe that the result on the Arellano model is caused by the cold start of the JVM for BioLQM.
+Moreover, we believe that the result on the arellano\_rootstem model is caused by the cold start of the JVM for BioLQM.
 On the contrary, on every model that was a bit challenging for PyBoolNet, the new method was far more efficient with speedups between one and two orders of magnitude.
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
-    Columns \(n\) and \(s\) denote the numbers of nodes (resp.\ input nodes) of each model.
+  \caption{Timing comparisons between the proposed method (i.e., Trappist) and PyBoolNet on selected models from the literature.
+    Column \(n\) (resp.\ \(s\)) denotes the number of nodes (resp.\ source nodes) of each model.
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
@@
 
-Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-invariants.
+Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
 Hence, not even a single minimal trap-space was output by the method.
 
-The first observation is that the proposed method much outperforms PyBoolNet in computational time. More specifically, ...
-% \vangiang{In the journal version, we can measure the memory consumption of the two methods}
+%%% The proposed method outperforms PyBoolNet %%%
+The first observation is that the proposed method much outperforms PyBoolNet in computational time.
+More specifically, ...
 
+%%% It is not clear to show the advantage in ASP encoding %%%
 The second observation is that ...
+
+%%% For very huge models %%%
+The third observation is that for very huge models (), the proposed method did not manage to finish the Petri net conversion before the timeout, as marked by ``???'' in Table~\ref{tab:result_real}.
 
@@
 The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it has to be evaluated in depth.
+Second, with the experimental results shown in this paper, it is not enough information to distinctly expose the practical differences between our ASP encoding and that of PyBoolNet.
+Hence, we plan to conduct experiments on more real-world models or maybe random models that can be randomly generated by using BoolNet~\cite{mussel2010boolnet}.
 

### 2022-05-12 cmsb22.tex
Small things

@@
 
-  Boolean modelling of gene regulation but also post-transcri\-ptomic models has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
+  Boolean modelling of gene regulation but also of post-transcri\-ptomic systems has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
   This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
@@
 
-Petri nets were introduced in the 60s as a simple formalism for describing and analyzing information processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
+Petri nets were introduced in the 60s as simple formalism for describing and analyzing information-processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
 The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, was introduced quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
-Siphons are such a concept, but they have not been used a lot for the analysis of biochemical systems even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{nabli2016enumerating}.
+Siphons are such a concept, but they have not been used a lot for the study of biochemical systems even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{nabli2016enumerating}.
 
@@
 
-All models used for evaluation and the implementation of the presented method are available at\dots.
+All models used for evaluation and the implementation of the presented method are available at \url{https://github.com/soli/trap-spaces-as-siphons} and executable as a CoLoMoTo docker image.
 
@@
 \end{subfigure}%
-\caption{(a) Dynamics under the fully asynchronous update of the Boolean model. (b) Petri net encoding of the Boolean model. Circles denotes places, whereas rectangles denotes transitions.}
+\caption{(a) Dynamics under the fully asynchronous update of the Boolean model. (b) Petri net encoding of the Boolean model. Circles denote places, whereas rectangles denote transitions.}
 \label{fig:stg_and_PN}

### 2022-05-12 cmsb22.tex
More small things

@@
 
-  Boolean modelling of gene regulation but also of post-transcri\-ptomic systems has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
+  Boolean modelling of gene regulation but also of post-trans\-criptomic systems has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
   This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
@@
 \subsection{Traps spaces}
+
+\sylvain{This section might be better with real ``Definitions'', but when I tried rewriting it it led to a lot of changes\dots I guess we can do that for the journal version but for now keep things as they are.}
 
@@
 Figure~\ref{fig:stg_and_PN}(b) shows the Petri net encoding of this Boolean model.
-Place \(p_{x_1}\) (resp. \(\overline{p_{x_1}}\)) in \(\mathcal{P}\) represents the activation (resp. the inactivation) of node \(x_1\) in \(\mathcal{M}\).
+Place \(p_{x_1}\) (resp. \(\overline{p_{x_1}}\)) in \(\mathcal{P}\) represents the activation (resp.\ the inactivation) of node \(x_1\) in \(\mathcal{M}\).
 Marking \(\{p_{x_1}, \overline{p_{x_2}}\}\) in \(\mathcal{P}\) represents state 10 in \(\mathcal{M}\). 
 Transitions \(t^{1}_{x_1}\) and \(t^{2}_{x_1}\) represent the update of node \(x_1\).
-Of course, at any marking \(t^{1}_{x_1}\) and \(t^{2}_{x_1}\) cannot be both enabled.
-Then, the fully asynchronous update scheme in \(\mathcal{M}\) corresponds to the classical firing scheme in \(\mathcal{P}\) where only one of the enabled transitions at a marking will be fired~\cite{Murata1989}.
+Of course, in any marking \(t^{1}_{x_1}\) and \(t^{2}_{x_1}\) cannot be both enabled.
+Then, the fully asynchronous update scheme in \(\mathcal{M}\) corresponds to the classical firing scheme in \(\mathcal{P}\) where only one of the enabled transitions for a given marking will be fired~\cite{Murata1989}.
 
-Note that given a Logical model in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., one of the packages of SBML v3~\cite{keating2020sbml}, one can easily obtain its Petri net encoding in the PNML\footnote{\url{https://www.pnml.org/}} standard using the BioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
-This piece of software extracted from GINsim~\cite{chaouiya2012logical} and part of the CoLoMoTo\footnote{\url{http://colomoto.org/}} software suite allows for easy conversion between standard formats.
+Note that given a Logical model in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the BioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
+This piece of software extracted from GINsim~\cite{chaouiya2012logical} and part of the CoLoMoTo\footnote{\url{http://colomoto.org/}}~\cite{naldi2015cooperative} software suite allows for easy conversion between standard formats.
 It also accepts many other common formats for Logical models, notably the \verb|.bnet| files of the  BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} tools.
@@
 Naturally, a set-inclusion maximal answer set of \(\mathcal{L}\) is equivalent to a maximal conflict-free siphon of \(\mathcal{P}\), thus a minimal trap space of \(\mathcal{M}\).
-By using this built-in option, we can compute all the set-inclusion maximal answer sets of \(\mathcal{L}\) (resp.
-all the minimal trap spaces of \(\mathcal{M}\)) in one execution.
+By using this built-in option, we can compute all the set-inclusion maximal answer sets of \(\mathcal{L}\) (resp.\ all the minimal trap spaces of \(\mathcal{M}\)) in one execution.
 

### 2022-05-12 cmsb22.tex
Still going…

@@
 
-  A set of places of Petri net \(P\) encoding logical model \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{gene} of \(\mathcal{M}\).
+  A set of places of Petri net \(\mathcal{P}\) encoding logical model \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{gene} of \(\mathcal{M}\).
 Then, a conflict-free siphon \(S\) is said to be \emph{maximal} if and only if there is no other conflict-free siphon \(S'\) such that \(S \subset S'\).
@@
 
-For a few years now the second author has been collaborating with biologists who have been building very large detailed and annotated maps and now wish to analyze the dynamics of the corresponding models.
-One of the main maps studied this way is a map representing knowledge about the Rheumatoïd Arthritis~\cite{singh2018computational}, which was the main motivation for the development of a tool to automatically transform it into an executable Boolean model~\cite{aghamiri2020automated}.
-In the supplementary material of the paper, an excerpt of the map, focused around the apoptosis (cell death) module is transformed into a model of \emph{reasonable} size, namely 181 Boolean variables (model \verb|F5_RA_apoptosis_executable_module.sbml| of supplementary material S3, and model ``RA-apoptosis'' of Section~\ref{sec:eval}).
+For a few years now we have been collaborating with biologists who build very large detailed and annotated maps and now wish to analyze the dynamics of the corresponding models.
+One of the main maps studied this way represents knowledge about the Rheumatoïd Arthritis~\cite{singh2018computational}, and was the main motivation for the development of a tool to automatically transform it into an executable Boolean model~\cite{aghamiri2020automated}.
+In the supplementary material of the paper, an excerpt of the map, focused around the apoptosis (cell death) module is transformed into a model of \emph{reasonable} size, namely 180 Boolean variables (model \verb|F5_RA_apoptosis_executable_module.sbml| of supplementary material S3, and model ``RA-apoptosis'' of Section~\ref{sec:eval}).
 The study of such model, though, is a big hurdle.
@@
 The current practice usually revolves then around fixing some inputs to plausible values and reducing the model accordingly.
-While this approach makes sense, it relies on potentially arbitrary decisions, and \emph{hides away} critical modelling choices that were actually not part of the original Boolean model.
+While this approach makes sense, it relies on potentially arbitrary decisions, and \emph{hides away} critical modelling choices that were actually not part of the original Boolean model or even of the starting map.
 
@@
 
-To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
+To assess the efficiency of the proposed method, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
 
@@
 All the models and a CoLoMoTo notebook realizing the benchmarks can be found at \url{https://github.com/soli/trap-spaces-as-siphons}.
-These can be run on a Docker image in the cloud by clicking the Binder button.
+These can be run on a Docker image in the cloud by clicking the ``Binder'' button.
 

### 2022-05-12 cmsb22.tex
Some words in the conclusion mostly

@@
 \begin{table}[!htb]
-  \caption{Timing comparisons between the proposed method (i.e., Trappist) and PyBoolNet on the PyBoolNet's repository.
-    Column \(n\) denotes the number of nodes of each model, Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
+  \caption{Timing comparisons between the proposed method (i.e., Trappist) and PyBoolNet on the PyBoolNet repository.
+    Column \(n\) denotes the number of nodes of each model, Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
     A number in bold indicates a ratio greater than three.}
@@
 As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}, the results are comparable with all minimal trap spaces found very fast.
-Note however that on some very small models, our proposed method was sometimes slower than PyBoolNet, but still way under one second.
+Note however that on some very small models, our proposed method is sometimes slower than PyBoolNet, but still significantly under one second.
 Moreover, we believe that the result on the arellano\_rootstem model is caused by the cold start of the JVM for BioLQM.
-On the contrary, on every model that was a bit challenging for PyBoolNet, the new method was far more efficient with speedups between one and two orders of magnitude.
+On the contrary, on every model that was a bit challenging for PyBoolNet, the new method is far more efficient with speedups between one and two orders of magnitude.
 
@@
 It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
-Table~\ref{tab:result_real} shows the experimental results on real-world models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
+Table~\ref{tab:result_real} shows the experimental results on those models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
 
@@
 Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
-The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it has to be evaluated in depth.
-Second, with the experimental results shown in this paper, it is not enough information to distinctly expose the practical differences between our ASP encoding and that of PyBoolNet.
+The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it (in other words, the tradeoff of allowing redundant constraints) has to be evaluated in depth.
+
+Second, the experimental results shown in this paper mostly expose the differences caused by the prime-implicant phase of PyBoolNet.
+There is much more information required to distinctly study the practical differences between our ASP encoding and that of PyBoolNet, and notably their size/efficiency ratio.
 Hence, we plan to conduct experiments on more real-world models or maybe random models that can be randomly generated by using BoolNet~\cite{mussel2010boolnet}.
+Such experiments should include a time-course of the number of ASP solutions found for proper comparison of the APS encodings.
 
@@
 
-However, the current method already appears to perform very well even on the biggest models we have considered.
+However, the current method already appears to already perform very well even on the biggest models we have considered.
 We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.

### 2022-05-12 cmsb22.tex
Some observations

@@
 
+%%% The proposed method outperforms PyBoolNet %%%
+The first observation is that the proposed method vastly outperforms PyBoolNet in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100-1000 nodes size range).
 Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
 Hence, not even a single minimal trap-space was output by the method.
+The computational advantage is therefore immediately a practical advantage since on the one hand the state-of-the-art method did not allow any analysis whatsoever of the models, and on the other hand the proposed method could provide, very often under one second, the first thousand minimal trap spaces.
+For modellers having a critical look at a model and in a \emph{model, invalidate, refine} loop this means a huge difference in the models that are amenable to study.
 
-%%% The proposed method outperforms PyBoolNet %%%
-The first observation is that the proposed method much outperforms PyBoolNet in computational time.
-More specifically, ...
+Note that even with a very restricted time-limit of two minutes, it was possible with the proposed technique to find all minimal trap spaces of small models (roughly under 130 nodes, i.e., considered as quite big up to now).
+Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
 

### 2022-05-12 cmsb22.tex
Large models observation

@@
 %%% For very huge models %%%
-The third observation is that for very huge models (), the proposed method did not manage to finish the Petri net conversion before the timeout, as marked by ``???'' in Table~\ref{tab:result_real}.
+The third observation is that for very large models (i.e., more than two thousand nodes) the proposed method did not manage to finish the Petri net conversion before the timeout, as marked by ``???'' in Table~\ref{tab:result_real}.
+This points to the fact that our current choice of using a BDD-based translation to obtain that Petri net encoding, though it provides a small/efficient ASP might be too costly to handle the largest models.
+In such a case, a more \emph{naive} encoding might provide a much larger ASP program, with many redundant rules, but easier/faster to obtain.
+The evaluation of the feasibility of such strategy, and of its impact on smaller instances, remains to be done and is out of the scope of this first article.
 

### 2022-05-12 cmsb22.tex
Playing with Fig 1

@@
 \usetikzlibrary{positioning,automata} % drawing automa
-\usetikzlibrary{calc,fit,tikzmark}
+% \usetikzlibrary{calc,fit,tikzmark}
 
 % for captions in figures
-\usepackage{subcaption}
+\usepackage[labelformat=simple]{subcaption}
+\renewcommand\thesubfigure{(\alph{subfigure})}
 
@@
 For example, let us consider the Boolean model shown in Example~\ref{example:BN}.
-Figure~\ref{fig:stg_and_PN}(a) shows the dynamics of this model under the fully asynchronous update (i.e., only one node is nondeterministically selected in order to be updated at each time step).
+Figure~\ref{fig:stg} shows the dynamics of this model under the fully asynchronous update (i.e., only one node is nondeterministically selected in order to be updated at each time step).
 The model has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
@@
 \centering
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.3\textwidth}
 \centering
@@
 \end{tikzpicture}
-\caption{}
-\end{subfigure}%
+\caption{State transition graph, under the fully asynchronous update.}
+\label{fig:stg}
+\end{subfigure}\hfill%
 \begin{subfigure}[b]{0.6\textwidth}
@@
 \end{tikzpicture}
-\caption{}
+\caption{Petri net encoding of the model. Circles denote places, whereas rectangles denote transitions.}
+\label{fig:PN}
 \end{subfigure}%
-\caption{(a) Dynamics under the fully asynchronous update of the Boolean model. (b) Petri net encoding of the Boolean model. Circles denote places, whereas rectangles denote transitions.}
+\caption{Dynamics and encoding of the Boolean model of Example~\ref{example:BN}}
 \label{fig:stg_and_PN}
@@
 For illustration, let us reconsider the Boolean model shown in Example~\ref{example:BN}.
-Figure~\ref{fig:stg_and_PN}(b) shows the Petri net encoding of this Boolean model.
+Figure~\ref{fig:PN} shows the Petri net encoding of this Boolean model.
 Place \(p_{x_1}\) (resp. \(\overline{p_{x_1}}\)) in \(\mathcal{P}\) represents the activation (resp.\ the inactivation) of node \(x_1\) in \(\mathcal{M}\).
@@
 
-% For illustration, we consider the Petri net encoding \(\mathcal{P}\) shown in Figure~\ref{fig:stg_and_PN}(b).
+% For illustration, we consider the Petri net encoding \(\mathcal{P}\) shown in Figure~\ref{fig:PN}.
 % If \(p_{x_1}\) is in a siphon \(S\), then \(\overline{p_{x_1}}\) or \(\overline{p_{x_2}}\) should be in \(S\).

### 2022-05-12 cmsb22.tex
chktex

@@
 
-\sylvain{This section might be better with real ``Definitions'', but when I tried rewriting it it led to a lot of changes\dots I guess we can do that for the journal version but for now keep things as they are.}
+\sylvain{This section might be better with real ``Definitions'', but when I tried rewriting it, this led to a lot of changes\dots I guess we can do that for the journal version but for now keep things as they are.}
 
@@
 We give a Boolean model \(\mathcal{M} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
-Herein, \(\land\), \(\lor\), and \(\neg\) denote the CONJUNCTION, DISJUNCTION, and NEGATION logical operators, respectively.
-\label{example:BN}
+Herein, \(\land\), \(\lor\), and \(\neg\) denote the CONJUNCTION, DISJUNCTION, and NEGATION logical operators, respectively.\label{example:BN}
 \end{example}
@@
 \end{tikzpicture}
-\caption{State transition graph, under the fully asynchronous update.}
+\caption{State transition graph, under the fully asynchronous update.}%
 \label{fig:stg}
@@
 \end{tikzpicture}
-\caption{Petri net encoding of the model. Circles denote places, whereas rectangles denote transitions.}
+\caption{Petri net encoding of the model. Circles denote places, whereas rectangles denote transitions.}%
 \label{fig:PN}
 \end{subfigure}%
-\caption{Dynamics and encoding of the Boolean model of Example~\ref{example:BN}}
+\caption{Dynamics and encoding of the Boolean model of Example~\ref{example:BN}}%
 \label{fig:stg_and_PN}
@@
 
-% \vangiang{In this small figure, we can show the STG under the fully asynchronous update and the Petri net encoding of the Boolean model.}
-
-\subsection{Petri net encoding of Logical models}
+\subsection{Petri net encoding of Logical models}%
 \label{sec:encoding}
@@
 
-\begin{theorem}
+\begin{theorem}%
 \label{theo:ts_2_sp}
@@
 
-  From (*) and (**), we can conclude the proof.\qed
-
+  From (*) and (**), we can conclude the proof.\qed%
 \end{proof}
 
-\begin{theorem}
+\begin{theorem}%
 \label{theo:min_ts_2_max_sp}
@@
 
-  From (*) and (**), we can conclude the proof.\qed
-
+  From (*) and (**), we can conclude the proof.\qed%
 \end{proof}
@@
 
-\section{Evaluation}
+\section{Evaluation}%
 \label{sec:eval}
@@
 Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by PyBoolNet).
-We ran all the benchmarks on an apple laptop whose environment is CPU: Intel\textregistered\ Core\texttrademark\ i7 1.20GHz x 4, 16 GB DDR4 RAM, MacOS 12.3.1.
+We ran all the benchmarks on an apple laptop whose environment is CPU:\@ Intel\textregistered\ Core\texttrademark\ i7 1.20GHz x 4, 16 GB DDR4 RAM, MacOS 12.3.1.
 Finally, we set a time limit of two minutes for each model.
@@
     A number in bold indicates a ratio greater than three.}
-  \centering
+  \centering%
   \label{tab:pyboolnet_repo}
@@
 Note however that on some very small models, our proposed method is sometimes slower than PyBoolNet, but still significantly under one second.
-Moreover, we believe that the result on the arellano\_rootstem model is caused by the cold start of the JVM for BioLQM.
+Moreover, we believe that the result on the arellano\_rootstem model is caused by the cold start of the JVM for BioLQM.\@
 On the contrary, on every model that was a bit challenging for PyBoolNet, the new method is far more efficient with speedups between one and two orders of magnitude.
@@
     ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of two minutes.}
-  \centering
+  \centering%
   \label{tab:result_real}
@@
 %%% The proposed method outperforms PyBoolNet %%%
-The first observation is that the proposed method vastly outperforms PyBoolNet in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100-1000 nodes size range).
+The first observation is that the proposed method vastly outperforms PyBoolNet in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100--1000 nodes size range).
 Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
@@
 %%% It is not clear to show the advantage in ASP encoding %%%
-The second observation is that ...
+The second observation is that\dots
 
@@
 Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
-The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it (in other words, the tradeoff of allowing redundant constraints) has to be evaluated in depth.
+The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it (in other words, the trade-off of allowing redundant constraints) has to be evaluated in depth.
 
@@
 Hence, we plan to conduct experiments on more real-world models or maybe random models that can be randomly generated by using BoolNet~\cite{mussel2010boolnet}.
-Such experiments should include a time-course of the number of ASP solutions found for proper comparison of the APS encodings.
+Such experiments should include a time-course of the number of ASP solutions found for proper comparison of the ASP encodings.
 

### 2022-05-12 cmsb22.tex
lacheck
Change the order of authors

@@
 \author{Van-Giang Trinh\inst{1}\orcidID{0000-0001-6581-998X} \and \\
-  Sylvain Soliman\inst{2}\orcidID{0000-0001-5525-7418} \and \\
-  Kunihiko Hiraishi\inst{3}\orcidID{0000-0003-1750-1891} \and \\ Belaid Benhamou\inst{1}
+  Kunihiko Hiraishi\inst{3}\orcidID{0000-0003-1750-1891} \and \\
+  Belaid Benhamou\inst{1} \and \\
+  Sylvain Soliman\inst{2}\orcidID{0000-0001-5525-7418}
   }
@@
 
-\institute{
+\institute{%
   LIS, Aix-Marseille Université, Marseille, France\\
@@
 \sylvain{This section might be better with real ``Definitions'', but when I tried rewriting it, this led to a lot of changes\dots I guess we can do that for the journal version but for now keep things as they are.}
+\vangiang{Yes. This should be better. We can do that for the journal version.}
 

### 2022-05-12 cmsb22.tex
Finish the second observation

@@
 
-\sylvain{This section might be better with real ``Definitions'', but when I tried rewriting it, this led to a lot of changes\dots I guess we can do that for the journal version but for now keep things as they are.}
-\vangiang{Yes. This should be better. We can do that for the journal version.}
+% \sylvain{This section might be better with real ``Definitions'', but when I tried rewriting it, this led to a lot of changes\dots I guess we can do that for the journal version but for now keep things as they are.}
+% \vangiang{Yes. This should be better. We can do that for the journal version.}
 
@@
 Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
-Hence, not even a single minimal trap-space was output by the method.
+Hence, not even a single minimal trap space was output by the method.
 The computational advantage is therefore immediately a practical advantage since on the one hand the state-of-the-art method did not allow any analysis whatsoever of the models, and on the other hand the proposed method could provide, very often under one second, the first thousand minimal trap spaces.
@@
 %%% It is not clear to show the advantage in ASP encoding %%%
-The second observation is that\dots
+The second observation is that for all the models where PyBoolNet finished before the timeout, once PyBoolNet went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
+For these models, the ASP solving phase of the proposed method also took very short time, all under one second.
+Hence, with the experimental results shown in this paper, the practical differences between our ASP encoding and that of PyBoolNet are not distinctly exposed.
 

### 2022-05-13 cmsb22.tex
Some small wordings… Seems pretty good now

@@
 
-We used a set of real-world Boolean models lying in various scales collected from various bibliographic sources.
+We used a set of real-world Boolean models lying in various scales collected from numerous bibliographic sources.
 These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
@@
 Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
-Hence, not even a single minimal trap space was output by the method.
+Hence, not even a single minimal trap space was output by that method.
 The computational advantage is therefore immediately a practical advantage since on the one hand the state-of-the-art method did not allow any analysis whatsoever of the models, and on the other hand the proposed method could provide, very often under one second, the first thousand minimal trap spaces.
@@
 
-Note that even with a very restricted time-limit of two minutes, it was possible with the proposed technique to find all minimal trap spaces of small models (roughly under 130 nodes, i.e., considered as quite big up to now).
+Note that even with a very restricted time-limit of two minutes, it was possible with the proposed technique to find \emph{all} minimal trap spaces of small models (roughly under 130 nodes, i.e., considered as quite big up to now).
 Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
@@
 Hence, with the experimental results shown in this paper, the practical differences between our ASP encoding and that of PyBoolNet are not distinctly exposed.
+The fact that our new ASP encoding is guaranteed to be linear in the number of nodes of the original model does not seem to be crucial here, however a much deeper analysis of those cases remains to be done.
 

### 2022-05-13 cmsb22.tex
Update the order of authors again

@@
 \author{Van-Giang Trinh\inst{1}\orcidID{0000-0001-6581-998X} \and \\
-  Kunihiko Hiraishi\inst{3}\orcidID{0000-0003-1750-1891} \and \\
   Belaid Benhamou\inst{1} \and \\
-  Sylvain Soliman\inst{2}\orcidID{0000-0001-5525-7418}
+  Sylvain Soliman\inst{2}\orcidID{0000-0001-5525-7418} \and \\
+  Kunihiko Hiraishi\inst{3}\orcidID{0000-0003-1750-1891}
   }

### 2022-05-13 cmsb22.tex
Remove comments
Update the order authors ...

@@
 \titlerunning{Minimal trap spaces as maximal siphons}
-% \title{Scaling-up attractor computation for Boolean models}
-% \vangiang{I like the first title :).
-Since we only conduct experiments on real-world Boolean models of not too large size, it seems to make less sense to use "scaling-up" (this should be for the journal version).}
 
@@
   Belaid Benhamou\inst{1} \and \\
-  Sylvain Soliman\inst{2}\orcidID{0000-0001-5525-7418} \and \\
-  Kunihiko Hiraishi\inst{3}\orcidID{0000-0003-1750-1891}
+  Kunihiko Hiraishi\inst{2}\orcidID{0000-0003-1750-1891} \and \\
+  Sylvain Soliman\inst{3}\orcidID{0000-0001-5525-7418}
   }
@@
   \and
+  School of Information Science, Japan Advanced Institute of Science and Technology, Japan\\
+  \email{hira@jaist.ac.jp}
+  \and
   Lifeware team, Inria Saclay center, Palaiseau, France\\
   \email{Sylvain.Soliman@inria.fr}  
-  \and
-  School of Information Science, Japan Advanced Institute of Science and Technology, Japan\\
-  \email{hira@jaist.ac.jp}
 }
@@
 \subsection{Traps spaces}
-
-% \sylvain{This section might be better with real ``Definitions'', but when I tried rewriting it, this led to a lot of changes\dots I guess we can do that for the journal version but for now keep things as they are.}
-% \vangiang{Yes. This should be better. We can do that for the journal version.}
 
@@
 
-% The main property of this encoding is that it is completely faithful w.r.t.\ the asynchronous semantics.
-% Asynchronous transitions of the original model correspond one-to-one with firings of transitions in the Petri net.
-
 The main property of this encoding is that it is completely faithful with respect to the update scheme of the original Boolean model.
@@
 \label{theo:min_ts_2_max_sp}
-  %\(S\) is a minimal trap space of Boolean model \(\mathcal{M}\) if and only if the places corresponding to the fixed values in \(S\) (active or inactive) are a maximal conflict-free siphon of \((P, T, W)\), the Petri net encoding of \(S\).
 
@@
 Note that the number of atoms of \(\mathcal{L}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
-
-% For illustration, we consider the Petri net encoding \(\mathcal{P}\) shown in Figure~\ref{fig:PN}.
-% If \(p_{x_1}\) is in a siphon \(S\), then \(\overline{p_{x_1}}\) or \(\overline{p_{x_2}}\) should be in \(S\).
-% This rule is translated into the ASP rule
-% \[
-% \verb|p(x_1, 0) ; p(x_2, 0) :- p(x_1, 1).|
-% \] To forbid that \(S\) contains both \(p_{x_1}\) and \(\overline{p_{x_1}}\), we have the ASP rule
-% \[
-% \verb|:- p(x_1, 1) , p(x_2, 0).|
-% \] Analogously for other places in \(\mathcal{P}\), we obtain the final ASP as follows:
 

### 2022-05-13 cmsb22.tex
A small correction

@@
 A \emph{subspace} \(m\) of \(\mathcal{S}_{\mathcal{M}}\) is characterized by its fixed nodes (denoted by \(D_m\)) and free nodes.
-This subspace can be specified by an assignment \(D_m \mapsto \mathbb{B}\) where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
+This subspace can be specified by an assignment \(m : D_m \mapsto \mathbb{B}\) where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
 The remaining nodes of \(\mathcal{M}\), \(V \setminus D_m\), are said to be \emph{free}, i.e., they can receive any Boolean value.

### 2022-06-08 trappist/__init__.py
Preliminary .bnet parser and BDD builder

@@ -18,7 +18,7 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 # Don't forget to git tag v<version> && git push --tags
-version = "0.4.0"  # pylint: disable=invalid-name
+version = "0.5.0"  # pylint: disable=invalid-name
 
 
 from .trappist import compute_trap_spaces   # noqa

### 2022-06-08 trappist/bnet.py

@@ -1,0 +1,75 @@
+"""Compute minimal trap-spaces of a Petri-net encoded Boolean model.
+
+Copyright (C) 2022 Sylvain.Soliman@inria.fr
+
+This program is free software: you can redistribute it and/or modify
+it under the terms of the GNU General Public License as published by
+the Free Software Foundation, either version 3 of the License, or
+(at your option) any later version.
+
+This program is distributed in the hope that it will be useful,
+but WITHOUT ANY WARRANTY; without even the implied warranty of
+MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License
+along with this program.  If not, see <https://www.gnu.org/licenses/>.
+"""
+
+from typing import IO
+
+import networkx as nx  # TODO maybe replace with lists/dicts
+
+from pyeda.boolalg import boolfunc
+from pyeda.boolalg.bdd import bddvar, expr2bdd
+from pyeda.boolalg.expr import expr
+
+
+def add_edges(net: nx.DiGraph, tname: str, things: boolfunc, source: str):
+    """Add all edges from things to tname and back in net except for source."""
+    if source.startswith("-"):
+        nsource = source[1:]
+    else:
+        nsource = "-" + source
+
+    for i, t in enumerate(things.satisfy_all()):
+        name = f"{tname}_{i}"
+        net.add_node(name, kind="transition")
+        for p, v in t.items():
+            if v == 0:
+                pname = "-" + str(p)
+            else:
+                pname = str(p)
+            net.add_edge(pname, name)
+            if pname == source:  # don't add an edge back to "source" but add to Not(source)
+                net.add_edge(name, nsource)
+            else:
+                net.add_edge(name, pname)
+
+
+def read_bnet(fileobj: IO) -> nx.DiGraph:
+    """Parse a BoolNet .bnet file and build the corresponding Petri net."""
+    net = nx.DiGraph()
+
+    for line in fileobj.readlines():
+        if line.startswith("#") or line.startswith("targets, factors"):
+            continue
+        try:
+            x, fx = line.replace(" ", "").replace("!", "~").split(",", maxsplit=1)
+        except ValueError:  # not enough values to unpack
+            continue
+        net.add_node(x, kind="place")
+        net.add_node(
+            "-" + x, kind="place"
[…]

### 2022-06-08 trappist/trappist.py

@@ -23,11 +23,12 @@ import subprocess
 import sys
 import tempfile
 import xml.etree.ElementTree as etree
-from typing import Generator, IO, List, Optional
+from typing import Generator, IO, List
 
 import networkx as nx  # TODO maybe replace with lists/dicts
 
 from . import version
+from .bnet import read_bnet
 from .max_sat import get_sat_solutions
 
 
@@ -151,7 +152,22 @@ def compute_trap_spaces(
     method: str = "asp",
 ) -> Generator[List[str], None, None]:
     """Do the minimal trap-space computation on input file infile."""
-    petri_net = read_pnml(infile)
+    toclose = False
+    if isinstance(infile, str):
+        infile = open(infile, "r", encoding="utf-8")
+        toclose = True
+
+    if infile.name.endswith(".pnml"):
+        petri_net = read_pnml(infile)
+    elif infile.name.endswith(".bnet"):
+        petri_net = read_bnet(infile)
+    else:
+        infile.close()
+        raise ValueError("Currently limited to parsing PNML files")
+
+    if toclose:
+        infile.close()
+
     places = []
     for node, kind in petri_net.nodes(data="kind"):
         if kind == "place" and not node.startswith("-"):

### 2022-06-09 trappist/trappist.py
Just types…

@@ -134,7 +134,7 @@ def get_solutions(
 
 def get_asp_output(
     petri_net: nx.DiGraph, max_output: int, time_limit: int
-) -> List[str]:
+) -> str:
     """Generate and solve ASP file."""
     (_, tmpname) = tempfile.mkstemp(suffix=".lp", text=True)
     with open(tmpname, "wt") as asp_file:
@@ -176,8 +176,8 @@ def compute_trap_spaces(
         print(" ".join(places))
 
     if method == "asp":
-        solutions = get_asp_output(petri_net, max_output, time_limit)
-        solutions = get_solutions(solutions, places)
+        solutions_output = get_asp_output(petri_net, max_output, time_limit)
+        solutions = get_solutions(solutions_output, places)
     if method == "sat":
         solutions = get_sat_solutions(petri_net, max_output, time_limit, places)
 

### 2022-06-10 cmsb22.tex
Starting to update benchmarks in the paper

@@
 \renewcommand\UrlFont{\color{blue}\rmfamily}
+
+% table row numbers
+\newcounter{rownumber}
+\newcommand\rownb{\stepcounter{rownumber}\arabic{rownumber}}
 
@@
 
-To assess the efficiency of the proposed method, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
+To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
 
@@
 
+\setlength{\tabcolsep}{4pt}
 \begin{table}[!htb]
-  \caption{Timing comparisons between the proposed method (i.e., Trappist) and PyBoolNet on the PyBoolNet repository.
-    Column \(n\) denotes the number of nodes of each model, Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
-    A number in bold indicates a ratio greater than three.}
+  \caption{Timing comparisons between PyBoolNet, mpbn and Trappist on the PyBoolNet repository.
+    Column \(n\) denotes the number of nodes of each model.
+Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
+    A number in bold indicates a ratio greater than three compared to the best result. ``NM'' indicates a non-locally-monotonic model.}
   \centering%
   \label{tab:pyboolnet_repo}
-  \begin{tabular}{lrrrrrr}
+  \begin{tabular}{rlrrrrrrr}
     \toprule
-    model & $n$ & $|M|$ & PyBoolNet & Trappist \\ \midrule
-    arellano\_rootstem & 9 & 4 & 0.04 & \textbf{0.31}\\
-    calzone\_cellfate & 28 & 27 & 0.03 & 0.08\\
-    dahlhaus\_neuroplastoma & 23 & 32 & 0.04 & 0.11\\
-    davidich\_yeast & 10 & 12 & 0.03 & 0.08\\
-    dinwoodie\_life & 15 & 7 & 0.03 & 0.08\\
-    dinwoodie\_stomatal & 13 & 1 & 0.04 & 0.05\\
-    faure\_cellcycle & 10 & 2 & 0.04 & 0.04\\
-    grieco\_mapk & 53 & 18 & 0.04 & 0.08\\
-    irons\_yeast & 18 & 1 & 0.05 & 0.04\\
-    jaoude\_thdiff & 103 & >1000 & \textbf{1.73} & 0.40\\
-    klamt\_tcr & 40 & 8 & 0.06 & 0.04\\
-    krumsiek\_myeloid & 11 & 6 & 0.03 & 0.03\\
-    multivalued & 13 & 4 & 0.02 & 0.02\\
-    n12c5 & 11 & 5 & \textbf{34.37} & 0.06\\
-    n3s1c1a & 2 & 2 & 0.02 & 0.03\\
-    n3s1c1b & 2 & 2 & 0.02 & 0.03\\
-    n5s3 & 4 & 3 & 0.03 & 0.03\\
-    n6s1c2 & 5 & 3 & 0.03 & 0.04\\
-    n7s3 & 6 & 3 & 0.02 & 0.02\\
-    raf & 3 & 2 & 0.03 & 0.02\\
-    randomnet\_n15k3 & 15 & 3 & 0.04 & 0.05\\
-    randomnet\_n7k3 & 7 & 10 & 0.03 & \textbf{0.19}\\
-    remy\_tumorigenesis & 34 & 25 & \textbf{1.79} & 0.13\\
-    saadatpour\_guardcell & 13 & 1 & 0.02 & 0.03\\
-    selvaggio\_emt & 56 & >1000 & \textbf{1.03} & 0.23\\
-    tournier\_apoptosis & 12 & 3 & 0.05 & 0.06\\
-    xiao\_wnt5a & 7 & 4 & 0.06 & 0.03\\
-    zhang\_tlgl & 60 & 156 & 0.26 & 0.17\\
-    zhang\_tlgl\_v2 & 60 & 258 & 0.19 & 0.17\\
+    & model & $n$ & $|M|$ & PyBoolNet & mpbn & Trappist \\ \midrule
+    \rownb & arellano\_rootstem & 9 & 4 & 0.05 & 0.01 & \textbf{0.25}\\
+    \rownb & calzone\_cellfate & 28 & 27 & 0.03 & \textbf{NM} & 0.03\\
+    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & 0.09 & 0.02 & 0.03\\
+    \rownb & davidich\_yeast & 10 & 12 & 0.03 & 0.01 & 0.02\\
+    \rownb & dinwoodie\_life & 15 & 7 & 0.03 & 0.01 & 0.01\\
+    \rownb & dinwoodie\_stomatal & 13 & 1 & 0.03 & 0.01 & 0.01\\
+    \rownb & faure\_cellcycle & 10 & 2 & 0.04 & 0.01 & 0.01\\
+    \rownb & grieco\_mapk & 53 & 18 & 0.04 & 0.04 & 0.03\\
+    \rownb & irons\_yeast & 18 & 1 & 0.04 & 0.01 & 0.02\\
+    \rownb & jaoude\_thdiff & 103 & >1000 & \textbf{1.43} & \textbf{1.20} & 0.06\\
+    \rownb & klamt\_tcr & 40 & 8 & 0.04 & 0.02 & 0.01\\
+    \rownb & krumsiek\_myeloid & 11 & 6 & 0.03 & 0.01 & 0.01\\
+    \rownb & multivalued & 13 & 4 & 0.02 & 0.01 & 0.01\\
+    \rownb & n12c5 & 11 & 5 & \textbf{35.21} & 0.02 & 0.02\\
+    \rownb & n3s1c1a & 2 & 2 & 0.03 & 0.01 & 0.01\\
+    \rownb & n3s1c1b & 2 & 2 & 0.03 & 0.01 & 0.01\\
+    \rownb & n5s3 & 4 & 3 & 0.03 & \textbf{NM} & 0.01\\
+    \rownb & n6s1c2 & 5 & 3 & 0.04 & 0.01 & 0.01\\
+    \rownb & n7s3 & 6 & 3 & 0.03 & 0.01 & 0.01\\
+    \rownb & raf & 3 & 2 & 0.03 & 0.01 & 0.01\\
+    \rownb & randomnet\_n15k3 & 15 & 3 & 0.04 & \textbf{NM} & 0.01\\
+    \rownb & randomnet\_n7k3 & 7 & 10 & 0.03 & \textbf{NM} & 0.01\\
+    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{2.41} & 0.03 & 0.02\\
+    \rownb & saadatpour\_guardcell & 13 & 1 & 0.05 & 0.01 & 0.01\\
+    \rownb & selvaggio\_emt & 56 & >1000 & \textbf{1.75} & \textbf{0.64} & 0.04\\
+    \rownb & tournier\_apoptosis & 12 & 3 & 0.06 & 0.01 & 0.01\\
+    \rownb & xiao\_wnt5a & 7 & 4 & 0.04 & 0.01 & 0.01\\
+    \rownb & zhang\_tlgl & 60 & 156 & 0.33 & \textbf{NM} & 0.04\\
+    \rownb & zhang\_tlgl\_v2 & 60 & 258 & 0.12 & 0.18 & 0.03\\
     \bottomrule
@@
 \end{table}
+\setcounter{rownumber}{0}
 
@@
 
-\setlength{\tabcolsep}{4pt}
 \begin{table}[!htb]
-  \caption{Timing comparisons between the proposed method (i.e., Trappist) and PyBoolNet on selected models from the literature.
+  \caption{Timing comparisons between PyBoolNet (PBN), mpbn and Trappist on selected models from the literature.
     Column \(n\) (resp.\ \(s\)) denotes the number of nodes (resp.\ source nodes) of each model.
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-    ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of two minutes.}
+    ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of two minutes.
+``NM'' indicates a non-locally-monotonic model.}
   \centering%
   \label{tab:result_real}
-  \begin{tabular}{lrrrrrr}
+  \begin{tabular}{rlrrrrrrr}
     \toprule
-    & & & & PyBoolNet & \multicolumn{2}{c}{Trappist} \\
-    \cmidrule(lr){6-7}
-    model & \(n\) & \(s\) & \(|M|\) & 1000 & 1000 & all \\
+    & & & & & PBN & mpbn & \multicolumn{2}{c}{Trappist} \\
+    \cmidrule(lr){8-9}
+    & model & \(n\) & \(s\) & \(|M|\) & 1000 & 1000 & 1000 & all \\
     \midrule
-    inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & 21.90 & 25.75\\
-    T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.73 & 0.14 & 0.14\\
-    butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.59 & 0.13 & 0.64\\
-    colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.16 & 0.08 & 0.06\\
-    mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & >1000 & 0.64 & 0.15 & DNF\\
-    IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 3.30 & 0.10 & 5.06\\
-    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & >1000 & DNF & 0.38 & DNF\\
-    Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 0.61 & 5.01\\ \midrule
-    adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 4 & 78 & 40.81 & 0.33 & 0.45\\
-    interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & >1000 & 8.36 & 0.19 & DNF\\
-    TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 0.88 & 0.12 & 0.11\\
-    influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 63.62 & 0.21 & 1.95\\
-    signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & 0.32 & 0.72\\
-    HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & 0.87 & 7.72\\
-    signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 9 & >1000 & DNF & 0.56 & DNF\\
-    HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & >1000 & 5.16 & 0.18 & DNF\\
-    kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & >1000 & DNF & 0.74 & DNF\\
-    virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & >1000 & DNF & 0.19 & DNF\\
-    immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & >1000 & DNF & 0.67 & DNF\\
-    RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & >1000 & DNF & 0.16 & DNF\\
-    MAPK~\cite{aghamiri2020automated} & 181 & 37 & >1000 & 68.62 & 0.64 & DNF\\
-    er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & >1000 & 28.65 & 0.21 & DNF\\
-    cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 92.75 &  0.65 & 0.83\\
-    CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & 1.01 & 1.99\\
-    T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & >1000 & 37.40 & 0.35 & DNF\\
-    ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & >1000 & DNF & 1.07 & DNF\\
-    signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & >1000 & 15.82 & 0.63 & DNF\\
-    cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & >1000 & 1.22 & 0.73 & DNF\\
-    Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & >1000 & DNF & 1.03 & DNF\\ \midrule
-    KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & >1000 & DNF & 44.85 & DNF\\
-    human-network~\cite{kim2011reduction} & 1953 & 669 & >1000 & DNF & 42.97 & DNF\\
-    SN-5~\cite{kim2013rmod} & 2746 & 829 & ??? & DNF & DNF & DNF\\
-    turei-2016~\cite{lee2019signal} & 4691 & 1257 & ??? & DNF & DNF & DNF\\
+    \rownb & inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & & 21.90 & 25.75\\
+    \rownb & T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.73 & & 0.14 & 0.14\\
+    \rownb & butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.59 & & 0.13 & 0.64\\
+    \rownb & colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.16 & & 0.08 & 0.06\\
+    \rownb & mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & >1000 & 0.64 & & 0.15 & DNF\\
+    \rownb & IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 3.30 & & 0.10 & 5.06\\
+    \rownb & Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & >1000 & DNF & & 0.38 & DNF\\
+    \rownb & Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & & 0.61 & 5.01\\ \midrule
+    \rownb & adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 4 & 78 & 40.81 & & 0.33 & 0.45\\
+    \rownb & interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & >1000 & 8.36 & & 0.19 & DNF\\
+    \rownb & TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 0.88 & & 0.12 & 0.11\\
+    \rownb & influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 63.62 & & 0.21 & 1.95\\
+    \rownb & signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & & 0.32 & 0.72\\
+    \rownb & HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & & 0.87 & 7.72\\
+    \rownb & signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 9 & >1000 & DNF & & 0.56 & DNF\\
+    \rownb & HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & >1000 & 5.16 & & 0.18 & DNF\\
+    \rownb & kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & >1000 & DNF & & 0.74 & DNF\\
+    \rownb & virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & >1000 & DNF & & 0.19 & DNF\\
+    \rownb & immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & >1000 & DNF & & 0.67 & DNF\\
+    \rownb & RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & >1000 & DNF & & 0.16 & DNF\\
+    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & 37 & >1000 & 68.62 & & 0.64 & DNF\\
+    \rownb & er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & >1000 & 28.65 & & 0.21 & DNF\\
+    \rownb & cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 92.75 & &  0.65 & 0.83\\
+    \rownb & CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & & 1.01 & 1.99\\
+    \rownb & T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & >1000 & 37.40 & & 0.35 & DNF\\
+    \rownb & ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & >1000 & DNF & & 1.07 & DNF\\
+    \rownb & signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & >1000 & 15.82 & & 0.63 & DNF\\
+    \rownb & cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & >1000 & 1.22 & & 0.73 & DNF\\
+    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & >1000 & DNF & & 1.03 & DNF\\ \midrule
+    \rownb & KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & >1000 & DNF & & 44.85 & DNF\\
+    \rownb & human-network~\cite{kim2011reduction} & 1953 & 669 & >1000 & DNF & & 42.97 & DNF\\
+    \rownb & SN-5~\cite{kim2013rmod} & 2746 & 829 & ??? & DNF & & DNF & DNF\\
+    \rownb & turei-2016~\cite{lee2019signal} & 4691 & 1257 & ??? & DNF & & DNF & DNF\\
     \bottomrule
@@
 
-However, the current method already appears to already perform very well even on the biggest models we have considered.
+However, the current method already appears to perform very well even on the biggest models we have considered.
 We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.

### 2022-06-10 cmsb22.tex
biblio

@@
 Petri nets were introduced in the 60s as simple formalism for describing and analyzing information-processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
-The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, was introduced quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
-Siphons are such a concept, but they have not been used a lot for the study of biochemical systems even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{nabli2016enumerating}.
+The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, hinted at already in~\cite{peterson1981, Murata1989} was used more thoroughly quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
+Siphons are such a concept, but they have not been used a lot for the study of biochemical systems~\cite{zevedei2003topological,blatke2015biomodel} even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{oanea2010new,nabli2016enumerating}.
 
@@
 We give a Boolean model \(\mathcal{M} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
-Herein, \(\land\), \(\lor\), and \(\neg\) denote the CONJUNCTION, DISJUNCTION, and NEGATION logical operators, respectively.\label{example:BN}
+Herein, \(\land\), \(\lor\), and \(\neg\) denote the conjunction, disjunction, and negation logical operators, respectively.\label{example:BN}
 \end{example}
@@
 
-\subsection{Petri net encoding of Logical models}%
+\subsection{Petri net encoding of Boolean models}%
 \label{sec:encoding}
@@
 
-The link between Logical models \emph{à la} Thomas and Petri nets was originally established in~\cite{chaouiya2004qualitative} in order to make available formal methods like model-checking for the analysis of such systems.
-The basic encoding into 1-safe (i.e., never more than one token in each place) nets only holds for purely Boolean models but was later extended to multivalued models in two ways, either in~\cite{chaouiya2011petri} with non 1-safe Petri nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.
+The link between Boolean models \emph{à la} Thomas and Petri nets was originally established in~\cite{chaouiya2004qualitative} in order to make available formal methods like model-checking for the analysis of such systems.
+The basic encoding into 1-safe (i.e., never more than one token in each place) nets only holds for purely Boolean models but was later extended to multivalued Logical models in two ways, either in~\cite{chaouiya2011petri} with non 1-safe Petri nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.
 
@@
 
-Note that given a Logical model in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the BioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
+Note that given a Boolean model in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the BioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
 This piece of software extracted from GINsim~\cite{chaouiya2012logical} and part of the CoLoMoTo\footnote{\url{http://colomoto.org/}}~\cite{naldi2015cooperative} software suite allows for easy conversion between standard formats.
-It also accepts many other common formats for Logical models, notably the \verb|.bnet| files of the  BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} tools.
+It also accepts many other common formats for Boolean models, notably the \verb|.bnet| files of the  BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} tools.
 The conversion is executed as follows:
@@
 \section{Minimal trap spaces as maximal conflict-free siphons}
-First, we add a definition related to any set of places of a Petri net encoding a Logical model, and notably a siphon of such a net.
+First, we add a definition related to any set of places of a Petri net encoding a Boolean model, and notably a siphon of such a net.
 
@@
 
-  A set of places of Petri net \(\mathcal{P}\) encoding logical model \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{gene} of \(\mathcal{M}\).
+  A set of places of Petri net \(\mathcal{P}\) encoding Boolean model \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{gene} of \(\mathcal{M}\).
 Then, a conflict-free siphon \(S\) is said to be \emph{maximal} if and only if there is no other conflict-free siphon \(S'\) such that \(S \subset S'\).

### 2022-06-10 cmsb22.tex
Integrated benchs
We are much faster than before, I don't know why.
mpbn 1.6 only works on 6 models…

@@
 Petri nets were introduced in the 60s as simple formalism for describing and analyzing information-processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
-The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, hinted at already in~\cite{peterson1981, Murata1989} was used more thoroughly quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
+The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, hinted at already in~\cite{peterson1981petri,Murata1989} was used more thoroughly quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
 Siphons are such a concept, but they have not been used a lot for the study of biochemical systems~\cite{zevedei2003topological,blatke2015biomodel} even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{oanea2010new,nabli2016enumerating}.
@@
 We recall here the basic definition establishing that to produce something in a siphon you must consume something from the siphon.
-This corresponds to the idea that a siphon is a set of places that once empty remains empty.
+This corresponds to the idea that a siphon is a set of places that once unmarked remains unmarked.
 
@@
 
-Intuitively, a siphon is a set of places that once empty remains so.
-If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the empty places remain empty).
+Intuitively, a siphon is a set of places that once unmarked remains so.
+If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the unmarked places remain unmarked).
 This is precisely the definition of a trap space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap space.
@@
 \end{equation}
-The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net~\cite{nabli2016enumerating}.
+The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net and has been used with SAT solvers in~\cite{oanea2010new,nabli2016enumerating}.
 To make \(S\) to be a conflict-free siphon, we need to add to the system the rule
@@
 As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}, the results are comparable with all minimal trap spaces found very fast.
-Note however that on some very small models, our proposed method is sometimes slower than PyBoolNet, but still significantly under one second.
+For 5 of the 29 models, mpbn does not give any answer since these models are not locally-monotonic.
+Note that on some very small models, Trappist is sometimes slower than PyBoolNet and/or mpbn, but still significantly under one second.
 Moreover, we believe that the result on the arellano\_rootstem model is caused by the cold start of the JVM for BioLQM.\@
-On the contrary, on every model that was a bit challenging for PyBoolNet, the new method is far more efficient with speedups between one and two orders of magnitude.
+On the contrary, on every model that was a bit challenging for PyBoolNet or mpbn, the new method is far more efficient with speedups between one and two orders of magnitude.
 
@@
     \midrule
-    \rownb & inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & & 21.90 & 25.75\\
-    \rownb & T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.73 & & 0.14 & 0.14\\
-    \rownb & butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.59 & & 0.13 & 0.64\\
-    \rownb & colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.16 & & 0.08 & 0.06\\
-    \rownb & mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & >1000 & 0.64 & & 0.15 & DNF\\
-    \rownb & IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 3.30 & & 0.10 & 5.06\\
-    \rownb & Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & >1000 & DNF & & 0.38 & DNF\\
-    \rownb & Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & & 0.61 & 5.01\\ \midrule
-    \rownb & adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 4 & 78 & 40.81 & & 0.33 & 0.45\\
-    \rownb & interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & >1000 & 8.36 & & 0.19 & DNF\\
-    \rownb & TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 0.88 & & 0.12 & 0.11\\
-    \rownb & influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 63.62 & & 0.21 & 1.95\\
-    \rownb & signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & & 0.32 & 0.72\\
-    \rownb & HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & & 0.87 & 7.72\\
-    \rownb & signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 9 & >1000 & DNF & & 0.56 & DNF\\
-    \rownb & HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & >1000 & 5.16 & & 0.18 & DNF\\
-    \rownb & kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & >1000 & DNF & & 0.74 & DNF\\
-    \rownb & virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & >1000 & DNF & & 0.19 & DNF\\
-    \rownb & immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & >1000 & DNF & & 0.67 & DNF\\
-    \rownb & RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & >1000 & DNF & & 0.16 & DNF\\
-    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & 37 & >1000 & 68.62 & & 0.64 & DNF\\
-    \rownb & er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & >1000 & 28.65 & & 0.21 & DNF\\
-    \rownb & cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 92.75 & &  0.65 & 0.83\\
-    \rownb & CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & & 1.01 & 1.99\\
-    \rownb & T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & >1000 & 37.40 & & 0.35 & DNF\\
-    \rownb & ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & >1000 & DNF & & 1.07 & DNF\\
-    \rownb & signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & >1000 & 15.82 & & 0.63 & DNF\\
-    \rownb & cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & >1000 & 1.22 & & 0.73 & DNF\\
-    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & >1000 & DNF & & 1.03 & DNF\\ \midrule
-    \rownb & KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & >1000 & DNF & & 44.85 & DNF\\
-    \rownb & human-network~\cite{kim2011reduction} & 1953 & 669 & >1000 & DNF & & 42.97 & DNF\\
-    \rownb & SN-5~\cite{kim2013rmod} & 2746 & 829 & ??? & DNF & & DNF & DNF\\
-    \rownb & turei-2016~\cite{lee2019signal} & 4691 & 1257 & ??? & DNF & & DNF & DNF\\
+    \rownb & inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & NM & 0.82 & 0.69\\
+    \rownb & T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.84 & NM & 0.01 & 0.01\\
+    \rownb & butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.71 & NM & 0.02 & 0.02\\
+    \rownb & colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.20 & NM & 0.01 & 0.01\\
+    \rownb & mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & >1000 & 0.76 & NM & 0.01 & DNF\\
+    \rownb & IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 1.29 & NM & 0.01 & 0.01\\
+    \rownb & Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & >1000 & DNF & NM & 0.06 & DNF\\
+    \rownb & Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 1.50 & 0.07 & 0.08\\ \midrule
+    \rownb & adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 4 & 78 & 36.80 & 0.35 & 0.08 & 0.09\\
+    \rownb & interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & >1000 & 10.11 & NM & 0.02 & DNF\\
+    \rownb & TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 2.06 & NM & 0.03 & 0.02\\
+    \rownb & influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 46.67 & NM & 0.03 & 0.02\\
+    \rownb & signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & NM & 0.05 & 0.04\\
+    \rownb & HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & NM & 0.03 & 0.06\\
+    \rownb & signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 9 & >1000 & DNF & NM & 0.06 & DNF\\
+    \rownb & HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & >1000 & 6.52 & NM & 0.02 & DNF\\
+    \rownb & kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & >1000 & DNF & NM & 0.11 & DNF\\
+    \rownb & virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & >1000 & DNF & NM & 0.19 & DNF\\
+    \rownb & immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & >1000 & DNF & NM & 0.11 & DNF\\
+    \rownb & RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & >1000 & DNF & NM & 0.05 & DNF\\
+    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & 37 & >1000 & 87.13 & NM & 0.04 & DNF\\
+    \rownb & er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & >1000 & 17.88 & NM & 0.03 & DNF\\
+    \rownb & cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 101.18 & NM & 0.40 & 0.07\\
+    \rownb & CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & 3.19 & 0.06 & 0.08\\
+    \rownb & T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & >1000 & 69.94 & NM & 0.05 & DNF\\
+    \rownb & ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & >1000 & DNF & NM & 0.23 & DNF\\
+    \rownb & signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & >1000 & 16.43 & NM & 0.04 & DNF\\
+    \rownb & cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & >1000 & 1.42 & NM & 0.14 & DNF\\
+    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & >1000 & DNF & NM & 0.27 & DNF\\ \midrule
+    \rownb & KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & >1000 & DNF & 23.44 & 3.60 & DNF\\
+    \rownb & human-network~\cite{kim2011reduction} & 1953 & 669 & >1000 & DNF & 25.63 & 5.77 & DNF\\
+    \rownb & SN-5~\cite{kim2013rmod} & 2746 & 829 & >1000 & DNF & 33.57 & DNF & DNF\\
+    \rownb & turei-2016~\cite{lee2019signal} & 4691 & 1257 & ??? & DNF & DNF & DNF & DNF\\
     \bottomrule
@@
 
+\sylvain{Question of reviewer 3:
+Thanks to the performance improvement coming with Trappist, we now do know for sure that logical models may come with an overwhelming number of trap spaces, often exceeding 1000, which leaves the intriguing question how to evaluate these 1000 or more (even with Trappist so far not always computable) attractors/phenotypes?}
+
 % \subsubsection{Acknowledgments}

### 2022-06-13 cmsb22.tex
Remove 'already'
Re-add one already, we had two…
Adjust Table 2 to fit the page margins

@@
     \midrule
-    \rownb & inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & NM & 0.82 & 0.69\\
-    \rownb & T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.84 & NM & 0.01 & 0.01\\
+    \rownb & inflammatory-bowel~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & NM & 0.82 & 0.69\\
+    \rownb & T-LGL-survival~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.84 & NM & 0.01 & 0.01\\
     \rownb & butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.71 & NM & 0.02 & 0.02\\
-    \rownb & colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.20 & NM & 0.01 & 0.01\\
+    \rownb & colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.20 & NM & 0.01 & 0.01\\
     \rownb & mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & >1000 & 0.76 & NM & 0.01 & DNF\\
@@
     \rownb & Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 1.50 & 0.07 & 0.08\\ \midrule
-    \rownb & adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 4 & 78 & 36.80 & 0.35 & 0.08 & 0.09\\
+    \rownb & adhesion-cip-migration~\cite{guberman2020boolean} & 121 & 4 & 78 & 36.80 & 0.35 & 0.08 & 0.09\\
     \rownb & interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & >1000 & 10.11 & NM & 0.02 & DNF\\
-    \rownb & TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 2.06 & NM & 0.03 & 0.02\\
-    \rownb & influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 46.67 & NM & 0.03 & 0.02\\
-    \rownb & signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & NM & 0.05 & 0.04\\
-    \rownb & HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & NM & 0.03 & 0.06\\
-    \rownb & signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 9 & >1000 & DNF & NM & 0.06 & DNF\\
+    \rownb & TCR-TLR5-signaling~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 2.06 & NM & 0.03 & 0.02\\
+    \rownb & influenza-replication~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 46.67 & NM & 0.03 & 0.02\\
+    \rownb & prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & NM & 0.05 & 0.04\\
+    \rownb & HIV-1~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & NM & 0.03 & 0.06\\
+    \rownb & fibroblasts~\cite{helikar2008emergent} & 139 & 9 & >1000 & DNF & NM & 0.06 & DNF\\
     \rownb & HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & >1000 & 6.52 & NM & 0.02 & DNF\\
@@
     \rownb & ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & >1000 & DNF & NM & 0.23 & DNF\\
-    \rownb & signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & >1000 & 16.43 & NM & 0.04 & DNF\\
+    \rownb & macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & >1000 & 16.43 & NM & 0.04 & DNF\\
     \rownb & cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & >1000 & 1.42 & NM & 0.14 & DNF\\
@@
 
-However, the current method already appears to perform very well even on the biggest models we have considered.
+However, the current method appears to already perform very well even on the biggest models we have considered.
 We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.

### 2022-06-13 cmsb22.tex
Some answers

@@
 Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
+Since the ASP code generated is also quite declarative by nature, it is also possible to add to it supplementary constraints coming from the modeller in case one is looking for specific attractors.
 
@@
 The evaluation of the feasibility of such strategy, and of its impact on smaller instances, remains to be done and is out of the scope of this first article.
+
+Note that though enumerating the extremal siphons of a Petri net is exponential(see~\cite{nabli2016enumerating} for instance) this is apparently not the bottleneck of the proposed method, showing once again that networks obtained from biochemical models do have a specific structure.
 
@@
 
-\sylvain{Question of reviewer 3:
-Thanks to the performance improvement coming with Trappist, we now do know for sure that logical models may come with an overwhelming number of trap spaces, often exceeding 1000, which leaves the intriguing question how to evaluate these 1000 or more (even with Trappist so far not always computable) attractors/phenotypes?}
-
 % \subsubsection{Acknowledgments}

### 2022-06-13 cmsb22.tex
Some more answers

@@
 By Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean model to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
+Note that in the case of stable states, this can be put in regard to the classical relationship between siphons and deadlocks in Petri nets.
+It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap space that might be useful for the analysis of Petri nets, but this is out of the scope of this article.
+
 It is noted that there are no existing methods specifically designed for computing maximal conflict-free siphons (even maximal siphons) of a Petri net.

### 2022-06-13 cmsb22.tex
Highlight changes

--- 2022-06-13 cmsb22.tex
@@
 \newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
+
+\newcommand{\revise}[1]{\textcolor{blue}{#1}}
 
@@
 Petri nets were introduced in the 60s as simple formalism for describing and analyzing information-processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
-The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, hinted at already in~\cite{peterson1981petri,Murata1989} was used more thoroughly quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
+\revise{The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, hinted at already in~\cite{peterson1981petri,Murata1989} was used more thoroughly quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.}
 Siphons are such a concept, but they have not been used a lot for the study of biochemical systems~\cite{zevedei2003topological,blatke2015biomodel} even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{oanea2010new,nabli2016enumerating}.
@@
 We give a Boolean model \(\mathcal{M} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
-Herein, \(\land\), \(\lor\), and \(\neg\) denote the conjunction, disjunction, and negation logical operators, respectively.\label{example:BN}
+Herein, \(\land\), \(\lor\), and \(\neg\) denote the \revise{conjunction, disjunction, and negation} logical operators, respectively.\label{example:BN}
 \end{example}
@@
 We recall here the basic definition establishing that to produce something in a siphon you must consume something from the siphon.
-This corresponds to the idea that a siphon is a set of places that once unmarked remains unmarked.
+\revise{This corresponds to the idea that a siphon is a set of places that once unmarked remains unmarked.}
 
@@
 
-Intuitively, a siphon is a set of places that once unmarked remains so.
-If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the unmarked places remain unmarked).
+\revise{Intuitively, a siphon is a set of places that once unmarked remains so.
+If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the unmarked places remain unmarked).}
 This is precisely the definition of a trap space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap space.
@@
 By Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean model to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
+\revise{Note that in the case of stable states, this can be put in regard to the classical relationship between siphons and deadlocks in Petri nets.
+It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap space that might be useful for the analysis of Petri nets, but this is out of the scope of this article.
+}
+
 It is noted that there are no existing methods specifically designed for computing maximal conflict-free siphons (even maximal siphons) of a Petri net.
@@
 \end{equation}
-The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net and has been used with SAT solvers in~\cite{oanea2010new,nabli2016enumerating}.
+\revise{The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net and has been used with SAT solvers in~\cite{oanea2010new,nabli2016enumerating}.}
 To make \(S\) to be a conflict-free siphon, we need to add to the system the rule
@@
 
-To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
+\revise{To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.}
 
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons between PyBoolNet, mpbn and Trappist on the PyBoolNet repository.
+  \caption{\revise{Timing comparisons between PyBoolNet, mpbn and Trappist on the PyBoolNet repository.
     Column \(n\) denotes the number of nodes of each model.
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
-    A number in bold indicates a ratio greater than three compared to the best result. ``NM'' indicates a non-locally-monotonic model.}
+    A number in bold indicates a ratio greater than three compared to the best result. ``NM'' indicates a non-locally-monotonic model.}}
   \centering%
@@
 As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}, the results are comparable with all minimal trap spaces found very fast.
-For 5 of the 29 models, mpbn does not give any answer since these models are not locally-monotonic.
-Note that on some very small models, Trappist is sometimes slower than PyBoolNet and/or mpbn, but still significantly under one second.
+\revise{For 5 of the 29 models, mpbn does not give any answer since these models are not locally-monotonic.
+Note that on some very small models, Trappist is sometimes slower than PyBoolNet and/or mpbn, but still significantly under one second.}
 Moreover, we believe that the result on the arellano\_rootstem model is caused by the cold start of the JVM for BioLQM.\@
-On the contrary, on every model that was a bit challenging for PyBoolNet or mpbn, the new method is far more efficient with speedups between one and two orders of magnitude.
+\revise{On the contrary, on every model that was a bit challenging for PyBoolNet or mpbn, the new method is far more efficient with speedups between one and two orders of magnitude.}
 
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons between PyBoolNet (PBN), mpbn and Trappist on selected models from the literature.
+  \caption{\revise{Timing comparisons between PyBoolNet (PBN), mpbn and Trappist on selected models from the literature.
     Column \(n\) (resp.\ \(s\)) denotes the number of nodes (resp.\ source nodes) of each model.
@@
     ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of two minutes.
-``NM'' indicates a non-locally-monotonic model.}
+``NM'' indicates a non-locally-monotonic model.}}
   \centering%
@@
 Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
+\revise{Since the ASP code generated is also quite declarative by nature, it is also possible to add to it supplementary constraints coming from the modeller in case one is looking for specific attractors.}
 
@@
 The evaluation of the feasibility of such strategy, and of its impact on smaller instances, remains to be done and is out of the scope of this first article.
+
+\revise{Note that though enumerating the extremal siphons of a Petri net is exponential(see~\cite{nabli2016enumerating} for instance) this is apparently not the bottleneck of the proposed method, showing once again that networks obtained from biochemical models do have a specific structure.}
 
@@
 
-However, the current method appears to already perform very well even on the biggest models we have considered.
+\revise{However, the current method appears to already perform very well even on the biggest models we have considered.}
 We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.
-
-\sylvain{Question of reviewer 3:
-Thanks to the performance improvement coming with Trappist, we now do know for sure that logical models may come with an overwhelming number of trap spaces, often exceeding 1000, which leaves the intriguing question how to evaluate these 1000 or more (even with Trappist so far not always computable) attractors/phenotypes?}
 

### 2022-06-13 cmsb22.tex
Sampling from ASP solutions

@@
 Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
-\revise{Since the ASP code generated is also quite declarative by nature, it is also possible to add to it supplementary constraints coming from the modeller in case one is looking for specific attractors.}
+\revise{Since the ASP code generated is also quite declarative by nature, it is also possible to add to it supplementary constraints coming from the modeller in case one is looking for specific attractors.
+Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration}
 

### 2022-06-13 cmsb22.tex
Correct some small errors

@@
 \revise{Since the ASP code generated is also quite declarative by nature, it is also possible to add to it supplementary constraints coming from the modeller in case one is looking for specific attractors.
-Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration}
+Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration.}
 
@@
 
-\revise{Note that though enumerating the extremal siphons of a Petri net is exponential(see~\cite{nabli2016enumerating} for instance) this is apparently not the bottleneck of the proposed method, showing once again that networks obtained from biochemical models do have a specific structure.}
+\revise{Note that though enumerating the extremal siphons of a Petri net is exponential (see~\cite{nabli2016enumerating} for instance) this is apparently not the bottleneck of the proposed method, showing once again that networks obtained from biochemical models do have a specific structure.}
 

### 2022-06-13 cmsb22.tex
Add descriptions for using mpbn

@@
 However, using a state-of-the-art tool like PyBoolNet~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
+\revise{Note that mpbn does not give any answer because that model is not locally-monotonic.}
 And hence, it is not possible to extract any (complex) attractor at all.
@@
 
-\revise{To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.}
+\revise{To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
+In addition, we also include the tool mpbn~\cite{Paulev2020}\footnote{We use the latest version of mpbn available at submission time through CoLoMoTo.} to the benchmarks although it only handle locally-monotonic models, whereas both Trappist and PyBoolNet can handle general models.}
 
@@
 These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
-We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
+\revise{We then applied PyBoolNet, mpbn, and Trappist} to computing minimal trap spaces of these real-world models.
 It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.

### 2022-06-14 cmsb22.tex
Observations

@@
 As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}, the results are comparable with all minimal trap spaces found very fast.
-\revise{For 5 of the 29 models, mpbn does not give any answer since these models are not locally-monotonic.
+\revise{For 5 of the 29 models, mpbn did not give any answer because it recognized that these models are not locally-monotonic.
 Note that on some very small models, Trappist is sometimes slower than PyBoolNet and/or mpbn, but still significantly under one second.}
@@
 
+%%% The proposed method and mpbn %%%
+\revise{The first observation is that for 27 of the 33 models (more than 78\%), mpbn did not give any answer because it recognized that these models are not locally-monotonic.
+For 6 of the 33 models where mpbn returned the answers, mpbn and Trappist are comparable in computation time.
+Along with the fact that mbpn can handle only locally-monotonic models and Trappist can handle general models, it is less insightful to further compare between them.
+Hence, we focus on only comparisons between PyBoolNet and Trappist in the following observations.}
+
 %%% The proposed method outperforms PyBoolNet %%%
-The first observation is that the proposed method vastly outperforms PyBoolNet in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100--1000 nodes size range).
+\revise{The second observation} is that the proposed method vastly outperforms PyBoolNet in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100--1000 nodes size range).
 Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
@@
 %%% It is not clear to show the advantage in ASP encoding %%%
-The second observation is that for all the models where PyBoolNet finished before the timeout, once PyBoolNet went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
+\revise{The third observation} is that for all the models where PyBoolNet finished before the timeout, once PyBoolNet went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
 For these models, the ASP solving phase of the proposed method also took very short time, all under one second.
@@
 %%% For very huge models %%%
-The third observation is that for very large models (i.e., more than two thousand nodes) the proposed method did not manage to finish the Petri net conversion before the timeout, as marked by ``???'' in Table~\ref{tab:result_real}.
+\revise{The last observation} is that for very large models (i.e., more than two thousand nodes) the proposed method did not manage to finish the Petri net conversion before the timeout, as marked by ``???'' in Table~\ref{tab:result_real}.
 This points to the fact that our current choice of using a BDD-based translation to obtain that Petri net encoding, though it provides a small/efficient ASP might be too costly to handle the largest models.

### 2022-06-14 cmsb22.tex
Stuff related to mpbn

@@
 However, with the continuing increase in model-size, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of implicants.
+
+\revise{It is worth noting that there has been a recent method~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces, which avoids the prime-implicants computation.
+This method has been implemented in the tool mpbn\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
+However, this method is only applicable for \emph{locally-monotonic} Boolean models, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for general Boolean models (i.e..
+including both locally-monotonic and non-locally-monotonic Boolean models).
+The study~\cite{DBLP:journals/tcs/NoualRS13} highlights the need of non-locally-monotonic Boolean models in both biological and theoretical aspects.
+Hence, it is still necessary to develop efficient methods for computing minimal trap spaces of large-scale general Boolean models.
+}
 
@@
 %%% The proposed method and mpbn %%%
-\revise{The first observation is that for 27 of the 33 models (more than 78\%), mpbn did not give any answer because it recognized that these models are not locally-monotonic.
+\revise{The first observation is that for 26 of the 33 models (more than 78\%), mpbn did not give any answer because it recognized that these models are not locally-monotonic.
 For 6 of the 33 models where mpbn returned the answers, mpbn and Trappist are comparable in computation time.
-Along with the fact that mbpn can handle only locally-monotonic models and Trappist can handle general models, it is less insightful to further compare between them.
+Along with the fact that mbpn can handle only locally-monotonic models and Trappist can handle general models, it is less meaningful to further compare between them.
 Hence, we focus on only comparisons between PyBoolNet and Trappist in the following observations.}

### 2022-06-14 cmsb22.tex
ASP encoding

@@
 This method has been implemented in the tool mpbn\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
-However, this method is only applicable for \emph{locally-monotonic} Boolean models, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for general Boolean models (i.e..
-including both locally-monotonic and non-locally-monotonic Boolean models).
-The study~\cite{DBLP:journals/tcs/NoualRS13} highlights the need of non-locally-monotonic Boolean models in both biological and theoretical aspects.
+However, this method is only applicable for \emph{locally-monotonic} Boolean models, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean models (i.e., including both locally-monotonic and non-locally-monotonic Boolean models).
+The study~\cite{DBLP:journals/tcs/NoualRS13} highlights the need for non-locally-monotonic Boolean models in both biological and theoretical aspects.
 Hence, it is still necessary to develop efficient methods for computing minimal trap spaces of large-scale general Boolean models.
@@
 
-In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, of a Boolean model.
+In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, of a general Boolean model.
 It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
@@
 Note that the number of atoms of \(\mathcal{L}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
+\revise{In~\cite{DBLP:conf/ictai/ChevalierFPZ19}, there is an ASP characterization of trap spaces that does not rely on minimal DNFs and thus seems very similar to the our ASP encoding.
+However, this encoding is only applicable for locally-monotonic Boolean models.
+Hence it might be theoretically insightful to compare in terms of complexity between it and our ASP encoding only for locally-monotonic BNs, but this is out of the scope of this article.
+}
 
@@
 However, using a state-of-the-art tool like PyBoolNet~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
-\revise{Note that mpbn does not give any answer because that model is not locally-monotonic.}
+\revise{Note that mpbn~\cite{Paulev2020} does not give any answer because it recognized that model non-locally-monotonic.}
 And hence, it is not possible to extract any (complex) attractor at all.

### 2022-06-14 cmsb22.tex
Mostly small changes

@@
 
-\revise{It is worth noting that there has been a recent method~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces, which avoids the prime-implicants computation.
+\revise{It is worth noting that the recent method presented in~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces avoids the prime-implicants computation by relying on the \emph{most-permissive} semantics of Boolean models.
 This method has been implemented in the tool mpbn\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
@@
 Note that the number of atoms of \(\mathcal{L}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
-\revise{In~\cite{DBLP:conf/ictai/ChevalierFPZ19}, there is an ASP characterization of trap spaces that does not rely on minimal DNFs and thus seems very similar to the our ASP encoding.
-However, this encoding is only applicable for locally-monotonic Boolean models.
-Hence it might be theoretically insightful to compare in terms of complexity between it and our ASP encoding only for locally-monotonic BNs, but this is out of the scope of this article.
+\revise{In~\cite{DBLP:conf/ictai/ChevalierFPZ19}, there is an ASP characterization of trap spaces that does not rely on minimal DNFs either and thus seems very similar to our ASP encoding.
+Remarkably it only requires the DNF for the \emph{activation} part, using the information that it will only be used for locally-monotonic Boolean models.
+We would therefore expect that, when available, it will have comparable performance on the ASP part (the ASP program would be twice smaller, but redundancy is not always bad in that field), but can also avoid combinatorial explosion of the Petri net encoding for some formula where the activation DNF is simple but the inhibition is not.
+Since mpbn is included in our benchmark this will be evaluated in our experiments.
 }
@@
 However, using a state-of-the-art tool like PyBoolNet~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
-\revise{Note that mpbn~\cite{Paulev2020} does not give any answer because it recognized that model non-locally-monotonic.}
+\revise{mpbn~\cite{Paulev2020} does not give any answer either because it recognizes that model as non-locally-monotonic.}
 And hence, it is not possible to extract any (complex) attractor at all.
@@
 \revise{To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
-In addition, we also include the tool mpbn~\cite{Paulev2020}\footnote{We use the latest version of mpbn available at submission time through CoLoMoTo.} to the benchmarks although it only handle locally-monotonic models, whereas both Trappist and PyBoolNet can handle general models.}
+In addition, we also include the tool mpbn~\cite{Paulev2020} to the benchmarks although it only handle locally-monotonic models, whereas both Trappist and PyBoolNet can handle general models.
+}
 
@@
 
+\revise{For all the above tools, namely PyBoolNet, mpbn and CLINGO, we used the version available in the latest CoLoMoTo docker image tagged \texttt{2022-05-01}.}
 All the models and a CoLoMoTo notebook realizing the benchmarks can be found at \url{https://github.com/soli/trap-spaces-as-siphons}.
@@
 As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}, the results are comparable with all minimal trap spaces found very fast.
-\revise{For 5 of the 29 models, mpbn did not give any answer because it recognized that these models are not locally-monotonic.
+\revise{For 5 of the 29 models, mpbn did not give any answer because it recognized these models as not locally-monotonic.
 Note that on some very small models, Trappist is sometimes slower than PyBoolNet and/or mpbn, but still significantly under one second.}
@@
 %%% The proposed method and mpbn %%%
-\revise{The first observation is that for 26 of the 33 models (more than 78\%), mpbn did not give any answer because it recognized that these models are not locally-monotonic.
-For 6 of the 33 models where mpbn returned the answers, mpbn and Trappist are comparable in computation time.
-Along with the fact that mbpn can handle only locally-monotonic models and Trappist can handle general models, it is less meaningful to further compare between them.
+\revise{The first observation is that for 26 of the 33 models (more than 78\%), mpbn did not give any answer because it recognized that these models as not locally-monotonic.
+For 6 of the 33 models where mpbn returned the answers, mpbn and Trappist are comparable in computation time, though surprisingly a bit slower on average.
+Note however that mpbn was the only tool to provide a solution for the SN-5 model, thus confirming that if the activation function is in the right form, not having to compute the inactivation function's disjunctive normal form can render a difficult problem tractable.
+However, since mbpn can handle only locally-monotonic models and Trappist can handle general models, it is difficult to further compare between them.
 Hence, we focus on only comparisons between PyBoolNet and Trappist in the following observations.}
@@
 \end{document}
+
+We believe that an improved test for local-monotonicity will soon be available through CoLoMoTo and intend to redo our comparison then.

### 2022-06-15 cmsb22.tex
Still adding answers

@@
 Though this might appear quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
-In practice this transformation, here using BDDs in BioLQM, seems to behave quite well on even quite big models as will be shown in the Section~\ref{sec:eval} on evaluation.
+\revise{One way to look at this is to consider that this amounts to a similar approach as that used in~\cite{DBLP:conf/ictai/ChevalierFPZ19} but with the encoding of both activation and inhibition functions as DNFs in order to take into account a possible non-local-monotonicity.
+This does not change the worst-case-complexity (obtaining a single DNF being exponential) but might matter a lot in practice.
+As such, we will explore how this transformation, here using BDDs in BioLQM, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.}
 

### 2022-06-15 cmsb22.tex
Conclusion
v_i(t+1)

@@
 Given a Boolean model \(\mathcal{M} = (V, F)\) with nodes \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\mathbb{B}^{n}\) with \(\mathbb{B} = \{0, 1\}\).
-A state \(s \in \mathbb{B}^{n}\) is a mapping \(s : V \mapsto \mathbb{B}\) that assigns either 0 (inactive) or 1 (active) to each node.
+A state \(s \in \mathbb{B}^{n}\) is a mapping \(s \colon V \mapsto \mathbb{B}\) that assigns either 0 (inactive) or 1 (active) to each node.
 At each time step \(t\), node \(v_i\) can update its state by
-\[v_i(t + 1) = f_i(v(t)),\] where \(v(t)\) is the state of \(\mathcal{M}\) at time \(t\) and \(v_i(t + 1)\) the state of node \(v_i\) at time \(t + 1\).
-An update scheme of a Boolean model specifies the way that the nodes of the model update their states through time evolution~\cite{thomas1991regulatory}.
-There are two main types of update schemes: synchronous and asynchronous.
+\revise{\[v_i(t + 1) \in \{v_{i}(t), f_i(v(t))\}\]
+where \(v(t)\) is the state of \(\mathcal{M}\) at time \(t\), \(v_i(t + 1)\) the state of node \(v_i\) at time \(t + 1\) and an update scheme specifies which value is chosen~\cite{thomas1991regulatory}.
+There are two main types of update schemes: synchronous, where for all \(i, v_{i}(t+1) = f_{i}(v(t))\), and asynchronous, where an \(i\) is chosen among the possible updates (\(f_{i}(v(t))\neq v_{i}(t)\)), and all the other nodes \(j\) remain fixed at \(v_{j}(t)\).}
 
@@
 
+\revise{Finally, we think that the links between Petri nets and Boolean models that we stumbled upon in this method might have deeper roots.
+Exploring those connections might lead both to interesting topics of research for Petri nets, like a notion of trap-spaces, and for Boolean models.}
+
 % \subsubsection{Acknowledgments}

### 2022-06-15 cmsb22.tex
Some small corrections

@@
 This method has been implemented in the tool mpbn\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
-However, this method is only applicable for \emph{locally-monotonic} Boolean models, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean models (i.e., including both locally-monotonic and non-locally-monotonic Boolean models).
+However, this method is only applicable for \emph{locally-monotonic} Boolean models, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean models (i.e., including both locally-monotonic and non-locally-monotonic ones).
 The study~\cite{DBLP:journals/tcs/NoualRS13} highlights the need for non-locally-monotonic Boolean models in both biological and theoretical aspects.
@@
 \revise{\[v_i(t + 1) \in \{v_{i}(t), f_i(v(t))\}\]
-where \(v(t)\) is the state of \(\mathcal{M}\) at time \(t\), \(v_i(t + 1)\) the state of node \(v_i\) at time \(t + 1\) and an update scheme specifies which value is chosen~\cite{thomas1991regulatory}.
+where \(v(t)\) is the state of \(\mathcal{M}\) at time \(t\), \(v_i(t + 1)\) is the state of node \(v_i\) at time \(t + 1\) and an update scheme specifies which value is chosen~\cite{thomas1991regulatory}.
 There are two main types of update schemes: synchronous, where for all \(i, v_{i}(t+1) = f_{i}(v(t))\), and asynchronous, where an \(i\) is chosen among the possible updates (\(f_{i}(v(t))\neq v_{i}(t)\)), and all the other nodes \(j\) remain fixed at \(v_{j}(t)\).}
@@
 Though this might appear quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
-\revise{One way to look at this is to consider that this amounts to a similar approach as that used in~\cite{DBLP:conf/ictai/ChevalierFPZ19} but with the encoding of both activation and inhibition functions as DNFs in order to take into account a possible non-local-monotonicity.
+\revise{One way to look at this is to consider that this amounts to a similar approach as that used in~\cite{DBLP:conf/ictai/ChevalierFPZ19} but with the encoding of both activation and inhibition functions as DNFs in order to take into account possible non-local-monotonicity.
 This does not change the worst-case-complexity (obtaining a single DNF being exponential) but might matter a lot in practice.

### 2022-06-15 cmsb22.tex
Updates from Overleaf

@@
 Petri nets were introduced in the 60s as simple formalism for describing and analyzing information-processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
-\revise{The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, hinted at already in~\cite{peterson1981petri,Murata1989} was used more thoroughly quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.}
-Siphons are such a concept, but they have not been used a lot for the study of biochemical systems~\cite{zevedei2003topological,blatke2015biomodel} even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{oanea2010new,nabli2016enumerating}.
+\revise{The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, hinted at already in~\cite{peterson1981petri,Murata1989} was used more thoroughly quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
+Siphons are such a concept, but they have not been used a lot for the study of biochemical systems~\cite{zevedei2003topological,blatke2015biomodel} even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{oanea2010new,nabli2016enumerating}.}
 

### 2022-06-16 cmsb22.tex
PDF and useless sentence
Final touches

@@
 Remarkably it only requires the DNF for the \emph{activation} part, using the information that it will only be used for locally-monotonic Boolean models.
-We would therefore expect that, when available, it will have comparable performance on the ASP part (the ASP program would be twice smaller, but redundancy is not always bad in that field), but can also avoid combinatorial explosion of the Petri net encoding for some formula where the activation DNF is simple but the inhibition is not.
+We would therefore expect that, when available, it will have comparable performance on the ASP part (the ASP program would be approximately twice smaller, though redundancy is not always bad in that field), but can also avoid combinatorial explosion of the Petri net encoding for some formula where the activation DNF is simple but the inhibition is not.
 Since mpbn is included in our benchmark this will be evaluated in our experiments.
@@
 \revise{To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
-In addition, we also include the tool mpbn~\cite{Paulev2020} to the benchmarks although it only handle locally-monotonic models, whereas both Trappist and PyBoolNet can handle general models.
+In addition, we also include the tool mpbn~\cite{Paulev2020} to the benchmarks although it only handles locally-monotonic models, whereas both Trappist and PyBoolNet can handle general models.
 }
@@
 \revise{The first observation is that for 26 of the 33 models (more than 78\%), mpbn did not give any answer because it recognized that these models as not locally-monotonic.
-For 6 of the 33 models where mpbn returned the answers, mpbn and Trappist are comparable in computation time, though surprisingly a bit slower on average.
+For 6 of the 33 models where mpbn returned the answers, mpbn and Trappist are comparable in computation time, though surprisingly mpbn appears a bit slower on average.
 Note however that mpbn was the only tool to provide a solution for the SN-5 model, thus confirming that if the activation function is in the right form, not having to compute the inactivation function's disjunctive normal form can render a difficult problem tractable.
@@
 The evaluation of the feasibility of such strategy, and of its impact on smaller instances, remains to be done and is out of the scope of this first article.
+\revise{Recognizing that a model is locally-monotonic and applying in that specific case dedicated strategies as those of mpbn might also be a partial solution as shown in the SN-5 case, but not for the turei-2016 model.}
 
@@
 \bibliographystyle{splncs04}
-\bibliography{cmsb22.bib}
+\bibliography{cmsb22}
 
 \end{document}
-
-We believe that an improved test for local-monotonicity will soon be available through CoLoMoTo and intend to redo our comparison then.

### 2022-06-16 cmsb22.tex
Updated Table 1
Updated Table 2

@@
     & model & $n$ & $|M|$ & PyBoolNet & mpbn & Trappist \\ \midrule
-    \rownb & arellano\_rootstem & 9 & 4 & 0.05 & 0.01 & \textbf{0.25}\\
+    \rownb & arellano\_rootstem & 9 & 4 & 0.05 & 0.01 & \textbf{0.20}\\
     \rownb & calzone\_cellfate & 28 & 27 & 0.03 & \textbf{NM} & 0.03\\
-    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & 0.09 & 0.02 & 0.03\\
-    \rownb & davidich\_yeast & 10 & 12 & 0.03 & 0.01 & 0.02\\
+    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & 0.05 & 0.02 & 0.03\\
+    \rownb & davidich\_yeast & 10 & 12 & 0.04 & 0.01 & 0.02\\
     \rownb & dinwoodie\_life & 15 & 7 & 0.03 & 0.01 & 0.01\\
     \rownb & dinwoodie\_stomatal & 13 & 1 & 0.03 & 0.01 & 0.01\\
-    \rownb & faure\_cellcycle & 10 & 2 & 0.04 & 0.01 & 0.01\\
-    \rownb & grieco\_mapk & 53 & 18 & 0.04 & 0.04 & 0.03\\
-    \rownb & irons\_yeast & 18 & 1 & 0.04 & 0.01 & 0.02\\
-    \rownb & jaoude\_thdiff & 103 & >1000 & \textbf{1.43} & \textbf{1.20} & 0.06\\
-    \rownb & klamt\_tcr & 40 & 8 & 0.04 & 0.02 & 0.01\\
+    \rownb & faure\_cellcycle & 10 & 2 & 0.04 & 0.01 & 0.02\\
+    \rownb & grieco\_mapk & 53 & 18 & 0.04 & 0.02 & 0.02\\
+    \rownb & irons\_yeast & 18 & 1 & 0.05 & 0.01 & 0.02\\
+    \rownb & jaoude\_thdiff & 103 & >1000 & \textbf{1.44} & \textbf{0.90} & 0.09\\
+    \rownb & klamt\_tcr & 40 & 8 & 0.04 & 0.01 & 0.03\\
     \rownb & krumsiek\_myeloid & 11 & 6 & 0.03 & 0.01 & 0.01\\
-    \rownb & multivalued & 13 & 4 & 0.02 & 0.01 & 0.01\\
-    \rownb & n12c5 & 11 & 5 & \textbf{35.21} & 0.02 & 0.02\\
-    \rownb & n3s1c1a & 2 & 2 & 0.03 & 0.01 & 0.01\\
-    \rownb & n3s1c1b & 2 & 2 & 0.03 & 0.01 & 0.01\\
+    \rownb & multivalued & 13 & 4 & 0.03 & 0.01 & 0.01\\
+    \rownb & n12c5 & 11 & 5 & \textbf{35.16} & 0.01 & 0.02\\
+    \rownb & n3s1c1a & 2 & 2 & 0.02 & 0.01 & 0.01\\
+    \rownb & n3s1c1b & 2 & 2 & 0.02 & 0.01 & 0.01\\
     \rownb & n5s3 & 4 & 3 & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & n6s1c2 & 5 & 3 & 0.04 & 0.01 & 0.01\\
-    \rownb & n7s3 & 6 & 3 & 0.03 & 0.01 & 0.01\\
-    \rownb & raf & 3 & 2 & 0.03 & 0.01 & 0.01\\
-    \rownb & randomnet\_n15k3 & 15 & 3 & 0.04 & \textbf{NM} & 0.01\\
+    \rownb & n6s1c2 & 5 & 3 & 0.03 & 0.01 & 0.01\\
+    \rownb & n7s3 & 6 & 3 & 0.02 & 0.01 & 0.01\\
+    \rownb & raf & 3 & 2 & 0.02 & 0.01 & 0.01\\
+    \rownb & randomnet\_n15k3 & 15 & 3 & 0.03 & \textbf{NM} & 0.01\\
     \rownb & randomnet\_n7k3 & 7 & 10 & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{2.41} & 0.03 & 0.02\\
-    \rownb & saadatpour\_guardcell & 13 & 1 & 0.05 & 0.01 & 0.01\\
-    \rownb & selvaggio\_emt & 56 & >1000 & \textbf{1.75} & \textbf{0.64} & 0.04\\
-    \rownb & tournier\_apoptosis & 12 & 3 & 0.06 & 0.01 & 0.01\\
-    \rownb & xiao\_wnt5a & 7 & 4 & 0.04 & 0.01 & 0.01\\
-    \rownb & zhang\_tlgl & 60 & 156 & 0.33 & \textbf{NM} & 0.04\\
-    \rownb & zhang\_tlgl\_v2 & 60 & 258 & 0.12 & 0.18 & 0.03\\
+    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{2.14} & 0.02 & 0.02\\
+    \rownb & saadatpour\_guardcell & 13 & 1 & 0.03 & 0.01 & 0.01\\
+    \rownb & selvaggio\_emt & 56 & >1000 & \textbf{1.02} & \textbf{0.52} & 0.09\\
+    \rownb & tournier\_apoptosis & 12 & 3 & 0.04 & 0.01 & 0.01\\
+    \rownb & xiao\_wnt5a & 7 & 4 & 0.03 & 0.01 & 0.01\\
+    \rownb & zhang\_tlgl & 60 & 156 & 0.22 & \textbf{NM} & 0.05\\
+    \rownb & zhang\_tlgl\_v2 & 60 & 258 & 0.09 & 0.15 & 0.02\\
     \bottomrule
@@
     \midrule
-    \rownb & inflammatory-bowel~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & NM & 0.82 & 0.69\\
+    \rownb & inflammatory-bowel~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & NM & 1.26 & 0.69\\
     \rownb & T-LGL-survival~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.84 & NM & 0.01 & 0.01\\
-    \rownb & butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.71 & NM & 0.02 & 0.02\\
-    \rownb & colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.20 & NM & 0.01 & 0.01\\
+    \rownb & butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.71 & NM & 0.01 & 0.02\\
+    \rownb & colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.20 & NM & 0.02 & 0.01\\
     \rownb & mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & >1000 & 0.76 & NM & 0.01 & DNF\\
     \rownb & IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 1.29 & NM & 0.01 & 0.01\\
-    \rownb & Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & >1000 & DNF & NM & 0.06 & DNF\\
-    \rownb & Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 1.50 & 0.07 & 0.08\\ \midrule
-    \rownb & adhesion-cip-migration~\cite{guberman2020boolean} & 121 & 4 & 78 & 36.80 & 0.35 & 0.08 & 0.09\\
+    \rownb & Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & >1000 & DNF & NM & 0.03 & DNF\\
+    \rownb & Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 1.04 & 0.06 & 0.08\\ \midrule
+    \rownb & adhesion-cip-migration~\cite{guberman2020boolean} & 121 & 4 & 78 & 36.80 & 0.35 & 0.27 & 0.09\\
     \rownb & interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & >1000 & 10.11 & NM & 0.02 & DNF\\
     \rownb & TCR-TLR5-signaling~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 2.06 & NM & 0.03 & 0.02\\
-    \rownb & influenza-replication~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 46.67 & NM & 0.03 & 0.02\\
-    \rownb & prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & NM & 0.05 & 0.04\\
-    \rownb & HIV-1~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & NM & 0.03 & 0.06\\
-    \rownb & fibroblasts~\cite{helikar2008emergent} & 139 & 9 & >1000 & DNF & NM & 0.06 & DNF\\
+    \rownb & influenza-replication~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 46.67 & NM & 0.02 & 0.02\\
+    \rownb & prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & NM & 0.16 & 0.04\\
+    \rownb & HIV-1~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & NM & 0.06 & 0.06\\
+    \rownb & fibroblasts~\cite{helikar2008emergent} & 139 & 9 & >1000 & DNF & NM & 0.08 & DNF\\
     \rownb & HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & >1000 & 6.52 & NM & 0.02 & DNF\\
-    \rownb & kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & >1000 & DNF & NM & 0.11 & DNF\\
-    \rownb & virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & >1000 & DNF & NM & 0.19 & DNF\\
-    \rownb & immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & >1000 & DNF & NM & 0.11 & DNF\\
-    \rownb & RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & >1000 & DNF & NM & 0.05 & DNF\\
-    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & 37 & >1000 & 87.13 & NM & 0.04 & DNF\\
+    \rownb & kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & >1000 & DNF & NM & 0.19 & DNF\\
+    \rownb & virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & >1000 & DNF & NM & 0.03 & DNF\\
+    \rownb & immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & >1000 & DNF & NM & 0.08 & DNF\\
+    \rownb & RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & >1000 & DNF & NM & 0.02 & DNF\\
+    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & 37 & >1000 & 87.13 & NM & 0.03 & DNF\\
     \rownb & er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & >1000 & 17.88 & NM & 0.03 & DNF\\
     \rownb & cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 101.18 & NM & 0.40 & 0.07\\
-    \rownb & CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & 3.19 & 0.06 & 0.08\\
+    \rownb & CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & 2.45 & 0.07 & 0.08\\
     \rownb & T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & >1000 & 69.94 & NM & 0.05 & DNF\\
-    \rownb & ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & >1000 & DNF & NM & 0.23 & DNF\\
-    \rownb & macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & >1000 & 16.43 & NM & 0.04 & DNF\\
-    \rownb & cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & >1000 & 1.42 & NM & 0.14 & DNF\\
-    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & >1000 & DNF & NM & 0.27 & DNF\\ \midrule
-    \rownb & KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & >1000 & DNF & 23.44 & 3.60 & DNF\\
-    \rownb & human-network~\cite{kim2011reduction} & 1953 & 669 & >1000 & DNF & 25.63 & 5.77 & DNF\\
-    \rownb & SN-5~\cite{kim2013rmod} & 2746 & 829 & >1000 & DNF & 33.57 & DNF & DNF\\
+    \rownb & ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & >1000 & DNF & NM & 0.24 & DNF\\
+    \rownb & macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & >1000 & 16.43 & NM & 0.05 & DNF\\
+    \rownb & cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & >1000 & 1.42 & NM & 0.08 & DNF\\
+    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & >1000 & DNF & NM & 0.29 & DNF\\ \midrule
+    \rownb & KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & >1000 & DNF & 20.76 & 3.40 & DNF\\
+    \rownb & human-network~\cite{kim2011reduction} & 1953 & 669 & >1000 & DNF & 23.62 & 5.93 & DNF\\
+    \rownb & SN-5~\cite{kim2013rmod} & 2746 & 829 & >1000 & DNF & 28.94 & DNF & DNF\\
     \rownb & turei-2016~\cite{lee2019signal} & 4691 & 1257 & ??? & DNF & DNF & DNF & DNF\\

### 2022-06-16 cmsb22.tex
Comments on the benchmark conditions and back =15p

@@
 Finally, we set a time limit of two minutes for each model.
+\revise{Note that we did not get the opportunity to run this benchmark on a proper dedicated computing workstation, the results are therefore only indicative of global trends and should not be interpreted as precise performance.
+This is also why in some rare cases finding all minimal trap spaces was faster than with the limit to 1000.}
 
@@
 Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
-\revise{Since the ASP code generated is also quite declarative by nature, it is also possible to add to it supplementary constraints coming from the modeller in case one is looking for specific attractors.
+\revise{Since the ASP code is declarative by nature, it is also possible to add to it supplementary constraints coming from the modeler in case one is looking for specific attractors.
 Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration.}
@@
 
-Though we benchmarked this new approach against the state-of-the-art tool PyBoolNet, there are many more evaluations that we plan to do in the future.
+Though we benchmarked this new approach against state-of-the-art \revise{tools}, there are many more evaluations that we plan to do in the future.
 First, the BioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
@@
 They could replace our ASP program if they outperform it.
-
 \revise{However, the current method appears to already perform very well even on the biggest models we have considered.}
-We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.
+% We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.
 

### 2022-06-27 cmsb22.tex
Cleaned TeX file

--- 2022-06-16 cmsb22.tex
@@
 
-\usetikzlibrary{positioning,automata} % drawing automa
-% \usetikzlibrary{calc,fit,tikzmark}
+% drawing automa
+\usetikzlibrary{positioning,automata}
 
@@
 \newcommand\rownb{\stepcounter{rownumber}\arabic{rownumber}}
-
-\newcommand{\vangiang}[1]{\textcolor{magenta}{#1}}
-\newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
-
-\newcommand{\revise}[1]{\textcolor{blue}{#1}}
 
@@
 
-\revise{It is worth noting that the recent method presented in~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces avoids the prime-implicants computation by relying on the \emph{most-permissive} semantics of Boolean models.
+It is worth noting that the recent method presented in~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces avoids the prime-implicants computation by relying on the \emph{most-permissive semantics of Boolean models.
 This method has been implemented in the tool mpbn\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
@@
 Petri nets were introduced in the 60s as simple formalism for describing and analyzing information-processing systems that are characterized as being concurrent, asynchronous, non-deterministic and possibly distributed~\cite{peterson1981petri,Murata1989}.
-\revise{The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, hinted at already in~\cite{peterson1981petri,Murata1989} was used more thoroughly quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
-Siphons are such a concept, but they have not been used a lot for the study of biochemical systems~\cite{zevedei2003topological,blatke2015biomodel} even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{oanea2010new,nabli2016enumerating}.}
+The use of Petri nets for representing biochemical reaction systems, by mapping molecular species to places and reactions to transitions, hinted at already in~\cite{peterson1981petri,Murata1989} was used more thoroughly quite late in~\cite{reddy1993petri}, together with some Petri net concepts and tools for the analysis of metabolic networks.
+Siphons are such a concept, but they have not been used a lot for the study of biochemical systems~\cite{zevedei2003topological,blatke2015biomodel} even if the practical cost of computing their minimal/maximal elements appear much more manageable than the theoretical complexity would indicate~\cite{oanea2010new,nabli2016enumerating}.
 
@@
 
-%%% Boolean model %%%
 Given a Boolean model \(\mathcal{M} = (V, F)\) with nodes \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\mathbb{B}^{n}\) with \(\mathbb{B} = \{0, 1\}\).
@@
 At each time step \(t\), node \(v_i\) can update its state by
-\revise{\[v_i(t + 1) \in \{v_{i}(t), f_i(v(t))\}\]
+\[v_i(t + 1) \in \{v_{i}(t), f_i(v(t))\}\]
 where \(v(t)\) is the state of \(\mathcal{M}\) at time \(t\), \(v_i(t + 1)\) is the state of node \(v_i\) at time \(t + 1\) and an update scheme specifies which value is chosen~\cite{thomas1991regulatory}.
-There are two main types of update schemes: synchronous, where for all \(i, v_{i}(t+1) = f_{i}(v(t))\), and asynchronous, where an \(i\) is chosen among the possible updates (\(f_{i}(v(t))\neq v_{i}(t)\)), and all the other nodes \(j\) remain fixed at \(v_{j}(t)\).}
+There are two main types of update schemes: synchronous, where for all \(i, v_{i}(t+1) = f_{i}(v(t))\), and asynchronous, where an \(i\) is chosen among the possible updates (\(f_{i}(v(t))\neq v_{i}(t)\)), and all the other nodes \(j\) remain fixed at \(v_{j}(t)\).
 
-%%% Trap set %%%
 A non-empty set \(T \subseteq S_{\mathcal{M}}\) is a \emph{trap set} of \(\mathcal{M}\) if and only if for every \(x \in T\) and \(y \in S_{\mathcal{M}}\) with \(y\) is reachable from \(x\) it holds that \(y \in T\).
 
-%%% Subspace %%%
 A \emph{subspace} \(m\) of \(\mathcal{S}_{\mathcal{M}}\) is characterized by its fixed nodes (denoted by \(D_m\)) and free nodes.
@@
 
-%%% Minimal trap space and example %%%
 A \emph{trap space} is defined as a subspace that is also a trap set.
@@
 
-%%% Example %%%
 For example, let us consider the Boolean model shown in Example~\ref{example:BN}.
@@
 We give a Boolean model \(\mathcal{M} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
-Herein, \(\land\), \(\lor\), and \(\neg\) denote the \revise{conjunction, disjunction, and negation} logical operators, respectively.\label{example:BN}
+Herein, \(\land\), \(\lor\), and \(\neg\) denote the conjunction, disjunction, and negation logical operators, respectively.\label{example:BN}
 \end{example}
@@
 Though this might appear quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
-\revise{One way to look at this is to consider that this amounts to a similar approach as that used in~\cite{DBLP:conf/ictai/ChevalierFPZ19} but with the encoding of both activation and inhibition functions as DNFs in order to take into account possible non-local-monotonicity.
+One way to look at this is to consider that this amounts to a similar approach as that used in~\cite{DBLP:conf/ictai/ChevalierFPZ19} but with the encoding of both activation and inhibition functions as DNFs in order to take into account possible non-local-monotonicity.
 This does not change the worst-case-complexity (obtaining a single DNF being exponential) but might matter a lot in practice.
-As such, we will explore how this transformation, here using BDDs in BioLQM, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.}
+As such, we will explore how this transformation, here using BDDs in BioLQM, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.
 
@@
 We recall here the basic definition establishing that to produce something in a siphon you must consume something from the siphon.
-\revise{This corresponds to the idea that a siphon is a set of places that once unmarked remains unmarked.}
+This corresponds to the idea that a siphon is a set of places that once unmarked remains unmarked.
 
@@
 
-\revise{Intuitively, a siphon is a set of places that once unmarked remains so.
-If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the unmarked places remain unmarked).}
+Intuitively, a siphon is a set of places that once unmarked remains so.
+If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the unmarked places remain unmarked).
 This is precisely the definition of a trap space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap space.
@@
 By Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean model to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
-\revise{Note that in the case of stable states, this can be put in regard to the classical relationship between siphons and deadlocks in Petri nets.
+Note that in the case of stable states, this can be put in regard to the classical relationship between siphons and deadlocks in Petri nets.
 It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap space that might be useful for the analysis of Petri nets, but this is out of the scope of this article.
-}
 
@@
 \end{equation}
-\revise{The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net and has been used with SAT solvers in~\cite{oanea2010new,nabli2016enumerating}.}
+The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net and has been used with SAT solvers in~\cite{oanea2010new,nabli2016enumerating}.
 To make \(S\) to be a conflict-free siphon, we need to add to the system the rule
@@
 Note that the number of atoms of \(\mathcal{L}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
-\revise{In~\cite{DBLP:conf/ictai/ChevalierFPZ19}, there is an ASP characterization of trap spaces that does not rely on minimal DNFs either and thus seems very similar to our ASP encoding.
+In~\cite{DBLP:conf/ictai/ChevalierFPZ19}, there is an ASP characterization of trap spaces that does not rely on minimal DNFs either and thus seems very similar to our ASP encoding.
 Remarkably it only requires the DNF for the \emph{activation} part, using the information that it will only be used for locally-monotonic Boolean models.
@@
 Since mpbn is included in our benchmark this will be evaluated in our experiments.
-}
 
@@
 However, using a state-of-the-art tool like PyBoolNet~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
-\revise{mpbn~\cite{Paulev2020} does not give any answer either because it recognizes that model as non-locally-monotonic.}
+mpbn~\cite{Paulev2020} does not give any answer either because it recognizes that model as non-locally-monotonic.
 And hence, it is not possible to extract any (complex) attractor at all.
@@
 
-\revise{To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
+To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
 In addition, we also include the tool mpbn~\cite{Paulev2020} to the benchmarks although it only handles locally-monotonic models, whereas both Trappist and PyBoolNet can handle general models.
-}
 
@@
 Finally, we set a time limit of two minutes for each model.
+Note that we did not get the opportunity to run this benchmark on a proper dedicated computing workstation, the results are therefore only indicative of global trends and should not be interpreted as precise performance.
+This is also why in some rare cases finding all minimal trap spaces was faster than with the limit to 1000.
 
-\revise{For all the above tools, namely PyBoolNet, mpbn and CLINGO, we used the version available in the latest CoLoMoTo docker image tagged \texttt{2022-05-01}.}
+For all the above tools, namely PyBoolNet, mpbn and CLINGO, we used the version available in the latest CoLoMoTo docker image tagged \texttt{2022-05-01}.
 All the models and a CoLoMoTo notebook realizing the benchmarks can be found at \url{https://github.com/soli/trap-spaces-as-siphons}.
@@
 \begin{table}[!htb]
-  \caption{\revise{Timing comparisons between PyBoolNet, mpbn and Trappist on the PyBoolNet repository.
+  \caption{Timing comparisons between PyBoolNet, mpbn and Trappist on the PyBoolNet repository.
     Column \(n\) denotes the number of nodes of each model.
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
-    A number in bold indicates a ratio greater than three compared to the best result. ``NM'' indicates a non-locally-monotonic model.}}
+    A number in bold indicates a ratio greater than three compared to the best result. ``NM'' indicates a non-locally-monotonic model.}
   \centering%
@@
 As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}, the results are comparable with all minimal trap spaces found very fast.
-\revise{For 5 of the 29 models, mpbn did not give any answer because it recognized these models as not locally-monotonic.
-Note that on some very small models, Trappist is sometimes slower than PyBoolNet and/or mpbn, but still significantly under one second.}
+For 5 of the 29 models, mpbn did not give any answer because it recognized these models as not locally-monotonic.
+Note that on some very small models, Trappist is sometimes slower than PyBoolNet and/or mpbn, but still significantly under one second.
 Moreover, we believe that the result on the arellano\_rootstem model is caused by the cold start of the JVM for BioLQM.\@
-\revise{On the contrary, on every model that was a bit challenging for PyBoolNet or mpbn, the new method is far more efficient with speedups between one and two orders of magnitude.}
+On the contrary, on every model that was a bit challenging for PyBoolNet or mpbn, the new method is far more efficient with speedups between one and two orders of magnitude.
 
@@
 These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
-\revise{We then applied PyBoolNet, mpbn, and Trappist} to computing minimal trap spaces of these real-world models.
+We then applied PyBoolNet, mpbn, and Trappist to computing minimal trap spaces of these real-world models.
 It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
@@
 \begin{table}[!htb]
-  \caption{\revise{Timing comparisons between PyBoolNet (PBN), mpbn and Trappist on selected models from the literature.
+  \caption{Timing comparisons between PyBoolNet (PBN), mpbn and Trappist on selected models from the literature.
     Column \(n\) (resp.\ \(s\)) denotes the number of nodes (resp.\ source nodes) of each model.
@@
     ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of two minutes.
-``NM'' indicates a non-locally-monotonic model.}}
+``NM'' indicates a non-locally-monotonic model.}
   \centering%
@@
 \end{table}
-% s obtained by: grep -c '^\([a-zA-Z0-9_-]*\),\s*\1\s*\(#.*\)\?$' models/*.bnet
 
-%%% The proposed method and mpbn %%%
-\revise{The first observation is that for 26 of the 33 models (more than 78\%), mpbn did not give any answer because it recognized that these models as not locally-monotonic.
+The first observation is that for 26 of the 33 models (more than 78\%), mpbn did not give any answer because it recognized that these models as not locally-monotonic.
 For 6 of the 33 models where mpbn returned the answers, mpbn and Trappist are comparable in computation time, though surprisingly mpbn appears a bit slower on average.
@@
 However, since mbpn can handle only locally-monotonic models and Trappist can handle general models, it is difficult to further compare between them.
-Hence, we focus on only comparisons between PyBoolNet and Trappist in the following observations.}
+Hence, we focus on only comparisons between PyBoolNet and Trappist in the following observations.
 
-%%% The proposed method outperforms PyBoolNet %%%
-\revise{The second observation} is that the proposed method vastly outperforms PyBoolNet in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100--1000 nodes size range).
+The second observation is that the proposed method vastly outperforms PyBoolNet in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100--1000 nodes size range).
 Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
@@
 Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
-\revise{Since the ASP code generated is also quite declarative by nature, it is also possible to add to it supplementary constraints coming from the modeller in case one is looking for specific attractors.
-Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration.}
+Since the ASP code is declarative by nature, it is also possible to add to it supplementary constraints coming from the modeler in case one is looking for specific attractors.
+Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration.
 
-%%% It is not clear to show the advantage in ASP encoding %%%
-\revise{The third observation} is that for all the models where PyBoolNet finished before the timeout, once PyBoolNet went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
+The third observation is that for all the models where PyBoolNet finished before the timeout, once PyBoolNet went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
 For these models, the ASP solving phase of the proposed method also took very short time, all under one second.
@@
 
-%%% For very huge models %%%
-\revise{The last observation} is that for very large models (i.e., more than two thousand nodes) the proposed method did not manage to finish the Petri net conversion before the timeout, as marked by ``???'' in Table~\ref{tab:result_real}.
+The last observation is that for very large models (i.e., more than two thousand nodes) the proposed method did not manage to finish the Petri net conversion before the timeout, as marked by ``???'' in Table~\ref{tab:result_real}.
 This points to the fact that our current choice of using a BDD-based translation to obtain that Petri net encoding, though it provides a small/efficient ASP might be too costly to handle the largest models.
@@
 The evaluation of the feasibility of such strategy, and of its impact on smaller instances, remains to be done and is out of the scope of this first article.
-\revise{Recognizing that a model is locally-monotonic and applying in that specific case dedicated strategies as those of mpbn might also be a partial solution as shown in the SN-5 case, but not for the turei-2016 model.}
+Recognizing that a model is locally-monotonic and applying in that specific case dedicated strategies as those of mpbn might also be a partial solution as shown in the SN-5 case, but not for the turei-2016 model.
 
-\revise{Note that though enumerating the extremal siphons of a Petri net is exponential (see~\cite{nabli2016enumerating} for instance) this is apparently not the bottleneck of the proposed method, showing once again that networks obtained from biochemical models do have a specific structure.}
+Note that though enumerating the extremal siphons of a Petri net is exponential (see~\cite{nabli2016enumerating} for instance) this is apparently not the bottleneck of the proposed method, showing once again that networks obtained from biochemical models do have a specific structure.
 
@@
 
-Though we benchmarked this new approach against the state-of-the-art tool PyBoolNet, there are many more evaluations that we plan to do in the future.
+Though we benchmarked this new approach against state-of-the-art tools, there are many more evaluations that we plan to do in the future.
 First, the BioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
@@
 They could replace our ASP program if they outperform it.
+However, the current method appears to already perform very well even on the biggest models we have considered.
 
-\revise{However, the current method appears to already perform very well even on the biggest models we have considered.}
-We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.
-
-\revise{Finally, we think that the links between Petri nets and Boolean models that we stumbled upon in this method might have deeper roots.
-Exploring those connections might lead both to interesting topics of research for Petri nets, like a notion of trap-spaces, and for Boolean models.}
-
-% \subsubsection{Acknowledgments}
+Finally, we think that the links between Petri nets and Boolean models that we stumbled upon in this method might have deeper roots.
+Exploring those connections might lead both to interesting topics of research for Petri nets, like a notion of trap-spaces, and for Boolean models.
 

### 2022-11-14 manuscripts/CMSB2022/cmsb22.tex
manuscripts

--- 2022-06-27 cmsb22.tex
@@
 
-It is worth noting that the recent method presented in~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces avoids the prime-implicants computation by relying on the \emph{most-permissive semantics of Boolean models.
+It is worth noting that the recent method presented in~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces avoids the prime-implicants computation by relying on the \emph{most-permissive} semantics of Boolean models.
 This method has been implemented in the tool mpbn\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
@@
 Hence, it is still necessary to develop efficient methods for computing minimal trap spaces of large-scale general Boolean models.
-}
 

### 2022-11-14 manuscripts/TCS2022/tcs2022.tex
manuscripts

--- 2022-11-14 manuscripts/CMSB2022/cmsb22.tex
@@
-\documentclass[runningheads]{llncs}
-\usepackage[T1]{fontenc}
+%% 
+%% Copyright 2007-2020 Elsevier Ltd
+%% 
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%% 
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%% 
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%% 
+%% Template article for Elsevier's document class `elsarticle'
+%% with harvard style bibliographic references
+
+\documentclass[preprint,12pt]{elsarticle}
+
+%% Use the option review to obtain double line spacing
+%% \documentclass[preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn]{elsarticle}
+%% \documentclass[final,3p,times]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn]{elsarticle}
+%% \documentclass[final,5p,times]{elsarticle}
+%% \documentclass[final,5p,times,twocolumn]{elsarticle}
+
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
+
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+\usepackage{amsthm}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{example}{Example}[section]
+\newtheorem{definition}{Definition}[section]
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+\usepackage{lineno}
+
 \usepackage{hyperref}
-\usepackage{xcolor}
-\usepackage{booktabs}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-% hyperref for LNCS
-\renewcommand\UrlFont{\color{blue}\rmfamily}
+\usepackage{booktabs}
 
@@
 
+\journal{Theoretical Computer Science}
+
 \begin{document}
 
-\title{Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding}
-\titlerunning{Minimal trap spaces as maximal siphons}
+\begin{frontmatter}
 
-\author{Van-Giang Trinh\inst{1}\orcidID{0000-0001-6581-998X} \and \\
-  Belaid Benhamou\inst{1} \and \\
-  Kunihiko Hiraishi\inst{2}\orcidID{0000-0003-1750-1891} \and \\
-  Sylvain Soliman\inst{3}\orcidID{0000-0001-5525-7418}
-  }
+%% Title, authors and addresses
 
-\authorrunning{Trinh et al.}
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \affiliation{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%% \fntext[label3]{}
 
-\institute{%
-  LIS, Aix-Marseille Université, Marseille, France\\
-  \email{\{trinh.van-giang, belaid.benhamou\}@lis-lab.fr}
-  \and
-  School of Information Science, Japan Advanced Institute of Science and Technology, Japan\\
-  \email{hira@jaist.ac.jp}
-  \and
-  Lifeware team, Inria Saclay center, Palaiseau, France\\
-  \email{Sylvain.Soliman@inria.fr}
-}
+\title{Minimal trap spaces of Boolean networks are maximal siphons of their Petri net encoding}
 
-\maketitle
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%% \affiliation[label1]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%%
+%% \affiliation[label2]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+
+\author{Van-Giang Trinh\fnref{1}}
+\ead{trinh.van-giang@lis-lab.fr}
+
+\author{Belaid Benhamou\fnref{1}}
+\ead{belaid.benhamou@lis-lab.fr}
+
+\affiliation[1]{organization={LIS, Aix-Marseille University},%Department and Organization
+            %addressline={}, 
+            city={Marseille},
+            %postcode={}, 
+            %state={},
+            country={France}}
+
+\author{Sylvain Soliman\corref{cor1}\fnref{2}}
+\ead{Sylvain.Soliman@inria.fr}
+
+\affiliation[2]{organization={Lifeware team, Inria Saclay center},%Department and Organization
+            %addressline={}, 
+            city={Palaiseau},
+            %postcode={}, 
+            %state={},
+            country={France}}
+
+\cortext[cor1]{Corresponding author.}
 
 \begin{abstract}
+%% Text of abstract
 
-  Boolean modelling of gene regulation but also of post-trans\-criptomic systems has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
+Boolean networkling of gene regulation but also of post-trans\-criptomic systems has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
 This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
@@
 
-  In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, of a Boolean model.
+In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, of a Boolean network.
 It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
 After some technical preliminaries, we expose the concrete need for such a method and detail its implementation using Answer Set Programming.
-  We then demonstrate its efficiency and compare it to implicant-based methods on some large Boolean models from the literature.
+We then demonstrate its efficiency and compare it to implicant-based methods on some large Boolean networks from the literature.
 
-\keywords{Logical models \and Boolean models \and Trap spaces \and Attractor computation \and Petri nets \and Siphons}
 \end{abstract}
 
+%%Graphical abstract
+% \begin{graphicalabstract}
+% %\includegraphics{grabs}
+% \end{graphicalabstract}
+
+%%Research highlights
+% \begin{highlights}
+% \item Research highlight 1
+% \item Research highlight 2
+% \end{highlights}
+
+\begin{keyword}
+%% keywords here, in the form: keyword \sep keyword
+
+%% PACS codes here, in the form: \PACS code \sep code
+
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+
+Logical model \sep Boolean network \sep Trap space \sep Attractor computation \sep Petri net \sep Siphon \sep Answer set programming \sep Systems biology
+\end{keyword}
+
+\end{frontmatter}
+
+\linenumbers
+
+%% main text
 \section{Introduction}
@@
 
-Boolean modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}, even for modelling post-transcriptional mechanisms.
+Boolean networkling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}, even for modelling post-transcriptional mechanisms.
 This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas~\cite{aghamiri2020automated}.
@@
 
-It is worth noting that the recent method presented in~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces avoids the prime-implicants computation by relying on the \emph{most-permissive} semantics of Boolean models.
+It is worth noting that the recent method presented in~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces avoids the prime-implicants computation by relying on the \emph{most-permissive} semantics of Boolean networks.
 This method has been implemented in the tool mpbn\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
-However, this method is only applicable for \emph{locally-monotonic} Boolean models, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean models (i.e., including both locally-monotonic and non-locally-monotonic ones).
-The study~\cite{DBLP:journals/tcs/NoualRS13} highlights the need for non-locally-monotonic Boolean models in both biological and theoretical aspects.
-Hence, it is still necessary to develop efficient methods for computing minimal trap spaces of large-scale general Boolean models.
+However, this method is only applicable for \emph{locally-monotonic} Boolean networks, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean networks (i.e., including both locally-monotonic and non-locally-monotonic ones).
+The study~\cite{DBLP:journals/tcs/NoualRS13} highlights the need for non-locally-monotonic Boolean networks in both biological and theoretical aspects.
+Hence, it is still necessary to develop efficient methods for computing minimal trap spaces of large-scale general Boolean networks.
 
@@
 
-In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, of a general Boolean model.
-It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
-After some technical preliminaries, we expose the concrete need for such a method and detail its implementation using Answer Set Programming.
-We then demonstrate its efficiency and compare it to implicant-based methods on several large Boolean models from the literature.
+In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, of a general Boolean network.
+It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model (Section~\ref{sec:Main_finding}).
+After some technical preliminaries (Section~\ref{sec:Preliminaries}), we expose the concrete need for such a method and detail its implementation using Answer Set Programming~\ref{sec:Computation}.
+We then demonstrate its efficiency and compare it to implicant-based methods on several large Boolean networks from the literature (Section~\ref{sec:eval}).
 
@@
 
+% Mention the extensions for the journal version
+Herein we extend our previous work in~\cite{DBLP:conf/cmsb/TrinhBHS22} as follows.
+First, beyond the proposed ASP method, we propose several other possible methods using MaxSAT, CLP, and ILP.
+Second, we conduct a more comprehensive benchmark to evaluate all the proposed methods.
+Third, regarding the implementation, we have developed a converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of BioLQM.
+Finally, we explore several complexity analysis on maximal (conflict-free) siphon problems.
+
 \section{Preliminaries}
+\label{sec:Preliminaries}
 
-We will briefly recall here some preliminaries on Boolean models related to trap spaces and Petri nets.
-In the case of multi-level Logical models, an encoding into a Boolean model is always possible~\cite{Didier2011}.
+We will briefly recall here some preliminaries on Boolean networks related to trap spaces and Petri nets.
+In the case of multi-level Logical models, an encoding into a Boolean network is always possible~\cite{Didier2011}.
 
@@
 
+% TODO give formal definitions
+
 We recall here some definitions from~\cite{klarner2015computing} for the introduction of \emph{trap spaces}.
-Minimal trap spaces prove to be a very good approximation of the attractors of a Boolean model under asynchronous update schemes and have become the \emph{de facto} standard way to analyze models of a few tens of \emph{genes}~\cite{klarner2017pyboolnet,cifuentes2020control}.
+Minimal trap spaces prove to be a very good approximation of the attractors of a Boolean network under asynchronous update schemes and have become the \emph{de facto} standard way to analyze models of a few tens of \emph{genes}~\cite{klarner2017pyboolnet,cifuentes2020control}.
 
-Given a Boolean model \(\mathcal{M} = (V, F)\) with nodes \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\mathbb{B}^{n}\) with \(\mathbb{B} = \{0, 1\}\).
+Given a Boolean network \(\mathcal{M} = (V, F)\) with nodes \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\mathbb{B}^{n}\) with \(\mathbb{B} = \{0, 1\}\).
 A state \(s \in \mathbb{B}^{n}\) is a mapping \(s \colon V \mapsto \mathbb{B}\) that assigns either 0 (inactive) or 1 (active) to each node.
@@
 A \emph{trap space} is defined as a subspace that is also a trap set.
-It is noted that trap spaces of a Boolean model are independent of the update scheme of this model~\cite{klarner2015computing}.
+It is noted that trap spaces of a Boolean network are independent of the update scheme of this model~\cite{klarner2015computing}.
 Then, we define a partial order \(<\) on \(\mathcal{S}_{\mathcal{M}}^{\star}\) as: \(m < m'\) if and only if \(\mathcal{S}_{\mathcal{M}}[m] \subseteq \mathcal{S}_{\mathcal{M}}[m']\) and \(\mathcal{S}_{\mathcal{M}}[m] \neq \mathcal{S}_{\mathcal{M}}[m']\).
@@
 
-For example, let us consider the Boolean model shown in Example~\ref{example:BN}.
+For example, let us consider the Boolean network shown in Example~\ref{example:BN}.
 Figure~\ref{fig:stg} shows the dynamics of this model under the fully asynchronous update (i.e., only one node is nondeterministically selected in order to be updated at each time step).
 The model has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
-Since \(m_1 < m_2\), \(m_1\) is a minimal trap space of the Boolean model.
+Since \(m_1 < m_2\), \(m_1\) is a minimal trap space of the Boolean network.
 
 \begin{example}
-We give a Boolean model \(\mathcal{M} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
+We give a Boolean network \(\mathcal{M} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
 Herein, \(\land\), \(\lor\), and \(\neg\) denote the conjunction, disjunction, and negation logical operators, respectively.\label{example:BN}
@@
 \end{subfigure}%
-\caption{Dynamics and encoding of the Boolean model of Example~\ref{example:BN}}%
+\caption{Dynamics and encoding of the Boolean network of Example~\ref{example:BN}}%
 \label{fig:stg_and_PN}
@@
 
-\subsection{Petri net encoding of Boolean models}%
+\subsection{Petri net encoding of Boolean networks}%
 \label{sec:encoding}
@@
 
-The link between Boolean models \emph{à la} Thomas and Petri nets was originally established in~\cite{chaouiya2004qualitative} in order to make available formal methods like model-checking for the analysis of such systems.
-The basic encoding into 1-safe (i.e., never more than one token in each place) nets only holds for purely Boolean models but was later extended to multivalued Logical models in two ways, either in~\cite{chaouiya2011petri} with non 1-safe Petri nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.
+The link between Boolean networks \emph{à la} Thomas and Petri nets was originally established in~\cite{chaouiya2004qualitative} in order to make available formal methods like model-checking for the analysis of such systems.
+The basic encoding into 1-safe (i.e., never more than one token in each place) nets only holds for purely Boolean networks but was later extended to multivalued Logical models in two ways, either in~\cite{chaouiya2011petri} with non 1-safe Petri nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.
 
-Since our study is focused on Boolean models, we briefly recall the original encoding here.
+Since our study is focused on Boolean networks, we briefly recall the original encoding here.
 Its basis is that every node (\emph{gene}) \(v\) of the original model \(\mathcal{M} = (V, F)\) is represented by two separate places (\(p_v\) and \(\overline{p}_v\)), corresponding to its two states, active, and inactive, respectively.
@@
 And conversely for the inactivation.
-Let \(s\) be a state of the Boolean model and \(m_s\) be its corresponding marking in the encoded Petri net.
+Let \(s\) be a state of the Boolean network and \(m_s\) be its corresponding marking in the encoded Petri net.
 It holds that \(\forall v \in V\), \(s(v) = 0\) if and only if \(m_s(\overline{p}_v) = 1\) and \(s(v) = 1\) if and only if \(m_s(p_v) = 1\).
-Note also that at any marking \(m\) of the Petri net encoding a Boolean model, it always holds that \(m(p_v) + m(\overline{p}_v) = 1\).
+Note also that at any marking \(m\) of the Petri net encoding a Boolean network, it always holds that \(m(p_v) + m(\overline{p}_v) = 1\).
 
-The main property of this encoding is that it is completely faithful with respect to the update scheme of the original Boolean model.
+The main property of this encoding is that it is completely faithful with respect to the update scheme of the original Boolean network.
 For each node \(v\) of \(\mathcal{M}\), only transitions corresponding to \(v\) can change the current marking of \(p_v\) or \(\overline{p_v}\).
@@
 
-For illustration, let us reconsider the Boolean model shown in Example~\ref{example:BN}.
-Figure~\ref{fig:PN} shows the Petri net encoding of this Boolean model.
+For illustration, let us reconsider the Boolean network shown in Example~\ref{example:BN}.
+Figure~\ref{fig:PN} shows the Petri net encoding of this Boolean network.
 Place \(p_{x_1}\) (resp. \(\overline{p_{x_1}}\)) in \(\mathcal{P}\) represents the activation (resp.\ the inactivation) of node \(x_1\) in \(\mathcal{M}\).
@@
 
-Note that given a Boolean model in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the BioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
+Note that given a Boolean network in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the BioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
 This piece of software extracted from GINsim~\cite{chaouiya2012logical} and part of the CoLoMoTo\footnote{\url{http://colomoto.org/}}~\cite{naldi2015cooperative} software suite allows for easy conversion between standard formats.
-It also accepts many other common formats for Boolean models, notably the \verb|.bnet| files of the  BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} tools.
+It also accepts many other common formats for Boolean networks, notably the \verb|.bnet| files of the  BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} tools.
 The conversion is executed as follows:
@@
 
-Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining conditions for the activation and inactivation of the states.
+Note that transforming a Boolean network defined by its functions into its Petri net encoding roughly relies on obtaining conditions for the activation and inactivation of the states.
 In~\cite{chaouiya2004qualitative} this took the form of the whole truth table of the Boolean functions, but as shown in Appendix 1 of~\cite{chatain2014characterization} computing Disjunctive Normal Forms (DNF) of each Boolean function is enough.
@@
 \section{Minimal trap spaces as maximal conflict-free siphons}
-First, we add a definition related to any set of places of a Petri net encoding a Boolean model, and notably a siphon of such a net.
+\label{sec:Main_finding}
+
+First, we add a definition related to any set of places of a Petri net encoding a Boolean network, and notably a siphon of such a net.
 
@@
 
-  A set of places of Petri net \(\mathcal{P}\) encoding Boolean model \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{gene} of \(\mathcal{M}\).
+  A set of places of Petri net \(\mathcal{P}\) encoding Boolean network \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{node} of \(\mathcal{M}\).
   Then, a conflict-free siphon \(S\) is said to be \emph{maximal} if and only if there is no other conflict-free siphon \(S'\) such that \(S \subset S'\).
@@
 This is precisely the definition of a trap space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap space.
-For example, the Boolean model given in Example~\ref{example:BN} has two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
-The Petri net encoding of this Boolean model has five generic siphons, \(S_1 = \emptyset\), \(S_2 = \{p_{x_1}, \overline{p_{x_1}}\}\), \(S_3 = \{p_{x_2}, \overline{p_{x_2}}\}\), \(S_4 = \{\overline{p_{x_1}}, \overline{p_{x_2}}\}\), and \(S_5 = \{p_{x_1}, \overline{p_{x_1}}, p_{x_2}, \overline{p_{x_2}}\}\).
+For example, the Boolean network given in Example~\ref{example:BN} has two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
+The Petri net encoding of this Boolean network has five generic siphons, \(S_1 = \emptyset\), \(S_2 = \{p_{x_1}, \overline{p_{x_1}}\}\), \(S_3 = \{p_{x_2}, \overline{p_{x_2}}\}\), \(S_4 = \{\overline{p_{x_1}}, \overline{p_{x_2}}\}\), and \(S_5 = \{p_{x_1}, \overline{p_{x_1}}, p_{x_2}, \overline{p_{x_2}}\}\).
 However, only \(S_1\) and \(S_4\) are conflict-free siphons and correspond to \(m_2\) and \(m_1\), respectively.
@@
 
-  Let \(m\) be a subspace of Boolean model \(\mathcal{M} = (V, F)\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) such that:
+  Let \(m\) be a subspace of Boolean network \(\mathcal{M} = (V, F)\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) such that:
   \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S\] and \[\forall v \in V \setminus D_m, p_v \not \in S, \overline{p}_v \not \in S.\]
@@
 
-  Let \(\mathcal{M} = (V, F)\) be a Boolean model and \(\mathcal{P}\) be its Petri net encoding.
+  Let \(\mathcal{M} = (V, F)\) be a Boolean network and \(\mathcal{P}\) be its Petri net encoding.
 A subspace \(m\) is a trap space of \(\mathcal{M}\) if and only if its mirror \(S\) is a conflict-free siphon of \(\mathcal{P}\).
@@
 
-  Let \(\mathcal{M}\) be a Boolean model and \(\mathcal{P}\) be its Petri net encoding.
+  Let \(\mathcal{M}\) be a Boolean network and \(\mathcal{P}\) be its Petri net encoding.
 A subspace \(m\) is a minimal trap space of \(\mathcal{M}\) if and only if its mirror \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\).
@@
 
-By Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean model to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
+By Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean network to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
 Note that in the case of stable states, this can be put in regard to the classical relationship between siphons and deadlocks in Petri nets.
@@
 
-\section{Answer set programming-based method}
+\section{Computation methods}
+\label{sec:Computation}
+
 First, we show the characterization of all conflict-free siphons of the encoded Petri net \(\mathcal{P} = (P, T, W)\).
@@
 for each \(v \in V\).
-Note that the number of atoms of \(\mathcal{L}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).
+Note that the number of atoms of \(\mathcal{L}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean network and that number might be exponential in \(n\).
 In~\cite{DBLP:conf/ictai/ChevalierFPZ19}, there is an ASP characterization of trap spaces that does not rely on minimal DNFs either and thus seems very similar to our ASP encoding.
-Remarkably it only requires the DNF for the \emph{activation} part, using the information that it will only be used for locally-monotonic Boolean models.
+Remarkably it only requires the DNF for the \emph{activation} part, using the information that it will only be used for locally-monotonic Boolean networks.
 We would therefore expect that, when available, it will have comparable performance on the ASP part (the ASP program would be approximately twice smaller, though redundancy is not always bad in that field), but can also avoid combinatorial explosion of the Petri net encoding for some formula where the activation DNF is simple but the inhibition is not.
@@
 
+% TODO Choose another motivating example because this Boolean model can be easily handled by mpbn 1.7
+
 For a few years now we have been collaborating with biologists who build very large detailed and annotated maps and now wish to analyze the dynamics of the corresponding models.
-One of the main maps studied this way represents knowledge about the Rheumatoïd Arthritis~\cite{singh2018computational}, and was the main motivation for the development of a tool to automatically transform it into an executable Boolean model~\cite{aghamiri2020automated}.
+One of the main maps studied this way represents knowledge about the Rheumatoïd Arthritis~\cite{singh2018computational}, and was the main motivation for the development of a tool to automatically transform it into an executable Boolean network~\cite{aghamiri2020automated}.
 In the supplementary material of the paper, an excerpt of the map, focused around the apoptosis (cell death) module is transformed into a model of \emph{reasonable} size, namely 180 Boolean variables (model \verb|F5_RA_apoptosis_executable_module.sbml| of supplementary material S3, and model ``RA-apoptosis'' of Section~\ref{sec:eval}).
@@
 The current practice usually revolves then around fixing some inputs to plausible values and reducing the model accordingly.
-While this approach makes sense, it relies on potentially arbitrary decisions, and \emph{hides away} critical modelling choices that were actually not part of the original Boolean model or even of the starting map.
+While this approach makes sense, it relies on potentially arbitrary decisions, and \emph{hides away} critical modelling choices that were actually not part of the original Boolean network or even of the starting map.
 
@@
 
-We used a set of real-world Boolean models lying in various scales collected from numerous bibliographic sources.
+We used a set of real-world Boolean networks lying in various scales collected from numerous bibliographic sources.
 These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
@@
 
+\subsection{Randomly generated models}
+
 \section{Conclusion}
 
-In this article we proposed a new method for the computation of minimal trap spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
+In this article we proposed a new method for the computation of minimal trap spaces of Boolean networks, based on a new concept called maximal conflict-free siphons.
 This method is evaluated on large models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
-We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
+We believe that this opens up the way to a much better analysis of large Boolean networks, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
 
@@
 
-Finally, we think that the links between Petri nets and Boolean models that we stumbled upon in this method might have deeper roots.
-Exploring those connections might lead both to interesting topics of research for Petri nets, like a notion of trap-spaces, and for Boolean models.
+Finally, we think that the links between Petri nets and Boolean networks that we stumbled upon in this method might have deeper roots.
+Exploring those connections might lead both to interesting topics of research for Petri nets, like a notion of trap-spaces, and for Boolean networks.
 
-\bibliographystyle{splncs04}
-\bibliography{cmsb22}
+%% The Appendices part is started with the command \appendix;
+%% appendix sections are then done as normal sections
+%% \appendix
 
+%% \section{}
+%% \label{}
+
+%% For citations use: 
+%%       \citet{<label>} ==> Jones et al. [21]
+%%       \citep{<label>} ==> [21]
+%%
+
+%% If you have bibdatabase file and want bibtex to generate the
+%% bibitems, please use
+%%
+\bibliographystyle{elsarticle-num-names} 
+\bibliography{tcs2022.bib}
+
+%% else use the following coding to input the bibitems directly in the
+%% TeX file.
+
+% \begin{thebibliography}{00}
+
+% %% \bibitem[Author(year)]{label}
+% %% Text of bibliographic item
+
+% \bibitem[ ()]{}
+
+% \end{thebibliography}
 \end{document}
+
+\endinput
+%%
+%% End of file `elsarticle-template-num-names.tex'.

### 2022-11-14 manuscripts/TCS2022/tcs2022.tex
Several possible methods for computing maximal conflict-free siphons

@@
 \newtheorem{definition}{Definition}[section]
+\newtheorem{proposition}{Proposition}[section]
 
@@
 \newcommand\rownb{\stepcounter{rownumber}\arabic{rownumber}}
+
+\newcommand{\vangiang}[1]{\textcolor{magenta}{#1}}
+\newcommand{\belaid}[1]{\textcolor{blue}{#1}}
+\newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
 
@@
 
+\subsection{Characterization}
+\label{subsec:siphon_characterization}
+
 First, we show the characterization of all conflict-free siphons of the encoded Petri net \(\mathcal{P} = (P, T, W)\).
@@
 
-Then, we translate the above characterization into the ASP \(\mathcal{L}\) as follows.
-We introduce atom \verb|p-v| (resp.
-\verb|n-v|) to denote place \(p_v\) (resp.
-\(\overline{p_v}\)), \(\forall v \in V\).
+\subsection{Constraint Satisfaction Problem}
+\label{subsec:computation_csp}
+
+The following Boolean CSP directly derives from the above characterization:
+\begin{definition}
+
+  Given a Petri net \(\mathcal{P} = (P, T, W)\) encoding a Boolean model \(\mathcal{M} = (V, F)\).
+  The CSP \(\mathcal{C}(\mathcal{P})\) is the triple \((R, D, C)\) where
+  \begin{itemize}%
+    \item \(R = P\), i.e., a variable is introduced for each place of \(\mathcal{P}\),
+    \item \(D(p) = \mathbb{B}\) for all \(p \in R\), i.e., the variables are Boolean,
+    \item \(C = \{\neg p_v \vee \neg \overline{p_v} = 1 \mid \forall v \in V\} \bigwedge
+\{(p = 1 \rightarrow \bigvee_{p' \in pred(t)}p' = 1) \mid p \in P, t \in pred(p)\}\).
+  \end{itemize}
+
+\end{definition}
+
+\begin{proposition}
+\label{prop:csp_conflict_free_siphon}
+  \(\mathcal{C}(\mathcal{P})\) is satisfied by a valuation \(r\) if and only if
+  \[
+    \{p \in P \;|\; r(p) = 1\}\
+  \]
+  is a conflict-free siphon of \(\mathcal{P}\).
+
+\end{proposition}
+
+\begin{proof}
+
+  By the former part \(\neg p_v \vee \neg \overline{p_v}\) = 1 of \(C\), the conflict-freeness is imposed because for any satisfable valuation \(r\), \(r(p_v) = r(\overline{p_v}) = 1\) is impossible for all \(v \in V\).
+  As shown in~\cite{nabli2016enumerating}, the latter part of \(C\) can characterize the set of all generic siphons of \(\mathcal{P}\).
+  Hence, we can conclude the proof.
+
+\end{proof}
+
+In~\cite{nabli2016enumerating}, the set of all siphons of a given Petri net is characterized by a similar Boolean Constraint Satisfcation Problem (CSP) except the conflict-freeness constraint.
+From the encoded CSP, the set of all \emph{minimal} siphons of the Petri net can be enumerated in the set inclusion order.
+For enumerating siphons in the set inclusion order, the proposed method by~\cite{nabli2016enumerating} uses the technique that labels directly the Boolean variables with increasing value selection (i.e., to test first the absence, then the presence of a place in the candidate solution).
+The method has two implementations, one uses an iterated SAT procedure and the other uses Constraint Logic Programming with backtracking.
+
+One natural question is that how to use the CSP-based method for enumerating all the maximal conflict-free siphons of a Petri net encoding a Boolean model?
+Of course, the set of all conflict-free siphons of the Petri net can easily characterized by the CSP model presented in~\cite{nabli2016enumerating} along with the additional constraint \(\neg p_v \vee \neg \overline{p_v} = 1\), for each \(v \in V\), which represents the conflict-freeness.
+However, the main concern is to enumerate all the \emph{maximal} ones, which is not trivial to adapt from the CSP-based method.
+By Proposition~\ref{prop:csp_conflict_free_siphon}, the set of all maximal conflict-free siphons of \(\mathcal{P}\) can be enumerated in the (maximality) set inclusion order, by restarting the search each time a conflict-free siphon \(S\) is found, with the following additional constraint for disallowing any subset of that conflict-free siphon: \(\bigvee_{p \not \in S} p = 1\).
+For enumerating conflict-free siphons in the set inclusion order, we can use the same technique as used in~\cite{nabli2016enumerating} but with the opposite setting, i.e., labeling directly the Boolean variables with decreasing value selection.
+The correctness of this technique comes from the fact that once \(S\) is found, it is the conflict-free siphon of maximum cardinality among all the remaining feasible conflict-free siphons.
+Similar to~\cite{nabli2016enumerating}, the newly CSP-based method can also be implemented with SAT and CLP solvers.
+
+However a more direct method is to use a MaxSAT solver.
+We construct a MaxSAT problem with the following hard clauses:
+\[
+  (\neg p_v \vee \neg \overline{p_v}), \forall v \in V
+\]
+and
+\[
+  (\neg p \vee \bigvee_{p' \in pred(t)}p'), \forall p \in P, \forall t \in pred(p).
+\]
+We set a soft clause for each variable of the CSP and then use a ``minimal correction subset'' blocking strategy, which will ensure set-inclusion maximality of the solutions.
+This is what is implemented in \texttt{trappist} using the RC2 MaxSAT solver~\cite{DBLP:journals/jsat/IgnatievMM19} available through the \texttt{python-sat} package\footnote{\url{https://pysathq.github.io/docs/html/api/examples/rc2.html}}.
+
+\subsection{Answer set programming-based method}
+\label{subsec:computation_asp}
+
+Another possible method is to translate the characterization shown in Subsection~\ref{subsec:siphon_characterization} into the ASP \(\mathcal{L}\) as follows.
+We introduce atom \verb|p-v| (resp. \verb|n-v|) to denote place \(p_v\) (resp. \(\overline{p_v}\)), \(\forall v \in V\). 
 The set of all atoms in \(\mathcal{L}\) is given as \(\mathcal{A} = \bigcup_{v \in V}\{\verb|p-v|, \verb|n-v|\}\).
@@
 \[\verb|:- p-v, n-v.|\]
-for each \(v \in V\). This ASP rule guarantees that two places representing the same node in \(\mathcal{M}\) never belong to the same siphon of \(\mathcal{P}\), representing the conflict-freeness.
+for each \(v \in V\). 
+This ASP rule guarantees that two places representing the same node in \(\mathcal{M}\) never belong to the same siphon of \(\mathcal{P}\), representing the conflict-freeness.
 Naturally, a Herbrand model (see, e.g,~\cite{DBLP:journals/aicom/GebserKKOSS11}) of \(\mathcal{L}\) is equivalent to a conflict-free siphon of \(\mathcal{P}\).
@@
 By using this built-in option, we can compute all the set-inclusion maximal answer sets of \(\mathcal{L}\) (resp.\ all the minimal trap spaces of \(\mathcal{M}\)) in one execution.
+
+\subsection{Integer linear programming-based method}
+\label{subsec:computation_ilp}
+
+We first show how an Integer Linear Programming (ILP) \(\mathcal{I}\) can define a set of all conflict-free siphons of the encoded Petri net \(\mathcal{P}\).
+We introduce \emph{binary} variable \verb|p-v| (resp. \verb|n-v|) to denote place \(p_v\) (resp. \(\overline{p_v}\)), \(\forall v \in V\).
+The set of all binary variables in \(\mathcal{I}\) is \(\bigcup_{v \in V}\{\verb|p-v|, \verb|n-v|\}\).
+For each pair \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\), we translate the rule (\ref{eq:siphon}) into the ILP inequality
+\[
+\verb|a <= a_1 + ... + a_k|
+\]
+where \verb|a| is the binary variable representing place \(p\) and \{\verb|a_1|, \dots,\verb|a_k|\} is the set of binary variable representing places in \(pred(t)\).
+The rule (\ref{eq:conflict}) is translated into the ILP inequality
+\[\verb|p-v + n-v <= 1|\]
+for each \(v \in V\).
+This inequality forbids both \verb|p-v| and \verb|n-p| receive the value 1, thus representing the conflict-freeness.
+Since we only consider feasible solutions, the objective function is set to \verb|max p-v| for some \(v \in V\).
+Naturally, a solution \(I\) of \(\mathcal{I}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\).
+The conversion is that
+\[
+  S = \{p \in P \;|\; I(\verb|a-p|) = 1\}
+\]
+where \verb|a-p| is the binary variable presenting place \(p\).
+
+We can see the similarity between \(\mathcal{I}\) and the encoded ASP shown in the previous subsection.
+However, due to the nature of solutions of an ILP, it is hard to compute all the set-inclusion maximal solutions of \(\mathcal{I}\) in one execution of an ILP solver.
+Hence, we propose an iterative approach as follows.
+
+The conflict-free siphon of maximum cardinality is of course maximal.
+Therefore, we impose the following objective function:
+\[
+  \verb|max| \sum_{v \in V}(\verb|p-v| + \verb|n-v|).
+\]
+Now, \(\mathcal{I}\) can be solved using a general purpose ILP solver.
+If it admits any solution \(I^{*}\), the corresponding conflict-free siphon (say \(S^{*}\)) is maximal.
+Hence, it makes sense that it does not need to find any other conflict-free siphon of the net that is strictly contained in \(S^{*}\).
+To do this, we add to \(\mathcal{I}\) a new inequality
+\[
+  \verb|1 <=| \sum_{p \in P \setminus S^{*}} \verb|a-p|
+\]
+where \verb|a-p| is the binary variable presenting place \(p\).
+Now, we solve \(\mathcal{I}\) again to find a new solution.
+If a new solution \(I'\) exists, then let \(S'\) be its corresponding conflict-free siphon.
+Indeed, abide by the newly added inequality, we have \(S' \cap (P \setminus S^{*}) \neq \emptyset\) because there is some \verb|a-p| with \(p \in P \setminus S^{*}\) such that \(I'(\verb|a-p|) = 1\).
+This implies that it is impossible that \(S' = S^{*}\) or \(S' \subset S^{*}\).
+By the objective function, it means that \(S'\) is the conflict-free siphon of maximum cardinality among the conflict-free siphons that are not contained in \(S^{*}\).
+Hence, \(S'\) is also a maximal conflict-free siphon.
+Again, we add to \(\mathcal{I}\) a new inequality with respect to the newly found siphon.
+The above process is iterated until \(\mathcal{I}\) becomes unfeasible, this means that there is no further maximal conflict-free siphon.
+Thus, all the maximal conflict-free siphons of the Petri net have been found.
 
@@
     \rownb & irons\_yeast & 18 & 1 & 0.05 & 0.01 & 0.02\\
-    \rownb & jaoude\_thdiff & 103 & >1000 & \textbf{1.44} & \textbf{0.90} & 0.09\\
+    \rownb & jaoude\_thdiff & 103 & $>1000$ & \textbf{1.44} & \textbf{0.90} & 0.09\\
     \rownb & klamt\_tcr & 40 & 8 & 0.04 & 0.01 & 0.03\\
@@
     \rownb & saadatpour\_guardcell & 13 & 1 & 0.03 & 0.01 & 0.01\\
-    \rownb & selvaggio\_emt & 56 & >1000 & \textbf{1.02} & \textbf{0.52} & 0.09\\
+    \rownb & selvaggio\_emt & 56 & $>1000$ & \textbf{1.02} & \textbf{0.52} & 0.09\\
     \rownb & tournier\_apoptosis & 12 & 3 & 0.04 & 0.01 & 0.01\\
@@
     \rownb & colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.20 & NM & 0.02 & 0.01\\
-    \rownb & mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & >1000 & 0.76 & NM & 0.01 & DNF\\
+    \rownb & mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & $>1000$ & 0.76 & NM & 0.01 & DNF\\
     \rownb & IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 1.29 & NM & 0.01 & 0.01\\
-    \rownb & Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & >1000 & DNF & NM & 0.03 & DNF\\
+    \rownb & Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & $>1000$ & DNF & NM & 0.03 & DNF\\
     \rownb & Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 1.04 & 0.06 & 0.08\\ \midrule
     \rownb & adhesion-cip-migration~\cite{guberman2020boolean} & 121 & 4 & 78 & 36.80 & 0.35 & 0.27 & 0.09\\
-    \rownb & interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & >1000 & 10.11 & NM & 0.02 & DNF\\
+    \rownb & interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & $>1000$ & 10.11 & NM & 0.02 & DNF\\
     \rownb & TCR-TLR5-signaling~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 2.06 & NM & 0.03 & 0.02\\
@@
     \rownb & HIV-1~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & NM & 0.06 & 0.06\\
-    \rownb & fibroblasts~\cite{helikar2008emergent} & 139 & 9 & >1000 & DNF & NM & 0.08 & DNF\\
-    \rownb & HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & >1000 & 6.52 & NM & 0.02 & DNF\\
-    \rownb & kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & >1000 & DNF & NM & 0.19 & DNF\\
-    \rownb & virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & >1000 & DNF & NM & 0.03 & DNF\\
-    \rownb & immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & >1000 & DNF & NM & 0.08 & DNF\\
-    \rownb & RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & >1000 & DNF & NM & 0.02 & DNF\\
-    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & 37 & >1000 & 87.13 & NM & 0.03 & DNF\\
-    \rownb & er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & >1000 & 17.88 & NM & 0.03 & DNF\\
+    \rownb & fibroblasts~\cite{helikar2008emergent} & 139 & 9 & $>1000$ & DNF & NM & 0.08 & DNF\\
+    \rownb & HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & $>1000$ & 6.52 & NM & 0.02 & DNF\\
+    \rownb & kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & $>1000$ & DNF & NM & 0.19 & DNF\\
+    \rownb & virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & $>1000$ & DNF & NM & 0.03 & DNF\\
+    \rownb & immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & $>1000$ & DNF & NM & 0.08 & DNF\\
+    \rownb & RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & $>1000$ & DNF & NM & 0.02 & DNF\\
+    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & 37 & $>1000$ & 87.13 & NM & 0.03 & DNF\\
+    \rownb & er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & $>1000$ & 17.88 & NM & 0.03 & DNF\\
     \rownb & cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 101.18 & NM & 0.40 & 0.07\\
     \rownb & CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & 2.45 & 0.07 & 0.08\\
-    \rownb & T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & >1000 & 69.94 & NM & 0.05 & DNF\\
-    \rownb & ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & >1000 & DNF & NM & 0.24 & DNF\\
-    \rownb & macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & >1000 & 16.43 & NM & 0.05 & DNF\\
-    \rownb & cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & >1000 & 1.42 & NM & 0.08 & DNF\\
-    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & >1000 & DNF & NM & 0.29 & DNF\\ \midrule
-    \rownb & KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & >1000 & DNF & 20.76 & 3.40 & DNF\\
-    \rownb & human-network~\cite{kim2011reduction} & 1953 & 669 & >1000 & DNF & 23.62 & 5.93 & DNF\\
-    \rownb & SN-5~\cite{kim2013rmod} & 2746 & 829 & >1000 & DNF & 28.94 & DNF & DNF\\
+    \rownb & T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & $>1000$ & 69.94 & NM & 0.05 & DNF\\
+    \rownb & ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & $>1000$ & DNF & NM & 0.24 & DNF\\
+    \rownb & macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & $>1000$ & 16.43 & NM & 0.05 & DNF\\
+    \rownb & cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & $>1000$ & 1.42 & NM & 0.08 & DNF\\
+    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & $>1000$ & DNF & NM & 0.29 & DNF\\ \midrule
+    \rownb & KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & $>1000$ & DNF & 20.76 & 3.40 & DNF\\
+    \rownb & human-network~\cite{kim2011reduction} & 1953 & 669 & $>1000$ & DNF & 23.62 & 5.93 & DNF\\
+    \rownb & SN-5~\cite{kim2013rmod} & 2746 & 829 & $>1000$ & DNF & 28.94 & DNF & DNF\\
     \rownb & turei-2016~\cite{lee2019signal} & 4691 & 1257 & ??? & DNF & DNF & DNF & DNF\\

### 2022-12-21 trappist/cp.py
Minizinc install (strange PATH issues)

@@ -29,23 +29,34 @@ def create_cp(petri_net: nx.DiGraph) -> Model:
     for node in (
             n for n, k in petri_net.nodes(data="kind") if k == "place"
     ):
-        model.add_string(f"var bool: {node};\n")
-        model.add_string(f"var bool: -{node};\n")
+        model.add_string(f"var bool: {zincify(node)};\n")
+        print(f"var bool: {zincify(node)};\n")
     for node, kind in petri_net.nodes(data="kind"):
         if kind == "place":
             if not node.startswith("-"):
                 # conflict-freeness
-                model.add_string(f"constraint not {node} \\/ not -{node};\n")
+                model.add_string(f"constraint not {node} \\/ not not_{node};\n")
+                print(f"constraint not {node} \\/ not not_{node};\n")
         else:
             # it's a transition, apply siphon
             # (if one succ is true, one pred must be true)
-            or_preds = petri_net.predecessors(node)
+            or_preds = [zincify(p) for p in petri_net.predecessors(node)]
             for succ in petri_net.successors(node):
-                if succ not in or_preds:  # optimize obvious tautologies
-                    or_string = (or_preds + ["-" + succ]).join("\\/")
-                    model.add_string(f"constraint {or_string};\n")
+                zsucc = zincify(succ)
+                if zsucc not in or_preds:  # optimize obvious tautologies
+                    or_string = " \\/ ".join(or_preds)
+                    model.add_string(f"constraint {zsucc} -> {or_string};\n")
+                    print(f"constraint {zsucc} -> {or_string};\n")
     model.add_string("solve satisfy;\n")
+    print("solve satisfy;\n")
     return model
+
+
+def zincify(variable: str) -> str:
+    """Return the name of the corresponding CP variable."""
+    if variable.startswith("-"):
+        return f"not_{variable[1:]}"
+    return variable
 
 
 def solve_cp(

### 2022-12-21 trappist/bnet.py
Only solving first trap-space

@@ -67,7 +67,7 @@ def read_bnet(fileobj: IO, method: str) -> nx.DiGraph:
         net.add_node(
             "-" + x, kind="place"
         )  # convention in PNML files obtained from bnet
-        if method in ("asp", "sat"):
+        if method in ("asp", "sat", "cp"):
             vx = bddvar(x)
             fx = expr2bdd(expr(fx))
             # print(x, fx)

### 2022-12-21 trappist/cp.py

@@ -18,20 +18,24 @@ along with this program.  If not, see <https://www.gnu.org/licenses/>.
 
 from typing import Generator, List
 
-import networkx as nx  # TODO maybe replace with lists/dicts
+from minizinc import Instance, Model, Result, Solver, Status
 
-from minizinc import Instance, Model, Solver
+import networkx as nx  # TODO maybe replace with lists/dicts
 
 
 def create_cp(petri_net: nx.DiGraph) -> Model:
     """Create the CP program for the max conflict-free siphons of petri_net."""
     model = Model()
+    vars = []
     for node in (
             n for n, k in petri_net.nodes(data="kind") if k == "place"
     ):
-        model.add_string(f"var bool: {zincify(node)};\n")
-        print(f"var bool: {zincify(node)};\n")
+        v = zincify(node)
+        vars.append(v)
+        model.add_string(f"var bool: {v};\n")
+        print(f"var bool: {v};\n")
     for node, kind in petri_net.nodes(data="kind"):
+        print(f"node, kind: {node}, {kind}")
         if kind == "place":
             if not node.startswith("-"):
                 # conflict-freeness
@@ -47,8 +51,8 @@ def create_cp(petri_net: nx.DiGraph) -> Model:
                     or_string = " \\/ ".join(or_preds)
                     model.add_string(f"constraint {zsucc} -> {or_string};\n")
                     print(f"constraint {zsucc} -> {or_string};\n")
-    model.add_string("solve satisfy;\n")
-    print("solve satisfy;\n")
+    model.add_string(f"solve :: bool_search([{', '.join(vars)}], input_order, indomain_max) satisfy;\n")
+    print(f"solve :: bool_search([{', '.join(vars)}], input_order, indomain_max) satisfy;\n")
     return model
 
 
@@ -69,25 +73,25 @@ def solve_cp(
     # TODO handle time limit?
     while max_output == 0 or nsol < max_output:
         result = inst.solve()
-        if result.solution is not None:
+        if result.status == Status.SATISFIED:
             nsol += 1
             print(result.solution)
-            yield result.solution
+            yield result
             # lexicographic constraint and non-superset
             # inst.add_string("")
         else:
             break
 
 
-def sat_to_bool(
-    places: List[str], all_places: List[str], model: List[int]
+def cp_to_bool(
+    places: List[str], res: Result
 ) -> List[str]:
-    """Transform a model to a list of 0/1/- strings."""
[…]

### 2022-12-22 manuscripts/TCS2022/tcs2022.tex
Formal definitions of Boolean networks and trap spaces

@@
 
-Boolean networkling of gene regulation but also of post-trans\-criptomic systems has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
+Boolean modeling of gene regulation but also of post-trans\-criptomic systems has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
 This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
@@
 
-Boolean networkling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}, even for modelling post-transcriptional mechanisms.
+Boolean modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}, even for modelling post-transcriptional mechanisms.
 This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas~\cite{aghamiri2020automated}.
@@
 It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model (Section~\ref{sec:Main_finding}).
-After some technical preliminaries (Section~\ref{sec:Preliminaries}), we expose the concrete need for such a method and detail its implementation using Answer Set Programming~\ref{sec:Computation}.
+After some technical preliminaries (Section~\ref{sec:Preliminaries}), we expose the concrete need for such a method and detail its implementation using Answer Set Programming (Section~\ref{sec:Computation}).
 We then demonstrate its efficiency and compare it to implicant-based methods on several large Boolean networks from the literature (Section~\ref{sec:eval}).
@@
 Herein we extend our previous work in~\cite{DBLP:conf/cmsb/TrinhBHS22} as follows.
-First, beyond the proposed ASP method, we propose several other possible methods using MaxSAT, CLP, and ILP.
-Second, we conduct a more comprehensive benchmark to evaluate all the proposed methods.
-Third, regarding the implementation, we have developed a converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of BioLQM.
+First, we showcase one theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
+Second, beyond the proposed ASP method, we propose several other possible methods using MaxSAT, CLP, and ILP.
+Third, we conduct a more comprehensive benchmark on both real-world and randomly generated models to evaluate all the proposed methods.
+Forth, regarding the implementation, we have developed a converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of BioLQM.
 Finally, we explore several complexity analysis on maximal (conflict-free) siphon problems.
@@
 
-We will briefly recall here some preliminaries on Boolean networks related to trap spaces and Petri nets.
-In the case of multi-level Logical models, an encoding into a Boolean network is always possible~\cite{Didier2011}.
+We shall briefly recall here some preliminaries on Boolean networks related to trap spaces and Petri nets.
+In the case of multi-level logical models, an encoding into a Boolean network is always possible~\cite{Didier2011}.
+
+\subsection{Boolean networks}
+
+\begin{definition}
+
+  A Boolean Network (BN) is a pair \(\mathcal{N} = (V, F)\) where:
+  \begin{itemize}
+    \item \(V = \{v_1, \dots, v_n\}\) is the set of nodes.
+    We use \(v_i\) to denote both the node \(v_i\) and its associated Boolean variable.
+
+    \item \(F = \{f_1, \dots, f_n\}\) is the set of update functions.
+    Each function \(f_i\) is associated with node \(v_i\) and satisfies \(f_i \colon \mathbb{B}^{\vert IN(v_i)\vert} \mapsto \mathbb{B}\) where \(\mathbb{B} = \{0, 1\}\) and \(IN(v_i)\) denotes the set of input nodes of \(v_i\).
+  \end{itemize}
+
+\label{def:BN}
+\end{definition}
+
+A Boolean function is \emph{locally-monotonic} if it can be represented by a formula in disjunctive normal form in which all occurrences of any given literal are either negated or non-negated~\cite{Paulev2020}.
+A Boolean network is said to be locally-monotonic if all its Boolean functions are locally-monotonic. 
+Otherwise, this model is said to be non-locally-monotonic.
+
+A state \(v \in \mathbb{B}^{n}\) is as a mapping \(v \colon V \mapsto \mathbb{B}\) that assigns either 0 (inactive) or 1 (active) to each node.
+We denote the set of all possible states of a Boolean network \(\mathcal{N}\) by \(\mathcal{S}_{\mathcal{N}} = \mathbb{B}^n\).
+At each time step \(t\), node \(v_i\) can update its state by
+\[v_i(t + 1) = f_i(v(t))\]
+where \(v(t)\) is the state of \(\mathcal{N}\) at time \(t\) and \(v_i(t + 1)\) is the state of node \(v_i\) at time \(t + 1\).
+Note that for simplicity, we write \(f_i(v(t))\) even \(IN(v_i) \subset V\).
+An update scheme of a Boolean network specifies the way that the nodes update their states through time evolution~\cite{thomas1991regulatory}. 
+Following the update scheme, the Boolean network transits from a state to another state (possibly identical).
+This transition is called the \emph{state transition} and denoted by \(\rightarrow \subseteq \mathcal{S}_{\mathcal{N}} \times \mathcal{S}_{\mathcal{N}}\).
+Then the dynamics of \(\mathcal{N}\) is captured by the directed graph \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\) called the State Transition Graph (STG).
+There are two main types of update schemes~\cite{thomas1991regulatory}: synchronous, where all the nodes are update simultaneously, and fully asynchronous, where only one node is nondeterministically selected to be updated.
 
 \subsection{Traps spaces}
-
-% TODO give formal definitions
 
@@
 
-Given a Boolean network \(\mathcal{M} = (V, F)\) with nodes \(V=(v_{1},\dots,v_{n})\) and Boolean functions \(F=(f_{1},\dots,f_{n})\), its state-space is \(\mathcal{S}_{\mathcal{M}} =\mathbb{B}^{n}\) with \(\mathbb{B} = \{0, 1\}\).
-A state \(s \in \mathbb{B}^{n}\) is a mapping \(s \colon V \mapsto \mathbb{B}\) that assigns either 0 (inactive) or 1 (active) to each node.
-At each time step \(t\), node \(v_i\) can update its state by
-\[v_i(t + 1) \in \{v_{i}(t), f_i(v(t))\}\]
-where \(v(t)\) is the state of \(\mathcal{M}\) at time \(t\), \(v_i(t + 1)\) is the state of node \(v_i\) at time \(t + 1\) and an update scheme specifies which value is chosen~\cite{thomas1991regulatory}.
-There are two main types of update schemes: synchronous, where for all \(i, v_{i}(t+1) = f_{i}(v(t))\), and asynchronous, where an \(i\) is chosen among the possible updates (\(f_{i}(v(t))\neq v_{i}(t)\)), and all the other nodes \(j\) remain fixed at \(v_{j}(t)\).
+An non-empty set \(T \subseteq \mathcal{S}_{\mathcal{N}}\) is a trap set with respect to \(\rightarrow\) if for every \(x \in T\) and \(y \in S\) with \(x \rightarrow y\) it holds that \(y \in T\)~\cite{klarner2015computing}.
+An attractor of \(\mathcal{N}\) with respect to \(\rightarrow\) can be defined as an inclusion-wise minimal trap set of \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\).
+An attractor can be also seen as a terminal strongly connected component of \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\)~\cite{chatain2014characterization}.
+An attractor of size 1 is called a fixed point, otherwise a cyclic attractor~\cite{klarner2015computing}.
 
-A non-empty set \(T \subseteq S_{\mathcal{M}}\) is a \emph{trap set} of \(\mathcal{M}\) if and only if for every \(x \in T\) and \(y \in S_{\mathcal{M}}\) with \(y\) is reachable from \(x\) it holds that \(y \in T\).
-
-A \emph{subspace} \(m\) of \(\mathcal{S}_{\mathcal{M}}\) is characterized by its fixed nodes (denoted by \(D_m\)) and free nodes.
-This subspace can be specified by an assignment \(m : D_m \mapsto \mathbb{B}\) where \(D_m \subseteq V\) and \(m(u)\) is the value of node \(u \in D_m\).
-The remaining nodes of \(\mathcal{M}\), \(V \setminus D_m\), are said to be \emph{free}, i.e., they can receive any Boolean value.
-We write subspaces like states but use in addition the symbol \(\star\) to indicate that a node is free.
-A subspace \(m\) thus corresponds to the set of states \(\mathcal{S}_{\mathcal{M}}[m] := \{s \in \mathcal{S}_{\mathcal{M}}\;|\;\forall v \in D_m : s(v) = m(v)\}\).
-For example, \(m = \star\star1\) means that \(D_m = \{v_3\}\), \(m(v_3) = 1\), and corresponds to the set of states \(\{001, 011, 101, 111\}\).
-Let \(\mathcal{S}_{\mathcal{M}}^{\star}\) denote the set of all possible subspaces of \(\mathcal{M}\).
-Note that \(\left|\mathcal{S}_{\mathcal{M}}^{\star}\right| = 3^n\) and \(S_{\mathcal{M}} \subset \mathcal{S}_{\mathcal{M}}^{\star}\)~\cite{klarner2015computing}.
+A subspace \(m\) of a Boolean network \(\mathcal{N} = (V, F)\) is a mapping \(m \colon V \mapsto \mathbb{B} \cup \{\star\}\).
+\(m(v_i) \in \mathbb{B}\) means that the value of \(v_i\) is fixed in \(m\) and \(v_i\) is called a fixed variable.
+\(m(v_i) \in \star\) means that the value of \(v_i\) is free in \(m\) and \(v_i\) is called a free variable.
+We denote \(D_m\) the set of all fixed variables of \(m\).
+A subspace \(m\) is equivalent to a set of states:
+\[
+\mathcal{S}_{\mathcal{N}}[m] := \{s \in \mathcal{S}_{\mathcal{N}} \mid \forall v \in D_m \colon s(v) = m(v)\}.
+\] For example, \(m = \star\star1\) (for simplicity, we write subspaces likes states) means that \(D_m = \{v_3\}, m(v_3) = 1\), and it is equivalent to the set of states \(\{001, 011, 101, 111\}\).
+We denote \(\mathcal{S}_{\mathcal{N}}^{\star} = (\mathbb{B} \cup \{\star\})^n\) the set of all possible subspaces of \(\mathcal{N}\).
+Note that \(\left\vert\mathcal{S}_{\mathcal{M}}^{\star}\right\vert = 3^n\) and \(S_{\mathcal{N}} \subset \mathcal{S}_{\mathcal{N}}^{\star}\)~\cite{klarner2015computing}.
 
@@
 It is noted that trap spaces of a Boolean network are independent of the update scheme of this model~\cite{klarner2015computing}.
-Then, we define a partial order \(<\) on \(\mathcal{S}_{\mathcal{M}}^{\star}\) as: \(m < m'\) if and only if \(\mathcal{S}_{\mathcal{M}}[m] \subseteq \mathcal{S}_{\mathcal{M}}[m']\) and \(\mathcal{S}_{\mathcal{M}}[m] \neq \mathcal{S}_{\mathcal{M}}[m']\).
-Consequently, a trap space \(m\) is minimal if and only if there is no trap space \(m' \in \mathcal{S}_{\mathcal{M}}^{\star}\) such that \(m' < m\).
+Then, we define a partial order \(<\) on \(\mathcal{S}_{\mathcal{N}}^{\star}\) as: \(m < m'\) if and only if \(\mathcal{S}_{\mathcal{N}}[m] \subseteq \mathcal{S}_{\mathcal{N}}[m']\) and \(\mathcal{S}_{\mathcal{N}}[m] \neq \mathcal{S}_{\mathcal{N}}[m']\).
+Consequently, a trap space \(m\) is minimal if and only if there is no trap space \(m' \in \mathcal{S}_{\mathcal{N}}^{\star}\) such that \(m' < m\).
 

### 2022-12-22 manuscripts/TCS2022/tcs2022.tex
Petri net definition

@@
 A place \(p\) is marked by a marking \(m\) if and only if \(m(p) > 0\). 
+Marking \(m\) can be seen as a subset of \(P\) that contains all marked places by \(m\).
 We shall write \(pred(x)\) (resp.\ \(succ(x)\)) to represent the set of vertices that have a (non-zero weighted) arc leading to (resp.\ coming from) \(x\).
+In this work, we consider a class of Petri nets called 1-safe Petri nets where every place has at most 1 token and all arcs are of weight 1.
+In this case, weights are implicitly omitted in the arcs of a Petri net.
+Then, a transition \(t \in T\) is \emph{enabled} at a marking \(m\) if and only if \(pred(t) \subseteq m\). 
+The firing of \(t\) leads to a new marking \(m'\) specified by \(m' = (m \backslash pred(t)) \cup succ(t)\).
+Note that when multiple transitions are enabled, we need to embed one firing scheme (similar to the update scheme of a Boolean network) to the Petri net.
+The classical firing scheme is that only one of the enabled transition is non-deterministically chosen to fire~\cite{Murata1989}.
 
 The link between Boolean networks \emph{à la} Thomas and Petri nets was originally established in~\cite{chaouiya2004qualitative} in order to make available formal methods like model-checking for the analysis of such systems.
-The basic encoding into 1-safe (i.e., never more than one token in each place) nets only holds for purely Boolean networks but was later extended to multivalued Logical models in two ways, either in~\cite{chaouiya2011petri} with non 1-safe Petri nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.
+The basic encoding into 1-safe (i.e., never more than one token in each place) nets only holds for purely Boolean networks but was later extended to multivalued logical models in two ways, either in~\cite{chaouiya2011petri} with non 1-safe Petri nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.
 
 Since our study is focused on Boolean networks, we briefly recall the original encoding here.
-Its basis is that every node (\emph{gene}) \(v\) of the original model \(\mathcal{M} = (V, F)\) is represented by two separate places (\(p_v\) and \(\overline{p}_v\)), corresponding to its two states, active, and inactive, respectively.
+Its basis is that every node (\emph{gene}) \(v\) of the original model \(\mathcal{N} = (V, F)\) is represented by two separate places (\(p_v\) and \(\overline{p}_v\)), corresponding to its two states, active, and inactive, respectively.
 Each conjunct of the logical function that activates the \emph{gene} will lead to a transition \(t\), consuming the inactive place (i.e., a directional arc from \(\overline{p}_v\) to \(t\)), producing the active place (i.e., a directional arc from \(t\) to \(p_v\)), and with all other literals both consumed and produced (i.e., a bidirectional arc).
@@
 The main property of this encoding is that it is completely faithful with respect to the update scheme of the original Boolean network.
-For each node \(v\) of \(\mathcal{M}\), only transitions corresponding to \(v\) can change the current marking of \(p_v\) or \(\overline{p_v}\).
+For each node \(v\) of \(\mathcal{N}\), only transitions corresponding to \(v\) can change the current marking of \(p_v\) or \(\overline{p_v}\).
 In addition, at any marking at most one of such transitions is enabled because \(m(p_v) + m(\overline{p}_v) = 1\) holds.
-Hence, for any update scheme in \(\mathcal{M}\), we have a corresponding firing scheme in \(\mathcal{P}\), which preserves the equivalence between the dynamics of \(\mathcal{M}\) and \(\mathcal{P}\)~\cite{DBLP:journals/nc/ChatainHKPT20}.
+Hence, for any update scheme in \(\mathcal{N}\), we have a corresponding firing scheme in \(\mathcal{P}\), which preserves the equivalence between the dynamics of \(\mathcal{N}\) and \(\mathcal{P}\)~\cite{DBLP:journals/nc/ChatainHKPT20}.
 
@@
 Figure~\ref{fig:PN} shows the Petri net encoding of this Boolean network.
-Place \(p_{x_1}\) (resp. \(\overline{p_{x_1}}\)) in \(\mathcal{P}\) represents the activation (resp.\ the inactivation) of node \(x_1\) in \(\mathcal{M}\).
-Marking \(\{p_{x_1}, \overline{p_{x_2}}\}\) in \(\mathcal{P}\) represents state 10 in \(\mathcal{M}\).
+Place \(p_{x_1}\) (resp. \(\overline{p_{x_1}}\)) in \(\mathcal{P}\) represents the activation (resp.\ the inactivation) of node \(x_1\) in \(\mathcal{N}\).
+Marking \(\{p_{x_1}, \overline{p_{x_2}}\}\) in \(\mathcal{P}\) represents state 10 in \(\mathcal{N}\).
 Transitions \(t^{1}_{x_1}\) and \(t^{2}_{x_1}\) represent the update of node \(x_1\).
 Of course, in any marking \(t^{1}_{x_1}\) and \(t^{2}_{x_1}\) cannot be both enabled.
-Then, the fully asynchronous update scheme in \(\mathcal{M}\) corresponds to the classical firing scheme in \(\mathcal{P}\) where only one of the enabled transitions for a given marking will be fired~\cite{Murata1989}.
+Then, the fully asynchronous update scheme in \(\mathcal{N}\) corresponds to the classical firing scheme in \(\mathcal{P}\) where only one of the enabled transitions for a given marking will be fired~\cite{Murata1989}.
 

### 2022-12-22 manuscripts/TCS2022/tcs2022.tex
New theoretical application

@@
 
-  A set of places of Petri net \(\mathcal{P}\) encoding Boolean network \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{node} of \(\mathcal{M}\).
+  A set of places of Petri net \(\mathcal{P}\) encoding Boolean network \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{node} of \(\mathcal{N}\).
   Then, a conflict-free siphon \(S\) is said to be \emph{maximal} if and only if there is no other conflict-free siphon \(S'\) such that \(S \subset S'\).
@@
 
-  Let \(m\) be a subspace of Boolean network \(\mathcal{M} = (V, F)\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) such that:
+  Let \(m\) be a subspace of Boolean network \(\mathcal{N} = (V, F)\). 
+  A \emph{mirror} of \(m\) is a set of places \(S\) in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{N}\) such that:
   \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S\] and \[\forall v \in V \setminus D_m, p_v \not \in S, \overline{p}_v \not \in S.\]
@@
 
-  From (*) and (**), we can conclude the proof.\qed%
+  From (*) and (**), we can conclude the proof.
 \end{proof}
@@
 
-  From (*) and (**), we can conclude the proof.\qed%
+  From (*) and (**), we can conclude the proof.
 \end{proof}
 
-By Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean network to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
+We here showcase a theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
+We use it to prove a property of minimal trap spaces, which has surprisingly not been formally proved.
+Specifically, all minimal trap spaces of a Boolean network are mutually disjoint.
+This property is important because we can use it to approximate the set of attractors of the Boolean network~\cite{klarner2015computing}.
+
+\begin{theorem}%
+\label{theo:separation_min_ts}
+
+  Let \(\mathcal{N} = (V, F)\) be a Boolean network.
+  For any two distinct minimal trap spaces \(m_1\) and \(m_2\) of \(\mathcal{N}\), we have that \(\mathcal{S}_{\mathcal{N}}[m_1] \cap \mathcal{S}_{\mathcal{N}}[m_2] = \emptyset\).
+
+\end{theorem}
+
+\begin{proof}
+
+  Let \(\mathcal{P}\) be the Petri net encoding of \(\mathcal{N}\).
+  If \(\mathcal{N}\) has only one minimal trap space, then the theorem trivially holds.
+  Note that by Theorem~\ref{theo:min_ts_2_max_sp}, \(\mathcal{N}\) always has at least one minimal trap space because \(\mathcal{P}\) has at least one maximal conflict-free siphon.
+  Hence, we consider the case that \(\mathcal{N}\) has at least two minimal trap spaces.
+
+  Consider two any distinct minimal trap spaces \(m_1\) and \(m_2\).
+  Assume that \(\mathcal{S}_{\mathcal{N}}[m_1] \cap \mathcal{S}_{\mathcal{N}}[m_2] \neq \emptyset\).
+  Let \(S_1\) and \(S_2\) be the mirrors of \(m_1\) and \(m_2\), respectively.
+  By Theorem~\ref{theo:min_ts_2_max_sp}, \(S_1\) and \(S_2\) are maximal conflict-free siphons of \(\mathcal{P}\).
+  We have that \(S = S_1 \cup S_2\) is also a siphon because of the property of siphons in Petri nets~\cite{DBLP:journals/isci/LiuB16}.
+  For every node \(v \in V\), assume that \(p_v \in S\) and \(\overline{p}_v \in S\) hold.
+  Since \(S_1\) and \(S_2\) are conflict-free, there are all two cases.
+  Case 1: \(p_v \in S_1\) and \(\overline{p}_v \in S_2\).
+  Case 2: \(p_v \in S_2\) and \(\overline{p}_v \in S_1\).
+  These two cases lead to \(m_1(v) \neq m_2(v), m_1(v) \neq \star, m_2(v) \neq \star\), then  \(\mathcal{S}_{\mathcal{N}}[m_1] \cap \mathcal{S}_{\mathcal{N}}[m_2] = \emptyset\).
+  This is a contradiction.
+  Hence, for every node \(v \in V\), \(p_v \in S\) and \(\overline{p}_v \in S\) cannot hold together.
+  Therefore, \(S\) is conflict-free.
+  Now, we have that \(S\) is a conflict-free siphon but \(S_1 \subset S\) or \(S_2 \subset S\) holds because \(S_1 \neq S_2\).
+  This contradicts to the maximality of \(S_1\) and \(S_2\).
+  Hence, \(\mathcal{S}_{\mathcal{N}}[m_1] \cap \mathcal{S}_{\mathcal{N}}[m_2] = \emptyset\) holds.
+
+\end{proof}
+
+As a computational application, by Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean network to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
 Note that in the case of stable states, this can be put in regard to the classical relationship between siphons and deadlocks in Petri nets.
-It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap space that might be useful for the analysis of Petri nets, but this is out of the scope of this article.
+It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap spaces that might be useful for the analysis of Petri nets, but this is out of the scope of this article.
 

### 2022-12-22 trappist/max_sat.py
max_output not used properly by max_sat

@@ -56,7 +56,7 @@ def solve_maxsat(
     # print(cnf.hard)
     rc2 = RC2(cnf)
     sol = rc2.enumerate(block=-1)  # minimal correction subsets
-    yield from islice(sol, None if max_output == 0 else 1000)
+    yield from islice(sol, None if max_output == 0 else max_output)
 
 
 def sat_to_bool(

### 2022-12-22 manuscripts/TCS2022/tcs2022.tex
Result tables

@@
 As a computational application, by Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean network to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
-Note that in the case of stable states, this can be put in regard to the classical relationship between siphons and deadlocks in Petri nets.
+Note that in the case of special types of trap spaces (e.g., fixed points), this can be put in regard to special types of siphons in Petri nets.
+See Subsection~\ref{subsec:computation_special_ts} for more discussions about many special types of trap spaces.
 It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap spaces that might be useful for the analysis of Petri nets, but this is out of the scope of this article.
@@
 It is noted that there are no existing methods specifically designed for computing maximal conflict-free siphons (even maximal siphons) of a Petri net.
-The reason might be that researchers mainly focus on minimal siphons~\cite{DBLP:journals/isci/LiuB16}.
-Hence, we here propose a new method based on Answer Set Programming (ASP)~\cite{DBLP:journals/aicom/GebserKKOSS11} for computing maximal conflict-free siphons of a Petri net.
-The details of the proposed method shall be given in the next section.
+The reason might be that researchers mainly focus on minimal generic siphons~\cite{DBLP:journals/isci/LiuB16} in the field of Petri nets. 
+Hence, we here propose several methods for computing maximal conflict-free siphons of a Petri net. 
+The details of the proposed methods shall be given in the next section.
 
@@
 
-\subsection{Constraint Satisfaction Problem}
+\subsection{Constraint satisfaction problem}
 \label{subsec:computation_csp}
 
-The following Boolean CSP directly derives from the above characterization:
+The following Boolean Constraint Satisfaction Problem (CSP) directly derives from the above characterization:
 \begin{definition}
 
-  Given a Petri net \(\mathcal{P} = (P, T, W)\) encoding a Boolean model \(\mathcal{M} = (V, F)\).
+  Given a Petri net \(\mathcal{P} = (P, T, W)\) encoding a Boolean network \(\mathcal{N} = (V, F)\).
   The CSP \(\mathcal{C}(\mathcal{P})\) is the triple \((R, D, C)\) where
@@
 This ASP rule guarantees that two places representing the same node in \(\mathcal{M}\) never belong to the same siphon of \(\mathcal{P}\), representing the conflict-freeness.
-Naturally, a Herbrand model (see, e.g,~\cite{DBLP:journals/aicom/GebserKKOSS11}) of \(\mathcal{L}\) is equivalent to a conflict-free siphon of \(\mathcal{P}\).
+Naturally, a Herbrand model (see, e.g.,~\cite{DBLP:journals/aicom/GebserKKOSS11}) of \(\mathcal{L}\) is equivalent to a conflict-free siphon of \(\mathcal{P}\).
 To guarantee that a Herbrand model is also a stable model (an answer set), we need to add to \(\mathcal{L}\) the two choice rules
@@
 
+\subsection{Computation of special types of trap spaces}
+\label{subsec:computation_special_ts}
+
 \section{Motivating example}
@@
 
-To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
-In addition, we also include the tool mpbn~\cite{Paulev2020} to the benchmarks although it only handles locally-monotonic models, whereas both Trappist and PyBoolNet can handle general models.
+To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compared it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
+In addition, we also included the tool mpbn~\cite{Paulev2020} to the benchmarks although it only handles locally-monotonic models, whereas both Trappist and PyBoolNet can handle general models.
+To obtain a more comprehensive comparison, we also used random models generated by a third-party software (i.e., BoolNet R package~\cite{mussel2010boolnet}).
 
@@
   \label{tab:pyboolnet_repo}
-  \begin{tabular}{rlrrrrrrr}
+  \begin{tabular}{rlrrrrrr}
     \toprule
-    & model & $n$ & $|M|$ & PyBoolNet & mpbn & Trappist \\ \midrule
-    \rownb & arellano\_rootstem & 9 & 4 & 0.05 & 0.01 & \textbf{0.20}\\
-    \rownb & calzone\_cellfate & 28 & 27 & 0.03 & \textbf{NM} & 0.03\\
-    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & 0.05 & 0.02 & 0.03\\
-    \rownb & davidich\_yeast & 10 & 12 & 0.04 & 0.01 & 0.02\\
-    \rownb & dinwoodie\_life & 15 & 7 & 0.03 & 0.01 & 0.01\\
-    \rownb & dinwoodie\_stomatal & 13 & 1 & 0.03 & 0.01 & 0.01\\
-    \rownb & faure\_cellcycle & 10 & 2 & 0.04 & 0.01 & 0.02\\
-    \rownb & grieco\_mapk & 53 & 18 & 0.04 & 0.02 & 0.02\\
-    \rownb & irons\_yeast & 18 & 1 & 0.05 & 0.01 & 0.02\\
-    \rownb & jaoude\_thdiff & 103 & $>1000$ & \textbf{1.44} & \textbf{0.90} & 0.09\\
-    \rownb & klamt\_tcr & 40 & 8 & 0.04 & 0.01 & 0.03\\
-    \rownb & krumsiek\_myeloid & 11 & 6 & 0.03 & 0.01 & 0.01\\
-    \rownb & multivalued & 13 & 4 & 0.03 & 0.01 & 0.01\\
-    \rownb & n12c5 & 11 & 5 & \textbf{35.16} & 0.01 & 0.02\\
-    \rownb & n3s1c1a & 2 & 2 & 0.02 & 0.01 & 0.01\\
-    \rownb & n3s1c1b & 2 & 2 & 0.02 & 0.01 & 0.01\\
-    \rownb & n5s3 & 4 & 3 & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & n6s1c2 & 5 & 3 & 0.03 & 0.01 & 0.01\\
-    \rownb & n7s3 & 6 & 3 & 0.02 & 0.01 & 0.01\\
-    \rownb & raf & 3 & 2 & 0.02 & 0.01 & 0.01\\
-    \rownb & randomnet\_n15k3 & 15 & 3 & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & randomnet\_n7k3 & 7 & 10 & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{2.14} & 0.02 & 0.02\\
-    \rownb & saadatpour\_guardcell & 13 & 1 & 0.03 & 0.01 & 0.01\\
-    \rownb & selvaggio\_emt & 56 & $>1000$ & \textbf{1.02} & \textbf{0.52} & 0.09\\
-    \rownb & tournier\_apoptosis & 12 & 3 & 0.04 & 0.01 & 0.01\\
-    \rownb & xiao\_wnt5a & 7 & 4 & 0.03 & 0.01 & 0.01\\
-    \rownb & zhang\_tlgl & 60 & 156 & 0.22 & \textbf{NM} & 0.05\\
-    \rownb & zhang\_tlgl\_v2 & 60 & 258 & 0.09 & 0.15 & 0.02\\
+    & model & $n$ & $|M|$ & LQM & PBN & mpbn & Trappist \\ \midrule
+    \rownb & arellano\_rootstem & 9 & 4 & - & 0.05 & 0.01 & \textbf{0.20}\\
+    \rownb & calzone\_cellfate & 28 & 27 & - & 0.03 & \textbf{NM} & 0.03\\
+    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & - & 0.05 & 0.02 & 0.03\\
+    \rownb & davidich\_yeast & 10 & 12 & - & 0.04 & 0.01 & 0.02\\
+    \rownb & dinwoodie\_life & 15 & 7 & - & 0.03 & 0.01 & 0.01\\
+    \rownb & dinwoodie\_stomatal & 13 & 1 & - & 0.03 & 0.01 & 0.01\\
+    \rownb & faure\_cellcycle & 10 & 2 & - & 0.04 & 0.01 & 0.02\\
+    \rownb & grieco\_mapk & 53 & 18 & - & 0.04 & 0.02 & 0.02\\
+    \rownb & irons\_yeast & 18 & 1 & - & 0.05 & 0.01 & 0.02\\
+    \rownb & jaoude\_thdiff & 103 & $>1000$ & - & \textbf{1.44} & \textbf{0.90} & 0.09\\
+    \rownb & klamt\_tcr & 40 & 8 & - & 0.04 & 0.01 & 0.03\\
+    \rownb & krumsiek\_myeloid & 11 & 6 & - & 0.03 & 0.01 & 0.01\\
+    \rownb & multivalued & 13 & 4 & - & 0.03 & 0.01 & 0.01\\
+    \rownb & n12c5 & 11 & 5 & - & \textbf{35.16} & 0.01 & 0.02\\
+    \rownb & n3s1c1a & 2 & 2 & - & 0.02 & 0.01 & 0.01\\
+    \rownb & n3s1c1b & 2 & 2 & - & 0.02 & 0.01 & 0.01\\
+    \rownb & n5s3 & 4 & 3 & - & 0.03 & \textbf{NM} & 0.01\\
+    \rownb & n6s1c2 & 5 & 3 & - & 0.03 & 0.01 & 0.01\\
+    \rownb & n7s3 & 6 & 3 & - & 0.02 & 0.01 & 0.01\\
+    \rownb & raf & 3 & 2 & - & 0.02 & 0.01 & 0.01\\
+    \rownb & randomnet\_n15k3 & 15 & 3 & - & 0.03 & \textbf{NM} & 0.01\\
+    \rownb & randomnet\_n7k3 & 7 & 10 & - & 0.03 & \textbf{NM} & 0.01\\
+    \rownb & remy\_tumorigenesis & 34 & 25 & - & \textbf{2.14} & 0.02 & 0.02\\
+    \rownb & saadatpour\_guardcell & 13 & 1 & - & 0.03 & 0.01 & 0.01\\
+    \rownb & selvaggio\_emt & 56 & $>1000$ & - & \textbf{1.02} & \textbf{0.52} & 0.09\\
+    \rownb & tournier\_apoptosis & 12 & 3 & - & 0.04 & 0.01 & 0.01\\
+    \rownb & xiao\_wnt5a & 7 & 4 & - & 0.03 & 0.01 & 0.01\\
+    \rownb & zhang\_tlgl & 60 & 156 & - & 0.22 & \textbf{NM} & 0.05\\
+    \rownb & zhang\_tlgl\_v2 & 60 & 258 & - & 0.09 & 0.15 & 0.02\\
     \bottomrule
@@
 On the contrary, on every model that was a bit challenging for PyBoolNet or mpbn, the new method is far more efficient with speedups between one and two orders of magnitude.
+
+\subsection{BBM repository}
+
+\setlength{\tabcolsep}{4pt}
+\begin{table}[!h]
+  \caption{Results on the real-world models from the BBM repository.}
+  \centering%
+  \label{tab:bbm_repo}
+  \begin{tabular}{l||r|r|r||r}
+    \toprule
+    Method & 1 min & 2 min & 3 min & Time average (s) \\ \midrule
+    PyBoolNet & - & - & - & 0.05 \\
+    mpbn & - & - & - & 0.05 \\
+    Trappist & - & - & - & 0.05 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+\setcounter{rownumber}{0}
 
@@
 
+\begin{table}[!htb]
+  \caption{Results on N-K models.}
+  \centering%
+  \label{tab:result_N_K}
+  \begin{tabular}{rrrrrrr}
+    \toprule
+    & \multicolumn{2}{c}{1 min} & \multicolumn{2}{c}{2 min} & \multicolumn{2}{c}{3 min} \\
+    \cmidrule(lr){2-7}
+    \(n\) & PBN & Trappist & PBN & Trappist & PBN & Trappist \\
+    \midrule
+    100 & - & - & - & - & - & - \\
+    150 & - & - & - & - & - & - \\
+    200 & - & - & - & - & - & - \\
+    250 & - & - & - & - & - & - \\
+    300 & - & - & - & - & - & - \\
+    350 & - & - & - & - & - & - \\
+    400 & - & - & - & - & - & - \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\begin{table}[!htb]
+  \caption{Results on scale-free models.}
+  \centering%
+  \label{tab:result_N_K}
+  \begin{tabular}{rrrrrrr}
+    \toprule
+    & \multicolumn{2}{c}{1 min} & \multicolumn{2}{c}{2 min} & \multicolumn{2}{c}{3 min} \\
+    \cmidrule(lr){2-7}
+    \(n\) & PBN & Trappist & PBN & Trappist & PBN & Trappist \\
+    \midrule
+    500 & - & - & - & - & - & - \\
+    750 & - & - & - & - & - & - \\
+    1000 & - & - & - & - & - & - \\
+    1250 & - & - & - & - & - & - \\
+    1500 & - & - & - & - & - & - \\
+    1750 & - & - & - & - & - & - \\
+    2000 & - & - & - & - & - & - \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
 \section{Conclusion}

### 2022-12-22 trappist/cp.py
Works but seems very slow…

@@ -16,6 +16,9 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 
+from dataclasses import asdict
+from datetime import timedelta
+from time import perf_counter
 from typing import Generator, List
 
 from minizinc import Instance, Model, Result, Solver, Status
@@ -26,21 +29,17 @@ import networkx as nx  # TODO maybe replace with lists/dicts
 def create_cp(petri_net: nx.DiGraph) -> Model:
     """Create the CP program for the max conflict-free siphons of petri_net."""
     model = Model()
-    vars = []
-    for node in (
-            n for n, k in petri_net.nodes(data="kind") if k == "place"
-    ):
+    model.add_string('include "globals.mzn";\n')
+    variables = []
+    for node in (n for n, k in petri_net.nodes(data="kind") if k == "place"):
         v = zincify(node)
-        vars.append(v)
+        variables.append(v)
         model.add_string(f"var bool: {v};\n")
-        print(f"var bool: {v};\n")
     for node, kind in petri_net.nodes(data="kind"):
-        print(f"node, kind: {node}, {kind}")
         if kind == "place":
             if not node.startswith("-"):
                 # conflict-freeness
                 model.add_string(f"constraint not {node} \\/ not not_{node};\n")
-                print(f"constraint not {node} \\/ not not_{node};\n")
         else:
             # it's a transition, apply siphon
             # (if one succ is true, one pred must be true)
@@ -50,9 +49,9 @@ def create_cp(petri_net: nx.DiGraph) -> Model:
                 if zsucc not in or_preds:  # optimize obvious tautologies
                     or_string = " \\/ ".join(or_preds)
                     model.add_string(f"constraint {zsucc} -> {or_string};\n")
-                    print(f"constraint {zsucc} -> {or_string};\n")
-    model.add_string(f"solve :: bool_search([{', '.join(vars)}], input_order, indomain_max) satisfy;\n")
-    print(f"solve :: bool_search([{', '.join(vars)}], input_order, indomain_max) satisfy;\n")
+    model.add_string(
+        f"solve :: bool_search([{', '.join(variables)}], input_order, indomain_max) satisfy;\n"
+    )
     return model
 
 
@@ -64,28 +63,57 @@ def zincify(variable: str) -> str:
 
 
 def solve_cp(
-        model: Model, max_output: int, time_limit: int
+    model: Model,
[…]

### 2022-12-22 trappist/trappist.py

@@ -192,7 +192,7 @@ def compute_trap_spaces(
     if method == "sat":
         solutions = get_sat_solutions(petri_net, max_output, time_limit, places)
     elif method == "cp":
-        solutions = get_cp_solutions(petri_net, max_output, time_limit, places)
+        solutions = get_cp_solutions(petri_net, max_output, time_limit, places, nprocs)
     else:
         solutions_output = get_asp_output(
             petri_net, max_output, time_limit, method, debug, nprocs

### 2022-12-22 trappist/__init__.py
bump

@@ -18,7 +18,7 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 # Don't forget to git tag v<version> && git push --tags
-version = "0.5.0"  # pylint: disable=invalid-name
+version = "0.6.0"  # pylint: disable=invalid-name
 
 
 def pnml_to_asp(name: str) -> str:

### 2022-12-22 trappist/__init__.py
hatch version minor

@@ -18,7 +18,7 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 # Don't forget to git tag v<version> && git push --tags
-version = "0.6.0"  # pylint: disable=invalid-name
+version = "0.7.0"  # pylint: disable=invalid-name
 
 
 def pnml_to_asp(name: str) -> str:

### 2022-12-22 manuscripts/TCS2022/tcs2022.tex
Updates from Overleaf

@@
     & model & $n$ & $|M|$ & LQM & PBN & mpbn & Trappist \\ \midrule
-    \rownb & arellano\_rootstem & 9 & 4 & - & 0.05 & 0.01 & \textbf{0.20}\\
-    \rownb & calzone\_cellfate & 28 & 27 & - & 0.03 & \textbf{NM} & 0.03\\
-    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & - & 0.05 & 0.02 & 0.03\\
-    \rownb & davidich\_yeast & 10 & 12 & - & 0.04 & 0.01 & 0.02\\
-    \rownb & dinwoodie\_life & 15 & 7 & - & 0.03 & 0.01 & 0.01\\
-    \rownb & dinwoodie\_stomatal & 13 & 1 & - & 0.03 & 0.01 & 0.01\\
-    \rownb & faure\_cellcycle & 10 & 2 & - & 0.04 & 0.01 & 0.02\\
-    \rownb & grieco\_mapk & 53 & 18 & - & 0.04 & 0.02 & 0.02\\
-    \rownb & irons\_yeast & 18 & 1 & - & 0.05 & 0.01 & 0.02\\
-    \rownb & jaoude\_thdiff & 103 & $>1000$ & - & \textbf{1.44} & \textbf{0.90} & 0.09\\
-    \rownb & klamt\_tcr & 40 & 8 & - & 0.04 & 0.01 & 0.03\\
-    \rownb & krumsiek\_myeloid & 11 & 6 & - & 0.03 & 0.01 & 0.01\\
-    \rownb & multivalued & 13 & 4 & - & 0.03 & 0.01 & 0.01\\
-    \rownb & n12c5 & 11 & 5 & - & \textbf{35.16} & 0.01 & 0.02\\
-    \rownb & n3s1c1a & 2 & 2 & - & 0.02 & 0.01 & 0.01\\
-    \rownb & n3s1c1b & 2 & 2 & - & 0.02 & 0.01 & 0.01\\
-    \rownb & n5s3 & 4 & 3 & - & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & n6s1c2 & 5 & 3 & - & 0.03 & 0.01 & 0.01\\
-    \rownb & n7s3 & 6 & 3 & - & 0.02 & 0.01 & 0.01\\
-    \rownb & raf & 3 & 2 & - & 0.02 & 0.01 & 0.01\\
-    \rownb & randomnet\_n15k3 & 15 & 3 & - & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & randomnet\_n7k3 & 7 & 10 & - & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & remy\_tumorigenesis & 34 & 25 & - & \textbf{2.14} & 0.02 & 0.02\\
-    \rownb & saadatpour\_guardcell & 13 & 1 & - & 0.03 & 0.01 & 0.01\\
-    \rownb & selvaggio\_emt & 56 & $>1000$ & - & \textbf{1.02} & \textbf{0.52} & 0.09\\
-    \rownb & tournier\_apoptosis & 12 & 3 & - & 0.04 & 0.01 & 0.01\\
-    \rownb & xiao\_wnt5a & 7 & 4 & - & 0.03 & 0.01 & 0.01\\
-    \rownb & zhang\_tlgl & 60 & 156 & - & 0.22 & \textbf{NM} & 0.05\\
-    \rownb & zhang\_tlgl\_v2 & 60 & 258 & - & 0.09 & 0.15 & 0.02\\
+    \rownb & arellano\_rootstem & 9 & 4 & 0.13 & 0.05 & 0.01 & \textbf{0.20}\\
+    \rownb & calzone\_cellfate & 28 & 27 & 0.12 & 0.03 & \textbf{NM} & 0.03\\
+    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & 0.11 & 0.05 & 0.02 & 0.03\\
+    \rownb & davidich\_yeast & 10 & 12 & 0.11 & 0.04 & 0.01 & 0.02\\
+    \rownb & dinwoodie\_life & 15 & 7 & 0.11 & 0.03 & 0.01 & 0.01\\
+    \rownb & dinwoodie\_stomatal & 13 & 1 & 0.10 & 0.03 & 0.01 & 0.01\\
+    \rownb & faure\_cellcycle & 10 & 2 & 0.11 & 0.04 & 0.01 & 0.02\\
+    \rownb & grieco\_mapk & 53 & 18 & 0.19 & 0.04 & 0.02 & 0.02\\
+    \rownb & irons\_yeast & 18 & 1 & 0.12 & 0.05 & 0.01 & 0.02\\
+    \rownb & jaoude\_thdiff & 103 & $>1000$ & N/A & \textbf{1.44} & \textbf{0.90} & 0.09\\
+    \rownb & klamt\_tcr & 40 & 8 & 0.11 & 0.04 & 0.01 & 0.03\\
+    \rownb & krumsiek\_myeloid & 11 & 6 & 0.10 & 0.03 & 0.01 & 0.01\\
+    \rownb & multivalued & 13 & 4 & 0.10 & 0.03 & 0.01 & 0.01\\
+    \rownb & n12c5 & 11 & 5 & 0.11 & \textbf{35.16} & 0.01 & 0.02\\
+    \rownb & n3s1c1a & 2 & 2 & 0.10 & 0.02 & 0.01 & 0.01\\
+    \rownb & n3s1c1b & 2 & 2 & 0.09 & 0.02 & 0.01 & 0.01\\
+    \rownb & n5s3 & 4 & 3 & 0.10 & 0.03 & \textbf{NM} & 0.01\\
+    \rownb & n6s1c2 & 5 & 3 & 0.10 & 0.03 & 0.01 & 0.01\\
+    \rownb & n7s3 & 6 & 3 & 0.11 & 0.02 & 0.01 & 0.01\\
+    \rownb & raf & 3 & 2 & 0.10 & 0.02 & 0.01 & 0.01\\
+    \rownb & randomnet\_n15k3 & 15 & 3 & 0.10 & 0.03 & \textbf{NM} & 0.01\\
+    \rownb & randomnet\_n7k3 & 7 & 10 & 0.10 & 0.03 & \textbf{NM} & 0.01\\
+    \rownb & remy\_tumorigenesis & 34 & 25 & 0.15 & \textbf{2.14} & 0.02 & 0.02\\
+    \rownb & saadatpour\_guardcell & 13 & 1 & 0.10 & 0.03 & 0.01 & 0.01\\
+    \rownb & selvaggio\_emt & 56 & $>1000$ & N/A & \textbf{1.02} & \textbf{0.52} & 0.09\\
+    \rownb & tournier\_apoptosis & 12 & 3 & 0.10 & 0.04 & 0.01 & 0.01\\
+    \rownb & xiao\_wnt5a & 7 & 4 & 0.10 & 0.03 & 0.01 & 0.01\\
+    \rownb & zhang\_tlgl & 60 & 156 & 0.60 & 0.22 & \textbf{NM} & 0.05\\
+    \rownb & zhang\_tlgl\_v2 & 60 & 258 & 0.64 & 0.09 & 0.15 & 0.02\\
     \bottomrule
@@
     mpbn & - & - & - & 0.05 \\
-    Trappist & - & - & - & 0.05 \\
+    Trappist-maxSAT & - & - & - & 0.05 \\
+    Trappist-CLP & - & - & - & 0.05 \\
+    Trappist-ILP & - & - & - & 0.05 \\
+    Trappist-ASP & - & - & - & 0.05 \\
     \bottomrule
@@
 
-\begin{table}[!htb]
-  \caption{Timing comparisons between PyBoolNet (PBN), mpbn and Trappist on selected models from the literature.
 Column \(n\) (resp.\ \(s\)) denotes the number of nodes (resp.\ source nodes) of each model.
@@
 ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of two minutes.
-``NM'' indicates a non-locally-monotonic model.}
+``NM'' indicates a non-locally-monotonic model.
+
+\begin{table}[!htb]
+  \caption{Timing comparisons between bioLQM (LQM), PyBoolNet (PBN), mpbn and the four variants of Trappist on selected models from the literature.}
   \centering%
   \label{tab:result_real}
-  \begin{tabular}{rlrrrrrrr}
+  \begin{tabular}{rlrrrrrrrrr}
     \toprule
-    & & & & & PBN & mpbn & \multicolumn{2}{c}{Trappist} \\
-    \cmidrule(lr){8-9}
-    & model & \(n\) & \(s\) & \(|M|\) & 1000 & 1000 & 1000 & all \\
-    \midrule
-    \rownb & inflammatory-bowel~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & NM & 1.26 & 0.69\\
-    \rownb & T-LGL-survival~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.84 & NM & 0.01 & 0.01\\
-    \rownb & butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.71 & NM & 0.01 & 0.02\\
-    \rownb & colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.20 & NM & 0.02 & 0.01\\
-    \rownb & mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & $>1000$ & 0.76 & NM & 0.01 & DNF\\
-    \rownb & IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 1.29 & NM & 0.01 & 0.01\\
-    \rownb & Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & $>1000$ & DNF & NM & 0.03 & DNF\\
-    \rownb & Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 1.04 & 0.06 & 0.08\\ \midrule
-    \rownb & adhesion-cip-migration~\cite{guberman2020boolean} & 121 & 4 & 78 & 36.80 & 0.35 & 0.27 & 0.09\\
-    \rownb & interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & $>1000$ & 10.11 & NM & 0.02 & DNF\\
-    \rownb & TCR-TLR5-signaling~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 2.06 & NM & 0.03 & 0.02\\
-    \rownb & influenza-replication~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 46.67 & NM & 0.02 & 0.02\\
-    \rownb & prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & NM & 0.16 & 0.04\\
-    \rownb & HIV-1~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & NM & 0.06 & 0.06\\
-    \rownb & fibroblasts~\cite{helikar2008emergent} & 139 & 9 & $>1000$ & DNF & NM & 0.08 & DNF\\
-    \rownb & HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & $>1000$ & 6.52 & NM & 0.02 & DNF\\
-    \rownb & kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & $>1000$ & DNF & NM & 0.19 & DNF\\
-    \rownb & virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & $>1000$ & DNF & NM & 0.03 & DNF\\
-    \rownb & immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & $>1000$ & DNF & NM & 0.08 & DNF\\
-    \rownb & RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & $>1000$ & DNF & NM & 0.02 & DNF\\
-    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & 37 & $>1000$ & 87.13 & NM & 0.03 & DNF\\
-    \rownb & er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & $>1000$ & 17.88 & NM & 0.03 & DNF\\
-    \rownb & cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 101.18 & NM & 0.40 & 0.07\\
-    \rownb & CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & 2.45 & 0.07 & 0.08\\
-    \rownb & T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & $>1000$ & 69.94 & NM & 0.05 & DNF\\
-    \rownb & ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & $>1000$ & DNF & NM & 0.24 & DNF\\
-    \rownb & macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & $>1000$ & 16.43 & NM & 0.05 & DNF\\
-    \rownb & cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & $>1000$ & 1.42 & NM & 0.08 & DNF\\
-    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & $>1000$ & DNF & NM & 0.29 & DNF\\ \midrule
-    \rownb & KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & $>1000$ & DNF & 20.76 & 3.40 & DNF\\
-    \rownb & human-network~\cite{kim2011reduction} & 1953 & 669 & $>1000$ & DNF & 23.62 & 5.93 & DNF\\
-    \rownb & SN-5~\cite{kim2013rmod} & 2746 & 829 & $>1000$ & DNF & 28.94 & DNF & DNF\\
-    \rownb & turei-2016~\cite{lee2019signal} & 4691 & 1257 & ??? & DNF & DNF & DNF & DNF\\
+    & & & & & & & \multicolumn{4}{c}{Trappist}\\
+    \cmidrule(rr){8-11}
+    & model & \(n\) & \(|M|\) & LQM & PBN & mpbn & SAT & CLP & ILP & ASP \\
+    \midrule % n < 100
+    \rownb & Arabidopsis\_thaliana & 15 & 8 & 0.15 & - & - & - & - & - & - \\
+    \rownb & Bcell & 73 & 72 & 0.21 & - & - & - & - & - & - \\
+    \rownb & Cacace\_TdevModel & 61 & 28 & 1.21 & - & NM & - & - & - & - \\
+    \rownb & Corral\_ThIL17diff & 92 & - & N/A & - & - & - & - & - & - \\
+    \rownb & DNA\_damage & 26 & - & 0.23 & - & NM & - & - & - & - \\
+    \rownb & Drosophila & 52 & - & 0.33 & - & - & - & - & - & - \\
+    \rownb & EMT & 69 & - & - & - & - & - & - & - & - \\
+    \rownb & mast\_cell & 73 & - & N/A & - & - & - & - & - & - \\
+    \rownb & FT-GRN & 23 & - & DNF & - & NM & - & - & - & - \\
+    \rownb & hedgehog & 65 & - & N/A & - & - & - & - & - & - \\
+    \rownb & metastatic & 10 & - & 0.11 & - & NM & - & - & - & - \\
+    \rownb & Pancreatic\_Cancer & 43 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Pluripotent & 36 & - & 0.37 & - & NM & - & - & - & - \\
+    \rownb & Rho-GTPases & 33 & 1 & 0.21 & - & - & - & - & - & - \\
+    \rownb & Pluripotency & 36 & - & DNF & - & NM & - & - & - & - \\
+    \rownb & p53\_high\_dna & 16 & 1 & 0.40 & - & NM & - & - & - & - \\
+    \rownb & p53\_low\_dna & 16 & 1 & 0.38 & - & NM & - & - & - & - \\
+
+    \midrule % 100 <= n < 200
+    \rownb & angiofull & 142 & - & N/A & - & - & - & - & - & - \\
+    \rownb & COVID\_19 & 111 & - & N/A & - & - & - & - & - & - \\
+    \rownb & EMT\_Mech & 136 & - & DNF & - & - & - & - & - & - \\
+    \rownb & EMT\_Mech\_TGFbeta & 150 & - & DNF & - & - & - & - & - & - \\
+    \rownb & MAPK & 181 & - & N/A & - & - & - & - & - & - \\
+    \rownb & macrophage & 136 & - & N/A & - & - & - & - & - & - \\
+    \rownb & RA\_apoptosis & 180 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Adhesion\_CIP & 121 & - & 57.33 & - & - & - & - & - & - \\
+    \rownb & angiogenesis & 141 & - & N/A & - & - & - & - & - & - \\
+
+    \midrule % 200 <= n
+    \rownb & T-cell-co-receptor & 206 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Snf1-pathway & 202 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Mycobacterium & 317 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Leishmania & 342 & - & N/A & - & - & - & - & - & - \\
+    \rownb & TcellCheckPoint & 218 & - & N/A & - & NM & - & - & - & - \\
+    \rownb & Cholocystokinin & 383 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Alzheimer & 762 & - & N/A & - & NM & - & - & - & - \\
+
     \bottomrule
@@
   \label{tab:result_N_K}
-  \begin{tabular}{rrrrrrr}
+  \begin{tabular}{rrrrrrrr}
     \toprule
-    & \multicolumn{2}{c}{1 min} & \multicolumn{2}{c}{2 min} & \multicolumn{2}{c}{3 min} \\
-    \cmidrule(lr){2-7}
-    \(n\) & PBN & Trappist & PBN & Trappist & PBN & Trappist \\
+    & & & & \multicolumn{4}{c}{Trappist}\\
+    \cmidrule(rr){5-8}
+    \(n\) & LQM & mpbn & PBN & maxSAT & CLP & ILP & ASP \\
     \midrule
-    100 & - & - & - & - & - & - \\
-    150 & - & - & - & - & - & - \\
-    200 & - & - & - & - & - & - \\
-    250 & - & - & - & - & - & - \\
-    300 & - & - & - & - & - & - \\
-    350 & - & - & - & - & - & - \\
-    400 & - & - & - & - & - & - \\
+    100 & - & - & - & - & - & - & - \\
+    150 & - & - & - & - & - & - & - \\
+    200 & - & - & - & - & - & - & - \\
+    250 & - & - & - & - & - & - & - \\
+    300 & - & - & - & - & - & - & - \\
+    350 & - & - & - & - & - & - & - \\
+    400 & - & - & - & - & - & - & - \\
     \bottomrule
@@
   \centering%
-  \label{tab:result_N_K}
-  \begin{tabular}{rrrrrrr}
+  \label{tab:result_scale_free}
+  \begin{tabular}{rrrrrrrr}
     \toprule
-    & \multicolumn{2}{c}{1 min} & \multicolumn{2}{c}{2 min} & \multicolumn{2}{c}{3 min} \\
-    \cmidrule(lr){2-7}
-    \(n\) & PBN & Trappist & PBN & Trappist & PBN & Trappist \\
+    & & & & \multicolumn{4}{c}{Trappist}\\
+    \cmidrule(rr){5-8}
+    \(n\) & LQM & mpbn & PBN & maxSAT & CLP & ILP & ASP \\
     \midrule
-    500 & - & - & - & - & - & - \\
-    750 & - & - & - & - & - & - \\
-    1000 & - & - & - & - & - & - \\
-    1250 & - & - & - & - & - & - \\
-    1500 & - & - & - & - & - & - \\
-    1750 & - & - & - & - & - & - \\
-    2000 & - & - & - & - & - & - \\
+    500 & - & - & - & - & - & - & - \\
+    750 & - & - & - & - & - & - & - \\
+    1000 & - & - & - & - & - & - & - \\
+    1250 & - & - & - & - & - & - & - \\
+    1500 & - & - & - & - & - & - & - \\
+    1750 & - & - & - & - & - & - & - \\
+    2000 & - & - & - & - & - & - & - \\
     \bottomrule

### 2022-12-22 manuscripts/TCS2022/tcs2022.tex
Update references

@@
 \begin{table}[!htb]
-  \caption{Timing comparisons between PyBoolNet, mpbn and Trappist on the PyBoolNet repository.
-    Column \(n\) denotes the number of nodes of each model.
-Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
-    A number in bold indicates a ratio greater than three compared to the best result. ``NM'' indicates a non-locally-monotonic model.}
+  \caption{Timing comparisons between bioLQM (LQM), PyBoolNet (PBN), mpbn and the four variants of Trappist on the PyBoolNet repository.}
   \centering%
   \label{tab:pyboolnet_repo}
-  \begin{tabular}{rlrrrrrr}
+  \begin{tabular}{rlrrrrrrrrr}
     \toprule
-    & model & $n$ & $|M|$ & LQM & PBN & mpbn & Trappist \\ \midrule
-    \rownb & arellano\_rootstem & 9 & 4 & 0.13 & 0.05 & 0.01 & \textbf{0.20}\\
-    \rownb & calzone\_cellfate & 28 & 27 & 0.12 & 0.03 & \textbf{NM} & 0.03\\
-    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & 0.11 & 0.05 & 0.02 & 0.03\\
-    \rownb & davidich\_yeast & 10 & 12 & 0.11 & 0.04 & 0.01 & 0.02\\
-    \rownb & dinwoodie\_life & 15 & 7 & 0.11 & 0.03 & 0.01 & 0.01\\
-    \rownb & dinwoodie\_stomatal & 13 & 1 & 0.10 & 0.03 & 0.01 & 0.01\\
-    \rownb & faure\_cellcycle & 10 & 2 & 0.11 & 0.04 & 0.01 & 0.02\\
-    \rownb & grieco\_mapk & 53 & 18 & 0.19 & 0.04 & 0.02 & 0.02\\
-    \rownb & irons\_yeast & 18 & 1 & 0.12 & 0.05 & 0.01 & 0.02\\
-    \rownb & jaoude\_thdiff & 103 & $>1000$ & N/A & \textbf{1.44} & \textbf{0.90} & 0.09\\
-    \rownb & klamt\_tcr & 40 & 8 & 0.11 & 0.04 & 0.01 & 0.03\\
-    \rownb & krumsiek\_myeloid & 11 & 6 & 0.10 & 0.03 & 0.01 & 0.01\\
-    \rownb & multivalued & 13 & 4 & 0.10 & 0.03 & 0.01 & 0.01\\
-    \rownb & n12c5 & 11 & 5 & 0.11 & \textbf{35.16} & 0.01 & 0.02\\
-    \rownb & n3s1c1a & 2 & 2 & 0.10 & 0.02 & 0.01 & 0.01\\
-    \rownb & n3s1c1b & 2 & 2 & 0.09 & 0.02 & 0.01 & 0.01\\
-    \rownb & n5s3 & 4 & 3 & 0.10 & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & n6s1c2 & 5 & 3 & 0.10 & 0.03 & 0.01 & 0.01\\
-    \rownb & n7s3 & 6 & 3 & 0.11 & 0.02 & 0.01 & 0.01\\
-    \rownb & raf & 3 & 2 & 0.10 & 0.02 & 0.01 & 0.01\\
-    \rownb & randomnet\_n15k3 & 15 & 3 & 0.10 & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & randomnet\_n7k3 & 7 & 10 & 0.10 & 0.03 & \textbf{NM} & 0.01\\
-    \rownb & remy\_tumorigenesis & 34 & 25 & 0.15 & \textbf{2.14} & 0.02 & 0.02\\
-    \rownb & saadatpour\_guardcell & 13 & 1 & 0.10 & 0.03 & 0.01 & 0.01\\
-    \rownb & selvaggio\_emt & 56 & $>1000$ & N/A & \textbf{1.02} & \textbf{0.52} & 0.09\\
-    \rownb & tournier\_apoptosis & 12 & 3 & 0.10 & 0.04 & 0.01 & 0.01\\
-    \rownb & xiao\_wnt5a & 7 & 4 & 0.10 & 0.03 & 0.01 & 0.01\\
-    \rownb & zhang\_tlgl & 60 & 156 & 0.60 & 0.22 & \textbf{NM} & 0.05\\
-    \rownb & zhang\_tlgl\_v2 & 60 & 258 & 0.64 & 0.09 & 0.15 & 0.02\\
+    & & & & & & & \multicolumn{4}{c}{Trappist}\\
+    \cmidrule(rr){8-11}
+    & model & \(n\) & \(|M|\) & LQM & PBN & mpbn & SAT & CLP & ILP & ASP \\
+    \rownb & arellano\_rootstem & 9 & 4 & \textbf{0.13} & 0.05 & 0.01 & - & - & - & 0.02\\
+    \rownb & calzone\_cellfate & 28 & 27 & \textbf{0.12} & 0.03 & \textbf{NM} & - & - & - & 0.03\\
+    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & \textbf{0.11} & 0.05 & 0.02 & - & - & - & 0.03\\
+    \rownb & davidich\_yeast & 10 & 12 & \textbf{0.11} & 0.04 & 0.01 & - & - & - & 0.03\\
+    \rownb & dinwoodie\_life & 15 & 7 & \textbf{0.11} & 0.03 & 0.01 & - & - & - & 0.03\\
+    \rownb & dinwoodie\_stomatal & 13 & 1 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
+    \rownb & faure\_cellcycle & 10 & 2 & \textbf{0.11} & 0.04 & 0.01 & - & - & - & 0.03\\
+    \rownb & grieco\_mapk & 53 & 18 & \textbf{0.19} & 0.04 & 0.02 & - & - & - & 0.05\\
+    \rownb & irons\_yeast & 18 & 1 & \textbf{0.12} & 0.05 & 0.01 & - & - & - & 0.03\\
+    \rownb & jaoude\_thdiff & 103 & $>1000$ & N/A & \textbf{1.44} & \textbf{0.90} & - & - & - & 0.20\\
+    \rownb & klamt\_tcr & 40 & 8 & 0.11 & 0.04 & 0.01 & - & - & - & 0.04\\
+    \rownb & krumsiek\_myeloid & 11 & 6 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
+    \rownb & multivalued & 13 & 4 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
+    \rownb & n12c5 & 11 & 5 & 0.11 & \textbf{35.16} & 0.01 & - & - & - & 0.04\\
+    \rownb & n3s1c1a & 2 & 2 & \textbf{0.10} & 0.02 & 0.01 & - & - & - & 0.02\\
+    \rownb & n3s1c1b & 2 & 2 & \textbf{0.09} & 0.02 & 0.01 & - & - & - & 0.01\\
+    \rownb & n5s3 & 4 & 3 & \textbf{0.10} & 0.03 & \textbf{NM} & - & - & - & 0.02\\
+    \rownb & n6s1c2 & 5 & 3 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
+    \rownb & n7s3 & 6 & 3 & \textbf{0.11} & 0.02 & 0.01 & - & - & - & 0.02\\
+    \rownb & raf & 3 & 2 & \textbf{0.10} & 0.02 & 0.01 & - & - & - & 0.02\\
+    \rownb & randomnet\_n15k3 & 15 & 3 & 0.10 & 0.03 & \textbf{NM} & - & - & - & 0.04\\
+    \rownb & randomnet\_n7k3 & 7 & 10 & \textbf{0.10} & 0.03 & \textbf{NM} & - & - & - & 0.02\\
+    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{0.15} & \textbf{2.14} & 0.02 & - & - & - & 0.06\\
+    \rownb & saadatpour\_guardcell & 13 & 1 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
+    \rownb & selvaggio\_emt & 56 & $>1000$ & N/A & \textbf{1.02} & \textbf{0.52} & - & - & - & 0.18\\
+    \rownb & tournier\_apoptosis & 12 & 3 & \textbf{0.10} & 0.04 & 0.01 & - & - & - & 0.02\\
+    \rownb & xiao\_wnt5a & 7 & 4 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
+    \rownb & zhang\_tlgl & 60 & 156 & \textbf{0.60} & 0.22 & \textbf{NM} & - & - & - & 0.09\\
+    \rownb & zhang\_tlgl\_v2 & 60 & 258 & \textbf{0.64} & 0.09 & 0.15 & - & - & - & 0.08\\
     \bottomrule
@@
 \setcounter{rownumber}{0}
+
+Column \(n\) denotes the number of nodes of each model.
+Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
+A number in bold indicates a ratio greater than three compared to the best result. ``NM'' indicates a non-locally-monotonic model.
 
@@
     Method & 1 min & 2 min & 3 min & Time average (s) \\ \midrule
-    PyBoolNet & - & - & - & 0.05 \\
-    mpbn & - & - & - & 0.05 \\
-    Trappist-maxSAT & - & - & - & 0.05 \\
-    Trappist-CLP & - & - & - & 0.05 \\
-    Trappist-ILP & - & - & - & 0.05 \\
-    Trappist-ASP & - & - & - & 0.05 \\
+    bioLQM & - & - & - & - \\
+    PyBoolNet & - & - & - & - \\
+    mpbn & - & - & - & - \\
+    Trappist-maxSAT & - & - & - & - \\
+    Trappist-CLP & - & - & - & - \\
+    Trappist-ILP & - & - & - & - \\
+    Trappist-ASP & - & - & - & - \\
     \bottomrule
@@
     \midrule % n < 100
-    \rownb & Arabidopsis\_thaliana & 15 & 8 & 0.15 & - & - & - & - & - & - \\
-    \rownb & Bcell & 73 & 72 & 0.21 & - & - & - & - & - & - \\
-    \rownb & Cacace\_TdevModel & 61 & 28 & 1.21 & - & NM & - & - & - & - \\
-    \rownb & Corral\_ThIL17diff & 92 & - & N/A & - & - & - & - & - & - \\
-    \rownb & DNA\_damage & 26 & - & 0.23 & - & NM & - & - & - & - \\
-    \rownb & Drosophila & 52 & - & 0.33 & - & - & - & - & - & - \\
-    \rownb & EMT & 69 & - & - & - & - & - & - & - & - \\
-    \rownb & mast\_cell & 73 & - & N/A & - & - & - & - & - & - \\
-    \rownb & FT-GRN & 23 & - & DNF & - & NM & - & - & - & - \\
-    \rownb & hedgehog & 65 & - & N/A & - & - & - & - & - & - \\
-    \rownb & metastatic & 10 & - & 0.11 & - & NM & - & - & - & - \\
-    \rownb & Pancreatic\_Cancer & 43 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Pluripotent & 36 & - & 0.37 & - & NM & - & - & - & - \\
-    \rownb & Rho-GTPases & 33 & 1 & 0.21 & - & - & - & - & - & - \\
-    \rownb & Pluripotency & 36 & - & DNF & - & NM & - & - & - & - \\
-    \rownb & p53\_high\_dna & 16 & 1 & 0.40 & - & NM & - & - & - & - \\
-    \rownb & p53\_low\_dna & 16 & 1 & 0.38 & - & NM & - & - & - & - \\
+    \rownb & Arabidopsis\_thaliana~\cite{DesignPrinciplesGeneNetworks} & 15 & 8 & 0.15 & - & - & - & - & - & - \\
+    \rownb & Bcell~\cite{Dutta2019} & 73 & 72 & 0.21 & - & - & - & - & - & - \\
+    \rownb & Cacace\_TdevModel~\cite{Cacace2020} & 61 & 28 & 1.21 & - & NM & - & - & - & - \\
+    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & - & N/A & - & - & - & - & - & - \\
+    \rownb & DNA\_damage~\cite{DesignPrinciplesGeneNetworks} & 26 & - & 0.23 & - & NM & - & - & - & - \\
+    \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & - & 0.33 & - & - & - & - & - & - \\
+    \rownb & EMT~\cite{Rozum2021} & 69 & - & - & - & - & - & - & - & - \\
+    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & - & N/A & - & - & - & - & - & - \\
+    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & - & DNF & - & NM & - & - & - & - \\
+    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & - & N/A & - & - & - & - & - & - \\
+    \rownb & metastatic~\cite{DesignPrinciplesGeneNetworks} & 10 & - & 0.11 & - & NM & - & - & - & - \\
+    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & - & 0.37 & - & NM & - & - & - & - \\
+    \rownb & Rho-GTPases~\cite{DesignPrinciplesGeneNetworks} & 33 & 1 & 0.21 & - & - & - & - & - & - \\
+    \rownb & Pluripotency~\cite{YachieKinoshita2018} & 36 & - & DNF & - & NM & - & - & - & - \\
+    \rownb & p53\_high\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.40 & - & NM & - & - & - & - \\
+    \rownb & p53\_low\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.38 & - & NM & - & - & - & - \\
 
     \midrule % 100 <= n < 200
-    \rownb & angiofull & 142 & - & N/A & - & - & - & - & - & - \\
-    \rownb & COVID\_19 & 111 & - & N/A & - & - & - & - & - & - \\
-    \rownb & EMT\_Mech & 136 & - & DNF & - & - & - & - & - & - \\
-    \rownb & EMT\_Mech\_TGFbeta & 150 & - & DNF & - & - & - & - & - & - \\
-    \rownb & MAPK & 181 & - & N/A & - & - & - & - & - & - \\
-    \rownb & macrophage & 136 & - & N/A & - & - & - & - & - & - \\
-    \rownb & RA\_apoptosis & 180 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Adhesion\_CIP & 121 & - & 57.33 & - & - & - & - & - & - \\
-    \rownb & angiogenesis & 141 & - & N/A & - & - & - & - & - & - \\
+    \rownb & angiofull~\cite{Weinstein2017} & 142 & - & N/A & - & - & - & - & - & - \\
+    %\rownb & COVID\_19 & 111 & - & N/A & - & - & - & - & - & - \\
+    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & - & DNF & - & - & - & - & - & - \\
+    \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & - & DNF & - & - & - & - & - & - \\
+    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & - & N/A & - & - & - & - & - & - \\
+    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & - & N/A & - & - & - & - & - & - \\
+    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & - & 57.33 & - & - & - & - & - & - \\
+    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & - & N/A & - & - & - & - & - & - \\
 
     \midrule % 200 <= n
-    \rownb & T-cell-co-receptor & 206 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Snf1-pathway & 202 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Mycobacterium & 317 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Leishmania & 342 & - & N/A & - & - & - & - & - & - \\
-    \rownb & TcellCheckPoint & 218 & - & N/A & - & NM & - & - & - & - \\
-    \rownb & Cholocystokinin & 383 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Alzheimer & 762 & - & N/A & - & NM & - & - & - & - \\
+    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & - & N/A & - & - & - & - & - & - \\
+    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & - & N/A & - & NM & - & - & - & - \\
+    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & - & N/A & - & - & - & - & - & - \\
+    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & - & N/A & - & NM & - & - & - & - \\
 

### 2022-12-23 manuscripts/TCS2022/tcs2022.tex
A computational application

@@
 First, we showcase one theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
-Second, beyond the proposed ASP method, we propose several other possible methods using MaxSAT, CLP, and ILP.
-Third, we conduct a more comprehensive benchmark on both real-world and randomly generated models to evaluate all the proposed methods.
+Second, beyond the proposed ASP method, we propose several other possible methods for computing minimal trap spaces using MaxSAT, CLP, and ILP.
+Third, we discuss in detail how to compute several special types of trap space in a Boolean network.
 Forth, regarding the implementation, we have developed a converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of BioLQM.
-Finally, we explore several complexity analysis on maximal (conflict-free) siphon problems.
+%Finally, we explore several complexity analysis on maximal (conflict-free) siphon problems.
+Finally, we conduct a more comprehensive benchmark on both real-world and randomly generated models to evaluate all the proposed methods.
 
@@
   \[\forall t\in T, S\cap succ(t)\not =\emptyset\Rightarrow S\cap pred(t)\not =\emptyset.\]
+  Note that \(\emptyset\) is trivially a siphon.
 
@@
 
-Note that \(\emptyset\) is trivially a siphon.
+Let \(pred(S) := \bigcup_{s \in S}pred(s)\) and \(succ(S) := \bigcup_{s \in S}succ(s)\).
+We have an important property~\cite{DBLP:journals/isci/LiuB16} as follows.
+
+\begin{proposition}
+\label{prop:siphon_set}
+  Let \(S\) be a siphon of a Petri net \((P, T, W)\).
+  Then \(pred(S) \subseteq succ(S)\).
+
+\end{proposition}
 
@@
 
-As a computational application, by Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean network to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
+One naturally computational application of Theorem~\ref{theo:ts_2_sp} is that we can efficiently decide whether a subspace \(m\) is a trap space.
+In PyBoolNet~\cite{klarner2017pyboolnet}, this is checked by using the percolation on the prime-implicants of the Boolean functions.
+As we have mentioned at the beginning of this article, there may be a huge number of prime-implicants.
+Hence, the checking method in~\cite{klarner2017pyboolnet} shows its limitations.
+Instead, we can first compute the mirror \(S_m\) of \(m\) in the Petri net encoding.
+Then, by Proposition~\ref{prop:siphon_set} and Theorem~\ref{theo:ts_2_sp}, we check if \(pred(S_m) \subseteq succ(S_m)\).
+Note that the above computation (also the Petri net construction) is linear in the number of transitions of the Petri net, which can be much smaller than the number of prime-implicants~\cite{chatain2014characterization}.
+
+Furthermore, by Theorem~\ref{theo:min_ts_2_max_sp}, we can reduce the problem of computing all minimal trap spaces of a Boolean network to the problem of computing all maximal conflict-free siphons of its Petri net encoding.
 Note that in the case of special types of trap spaces (e.g., fixed points), this can be put in regard to special types of siphons in Petri nets.
@@
 Suppose that \(S\) is a generic siphon of \(\mathcal{P}\).
-If a place \(p\) should belong to \(S\), then by definition all the transitions in \(pred(p)\) must belong to \(succ(S)\).
+If a place \(p\) should belong to \(S\), then by Proposition~\ref{prop:siphon_set} all the transitions in \(pred(p)\) must belong to \(succ(S)\).
 Note that \(succ(S) = \bigcup_{p \in S}succ(p)\).
@@
 
+In the field of systems biology, biologists may want to compute more special types of trap spaces beyond minimal trap spaces~\cite{klarner2017pyboolnet}.
+We shall show that our proposed methods can be easily adjusted to compute popular types of trap spaces.
+We illustrate the adjustments via the ASP-based method (see Subsection~\ref{subsec:computation_asp}), but these adjustments are completely applicable for other approaches such as MaxSAT, CSP, and ILP.
+
+First, maximal trap spaces // TODO:
+
+Second, fixed points // TODO:
+
+Third, trap spaces inside a subspace // TODO:
+
+Finally, trap spaces that intersect a subspace // TODO:
+
 \section{Motivating example}
@@
     & model & \(n\) & \(|M|\) & LQM & PBN & mpbn & SAT & CLP & ILP & ASP \\
-    \rownb & arellano\_rootstem & 9 & 4 & \textbf{0.13} & 0.05 & 0.01 & - & - & - & 0.02\\
-    \rownb & calzone\_cellfate & 28 & 27 & \textbf{0.12} & 0.03 & \textbf{NM} & - & - & - & 0.03\\
-    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & \textbf{0.11} & 0.05 & 0.02 & - & - & - & 0.03\\
-    \rownb & davidich\_yeast & 10 & 12 & \textbf{0.11} & 0.04 & 0.01 & - & - & - & 0.03\\
-    \rownb & dinwoodie\_life & 15 & 7 & \textbf{0.11} & 0.03 & 0.01 & - & - & - & 0.03\\
-    \rownb & dinwoodie\_stomatal & 13 & 1 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
-    \rownb & faure\_cellcycle & 10 & 2 & \textbf{0.11} & 0.04 & 0.01 & - & - & - & 0.03\\
-    \rownb & grieco\_mapk & 53 & 18 & \textbf{0.19} & 0.04 & 0.02 & - & - & - & 0.05\\
-    \rownb & irons\_yeast & 18 & 1 & \textbf{0.12} & 0.05 & 0.01 & - & - & - & 0.03\\
-    \rownb & jaoude\_thdiff & 103 & $>1000$ & N/A & \textbf{1.44} & \textbf{0.90} & - & - & - & 0.20\\
-    \rownb & klamt\_tcr & 40 & 8 & 0.11 & 0.04 & 0.01 & - & - & - & 0.04\\
-    \rownb & krumsiek\_myeloid & 11 & 6 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
-    \rownb & multivalued & 13 & 4 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
-    \rownb & n12c5 & 11 & 5 & 0.11 & \textbf{35.16} & 0.01 & - & - & - & 0.04\\
-    \rownb & n3s1c1a & 2 & 2 & \textbf{0.10} & 0.02 & 0.01 & - & - & - & 0.02\\
-    \rownb & n3s1c1b & 2 & 2 & \textbf{0.09} & 0.02 & 0.01 & - & - & - & 0.01\\
+    \rownb & arellano\_rootstem & 9 & 4 & \textbf{0.13} & 0.05 & - & - & - & - & 0.02\\
+    \rownb & calzone\_cellfate & 28 & 27 & \textbf{0.12} & 0.03 & - & - & - & - & 0.03\\
+    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & \textbf{0.11} & 0.06 & - & - & - & - & 0.03\\
+    \rownb & davidich\_yeast & 10 & 12 & \textbf{0.11} & 0.04 & - & - & - & - & 0.04\\
+    \rownb & dinwoodie\_life & 15 & 7 & \textbf{0.11} & 0.03 & - & - & - & - & 0.03\\
+    \rownb & dinwoodie\_stomatal & 13 & 1 & \textbf{0.10} & 0.03 & - & - & - & - & 0.02\\
+    \rownb & faure\_cellcycle & 10 & 2 & \textbf{0.11} & 0.04 & - & - & - & - & 0.05\\
+    \rownb & grieco\_mapk & 53 & 18 & \textbf{0.19} & 0.04 & - & - & - & - & 0.06\\
+    \rownb & irons\_yeast & 18 & 1 & \textbf{0.12} & 0.05 & - & - & - & - & 0.06\\
+    \rownb & jaoude\_thdiff & 103 & $>1000$ & N/A & \textbf{1.92} & - & - & - & - & 0.20\\
+    \rownb & klamt\_tcr & 40 & 8 & 0.11 & 0.04 & - & - & - & - & 0.04\\
+    \rownb & krumsiek\_myeloid & 11 & 6 & \textbf{0.10} & 0.03 & - & - & - & - & 0.02\\
+    \rownb & multivalued & 13 & 4 & \textbf{0.10} & 0.03 & - & - & - & - & 0.02\\
+    \rownb & n12c5 & 11 & 5 & 0.11 & \textbf{38.80} & - & - & - & - & 0.04\\
+    \rownb & n3s1c1a & 2 & 2 & \textbf{0.10} & 0.02 & - & - & - & - & 0.02\\
+    \rownb & n3s1c1b & 2 & 2 & \textbf{0.09} & 0.02 & - & - & - & - & 0.01\\
     \rownb & n5s3 & 4 & 3 & \textbf{0.10} & 0.03 & \textbf{NM} & - & - & - & 0.02\\
-    \rownb & n6s1c2 & 5 & 3 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
-    \rownb & n7s3 & 6 & 3 & \textbf{0.11} & 0.02 & 0.01 & - & - & - & 0.02\\
-    \rownb & raf & 3 & 2 & \textbf{0.10} & 0.02 & 0.01 & - & - & - & 0.02\\
-    \rownb & randomnet\_n15k3 & 15 & 3 & 0.10 & 0.03 & \textbf{NM} & - & - & - & 0.04\\
+    \rownb & n6s1c2 & 5 & 3 & \textbf{0.10} & 0.04 & - & - & - & - & 0.02\\
+    \rownb & n7s3 & 6 & 3 & \textbf{0.11} & 0.02 & - & - & - & - & 0.02\\
+    \rownb & raf & 3 & 2 & \textbf{0.10} & 0.02 & - & - & - & - & 0.02\\
+    \rownb & randomnet\_n15k3 & 15 & 3 & 0.10 & 0.04 & \textbf{NM} & - & - & - & 0.04\\
     \rownb & randomnet\_n7k3 & 7 & 10 & \textbf{0.10} & 0.03 & \textbf{NM} & - & - & - & 0.02\\
-    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{0.15} & \textbf{2.14} & 0.02 & - & - & - & 0.06\\
-    \rownb & saadatpour\_guardcell & 13 & 1 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
-    \rownb & selvaggio\_emt & 56 & $>1000$ & N/A & \textbf{1.02} & \textbf{0.52} & - & - & - & 0.18\\
-    \rownb & tournier\_apoptosis & 12 & 3 & \textbf{0.10} & 0.04 & 0.01 & - & - & - & 0.02\\
-    \rownb & xiao\_wnt5a & 7 & 4 & \textbf{0.10} & 0.03 & 0.01 & - & - & - & 0.02\\
-    \rownb & zhang\_tlgl & 60 & 156 & \textbf{0.60} & 0.22 & \textbf{NM} & - & - & - & 0.09\\
-    \rownb & zhang\_tlgl\_v2 & 60 & 258 & \textbf{0.64} & 0.09 & 0.15 & - & - & - & 0.08\\
+    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{0.15} & \textbf{2.05} & - & - & - & - & 0.06\\
+    \rownb & saadatpour\_guardcell & 13 & 1 & \textbf{0.10} & 0.02 & - & - & - & - & 0.02\\
+    \rownb & selvaggio\_emt & 56 & $>1000$ & N/A & \textbf{1.09} & - & - & - & - & 0.18\\
+    \rownb & tournier\_apoptosis & 12 & 3 & \textbf{0.10} & 0.03 & - & - & - & - & 0.02\\
+    \rownb & xiao\_wnt5a & 7 & 4 & \textbf{0.10} & 0.02 & - & - & - & - & 0.02\\
+    \rownb & zhang\_tlgl & 60 & 156 & \textbf{0.60} & 0.22 & - & - & - & - & 0.09\\
+    \rownb & zhang\_tlgl\_v2 & 60 & 258 & \textbf{0.64} & 0.11 & - & - & - & - & 0.08\\
     \bottomrule
@@
     \midrule % n < 100
-    \rownb & Arabidopsis\_thaliana~\cite{DesignPrinciplesGeneNetworks} & 15 & 8 & 0.15 & - & - & - & - & - & - \\
+    \rownb & Arabidopsis\_thaliana~\cite{DesignPrinciplesGeneNetworks} & 15 & 8 & 0.15 & - & NM & - & - & - & - \\
     \rownb & Bcell~\cite{Dutta2019} & 73 & 72 & 0.21 & - & - & - & - & - & - \\
     \rownb & Cacace\_TdevModel~\cite{Cacace2020} & 61 & 28 & 1.21 & - & NM & - & - & - & - \\
-    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & - & N/A & - & - & - & - & - & - \\
-    \rownb & DNA\_damage~\cite{DesignPrinciplesGeneNetworks} & 26 & - & 0.23 & - & NM & - & - & - & - \\
-    \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & - & 0.33 & - & - & - & - & - & - \\
-    \rownb & EMT~\cite{Rozum2021} & 69 & - & - & - & - & - & - & - & - \\
-    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & - & N/A & - & - & - & - & - & - \\
-    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & - & DNF & - & NM & - & - & - & - \\
-    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & - & N/A & - & - & - & - & - & - \\
-    \rownb & metastatic~\cite{DesignPrinciplesGeneNetworks} & 10 & - & 0.11 & - & NM & - & - & - & - \\
-    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & - & 0.37 & - & NM & - & - & - & - \\
-    \rownb & Rho-GTPases~\cite{DesignPrinciplesGeneNetworks} & 33 & 1 & 0.21 & - & - & - & - & - & - \\
-    \rownb & Pluripotency~\cite{YachieKinoshita2018} & 36 & - & DNF & - & NM & - & - & - & - \\
+    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & DNA\_damage~\cite{DesignPrinciplesGeneNetworks} & 26 & 16 & 0.23 & - & NM & - & - & - & - \\
+    \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & 128 & 0.33 & - & - & - & - & - & - \\
+    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & - & - & - & - & - & - & - \\
+    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & 32 & DNF & - & NM & - & - & - & - \\
+    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & metastatic~\cite{DesignPrinciplesGeneNetworks} & 10 & 4 & 0.11 & - & NM & - & - & - & - \\
+    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & 276 & 0.37 & - & NM & - & - & - & - \\
+    \rownb & Rho-GTPases~\cite{DesignPrinciplesGeneNetworks} & 33 & 2 & 0.21 & - & - & - & - & - & - \\
+    \rownb & Pluripotency~\cite{YachieKinoshita2018} & 36 & 440 & DNF & - & NM & - & - & - & - \\
     \rownb & p53\_high\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.40 & - & NM & - & - & - & - \\
@@
     \midrule % 100 <= n < 200
-    \rownb & angiofull~\cite{Weinstein2017} & 142 & - & N/A & - & - & - & - & - & - \\
+    \rownb & angiofull~\cite{Weinstein2017} & 142 & $>1000$ & N/A & - & - & - & - & - & - \\
     %\rownb & COVID\_19 & 111 & - & N/A & - & - & - & - & - & - \\
-    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & - & DNF & - & - & - & - & - & - \\
-    \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & - & DNF & - & - & - & - & - & - \\
-    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & - & N/A & - & - & - & - & - & - \\
-    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & - & N/A & - & - & - & - & - & - \\
-    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & - & 57.33 & - & - & - & - & - & - \\
-    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & - & N/A & - & - & - & - & - & - \\
+    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & 82 & DNF & - & - & - & - & - & - \\
+    \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & 492 & DNF & - & - & - & - & - & - \\
+    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & 78 & 57.33 & - & - & - & - & - & - \\
+    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $>1000$ & N/A & - & - & - & - & - & - \\
 
     \midrule % 200 <= n
-    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & - & N/A & - & - & - & - & - & - \\
-    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & - & N/A & - & NM & - & - & - & - \\
-    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & - & N/A & - & - & - & - & - & - \\
-    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & - & N/A & - & NM & - & - & - & - \\
+    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $>1000$ & N/A & - & NM & - & - & - & - \\
+    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & $>1000$ & N/A & - & NM & - & - & - & - \\
 
@@
     \midrule
-    100 & - & - & - & - & - & - & - \\
-    150 & - & - & - & - & - & - & - \\
-    200 & - & - & - & - & - & - & - \\
-    250 & - & - & - & - & - & - & - \\
-    300 & - & - & - & - & - & - & - \\
-    350 & - & - & - & - & - & - & - \\
-    400 & - & - & - & - & - & - & - \\
+    100 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    150 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    200 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    250 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    300 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    350 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    400 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
     \bottomrule
@@
     \midrule
-    500 & - & - & - & - & - & - & - \\
-    750 & - & - & - & - & - & - & - \\
-    1000 & - & - & - & - & - & - & - \\
-    1250 & - & - & - & - & - & - & - \\
-    1500 & - & - & - & - & - & - & - \\
-    1750 & - & - & - & - & - & - & - \\
-    2000 & - & - & - & - & - & - & - \\
+    500 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    750 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    1000 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    1250 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    1500 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    1750 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    2000 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
     \bottomrule

### 2022-12-23 manuscripts/TCS2022/tcs2022.tex
Computation of fixed points

@@
 Then, a transition \(t \in T\) is \emph{enabled} at a marking \(m\) if and only if \(pred(t) \subseteq m\). 
+A marking \(m\) is called a deadlock if there are no enabled transitions at \(m\).
 The firing of \(t\) leads to a new marking \(m'\) specified by \(m' = (m \backslash pred(t)) \cup succ(t)\).
@@
 
-Second, fixed points // TODO:
+Second, we consider fixed points in Boolean networks.
+Let \(s\) be a fixed point of a Boolean network \(\mathcal{N}\).
+We have a subspace \(m\) corresponding to \(s\) as follows: \(\forall v \in V, m(v) = s(v)\), i.e., all nodes are fixed in \(m\).
+Clearly, \(s\) is a trap set of \(\mathcal{N}\) regardless of the update scheme.
+Hence, \(m\) is a trap space of \(\mathcal{N}\).
+In addition, since \(\left\vert S_{\mathcal{N}}[m]\right\vert = 1\), \(m\) is also a minimal trap space.
+To compute all fixed points of \(\mathcal{N}\), we can add more constraints to the encoded ASP characterizing all conflict-free siphons (equivalently trap spaces).
+For every \(v \in V\), we add to the encoded ASP the rule
+\[
+  \verb|p-v| ; \verb|n-v|.
+\]
+that ensures that for every conflict-free siphon \(S\), it contains either \(\verb|p-v|\) or \(\verb|n-v|\) for every \(v \in V\).
+Equivalently, the trap space corresponding to \(S\) is always a fixed point.
+Now, the set of answer sets of the encoded ASP is equivalent to the set of fixed points of \(\mathcal{N}\).
+In particular, when solving the encoded ASP using an ASP solver, we do not need to use the built-in option for computing set-inclusion maximal answer sets.
+Note that we can also build another ASP characterizing all fixed points of \(\mathcal{N}\) based on the equivalence between a fixed point of \(\mathcal{N}\) and a deadlock of its Petri net encoding~\cite{chatain2014characterization}.
+This approach may give a more compact ASP.
 
@@
     \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & 128 & 0.33 & - & - & - & - & - & - \\
-    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & - & - & - & - & - & - & - \\
+    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & 39.22 & - & - & - & - & - & - \\
     \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $>1000$ & N/A & - & - & - & - & - & - \\

### 2022-12-30 trappist/bnet.py
Benchmarks

@@ -54,7 +54,7 @@ def read_bnet(fileobj: IO, method: str) -> nx.DiGraph:
     net = nx.DiGraph()
 
     # big bnets…
-    setrecursionlimit(2048)
+    setrecursionlimit(20480)
     for line in fileobj.readlines():
         if line.startswith("#") or line.startswith("targets, factors"):
             continue

### 2022-12-30 manuscripts/TCS2022/tcs2022.tex
Results on random models

@@
   \label{tab:pyboolnet_repo}
+  \scalebox{0.85}{
   \begin{tabular}{rlrrrrrrrrr}
@@
     & model & \(n\) & \(|M|\) & LQM & PBN & mpbn & SAT & CLP & ILP & ASP \\
-    \rownb & arellano\_rootstem & 9 & 4 & \textbf{0.13} & 0.05 & - & - & - & - & 0.02\\
-    \rownb & calzone\_cellfate & 28 & 27 & \textbf{0.12} & 0.03 & - & - & - & - & 0.03\\
-    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & \textbf{0.11} & 0.06 & - & - & - & - & 0.03\\
-    \rownb & davidich\_yeast & 10 & 12 & \textbf{0.11} & 0.04 & - & - & - & - & 0.04\\
-    \rownb & dinwoodie\_life & 15 & 7 & \textbf{0.11} & 0.03 & - & - & - & - & 0.03\\
-    \rownb & dinwoodie\_stomatal & 13 & 1 & \textbf{0.10} & 0.03 & - & - & - & - & 0.02\\
-    \rownb & faure\_cellcycle & 10 & 2 & \textbf{0.11} & 0.04 & - & - & - & - & 0.05\\
-    \rownb & grieco\_mapk & 53 & 18 & \textbf{0.19} & 0.04 & - & - & - & - & 0.06\\
-    \rownb & irons\_yeast & 18 & 1 & \textbf{0.12} & 0.05 & - & - & - & - & 0.06\\
-    \rownb & jaoude\_thdiff & 103 & $>1000$ & N/A & \textbf{1.92} & - & - & - & - & 0.20\\
-    \rownb & klamt\_tcr & 40 & 8 & 0.11 & 0.04 & - & - & - & - & 0.04\\
-    \rownb & krumsiek\_myeloid & 11 & 6 & \textbf{0.10} & 0.03 & - & - & - & - & 0.02\\
-    \rownb & multivalued & 13 & 4 & \textbf{0.10} & 0.03 & - & - & - & - & 0.02\\
-    \rownb & n12c5 & 11 & 5 & 0.11 & \textbf{38.80} & - & - & - & - & 0.04\\
-    \rownb & n3s1c1a & 2 & 2 & \textbf{0.10} & 0.02 & - & - & - & - & 0.02\\
-    \rownb & n3s1c1b & 2 & 2 & \textbf{0.09} & 0.02 & - & - & - & - & 0.01\\
-    \rownb & n5s3 & 4 & 3 & \textbf{0.10} & 0.03 & \textbf{NM} & - & - & - & 0.02\\
-    \rownb & n6s1c2 & 5 & 3 & \textbf{0.10} & 0.04 & - & - & - & - & 0.02\\
-    \rownb & n7s3 & 6 & 3 & \textbf{0.11} & 0.02 & - & - & - & - & 0.02\\
-    \rownb & raf & 3 & 2 & \textbf{0.10} & 0.02 & - & - & - & - & 0.02\\
-    \rownb & randomnet\_n15k3 & 15 & 3 & 0.10 & 0.04 & \textbf{NM} & - & - & - & 0.04\\
-    \rownb & randomnet\_n7k3 & 7 & 10 & \textbf{0.10} & 0.03 & \textbf{NM} & - & - & - & 0.02\\
-    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{0.15} & \textbf{2.05} & - & - & - & - & 0.06\\
-    \rownb & saadatpour\_guardcell & 13 & 1 & \textbf{0.10} & 0.02 & - & - & - & - & 0.02\\
-    \rownb & selvaggio\_emt & 56 & $>1000$ & N/A & \textbf{1.09} & - & - & - & - & 0.18\\
-    \rownb & tournier\_apoptosis & 12 & 3 & \textbf{0.10} & 0.03 & - & - & - & - & 0.02\\
-    \rownb & xiao\_wnt5a & 7 & 4 & \textbf{0.10} & 0.02 & - & - & - & - & 0.02\\
-    \rownb & zhang\_tlgl & 60 & 156 & \textbf{0.60} & 0.22 & - & - & - & - & 0.09\\
-    \rownb & zhang\_tlgl\_v2 & 60 & 258 & \textbf{0.64} & 0.11 & - & - & - & - & 0.08\\
+    \midrule
+    \rownb & arellano\_rootstem & 9 & 4 & \textbf{0.13} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & calzone\_cellfate & 28 & 27 & \textbf{0.12} & 0.02 & 0.01 & 0.01 & - & - & 0.01\\
+    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & \textbf{0.11} & 0.03 & 0.01 & 0.01 & - & - & 0.01\\
+    \rownb & davidich\_yeast & 10 & 12 & \textbf{0.11} & 0.02 & 0.01 & 0.01 & - & - & 0.01\\
+    \rownb & dinwoodie\_life & 15 & 7 & \textbf{0.11} & 0.01 & 0.00 & 0.01 & - & - & 0.01\\
+    \rownb & dinwoodie\_stomatal & 13 & 1 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & faure\_cellcycle & 10 & 2 & \textbf{0.11} & 0.02 & 0.01 & 0.01 & - & - & 0.01\\
+    \rownb & grieco\_mapk & 53 & 18 & \textbf{0.19} & 0.03 & 0.02 & 0.03 & - & - & 0.02\\
+    \rownb & irons\_yeast & 18 & 1 & \textbf{0.12} & 0.03 & 0.01 & 0.01 & - & - & 0.02\\
+    \rownb & jaoude\_thdiff & 103 & $>1000$ & N/A & \textbf{0.85} & \textbf{0.45} & \textbf{0.56} & - & - & 0.09\\
+    \rownb & klamt\_tcr & 40 & 8 & \textbf{0.11} & 0.01 & 0.01 & 0.01 & - & - & 0.02\\
+    \rownb & krumsiek\_myeloid & 11 & 6 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & multivalued & 13 & 4 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & n12c5 & 11 & 5 & \textbf{0.11} & \textbf{17.83} & 0.01 & 0.01 & - & - & 0.01\\
+    \rownb & n3s1c1a & 2 & 2 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & n3s1c1b & 2 & 2 & \textbf{0.09} & 0.02 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & n5s3 & 4 & 3 & \textbf{0.10} & 0.02 & \textbf{NM} & 0.00 & - & - & 0.01\\
+    \rownb & n6s1c2 & 5 & 3 & \textbf{0.10} & 0.02 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & n7s3 & 6 & 3 & \textbf{0.11} & 0.02 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & raf & 3 & 2 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & randomnet\_n15k3 & 15 & 3 & \textbf{0.10} & 0.02 & \textbf{NM} & 0.01 & - & - & 0.01\\
+    \rownb & randomnet\_n7k3 & 7 & 10 & \textbf{0.10} & 0.01 & \textbf{NM} & 0.00 & - & - & 0.01\\
+    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{0.15} & \textbf{0.94} & 0.02 & 0.02 & - & - & 0.02\\
+    \rownb & saadatpour\_guardcell & 13 & 1 & \textbf{0.10} & 0.06 & 0.00 & 0.00 & - & - & 0.02\\
+    \rownb & selvaggio\_emt & 56 & $>1000$ & N/A & \textbf{0.48} & \textbf{0.28} & \textbf{0.28} & - & - & 0.09\\
+    \rownb & tournier\_apoptosis & 12 & 3 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & xiao\_wnt5a & 7 & 4 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
+    \rownb & zhang\_tlgl & 60 & 156 & \textbf{0.60} & 0.09 & 0.09 & 0.07 & - & - & 0.04\\
+    \rownb & zhang\_tlgl\_v2 & 60 & 258 & \textbf{0.64} & 0.04 & 0.08 & 0.11 & - & - & 0.04\\
     \bottomrule
   \end{tabular}
+  }
 \end{table}
@@
   \label{tab:bbm_repo}
-  \begin{tabular}{l||r|r|r||r}
+  \begin{tabular}{l||r|r|r||r|r}
     \toprule
-    Method & 1 min & 2 min & 3 min & Time average (s) \\ \midrule
-    bioLQM & - & - & - & - \\
-    PyBoolNet & - & - & - & - \\
-    mpbn & - & - & - & - \\
-    Trappist-maxSAT & - & - & - & - \\
-    Trappist-CLP & - & - & - & - \\
-    Trappist-ILP & - & - & - & - \\
-    Trappist-ASP & - & - & - & - \\
+    Method & 10s & 60s & 180s & avg-mono (s) & avg-all (s) \\ \midrule
+    bioLQM & 12 (134) & 9 (134) & 9 (134) & N/A & N/A \\
+    PyBoolNet & - & - & - & - & - \\
+    mpbn & 28 (187) & 26 (187) & 26 (187) & 2.31 & N/A \\
+    Trappist-maxSAT & 1 & 1 & 1 & 1.09 & 1.01 \\
+    Trappist-CLP & - & - & - & - & - \\
+    Trappist-ILP & - & - & - & - & - \\
+    Trappist-ASP & 1 & 1 & 1 & 1.02 & 0.93 \\
     \bottomrule
@@
   \caption{Timing comparisons between bioLQM (LQM), PyBoolNet (PBN), mpbn and the four variants of Trappist on selected models from the literature.}
+  % in bold if greater than or equal to 10
   \centering%
   \label{tab:result_real}
+  \scalebox{0.8}{
   \begin{tabular}{rlrrrrrrrrr}
@@
     \midrule % n < 100
-    \rownb & Arabidopsis\_thaliana~\cite{DesignPrinciplesGeneNetworks} & 15 & 8 & 0.15 & - & NM & - & - & - & - \\
-    \rownb & Bcell~\cite{Dutta2019} & 73 & 72 & 0.21 & - & - & - & - & - & - \\
-    \rownb & Cacace\_TdevModel~\cite{Cacace2020} & 61 & 28 & 1.21 & - & NM & - & - & - & - \\
-    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & DNA\_damage~\cite{DesignPrinciplesGeneNetworks} & 26 & 16 & 0.23 & - & NM & - & - & - & - \\
-    \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & 128 & 0.33 & - & - & - & - & - & - \\
-    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & 39.22 & - & - & - & - & - & - \\
-    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & 32 & DNF & - & NM & - & - & - & - \\
-    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & metastatic~\cite{DesignPrinciplesGeneNetworks} & 10 & 4 & 0.11 & - & NM & - & - & - & - \\
-    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & 276 & 0.37 & - & NM & - & - & - & - \\
-    \rownb & Rho-GTPases~\cite{DesignPrinciplesGeneNetworks} & 33 & 2 & 0.21 & - & - & - & - & - & - \\
-    \rownb & Pluripotency~\cite{YachieKinoshita2018} & 36 & 440 & DNF & - & NM & - & - & - & - \\
-    \rownb & p53\_high\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.40 & - & NM & - & - & - & - \\
-    \rownb & p53\_low\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.38 & - & NM & - & - & - & - \\
+    \rownb & Arabidopsis\_thaliana~\cite{DesignPrinciplesGeneNetworks} & 15 & 8 & \textbf{0.10} & 0.06 & NM & 0.01 & - & - & 0.02 \\
+    \rownb & Bcell~\cite{Dutta2019} & 73 & 72 & 0.23 & 0.04 & 0.08 & 0.06 & - & - & 0.05 \\
+    \rownb & Cacace\_TdevModel~\cite{Cacace2020} & 61 & 28 & \textbf{1.29} & \textbf{5.67} & NM & 0.06 & - & - & 0.08 \\
+    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $>1000$ & N/A & \textbf{107.57} & 0.76 & 0.56 & - & - & 0.16 \\
+    \rownb & DNA\_damage~\cite{DesignPrinciplesGeneNetworks} & 26 & 16 & \textbf{0.24} & \textbf{0.33} & NM & 0.02 & - & - & 0.05 \\
+    \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & 128 & 0.33 & 0.05 & 0.07 & 0.06 & - & - & 0.05 \\
+    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & \textbf{39.22} & \textbf{1.01} & 0.20 & 0.12 & - & - & 0.05 \\
+    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $>1000$ & N/A & 0.09 & 0.55 & 0.37 & - & - & 0.15 \\
+    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & 32 & \textbf{DNF} & \textbf{DNF} & NM & 0.03 & - & - & 0.19 \\
+    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $>1000$ & N/A & \textbf{DNF} & 0.50 & 0.34 & - & - & 0.33 \\
+    \rownb & metastatic~\cite{DesignPrinciplesGeneNetworks} & 10 & 4 & \textbf{0.10} & 0.04 & NM & 0.01 & - & - & 0.02 \\
+    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $>1000$ & N/A & 0.11 & 0.36 & 0.17 & - & - & 0.06 \\
+    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & 276 & 0.37 & 0.43 & NM & 0.07 & - & - & 0.06 \\
+    \rownb & Rho-GTPases~\cite{DesignPrinciplesGeneNetworks} & 33 & 2 & 0.17 & 0.57 & \textbf{40.39} & 0.07 & - & - & 0.11 \\
+    \rownb & Pluripotency~\cite{YachieKinoshita2018} & 36 & 440 & \textbf{DNF} & \textbf{DNF} & NM & 0.16 & - & - & 0.28 \\
+    \rownb & p53\_high\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.38 & \textbf{1.76} & NM & 0.08 & - & - & 0.14 \\
+    \rownb & p53\_low\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.41 & \textbf{1.76} & NM & 0.07 & - & - & 0.14 \\
 
     \midrule % 100 <= n < 200
-    \rownb & angiofull~\cite{Weinstein2017} & 142 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & angiofull~\cite{Weinstein2017} & 142 & $>1000$ & N/A & 0.17 & 1.06 & 0.88 & - & - & 0.23 \\
     %\rownb & COVID\_19 & 111 & - & N/A & - & - & - & - & - & - \\
-    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & 82 & DNF & - & - & - & - & - & - \\
-    \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & 492 & DNF & - & - & - & - & - & - \\
-    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & 78 & 57.33 & - & - & - & - & - & - \\
-    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $>1000$ & N/A & - & - & - & - & - & - \\
+    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & 82 & \textbf{DNF} & \textbf{14.01} & 0.27 & 0.20 & - & - & 0.25 \\
+    \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & 492 & \textbf{DNF} & \textbf{11.28} & 0.78 & 0.69 & - & - & 0.35 \\
+    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & $>1000$ & N/A & \textbf{13.58} & 1.76 & 1.51 & - & - & 0.27 \\
+    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $>1000$ & N/A & 0.54 & 1.09 & 0.84 & - & - & 0.27 \\
+    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $>1000$ & N/A & \textbf{DNF} & 1.43 & 1.55 & - & - & 0.19 \\
+    \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & 78 & \textbf{56.81} & \textbf{4.25} & 0.23 & 0.17 & - & - & 0.19 \\
+    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $>1000$ & N/A & 0.16 & 1.07 & 1.06 & - & - & 0.16 \\
 
     \midrule % 200 <= n
-    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $>1000$ & N/A & - & NM & - & - & - & - \\
-    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & $>1000$ & N/A & - & - & - & - & - & - \\
-    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & $>1000$ & N/A & - & NM & - & - & - & - \\
+    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & $>1000$ & N/A & \textbf{DNF} & 1.52 & 2.26 & - & - & 0.35 \\
+    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $>1000$ & N/A & 1.13 & 1.47 & 1.43 & - & - & 0.31 \\
+    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & $>1000$ & N/A & 0.42 & 2.36 & \textbf{4.91} & - & - & 0.44 \\
+    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & $>1000$ & N/A & \textbf{DNF} & 2.56 & \textbf{5.62} & - & - & 0.46 \\
+    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $>1000$ & N/A & \textbf{4.99} & NM & 1.96 & - & - & 0.28 \\
+    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & $>1000$ & N/A & 0.36 & 2.99 & \textbf{4.81} & - & - & 0.37 \\
+    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & $>1000$ & N/A & \textbf{DNF} & NM & \textbf{18.21} & - & - & 0.79 \\
 
@@
   \end{tabular}
+  }
 \end{table}
@@
 \subsection{Randomly generated models}
+
+We randomly generated a set of N-K models~\cite{glass1973logical} with network size \(n\) in the set \{100, 150, 200, 250, 300, 350, 400\} and \(K = 3\) (i.e., each node has exactly \(K = 3\) input nodes).
+We chose N-K models because they are an useful tool for studying the dynamics of Boolean networks~\cite{glass1973logical,klarner2015computing}).
+For each network size, 50 instances were generated using the \verb|generateRandomNKNetwork| function.
+In total, we have 350 random models. 
+We then applied the compared methods to these models and recorded the numbers of failures (i.e., failed to obtain the result within the time limit) of each method.
+We considered three time limits: 10 seconds, 60 seconds, and 180 seconds.
+It is worth noting that N-K models usually have small numbers of minimal trap spaces~\cite{klarner2015computing}.
+Hence, we searched for all solutions in each model, which makes the comparison to bioLQM more comprehensive.
+In addition, each node has only 3 input nodes, i.e., the number of prime-implicants of the associated Boolean function is small.
+Hence, PyBoolNet always passed the phase of computing prime-implicants in every model even with the time limit of 10s, which enables us to compare the ASP encoding of PyBoolNet and that of Trappist.
 
@@
     \midrule
-    100 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    150 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    200 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    250 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    300 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    350 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    400 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+    100 & 50-50-50 & 50-50-50 & 0-0-0 & 0-0-0 & - & - & 0-0-0 \\
+    150 & 50-50-50 & 50-50-50 & 0-0-0 & 0-0-0 & - & - & 0-0-0 \\
+    200 & 50-50-50 & 50-50-50 & 0-0-0 & 0-0-0 & - & - & 0-0-0 \\
+    250 & 50-50-50 & 50-50-50 & 1-0-0 & 0-0-0 & - & - & 0-0-0 \\
+    300 & 50-50-50 & 50-50-50 & 10-1-0 & 6-0-0 & - & - & 0-0-0 \\
+    350 & 50-50-50 & 50-50-50 & 43-14-1 & 31-3-0 & - & - & 7-0-0 \\
+    400 & 50-50-50 & 50-50-50 & 49-44-29 & 47-26-12 & - & - & 39-8-1 \\
     \bottomrule
@@
 
-\begin{table}[!htb]
-  \caption{Results on scale-free models.}
-  \centering%
-  \label{tab:result_scale_free}
-  \begin{tabular}{rrrrrrrr}
-    \toprule
-    & & & & \multicolumn{4}{c}{Trappist}\\
-    \cmidrule(rr){5-8}
-    \(n\) & LQM & mpbn & PBN & maxSAT & CLP & ILP & ASP \\
-    \midrule
-    500 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    750 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    1000 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    1250 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    1500 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    1750 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    2000 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+% \begin{table}[!htb]
+%   \caption{Results on scale-free models.}
+%   \centering%
+%   \label{tab:result_scale_free}
+%   \begin{tabular}{rrrrrrrr}
+%     \toprule
+%     & & & & \multicolumn{4}{c}{Trappist}\\
+%     \cmidrule(rr){5-8}
+%     \(n\) & LQM & mpbn & PBN & maxSAT & CLP & ILP & ASP \\
+%     \midrule
+%     500 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+%     750 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+%     1000 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+%     1250 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+%     1500 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+%     1750 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+%     2000 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
+%     \bottomrule
+%   \end{tabular}
+% \end{table}
 

### 2022-12-30 trappist/bnet.py

@@ -54,7 +54,7 @@ def read_bnet(fileobj: IO, method: str) -> nx.DiGraph:
     net = nx.DiGraph()
 
     # big bnets…
-    setrecursionlimit(20480)
+    setrecursionlimit(2048)
     for line in fileobj.readlines():
         if line.startswith("#") or line.startswith("targets, factors"):
             continue

### 2023-01-02 manuscripts/TCS2022/tcs2022.tex
Small corrections

@@
 After some technical preliminaries (Section~\ref{sec:Preliminaries}), we expose the concrete need for such a method and detail its implementation using Answer Set Programming (Section~\ref{sec:Computation}).
-We then demonstrate its efficiency and compare it to implicant-based methods on several large Boolean networks from the literature (Section~\ref{sec:eval}).
+We then demonstrate its efficiency and compare it to implicant-based methods on several large real-world models from the literature and randomly generated models (Section~\ref{sec:eval}).
 
@@
 % Mention the extensions for the journal version
-Herein we extend our previous work in~\cite{DBLP:conf/cmsb/TrinhBHS22} as follows.
-First, we showcase one theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
-Second, beyond the proposed ASP method, we propose several other possible methods for computing minimal trap spaces using MaxSAT, CLP, and ILP.
-Third, we discuss in detail how to compute several special types of trap space in a Boolean network.
-Forth, regarding the implementation, we have developed a converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of BioLQM.
+Herein we revise and extend our previous work in~\cite{DBLP:conf/cmsb/TrinhBHS22} as follows.
+First, more formal definitions are given and the existing proofs are compacted.
+Second, we showcase one theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
+Third, beyond the proposed ASP method, we propose several other possible methods for computing minimal trap spaces using MaxSAT, CLP, and ILP.
+Forth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
+Fifth, regarding the implementation, we have developed a converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of BioLQM~\cite{DBLP:conf/cmsb/TrinhBHS22}.
 %Finally, we explore several complexity analysis on maximal (conflict-free) siphon problems.
-Finally, we conduct a more comprehensive benchmark on both real-world and randomly generated models to evaluate all the proposed methods.
+Finally, we conduct a more comprehensive benchmark on both real-world models from various sources and randomly generated models to evaluate all the proposed methods.
 
@@
 Let \(pred(S) := \bigcup_{s \in S}pred(s)\) and \(succ(S) := \bigcup_{s \in S}succ(s)\).
-We have an important property~\cite{DBLP:journals/isci/LiuB16} as follows.
+If \(S = \emptyset\), then conventionally \(pred(S) = succ(S) = \emptyset\).
+We have an important property on siphons~\cite{DBLP:journals/isci/LiuB16} as follows.
 
@@
 If a place \(p\) should belong to \(S\), then by Proposition~\ref{prop:siphon_set} all the transitions in \(pred(p)\) must belong to \(succ(S)\).
-Note that \(succ(S) = \bigcup_{p \in S}succ(p)\).
 A transition \(t\) belongs to \(succ(S)\) if and only if there is at least one place \(p'\) in \(S\) such that \(p' \in pred(t)\).

### 2023-01-02 manuscripts/TCS2022/tcs2022.tex
BDD-based method provided by bioLQM

@@
 However, this method is only applicable for \emph{locally-monotonic} Boolean networks, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean networks (i.e., including both locally-monotonic and non-locally-monotonic ones).
+In addition, the BioLQM platform also provides another method using Binary Decision Diagrams (BDDs) in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html}.
+This method avoids the prime-implicants computation as it characterizes the set of generic trap spaces of a Boolean network by a BDD, then filters this set to get the set of all minimal trap spaces.
+By this characteristic, it requires the computation of all solutions, while the ASP-based methods~\cite{klarner2015computing,Paulev2020} can start enumerating them as they are found.
+Moreover, the main issue with the BDD-based method is that the number of generic trap spaces of a Boolean network may be extremely larger than the number of minimal trap spaces of this Boolean network.
+This issue limits the efficiency of the BDD-based method.
 The study~\cite{DBLP:journals/tcs/NoualRS13} highlights the need for non-locally-monotonic Boolean networks in both biological and theoretical aspects.

### 2023-01-02 manuscripts/TCS2022/tcs2022.tex
Theorem 3.2

@@
   First, we show that if \(m\) is a minimal trap space of \(\mathcal{M}\), then \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\) (*). 
-Since \(m\) is a trap space of \(\mathcal{M}\), \(S\) is a conflict-free siphon of \(\mathcal{P}\) by Theorem~\ref{theo:ts_2_sp}.
-Assume that \(S\) is not maximal.
+  Since \(m\) is a trap space of \(\mathcal{M}\), \(S\) is a conflict-free siphon of \(\mathcal{P}\) by Theorem~\ref{theo:ts_2_sp}. Assume that \(S\) is not maximal. 
   Then, there is another conflict-free siphon \(S'\) such that \(S \subset S'\). 
   By Theorem~\ref{theo:ts_2_sp}, there is a trap space \(m'\) corresponding to \(S'\). 
-Following the definition of a mirror, \(\mathcal{S}_{\mathcal{M}}[m'] \subset \mathcal{S}_{\mathcal{M}}[m]\), thus \(m' < m\).
-This is a contradiction because \(m\) is a minimal trap space.
+  Following the definition of a mirror, \(D_m \subset D_{m'}\) and \(m(v) = m'(v), \forall v \in D_m\).
+  It follows that \(S_{\mathcal{N}}[m'] \subset S_{\mathcal{N}}[m]\), thus \(m' < m\).
+  This contradicts to the minimality of \(m\). 
   Hence, \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\).
@@
   Then, there is another trap space \(m'\) such that \(m' < m\).
-In other words, \(\mathcal{S}_{\mathcal{M}}[m'] \subset \mathcal{S}_{\mathcal{M}}[m]\).
-Let \(S'\) be the mirror of \(m'\).
+  By the definition of the partial order \(<\) on subspaces, \(\mathcal{S}_{\mathcal{N}}[m'] \subset \mathcal{S}_{\mathcal{N}}[m]\). Let \(S'\) be the mirror of \(m'\).
   \(S'\) is a conflict-free siphon by Theorem~\ref{theo:ts_2_sp}. 
-Following the definition of a mirror, \(S \subset S'\), which is a contradiction because \(S\) is a maximal conflict-free siphon.
+  Following the definition of a mirror, \(S \subset S'\), which contradicts to the maximality of \(S\).
   Hence, \(m\) is a minimal trap space of \(\mathcal{M}\).
@@
 Then, by Proposition~\ref{prop:siphon_set} and Theorem~\ref{theo:ts_2_sp}, we check if \(pred(S_m) \subseteq succ(S_m)\).
-Note that the above computation (also the Petri net construction) is linear in the number of transitions of the Petri net, which can be much smaller than the number of prime-implicants~\cite{chatain2014characterization}.
+Note that the above computation (also the Petri net construction) is quadratic in the number of transitions of the Petri net, which can be much smaller than the number of prime-implicants~\cite{chatain2014characterization}.
 
@@
 See Subsection~\ref{subsec:computation_special_ts} for more discussions about many special types of trap spaces.
-It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap spaces that might be useful for the analysis of Petri nets, but this is out of the scope of this article.
+It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap spaces that might be useful for the analysis of Petri nets, but this is out of the scope of this present article.
 

### 2023-01-02 manuscripts/TCS2022/tcs2022.tex
Finished the computation of special types of trap spaces

@@
 Now, a solution (simply an answer set) \(A \subseteq \mathcal{A}\) of \(\mathcal{L}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a trap space \(m\) of \(\mathcal{M}\). 
-The conversion from \(A\) to \(m\) is straightforward.
-If \(\verb|p-v| \in A\) then \(v \in D_m\) and \(m(v) = 0\).
+The conversion from \(A\) to \(m\) is straightforward. If \(\verb|p-v| \in A\) then \(v \in D_m\) and \(m(v) = 0\). 
 Conversely, if \(\verb|n-v| \in A\) then \(v \in D_m\) and \(m(v) = 1\). 
@@
 
-First, maximal trap spaces // TODO:
+First, the work by~\cite{Rozum2021} uses the concept of stable motifs to build the succession diagram of a Boolean network, a summary of the decisions in the network dynamics that lead to successively more restrictive nested stable motifs.
+The succession diagram is useful for control and decision making on this Boolean network.
+In particular, the proposed control methods are independent to the update scheme.
+It has been shown that a stable motif of a Boolean network is equivalent to a maximal trap space of this Boolean network~\cite{Rozum2021}.
+Hence, it is necessary to develop an efficient method for computing maximal trap spaces of a Boolean network.
+We shall show how to adjust the ASP-method presented in Subsection~\ref{subsec:computation_asp} to compute maximal trap spaces.
+
+We first provide the definition of maximal trap spaces.
+Let \(\varepsilon\) be the special trap space of \(\mathcal{N}\) where all the nodes are free.
+Of course, \(\varepsilon\) corresponds to the special conflict-free siphon \(\emptyset\).
+A trap space \(m\) is called maximal if \(m \neq \varepsilon\) and there is no other trap space \(m'\) such that \(m' \neq \varepsilon\) and \(m < m'\).
+Analogously, a conflict-free siphon \(S\) is called minimal if \(S \neq \emptyset\) and there is no other trap space \(S'\) such that \(S' \neq \emptyset\) and \(S' \subset S\).
+By using the reasoning similar to the proof of Theorem~\ref{theo:min_ts_2_max_sp}, we can easily conclude that a maximal trap space of \(\mathcal{N}\) is equivalent to a minimal conflict-free siphon of its Petri net \(\mathcal{P}\).
+Let \(\mathcal{L}\) be the ASP characterizing all siphons of \(\mathcal{P}\) (see Subsection~\ref{subsec:computation_asp}).
+Naturally, we need to exclude \(\emptyset\) from the solution space of \(\mathcal{L}\) (equivalently exclude \(\varepsilon\) from the set of trap spaces).
+To do this, we add to \(\mathcal{L}\) the ASP rule
+\[
+  \verb|p-v_1| ; \verb|n-v_1| ; \dots ; \verb|p-v_n| ; \verb|n-v_n|.
+\]
+that ensures that every answer set of \(\mathcal{L}\) cannot be empty.
+Then a set-inclusion minimal answer set of \(\mathcal{L}\) is equivalent to a minimal conflict-free siphon of \(\mathcal{P}\), thus a maximal trap space of \(\mathcal{N}\)
 
@@
 \[
-  \verb|p-v| ; \verb|n-v|.
+  \verb|p-v ; n-v.|
 \]
@@
 
-Third, trap spaces inside a subspace // TODO:
-
-Finally, trap spaces that intersect a subspace // TODO:
+Third, we consider the trap spaces intersecting a given subspace \(m^{*}\) of a Boolean network.
+A trap space \(m\) intersects \(m^{*}\) if and only if \(S_{\mathcal{N}}[m] \cap S_{\mathcal{N}}[m^{*}] \neq \emptyset\).
+It follows that for every \(v\), if \(m^{*}(v) = 0\) then \(m(v) = 0\) or \(m(v) = \star\), if \(m^{*}(v) = 1\) then \(m(v) = 1\) or \(m(v) = \star\).
+For the former case, we add to \(\mathcal{L}\) the ASP rule
+\[
+  \verb|:- n-v.|
+\]
+that ensures that \(m(v)\) cannot be 1.
+For the latter case, we add to \(\mathcal{L}\) the ASP rule
+\[
+  \verb|:- p-v.|
+\]
+that ensures that \(m(v)\) cannot be 0.
+Now \(\mathcal{L}\) characterizes all trap spaces that intersect \(m^{*}\).
+
+Finally, we consider the trap spaces that are inside a given subspace \(m^{*}\) of a Boolean network.
+We first adjust \(\mathcal{L}\) to characterize all such trap spaces.
+A trap space \(m\) is inside \(m^{*}\) if and only if \(m(v) = m^{*}(v)\) for every \(v \in D_m\).
+If \(m^{*}(v) = 0\), we add to \(\mathcal{L}\) the ASP rule
+\[
+  \verb|p-v.|
+\]
+that ensures that \(m(v) = 0\).
+If \(m^{*}(v) = 1\), we add to \(\mathcal{L}\) the ASP rule
+\[
+  \verb|n-v.|
+\]
+that ensures that \(m(v) = 1\).
+It is noted that if we want to compute maximal trap spaces inside \(m^{*}\), we need to exclude the conflict-free siphon corresponding \(m^{*}\) from the solution space.
+Specifically, we need to add to \(\mathcal{L}\) the ASP rule
+\[
+  \verb|p-v_i1| ; \verb|n-v_i1| ; \dots ; \verb|p-v_ik| ; \verb|n-v_ik|.
+\]
+where \(\{v_{i_1}, \dots, v_{i_k}\}\) is the set of free nodes of \(m^{*}\).
+This rule ensures that \(m \neq m^{*}\).
+In the case that \(m^{*} = \varepsilon\), we have all maximal trap spaces of the original Boolean network.
 

### 2023-01-04 manuscripts/TCS2022/tcs2022.tex
Theorem 3.1

@@
 However, this method is only applicable for \emph{locally-monotonic} Boolean networks, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean networks (i.e., including both locally-monotonic and non-locally-monotonic ones).
-In addition, the BioLQM platform also provides another method using Binary Decision Diagrams (BDDs) in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html}.
+In addition, the bioLQM platform also provides another method using Binary Decision Diagrams (BDDs) in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html}.
 This method avoids the prime-implicants computation as it characterizes the set of generic trap spaces of a Boolean network by a BDD, then filters this set to get the set of all minimal trap spaces.
@@
 Herein we revise and extend our previous work in~\cite{DBLP:conf/cmsb/TrinhBHS22} as follows.
-First, more formal definitions are given and the existing proofs are compacted.
+First, more formal definitions are given and the existing proofs are made more detailed.
+In particular, the updated proof provides another way to prove the independence of trap spaces of a Boolean network on its update scheme, which was originally proved in~\cite{klarner2015computing}.
 Second, we showcase one theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
@@
 Forth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
-Fifth, regarding the implementation, we have developed a converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of BioLQM~\cite{DBLP:conf/cmsb/TrinhBHS22}.
+Fifth, regarding the implementation, we have developed a converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of bioLQM~\cite{DBLP:conf/cmsb/TrinhBHS22}.
 %Finally, we explore several complexity analysis on maximal (conflict-free) siphon problems.
@@
 We denote \(\mathcal{S}_{\mathcal{N}}^{\star} = (\mathbb{B} \cup \{\star\})^n\) the set of all possible subspaces of \(\mathcal{N}\).
-Note that \(\left\vert\mathcal{S}_{\mathcal{M}}^{\star}\right\vert = 3^n\) and \(S_{\mathcal{N}} \subset \mathcal{S}_{\mathcal{N}}^{\star}\)~\cite{klarner2015computing}.
+Note that \(\left\vert\mathcal{S}_{\mathcal{N}}^{\star}\right\vert = 3^n\) and \(S_{\mathcal{N}} \subset \mathcal{S}_{\mathcal{N}}^{\star}\)~\cite{klarner2015computing}.
 
@@
 \begin{example}
-We give a Boolean network \(\mathcal{M} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
+We give a Boolean network \(\mathcal{N} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
 Herein, \(\land\), \(\lor\), and \(\neg\) denote the conjunction, disjunction, and negation logical operators, respectively.\label{example:BN}
@@
 
-Note that given a Boolean network in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the BioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
+Note that given a Boolean network in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the bioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
 This piece of software extracted from GINsim~\cite{chaouiya2012logical} and part of the CoLoMoTo\footnote{\url{http://colomoto.org/}}~\cite{naldi2015cooperative} software suite allows for easy conversion between standard formats.
@@
 This does not change the worst-case-complexity (obtaining a single DNF being exponential) but might matter a lot in practice.
-As such, we will explore how this transformation, here using BDDs in BioLQM, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.
+As such, we will explore how this transformation, here using BDDs in bioLQM, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.
 
@@
 
-  A set of places of Petri net \(\mathcal{P}\) encoding Boolean network \(\mathcal{M}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{node} of \(\mathcal{N}\).
+  A set of places of Petri net \(\mathcal{P}\) encoding Boolean network \(\mathcal{N}\) is \emph{conflict-free} if it does not contain any two places corresponding to the active and inactive states of the same \emph{node} of \(\mathcal{N}\).
   Then, a conflict-free siphon \(S\) is said to be \emph{maximal} if and only if there is no other conflict-free siphon \(S'\) such that \(S \subset S'\).
@@
 
-  Let \(\mathcal{M} = (V, F)\) be a Boolean network and \(\mathcal{P}\) be its Petri net encoding.
-A subspace \(m\) is a trap space of \(\mathcal{M}\) if and only if its mirror \(S\) is a conflict-free siphon of \(\mathcal{P}\).
+  Let \(\mathcal{N} = (V, F)\) be a Boolean network and \(\mathcal{P}\) be its Petri net encoding.
+A subspace \(m\) is a trap space of \(\mathcal{N}\) if and only if its mirror \(S\) is a conflict-free siphon of \(\mathcal{P}\).
 
@@
 
-  First, we show that if \(m\) is a trap space of \(\mathcal{M}\), then \(S\) is a conflict-free siphon of \(\mathcal{P}\) (*).
+  First, we show that if \(m\) is a trap space of \(\mathcal{N}\), then \(S\) is a conflict-free siphon of \(\mathcal{P}\) (*). 
   If \(D_m = \emptyset\), then \(S = \emptyset\) is trivially a conflict-free siphon of \(\mathcal{P}\). 
-Thus, we consider the case that \(D_m \neq \emptyset\) (resp.
-\(S \neq \emptyset\)).
+  Thus, we consider the case that \(D_m \neq \emptyset\) (resp. \(S \neq \emptyset\)).
   Assume that \(S\) is not a siphon of \(\mathcal{P}\). 
-Then, there is a transition \(t \in T\) such that \(S\cap succ(t)\not =\emptyset\) but \(S\cap pred(t)=\emptyset\).
-In other words, there is a place \(p \in S\) such that \(p \in succ(t)\) but \(p \not \in pred(t)\).
-Let \(v\) be the corresponding node in \(\mathcal{M}\) of \(p\).
-By the characterization of the encoding~\cite{chaouiya2004qualitative}, there is a directional arc from \(t\) to \(p\) and a directional arc from the complementary place of \(p\) to \(t\).
+  Then, there is a transition \(t \in T\) such that \(S \cap succ(t) \neq \emptyset\) but \(S \cap pred(t) = \emptyset\).
+  This implies that there is a place \(p \in S\) such that \(p \in succ(t)\) but \(p \not \in pred(t)\).
+  Let \(v\) be the corresponding node in \(\mathcal{N}\) of \(p\). 
+  By the characteristics of the encoding~\cite{chaouiya2004qualitative}, there is a directional arc from \(t\) to \(p\) and a directional arc from the complementary place of \(p\) to \(t\).
   Without loss of generality, we assume that \(p = p_v\), then there is a directional arc from \(t\) to \(p_v\) and a directional arc from \(\overline{p}_v\) to \(t\).
-In addition, there is also no arc or a bidirectional arc between \(t\) and another place rather than \(p_v\) and \(\overline{p}_v\).
-Thus, there is no connecting arc between \(t\) and any place in \(S \setminus \{p_v\}\) because \(S\cap pred(t)\not =\emptyset\).
-In \(\mathcal{S}_{\mathcal{M}}[m]\), a node in \(V \setminus D_m\) can receive any Boolean value.
-Hence, there is a state \(s \in \mathcal{S}_{\mathcal{M}}[m]\) such that \(m_s(p') = 1, \forall p' \in pred(t) \setminus \{\overline{p_v}\}\) where \(m_s\) is the corresponding marking in \(\mathcal{P}\) of \(s\).
-We also have \(m_s(p_v) = 0\), leading to \(m_s(\overline{p_v}) = 1\) by the characterization of the encoding~\cite{chaouiya2004qualitative}.
+  %In addition, there is also no arc or a bidirectional arc between \(t\) and another place rather than \(p_v\) and \(\overline{p}_v\).
+  We follow the following procedure to find a state \(s \in \mathcal{S}_{\mathcal{N}}[m]\) such that \(m_s(p') = 1, \forall p' \in pred(t)\) where \(m_s\) is the corresponding marking in \(\mathcal{P}\) of \(s\).
+  For every place \(p' \in pred(t)\), let \(p''\) be the complementary place of \(p'\) and \(v'\) be the corresponding node in \(\mathcal{N}\) of \(p'\) and \(p''\).
+  If \(p'' \not \in S\), then \(v' \not \in D_m\) and we can always set a Boolean value to \(s(v')\) such that \(s \in \mathcal{S}_{\mathcal{N}}[m]\) and \(m_s(p') = 1\).
+  If \(p'' \in S\), then \(v' \in D_m\) and we set \(s(v') = m(v')\).
+  In this case, if \(p' = p_{v'}\) then \(s(v') = m(v') = 1\) leading to \(m_s(p') = 1\), if \(p' = \overline{p}_{v'}\) then \(s(v') = m(v') = 0\) leading to \(m_s(p') = 1\).
+  For the remaining nodes of \(\mathcal{N}\), we can always set Boolean values to these nodes to preserve that \(s \in \mathcal{S}_{\mathcal{N}}[m]\).
+  We also have \(m_s(p_v) = 0\) by the characteristics of the encoding~\cite{chaouiya2004qualitative}.
   Now, \(t\) is enabled at marking \(m_s\).
   Its firing leads to a new marking \(m'_s\) such that \(m'_s(p_v) = 1\) and \(m'_s(\overline{p}_v) = 0\). 
-Let \(s'\) be the corresponding state in \(\mathcal{M}\) of \(m'_s\).
-Since \(m\) is a trap space of \(\mathcal{M}\), \(s' \in \mathcal{S}_{\mathcal{M}}[m]\).
-Then, \(s'(v) = m(v)\), leading to \(m'_s(p_v) = 0\), which is a contradiction.
+  Let \(s'\) be the corresponding state in \(\mathcal{N}\) of \(m'_s\).
+  We have \(s'(v) = 1\) because \(m'_s(p_v) = 1\) and \(m(v) = 0\) because \(p_v \in S\).
+  This implies that \(s' \not \in \mathcal{S}_{\mathcal{N}}[m]\).
+  For any firing scheme of \(\mathcal{P}\), the firing of \(t\) always happens.
+  Since a firing scheme of \(\mathcal{P}\) is equivalent to an update scheme of \(\mathcal{N}\), \(s\) can escape from the trap space \(m\) for any update scheme of \(\mathcal{N}\), which contradicts to the property of a trap space.
   Hence, \(S\) is a siphon of \(\mathcal{P}\).
@@
 
-  Second, we show that if \(S\) is a conflict-free siphon of \(\mathcal{P}\), then \(m\) is a trap space of \(\mathcal{M}\) (**).
-By the definition of a mirror, \(m\) is a subspace of \(\mathcal{M}\).
-Let \(s\) be an arbitrary state in \(\mathcal{S}_{\mathcal{M}}[m]\) and \(m_s\) be its corresponding marking in \(\mathcal{P}\).
-By the characterization of the encoding~\cite{chaouiya2004qualitative}, \(m_s(p) = 0, \forall p \in S\).
+  Second, we show that if \(S\) is a conflict-free siphon of \(\mathcal{P}\), then \(m\) is a trap space of \(\mathcal{N}\) (**). 
+  By the definition of a mirror, \(m\) is a subspace of \(\mathcal{N}\). 
+  Let \(s\) be an arbitrary state in \(\mathcal{S}_{\mathcal{N}}[m]\) and \(m_s\) be its corresponding marking in \(\mathcal{P}\).
+  Assume that there is a place \(p \in S\) such that \(m_s(p) = 1\).
+  Let \(v\) be the corresponding node in \(\mathcal{N}\) of \(p\).
+  Since \(p \in S\), \(v \in D_m\) and \(m(v) = s(v)\).
+  If \(p = p_{v}\), then \(m_s(p_{v}) = 1\) leading to \(m(v) = s(v) = 1\) by the characteristics of the encoding~\cite{chaouiya2004qualitative}.
+  By the definition of a mirror, \(m(v) = 0\) because \(p_{v} \in S\), which is a contradiction.
+  It is symmetric for the case that \(p = \overline{p}_{v}\).
+  Hence, \(m_s(p) = 0, \forall p \in S\).
   In any marking \(m'_s\) reachable from \(m_s\) regardless of the firing scheme of \(\mathcal{P}\), we have \(m'_s(p) = 0, \forall p \in S\) by the dynamical property on markings of a siphon~\cite{DBLP:journals/isci/LiuB16}.
-Equivalently, in any state \(s'\) reachable from \(s\) regardless of the update scheme of \(\mathcal{M}\), we have \(s'(v) = s(v) = m(v), \forall v \in D_m\).
-Then, \(s' \in \mathcal{S}_{\mathcal{M}}[m]\).
-By the definition of a trap space and the arbitrariness of \(s\), \(m\) is a trap space of \(\mathcal{M}\).
+  Let \(s'\) be the corresponding state in \(\mathcal{N}\) of \(m'_s\).
+  For every node \(v \in D_m\), we have all two cases as follows.
+  Case 1: \(p_v \in S\), then \(m'_s(p_v) = 0\), thus \(s'(v) = 0 = m(v)\).
+  Case 2: \(\overline{p}_v \in S\), then \(m'_s(\overline{p}_v) = 0\), thus \(s'(v) = 1 = m(v)\).
+  Hence, \(s'(v) = m(v)\) for every \(v \in D_m\).
+  Then, \(s' \in \mathcal{S}_{\mathcal{N}}[m]\).
+  By the definition of a trap space and the arbitrariness of \(s\), \(m\) is a trap space of \(\mathcal{N}\).
 
@@
 
+From the proof of Theorem~\ref{theo:ts_2_sp}, we can see that this theorem still holds for any update scheme of the Boolean network.
+Since the Petri net encoding of a Boolean network is independent of its update scheme and siphons are a static property of a Petri net, we can imply that trap spaces of a Boolean network are independent of its update scheme.
+Note that the original proof for this property of trap spaces (see Theorem 1 of~\cite{klarner2015computing}) only considers the two popular update schemes (i.e., synchronous and fully asynchronous).
+This exhibits the very first theoretical application of the connection between trap spaces of Boolean networks and siphons of Petri nets.
+
 \begin{theorem}%
@@
 
-  Let \(\mathcal{M}\) be a Boolean network and \(\mathcal{P}\) be its Petri net encoding.
-A subspace \(m\) is a minimal trap space of \(\mathcal{M}\) if and only if its mirror \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\).
+  Let \(\mathcal{N}\) be a Boolean network and \(\mathcal{P}\) be its Petri net encoding.
+A subspace \(m\) is a minimal trap space of \(\mathcal{N}\) if and only if its mirror \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\).
 
@@
 
-  First, we show that if \(m\) is a minimal trap space of \(\mathcal{M}\), then \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\) (*). 
-  Since \(m\) is a trap space of \(\mathcal{M}\), \(S\) is a conflict-free siphon of \(\mathcal{P}\) by Theorem~\ref{theo:ts_2_sp}. Assume that \(S\) is not maximal. 
+  First, we show that if \(m\) is a minimal trap space of \(\mathcal{N}\), then \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\) (*). 
+  Since \(m\) is a trap space of \(\mathcal{N}\), \(S\) is a conflict-free siphon of \(\mathcal{P}\) by Theorem~\ref{theo:ts_2_sp}. Assume that \(S\) is not maximal. 
   Then, there is another conflict-free siphon \(S'\) such that \(S \subset S'\). 
@@
 
-  Second, we show that if \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\), then \(m\) is a minimal trap space of \(\mathcal{M}\) (**).
-  Since \(S\) is a conflict-free siphon of \(\mathcal{P}\), \(m\) is a trap space of \(\mathcal{M}\) by Theorem~\ref{theo:ts_2_sp}. 
+  Second, we show that if \(S\) is a maximal conflict-free siphon of \(\mathcal{P}\), then \(m\) is a minimal trap space of \(\mathcal{N}\) (**).
+  Since \(S\) is a conflict-free siphon of \(\mathcal{P}\), \(m\) is a trap space of \(\mathcal{N}\) by Theorem~\ref{theo:ts_2_sp}. 
   Assume that \(m\) is not minimal.
@@
   Following the definition of a mirror, \(S \subset S'\), which contradicts to the maximality of \(S\).
-  Hence, \(m\) is a minimal trap space of \(\mathcal{M}\).
+  Hence, \(m\) is a minimal trap space of \(\mathcal{N}\).
 
@@
 for each \(v \in V\). 
-This ASP rule guarantees that two places representing the same node in \(\mathcal{M}\) never belong to the same siphon of \(\mathcal{P}\), representing the conflict-freeness.
+This ASP rule guarantees that two places representing the same node in \(\mathcal{N}\) never belong to the same siphon of \(\mathcal{P}\), representig the conflict-freeness.
 Naturally, a Herbrand model (see, e.g.,~\cite{DBLP:journals/aicom/GebserKKOSS11}) of \(\mathcal{L}\) is equivalent to a conflict-free siphon of \(\mathcal{P}\).
@@
 
-Now, a solution (simply an answer set) \(A \subseteq \mathcal{A}\) of \(\mathcal{L}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a trap space \(m\) of \(\mathcal{M}\). 
+Now, a solution (simply an answer set) \(A \subseteq \mathcal{A}\) of \(\mathcal{L}\) is equivalent to a conflict-free siphon \(S\) of \(\mathcal{P}\), thus a trap space \(m\) of \(\mathcal{N}\). 
 The conversion from \(A\) to \(m\) is straightforward. If \(\verb|p-v| \in A\) then \(v \in D_m\) and \(m(v) = 0\). 
@@
 Computing multiple answer sets is built into ASP solvers and the solving collection POTASSCO~\cite{DBLP:journals/aicom/GebserKKOSS11} also features the option to find set-inclusion maximal answer sets with respect to the set of atoms.
-Naturally, a set-inclusion maximal answer set of \(\mathcal{L}\) is equivalent to a maximal conflict-free siphon of \(\mathcal{P}\), thus a minimal trap space of \(\mathcal{M}\). 
-By using this built-in option, we can compute all the set-inclusion maximal answer sets of \(\mathcal{L}\) (resp.\ all the minimal trap spaces of \(\mathcal{M}\)) in one execution.
+Naturally, a set-inclusion maximal answer set of \(\mathcal{L}\) is equivalent to a maximal conflict-free siphon of \(\mathcal{P}\), thus a minimal trap space of \(\mathcal{N}\). 
+By using this built-in option, we can compute all the set-inclusion maximal answer sets of \(\mathcal{L}\) (resp.\ all the minimal trap spaces of \(\mathcal{N}\)) in one execution.
 
@@
 Note that on some very small models, Trappist is sometimes slower than PyBoolNet and/or mpbn, but still significantly under one second.
-Moreover, we believe that the result on the arellano\_rootstem model is caused by the cold start of the JVM for BioLQM.\@
+Moreover, we believe that the result on the arellano\_rootstem model is caused by the cold start of the JVM for bioLQM.\@
 On the contrary, on every model that was a bit challenging for PyBoolNet or mpbn, the new method is far more efficient with speedups between one and two orders of magnitude.
@@
 
-We randomly generated a set of N-K models~\cite{glass1973logical} with network size \(n\) in the set \{100, 150, 200, 250, 300, 350, 400\} and \(K = 3\) (i.e., each node has exactly \(K = 3\) input nodes).
-We chose N-K models because they are an useful tool for studying the dynamics of Boolean networks~\cite{glass1973logical,klarner2015computing}).
+We randomly generated a set of N-K models~\cite{glass1973logical} with network size \(n\) in the set \{100, 150, 200, 250, 300, 350, 400\} and \(K = 3\) (i.e., each node has exactly three input nodes).
+We chose N-K models because they are a useful tool for studying the dynamics of Boolean networks~\cite{glass1973logical,klarner2015computing}).
 For each network size, 50 instances were generated using the \verb|generateRandomNKNetwork| function.
@@
 Hence, we searched for all solutions in each model, which makes the comparison to bioLQM more comprehensive.
-In addition, each node has only 3 input nodes, i.e., the number of prime-implicants of the associated Boolean function is small.
+In addition, each node has only three input nodes, i.e., the number of prime-implicants of the associated Boolean function is small.
 Hence, PyBoolNet always passed the phase of computing prime-implicants in every model even with the time limit of 10s, which enables us to compare the ASP encoding of PyBoolNet and that of Trappist.
@@
 Though we benchmarked this new approach against state-of-the-art tools, there are many more evaluations that we plan to do in the future.
-First, the BioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
-Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
+First, the bioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
+Note that this also raises the question of replacing altogether the bioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
 The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it (in other words, the trade-off of allowing redundant constraints) has to be evaluated in depth.

### 2023-01-04 manuscripts/TCS2022/tcs2022.tex
Fix warnings
Add pages to several references

@@
 \usepackage{hyperref}
+\hypersetup{pdfauthor={Name}} % avoid warning messages from hyperref package in LaTex
 
@@
 for each \(v \in V\). 
-This ASP rule guarantees that two places representing the same node in \(\mathcal{N}\) never belong to the same siphon of \(\mathcal{P}\), representig the conflict-freeness.
+This ASP rule guarantees that two places representing the same node in \(\mathcal{N}\) never belong to the same siphon of \(\mathcal{P}\), representing the conflict-freeness.
 Naturally, a Herbrand model (see, e.g.,~\cite{DBLP:journals/aicom/GebserKKOSS11}) of \(\mathcal{L}\) is equivalent to a conflict-free siphon of \(\mathcal{P}\).

### 2023-01-05 manuscripts/TCS2022/tcs2022.tex
Complementary places

@@
 The main property of this encoding is that it is completely faithful with respect to the update scheme of the original Boolean network.
-For each node \(v\) of \(\mathcal{N}\), only transitions corresponding to \(v\) can change the current marking of \(p_v\) or \(\overline{p_v}\).
+For each node \(v\) of \(\mathcal{N}\), only transitions corresponding to \(v\) can change the current marking of \(p_v\) or \(\overline{p}_v\).
 In addition, at any marking at most one of such transitions is enabled because \(m(p_v) + m(\overline{p}_v) = 1\) holds.
@@
 Figure~\ref{fig:PN} shows the Petri net encoding of this Boolean network.
-Place \(p_{x_1}\) (resp. \(\overline{p_{x_1}}\)) in \(\mathcal{P}\) represents the activation (resp.\ the inactivation) of node \(x_1\) in \(\mathcal{N}\).
-Marking \(\{p_{x_1}, \overline{p_{x_2}}\}\) in \(\mathcal{P}\) represents state 10 in \(\mathcal{N}\).
+Place \(p_{x_1}\) (resp. \(\overline{p}_{x_1}\)) in \(\mathcal{P}\) represents the activation (resp.\ the inactivation) of node \(x_1\) in \(\mathcal{N}\).
+Marking \(\{p_{x_1}, \overline{p}_{x_2}\}\) in \(\mathcal{P}\) represents state 10 in \(\mathcal{N}\).
 Transitions \(t^{1}_{x_1}\) and \(t^{2}_{x_1}\) represent the update of node \(x_1\).
@@
 For example, the Boolean network given in Example~\ref{example:BN} has two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
-The Petri net encoding of this Boolean network has five generic siphons, \(S_1 = \emptyset\), \(S_2 = \{p_{x_1}, \overline{p_{x_1}}\}\), \(S_3 = \{p_{x_2}, \overline{p_{x_2}}\}\), \(S_4 = \{\overline{p_{x_1}}, \overline{p_{x_2}}\}\), and \(S_5 = \{p_{x_1}, \overline{p_{x_1}}, p_{x_2}, \overline{p_{x_2}}\}\).
+The Petri net encoding of this Boolean network has five generic siphons, \(S_1 = \emptyset\), \(S_2 = \{p_{x_1}, \overline{p}_{x_1}\}\), \(S_3 = \{p_{x_2}, \overline{p}_{x_2}\}\), \(S_4 = \{\overline{p}_{x_1}, \overline{p}_{x_2}\}\), and \(S_5 = \{p_{x_1}, \overline{p}_{x_1}, p_{x_2}, \overline{p}_{x_2}\}\).
 However, only \(S_1\) and \(S_4\) are conflict-free siphons and correspond to \(m_2\) and \(m_1\), respectively.
@@
 \label{eq:conflict}
-p_v \in S \Rightarrow \overline{p_v} \not \in S \wedge \overline{p_v} \in S \Rightarrow p_v \not \in S
+p_v \in S \Rightarrow \overline{p}_v \not \in S \wedge \overline{p}_v \in S \Rightarrow p_v \not \in S
 \end{equation}for each node \(v \in V\).
@@
     \item \(D(p) = \mathbb{B}\) for all \(p \in R\), i.e., the variables are Boolean,
-    \item \(C = \{\neg p_v \vee \neg \overline{p_v} = 1 \mid \forall v \in V\} \bigwedge
+    \item \(C = \{\neg p_v \vee \neg \overline{p}_v = 1 \mid \forall v \in V\} \bigwedge
 \{(p = 1 \rightarrow \bigvee_{p' \in pred(t)}p' = 1) \mid p \in P, t \in pred(p)\}\).
@@
 
-  By the former part \(\neg p_v \vee \neg \overline{p_v}\) = 1 of \(C\), the conflict-freeness is imposed because for any satisfable valuation \(r\), \(r(p_v) = r(\overline{p_v}) = 1\) is impossible for all \(v \in V\).
+  By the former part \(\neg p_v \vee \neg \overline{p}_v\) = 1 of \(C\), the conflict-freeness is imposed because for any satisfable valuation \(r\), \(r(p_v) = r(\overline{p}_v) = 1\) is impossible for all \(v \in V\).
   As shown in~\cite{nabli2016enumerating}, the latter part of \(C\) can characterize the set of all generic siphons of \(\mathcal{P}\).
@@
 One natural question is that how to use the CSP-based method for enumerating all the maximal conflict-free siphons of a Petri net encoding a Boolean model?
-Of course, the set of all conflict-free siphons of the Petri net can easily characterized by the CSP model presented in~\cite{nabli2016enumerating} along with the additional constraint \(\neg p_v \vee \neg \overline{p_v} = 1\), for each \(v \in V\), which represents the conflict-freeness.
+Of course, the set of all conflict-free siphons of the Petri net can easily characterized by the CSP model presented in~\cite{nabli2016enumerating} along with the additional constraint \(\neg p_v \vee \neg \overline{p}_v = 1\), for each \(v \in V\), which represents the conflict-freeness.
 However, the main concern is to enumerate all the \emph{maximal} ones, which is not trivial to adapt from the CSP-based method.
@@
 \[
-  (\neg p_v \vee \neg \overline{p_v}), \forall v \in V
+  (\neg p_v \vee \neg \overline{p}_v), \forall v \in V
 \]
@@
 Another possible method is to translate the characterization shown in Subsection~\ref{subsec:siphon_characterization} into the ASP \(\mathcal{L}\) as follows.
-We introduce atom \verb|p-v| (resp. \verb|n-v|) to denote place \(p_v\) (resp. \(\overline{p_v}\)), \(\forall v \in V\). 
+We introduce atom \verb|p-v| (resp. \verb|n-v|) to denote place \(p_v\) (resp. \(\overline{p}_v\)), \(\forall v \in V\). 
 The set of all atoms in \(\mathcal{L}\) is given as \(\mathcal{A} = \bigcup_{v \in V}\{\verb|p-v|, \verb|n-v|\}\).
@@
 We first show how an Integer Linear Programming (ILP) \(\mathcal{I}\) can define a set of all conflict-free siphons of the encoded Petri net \(\mathcal{P}\).
-We introduce \emph{binary} variable \verb|p-v| (resp. \verb|n-v|) to denote place \(p_v\) (resp. \(\overline{p_v}\)), \(\forall v \in V\).
+We introduce \emph{binary} variable \verb|p-v| (resp. \verb|n-v|) to denote place \(p_v\) (resp. \(\overline{p}_v\)), \(\forall v \in V\).
 The set of all binary variables in \(\mathcal{I}\) is \(\bigcup_{v \in V}\{\verb|p-v|, \verb|n-v|\}\).

### 2023-01-05 manuscripts/TCS2022/tcs2022.tex
Source nodes

@@
     Each function \(f_i\) is associated with node \(v_i\) and satisfies \(f_i \colon \mathbb{B}^{\vert IN(v_i)\vert} \mapsto \mathbb{B}\) where \(\mathbb{B} = \{0, 1\}\) and \(IN(v_i)\) denotes the set of input nodes of \(v_i\).
+    Note that a node \(v_i \in V\) is called a \emph{source} node if and only if \(f_{i} = v_i\).
   \end{itemize}
@@
 This is also true for the Alzheimer model also mentioned in that same article and originally from~\cite{ogishima2016alzpathway} (\verb|F4| file in the original supplementary material, and ``Alzheimer'' in Table~\ref{tab:result_real}), but actually not for the MAPK model for which the first trap spaces can be obtained in reasonable time.
-The current practice usually revolves then around fixing some inputs to plausible values and reducing the model accordingly.
+The current practice usually revolves then around fixing some source nodes to plausible values and reducing the model accordingly.
 While this approach makes sense, it relies on potentially arbitrary decisions, and \emph{hides away} critical modelling choices that were actually not part of the original Boolean network or even of the starting map.
@@
 
+Currently, a research group has made a great effort for building a collection (called BBM) of real-world Boolean models from various sources used in systems biology.
+It aims to be a comprehensive collection suitable for benchmarking and testing new tools and methods.
+This is released and maintained at \url{https://github.com/sybila/biodivine-boolean-models}.
+We here tested all the compared methods on this model repository.
+
 \setlength{\tabcolsep}{4pt}
@@
     bioLQM & 12 (134) & 9 (134) & 9 (134) & N/A & N/A \\
-    PyBoolNet & - & - & - & - & - \\
-    mpbn & 28 (187) & 26 (187) & 26 (187) & 2.31 & N/A \\
+    PyBoolNet & 22 & 17 & 12 & 11.00 & 13.59 \\
+    mpbn & 4 (187) & 2 (187) & 2 (187) & 2.31 & N/A \\
     Trappist-maxSAT & 1 & 1 & 1 & 1.09 & 1.01 \\
@@
 
+Table~\ref{tab:bbm_repo} shows the experimental results on the 211 real-world models from the BBM repository.
+
 \subsection{Selected models}
@@
 We then applied PyBoolNet, mpbn, and Trappist to computing minimal trap spaces of these real-world models.
-It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
+It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes in these models.
 Table~\ref{tab:result_real} shows the experimental results on those models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.

### 2023-01-05 manuscripts/TCS2022/tcs2022.tex
Running machine

@@
 Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by PyBoolNet).
-We ran all the benchmarks on an apple laptop whose environment is CPU:\@ Intel\textregistered\ Core\texttrademark\ i7 1.20GHz x 4, 16 GB DDR4 RAM, MacOS 12.3.1.
+We ran all the benchmarks on a machine whose environment is CPU: \@ Intel\textregistered\ Core\texttrademark\ i9-11950H 2.60GHz × 16, 16 GB DDR4 RAM, Ubuntu 20.04.5 LTS.
 Finally, we set a time limit of two minutes for each model.

### 2023-01-05 manuscripts/TCS2022/tcs2022.tex
Small corrections in Section "Evaluation"

@@
 \begin{table}[!htb]
-  \caption{Timing comparisons between bioLQM (LQM), PyBoolNet (PBN), mpbn and the four variants of Trappist on the PyBoolNet repository.}
+  \caption{Timing comparisons (in seconds) between bioLQM (LQM), PyBoolNet (PBN), mpbn and the four variants of Trappist on the PyBoolNet repository.}
   \centering%
@@
 
+Table~\ref{tab:pyboolnet_repo} shows the experimental results on the models from the PyBoolNet repository.
 Column \(n\) denotes the number of nodes of each model. 
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
-A number in bold indicates a ratio greater than three compared to the best result. ``NM'' indicates a non-locally-monotonic model.
+A number in bold indicates a ratio greater than three compared to the best result. 
+``NM'' indicates a non-locally-monotonic model.
 
@@
 Note that on some very small models, Trappist is sometimes slower than PyBoolNet and/or mpbn, but still significantly under one second.
-Moreover, we believe that the result on the arellano\_rootstem model is caused by the cold start of the JVM for bioLQM.\@
 On the contrary, on every model that was a bit challenging for PyBoolNet or mpbn, the new method is far more efficient with speedups between one and two orders of magnitude.
@@
 These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
-We then applied PyBoolNet, mpbn, and Trappist to computing minimal trap spaces of these real-world models.
+We then applied bioLQM, PyBoolNet, mpbn, and the four variants of Trappist to computing minimal trap spaces of these real-world models.
 It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes in these models.
-Table~\ref{tab:result_real} shows the experimental results on those models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
-
-Column \(n\) (resp.\ \(s\)) denotes the number of nodes (resp.\ source nodes) of each model.
-Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
-``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of two minutes.
-``NM'' indicates a non-locally-monotonic model.
+Table~\ref{tab:result_real} shows the experimental results on those models. 
+``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces) within the timeout of two minutes.
+In the case of bioLQM, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of bioLQM because it always requires to compute all minimal trap spaces.
+A number in bold indicates a ratio greater than or equal to 10 compared to the best result.
+Other notations are similar to those in Table~\ref{tab:pyboolnet_repo}.
+Hereafter, we analyze in detail the results with respect to minimal trap space computation.
 
 \begin{table}[!htb]
-  \caption{Timing comparisons between bioLQM (LQM), PyBoolNet (PBN), mpbn and the four variants of Trappist on selected models from the literature.}
+  \caption{Timing comparisons (in seconds) between bioLQM (LQM), PyBoolNet (PBN), mpbn and the four variants of Trappist on selected models from the literature.}
   % in bold if greater than or equal to 10

### 2023-01-06 manuscripts/TCS2022/tcs2022.tex
Evaluation section

@@
 Note however that the use of siphons for the analysis of biological models, though it is not new, has been mostly relevant to the ODE-based continuous semantics of Chemical Reaction Networks~\cite{angeli2007petri,angeli2011persistence,degrand2020graphical}.
-
 We recall here the basic definition establishing that to produce something in a siphon you must consume something from the siphon.
@@
 Since \(S_1 \subset S_4\), \(S_4\) is a maximal siphon corresponding to the minimal trap space \(m_1\).
-Hereafter, we formally prove that a maximal conflict-free siphon is equivalent to a minimal trap space.
+Hereafter, we formally prove that a (maximal) conflict-free siphon is equivalent to a (minimal) trap space.
 
@@
   By Theorem~\ref{theo:min_ts_2_max_sp}, \(S_1\) and \(S_2\) are maximal conflict-free siphons of \(\mathcal{P}\).
-  We have that \(S = S_1 \cup S_2\) is also a siphon because of the property of siphons in Petri nets~\cite{DBLP:journals/isci/LiuB16}.
+  We have that \(S = S_1 \cup S_2\) is also a siphon because of Proposition~\ref{prop:siphon_set}.
   For every node \(v \in V\), assume that \(p_v \in S\) and \(\overline{p}_v \in S\) hold.
@@
 Then, by Proposition~\ref{prop:siphon_set} and Theorem~\ref{theo:ts_2_sp}, we check if \(pred(S_m) \subseteq succ(S_m)\).
-Note that the above computation (also the Petri net construction) is quadratic in the number of transitions of the Petri net, which can be much smaller than the number of prime-implicants~\cite{chatain2014characterization}.
+Note that the above computation (also the Petri net construction) is quadratic in the number of transitions of the Petri net in worst cases, which can be much smaller than the number of prime-implicants~\cite{chatain2014characterization}.
 
@@
 Analogously, a conflict-free siphon \(S\) is called minimal if \(S \neq \emptyset\) and there is no other trap space \(S'\) such that \(S' \neq \emptyset\) and \(S' \subset S\).
-By using the reasoning similar to the proof of Theorem~\ref{theo:min_ts_2_max_sp}, we can easily conclude that a maximal trap space of \(\mathcal{N}\) is equivalent to a minimal conflict-free siphon of its Petri net \(\mathcal{P}\).
-Let \(\mathcal{L}\) be the ASP characterizing all siphons of \(\mathcal{P}\) (see Subsection~\ref{subsec:computation_asp}).
+By using the reasoning similar to the proof of Theorem~\ref{theo:min_ts_2_max_sp}, we can easily conclude that a maximal trap space of \(\mathcal{N}\) is equivalent to a minimal conflict-free siphon of its encoded Petri net \(\mathcal{P}\).
+Let \(\mathcal{L}\) be the ASP characterizing all conflict-free siphons of \(\mathcal{P}\) (see Subsection~\ref{subsec:computation_asp}).
 Naturally, we need to exclude \(\emptyset\) from the solution space of \(\mathcal{L}\) (equivalently exclude \(\varepsilon\) from the set of trap spaces).
@@
 that ensures that every answer set of \(\mathcal{L}\) cannot be empty.
-Then a set-inclusion minimal answer set of \(\mathcal{L}\) is equivalent to a minimal conflict-free siphon of \(\mathcal{P}\), thus a maximal trap space of \(\mathcal{N}\)
+Then a set-inclusion minimal answer set of \(\mathcal{L}\) is equivalent to a minimal conflict-free siphon of \(\mathcal{P}\), thus a maximal trap space of \(\mathcal{N}\).
 
@@
 We first adjust \(\mathcal{L}\) to characterize all such trap spaces.
-A trap space \(m\) is inside \(m^{*}\) if and only if \(m(v) = m^{*}(v)\) for every \(v \in D_m\).
+A trap space \(m\) is inside \(m^{*}\) if and only if \(m(v) = m^{*}(v)\) for every \(v \in D_{m^{*}}\).
 If \(m^{*}(v) = 0\), we add to \(\mathcal{L}\) the ASP rule
@@
 \section{Motivating example}
-
-% TODO Choose another motivating example because this Boolean model can be easily handled by mpbn 1.7
+\label{sec:case_study}
 
@@
 
-To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compared it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
-In addition, we also included the tool mpbn~\cite{Paulev2020} to the benchmarks although it only handles locally-monotonic models, whereas both Trappist and PyBoolNet can handle general models.
+To evaluate the performance of the newly proposed methods (implemented as a Python package named Trappist) and the state-of-the-art methods (bioLQM\footnote{\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}}, PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet}, and mpbn~\cite{Paulev2020}), we compared them on both PyBoolNet's own model repository and many real-world models from various sources in the literature.
+It is worth noting that mpbn~\cite{Paulev2020} only handles locally-monotonic models, whereas the other methods can handle general models.
 To obtain a more comprehensive comparison, we also used random models generated by a third-party software (i.e., BoolNet R package~\cite{mussel2010boolnet}).
-
-To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet}.
+As explained in Section~\ref{sec:case_study}, in our benchmarks, we only searched for the first 1000 solutions for each model.
+
+To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} and mpbn~\cite{Paulev2020}.
 Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by PyBoolNet).
 We ran all the benchmarks on a machine whose environment is CPU: \@ Intel\textregistered\ Core\texttrademark\ i9-11950H 2.60GHz × 16, 16 GB DDR4 RAM, Ubuntu 20.04.5 LTS. 
-Finally, we set a time limit of two minutes for each model.
-Note that we did not get the opportunity to run this benchmark on a proper dedicated computing workstation, the results are therefore only indicative of global trends and should not be interpreted as precise performance.
-This is also why in some rare cases finding all minimal trap spaces was faster than with the limit to 1000.
-
-For all the above tools, namely PyBoolNet, mpbn and CLINGO, we used the version available in the latest CoLoMoTo docker image tagged \texttt{2022-05-01}.
-All the models and a CoLoMoTo notebook realizing the benchmarks can be found at \url{https://github.com/soli/trap-spaces-as-siphons}.
+Finally, we set a time limit of three minutes for each model.
+
+All the models and a Jupyter notebook realizing the benchmarks can be found at~\url{https://github.com/soli/trap-spaces-as-siphons}.
 These can be run on a Docker image in the cloud by clicking the ``Binder'' button.
@@
 
-Table~\ref{tab:pyboolnet_repo} shows the experimental results on the models from the PyBoolNet repository.
+Table~\ref{tab:pyboolnet_repo} shows the experimental results on the models from the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}.
 Column \(n\) denotes the number of nodes of each model. 
@@
 ``NM'' indicates a non-locally-monotonic model.
-
-As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}, the results are comparable with all minimal trap spaces found very fast.
+There are four variants of Trappist: SAT (i.e., the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), CLP (i.e., the CLP-based method shown in Subsection~\ref{subsec:computation_csp}), ILP (i.e., the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and ASP (i.e., the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
+
+As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the PyBoolNet repository, the results are comparable with all minimal trap spaces found very fast.
 For 5 of the 29 models, mpbn did not give any answer because it recognized these models as not locally-monotonic.
@@
 It aims to be a comprehensive collection suitable for benchmarking and testing new tools and methods.
-This is released and maintained at \url{https://github.com/sybila/biodivine-boolean-models}.
+It is released and maintained at \url{https://github.com/sybila/biodivine-boolean-models}.
 We here tested all the compared methods on this model repository.
@@
   \label{tab:bbm_repo}
-  \begin{tabular}{l||r|r|r||r|r}
+  \begin{tabular}{lrrrr}
     \toprule
-    Method & 10s & 60s & 180s & avg-mono (s) & avg-all (s) \\ \midrule
-    bioLQM & 12 (134) & 9 (134) & 9 (134) & N/A & N/A \\
-    PyBoolNet & 22 & 17 & 12 & 11.00 & 13.59 \\
-    mpbn & 4 (187) & 2 (187) & 2 (187) & 2.31 & N/A \\
-    Trappist-maxSAT & 1 & 1 & 1 & 1.09 & 1.01 \\
-    Trappist-CLP & - & - & - & - & - \\
-    Trappist-ILP & - & - & - & - & - \\
-    Trappist-ASP & 1 & 1 & 1 & 1.02 & 0.93 \\
+    Method & \# failures & avg-lqm (s) & avg-mono (s) & avg-all (s) \\ \midrule
+    bioLQM & 9 (134) & 12.87 & N/A & N/A \\
+    PyBoolNet & 12 & 8.87 & 11.00 & 13.59 \\
+    mpbn & 2 (187) & N/A & 2.31 & N/A \\
+    Trappist-maxSAT & 1 & 0.03 & 1.09 & 1.01 \\
+    Trappist-CLP & - & - & - & - \\
+    Trappist-ILP & - & - & - & - \\
+    Trappist-ASP & 1 & 0.05 & 1.02 & 0.93 \\
     \bottomrule
@@
 Table~\ref{tab:bbm_repo} shows the experimental results on the 211 real-world models from the BBM repository.
+Columns 2 expresses the numbers of failures (i.e., did not finish the computation within a time limit of three minutes) of each method.
+For the case of bioLQM, we only considered the models that have at most 1000 minimal trap spaces.
+The number of such models is 134 (per all 211 models) and is denoted inside the parentheses.
+For the case of mpbn, we only considered the models that are locally-monotonic.
+The number of such models is 187 (per all 211 models) and is denoted inside the parentheses.
+Columns 3-5 express the average running time (in seconds) of each method for the models having at most 1000 minimal trap spaces, the locally-monotonic models, and all the models, respectively.
+Note that when computing the average running time, if the running time exceeds 180s, it is considered as 180s.
+From the results shown in Table~\ref{tab:bbm_repo}, we reported several observations as follows.
 
@@
 These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
+Note that these models are not included in the PyBoolNet and BBM repositories.
 We then applied bioLQM, PyBoolNet, mpbn, and the four variants of Trappist to computing minimal trap spaces of these real-world models.
@@
 In total, we have 350 random models. 
-We then applied the compared methods to these models and recorded the numbers of failures (i.e., failed to obtain the result within the time limit) of each method.
-We considered three time limits: 10 seconds, 60 seconds, and 180 seconds.
+We then applied the compared methods to these models and recorded the numbers of failures (i.e., failed to obtain the result within a time limit of three minutes) as well as the average running time (inside the parentheses) in each method for each network size \(n\).
 It is worth noting that N-K models usually have small numbers of minimal trap spaces~\cite{klarner2015computing}.
@@
 In addition, each node has only three input nodes, i.e., the number of prime-implicants of the associated Boolean function is small.
-Hence, PyBoolNet always passed the phase of computing prime-implicants in every model even with the time limit of 10s, which enables us to compare the ASP encoding of PyBoolNet and that of Trappist.
+Hence, PyBoolNet always passed the phase of computing prime-implicants in every model even within 1s, which enables us to compare the ASP encoding of PyBoolNet and that of Trappist.
 
@@
     \cmidrule(rr){5-8}
-    \(n\) & LQM & mpbn & PBN & maxSAT & CLP & ILP & ASP \\
+    \(n\) & LQM & mpbn & PBN & SAT & CLP & ILP & ASP \\
     \midrule
-    100 & 50-50-50 & 50-50-50 & 0-0-0 & 0-0-0 & - & - & 0-0-0 \\
-    150 & 50-50-50 & 50-50-50 & 0-0-0 & 0-0-0 & - & - & 0-0-0 \\
-    200 & 50-50-50 & 50-50-50 & 0-0-0 & 0-0-0 & - & - & 0-0-0 \\
-    250 & 50-50-50 & 50-50-50 & 1-0-0 & 0-0-0 & - & - & 0-0-0 \\
-    300 & 50-50-50 & 50-50-50 & 10-1-0 & 6-0-0 & - & - & 0-0-0 \\
-    350 & 50-50-50 & 50-50-50 & 43-14-1 & 31-3-0 & - & - & 7-0-0 \\
-    400 & 50-50-50 & 50-50-50 & 49-44-29 & 47-26-12 & - & - & 39-8-1 \\
+    100 & 50 ($>$ 180) & 50 (N/A) & 0 (0.07) & 0 (0.05) & - () & - () & 0 (0.09) \\
+    150 & 50 ($>$ 180) & 50 (N/A) & 0 (0.14) & 0 (0.10) & - () & - () & 0 (0.14) \\
+    200 & 50 ($>$ 180) & 50 (N/A) & 0 (0.43) & 0 (0.25) & - () & - () & 0 (0.24) \\
+    250 & 50 ($>$ 180) & 50 (N/A) & 0 (1.92) & 0 (1.04) & - () & - () & 0 (0.56) \\
+    300 & 50 ($>$ 180) & 50 (N/A) & 0 (9.68) & 0 (4.46) & - () & - () & 0 (1.83) \\
+    350 & 50 ($>$ 180) & 50 (N/A) & 1 (46.54) & 0 (20.09) & - () & - () & 0 (6.10) \\
+    400 & 50 ($>$ 180) & 50 (N/A) & 29 (144.09) & 12 (90.36) & - () & - () & 1 (33.01) \\
     \bottomrule

### 2023-01-06 trappist/cp.py
Chuffed seems more consistent

@@ -69,7 +69,7 @@ def solve_cp(
     nprocs: int,
 ) -> Generator[List[int], None, None]:
     """Run an CP solver on given WCNF and get the solutions."""
-    solver = Solver.lookup("gecode")
+    solver = Solver.lookup("chuffed")
     inst = Instance(solver, model)
     nsol = 0
     if time_limit > 0:
@@ -109,7 +109,8 @@ def solve_cp(
             model.add_string(f"constraint {or_string};")
             inst = Instance(solver, model)
         else:
-            print(result.status)
+            if result.status != Status.UNSATISFIABLE:
+                print(result.status)
             break
 
 

### 2023-01-06 trappist/cp.py
More up-to-date for argparse, plus black

@@ -77,8 +77,9 @@ def solve_cp(
     else:
         remaining = None
     # TODO handle time limit?
-    while (remaining is None or remaining > timedelta()) and \
-          (max_output == 0 or nsol < max_output):
+    while (remaining is None or remaining > timedelta()) and (
+        max_output == 0 or nsol < max_output
+    ):
         start = perf_counter()
         result = inst.solve(timeout=remaining, processes=nprocs)
         end = perf_counter()

### 2023-01-06 trappist/trappist.py

@@ -216,7 +216,7 @@ def main():
     parser.add_argument(
         "-d",
         "--debug",
-        action=argparse.store_true,
+        action="store_true",
         help="Print debugging information.",
     )
     parser.add_argument(
@@ -253,7 +253,7 @@ def main():
         default="asp",
         type=str,
         help="Solver to compute the Maximal conflict-free siphons.\n"
-             "'asp' requires clingo, 'cp' requires minizinc.",
+        "'asp' requires clingo, 'cp' requires minizinc.",
     )
     parser.add_argument(
         "infile",

### 2023-01-09 manuscripts/TCS2022/tcs2022.tex
Updates from Overleaf

@@
 
-In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, of a general Boolean network.
-It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model (Section~\ref{sec:Main_finding}).
-After some technical preliminaries (Section~\ref{sec:Preliminaries}), we expose the concrete need for such a method and detail its implementation using Answer Set Programming (Section~\ref{sec:Computation}).
-We then demonstrate its efficiency and compare it to implicant-based methods on several large real-world models from the literature and randomly generated models (Section~\ref{sec:eval}).
-
-All models used for evaluation and the implementation of the presented method are available at \url{https://github.com/soli/trap-spaces-as-siphons} and executable as a CoLoMoTo docker image.
+In this article we explore and prove for the first time a connection between trap spaces of a general Boolean network and siphons of its Petri net encoding.
+Not only having important theoretical applications in studying properties of trap spaces in Boolean networks, the connection has important applications in the trap space computation.
+Specifically, based on the connection, we propose an alternative approach to compute minimal trap spaces, and hence complex attractors, of a general Boolean network.
+It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
+We then demonstrate its efficiency and compare it to the state-of-the-art methods for computing minimal trap spaces in Boolean networks on many real-world models from various sources in the literature and randomly generated models.
+
+%After recalling several technical preliminaries (Section~\ref{sec:Preliminaries}), we expose the concrete need for such a method and detail its implementation using Answer Set Programming (Section~\ref{sec:Computation}).
+
+%All models used for evaluation and the implementation of the presented method are available at \url{https://github.com/soli/trap-spaces-as-siphons} and executable as a CoLoMoTo docker image.
 
@@
 First, more formal definitions are given and the existing proofs are made more detailed.
-In particular, the updated proof provides another way to prove the independence of trap spaces of a Boolean network on its update scheme, which was originally proved in~\cite{klarner2015computing}.
-Second, we showcase one theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
-Third, beyond the proposed ASP method, we propose several other possible methods for computing minimal trap spaces using MaxSAT, CLP, and ILP.
+In particular, an updated proof provides another way to prove the independence of trap spaces of a Boolean network on its update scheme, which was originally proved in~\cite{klarner2015computing}.
+Second, we showcase a theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
+Third, beyond the proposed ASP method implementing the alternative approach~\cite{DBLP:conf/cmsb/TrinhBHS22}, we propose several other possible methods for computing minimal trap spaces using MaxSAT, CLP, and ILP.
 Forth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
-Fifth, regarding the implementation, we have developed a converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of bioLQM~\cite{DBLP:conf/cmsb/TrinhBHS22}.
-%Finally, we explore several complexity analysis on maximal (conflict-free) siphon problems.
-Finally, we conduct a more comprehensive benchmark on both real-world models from various sources and randomly generated models to evaluate all the proposed methods.
+Fifth, regarding the implementation, we have developed a new converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of bioLQM~\cite{DBLP:conf/cmsb/TrinhBHS22}.
+Finally, we conduct a more comprehensive benchmark on more real-world models from various sources and randomly generated models to evaluate all the proposed methods (the benchmark conducted in~\cite{DBLP:conf/cmsb/TrinhBHS22} considers only real-world models).
+
+%%% Paper structure %%%
+The rest of this paper is organized as follows: Section~\ref{sec:Preliminaries} recalls the basic concepts including Boolean networks, attractors, trap spaces, Petri nets, and siphons.
+Section~\ref{sec:Main_finding} presents the main finding, the connection between trap spaces in Boolean networks and siphons in Petri nets.
+Section~\ref{sec:Computation} presents the alternative approach for computing minimal trap spaces and the four possible methods implementing it.
+Section~\ref{sec:case_study} shows an important biological case study showing the applicability of the new approach.
+Section~\ref{sec:eval} reports the experimental results for evaluating the efficiency of the proposed methods.
+Finally, Section~\ref{sec:Conclusion} concludes the paper and draws future work.
 
@@
 We shall briefly recall here some preliminaries on Boolean networks related to trap spaces and Petri nets.
-In the case of multi-level logical models, an encoding into a Boolean network is always possible~\cite{Didier2011}.
+%In the case of multi-level logical models, an encoding into a Boolean network is always possible~\cite{Didier2011}.
+\vangiang{Remove this statement because there is not sure if the encoded Boolean network preserves the trap spaces of the original multi-level logical model.}
 
@@
 where \(v(t)\) is the state of \(\mathcal{N}\) at time \(t\) and \(v_i(t + 1)\) is the state of node \(v_i\) at time \(t + 1\).
-Note that for simplicity, we write \(f_i(v(t))\) even \(IN(v_i) \subset V\).
+Note that for simplicity, we write \(f_i(v(t))\) even \(IN(v_i) \subset V\) (i.e., \(IN(v_i)\) does not contain some nodes of \(V\)).
 An update scheme of a Boolean network specifies the way that the nodes update their states through time evolution~\cite{thomas1991regulatory}. 
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!htb]
 \centering
@@
 Instead, we can first compute the mirror \(S_m\) of \(m\) in the Petri net encoding.
-Then, by Proposition~\ref{prop:siphon_set} and Theorem~\ref{theo:ts_2_sp}, we check if \(pred(S_m) \subseteq succ(S_m)\).
+Then, by Proposition~\ref{prop:siphon_set} and Theorem~\ref{theo:ts_2_sp}, we can check if \(pred(S_m) \subseteq succ(S_m)\).
 Note that the above computation (also the Petri net construction) is quadratic in the number of transitions of the Petri net in worst cases, which can be much smaller than the number of prime-implicants~\cite{chatain2014characterization}.
@@
 
-In~\cite{nabli2016enumerating}, the set of all siphons of a given Petri net is characterized by a similar Boolean Constraint Satisfcation Problem (CSP) except the conflict-freeness constraint.
+In~\cite{nabli2016enumerating}, the set of all siphons of a given Petri net is characterized by a similar Boolean CSP except the conflict-freeness constraint.
 From the encoded CSP, the set of all \emph{minimal} siphons of the Petri net can be enumerated in the set inclusion order.
 For enumerating siphons in the set inclusion order, the proposed method by~\cite{nabli2016enumerating} uses the technique that labels directly the Boolean variables with increasing value selection (i.e., to test first the absence, then the presence of a place in the candidate solution).
-The method has two implementations, one uses an iterated SAT procedure and the other uses Constraint Logic Programming with backtracking.
-
-One natural question is that how to use the CSP-based method for enumerating all the maximal conflict-free siphons of a Petri net encoding a Boolean model?
+The method has two implementations, one uses an iterated SAT procedure and the other uses Constraint Logic Programming (CLP) with backtracking.
+
+One natural question is that how to use the CSP-based method for enumerating all the maximal conflict-free siphons of a Petri net encoding a Boolean network?
 Of course, the set of all conflict-free siphons of the Petri net can easily characterized by the CSP model presented in~\cite{nabli2016enumerating} along with the additional constraint \(\neg p_v \vee \neg \overline{p}_v = 1\), for each \(v \in V\), which represents the conflict-freeness.
@@
 
-However a more direct method is to use a MaxSAT solver.
+For the SAT-based method, however a more direct method is to use a MaxSAT solver.
 We construct a MaxSAT problem with the following hard clauses:
@@
 We set a soft clause for each variable of the CSP and then use a ``minimal correction subset'' blocking strategy, which will ensure set-inclusion maximality of the solutions.
-This is what is implemented in \texttt{trappist} using the RC2 MaxSAT solver~\cite{DBLP:journals/jsat/IgnatievMM19} available through the \texttt{python-sat} package\footnote{\url{https://pysathq.github.io/docs/html/api/examples/rc2.html}}.
+This is what is implemented in Trappist using the RC2 MaxSAT solver~\cite{DBLP:journals/jsat/IgnatievMM19} available through the python-sat package\footnote{\url{https://pysathq.github.io/docs/html/api/examples/rc2.html}}.
+
+For the CLP-based method, we use \dots
 
@@
 Thus, all the maximal conflict-free siphons of the Petri net have been found.
+We use \dots as the ILP solver.
 
@@
 We shall show that our proposed methods can be easily adjusted to compute popular types of trap spaces.
-We illustrate the adjustments via the ASP-based method (see Subsection~\ref{subsec:computation_asp}), but these adjustments are completely applicable for other approaches such as MaxSAT, CSP, and ILP.
+We illustrate the adjustments via the ASP-based method (see Subsection~\ref{subsec:computation_asp}), but these adjustments are completely applicable for other approaches such as MaxSAT, CLP, and ILP.
 
@@
 
-To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} and mpbn~\cite{Paulev2020}.
+To solve the ASP problems, we used the same ASP solver Clingo~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} and mpbn~\cite{Paulev2020}.
 Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by PyBoolNet).
@@
 \setlength{\tabcolsep}{4pt}
-\begin{table}[!h]
+\begin{table}[!htb]
   \caption{Results on the real-world models from the BBM repository.}
@@
     mpbn & 2 (187) & N/A & 2.31 & N/A \\
-    Trappist-maxSAT & 1 & 0.03 & 1.09 & 1.01 \\
+    Trappist-MaxSAT & 1 & 0.03 & 1.09 & 1.01 \\
     Trappist-CLP & - & - & - & - \\
@@
 
-% \begin{table}[!htb]
-%   \caption{Results on scale-free models.}
-%   \centering%
-%   \label{tab:result_scale_free}
-%   \begin{tabular}{rrrrrrrr}
-%     \toprule
-%     & & & & \multicolumn{4}{c}{Trappist}\\
-%     \cmidrule(rr){5-8}
-%     \(n\) & LQM & mpbn & PBN & maxSAT & CLP & ILP & ASP \\
-%     \midrule
-%     500 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-%     750 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-%     1000 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-%     1250 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-%     1500 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-%     1750 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-%     2000 & 50-50-50 & 50-50-50 & - & - & - & - & - \\
-%     \bottomrule
-%   \end{tabular}
-% \end{table}
+Table~\ref{tab:result_N_K} shows the experimental results on N-K models.
+Column \(n\) denotes the network size.
+Columns ``LQM'' and ``PBN'' show the results of bioLQM and PyBoolNet, respectively.
+For each method, the number outside the parentheses indicates the number of failures, whereas the number inside the parentheses indicates the average running time (in seconds).
+Note that when computing the average running time, if the running time exceeds 180s, it is considered as 180s.
+From these results, we obtained several observations consistent with those obtained for real-world models.
+
+First, \dots
+
+Second, \dots
+
+Third, \dots
+
+Finally, \dots
 
 \section{Conclusion}
-
-In this article we proposed a new method for the computation of minimal trap spaces of Boolean networks, based on a new concept called maximal conflict-free siphons.
-This method is evaluated on large models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
+\label{sec:Conclusion}
+
+In this article we explored and proved for the first time the equivalence between (minimal) trap spaces of a general Boolean network and (maximal) conflict-free siphons of its Petri net encoding.
+From this equivalence, we proposed a new approach for the computation of minimal trap spaces of Boolean networks, based on the enumeration of maximal conflict-free siphons of Petri nets.
+We also proposed the four possible methods using MaxSAT, CLP, ILP, and ASP for implementing the new approach.
+The proposed methods were evaluated on many real-world models from the literature and randomly generated models.
+The experimental results show that \dots
 We believe that this opens up the way to a much better analysis of large Boolean networks, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
 
-Though we benchmarked this new approach against state-of-the-art tools, there are many more evaluations that we plan to do in the future.
-First, the bioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
-Note that this also raises the question of replacing altogether the bioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
-The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it (in other words, the trade-off of allowing redundant constraints) has to be evaluated in depth.
+%Though we benchmarked this new approach against state-of-the-art tools, there are many more evaluations that we plan to do in the future.
+%First, the bioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
+%Note that this also raises the question of replacing altogether the bioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
+%The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it (in other words, the trade-off of allowing redundant constraints) has to be evaluated in depth.
 

### 2023-01-11 trappist/bnet.py
ILP implementation

@@ -67,7 +67,7 @@ def read_bnet(fileobj: IO, method: str) -> nx.DiGraph:
         net.add_node(
             "-" + x, kind="place"
         )  # convention in PNML files obtained from bnet
-        if method in ("asp", "sat", "cp"):
+        if method in ("asp", "sat", "cp", "ilp"):
             vx = bddvar(x)
             fx = expr2bdd(expr(fx))
             # print(x, fx)

### 2023-01-11 trappist/ilp.py

@@ -25,11 +25,11 @@ from minizinc import Instance, Model, Result, Solver, Status
 
 import networkx as nx  # TODO maybe replace with lists/dicts
 
+from .cp import cp_to_bool, zincify
 
-def create_cp(petri_net: nx.DiGraph) -> Model:
-    """Create the CP program for the max conflict-free siphons of petri net."""
+def create_ilp(petri_net: nx.DiGraph) -> Model:
+    """Create the ILP program for the max conflict-free siphons of petri net."""
     model = Model()
-    model.add_string('include "globals.mzn";\n')
     variables = []
     for node in (n for n, k in petri_net.nodes(data="kind") if k == "place"):
         v = zincify(node)
@@ -39,7 +39,7 @@ def create_cp(petri_net: nx.DiGraph) -> Model:
         if kind == "place":
             if not node.startswith("-"):
                 # conflict-freeness
-                model.add_string(f"constraint not {node} \\/ not not_{node};\n")
+                model.add_string(f"constraint {node} + not_{node} <= 1;\n")
         else:
             # it's a transition, apply siphon
             # (if one succ is true, one pred must be true)
@@ -47,29 +47,22 @@ def create_cp(petri_net: nx.DiGraph) -> Model:
             for succ in petri_net.successors(node):
                 zsucc = zincify(succ)
                 if zsucc not in or_preds:  # optimize obvious tautologies
-                    or_string = " \\/ ".join(or_preds)
-                    model.add_string(f"constraint {zsucc} -> {or_string};\n")
+                    or_string = " + ".join(or_preds)
+                    model.add_string(f"constraint {zsucc} <= {or_string};\n")
     model.add_string(
-        f"solve :: bool_search([{', '.join(variables)}], input_order, indomain_max) satisfy;\n"
+        f"solve maximize {' + '.join(variables)};\n"
     )
     return model
 
 
-def zincify(variable: str) -> str:
-    """Return the name of the corresponding CP variable."""
-    if variable.startswith("-"):
-        return f"not_{variable[1:]}"
-    return variable
-
-
-def solve_cp(
+def solve_ilp(
     model: Model,
     max_output: int,
     time_limit: int,
     nprocs: int,
 ) -> Generator[List[int], None, None]:
-    """Run an CP solver on given Model and get the solutions."""
-    solver = Solver.lookup("chuffed")
+    """Run an ILP solver on given model and get the solutions."""
[…]

### 2023-01-11 trappist/trappist.py

@@ -30,6 +30,7 @@ import networkx as nx  # TODO maybe replace with lists/dicts
 from . import pnml_to_asp, version
 from .bnet import read_bnet
 from .cp import get_cp_solutions
+from .ilp import get_ilp_solutions
 from .max_sat import get_sat_solutions
 from .naive import write_naive_asp
 
@@ -168,7 +169,7 @@ def compute_trap_spaces(
         infile = open(infile, "r", encoding="utf-8")
         toclose = True
 
-    if infile.name.endswith(".pnml") and method in ("asp", "sat", "cp"):
+    if infile.name.endswith(".pnml") and method not in ("naive"):
         petri_net = read_pnml(infile)
     elif infile.name.endswith(".bnet"):
         petri_net = read_bnet(infile, method)
@@ -193,6 +194,8 @@ def compute_trap_spaces(
         solutions = get_sat_solutions(petri_net, max_output, time_limit, places)
     elif method == "cp":
         solutions = get_cp_solutions(petri_net, max_output, time_limit, places, nprocs)
+    elif method == "ilp":
+        solutions = get_ilp_solutions(petri_net, max_output, time_limit, places, nprocs)
     else:
         solutions_output = get_asp_output(
             petri_net, max_output, time_limit, method, debug, nprocs
@@ -249,7 +252,7 @@ def main():
     parser.add_argument(
         "-s",
         "--solver",
-        choices=["asp", "cp", "sat", "naive"],
+        choices=["asp", "cp", "ilp", "sat", "naive"],
         default="asp",
         type=str,
         help="Solver to compute the Maximal conflict-free siphons.\n"

### 2023-01-28 trappist/cp.py
error with CLP-based method

@@ -81,7 +81,8 @@ def solve_cp(
         max_output == 0 or nsol < max_output
     ):
         start = perf_counter()
-        result = inst.solve(timeout=remaining, processes=nprocs)
+        #result = inst.solve(timeout=remaining, processes=nprocs)
+        result = inst.solve_async(timeout=remaining, processes=nprocs)
         end = perf_counter()
         if remaining is not None:
             remaining -= timedelta(seconds=(end - start))

### 2023-01-28 manuscripts/TCS2022/tcs2022.tex
some corrections

@@
 
-\title{Minimal trap spaces of Boolean networks are maximal siphons of their Petri net encoding}
+\title{Trap spaces of Boolean networks are conflict-free siphons of their Petri net encoding}
 
@@
 
-Boolean modeling of gene regulation but also of post-trans\-criptomic systems has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
-This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
+Boolean network modeling of gene regulation but also of post-trans\-criptomic systems has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
+%This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
 Besides simulation, the analysis of such models is mostly based on attractor computation, since those correspond roughly to observable biological \emph{phenotypes}. 
 The recent use of trap spaces made a real breakthrough in that field allowing to consider medium-sized models that used to be out of reach.
-However, with the continuing increase in model-size, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of implicants.
-
-In this article we present an alternative method to compute minimal trap spaces, and hence complex attractors, of a Boolean network.
+However, with the continuing increase in model size and complexity of Boolean update functions, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits due to the difficulty of the prime-implicant computation.
+
+In this article we explore and prove for the first time a connection between trap spaces of a general Boolean network and siphons of its Petri net encoding.
+Besides important theoretical applications in studying properties of trap spaces, the connection enables us to propose an alternative approach to compute minimal trap spaces, and hence complex attractors, of a general Boolean network.
 It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
-After some technical preliminaries, we expose the concrete need for such a method and detail its implementation using Answer Set Programming.
-We then demonstrate its efficiency and compare it to implicant-based methods on some large Boolean networks from the literature.
+We then demonstrate its efficiency and compare it to the state-of-the-art methods on a large collection of real-world and randomly generated models.
 
@@
 
-Logical model \sep Boolean network \sep Trap space \sep Attractor computation \sep Petri net \sep Siphon \sep Answer set programming \sep Systems biology
+Logical model \sep Boolean network \sep Trap space \sep Attractor computation \sep Petri net \sep Siphon \sep Systems biology
 \end{keyword}
@@
 
-Boolean modelling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}, even for modelling post-transcriptional mechanisms.
+Boolean net modeling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}, even for modelling post-transcriptional mechanisms.
 This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas~\cite{aghamiri2020automated}.
@@
 The recent use of trap spaces~\cite{klarner2015computing} made a real breakthrough in that field allowing to consider medium-sized models that used to be out of reach.
-However, with the continuing increase in model-size, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits as there can be a huge number of implicants.
+However, with the continuing increase in model size and complexity of Boolean update functions, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits.
+More specifically, the number of prime implicants of a Boolean function is in general exponential in the number of input nodes of this function~\cite{klarner2015computing}.
+Moreover, the computation of prime implicants is a demanding task, especially for complex Boolean functions.
 
 It is worth noting that the recent method presented in~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces avoids the prime-implicants computation by relying on the \emph{most-permissive} semantics of Boolean networks.
-This method has been implemented in the tool mpbn\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
+This method has been implemented in the tool \texttt{mpbn}\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
 However, this method is only applicable for \emph{locally-monotonic} Boolean networks, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean networks (i.e., including both locally-monotonic and non-locally-monotonic ones).
-In addition, the bioLQM platform also provides another method using Binary Decision Diagrams (BDDs) in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html}.
+In addition, the \texttt{bioLQM} platform also provides another method using Binary Decision Diagrams (BDDs) in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html}.
 This method avoids the prime-implicants computation as it characterizes the set of generic trap spaces of a Boolean network by a BDD, then filters this set to get the set of all minimal trap spaces.
-By this characteristic, it requires the computation of all solutions, while the ASP-based methods~\cite{klarner2015computing,Paulev2020} can start enumerating them as they are found.
+By this approach, it requires the computation of all solutions, whereas the ASP-based methods~\cite{klarner2015computing,Paulev2020} can start enumerating them as they are found.
 Moreover, the main issue with the BDD-based method is that the number of generic trap spaces of a Boolean network may be extremely larger than the number of minimal trap spaces of this Boolean network.
@@
 In this article we explore and prove for the first time a connection between trap spaces of a general Boolean network and siphons of its Petri net encoding.
-Not only having important theoretical applications in studying properties of trap spaces in Boolean networks, the connection has important applications in the trap space computation.
+Not only having important theoretical applications in studying properties of trap spaces in Boolean networks, the connection has important practical applications in the trap space computation.
 Specifically, based on the connection, we propose an alternative approach to compute minimal trap spaces, and hence complex attractors, of a general Boolean network.
@@
 Second, we showcase a theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
-Third, beyond the proposed ASP method implementing the alternative approach~\cite{DBLP:conf/cmsb/TrinhBHS22}, we propose several other possible methods for computing minimal trap spaces using MaxSAT, CLP, and ILP.
+Third, beyond the proposed ASP method implementing the alternative approach~\cite{DBLP:conf/cmsb/TrinhBHS22}, we propose several other possible methods for computing minimal trap spaces using Maximum Satisfiability (MaxSAT), Constraint Logic Program (CLP), and Integer Linear Programming (ILP).
 Forth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
-Fifth, regarding the implementation, we have developed a new converter that directly reads a .bnet file and builds the Petri net encoding, instead of using the PNML conversion of bioLQM~\cite{DBLP:conf/cmsb/TrinhBHS22}.
+Fifth, regarding the implementation, we have developed a new converter that directly reads a \texttt{.bnet} file and builds the Petri net encoding, instead of using the \texttt{PNML} conversion of \texttt{bioLQM}~\cite{DBLP:conf/cmsb/TrinhBHS22}.
 Finally, we conduct a more comprehensive benchmark on more real-world models from various sources and randomly generated models to evaluate all the proposed methods (the benchmark conducted in~\cite{DBLP:conf/cmsb/TrinhBHS22} considers only real-world models).
@@
 
-Note that given a Boolean network in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the bioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
+Note that given a Boolean network in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the \texttt{bioLQM}\footnote{\url{http://www.colomoto.org/biolqm/}} library.
 This piece of software extracted from GINsim~\cite{chaouiya2012logical} and part of the CoLoMoTo\footnote{\url{http://colomoto.org/}}~\cite{naldi2015cooperative} software suite allows for easy conversion between standard formats.
@@
 This does not change the worst-case-complexity (obtaining a single DNF being exponential) but might matter a lot in practice.
-As such, we will explore how this transformation, here using BDDs in bioLQM, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.
+As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM}, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.
 
@@
 One naturally computational application of Theorem~\ref{theo:ts_2_sp} is that we can efficiently decide whether a subspace \(m\) is a trap space.
-In PyBoolNet~\cite{klarner2017pyboolnet}, this is checked by using the percolation on the prime-implicants of the Boolean functions.
-As we have mentioned at the beginning of this article, there may be a huge number of prime-implicants.
+In \texttt{PyBoolNet}~\cite{klarner2017pyboolnet}, this is checked by using the percolation on the prime-implicants of the Boolean functions.
+As we have mentioned at the beginning of this article, the computation of prime-implicants is a demanding task for complex Boolean networks, even is sometimes intractable.
 Hence, the checking method in~\cite{klarner2017pyboolnet} shows its limitations.
@@
 Then, by Proposition~\ref{prop:siphon_set} and Theorem~\ref{theo:ts_2_sp}, we can check if \(pred(S_m) \subseteq succ(S_m)\).
-Note that the above computation (also the Petri net construction) is quadratic in the number of transitions of the Petri net in worst cases, which can be much smaller than the number of prime-implicants~\cite{chatain2014characterization}.
+Note that the Petri net construction is less computationally demanding than the prime-implicant computation because it only requires computing generic (not prime) implicants of the Boolean functions~\cite{chatain2014characterization}.
+In addition, the time complexity of the above checking method is quadratic in the number of transitions of the Petri net in worst cases.
 
@@
 We would therefore expect that, when available, it will have comparable performance on the ASP part (the ASP program would be approximately twice smaller, though redundancy is not always bad in that field), but can also avoid combinatorial explosion of the Petri net encoding for some formula where the activation DNF is simple but the inhibition is not.
-Since mpbn is included in our benchmark this will be evaluated in our experiments.
+Since \texttt{mpbn} is included in our benchmark this will be evaluated in our experiments.
 
@@
 
-However, using a state-of-the-art tool like PyBoolNet~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
-mpbn~\cite{Paulev2020} does not give any answer either because it recognizes that model as non-locally-monotonic.
+However, using a state-of-the-art tool like \texttt{PyBoolNet}~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
+\texttt{mpbn}~\cite{Paulev2020} does not give any answer either because it recognizes that model as non-locally-monotonic.
 And hence, it is not possible to extract any (complex) attractor at all.
@@
 
-Using the method presented above, it is possible to convert the model to PNML in about one second and to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) in a few milliseconds.
+Using the method presented above, it is possible to convert the model to \texttt{PNML} in about one second and to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) in a few milliseconds.
 Unfortunately since this was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.
@@
 
-To evaluate the performance of the newly proposed methods (implemented as a Python package named Trappist) and the state-of-the-art methods (bioLQM\footnote{\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}}, PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet}, and mpbn~\cite{Paulev2020}), we compared them on both PyBoolNet's own model repository and many real-world models from various sources in the literature.
-It is worth noting that mpbn~\cite{Paulev2020} only handles locally-monotonic models, whereas the other methods can handle general models.
-To obtain a more comprehensive comparison, we also used random models generated by a third-party software (i.e., BoolNet R package~\cite{mussel2010boolnet}).
-As explained in Section~\ref{sec:case_study}, in our benchmarks, we only searched for the first 1000 solutions for each model.
-
-To solve the ASP problems, we used the same ASP solver Clingo~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} and mpbn~\cite{Paulev2020}.
-Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by PyBoolNet).
+To evaluate the performance of the newly proposed methods (implemented as a Python package named Trappist) and the state-of-the-art methods (\texttt{bioLQM}\footnote{\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}}, \texttt{PyBoolNet}~\cite{klarner2015computing,klarner2017pyboolnet}, and \texttt{mpbn}~\cite{Paulev2020}), we compared them on both \texttt{PyBoolNet}'s own model repository and many real-world models from various sources in the literature.
+It is worth noting that \texttt{mpbn}~\cite{Paulev2020} only handles locally-monotonic models, whereas the other methods can handle general models.
+To obtain a more comprehensive comparison, we also used random models generated by a third-party software (i.e., \texttt{BoolNet R} package~\cite{mussel2010boolnet}).
+As explained in Section~\ref{sec:case_study}, in our benchmarks, we only searched for the first 1000 minimal trap spaces for each model.
+
+To solve the ASP problems, we used the same ASP solver Clingo~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in \texttt{PyBoolNet}~\cite{klarner2015computing,klarner2017pyboolnet} and \texttt{mpbn}~\cite{Paulev2020}.
+Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by \texttt{PyBoolNet}).
 We ran all the benchmarks on a machine whose environment is CPU: \@ Intel\textregistered\ Core\texttrademark\ i9-11950H 2.60GHz × 16, 16 GB DDR4 RAM, Ubuntu 20.04.5 LTS. 
@@
 
-\subsection{PyBoolNet repository}
+\subsection{\texttt{PyBoolNet} repository}
 
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons (in seconds) between bioLQM (LQM), PyBoolNet (PBN), mpbn and the four variants of Trappist on the PyBoolNet repository.}
+  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (LQM), \texttt{PyBoolNet} (PBN), \texttt{mpbn} and the four variants of Trappist on the \texttt{PyBoolNet} repository.}
   \centering%
@@
     \cmidrule(rr){8-11}
-    & model & \(n\) & \(|M|\) & LQM & PBN & mpbn & SAT & CLP & ILP & ASP \\
+    & model & \(n\) & \(|M|\) & LQM & PBN & \texttt{mpbn} & SAT & CLP & ILP & ASP \\
     \midrule
@@
 
-Table~\ref{tab:pyboolnet_repo} shows the experimental results on the models from the official PyBoolNet repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}.
+Table~\ref{tab:pyboolnet_repo} shows the experimental results on the models from the official \texttt{PyBoolNet} repository\footnote{\url{https://github.com/hklarner/pyboolnet/tree/master/pyboolnet/repository}}.
 Column \(n\) denotes the number of nodes of each model. 
@@
 
-As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the PyBoolNet repository, the results are comparable with all minimal trap spaces found very fast.
-For 5 of the 29 models, mpbn did not give any answer because it recognized these models as not locally-monotonic.
-Note that on some very small models, Trappist is sometimes slower than PyBoolNet and/or mpbn, but still significantly under one second.
-On the contrary, on every model that was a bit challenging for PyBoolNet or mpbn, the new method is far more efficient with speedups between one and two orders of magnitude.
+As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the \texttt{PyBoolNet} repository, the results are comparable with all minimal trap spaces found very fast.
+For 5 of the 29 models, \texttt{mpbn} did not give any answer because it recognized these models as not locally-monotonic.
+Note that on some very small models, Trappist is sometimes slower than \texttt{PyBoolNet} and/or \texttt{mpbn}, but still significantly under one second.
+On the contrary, on every model that was a bit challenging for \texttt{PyBoolNet} or \texttt{mpbn}, the new method is far more efficient with speedups between one and two orders of magnitude.
 
@@
     Method & \# failures & avg-lqm (s) & avg-mono (s) & avg-all (s) \\ \midrule
-    bioLQM & 9 (134) & 12.87 & N/A & N/A \\
-    PyBoolNet & 12 & 8.87 & 11.00 & 13.59 \\
-    mpbn & 2 (187) & N/A & 2.31 & N/A \\
+    \texttt{bioLQM} & 9 (134) & 12.87 & N/A & N/A \\
+    \texttt{PyBoolNet} & 12 & 8.87 & 11.00 & 13.59 \\
+    \texttt{mpbn} & 2 (187) & N/A & 2.31 & N/A \\
     Trappist-MaxSAT & 1 & 0.03 & 1.09 & 1.01 \\
@@
 Columns 2 expresses the numbers of failures (i.e., did not finish the computation within a time limit of three minutes) of each method.
-For the case of bioLQM, we only considered the models that have at most 1000 minimal trap spaces.
+For the case of \texttt{bioLQM}, we only considered the models that have at most 1000 minimal trap spaces.
 The number of such models is 134 (per all 211 models) and is denoted inside the parentheses.
-For the case of mpbn, we only considered the models that are locally-monotonic.
+For the case of \texttt{mpbn}, we only considered the models that are locally-monotonic.
 The number of such models is 187 (per all 211 models) and is denoted inside the parentheses.
@@
 These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
-Note that these models are not included in the PyBoolNet and BBM repositories.
-We then applied bioLQM, PyBoolNet, mpbn, and the four variants of Trappist to computing minimal trap spaces of these real-world models.
+Note that these models are not included in the \texttt{PyBoolNet} and BBM repositories.
+We then applied \texttt{bioLQM}, \texttt{PyBoolNet}, \texttt{mpbn}, and the four variants of Trappist to computing minimal trap spaces of these real-world models.
 It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes in these models.
@@
 ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces) within the timeout of two minutes.
-In the case of bioLQM, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of bioLQM because it always requires to compute all minimal trap spaces.
+In the case of \texttt{bioLQM}, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of \texttt{bioLQM} because it always requires to compute all minimal trap spaces.
 A number in bold indicates a ratio greater than or equal to 10 compared to the best result.
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons (in seconds) between bioLQM (LQM), PyBoolNet (PBN), mpbn and the four variants of Trappist on selected models from the literature.}
+  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (LQM), \texttt{PyBoolNet} (PBN), \texttt{mpbn} and the four variants of Trappist on selected models from the literature.}
   % in bold if greater than or equal to 10
@@
     \cmidrule(rr){8-11}
-    & model & \(n\) & \(|M|\) & LQM & PBN & mpbn & SAT & CLP & ILP & ASP \\
+    & model & \(n\) & \(|M|\) & LQM & PBN & \texttt{mpbn} & SAT & CLP & ILP & ASP \\
     \midrule % n < 100
@@
 
-The first observation is that for 26 of the 33 models (more than 78\%), mpbn did not give any answer because it recognized that these models as not locally-monotonic.
-For 6 of the 33 models where mpbn returned the answers, mpbn and Trappist are comparable in computation time, though surprisingly mpbn appears a bit slower on average.
-Note however that mpbn was the only tool to provide a solution for the SN-5 model, thus confirming that if the activation function is in the right form, not having to compute the inactivation function's disjunctive normal form can render a difficult problem tractable.
+The first observation is that for 26 of the 33 models (more than 78\%), \texttt{mpbn} did not give any answer because it recognized that these models as not locally-monotonic.
+For 6 of the 33 models where \texttt{mpbn} returned the answers, \texttt{mpbn} and Trappist are comparable in computation time, though surprisingly \texttt{mpbn} appears a bit slower on average.
+Note however that \texttt{mpbn} was the only tool to provide a solution for the SN-5 model, thus confirming that if the activation function is in the right form, not having to compute the inactivation function's disjunctive normal form can render a difficult problem tractable.
 However, since mbpn can handle only locally-monotonic models and Trappist can handle general models, it is difficult to further compare between them.
-Hence, we focus on only comparisons between PyBoolNet and Trappist in the following observations.
-
-The second observation is that the proposed method vastly outperforms PyBoolNet in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100--1000 nodes size range).
-Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
+Hence, we focus on only comparisons between \texttt{PyBoolNet} and Trappist in the following observations.
+
+The second observation is that the proposed method vastly outperforms \texttt{PyBoolNet} in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100--1000 nodes size range).
+Note that for all the cases where \texttt{PyBoolNet} did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
 Hence, not even a single minimal trap space was output by that method.
@@
 
-The third observation is that for all the models where PyBoolNet finished before the timeout, once PyBoolNet went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
+The third observation is that for all the models where \texttt{PyBoolNet} finished before the timeout, once \texttt{PyBoolNet} went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
 For these models, the ASP solving phase of the proposed method also took very short time, all under one second.
-Hence, with the experimental results shown in this paper, the practical differences between our ASP encoding and that of PyBoolNet are not distinctly exposed.
+Hence, with the experimental results shown in this paper, the practical differences between our ASP encoding and that of \texttt{PyBoolNet} are not distinctly exposed.
 The fact that our new ASP encoding is guaranteed to be linear in the number of nodes of the original model does not seem to be crucial here, however a much deeper analysis of those cases remains to be done.
@@
 The evaluation of the feasibility of such strategy, and of its impact on smaller instances, remains to be done and is out of the scope of this first article.
-Recognizing that a model is locally-monotonic and applying in that specific case dedicated strategies as those of mpbn might also be a partial solution as shown in the SN-5 case, but not for the turei-2016 model.
+Recognizing that a model is locally-monotonic and applying in that specific case dedicated strategies as those of \texttt{mpbn} might also be a partial solution as shown in the SN-5 case, but not for the turei-2016 model.
 
@@
 It is worth noting that N-K models usually have small numbers of minimal trap spaces~\cite{klarner2015computing}.
-Hence, we searched for all solutions in each model, which makes the comparison to bioLQM more comprehensive.
+Hence, we searched for all solutions in each model, which makes the comparison to \texttt{bioLQM} more comprehensive.
 In addition, each node has only three input nodes, i.e., the number of prime-implicants of the associated Boolean function is small.
-Hence, PyBoolNet always passed the phase of computing prime-implicants in every model even within 1s, which enables us to compare the ASP encoding of PyBoolNet and that of Trappist.
+Hence, \texttt{PyBoolNet} always passed the phase of computing prime-implicants in every model even within 1s, which enables us to compare the ASP encoding of \texttt{PyBoolNet} and that of Trappist.
 
@@
     \cmidrule(rr){5-8}
-    \(n\) & LQM & mpbn & PBN & SAT & CLP & ILP & ASP \\
+    \(n\) & LQM & \texttt{mpbn} & PBN & SAT & CLP & ILP & ASP \\
     \midrule
@@
 Column \(n\) denotes the network size.
-Columns ``LQM'' and ``PBN'' show the results of bioLQM and PyBoolNet, respectively.
+Columns ``LQM'' and ``PBN'' show the results of \texttt{bioLQM} and \texttt{PyBoolNet}, respectively.
 For each method, the number outside the parentheses indicates the number of failures, whereas the number inside the parentheses indicates the average running time (in seconds).
@@
 %Though we benchmarked this new approach against state-of-the-art tools, there are many more evaluations that we plan to do in the future.
-%First, the bioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
-%Note that this also raises the question of replacing altogether the bioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
+%First, the \texttt{bioLQM} platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like \texttt{PyBoolNet}, that remains to be checked.
+%Note that this also raises the question of replacing altogether the \texttt{bioLQM} preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
 %The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it (in other words, the trade-off of allowing redundant constraints) has to be evaluated in depth.
 
-Second, the experimental results shown in this paper mostly expose the differences caused by the prime-implicant phase of PyBoolNet.
-There is much more information required to distinctly study the practical differences between our ASP encoding and that of PyBoolNet, and notably their size/efficiency ratio.
+Second, the experimental results shown in this paper mostly expose the differences caused by the prime-implicant phase of \texttt{PyBoolNet}.
+There is much more information required to distinctly study the practical differences between our ASP encoding and that of \texttt{PyBoolNet}, and notably their size/efficiency ratio.
 Hence, we plan to conduct experiments on more real-world models or maybe random models that can be randomly generated by using BoolNet~\cite{mussel2010boolnet}.

### 2023-01-28 manuscripts/TCS2022/tcs2022.tex
inconsistent and incomplete data

@@
 Second, we showcase a theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
-Third, beyond the proposed ASP method implementing the alternative approach~\cite{DBLP:conf/cmsb/TrinhBHS22}, we propose several other possible methods for computing minimal trap spaces using Maximum Satisfiability (MaxSAT), Constraint Logic Program (CLP), and Integer Linear Programming (ILP).
-Forth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
-Fifth, regarding the implementation, we have developed a new converter that directly reads a \texttt{.bnet} file and builds the Petri net encoding, instead of using the \texttt{PNML} conversion of \texttt{bioLQM}~\cite{DBLP:conf/cmsb/TrinhBHS22}.
+Third, beyond the proposed ASP method implementing the alternative approach~\cite{DBLP:conf/cmsb/TrinhBHS22}, we propose several other possible methods for computing minimal trap spaces using Maximum Satisfiability (MaxSAT), Constraint Logic Programming (CLP), and Integer Linear Programming (ILP).
+Fourth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
+Fifth, we present the idea for using our Petri net approach to handle the problem of inconsistent and incomplete data.
+Sixth, regarding the implementation, we have developed a new converter that directly reads a \texttt{.bnet} file and builds the Petri net encoding, instead of using the \texttt{PNML} conversion of \texttt{bioLQM}~\cite{DBLP:conf/cmsb/TrinhBHS22}.
 Finally, we conduct a more comprehensive benchmark on more real-world models from various sources and randomly generated models to evaluate all the proposed methods (the benchmark conducted in~\cite{DBLP:conf/cmsb/TrinhBHS22} considers only real-world models).
@@
   \node[circle,draw,label=left:$p_{x_1}$] (x1) {};
-  \node[circle,draw,label=left:$\overline{p_{x_1}}$] (nx1) [below=of x1, yshift=-0.5cm] {};
+  \node[circle,draw,label=left:$\overline{p}_{x_1}$] (nx1) [below=of x1, yshift=-0.5cm] {};
 
   \node[circle,draw,label=right:$p_{x_2}$] (x2) [right=of x1, xshift=2.0cm] {};
-  \node[circle,draw,label=right:$\overline{p_{x_2}}$] (nx2) [below=of x2, yshift=-0.5cm] {};
+  \node[circle,draw,label=right:$\overline{p}_{x_2}$] (nx2) [below=of x2, yshift=-0.5cm] {};
 
@@
 We set a soft clause for each variable of the CSP and then use a ``minimal correction subset'' blocking strategy, which will ensure set-inclusion maximality of the solutions.
-This is what is implemented in Trappist using the RC2 MaxSAT solver~\cite{DBLP:journals/jsat/IgnatievMM19} available through the python-sat package\footnote{\url{https://pysathq.github.io/docs/html/api/examples/rc2.html}}.
-
-For the CLP-based method, we use \dots
+This is what is implemented in \texttt{Trappist} using the \texttt{RC2} MaxSAT solver~\cite{DBLP:journals/jsat/IgnatievMM19} available through the python-sat package\footnote{\url{https://pysathq.github.io/docs/html/api/examples/rc2.html}}.
+
+For the CLP-based method, we use TODO: \dots
 
@@
 
+\subsection{Inconsistent and incomplete data}
+
+So far we have assumed that we are always able to derive complete
+and consistent truth-tables (i.e., Boolean functions) that correctly capture the behavior of each node in a Boolean network. 
+However, in practice it is rarely the
+case that a Boolean network of a biological system is fully understood and indeed, this is one important reason for modeling such networks. 
+The data provided may be incomplete in the sense that information is missing about what happens in certain states, or it may be inconsistent in that we have conflicting information.
+The result is that the behavior of some nodes under certain conditions may be unknown~\cite{Steggles2006}.
+
+Such incomplete and/or inconsistent behavioral information is
+problematic for Boolean networks with the synchronous update scheme, which are unable to represent the possibility of more than one next state.
+It can be addressed by using Boolean networks with asynchronous update schemes because of their non-determinism.
+However, multiple update functions may be considered for each node, which is problematic for existing analysis methods in Boolean networks.
+For example, if there are \(k\) nodes with two possible Boolean functions, then the analysis methods such as \texttt{PyBoolNet} and \texttt{mpbn} need to consider \(2^k\) possible Boolean networks due to they inputs are Boolean networks with complete Boolean functions.
+Moreover, one interesting question maybe raised from systems biologists is if there is a common behavioral property (e.g., minimal trap space, attractor) among all possible combinations.
+To answer this question, the existing analysis methods may consider \(2^k\) possible Boolean networks as well.
+
+We shall show how our Petri net approach presented the previous section can handle the problem of inconsistent and incomplete data.
+The idea is to build the Petri net encoding of the Boolean network from its partial truth tables.
+More specifically, for each unknown-output row in the truth table of a node, we create two conflicting transitions representing the two possible outputs (i.e., 0 or 1).
+For every remaining row, we create transitions as usual following the Petri net encoding~\cite{chatain2014characterization}.
+Because of the non–deterministic choice mechanisms of Petri nets, the new Petri net is still equivalent in dynamics to the incomplete Boolean network.
+One notable advantage of the connection between trap spaces of a Boolean network and conflict-free siphons of its Petri net encoding is that it still holds if the Petri net encoding changes (e.g., more or less transitions) but still maintain the equivalence in dynamics to the Boolean network.
+Hence, the minimal trap spaces of the incomplete Boolean network can be computed by computing maximal conflict-free siphons of the resulting Petri net.
+As a consequence, we can easily answer the question on common behavioral property.
+
+\begin{example}
+Consider an incomplete Boolean network \(\mathcal{N} = (V, F)\) (mentioned in~\cite{Steggles2006}), where \(V = (g_1, g_2, g_3)\) and \(F = (f_1, f_2, f_3)\) with \(f_1 = g_2, f_3 = \neg g_1\).
+Herein, \(f_2\) is an incomplete function and the partial truth table is as follows.
+\begin{center}
+	\begin{tabular}{rr|r}
+		\toprule
+		$g_1$ & $g_3$ & $g_2$ \\ \midrule
+		0 & 0 & 0 \\
+		0 & 1 & 0 \\
+		1 & 0 & ? \\
+		1 & 1 & 1 \\
+		\bottomrule
+	\end{tabular}
+\end{center}
+We have two possible update functions for \(g_2\): \(f_2 = g_1 \land g_3\) for the case that \(? = 0\) and \(f_2 = g_1\) for the case that \(? = 1\).
+Accordingly, we have two possible Boolean networks.
+The first one has one minimal trap space: 001.
+The second one has two minimal trap spaces: 001 and 110.
+Both have a common minimal trap space: 001.
+The asynchronous dynamics of the incomplete Boolean network is shown in Figure~\ref{fig:incomplete-dynamics-new-PN}a.
+The result Petri net following our idea is given in Figure~\ref{fig:incomplete-dynamics-new-PN}b.
+Herein, \(t^3_{g_2}\) and \(t^4_{g_2}\) are the two conflicting transitions representing the unknown-output row.
+This Petri net has only one maximal conflict-free siphon \(\{p_{g_1}, p_{g_2}, \overline{p}_{g_3}\}\) corresponding to the common minimal trap space 001.
+
+\label{example:incomplete-BN}
+\end{example}
+
+% TODO: finish the figure
+\begin{figure}[!htb]
+	\centering
+	\begin{subfigure}[b]{0.3\textwidth}
+		\centering
+		\begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]
+		\node[] (0) [] {000};
+		\node[] (1) [left=of 0] {001};
+		\node[] (4) [above=of 0] {100};
+		\node[] (6) [above=of 1] {110};
+		\node[] (5) [below=of 1] {101};
+		\node[] (2) [left=of 6] {010};
+		\node[] (3) [left=of 1] {011};
+		\node[] (7) [below=of 3] {111};
+
+		\draw[->] (0) edge [] (1);
+		\draw[->] (4) edge [] (0);
+		\draw[->] (4) edge [bend left] (6);
+		\draw[->] (6) edge [bend left] (4);
+
+		\draw[->] (5) edge [] (1);
+		\draw[->] (5) edge [] (7);
+
+		\draw[->] (2) edge [] (6);
+		\draw[->] (2) edge [] (0);
+		\draw[->] (2) edge [] (3);
+
+		\draw[->] (3) edge [] (1);
+		\draw[->] (3) edge [] (7);
+
+		\draw[->] (7) edge [] (6);
+
+		\end{tikzpicture}
+		\caption{}%
+		%\label{fig:stg}
+	\end{subfigure}\hfill%
+	\begin{subfigure}[b]{0.6\textwidth}
+		\centering
+		\begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]\footnotesize
+		\node[circle,draw,label=left:$p_{g_1}$] (x1) {};
+		\node[circle,draw,label=left:$\overline{p}_{g_1}$] (nx1) [below=of x1, yshift=-0.5cm] {};
+
+		\node[circle,draw,label=right:$p_{g_2}$] (x2) [right=of x1, xshift=2.5cm] {};
+		\node[circle,draw,label=right:$\overline{p}_{g_2}$] (nx2) [below=of x2, yshift=-0.5cm] {};
+
+		\node[circle,draw,label=left:$p_{g_3}$] (x3) [below=of nx1, yshift=-0.5cm] {};
+		\node[circle,draw,label=right:$\overline{p}_{g_3}$] (nx3) [below=of nx2, yshift=-0.5cm] {};
+
+		\node[rectangle,draw,label=above:$t^1_{g_2}$] (t1x2) [right=of x1] {};
+		\node[rectangle,draw,label=above:$t^2_{g_2}$] (t2x2) [right=of nx1] {};
+
+		\node[rectangle,draw,label=left:$t^3_{g_2}$] (t3x2) [left=of nx2] {};
+		\node[rectangle,draw,label=left:$t^4_{g_2}$] (t4x2) [left=of nx3] {};
+
+		\draw[->] (x2) -- (t1x2);
+		\draw[->] (t1x2) -- (nx2);
+		\draw[<->] (t1x2) -- (nx1);
+
+		\draw[->] (nx2) edge [bend left] (t2x2);
+		\draw[->] (t2x2) -- (x2);
+		\draw[<->] (t2x2) -- (x1);
+		\draw[<->] (t2x2) -- (x3);
+
+		\draw[->] (x2) -- (t3x2);
+		\draw[->] (t3x2) -- (nx2);
+		\draw[<->] (t3x2) -- (x1);
+		\draw[<->] (t3x2) -- (nx3);
+
+		\draw[->] (nx2) -- (t4x2);
+		\draw[->] (t4x2) -- (x2);
+		\draw[<->] (t4x2) edge [bend left=15] (x1);
+		\draw[<->] (t4x2) -- (nx3);
+
+		\end{tikzpicture}
+		\caption{}%
+		%\label{fig:PN}
+	\end{subfigure}%
+	\caption{Asynchronous dynamics (self arcs are omitted for simplicity) and new Petri net encoding (the transitions corresponding to \(g_1\) and \(g_3\) are omitted for simplicity) of the incomplete Boolean network of Example~\ref{example:incomplete-BN}}%
+	\label{fig:incomplete-dynamics-new-PN}
+\end{figure}
+
+In summary, we can still perform meaningful analysis on the resulting Petri net.
+Note that the new Petri net may have more transitions, however the number of additional transitions is only \(O(2k)\) where \(k\) is the number of unknown-output rows in the truth tables.
+This is actually a significant advantage, since we do not need to consider \(2^k\) possible Boolean networks.
+It is worth noting that the approach proposed in~\cite{Steggles2006} only deals with the case that \(k = 1\).
+Furthermore, as more data becomes available for the underlying biological system, the Petri net model can be refined to reduce the
+amount of non-determinism (i.e., the number of conflicting transitions) it contains.
+Hence, Petri nets provide an interesting means of documenting the development of knowledge about a biological system.
+
 \section{Motivating example}
@@
 
-To evaluate the performance of the newly proposed methods (implemented as a Python package named Trappist) and the state-of-the-art methods (\texttt{bioLQM}\footnote{\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}}, \texttt{PyBoolNet}~\cite{klarner2015computing,klarner2017pyboolnet}, and \texttt{mpbn}~\cite{Paulev2020}), we compared them on both \texttt{PyBoolNet}'s own model repository and many real-world models from various sources in the literature.
+To evaluate the performance of the newly proposed methods (implemented as a Python package named \texttt{Trappist}) and the state-of-the-art methods (\texttt{bioLQM}\footnote{\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}}, \texttt{PyBoolNet}~\cite{klarner2015computing,klarner2017pyboolnet}, and \texttt{mpbn}~\cite{Paulev2020}), we compared them on both \texttt{PyBoolNet}'s own model repository and many real-world models from various sources in the literature.
 It is worth noting that \texttt{mpbn}~\cite{Paulev2020} only handles locally-monotonic models, whereas the other methods can handle general models.
@@
 
-To solve the ASP problems, we used the same ASP solver Clingo~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in \texttt{PyBoolNet}~\cite{klarner2015computing,klarner2017pyboolnet} and \texttt{mpbn}~\cite{Paulev2020}.
+To solve the ASP problems, we used the same ASP solver \texttt{Clingo}~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in \texttt{PyBoolNet}~\cite{klarner2015computing,klarner2017pyboolnet} and \texttt{mpbn}~\cite{Paulev2020}.
 Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by \texttt{PyBoolNet}).
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (LQM), \texttt{PyBoolNet} (PBN), \texttt{mpbn} and the four variants of Trappist on the \texttt{PyBoolNet} repository.}
+  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (LQM), \texttt{PyBoolNet} (PBN), \texttt{mpbn} and the four variants of \texttt{Trappist} on the \texttt{PyBoolNet} repository.}
   \centering%
@@
     \toprule
-    & & & & & & & \multicolumn{4}{c}{Trappist}\\
+    & & & & & & & \multicolumn{4}{c}{\texttt{Trappist}}\\
     \cmidrule(rr){8-11}
@@
 ``NM'' indicates a non-locally-monotonic model.
-There are four variants of Trappist: SAT (i.e., the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), CLP (i.e., the CLP-based method shown in Subsection~\ref{subsec:computation_csp}), ILP (i.e., the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and ASP (i.e., the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
+There are four variants of \texttt{Trappist}: SAT (i.e., the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), CLP (i.e., the CLP-based method shown in Subsection~\ref{subsec:computation_csp}), ILP (i.e., the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and ASP (i.e., the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
 
@@
 For 5 of the 29 models, \texttt{mpbn} did not give any answer because it recognized these models as not locally-monotonic.
-Note that on some very small models, Trappist is sometimes slower than \texttt{PyBoolNet} and/or \texttt{mpbn}, but still significantly under one second.
+Note that on some very small models, \texttt{Trappist} is sometimes slower than \texttt{PyBoolNet} and/or \texttt{mpbn}, but still significantly under one second.
 On the contrary, on every model that was a bit challenging for \texttt{PyBoolNet} or \texttt{mpbn}, the new method is far more efficient with speedups between one and two orders of magnitude.
@@
     \texttt{mpbn} & 2 (187) & N/A & 2.31 & N/A \\
-    Trappist-MaxSAT & 1 & 0.03 & 1.09 & 1.01 \\
-    Trappist-CLP & - & - & - & - \\
-    Trappist-ILP & - & - & - & - \\
-    Trappist-ASP & 1 & 0.05 & 1.02 & 0.93 \\
+    \texttt{Trappist}-MaxSAT & 1 & 0.03 & 1.09 & 1.01 \\
+    \texttt{Trappist}-CLP & - & - & - & - \\
+    \texttt{Trappist}-ILP & - & - & - & - \\
+    \texttt{Trappist}-ASP & 1 & 0.05 & 1.02 & 0.93 \\
     \bottomrule
@@
 Note that these models are not included in the \texttt{PyBoolNet} and BBM repositories.
-We then applied \texttt{bioLQM}, \texttt{PyBoolNet}, \texttt{mpbn}, and the four variants of Trappist to computing minimal trap spaces of these real-world models.
+We then applied \texttt{bioLQM}, \texttt{PyBoolNet}, \texttt{mpbn}, and the four variants of \texttt{Trappist} to computing minimal trap spaces of these real-world models.
 It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes in these models.
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (LQM), \texttt{PyBoolNet} (PBN), \texttt{mpbn} and the four variants of Trappist on selected models from the literature.}
+  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (LQM), \texttt{PyBoolNet} (PBN), \texttt{mpbn} and the four variants of \texttt{Trappist} on selected models from the literature.}
   % in bold if greater than or equal to 10
@@
 The first observation is that for 26 of the 33 models (more than 78\%), \texttt{mpbn} did not give any answer because it recognized that these models as not locally-monotonic.
-For 6 of the 33 models where \texttt{mpbn} returned the answers, \texttt{mpbn} and Trappist are comparable in computation time, though surprisingly \texttt{mpbn} appears a bit slower on average.
+For 6 of the 33 models where \texttt{mpbn} returned the answers, \texttt{mpbn} and \texttt{Trappist} are comparable in computation time, though surprisingly \texttt{mpbn} appears a bit slower on average.
 Note however that \texttt{mpbn} was the only tool to provide a solution for the SN-5 model, thus confirming that if the activation function is in the right form, not having to compute the inactivation function's disjunctive normal form can render a difficult problem tractable.
-However, since mbpn can handle only locally-monotonic models and Trappist can handle general models, it is difficult to further compare between them.
-Hence, we focus on only comparisons between \texttt{PyBoolNet} and Trappist in the following observations.
+However, since mbpn can handle only locally-monotonic models and \texttt{Trappist} can handle general models, it is difficult to further compare between them.
+Hence, we focus on only comparisons between \texttt{PyBoolNet} and \texttt{Trappist} in the following observations.
 
@@
 In addition, each node has only three input nodes, i.e., the number of prime-implicants of the associated Boolean function is small.
-Hence, \texttt{PyBoolNet} always passed the phase of computing prime-implicants in every model even within 1s, which enables us to compare the ASP encoding of \texttt{PyBoolNet} and that of Trappist.
+Hence, \texttt{PyBoolNet} always passed the phase of computing prime-implicants in every model even within 1s, which enables us to compare the ASP encoding of \texttt{PyBoolNet} and that of \texttt{Trappist}.
 
@@
     \toprule
-    & & & & \multicolumn{4}{c}{Trappist}\\
+    & & & & \multicolumn{4}{c}{\texttt{Trappist}}\\
     \cmidrule(rr){5-8}
@@
 
+% Compare among various possible methods and obtained insights
 Second, the experimental results shown in this paper mostly expose the differences caused by the prime-implicant phase of \texttt{PyBoolNet}.
@@
 
+% Update: can borrow many advents in the field of Petri nets
 In addition, there are possibly other methods for computing maximal conflict-free siphons in Petri nets, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}.

### 2023-01-29 manuscripts/TCS2022/tcs2022.tex
new section for incomplete data problem

@@
 Fourth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
-Fifth, we present the idea for using our Petri net approach to handle the problem of inconsistent and incomplete data.
+Fifth, we present the idea for using our Petri net approach to handle the problem of inconsistent and incomplete data in modeling biological systems.
 Sixth, regarding the implementation, we have developed a new converter that directly reads a \texttt{.bnet} file and builds the Petri net encoding, instead of using the \texttt{PNML} conversion of \texttt{bioLQM}~\cite{DBLP:conf/cmsb/TrinhBHS22}.
@@
 Section~\ref{sec:Computation} presents the alternative approach for computing minimal trap spaces and the four possible methods implementing it.
+Section~\ref{sec:incomplete-data} presents the idea to deal with problem of inconsistent and incomplete data.
 Section~\ref{sec:case_study} shows an important biological case study showing the applicability of the new approach.
@@
 
-\subsection{Inconsistent and incomplete data}
+\section{Inconsistent and incomplete data}
+\label{sec:incomplete-data}
 

### 2023-01-29 manuscripts/TCS2022/tcs2022.tex
update motivating example

@@
 However, using a state-of-the-art tool like \texttt{PyBoolNet}~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
-\texttt{mpbn}~\cite{Paulev2020} does not give any answer either because it recognizes that model as non-locally-monotonic.
-And hence, it is not possible to extract any (complex) attractor at all.
-This is also true for the Alzheimer model also mentioned in that same article and originally from~\cite{ogishima2016alzpathway} (\verb|F4| file in the original supplementary material, and ``Alzheimer'' in Table~\ref{tab:result_real}), but actually not for the MAPK model for which the first trap spaces can be obtained in reasonable time.
+\texttt{mpbn}~\cite{Paulev2020} can return the first 1000 solution within 1.43s, but indeed, it limits the modeling range of the modelers as it does not permit using non-locally-monotonic Boolean functions.
+This is also true for the Alzheimer model also mentioned in that same article and originally from~\cite{ogishima2016alzpathway} (\verb|F4| file in the original supplementary material, and ``Alzheimer'' in Table~\ref{tab:result_real}), where \texttt{PyBoolNet} also fails at the prime-implicant computation and \texttt{mpbn} does not give any answer because this model is actually non-locally-monotonic.
 The current practice usually revolves then around fixing some source nodes to plausible values and reducing the model accordingly.
@@
 
-Using the method presented above, it is possible to convert the model to \texttt{PNML} in about one second and to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) in a few milliseconds.
+Using the ASP-based method presented in Section~\ref{subsec:computation_asp}, it is possible to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) within 0.19s, which is much quicker than \texttt{mpbn}.
 Unfortunately since this was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.

### 2023-01-29 manuscripts/TCS2022/tcs2022.tex
emphasize tool names

@@
 
-Note that given a Boolean network in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the \texttt{bioLQM}\footnote{\url{http://www.colomoto.org/biolqm/}} library.
-This piece of software extracted from GINsim~\cite{chaouiya2012logical} and part of the CoLoMoTo\footnote{\url{http://colomoto.org/}}~\cite{naldi2015cooperative} software suite allows for easy conversion between standard formats.
+Note that given a Boolean network in the standard \texttt{SBML-Qual} format~\cite{chaouiya2013sbml}, i.e., the package of SBML v3~\cite{keating2020sbml} for such models, one can easily obtain its Petri net encoding in the Petri Net Markup Language  (PNML)\footnote{\url{https://www.pnml.org/}} standard using the \texttt{bioLQM}\footnote{\url{http://www.colomoto.org/biolqm/}} library.
+This piece of software extracted from \texttt{GINsim}~\cite{chaouiya2012logical} and part of the \texttt{CoLoMoTo}\footnote{\url{http://colomoto.org/}}~\cite{naldi2015cooperative} software suite allows for easy conversion between standard formats.
 It also accepts many other common formats for Boolean networks, notably the \verb|.bnet| files of the  BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} tools.
@@
 We set a soft clause for each variable of the CSP and then use a ``minimal correction subset'' blocking strategy, which will ensure set-inclusion maximality of the solutions.
-This is what is implemented in \texttt{Trappist} using the \texttt{RC2} MaxSAT solver~\cite{DBLP:journals/jsat/IgnatievMM19} available through the python-sat package\footnote{\url{https://pysathq.github.io/docs/html/api/examples/rc2.html}}.
+This is what is implemented in \texttt{Trappist} using the \texttt{RC2} MaxSAT solver~\cite{DBLP:journals/jsat/IgnatievMM19} available through the \texttt{python-sat} package\footnote{\url{https://pysathq.github.io/docs/html/api/examples/rc2.html}}.
 
@@
 Otherwise, \(v \not \in D_m\).
-Computing multiple answer sets is built into ASP solvers and the solving collection POTASSCO~\cite{DBLP:journals/aicom/GebserKKOSS11} also features the option to find set-inclusion maximal answer sets with respect to the set of atoms.
+Computing multiple answer sets is built into ASP solvers and the solving collection \texttt{POTASSCO}~\cite{DBLP:journals/aicom/GebserKKOSS11} also features the option to find set-inclusion maximal answer sets with respect to the set of atoms.
 Naturally, a set-inclusion maximal answer set of \(\mathcal{L}\) is equivalent to a maximal conflict-free siphon of \(\mathcal{P}\), thus a minimal trap space of \(\mathcal{N}\). 
@@
 Thus, all the maximal conflict-free siphons of the Petri net have been found.
-We use \dots as the ILP solver.
+We use \dots as the ILP solver. TODO: \dots
 
@@
 
-\subsection{BBM repository}
-
-Currently, a research group has made a great effort for building a collection (called BBM) of real-world Boolean models from various sources used in systems biology.
+\subsection{\texttt{BBM} repository}
+
+Currently, a research group has made a great effort for building a collection (called \texttt{BBM}) of real-world Boolean models from various sources used in systems biology.
 It aims to be a comprehensive collection suitable for benchmarking and testing new tools and methods.
@@
 
-Table~\ref{tab:bbm_repo} shows the experimental results on the 211 real-world models from the BBM repository.
-Columns 2 expresses the numbers of failures (i.e., did not finish the computation within a time limit of three minutes) of each method.
+Table~\ref{tab:bbm_repo} shows the experimental results on the 211 real-world models from the \texttt{BBM} repository.
+Column 2 expresses the numbers of failures (i.e., did not finish the computation within a time limit of three minutes) of each method.
 For the case of \texttt{bioLQM}, we only considered the models that have at most 1000 minimal trap spaces.

### 2023-01-29 manuscripts/TCS2022/tcs2022.tex
update experimental descriptions

@@
 As explained in Section~\ref{sec:case_study}, in our benchmarks, we only searched for the first 1000 minimal trap spaces for each model.
+It is worth noting that unlike existing analysis shown in the literature, we did not fix specific values for source nodes in all the considered models.
 
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (LQM), \texttt{PyBoolNet} (PBN), \texttt{mpbn} and the four variants of \texttt{Trappist} on the \texttt{PyBoolNet} repository.}
+  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (\texttt{LQM}), \texttt{PyBoolNet} (\texttt{PBN}), \texttt{mpbn} and the four variants of \texttt{Trappist} on the \texttt{PyBoolNet} repository.}
   \centering%
@@
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
+In the case of \texttt{bioLQM}, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of \texttt{bioLQM} because it always requires to compute all minimal trap spaces.
 A number in bold indicates a ratio greater than three compared to the best result. 
@@
 We used a set of real-world Boolean networks lying in various scales collected from numerous bibliographic sources.
-These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
-Note that these models are not included in the \texttt{PyBoolNet} and BBM repositories.
+Most of these models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and have never been fully analyzed.
+Note that these models are not included in the \texttt{PyBoolNet} and \texttt{BBM} repositories.
 We then applied \texttt{bioLQM}, \texttt{PyBoolNet}, \texttt{mpbn}, and the four variants of \texttt{Trappist} to computing minimal trap spaces of these real-world models.
-It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes in these models.
-Table~\ref{tab:result_real} shows the experimental results on those models. 
+Table~\ref{tab:result_real} shows the obtained experimental results. 
 ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces) within the timeout of two minutes.
-In the case of \texttt{bioLQM}, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of \texttt{bioLQM} because it always requires to compute all minimal trap spaces.
+%In the case of \texttt{bioLQM}, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of \texttt{bioLQM} because it always requires to compute all minimal trap spaces.
 A number in bold indicates a ratio greater than or equal to 10 compared to the best result.
-Other notations are similar to those in Table~\ref{tab:pyboolnet_repo}.
+The remaining notations are similar to those in Table~\ref{tab:pyboolnet_repo}.
 Hereafter, we analyze in detail the results with respect to minimal trap space computation.
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (LQM), \texttt{PyBoolNet} (PBN), \texttt{mpbn} and the four variants of \texttt{Trappist} on selected models from the literature.}
+  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (\texttt{LQM}), \texttt{PyBoolNet} (\texttt{PBN}), \texttt{mpbn} and the four variants of \texttt{Trappist} on selected models from the literature.}
   % in bold if greater than or equal to 10
@@
     \cmidrule(rr){5-8}
-    \(n\) & LQM & \texttt{mpbn} & PBN & SAT & CLP & ILP & ASP \\
+    \(n\) & \texttt{LQM} & \texttt{mpbn} & \texttt{PBN} & SAT & CLP & ILP & ASP \\
     \midrule
@@
 Column \(n\) denotes the network size.
-Columns ``LQM'' and ``PBN'' show the results of \texttt{bioLQM} and \texttt{PyBoolNet}, respectively.
+Columns \texttt{LQM} and \texttt{PBN} show the results of \texttt{bioLQM} and \texttt{PyBoolNet}, respectively.
 For each method, the number outside the parentheses indicates the number of failures, whereas the number inside the parentheses indicates the average running time (in seconds).
@@
 From these results, we obtained several observations consistent with those obtained for real-world models.
+
+TODO: \dots
 

### 2023-01-29 manuscripts/TCS2022/tcs2022.tex
revising conclusion

@@
 
-In this article we explored and proved for the first time the equivalence between (minimal) trap spaces of a general Boolean network and (maximal) conflict-free siphons of its Petri net encoding.
-From this equivalence, we proposed a new approach for the computation of minimal trap spaces of Boolean networks, based on the enumeration of maximal conflict-free siphons of Petri nets.
-We also proposed the four possible methods using MaxSAT, CLP, ILP, and ASP for implementing the new approach.
-The proposed methods were evaluated on many real-world models from the literature and randomly generated models.
-The experimental results show that \dots
+In this article we have explored and proved for the first time the equivalence between (minimal) trap spaces of a general Boolean network and (maximal) conflict-free siphons of its Petri net encoding.
+We have shown several important applications of this finding to studying properties of trap spaces in Boolean networks. 
+As an important practical application of the equivalence, we have proposed a new approach for the computation of minimal trap spaces of Boolean networks, based on the enumeration of maximal conflict-free siphons of Petri nets.
+We have also proposed the four possible methods using MaxSAT, CLP, ILP, and ASP for implementing the new approach.
+The proposed methods have been evaluated on many real-world models from the literature and randomly generated models.
+The experimental results show that the new approach vastly outperforms all the state-of-the-art methods in terms of general Boolean networks and is comparable to the \texttt{mpbn} method even better in average in term of locally-monotonic Boolean networks.
 We believe that this opens up the way to a much better analysis of large Boolean networks, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
@@
 
-% Compare among various possible methods and obtained insights
+% TODO: Compare among various possible methods and obtained insights
 Second, the experimental results shown in this paper mostly expose the differences caused by the prime-implicant phase of \texttt{PyBoolNet}.
@@
 
-% Update: can borrow many advents in the field of Petri nets
+% TODO: can borrow many advents in the field of Petri nets
 In addition, there are possibly other methods for computing maximal conflict-free siphons in Petri nets, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}.

### 2023-01-30 manuscripts/TCS2022/tcs2022.tex
sorted by network size
finish conclusion

@@
     \midrule % n < 100
+    \rownb & metastatic~\cite{DesignPrinciplesGeneNetworks} & 10 & 4 & \textbf{0.10} & 0.04 & NM & 0.01 & - & - & 0.02 \\
     \rownb & Arabidopsis\_thaliana~\cite{DesignPrinciplesGeneNetworks} & 15 & 8 & \textbf{0.10} & 0.06 & NM & 0.01 & - & - & 0.02 \\
-    \rownb & Bcell~\cite{Dutta2019} & 73 & 72 & 0.23 & 0.04 & 0.08 & 0.06 & - & - & 0.05 \\
-    \rownb & Cacace\_TdevModel~\cite{Cacace2020} & 61 & 28 & \textbf{1.29} & \textbf{5.67} & NM & 0.06 & - & - & 0.08 \\
-    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $>1000$ & N/A & \textbf{107.57} & 0.76 & 0.56 & - & - & 0.16 \\
+    \rownb & p53\_high\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.38 & \textbf{1.76} & NM & 0.08 & - & - & 0.14 \\
+    \rownb & p53\_low\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.41 & \textbf{1.76} & NM & 0.07 & - & - & 0.14 \\
+    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & 32 & \textbf{DNF} & \textbf{DNF} & NM & 0.03 & - & - & 0.19 \\
     \rownb & DNA\_damage~\cite{DesignPrinciplesGeneNetworks} & 26 & 16 & \textbf{0.24} & \textbf{0.33} & NM & 0.02 & - & - & 0.05 \\
-    \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & 128 & 0.33 & 0.05 & 0.07 & 0.06 & - & - & 0.05 \\
-    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & \textbf{39.22} & \textbf{1.01} & 0.20 & 0.12 & - & - & 0.05 \\
-    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $>1000$ & N/A & 0.09 & 0.55 & 0.37 & - & - & 0.15 \\
-    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & 32 & \textbf{DNF} & \textbf{DNF} & NM & 0.03 & - & - & 0.19 \\
-    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $>1000$ & N/A & \textbf{DNF} & 0.50 & 0.34 & - & - & 0.33 \\
-    \rownb & metastatic~\cite{DesignPrinciplesGeneNetworks} & 10 & 4 & \textbf{0.10} & 0.04 & NM & 0.01 & - & - & 0.02 \\
-    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $>1000$ & N/A & 0.11 & 0.36 & 0.17 & - & - & 0.06 \\
-    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & 276 & 0.37 & 0.43 & NM & 0.07 & - & - & 0.06 \\
     \rownb & Rho-GTPases~\cite{DesignPrinciplesGeneNetworks} & 33 & 2 & 0.17 & 0.57 & \textbf{40.39} & 0.07 & - & - & 0.11 \\
     \rownb & Pluripotency~\cite{YachieKinoshita2018} & 36 & 440 & \textbf{DNF} & \textbf{DNF} & NM & 0.16 & - & - & 0.28 \\
-    \rownb & p53\_high\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.38 & \textbf{1.76} & NM & 0.08 & - & - & 0.14 \\
-    \rownb & p53\_low\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.41 & \textbf{1.76} & NM & 0.07 & - & - & 0.14 \\
+    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & 276 & 0.37 & 0.43 & NM & 0.07 & - & - & 0.06 \\
+    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $>1000$ & N/A & 0.11 & 0.36 & 0.17 & - & - & 0.06 \\
+    \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & 128 & 0.33 & 0.05 & 0.07 & 0.06 & - & - & 0.05 \\
+    \rownb & Cacace\_TdevModel~\cite{Cacace2020} & 61 & 28 & \textbf{1.29} & \textbf{5.67} & NM & 0.06 & - & - & 0.08 \\
+    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $>1000$ & N/A & \textbf{DNF} & 0.50 & 0.34 & - & - & 0.33 \\
+    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & \textbf{39.22} & \textbf{1.01} & 0.20 & 0.12 & - & - & 0.05 \\
+    \rownb & Bcell~\cite{Dutta2019} & 73 & 72 & 0.23 & 0.04 & 0.08 & 0.06 & - & - & 0.05 \\
+    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $>1000$ & N/A & 0.09 & 0.55 & 0.37 & - & - & 0.15 \\
+    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $>1000$ & N/A & \textbf{107.57} & 0.76 & 0.56 & - & - & 0.16 \\
 
     \midrule % 100 <= n < 200
+     \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & 78 & \textbf{56.81} & \textbf{4.25} & 0.23 & 0.17 & - & - & 0.19 \\
+    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & 82 & \textbf{DNF} & \textbf{14.01} & 0.27 & 0.20 & - & - & 0.25 \\
+    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $>1000$ & N/A & 0.54 & 1.09 & 0.84 & - & - & 0.27 \\
+    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $>1000$ & N/A & 0.16 & 1.07 & 1.06 & - & - & 0.16 \\
     \rownb & angiofull~\cite{Weinstein2017} & 142 & $>1000$ & N/A & 0.17 & 1.06 & 0.88 & - & - & 0.23 \\
-    %\rownb & COVID\_19 & 111 & - & N/A & - & - & - & - & - & - \\
-    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & 82 & \textbf{DNF} & \textbf{14.01} & 0.27 & 0.20 & - & - & 0.25 \\
     \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & 492 & \textbf{DNF} & \textbf{11.28} & 0.78 & 0.69 & - & - & 0.35 \\
+    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $>1000$ & N/A & \textbf{DNF} & 1.43 & 1.55 & - & - & 0.19 \\
     \rownb & MAPK~\cite{aghamiri2020automated} & 181 & $>1000$ & N/A & \textbf{13.58} & 1.76 & 1.51 & - & - & 0.27 \\
-    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $>1000$ & N/A & 0.54 & 1.09 & 0.84 & - & - & 0.27 \\
-    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $>1000$ & N/A & \textbf{DNF} & 1.43 & 1.55 & - & - & 0.19 \\
-    \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & 78 & \textbf{56.81} & \textbf{4.25} & 0.23 & 0.17 & - & - & 0.19 \\
-    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $>1000$ & N/A & 0.16 & 1.07 & 1.06 & - & - & 0.16 \\
 
     \midrule % 200 <= n
+    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $>1000$ & N/A & 1.13 & 1.47 & 1.43 & - & - & 0.31 \\
     \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & $>1000$ & N/A & \textbf{DNF} & 1.52 & 2.26 & - & - & 0.35 \\
-    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $>1000$ & N/A & 1.13 & 1.47 & 1.43 & - & - & 0.31 \\
+    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $>1000$ & N/A & \textbf{4.99} & NM & 1.96 & - & - & 0.28 \\
     \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & $>1000$ & N/A & 0.42 & 2.36 & \textbf{4.91} & - & - & 0.44 \\
     \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & $>1000$ & N/A & \textbf{DNF} & 2.56 & \textbf{5.62} & - & - & 0.46 \\
-    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $>1000$ & N/A & \textbf{4.99} & NM & 1.96 & - & - & 0.28 \\
     \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & $>1000$ & N/A & 0.36 & 2.99 & \textbf{4.81} & - & - & 0.37 \\
@@
 The fact that our new ASP encoding is guaranteed to be linear in the number of nodes of the original model does not seem to be crucial here, however a much deeper analysis of those cases remains to be done.
-
-The last observation is that for very large models (i.e., more than two thousand nodes) the proposed method did not manage to finish the Petri net conversion before the timeout, as marked by ``???'' in Table~\ref{tab:result_real}.
-This points to the fact that our current choice of using a BDD-based translation to obtain that Petri net encoding, though it provides a small/efficient ASP might be too costly to handle the largest models.
-In such a case, a more \emph{naive} encoding might provide a much larger ASP program, with many redundant rules, but easier/faster to obtain.
-The evaluation of the feasibility of such strategy, and of its impact on smaller instances, remains to be done and is out of the scope of this first article.
-Recognizing that a model is locally-monotonic and applying in that specific case dedicated strategies as those of \texttt{mpbn} might also be a partial solution as shown in the SN-5 case, but not for the turei-2016 model.
 
@@
 
-%Though we benchmarked this new approach against state-of-the-art tools, there are many more evaluations that we plan to do in the future.
-%First, the \texttt{bioLQM} platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like \texttt{PyBoolNet}, that remains to be checked.
-%Note that this also raises the question of replacing altogether the \texttt{bioLQM} preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
-%The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it (in other words, the trade-off of allowing redundant constraints) has to be evaluated in depth.
-
-% TODO: Compare among various possible methods and obtained insights
-Second, the experimental results shown in this paper mostly expose the differences caused by the prime-implicant phase of \texttt{PyBoolNet}.
-There is much more information required to distinctly study the practical differences between our ASP encoding and that of \texttt{PyBoolNet}, and notably their size/efficiency ratio.
-Hence, we plan to conduct experiments on more real-world models or maybe random models that can be randomly generated by using BoolNet~\cite{mussel2010boolnet}.
-Such experiments should include a time-course of the number of ASP solutions found for proper comparison of the ASP encodings.
-
-% TODO: can borrow many advents in the field of Petri nets
-In addition, there are possibly other methods for computing maximal conflict-free siphons in Petri nets, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}.
-Although these approaches do not directly support the maximal conflict-free siphon computation now, we plan to investigate them in the future.
-They could replace our ASP program if they outperform it.
-However, the current method appears to already perform very well even on the biggest models we have considered.
+Although the experimental results show the superiority of our approach to \texttt{mpbn} in general, we however note that there is a model in the \texttt{BBM} repository (with identifier 122) where all the four proposed methods for the new approach did not manage to finish the Petri net conversion before the timeout, whereas \texttt{mpbn} can still handle this model.
+The model is not very large but its Boolean functions are rather complicated.
+This points to the fact that our current choice of using a BDD-based translation to obtain that Petri net encoding, though it provides a small/efficient ASP might be too costly to handle the complex models.
+In such a case, a more \emph{naive} encoding might provide a much larger ASP program, with many redundant rules, but easier/faster to obtain.
+The evaluation of the feasibility of such strategy, and of its impact on smaller instances, remains to be done.
+Recognizing that a model is locally-monotonic and applying in that specific case dedicated strategies as those of \texttt{mpbn} might also be a partial solution.
+
+It is worth noting that there may be possibly other methods for computing minimal/maximal conflict-free siphons in Petri nets, like the methods for generic siphon computation in the field of Petri nets (see~\cite{DBLP:journals/isci/LiuB16} for a survey about these methods).
+Although these approaches do not directly support the minimal/maximal conflict-free siphon computation now, we plan to investigate them in the future. 
+They could replace our proposed methods if they give significantly better performance.
+However, the current methods appear to already perform very well even on the biggest models we have considered.
 

### 2023-01-30 manuscripts/TCS2022/tcs2022.tex
update section on incomplete data problem

@@
 \end{subfigure}%
-\caption{Dynamics and encoding of the Boolean network of Example~\ref{example:BN}}%
+\caption{Dynamics and encoding of the Boolean network of Example~\ref{example:BN}.}%
 \label{fig:stg_and_PN}
@@
 So far we have assumed that we are always able to derive complete
-and consistent truth-tables (i.e., Boolean functions) that correctly capture the behavior of each node in a Boolean network. 
+and consistent truth tables (i.e., Boolean functions) that correctly capture the behavior of each node in a Boolean network. 
 However, in practice it is rarely the
@@
 The data provided may be incomplete in the sense that information is missing about what happens in certain states, or it may be inconsistent in that we have conflicting information.
-The result is that the behavior of some nodes under certain conditions may be unknown~\cite{Steggles2006}.
-
-Such incomplete and/or inconsistent behavioral information is
-problematic for Boolean networks with the synchronous update scheme, which are unable to represent the possibility of more than one next state.
+The result is that the behavior of some nodes under certain conditions may be non-deterministic~\cite{Steggles2006}.
+
+Non-deterministic behavior is problematic for Boolean networks with the synchronous update scheme, which are unable to represent the possibility of more than one next state.
 It can be addressed by using Boolean networks with asynchronous update schemes because of their non-determinism.
 However, multiple update functions may be considered for each node, which is problematic for existing analysis methods in Boolean networks.
-For example, if there are \(k\) nodes with two possible Boolean functions, then the analysis methods such as \texttt{PyBoolNet} and \texttt{mpbn} need to consider \(2^k\) possible Boolean networks due to they inputs are Boolean networks with complete Boolean functions.
+For example, if the truth table of a node have \(k\) unknown-output rows, then this node may have \(2^k\) possible Boolean functions and the analysis methods such as \texttt{PyBoolNet} and \texttt{mpbn} need to consider \(2^k\) possible Boolean networks due to their inputs are Boolean networks with complete Boolean functions.
 Moreover, one interesting question maybe raised from systems biologists is if there is a common behavioral property (e.g., minimal trap space, attractor) among all possible combinations.
-To answer this question, the existing analysis methods may consider \(2^k\) possible Boolean networks as well.
+To answer this question, the existing analysis methods maybe need to consider \(2^k\) possible Boolean networks as well.
 
@@
 More specifically, for each unknown-output row in the truth table of a node, we create two conflicting transitions representing the two possible outputs (i.e., 0 or 1).
-For every remaining row, we create transitions as usual following the Petri net encoding~\cite{chatain2014characterization}.
+For the remaining rows, we create transitions as usual following the Petri net encoding~\cite{chatain2014characterization}.
 Because of the non–deterministic choice mechanisms of Petri nets, the new Petri net is still equivalent in dynamics to the incomplete Boolean network.
@@
 
-% TODO: finish the figure
 \begin{figure}[!htb]
@@
 	\end{subfigure}%
-	\caption{Asynchronous dynamics (self arcs are omitted for simplicity) and new Petri net encoding (the transitions corresponding to \(g_1\) and \(g_3\) are omitted for simplicity) of the incomplete Boolean network of Example~\ref{example:incomplete-BN}}%
+	\caption{Asynchronous dynamics (self arcs are omitted for simplicity) and new Petri net encoding (the transitions corresponding to \(g_1\) and \(g_3\) are omitted for simplicity) of the incomplete Boolean network of Example~\ref{example:incomplete-BN}.}%
 	\label{fig:incomplete-dynamics-new-PN}
@@
 This is actually a significant advantage, since we do not need to consider \(2^k\) possible Boolean networks.
-It is worth noting that the approach proposed in~\cite{Steggles2006} only deals with the case that \(k = 1\).
+It is worth noting that the approach proposed in~\cite{Steggles2006} is only shown to deal with the case that \(k = 1\), thus it is not clear how this approach can handle the case that \(k > 1\).
 Furthermore, as more data becomes available for the underlying biological system, the Petri net model can be refined to reduce the
 amount of non-determinism (i.e., the number of conflicting transitions) it contains.
-Hence, Petri nets provide an interesting means of documenting the development of knowledge about a biological system.
+Hence, Petri nets provide an interesting means of documenting the development of knowledge for a biological system.
 

### 2023-02-13 trappist/trappist.py
Small things to be precise

@@ -263,7 +263,7 @@ def main():
         type=argparse.FileType("r", encoding="utf-8"),
         nargs="?",
         default=sys.stdin,
-        help="Petri-net (PNML) file",
+        help="Boolnet (.bnet) file of the model or Petri-net (PNML) file of its Petri net encoding.",
     )
     args = parser.parse_args()
 

### 2023-02-15 manuscripts/TCS2022/tcs2022.tex
First pass (CP/ILP)

@@
-%% 
-%% Copyright 2007-2020 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for Elsevier's document class `elsarticle'
-%% with harvard style bibliographic references
-
 \documentclass[preprint,12pt]{elsarticle}
@@
 %% \documentclass[final,5p,times,twocolumn]{elsarticle}
-
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
 
@@
 
-\linenumbers
+\linenumbers%
 
@@
 Second, we showcase a theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
-Third, beyond the proposed ASP method implementing the alternative approach~\cite{DBLP:conf/cmsb/TrinhBHS22}, we propose several other possible methods for computing minimal trap spaces using Maximum Satisfiability (MaxSAT), Constraint Logic Programming (CLP), and Integer Linear Programming (ILP).
+Third, beyond the proposed ASP method implementing the alternative approach~\cite{DBLP:conf/cmsb/TrinhBHS22}, we propose several other possible methods for computing minimal trap spaces using Maximum Satisfiability (MaxSAT), Constraint Programming (CP), and Integer Linear Programming (ILP).
 Fourth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
@@
 
-\section{Preliminaries}
+\section{Preliminaries}%
 \label{sec:Preliminaries}
@@
 
-\begin{definition}
+\begin{definition}%
+\label{def:BN}
 
@@
 
-\label{def:BN}
 \end{definition}
@@
 This does not change the worst-case-complexity (obtaining a single DNF being exponential) but might matter a lot in practice.
-As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM}, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.
+As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM} and directly in our tool using the \texttt{pyeda}\footnote{\url{https://pyeda.readthedocs.io/en/latest/}} library, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.
 
@@
 
-\begin{proposition}
+\begin{proposition}%
 \label{prop:siphon_set}
@@
 
-\section{Minimal trap spaces as maximal conflict-free siphons}
+\section{Minimal trap spaces as maximal conflict-free siphons}%
 \label{sec:Main_finding}
@@
 
-\section{Computation methods}
+\section{Computation methods}%
 \label{sec:Computation}
 
-\subsection{Characterization}
+\subsection{Characterization}%
 \label{subsec:siphon_characterization}
@@
 
-\subsection{Constraint satisfaction problem}
+\subsection{Constraint satisfaction problem}%
 \label{subsec:computation_csp}
@@
 
-\begin{proposition}
+\begin{proposition}%
 \label{prop:csp_conflict_free_siphon}
@@
 For enumerating siphons in the set inclusion order, the proposed method by~\cite{nabli2016enumerating} uses the technique that labels directly the Boolean variables with increasing value selection (i.e., to test first the absence, then the presence of a place in the candidate solution).
-The method has two implementations, one uses an iterated SAT procedure and the other uses Constraint Logic Programming (CLP) with backtracking.
+The method has two implementations, one uses an iterated SAT procedure and the other uses Constraint Programming (CP) with backtracking.
 
@@
 The correctness of this technique comes from the fact that once \(S\) is found, it is the conflict-free siphon of maximum cardinality among all the remaining feasible conflict-free siphons.
-Similar to~\cite{nabli2016enumerating}, the newly CSP-based method can also be implemented with SAT and CLP solvers.
+Similar to~\cite{nabli2016enumerating}, the newly CSP-based method can also be implemented with SAT and CP solvers.
+
+This method was implemented using the state-of-the-art CP solver Chuffed\footnote{\url{https://github.com/chuffed/chuffed}}~\cite{demirovic2018solution} via its MiniZinc~\cite{nethercote2007minizinc} interface.
+Because it is a high-level interface, the backtrack-and-replay method of~\cite{nabli2016enumerating} was not used but rather the alternative implementation with two global constraints for lexicographic ordering (ensuring enumeration of solutions) and iterated non-subset of each already found solution (for maximality).
 
@@
 
-For the CLP-based method, we use TODO: \dots
-
-\subsection{Answer set programming-based method}
+\subsection{Answer set programming-based method}%
 \label{subsec:computation_asp}
@@
 
-\subsection{Integer linear programming-based method}
+\subsection{Integer linear programming-based method}%
 \label{subsec:computation_ilp}
@@
 Thus, all the maximal conflict-free siphons of the Petri net have been found.
-We use \dots as the ILP solver. TODO: \dots
-
-\subsection{Computation of special types of trap spaces}
+
+Since we used the MiniZinc framework to interface with the CP solver, it was simple to make the slight modifications described above and use that same interface to call the Coin-OR CBC solver\footnote{\url{https://github.com/coin-or/Cbc}}~\cite{john_forrest_2022_6522795}.
+
+\subsection{Computation of special types of trap spaces}%
 \label{subsec:computation_special_ts}
@@
 We shall show that our proposed methods can be easily adjusted to compute popular types of trap spaces.
-We illustrate the adjustments via the ASP-based method (see Subsection~\ref{subsec:computation_asp}), but these adjustments are completely applicable for other approaches such as MaxSAT, CLP, and ILP.
+We illustrate the adjustments via the ASP-based method (see Subsection~\ref{subsec:computation_asp}), but these adjustments are completely applicable for other approaches such as MaxSAT, CP, and ILP\@.
 
@@
 Note that we can also build another ASP characterizing all fixed points of \(\mathcal{N}\) based on the equivalence between a fixed point of \(\mathcal{N}\) and a deadlock of its Petri net encoding~\cite{chatain2014characterization}.
-This approach may give a more compact ASP.
+This approach may give a more compact ASP\@.
 
@@
 
-\section{Inconsistent and incomplete data}
+\section{Inconsistent and incomplete data}%
 \label{sec:incomplete-data}
@@
 
-\label{example:incomplete-BN}
+\label{example:incomplete-BN}%
 \end{example}
@@
 
-\section{Motivating example}
+\section{Motivating example}%
 \label{sec:case_study}
@@
 Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by \texttt{PyBoolNet}).
-We ran all the benchmarks on a machine whose environment is CPU: \@ Intel\textregistered\ Core\texttrademark\ i9-11950H 2.60GHz × 16, 16 GB DDR4 RAM, Ubuntu 20.04.5 LTS. 
+We ran all the benchmarks on a machine whose environment is CPU\@: Intel\textregistered\ Core\texttrademark\ i9-11950H 2.60GHz × 16, 16 GB DDR4 RAM, Ubuntu 20.04.5 LTS\@.
 Finally, we set a time limit of three minutes for each model.
@@
     \cmidrule(rr){8-11}
-    & model & \(n\) & \(|M|\) & LQM & PBN & \texttt{mpbn} & SAT & CLP & ILP & ASP \\
+    & model & \(n\) & \(|M|\) & LQM & PBN & \texttt{mpbn} & SAT & CP & ILP & ASP \\
     \midrule
@@
 ``NM'' indicates a non-locally-monotonic model.
-There are four variants of \texttt{Trappist}: SAT (i.e., the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), CLP (i.e., the CLP-based method shown in Subsection~\ref{subsec:computation_csp}), ILP (i.e., the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and ASP (i.e., the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
+There are four variants of \texttt{Trappist}: SAT (i.e., the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), CP (i.e., the CP-based method shown in Subsection~\ref{subsec:computation_csp}), ILP (i.e., the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and ASP (i.e., the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
 
@@
     \texttt{Trappist}-MaxSAT & 1 & 0.03 & 1.09 & 1.01 \\
-    \texttt{Trappist}-CLP & - & - & - & - \\
+    \texttt{Trappist}-CP & - & - & - & - \\
     \texttt{Trappist}-ILP & - & - & - & - \\
@@
     \cmidrule(rr){8-11}
-    & model & \(n\) & \(|M|\) & LQM & PBN & \texttt{mpbn} & SAT & CLP & ILP & ASP \\
+    & model & \(n\) & \(|M|\) & LQM & PBN & \texttt{mpbn} & SAT & CP & ILP & ASP \\
     \midrule % n < 100
@@
 Note however that \texttt{mpbn} was the only tool to provide a solution for the SN-5 model, thus confirming that if the activation function is in the right form, not having to compute the inactivation function's disjunctive normal form can render a difficult problem tractable.
-However, since mbpn can handle only locally-monotonic models and \texttt{Trappist} can handle general models, it is difficult to further compare between them.
+However, since \texttt{mbpn} can handle only locally-monotonic models and \texttt{Trappist} can handle general models, it is difficult to further compare between them.
 Hence, we focus on only comparisons between \texttt{PyBoolNet} and \texttt{Trappist} in the following observations.
@@
 We randomly generated a set of N-K models~\cite{glass1973logical} with network size \(n\) in the set \{100, 150, 200, 250, 300, 350, 400\} and \(K = 3\) (i.e., each node has exactly three input nodes).
-We chose N-K models because they are a useful tool for studying the dynamics of Boolean networks~\cite{glass1973logical,klarner2015computing}).
+We chose N-K models because they are a useful tool for studying the dynamics of Boolean networks~\cite{glass1973logical,klarner2015computing}.
 For each network size, 50 instances were generated using the \verb|generateRandomNKNetwork| function.
@@
     \cmidrule(rr){5-8}
-    \(n\) & \texttt{LQM} & \texttt{mpbn} & \texttt{PBN} & SAT & CLP & ILP & ASP \\
+    \(n\) & \texttt{LQM} & \texttt{mpbn} & \texttt{PBN} & SAT & CP & ILP & ASP \\
     \midrule
@@
 
-\section{Conclusion}
+\section{Conclusion}%
 \label{sec:Conclusion}
@@
 As an important practical application of the equivalence, we have proposed a new approach for the computation of minimal trap spaces of Boolean networks, based on the enumeration of maximal conflict-free siphons of Petri nets.
-We have also proposed the four possible methods using MaxSAT, CLP, ILP, and ASP for implementing the new approach.
+We have also proposed the four possible methods using MaxSAT, CP, ILP, and ASP for implementing the new approach.
 The proposed methods have been evaluated on many real-world models from the literature and randomly generated models.

### 2023-02-17 trappist/__init__.py
bump python version

@@ -18,7 +18,7 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 # Don't forget to git tag v<version> && git push --tags
-version = "0.7.0"  # pylint: disable=invalid-name
+version = "0.8.0"  # pylint: disable=invalid-name
 
 
 def pnml_to_asp(name: str) -> str:

### 2023-02-17 trappist/cp.py

@@ -81,7 +81,7 @@ def solve_cp(
         max_output == 0 or nsol < max_output
     ):
         start = perf_counter()
-        #result = inst.solve(timeout=remaining, processes=nprocs)
+        # result = inst.solve(timeout=remaining, processes=nprocs)
         result = inst.solve_async(timeout=remaining, processes=nprocs)
         end = perf_counter()
         if remaining is not None:

### 2023-02-17 trappist/ilp.py

@@ -21,11 +21,12 @@ from datetime import timedelta
 from time import perf_counter
 from typing import Generator, List
 
-from minizinc import Instance, Model, Result, Solver, Status
+from minizinc import Instance, Model, Solver, Status
 
 import networkx as nx  # TODO maybe replace with lists/dicts
 
 from .cp import cp_to_bool, zincify
+
 
 def create_ilp(petri_net: nx.DiGraph) -> Model:
     """Create the ILP program for the max conflict-free siphons of petri net."""
@@ -49,9 +50,7 @@ def create_ilp(petri_net: nx.DiGraph) -> Model:
                 if zsucc not in or_preds:  # optimize obvious tautologies
                     or_string = " + ".join(or_preds)
                     model.add_string(f"constraint {zsucc} <= {or_string};\n")
-    model.add_string(
-        f"solve maximize {' + '.join(variables)};\n"
-    )
+    model.add_string(f"solve maximize {' + '.join(variables)};\n")
     return model
 
 

### 2023-02-18 manuscripts/TCS2022/tcs2022.tex
position with ensembles of Boolean networks

@@
 \[
-  \verb|p-v ; n-v.|
+  \verb|p-v; n-v.|
 \]
@@
 However, multiple update functions may be considered for each node, which is problematic for existing analysis methods in Boolean networks.
-For example, if the truth table of a node have \(k\) unknown-output rows, then this node may have \(2^k\) possible Boolean functions and the analysis methods such as \texttt{PyBoolNet} and \texttt{mpbn} need to consider \(2^k\) possible Boolean networks due to their inputs are Boolean networks with complete Boolean functions.
-Moreover, one interesting question maybe raised from systems biologists is if there is a common behavioral property (e.g., minimal trap space, attractor) among all possible combinations.
+For example, if the truth table of a node have \(k\) unknown-output rows, then this node may have \(2^k\) possible Boolean functions and the analysis methods such as \texttt{PyBoolNet} and \texttt{mpbn} need to consider \(2^k\) possible Boolean networks due to they require complete Boolean functions for their underlying processing.
+Moreover, one interesting question maybe raised from systems biologists is if there is a common behavioral property (e.g., minimal trap space, fixed point, attractor) among all possible combinations.
 To answer this question, the existing analysis methods maybe need to consider \(2^k\) possible Boolean networks as well.
@@
 The asynchronous dynamics of the incomplete Boolean network is shown in Figure~\ref{fig:incomplete-dynamics-new-PN}a.
-The result Petri net following our idea is given in Figure~\ref{fig:incomplete-dynamics-new-PN}b.
+The resulting Petri net following our idea is given in Figure~\ref{fig:incomplete-dynamics-new-PN}b.
 Herein, \(t^3_{g_2}\) and \(t^4_{g_2}\) are the two conflicting transitions representing the unknown-output row.
@@
 It is worth noting that the approach proposed in~\cite{Steggles2006} is only shown to deal with the case that \(k = 1\), thus it is not clear how this approach can handle the case that \(k > 1\).
-Furthermore, as more data becomes available for the underlying biological system, the Petri net model can be refined to reduce the
+Moreover, in the context of Boolean models construction of biological systems from prior knowledge and experimental data, the work by~\cite{chevalier2020synthesis} also deals with some kind of uncertainty.
+It uses ASP to represent all possible Boolean networks possessing the same given property, where each solution of this ASP is equivalent to a distinct Boolean network.
+Then it samples through that solution space with heuristics to select \emph{diverse} Boolean networks, and proceed further simulations or analysis on every individual Boolean network.
+Compared to our Petri net approach presented above, it still requires to consider multiple Boolean networks as well as it is only applicable for locally-monotonic Boolean networks~\cite{chevalier2020synthesis}.
+
+To this end, as more data becomes available for the underlying biological system, the Petri net model can be refined to reduce the
 amount of non-determinism (i.e., the number of conflicting transitions) it contains.
-Hence, Petri nets provide an interesting means of documenting the development of knowledge for a biological system.
+Specifically, if we know the output for an unknown-output row of a node, we can remove one of the two conflicting transitions corresponding to this unknown-output row.
+Hence, Petri nets provide an interesting means of documenting the development of knowledge for a biological system, and our Petri net-based approach can be more useful in the model refinement loop.
 

### 2023-02-18 manuscripts/TCS2022/tcs2022.tex
remove incomplete and inconsistent data section

--- 2023-02-15 manuscripts/TCS2022/tcs2022.tex
@@
 Fourth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
-Fifth, we present the idea for using our Petri net approach to handle the problem of inconsistent and incomplete data in modeling biological systems.
-Sixth, regarding the implementation, we have developed a new converter that directly reads a \texttt{.bnet} file and builds the Petri net encoding, instead of using the \texttt{PNML} conversion of \texttt{bioLQM}~\cite{DBLP:conf/cmsb/TrinhBHS22}.
+%Fifth, we present the idea for using our Petri net approach to handle the problem of inconsistent and incomplete data in modeling biological systems.
+Fifth, regarding the implementation, we have developed a new converter that directly reads a \texttt{.bnet} file and builds the Petri net encoding, instead of using the \texttt{PNML} conversion of \texttt{bioLQM}~\cite{DBLP:conf/cmsb/TrinhBHS22}.
 Finally, we conduct a more comprehensive benchmark on more real-world models from various sources and randomly generated models to evaluate all the proposed methods (the benchmark conducted in~\cite{DBLP:conf/cmsb/TrinhBHS22} considers only real-world models).
@@
 Section~\ref{sec:Computation} presents the alternative approach for computing minimal trap spaces and the four possible methods implementing it.
-Section~\ref{sec:incomplete-data} presents the idea to deal with problem of inconsistent and incomplete data.
+%Section~\ref{sec:incomplete-data} presents the idea to deal with problem of inconsistent and incomplete data.
 Section~\ref{sec:case_study} shows an important biological case study showing the applicability of the new approach.
@@
 
-\section{Inconsistent and incomplete data}%
-\label{sec:incomplete-data}
-
-So far we have assumed that we are always able to derive complete
-and consistent truth tables (i.e., Boolean functions) that correctly capture the behavior of each node in a Boolean network.
-However, in practice it is rarely the
-case that a Boolean network of a biological system is fully understood and indeed, this is one important reason for modeling such networks.
-The data provided may be incomplete in the sense that information is missing about what happens in certain states, or it may be inconsistent in that we have conflicting information.
-The result is that the behavior of some nodes under certain conditions may be non-deterministic~\cite{Steggles2006}.
-
-Non-deterministic behavior is problematic for Boolean networks with the synchronous update scheme, which are unable to represent the possibility of more than one next state.
-It can be addressed by using Boolean networks with asynchronous update schemes because of their non-determinism.
-However, multiple update functions may be considered for each node, which is problematic for existing analysis methods in Boolean networks.
-For example, if the truth table of a node have \(k\) unknown-output rows, then this node may have \(2^k\) possible Boolean functions and the analysis methods such as \texttt{PyBoolNet} and \texttt{mpbn} need to consider \(2^k\) possible Boolean networks due to their inputs are Boolean networks with complete Boolean functions.
-Moreover, one interesting question maybe raised from systems biologists is if there is a common behavioral property (e.g., minimal trap space, attractor) among all possible combinations.
-To answer this question, the existing analysis methods maybe need to consider \(2^k\) possible Boolean networks as well.
-
-We shall show how our Petri net approach presented the previous section can handle the problem of inconsistent and incomplete data.
-The idea is to build the Petri net encoding of the Boolean network from its partial truth tables.
-More specifically, for each unknown-output row in the truth table of a node, we create two conflicting transitions representing the two possible outputs (i.e., 0 or 1).
-For the remaining rows, we create transitions as usual following the Petri net encoding~\cite{chatain2014characterization}.
-Because of the non–deterministic choice mechanisms of Petri nets, the new Petri net is still equivalent in dynamics to the incomplete Boolean network.
-One notable advantage of the connection between trap spaces of a Boolean network and conflict-free siphons of its Petri net encoding is that it still holds if the Petri net encoding changes (e.g., more or less transitions) but still maintain the equivalence in dynamics to the Boolean network.
-Hence, the minimal trap spaces of the incomplete Boolean network can be computed by computing maximal conflict-free siphons of the resulting Petri net.
-As a consequence, we can easily answer the question on common behavioral property.
-
-\begin{example}
-Consider an incomplete Boolean network \(\mathcal{N} = (V, F)\) (mentioned in~\cite{Steggles2006}), where \(V = (g_1, g_2, g_3)\) and \(F = (f_1, f_2, f_3)\) with \(f_1 = g_2, f_3 = \neg g_1\).
+%\section{Inconsistent and incomplete data}%
+%\label{sec:incomplete-data}
+%
+%So far we have assumed that we are always able to derive complete
+%and consistent truth tables (i.e., Boolean functions) that correctly capture the behavior of each node in a Boolean network.
+%However, in practice it is rarely the
+%case that a Boolean network of a biological system is fully understood and indeed, this is one important reason for modeling such networks.
+%The data provided may be incomplete in the sense that information is missing about what happens in certain states, or it may be inconsistent in that we have conflicting information.
+%The result is that the behavior of some nodes under certain conditions may be non-deterministic~\cite{Steggles2006}.
+%
+%Non-deterministic behavior is problematic for Boolean networks with the synchronous update scheme, which are unable to represent the possibility of more than one next state.
+%It can be addressed by using Boolean networks with asynchronous update schemes because of their non-determinism.
+%However, multiple update functions may be considered for each node, which is problematic for existing analysis methods in Boolean networks.
+%For example, if the truth table of a node have \(k\) unknown-output rows, then this node may have \(2^k\) possible Boolean functions and the analysis methods such as \texttt{PyBoolNet} and \texttt{mpbn} need to consider \(2^k\) possible Boolean networks due to they require complete Boolean functions for their underlying processing.
+%Moreover, one interesting question maybe raised from systems biologists is if there is a common behavioral property (e.g., minimal trap space, fixed point, attractor) among all possible combinations.
+%To answer this question, the existing analysis methods maybe need to consider \(2^k\) possible Boolean networks as well.
+%
+%We shall show how our Petri net approach presented the previous section can handle the problem of inconsistent and incomplete data.
+%The idea is to build the Petri net encoding of the Boolean network from its partial truth tables.
+%More specifically, for each unknown-output row in the truth table of a node, we create two conflicting transitions representing the two possible outputs (i.e., 0 or 1).
+%For the remaining rows, we create transitions as usual following the Petri net encoding~\cite{chatain2014characterization}.
+%Because of the non–deterministic choice mechanisms of Petri nets, the new Petri net is still equivalent in dynamics to the incomplete Boolean network.
+%One notable advantage of the connection between trap spaces of a Boolean network and conflict-free siphons of its Petri net encoding is that it still holds if the Petri net encoding changes (e.g., more or less transitions) but still maintain the equivalence in dynamics to the Boolean network.
+%Hence, the minimal trap spaces of the incomplete Boolean network can be computed by computing maximal conflict-free siphons of the resulting Petri net.
+%As a consequence, we can easily answer the question on common behavioral property.
+%
+%\begin{example}
+%Consider an incomplete Boolean network \(\mathcal{N} = (V, F)\) (mentioned in~\cite{Steggles2006}), where \(V = (g_1, g_2, g_3)\) and \(F = (f_1, f_2, f_3)\) with \(f_1 = g_2, f_3 = \neg g_1\).
 Herein, \(f_2\) is an incomplete function and the partial truth table is as follows.
-\begin{center}
-	\begin{tabular}{rr|r}
-		\toprule
-		$g_1$ & $g_3$ & $g_2$ \\ \midrule
-		0 & 0 & 0 \\
-		0 & 1 & 0 \\
-		1 & 0 & ? \\
-		1 & 1 & 1 \\
-		\bottomrule
-	\end{tabular}
-\end{center}
-We have two possible update functions for \(g_2\): \(f_2 = g_1 \land g_3\) for the case that \(? = 0\) and \(f_2 = g_1\) for the case that \(? = 1\).
-Accordingly, we have two possible Boolean networks.
-The first one has one minimal trap space: 001.
-The second one has two minimal trap spaces: 001 and 110.
-Both have a common minimal trap space: 001.
-The asynchronous dynamics of the incomplete Boolean network is shown in Figure~\ref{fig:incomplete-dynamics-new-PN}a.
-The result Petri net following our idea is given in Figure~\ref{fig:incomplete-dynamics-new-PN}b.
-Herein, \(t^3_{g_2}\) and \(t^4_{g_2}\) are the two conflicting transitions representing the unknown-output row.
-This Petri net has only one maximal conflict-free siphon \(\{p_{g_1}, p_{g_2}, \overline{p}_{g_3}\}\) corresponding to the common minimal trap space 001.
-
-\label{example:incomplete-BN}%
-\end{example}
-
-\begin{figure}[!htb]
-	\centering
-	\begin{subfigure}[b]{0.3\textwidth}
-		\centering
-		\begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]
-		\node[] (0) [] {000};
-		\node[] (1) [left=of 0] {001};
-		\node[] (4) [above=of 0] {100};
-		\node[] (6) [above=of 1] {110};
-		\node[] (5) [below=of 1] {101};
-		\node[] (2) [left=of 6] {010};
-		\node[] (3) [left=of 1] {011};
-		\node[] (7) [below=of 3] {111};
-
-		\draw[->] (0) edge [] (1);
-		\draw[->] (4) edge [] (0);
-		\draw[->] (4) edge [bend left] (6);
-		\draw[->] (6) edge [bend left] (4);
-
-		\draw[->] (5) edge [] (1);
-		\draw[->] (5) edge [] (7);
-
-		\draw[->] (2) edge [] (6);
-		\draw[->] (2) edge [] (0);
-		\draw[->] (2) edge [] (3);
-
-		\draw[->] (3) edge [] (1);
-		\draw[->] (3) edge [] (7);
-
-		\draw[->] (7) edge [] (6);
-
-		\end{tikzpicture}
-		\caption{}%
-		%\label{fig:stg}
-	\end{subfigure}\hfill%
-	\begin{subfigure}[b]{0.6\textwidth}
-		\centering
-		\begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]\footnotesize
-		\node[circle,draw,label=left:$p_{g_1}$] (x1) {};
-		\node[circle,draw,label=left:$\overline{p}_{g_1}$] (nx1) [below=of x1, yshift=-0.5cm] {};
-
-		\node[circle,draw,label=right:$p_{g_2}$] (x2) [right=of x1, xshift=2.5cm] {};
-		\node[circle,draw,label=right:$\overline{p}_{g_2}$] (nx2) [below=of x2, yshift=-0.5cm] {};
-
-		\node[circle,draw,label=left:$p_{g_3}$] (x3) [below=of nx1, yshift=-0.5cm] {};
-		\node[circle,draw,label=right:$\overline{p}_{g_3}$] (nx3) [below=of nx2, yshift=-0.5cm] {};
-
-		\node[rectangle,draw,label=above:$t^1_{g_2}$] (t1x2) [right=of x1] {};
-		\node[rectangle,draw,label=above:$t^2_{g_2}$] (t2x2) [right=of nx1] {};
-
-		\node[rectangle,draw,label=left:$t^3_{g_2}$] (t3x2) [left=of nx2] {};
-		\node[rectangle,draw,label=left:$t^4_{g_2}$] (t4x2) [left=of nx3] {};
-
-		\draw[->] (x2) -- (t1x2);
-		\draw[->] (t1x2) -- (nx2);
-		\draw[<->] (t1x2) -- (nx1);
-
-		\draw[->] (nx2) edge [bend left] (t2x2);
-		\draw[->] (t2x2) -- (x2);
-		\draw[<->] (t2x2) -- (x1);
-		\draw[<->] (t2x2) -- (x3);
-
-		\draw[->] (x2) -- (t3x2);
-		\draw[->] (t3x2) -- (nx2);
-		\draw[<->] (t3x2) -- (x1);
-		\draw[<->] (t3x2) -- (nx3);
-
-		\draw[->] (nx2) -- (t4x2);
-		\draw[->] (t4x2) -- (x2);
-		\draw[<->] (t4x2) edge [bend left=15] (x1);
-		\draw[<->] (t4x2) -- (nx3);
-
-		\end{tikzpicture}
-		\caption{}%
-		%\label{fig:PN}
-	\end{subfigure}%
-	\caption{Asynchronous dynamics (self arcs are omitted for simplicity) and new Petri net encoding (the transitions corresponding to \(g_1\) and \(g_3\) are omitted for simplicity) of the incomplete Boolean network of Example~\ref{example:incomplete-BN}.}%
-	\label{fig:incomplete-dynamics-new-PN}
-\end{figure}
-
-In summary, we can still perform meaningful analysis on the resulting Petri net.
-Note that the new Petri net may have more transitions, however the number of additional transitions is only \(O(2k)\) where \(k\) is the number of unknown-output rows in the truth tables.
-This is actually a significant advantage, since we do not need to consider \(2^k\) possible Boolean networks.
-It is worth noting that the approach proposed in~\cite{Steggles2006} is only shown to deal with the case that \(k = 1\), thus it is not clear how this approach can handle the case that \(k > 1\).
-Furthermore, as more data becomes available for the underlying biological system, the Petri net model can be refined to reduce the
-amount of non-determinism (i.e., the number of conflicting transitions) it contains.
-Hence, Petri nets provide an interesting means of documenting the development of knowledge for a biological system.
+%\begin{center}
+%	\begin{tabular}{rr|r}
+%		\toprule
+%		$g_1$ & $g_3$ & $g_2$ \\ \midrule
+%		0 & 0 & 0 \\
+%		0 & 1 & 0 \\
+%		1 & 0 & ? \\
+%		1 & 1 & 1 \\
+%		\bottomrule
+%	\end{tabular}
+%\end{center}
+%We have two possible update functions for \(g_2\): \(f_2 = g_1 \land g_3\) for the case that \(? = 0\) and \(f_2 = g_1\) for the case that \(? = 1\).
+%Accordingly, we have two possible Boolean networks.
+%The first one has one minimal trap space: 001.
+%The second one has two minimal trap spaces: 001 and 110.
+%Both have a common minimal trap space: 001.
+%The asynchronous dynamics of the incomplete Boolean network is shown in Figure~\ref{fig:incomplete-dynamics-new-PN}a.
+%The resulting Petri net following our idea is given in Figure~\ref{fig:incomplete-dynamics-new-PN}b.
+%Herein, \(t^3_{g_2}\) and \(t^4_{g_2}\) are the two conflicting transitions representing the unknown-output row.
+%This Petri net has only one maximal conflict-free siphon \(\{p_{g_1}, p_{g_2}, \overline{p}_{g_3}\}\) corresponding to the common minimal trap space 001.
+%
+%\label{example:incomplete-BN}%
+%\end{example}
+%
+%\begin{figure}[!htb]
+%	\centering
+%	\begin{subfigure}[b]{0.3\textwidth}
+%		\centering
+%		\begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]
+%		\node[] (0) [] {000};
+%		\node[] (1) [left=of 0] {001};
+%		\node[] (4) [above=of 0] {100};
+%		\node[] (6) [above=of 1] {110};
+%		\node[] (5) [below=of 1] {101};
+%		\node[] (2) [left=of 6] {010};
+%		\node[] (3) [left=of 1] {011};
+%		\node[] (7) [below=of 3] {111};
+%
+%		\draw[->] (0) edge [] (1);
+%		\draw[->] (4) edge [] (0);
+%		\draw[->] (4) edge [bend left] (6);
+%		\draw[->] (6) edge [bend left] (4);
+%
+%		\draw[->] (5) edge [] (1);
+%		\draw[->] (5) edge [] (7);
+%
+%		\draw[->] (2) edge [] (6);
+%		\draw[->] (2) edge [] (0);
+%		\draw[->] (2) edge [] (3);
+%
+%		\draw[->] (3) edge [] (1);
+%		\draw[->] (3) edge [] (7);
+%
+%		\draw[->] (7) edge [] (6);
+%
+%		\end{tikzpicture}
+%		\caption{}%
+%		%\label{fig:stg}
+%	\end{subfigure}\hfill%
+%	\begin{subfigure}[b]{0.6\textwidth}
+%		\centering
+%		\begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]\footnotesize
+%		\node[circle,draw,label=left:$p_{g_1}$] (x1) {};
+%		\node[circle,draw,label=left:$\overline{p}_{g_1}$] (nx1) [below=of x1, yshift=-0.5cm] {};
+%
+%		\node[circle,draw,label=right:$p_{g_2}$] (x2) [right=of x1, xshift=2.5cm] {};
+%		\node[circle,draw,label=right:$\overline{p}_{g_2}$] (nx2) [below=of x2, yshift=-0.5cm] {};
+%
+%		\node[circle,draw,label=left:$p_{g_3}$] (x3) [below=of nx1, yshift=-0.5cm] {};
+%		\node[circle,draw,label=right:$\overline{p}_{g_3}$] (nx3) [below=of nx2, yshift=-0.5cm] {};
+%
+%		\node[rectangle,draw,label=above:$t^1_{g_2}$] (t1x2) [right=of x1] {};
+%		\node[rectangle,draw,label=above:$t^2_{g_2}$] (t2x2) [right=of nx1] {};
+%
+%		\node[rectangle,draw,label=left:$t^3_{g_2}$] (t3x2) [left=of nx2] {};
+%		\node[rectangle,draw,label=left:$t^4_{g_2}$] (t4x2) [left=of nx3] {};
+%
+%		\draw[->] (x2) -- (t1x2);
+%		\draw[->] (t1x2) -- (nx2);
+%		\draw[<->] (t1x2) -- (nx1);
+%
+%		\draw[->] (nx2) edge [bend left] (t2x2);
+%		\draw[->] (t2x2) -- (x2);
+%		\draw[<->] (t2x2) -- (x1);
+%		\draw[<->] (t2x2) -- (x3);
+%
+%		\draw[->] (x2) -- (t3x2);
+%		\draw[->] (t3x2) -- (nx2);
+%		\draw[<->] (t3x2) -- (x1);
+%		\draw[<->] (t3x2) -- (nx3);
+%
+%		\draw[->] (nx2) -- (t4x2);
+%		\draw[->] (t4x2) -- (x2);
+%		\draw[<->] (t4x2) edge [bend left=15] (x1);
+%		\draw[<->] (t4x2) -- (nx3);
+%
+%		\end{tikzpicture}
+%		\caption{}%
+%		%\label{fig:PN}
+%	\end{subfigure}%
+%	\caption{Asynchronous dynamics (self arcs are omitted for simplicity) and new Petri net encoding (the transitions corresponding to \(g_1\) and \(g_3\) are omitted for simplicity) of the incomplete Boolean network of Example~\ref{example:incomplete-BN}.}%
+%	\label{fig:incomplete-dynamics-new-PN}
+%\end{figure}
+%
+%In summary, we can still perform meaningful analysis on the resulting Petri net.
+%Note that the new Petri net may have more transitions, however the number of additional transitions is only \(O(2k)\) where \(k\) is the number of unknown-output rows in the truth tables.
+%This is actually a significant advantage, since we do not need to consider \(2^k\) possible Boolean networks.
+%It is worth noting that the approach proposed in~\cite{Steggles2006} is only shown to deal with the case that \(k = 1\), thus it is not clear how this approach can handle the case that \(k > 1\).
+%Moreover, in the context of Boolean models construction of biological systems from prior knowledge and experimental data, the work by~\cite{chevalier2020synthesis} also deals with some kind of uncertainty.
+%It uses ASP to represent all possible Boolean networks possessing the same given property, where each solution of this ASP is equivalent to a distinct Boolean network.
+%Then it samples through that solution space with heuristics to select \emph{diverse} Boolean networks, and proceed further simulations or analysis on every individual Boolean network.
+%Compared to our Petri net approach presented above, it still requires to consider multiple Boolean networks as well as it is only applicable for locally-monotonic Boolean networks~\cite{chevalier2020synthesis}.
+%
+%To this end, as more data becomes available for the underlying biological system, the Petri net model can be refined to reduce the
+%amount of non-determinism (i.e., the number of conflicting transitions) it contains.
+%Specifically, if we know the output for an unknown-output row of a node, we can remove one of the two conflicting transitions corresponding to this unknown-output row.
+%Hence, Petri nets provide an interesting means of documenting the development of knowledge for a biological system, and our Petri net-based approach can be more useful in the model refinement loop.
 

### 2023-02-18 manuscripts/TCS2022/tcs2022.tex
add a new point to  conclusion

@@
 We have shown several important applications of this finding to studying properties of trap spaces in Boolean networks.
-As an important practical application of the equivalence, we have proposed a new approach for the computation of minimal trap spaces of Boolean networks, based on the enumeration of maximal conflict-free siphons of Petri nets.
+As an important practical application of the equivalence, we have proposed a new approach for the computation of minimal trap spaces in Boolean networks, based on the enumeration of maximal conflict-free siphons of Petri nets.
 We have also proposed the four possible methods using MaxSAT, CP, ILP, and ASP for implementing the new approach.
-The proposed methods have been evaluated on many real-world models from the literature and randomly generated models.
-The experimental results show that the new approach vastly outperforms all the state-of-the-art methods in terms of general Boolean networks and is comparable to the \texttt{mpbn} method even better in average in term of locally-monotonic Boolean networks.
+The proposed methods have been evaluated on many real-world models from the literature as well as randomly generated models.
+The experimental results show that the new approach vastly outperforms all the state-of-the-art methods in terms of general Boolean networks and is comparable to the \texttt{mpbn} method even better in average in terms of locally-monotonic Boolean networks.
 We believe that this opens up the way to a much better analysis of large Boolean networks, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
@@
 Exploring those connections might lead both to interesting topics of research for Petri nets, like a notion of trap-spaces, and for Boolean networks.
+We also believe that the connection between trap spaces of Boolean networks and siphons of Petri nets can be a very useful tool for exploring and proving more new properties of trap spaces in Boolean networks, as we have used it to successfully prove the separation of minimal trap spaces.
+Diving into this direction is one of our future work.
 

### 2023-02-18 manuscripts/TCS2022/tcs2022.tex
experimental analysis on random models

@@
 
-First, \dots
-
-Second, \dots
-
-Third, \dots
+First, \texttt{mpbn} did not be able to handle any model because all the models are non-locally-monotonic.
+Recall that a Boolean network is non-locally-monotonic if only one of its Boolean functions is non-locally-monotonic.
+Hence, it is apparent that all the randomly generated models are non-locally-monotonic.
+This observation confirms the limit on the applicable model class of \texttt{mpbn}.
+
+Second, surprisingly \texttt{bioLQM} could not handle any model.
+One of the reason may be that the BDD characterizing all trap spaces is too large, and its computation is slow.
+It is apparent because the network size is large (\(\geq 100\)) and the Boolean functions are not simple.
+
+Third, \texttt{PyBoolNet} could handle every model of network size less than or equal to 300.
+It only failed in one model of network size 350 but 29 models of network size 400.
+The average running time vastly increases as the network size increases.
+As compared to the four methods of our approach, the performance of \texttt{PyBoolNet} is comparable for the 100-node and 200-node models.
+However, from \(n = 250\), the performance difference is exhibited more clearly.
 

### 2023-02-24 manuscripts/TCS2022/tcs2022.tex
small changes

@@
 Fourth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
+Besides minimal trap spaces, these special types also play crucial roles in analyzing Boolean networks~\cite{Rozum2021}.
 %Fifth, we present the idea for using our Petri net approach to handle the problem of inconsistent and incomplete data in modeling biological systems.
 Fifth, regarding the implementation, we have developed a new converter that directly reads a \texttt{.bnet} file and builds the Petri net encoding, instead of using the \texttt{PNML} conversion of \texttt{bioLQM}~\cite{DBLP:conf/cmsb/TrinhBHS22}.
-Finally, we conduct a more comprehensive benchmark on more real-world models from various sources and randomly generated models to evaluate all the proposed methods (the benchmark conducted in~\cite{DBLP:conf/cmsb/TrinhBHS22} considers only real-world models).
+Finally, we conduct a more comprehensive benchmark on more real-world models from various sources and randomly generated models to evaluate all the proposed methods (the benchmark conducted in~\cite{DBLP:conf/cmsb/TrinhBHS22} considers only dozens of representative real-world models).
 
@@
 %In the case of multi-level logical models, an encoding into a Boolean network is always possible~\cite{Didier2011}.
-\vangiang{Remove this statement because there is not sure if the encoded Boolean network preserves the trap spaces of the original multi-level logical model.}
+%\vangiang{Remove this statement because there is not sure if the encoded Boolean network preserves the trap spaces of the original multi-level logical model.}
 
@@
   \label{tab:pyboolnet_repo}
-  \scalebox{0.85}{
+  \scalebox{0.9}{
   \begin{tabular}{rlrrrrrrrrr}
@@
     \rownb & irons\_yeast & 18 & 1 & \textbf{0.12} & 0.03 & 0.01 & 0.01 & - & - & 0.02\\
-    \rownb & jaoude\_thdiff & 103 & $>1000$ & N/A & \textbf{0.85} & \textbf{0.45} & \textbf{0.56} & - & - & 0.09\\
+    \rownb & jaoude\_thdiff & 103 & $1000^+$ & N/A & \textbf{0.85} & \textbf{0.45} & \textbf{0.56} & - & - & 0.09\\
     \rownb & klamt\_tcr & 40 & 8 & \textbf{0.11} & 0.01 & 0.01 & 0.01 & - & - & 0.02\\
@@
     \rownb & saadatpour\_guardcell & 13 & 1 & \textbf{0.10} & 0.06 & 0.00 & 0.00 & - & - & 0.02\\
-    \rownb & selvaggio\_emt & 56 & $>1000$ & N/A & \textbf{0.48} & \textbf{0.28} & \textbf{0.28} & - & - & 0.09\\
+    \rownb & selvaggio\_emt & 56 & $1000^+$ & N/A & \textbf{0.48} & \textbf{0.28} & \textbf{0.28} & - & - & 0.09\\
     \rownb & tournier\_apoptosis & 12 & 3 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
@@
     \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & 276 & 0.37 & 0.43 & NM & 0.07 & - & - & 0.06 \\
-    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $>1000$ & N/A & 0.11 & 0.36 & 0.17 & - & - & 0.06 \\
+    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $1000^+$ & N/A & 0.11 & 0.36 & 0.17 & - & - & 0.06 \\
     \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & 128 & 0.33 & 0.05 & 0.07 & 0.06 & - & - & 0.05 \\
     \rownb & Cacace\_TdevModel~\cite{Cacace2020} & 61 & 28 & \textbf{1.29} & \textbf{5.67} & NM & 0.06 & - & - & 0.08 \\
-    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $>1000$ & N/A & \textbf{DNF} & 0.50 & 0.34 & - & - & 0.33 \\
+    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $1000^+$ & N/A & \textbf{DNF} & 0.50 & 0.34 & - & - & 0.33 \\
     \rownb & EMT~\cite{Rozum2021} & 69 & 268 & \textbf{39.22} & \textbf{1.01} & 0.20 & 0.12 & - & - & 0.05 \\
     \rownb & Bcell~\cite{Dutta2019} & 73 & 72 & 0.23 & 0.04 & 0.08 & 0.06 & - & - & 0.05 \\
-    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $>1000$ & N/A & 0.09 & 0.55 & 0.37 & - & - & 0.15 \\
-    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $>1000$ & N/A & \textbf{107.57} & 0.76 & 0.56 & - & - & 0.16 \\
+    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $1000^+$ & N/A & 0.09 & 0.55 & 0.37 & - & - & 0.15 \\
+    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $1000^+$ & N/A & \textbf{107.57} & 0.76 & 0.56 & - & - & 0.16 \\
 
@@
     \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & 82 & \textbf{DNF} & \textbf{14.01} & 0.27 & 0.20 & - & - & 0.25 \\
-    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $>1000$ & N/A & 0.54 & 1.09 & 0.84 & - & - & 0.27 \\
-    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $>1000$ & N/A & 0.16 & 1.07 & 1.06 & - & - & 0.16 \\
-    \rownb & angiofull~\cite{Weinstein2017} & 142 & $>1000$ & N/A & 0.17 & 1.06 & 0.88 & - & - & 0.23 \\
+    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $1000^+$ & N/A & 0.54 & 1.09 & 0.84 & - & - & 0.27 \\
+    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $1000^+$ & N/A & 0.16 & 1.07 & 1.06 & - & - & 0.16 \\
+    \rownb & angiofull~\cite{Weinstein2017} & 142 & $1000^+$ & N/A & 0.17 & 1.06 & 0.88 & - & - & 0.23 \\
     \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & 492 & \textbf{DNF} & \textbf{11.28} & 0.78 & 0.69 & - & - & 0.35 \\
-    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $>1000$ & N/A & \textbf{DNF} & 1.43 & 1.55 & - & - & 0.19 \\
-    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & $>1000$ & N/A & \textbf{13.58} & 1.76 & 1.51 & - & - & 0.27 \\
+    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $1000^+$ & N/A & \textbf{DNF} & 1.43 & 1.55 & - & - & 0.19 \\
+    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & $1000^+$ & N/A & \textbf{13.58} & 1.76 & 1.51 & - & - & 0.27 \\
 
     \midrule % 200 <= n
-    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $>1000$ & N/A & 1.13 & 1.47 & 1.43 & - & - & 0.31 \\
-    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & $>1000$ & N/A & \textbf{DNF} & 1.52 & 2.26 & - & - & 0.35 \\
-    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $>1000$ & N/A & \textbf{4.99} & NM & 1.96 & - & - & 0.28 \\
-    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & $>1000$ & N/A & 0.42 & 2.36 & \textbf{4.91} & - & - & 0.44 \\
-    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & $>1000$ & N/A & \textbf{DNF} & 2.56 & \textbf{5.62} & - & - & 0.46 \\
-    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & $>1000$ & N/A & 0.36 & 2.99 & \textbf{4.81} & - & - & 0.37 \\
-    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & $>1000$ & N/A & \textbf{DNF} & NM & \textbf{18.21} & - & - & 0.79 \\
+    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $1000^+$ & N/A & 1.13 & 1.47 & 1.43 & - & - & 0.31 \\
+    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & $1000^+$ & N/A & \textbf{DNF} & 1.52 & 2.26 & - & - & 0.35 \\
+    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $1000^+$ & N/A & \textbf{4.99} & NM & 1.96 & - & - & 0.28 \\
+    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & $1000^+$ & N/A & 0.42 & 2.36 & \textbf{4.91} & - & - & 0.44 \\
+    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & $1000^+$ & N/A & \textbf{DNF} & 2.56 & \textbf{5.62} & - & - & 0.46 \\
+    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & $1000^+$ & N/A & 0.36 & 2.99 & \textbf{4.81} & - & - & 0.37 \\
+    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & $1000^+$ & N/A & \textbf{DNF} & NM & \textbf{18.21} & - & - & 0.79 \\
 

### 2023-02-24 manuscripts/TCS2022/tcs2022.tex
using plots instead of tables

@@
 \newtheorem{proposition}{Proposition}[section]
+
+\usepackage{pgfplots}
+\pgfplotsset{width=10cm,compat=1.13}
 
@@
     \cmidrule(rr){8-11}
-    & model & \(n\) & \(|M|\) & LQM & PBN & \texttt{mpbn} & SAT & CP & ILP & ASP \\
+    & model & \(n\) & \(|M|\) & \texttt{LQM} & \texttt{PBN} & \texttt{mpbn} & \texttt{SAT} & \texttt{CP} & \texttt{ILP} & \texttt{ASP} \\
     \midrule
@@
 ``NM'' indicates a non-locally-monotonic model.
-There are four variants of \texttt{Trappist}: SAT (i.e., the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), CP (i.e., the CP-based method shown in Subsection~\ref{subsec:computation_csp}), ILP (i.e., the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and ASP (i.e., the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
+There are four variants of \texttt{Trappist}: \texttt{SAT} (i.e., the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{CP} (i.e., the CP-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{ILP} (i.e., the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and \texttt{ASP} (i.e., the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
 
@@
   \label{tab:result_real}
-  \scalebox{0.8}{
+  \scalebox{0.82}{
   \begin{tabular}{rlrrrrrrrrr}
@@
     \cmidrule(rr){8-11}
-    & model & \(n\) & \(|M|\) & LQM & PBN & \texttt{mpbn} & SAT & CP & ILP & ASP \\
+    & model & \(n\) & \(|M|\) & \texttt{LQM} & \texttt{PBN} & \texttt{mpbn} & \texttt{SAT} & \texttt{CP} & \texttt{ILP} & \texttt{ASP} \\
     \midrule % n < 100
@@
 
-\begin{table}[!htb]
-  \caption{Results on N-K models.}
-  \centering%
-  \label{tab:result_N_K}
-  \begin{tabular}{rrrrrrrr}
-    \toprule
-    & & & & \multicolumn{4}{c}{\texttt{Trappist}}\\
-    \cmidrule(rr){5-8}
-    \(n\) & \texttt{LQM} & \texttt{mpbn} & \texttt{PBN} & SAT & CP & ILP & ASP \\
-    \midrule
-    100 & 50 ($>$ 180) & 50 (N/A) & 0 (0.07) & 0 (0.05) & - () & - () & 0 (0.09) \\
-    150 & 50 ($>$ 180) & 50 (N/A) & 0 (0.14) & 0 (0.10) & - () & - () & 0 (0.14) \\
-    200 & 50 ($>$ 180) & 50 (N/A) & 0 (0.43) & 0 (0.25) & - () & - () & 0 (0.24) \\
-    250 & 50 ($>$ 180) & 50 (N/A) & 0 (1.92) & 0 (1.04) & - () & - () & 0 (0.56) \\
-    300 & 50 ($>$ 180) & 50 (N/A) & 0 (9.68) & 0 (4.46) & - () & - () & 0 (1.83) \\
-    350 & 50 ($>$ 180) & 50 (N/A) & 1 (46.54) & 0 (20.09) & - () & - () & 0 (6.10) \\
-    400 & 50 ($>$ 180) & 50 (N/A) & 29 (144.09) & 12 (90.36) & - () & - () & 1 (33.01) \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+\begin{figure}[ht]
+	\centering
+	\begin{tikzpicture}
+	\begin{semilogxaxis}[
+		%title={},
+		xlabel={Time (s)},
+		ylabel={Models solved},
+		xmin=-1, xmax=180,
+		log basis x={10},
+		%scale only axis,
+		ymin=0, ymax=350,
+		%ymode=log,
+		%log basis y={10},
+		xtick={0.5, 1, 2, 5, 10, 20, 50, 100, 180},
+		xticklabels = {0.5, 1, 2, 5, 10, 20, 50, 100, 180},
+		ytick={0, 50, 100, 150, 200, 250, 300, 350},
+		y=0.5cm/30,
+		%legend style={font=\fontsize{7}{8}\selectfont},
+		%legend style={at={(0.0,-0.1)},anchor=north},
+		legend pos=outer north east,
+		ymajorgrids=true,
+		grid style=dashed,
+	]
+
+	\addplot[
+		smooth,
+		thick,
+		color=blue,
+		mark=*,
+	]
+	coordinates {
+		(0.5, 141)(1, 174)(2, 193)(5, 220)(10, 247)(20, 262)(50, 285)(100, 306)(180, 320)
+	}; \addlegendentry{\texttt{PyBoolNet}}
+
+	\addplot[
+		smooth,
+		thick,
+		%dashed,
+		color=cyan,
+		mark=triangle,
+	]
+	coordinates {
+		(0.5, 0)(1, 0)(2, 0)(5, 0)(10, 0)(20, 0)(50, 0)(100, 0)(180, 0)
+	}; \addlegendentry{\texttt{mpbn}}
+
+	\addplot[
+		smooth,
+		thick,
+		dashed,
+		color=orange,
+		mark=asterisk,
+	]
+	coordinates {
+		(0.5, 0)(1, 0)(2, 0)(5, 0)(10, 0)(20, 0)(50, 0)(100, 0)(180, 0)
+	}; \addlegendentry{\texttt{bioLQM}}
+
+	\addplot[
+		smooth,
+		thick,
+		%dashed,
+		color=magenta,
+		mark=square,
+	]
+	coordinates {
+		(0.5, 164)(1, 190)(2, 214)(5, 249)(10, 266)(20, 291)(50, 317)(100, 328)(180, 338)
+	}; \addlegendentry{\texttt{Trappist-maxSAT}}
+
+	\addplot[
+		smooth,
+		thick,
+		dashed,
+		color=violet,
+		mark=star,
+	]
+	coordinates {
+		(0.5, 179)(1, 211)(2, 244)(5, 274)(10, 304)(20, 327)(50, 339)(100, 347)(180, 349)
+	}; \addlegendentry{\texttt{Trappist-ASP}}
+
+	\end{semilogxaxis}
+	\end{tikzpicture}
+	\caption{Cumulative numbers of random models solved by the competing methods with respect to enumerating the first 1000 minimal trap spaces.}
+	\label{fig:random_cumul_first}
+\end{figure}
+
+%\begin{table}[!htb]
+%  \caption{Results on N-K models.}
+%  \centering%
+%  \label{tab:result_N_K}
+%  \begin{tabular}{rrrrrrrr}
+%    \toprule
+%    & & & & \multicolumn{4}{c}{\texttt{Trappist}}\\
+%    \cmidrule(rr){5-8}
+%    \(n\) & \texttt{LQM} & \texttt{mpbn} & \texttt{PBN} & \texttt{SAT} & \texttt{CP} & \texttt{ILP} & \texttt{ASP} \\
+%    \midrule
+%    100 & 50 ($>$ 180) & 50 (N/A) & 0 (0.07) & 0 (0.05) & - () & - () & 0 (0.09) \\
+%    150 & 50 ($>$ 180) & 50 (N/A) & 0 (0.14) & 0 (0.10) & - () & - () & 0 (0.14) \\
+%    200 & 50 ($>$ 180) & 50 (N/A) & 0 (0.43) & 0 (0.25) & - () & - () & 0 (0.24) \\
+%    250 & 50 ($>$ 180) & 50 (N/A) & 0 (1.92) & 0 (1.04) & - () & - () & 0 (0.56) \\
+%    300 & 50 ($>$ 180) & 50 (N/A) & 0 (9.68) & 0 (4.46) & - () & - () & 0 (1.83) \\
+%    350 & 50 ($>$ 180) & 50 (N/A) & 1 (46.54) & 0 (20.09) & - () & - () & 0 (6.10) \\
+%    400 & 50 ($>$ 180) & 50 (N/A) & 29 (144.09) & 12 (90.36) & - () & - () & 1 (33.01) \\
+%    \bottomrule
+%  \end{tabular}
+%\end{table}
 

### 2023-02-24 manuscripts/TCS2022/tcs2022.tex
more plots

@@
 \pgfplotsset{width=10cm,compat=1.13}
+\usepgfplotslibrary{groupplots}
 
@@
 Specifically, all minimal trap spaces of a Boolean network are mutually disjoint.
-This property is important because we can use it to approximate the set of attractors of the Boolean network~\cite{klarner2015computing}.
+This property is important because we can use it to approximate the set of attractors of the Boolean network under any update scheme~\cite{klarner2015computing} or to compute exactly the set of complex attractors of the Boolean network under the fully asynchronous update scheme~\cite{DBLP:conf/bcb/TrinhHB22}.
 
@@
 
-\setlength{\tabcolsep}{4pt}
-\begin{table}[!htb]
-  \caption{Results on the real-world models from the BBM repository.}
-  \centering%
-  \label{tab:bbm_repo}
-  \begin{tabular}{lrrrr}
-    \toprule
-    Method & \# failures & avg-lqm (s) & avg-mono (s) & avg-all (s) \\ \midrule
-    \texttt{bioLQM} & 9 (134) & 12.87 & N/A & N/A \\
-    \texttt{PyBoolNet} & 12 & 8.87 & 11.00 & 13.59 \\
-    \texttt{mpbn} & 2 (187) & N/A & 2.31 & N/A \\
-    \texttt{Trappist}-MaxSAT & 1 & 0.03 & 1.09 & 1.01 \\
-    \texttt{Trappist}-CP & - & - & - & - \\
-    \texttt{Trappist}-ILP & - & - & - & - \\
-    \texttt{Trappist}-ASP & 1 & 0.05 & 1.02 & 0.93 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-\setcounter{rownumber}{0}
+%\setlength{\tabcolsep}{4pt}
+%\begin{table}[!htb]
+%  \caption{Results on the real-world models from the BBM repository.}
+%  \centering%
+%  \label{tab:bbm_repo}
+%  \begin{tabular}{lrrrr}
+%    \toprule
+%    Method & \# failures & avg-lqm (s) & avg-mono (s) & avg-all (s) \\ \midrule
+%    \texttt{bioLQM} & 9 (134) & 12.87 & N/A & N/A \\
+%    \texttt{PyBoolNet} & 12 & 8.87 & 11.00 & 13.59 \\
+%    \texttt{mpbn} & 2 (187) & N/A & 2.31 & N/A \\
+%    \texttt{Trappist}-MaxSAT & 1 & 0.03 & 1.09 & 1.01 \\
+%    \texttt{Trappist}-CP & - & - & - & - \\
+%    \texttt{Trappist}-ILP & - & - & - & - \\
+%    \texttt{Trappist}-ASP & 1 & 0.05 & 1.02 & 0.93 \\
+%    \bottomrule
+%  \end{tabular}
+%\end{table}
+%\setcounter{rownumber}{0}
+
+\begin{figure}[ht]
+	\centering
+	\begin{tikzpicture}
+		\begin{semilogxaxis}[
+			%title={},
+			xlabel={Time (s)},
+			ylabel={Models solved},
+			xmin=-1, xmax=180,
+			log basis x={10},
+			%scale only axis,
+			ymin=100, ymax=134,
+			%ymode=log,
+			%log basis y={10},
+			xtick={0.5, 1, 2, 5, 10, 20, 50, 100, 180},
+			xticklabels = {0.5, 1, 2, 5, 10, 20, 50, 100, 180},
+			ytick={100, 110, 120, 130},
+			y=0.5cm/4,
+			%legend style={font=\fontsize{7}{8}\selectfont},
+			%legend style={at={(1.2,-0.1)},anchor=north west},
+			legend pos=outer north east,
+			ymajorgrids=true,
+			grid style=dashed,
+		]
+
+		\addplot[
+			smooth,
+			thick,
+			dashed,
+			color=orange,
+			mark=asterisk,
+		]
+		coordinates {
+			(0.5, 115)(1, 118)(2, 121)(5, 122)(10, 122)(20, 123)(50, 125)(100, 125)(180, 125)
+		}; \addlegendentry{\texttt{bioLQM}}
+
+		\addplot[
+			smooth,
+			thick,
+			color=blue,
+			mark=*,
+		]
+		coordinates {
+			(0.5, 108)(1, 118)(2, 121)(5, 126)(10, 126)(20, 126)(50, 128)(100, 128)(180, 128)
+		}; \addlegendentry{\texttt{PyBoolNet}}
+
+		\addplot[
+			smooth,
+			thick,
+			%dashed,
+			color=cyan,
+			mark=triangle,
+		]
+		coordinates {
+			(0.5, 116)(1, 117)(2, 117)(5, 117)(10, 117)(20, 117)(50, 118)(100, 118)(180, 118)
+		}; \addlegendentry{\texttt{mpbn}}
+
+		\addplot[
+			smooth,
+			thick,
+			%dashed,
+			color=teal,
+			mark=square,
+		]
+		coordinates {
+			(0.5, 133)(1, 134)(2, 134)(5, 134)(10, 134)(20, 134)(50, 134)(100, 134)(180, 134)
+		}; \addlegendentry{\texttt{maxSAT}}
+
+%		\addplot[
+%			smooth,
+%			thick,
+%			%dashed,
+%			color=green,
+%			mark=star,
+%		]
+%		coordinates {
+%			(0.5, 143)(1, 176)(2, 195)(5, 222)(10, 249)(20, 264)(50, 287)(100, 308)(180, 322)
+%		}; \addlegendentry{\texttt{CP}}
+
+%		\addplot[
+%			smooth,
+%			thick,
+%			dashed,
+%			color=magenta,
+%			mark=star,
+%		]
+%		coordinates {
+%			(0.5, 144)(1, 177)(2, 196)(5, 223)(10, 250)(20, 265)(50, 288)(100, 309)(180, 323)
+%		}; \addlegendentry{\texttt{ILP}}
+
+		\addplot[
+			smooth,
+			thick,
+			dashed,
+			color=violet,
+			mark=star,
+		]
+		coordinates {
+			(0.5, 131)(1, 134)(2, 134)(5, 134)(10, 134)(20, 134)(50, 134)(100, 134)(180, 134)
+		}; \addlegendentry{\texttt{ASP}}
+
+		\end{semilogxaxis}
+	\end{tikzpicture}
+	\begin{tikzpicture}
+		\begin{semilogxaxis}[
+			%title={},
+			xlabel={Time (s)},
+			ylabel={Models solved},
+			xmin=-1, xmax=180,
+			log basis x={10},
+			%scale only axis,
+			ymin=150, ymax=211,
+			%ymode=log,
+			%log basis y={10},
+			xtick={0.5, 1, 2, 5, 10, 20, 50, 100, 180},
+			xticklabels = {0.5, 1, 2, 5, 10, 20, 50, 100, 180},
+			ytick={150, 170, 190, 210},
+			y=0.5cm/7,
+			%legend style={font=\fontsize{7}{8}\selectfont},
+			%legend style={at={(0.0,-0.1)},anchor=north},
+			legend pos=outer north east,
+			ymajorgrids=true,
+			grid style=dashed,
+		]
+
+		\addplot[
+			smooth,
+			thick,
+			color=blue,
+			mark=*,
+		]
+		coordinates {
+			(0.5, 151)(1, 170)(2, 178)(5, 188)(10, 189)(20, 190)(50, 194)(100, 198)(180, 199)
+		}; \addlegendentry{\texttt{PyBoolNet}}
+
+		\addplot[
+			smooth,
+			thick,
+			%dashed,
+			color=cyan,
+			mark=triangle,
+		]
+		coordinates {
+			(0.5, 160)(1, 177)(2, 181)(5, 183)(10, 183)(20, 184)(50, 185)(100, 185)(180, 185)
+		}; \addlegendentry{\texttt{mpbn}}
+
+		\addplot[
+			smooth,
+			thick,
+			%dashed,
+			color=teal,
+			mark=square,
+		]
+		coordinates {
+			(0.5, 191)(1, 204)(2, 208)(5, 210)(10, 210)(20, 210)(50, 210)(100, 210)(180, 210)
+		}; \addlegendentry{\texttt{maxSAT}}
+
+		%		\addplot[
+		%			smooth,
+		%			thick,
+		%			%dashed,
+		%			color=green,
+		%			mark=star,
+		%		]
+		%		coordinates {
+		%			(0.5, 143)(1, 176)(2, 195)(5, 222)(10, 249)(20, 264)(50, 287)(100, 308)(180, 322)
+		%		}; \addlegendentry{\texttt{CP}}
+
+		%		\addplot[
+		%			smooth,
+		%			thick,
+		%			dashed,
+		%			color=magenta,
+		%			mark=star,
+		%		]
+		%		coordinates {
+		%			(0.5, 144)(1, 177)(2, 196)(5, 223)(10, 250)(20, 265)(50, 288)(100, 309)(180, 323)
+		%		}; \addlegendentry{\texttt{ILP}}
+
+		\addplot[
+			smooth,
+			thick,
+			dashed,
+			color=violet,
+			mark=star,
+		]
+		coordinates {
+			(0.5, 205)(1, 208)(2, 209)(5, 210)(10, 210)(20, 210)(50, 210)(100, 210)(180, 210)
+		}; \addlegendentry{\texttt{ASP}}
+
+	\end{semilogxaxis}
+	\end{tikzpicture}
+	\caption{Cumulative numbers of the \texttt{BBM} models that have less than 1000 minimal trap spaces (above) and \texttt{BBM} models (below) solved by the competing methods with respect to enumerating the first 1000 minimal trap spaces.}
+	\label{fig:BBM_cumul}
+\end{figure}
 
@@
 		thick,
+		dashed,
+		color=orange,
+		mark=asterisk,
+	]
+	coordinates {
+		(0.5, 0)(1, 0)(2, 0)(5, 0)(10, 0)(20, 0)(50, 0)(100, 0)(180, 0)
+	}; \addlegendentry{\texttt{bioLQM}}
+
+	\addplot[
+		smooth,
+		thick,
 		color=blue,
@@
 		thick,
-		dashed,
-		color=orange,
-		mark=asterisk,
+		%dashed,
+		color=teal,
+		mark=square,
 	]
 	coordinates {
-		(0.5, 0)(1, 0)(2, 0)(5, 0)(10, 0)(20, 0)(50, 0)(100, 0)(180, 0)
-	}; \addlegendentry{\texttt{bioLQM}}
+		(0.5, 164)(1, 190)(2, 214)(5, 249)(10, 266)(20, 291)(50, 317)(100, 328)(180, 338)
+	}; \addlegendentry{\texttt{maxSAT}}
 
@@
 		%dashed,
+		color=green,
+		mark=star,
+	]
+	coordinates {
+		(0.5, 143)(1, 176)(2, 195)(5, 222)(10, 249)(20, 264)(50, 287)(100, 308)(180, 322)
+	}; \addlegendentry{\texttt{CP}}
+
+	\addplot[
+		smooth,
+		thick,
+		dashed,
 		color=magenta,
-		mark=square,
+		mark=star,
 	]
 	coordinates {
-		(0.5, 164)(1, 190)(2, 214)(5, 249)(10, 266)(20, 291)(50, 317)(100, 328)(180, 338)
-	}; \addlegendentry{\texttt{Trappist-maxSAT}}
+		(0.5, 144)(1, 177)(2, 196)(5, 223)(10, 250)(20, 265)(50, 288)(100, 309)(180, 323)
+	}; \addlegendentry{\texttt{ILP}}
 
@@
 		(0.5, 179)(1, 211)(2, 244)(5, 274)(10, 304)(20, 327)(50, 339)(100, 347)(180, 349)
-	}; \addlegendentry{\texttt{Trappist-ASP}}
+	}; \addlegendentry{\texttt{ASP}}
 
@@
 	\end{tikzpicture}
-	\caption{Cumulative numbers of random models solved by the competing methods with respect to enumerating the first 1000 minimal trap spaces.}
+	\caption{Cumulative numbers of random models solved by the competing methods with respect to enumerating all the minimal trap spaces.}
 	\label{fig:random_cumul_first}
@@
 
-Table~\ref{tab:result_N_K} shows the experimental results on N-K models.
-Column \(n\) denotes the network size.
-Columns \texttt{LQM} and \texttt{PBN} show the results of \texttt{bioLQM} and \texttt{PyBoolNet}, respectively.
-For each method, the number outside the parentheses indicates the number of failures, whereas the number inside the parentheses indicates the average running time (in seconds).
-Note that when computing the average running time, if the running time exceeds 180s, it is considered as 180s.
-From these results, we obtained several observations consistent with those obtained for real-world models.
-
-TODO: \dots
+%Table~\ref{tab:result_N_K} shows the experimental results on N-K models.
+%Column \(n\) denotes the network size.
+%Columns \texttt{LQM} and \texttt{PBN} show the results of \texttt{bioLQM} and \texttt{PyBoolNet}, respectively.
+%For each method, the number outside the parentheses indicates the number of failures, whereas the number inside the parentheses indicates the average running time (in seconds).
+%Note that when computing the average running time, if the running time exceeds 180s, it is considered as 180s.
+%From these results, we obtained several observations consistent with those obtained for real-world models.
+
+Figure~\ref{fig:random_cumul_first} shows cumulative numbers of random models solved by the competing methods with respect to enumerating all the minimal trap spaces.
+The number of succeeded models within three minutes for each method is as follows: \texttt{bioLQM} (0), \texttt{PyBoolNet} (320), \texttt{mpbn} (0), \texttt{Trappist-maxSAT} (338), \texttt{Trappist-CP} (0), \texttt{Trappist-ILP} (0), \texttt{Trappist-ASP} (349).
+We can see that \texttt{Trappist-ASP} is the method that could handle the most number of models.
+Note that none of the other methods could handle the model failed by \texttt{Trappist-ASP}.
+Upon closer inspection, we obtained several observations consistent with those obtained for real-world models as follows.
 
@@
 One of the reason may be that the BDD characterizing all trap spaces is too large, and its computation is slow.
+In addition, having too many generic trap spaces before the filtering process may be also a reason.
 It is apparent because the network size is large (\(\geq 100\)) and the Boolean functions are not simple.

### 2023-02-25 manuscripts/TCS2022/tcs2022.tex
selected models: mpbn

@@
 
-One naturally computational application of Theorem~\ref{theo:ts_2_sp} is that we can efficiently decide whether a subspace \(m\) is a trap space.
+A naturally computational application of Theorem~\ref{theo:ts_2_sp} is that we can efficiently decide whether a subspace \(m\) is a trap space.
 In \texttt{PyBoolNet}~\cite{klarner2017pyboolnet}, this is checked by using the percolation on the prime-implicants of the Boolean functions.
@@
 
-In the field of systems biology, biologists may want to compute more special types of trap spaces beyond minimal trap spaces~\cite{klarner2017pyboolnet}.
+In the field of systems biology, biologists may want to compute more special types of trap spaces beyond minimal trap spaces~\cite{klarner2017pyboolnet}, which also play crucial roles in analysis of Boolean networks.
 We shall show that our proposed methods can be easily adjusted to compute popular types of trap spaces.
@@
 Column \(n\) denotes the number of nodes of each model.
-Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
+Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 minimal trap spaces.
 In the case of \texttt{bioLQM}, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of \texttt{bioLQM} because it always requires to compute all minimal trap spaces.
@@
 As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the \texttt{PyBoolNet} repository, the results are comparable with all minimal trap spaces found very fast.
-For 5 of the 29 models, \texttt{mpbn} did not give any answer because it recognized these models as not locally-monotonic.
-Note that on some very small models, \texttt{Trappist} is sometimes slower than \texttt{PyBoolNet} and/or \texttt{mpbn}, but still significantly under one second.
-On the contrary, on every model that was a bit challenging for \texttt{PyBoolNet} or \texttt{mpbn}, the new method is far more efficient with speedups between one and two orders of magnitude.
+For 3 of the 29 models, \texttt{mpbn} did not give any answer because these models are locally-monotonic.
+%Note that on some very small models, \texttt{Trappist} is sometimes slower than \texttt{PyBoolNet} and/or \texttt{mpbn}, but still significantly under one second.
+%On the contrary, on every model that was a bit challenging for \texttt{PyBoolNet} or \texttt{mpbn}, the new method is far more efficient with speedups between one and two orders of magnitude.
 
@@
 It aims to be a comprehensive collection suitable for benchmarking and testing new tools and methods.
+\texttt{BBM} consists of 211 models (24 out of them are non-locally-monotonic), peaking at 321 nodes, 1100 regulations among the nodes, and 133 source nodes, respectively.
 It is released and maintained at \url{https://github.com/sybila/biodivine-boolean-models}.
@@
 			(0.5, 133)(1, 134)(2, 134)(5, 134)(10, 134)(20, 134)(50, 134)(100, 134)(180, 134)
-		}; \addlegendentry{\texttt{maxSAT}}
+		}; \addlegendentry{\texttt{SAT}}
 
@@
 			(0.5, 191)(1, 204)(2, 208)(5, 210)(10, 210)(20, 210)(50, 210)(100, 210)(180, 210)
-		}; \addlegendentry{\texttt{maxSAT}}
+		}; \addlegendentry{\texttt{SAT}}
 
@@
 
-Table~\ref{tab:bbm_repo} shows the experimental results on the 211 real-world models from the \texttt{BBM} repository.
-Column 2 expresses the numbers of failures (i.e., did not finish the computation within a time limit of three minutes) of each method.
-For the case of \texttt{bioLQM}, we only considered the models that have at most 1000 minimal trap spaces.
-The number of such models is 134 (per all 211 models) and is denoted inside the parentheses.
-For the case of \texttt{mpbn}, we only considered the models that are locally-monotonic.
-The number of such models is 187 (per all 211 models) and is denoted inside the parentheses.
-Columns 3-5 express the average running time (in seconds) of each method for the models having at most 1000 minimal trap spaces, the locally-monotonic models, and all the models, respectively.
-Note that when computing the average running time, if the running time exceeds 180s, it is considered as 180s.
-From the results shown in Table~\ref{tab:bbm_repo}, we reported several observations as follows.
+Figure~\ref{fig:BBM_cumul} (above) shows cumulative numbers of the \texttt{BBM} models that have less than 1000 minimal trap spaces solved by the competing methods with respect to enumerating the first 1000 minimal trap spaces.
+The number of such models is 134 (per all 211 models), and 15 of them are non-locally-monotonic.
+We can first see that \dots
+
+Figure~\ref{fig:BBM_cumul} (below) shows cumulative numbers of the \texttt{BBM} models solved by the competing methods (except \texttt{bioLQM}) with respect to enumerating the first 1000 minimal trap spaces.
+\dots
+
+%Table~\ref{tab:bbm_repo} shows the experimental results on the 211 real-world models from the \texttt{BBM} repository.
+%Column 2 expresses the numbers of failures (i.e., did not finish the computation within a time limit of three minutes) of each method.
+%For the case of \texttt{bioLQM}, we only considered the models that have at most 1000 minimal trap spaces.
+%The number of such models is 134 (per all 211 models) and is denoted inside the parentheses.
+%For the case of \texttt{mpbn}, we only considered the models that are locally-monotonic.
+%The number of such models is 187 (per all 211 models) and is denoted inside the parentheses.
+%Columns 3-5 express the average running time (in seconds) of each method for the models having at most 1000 minimal trap spaces, the locally-monotonic models, and all the models, respectively.
+%Note that when computing the average running time, if the running time exceeds 180s, it is considered as 180s.
+%From the results shown in Table~\ref{tab:bbm_repo}, we reported several observations as follows.
 
@@
 
-We used a set of real-world Boolean networks lying in various scales collected from numerous bibliographic sources.
-Most of these models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and have never been fully analyzed.
+We used a set of real-world Boolean networks lying in various scales collected from numerous bibliographic sources in the literature.
+Most of these models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants), and have never been fully analyzed.
 Note that these models are not included in the \texttt{PyBoolNet} and \texttt{BBM} repositories.
@@
 Table~\ref{tab:result_real} shows the obtained experimental results.
-``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces) within the timeout of two minutes.
-%In the case of \texttt{bioLQM}, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of \texttt{bioLQM} because it always requires to compute all minimal trap spaces.
+``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces) within the time limit of three minutes.
 A number in bold indicates a ratio greater than or equal to 10 compared to the best result.
@@
 
-The first observation is that for 26 of the 33 models (more than 78\%), \texttt{mpbn} did not give any answer because it recognized that these models as not locally-monotonic.
-For 6 of the 33 models where \texttt{mpbn} returned the answers, \texttt{mpbn} and \texttt{Trappist} are comparable in computation time, though surprisingly \texttt{mpbn} appears a bit slower on average.
-Note however that \texttt{mpbn} was the only tool to provide a solution for the SN-5 model, thus confirming that if the activation function is in the right form, not having to compute the inactivation function's disjunctive normal form can render a difficult problem tractable.
+The first observation is that for 11 of the 32 models (more than 34\%), \texttt{mpbn} did not give any answer because these models as not locally-monotonic.
+For 22 of the 32 models where \texttt{mpbn} returned the answers, \texttt{mpbn} and \texttt{Trappist} are comparable in computation time, though surprisingly \texttt{mpbn} appears quite slower on average than the best variant of \texttt{Trappist}.
+In particular, for the Rho-GTPases model, \texttt{mpbn} is \(577\times\) slower than the best variant of \texttt{Trappist}.
+%Note however that \texttt{mpbn} was the only tool to provide a solution for the SN-5 model, thus confirming that if the activation function is in the right form, not having to compute the inactivation function's disjunctive normal form can render a difficult problem tractable.
 However, since \texttt{mbpn} can handle only locally-monotonic models and \texttt{Trappist} can handle general models, it is difficult to further compare between them.
-Hence, we focus on only comparisons between \texttt{PyBoolNet} and \texttt{Trappist} in the following observations.
+Hence, we focus on only comparisons among \texttt{bioLQM}, \texttt{PyBoolNet}, and \texttt{Trappist} in the following observations.
 

### 2023-02-25 manuscripts/TCS2022/tcs2022.tex
random models

@@
 In particular, for the Rho-GTPases model, \texttt{mpbn} is \(577\times\) slower than the best variant of \texttt{Trappist}.
-%Note however that \texttt{mpbn} was the only tool to provide a solution for the SN-5 model, thus confirming that if the activation function is in the right form, not having to compute the inactivation function's disjunctive normal form can render a difficult problem tractable.
 However, since \texttt{mbpn} can handle only locally-monotonic models and \texttt{Trappist} can handle general models, it is difficult to further compare between them.
@@
 
-The second observation is that the proposed method vastly outperforms \texttt{PyBoolNet} in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100--1000 nodes size range).
-Note that for all the cases where \texttt{PyBoolNet} did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
-Hence, not even a single minimal trap space was output by that method.
-The computational advantage is therefore immediately a practical advantage since on the one hand the state-of-the-art method did not allow any analysis whatsoever of the models, and on the other hand the proposed method could provide, very often under one second, the first thousand minimal trap spaces.
-For modellers having a critical look at a model and in a \emph{model, invalidate, refine} loop this means a huge difference in the models that are amenable to study.
-
-Note that even with a very restricted time-limit of two minutes, it was possible with the proposed technique to find \emph{all} minimal trap spaces of small models (roughly under 130 nodes, i.e., considered as quite big up to now).
-Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
-Since the ASP code is declarative by nature, it is also possible to add to it supplementary constraints coming from the modeler in case one is looking for specific attractors.
-Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration.
-
-The third observation is that for all the models where \texttt{PyBoolNet} finished before the timeout, once \texttt{PyBoolNet} went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
-For these models, the ASP solving phase of the proposed method also took very short time, all under one second.
-Hence, with the experimental results shown in this paper, the practical differences between our ASP encoding and that of \texttt{PyBoolNet} are not distinctly exposed.
-The fact that our new ASP encoding is guaranteed to be linear in the number of nodes of the original model does not seem to be crucial here, however a much deeper analysis of those cases remains to be done.
-
-Note that though enumerating the extremal siphons of a Petri net is exponential (see~\cite{nabli2016enumerating} for instance) this is apparently not the bottleneck of the proposed method, showing once again that networks obtained from biochemical models do have a specific structure.
+%The second observation is that the proposed method vastly outperforms \texttt{PyBoolNet} in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100--1000 nodes size range).
+%Note that for all the cases where \texttt{PyBoolNet} did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
+%Hence, not even a single minimal trap space was output by that method.
+%The computational advantage is therefore immediately a practical advantage since on the one hand the state-of-the-art method did not allow any analysis whatsoever of the models, and on the other hand the proposed method could provide, very often under one second, the first thousand minimal trap spaces.
+%For modellers having a critical look at a model and in a \emph{model, invalidate, refine} loop this means a huge difference in the models that are amenable to study.
+%
+%Note that even with a very restricted time-limit of two minutes, it was possible with the proposed technique to find \emph{all} minimal trap spaces of small models (roughly under 130 nodes, i.e., considered as quite big up to now).
+%Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
+%Since the ASP code is declarative by nature, it is also possible to add to it supplementary constraints coming from the modeler in case one is looking for specific attractors.
+%Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration.
+%
+%The third observation is that for all the models where \texttt{PyBoolNet} finished before the timeout, once \texttt{PyBoolNet} went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
+%For these models, the ASP solving phase of the proposed method also took very short time, all under one second.
+%Hence, with the experimental results shown in this paper, the practical differences between our ASP encoding and that of \texttt{PyBoolNet} are not distinctly exposed.
+%The fact that our new ASP encoding is guaranteed to be linear in the number of nodes of the original model does not seem to be crucial here, however a much deeper analysis of those cases remains to be done.
+%
+%Note that though enumerating the extremal siphons of a Petri net is exponential (see~\cite{nabli2016enumerating} for instance) this is apparently not the bottleneck of the proposed method, showing once again that networks obtained from biochemical models do have a specific structure.
 
@@
 We randomly generated a set of N-K models~\cite{glass1973logical} with network size \(n\) in the set \{100, 150, 200, 250, 300, 350, 400\} and \(K = 3\) (i.e., each node has exactly three input nodes).
-We chose N-K models because they are a useful tool for studying the dynamics of Boolean networks~\cite{glass1973logical,klarner2015computing}.
+We chose N-K models because they are a useful tool for studying the dynamics of Boolean networks~\cite{glass1973logical,klarner2015computing,Rozum2021}.
 For each network size, 50 instances were generated using the \verb|generateRandomNKNetwork| function.
 In total, we have 350 random models.
-We then applied the compared methods to these models and recorded the numbers of failures (i.e., failed to obtain the result within a time limit of three minutes) as well as the average running time (inside the parentheses) in each method for each network size \(n\).
+We then applied the compared methods to these models and recorded the running time of each method for each model.
 It is worth noting that N-K models usually have small numbers of minimal trap spaces~\cite{klarner2015computing}.
@@
 		(0.5, 164)(1, 190)(2, 214)(5, 249)(10, 266)(20, 291)(50, 317)(100, 328)(180, 338)
-	}; \addlegendentry{\texttt{maxSAT}}
+	}; \addlegendentry{\texttt{SAT}}
 
@@
 Recall that a Boolean network is non-locally-monotonic if only one of its Boolean functions is non-locally-monotonic.
-Hence, it is apparent that all the randomly generated models are non-locally-monotonic.
+Hence, it is apparent that all the randomly generated models are non-locally-monotonic because of the number of nodes is large (\(n \geq 100\)).
 This observation confirms the limit on the applicable model class of \texttt{mpbn}.
@@
 In addition, having too many generic trap spaces before the filtering process may be also a reason.
-It is apparent because the network size is large (\(\geq 100\)) and the Boolean functions are not simple.
-
-Third, \texttt{PyBoolNet} could handle every model of network size less than or equal to 300.
-It only failed in one model of network size 350 but 29 models of network size 400.
-The average running time vastly increases as the network size increases.
-As compared to the four methods of our approach, the performance of \texttt{PyBoolNet} is comparable for the 100-node and 200-node models.
-However, from \(n = 250\), the performance difference is exhibited more clearly.
+It is apparent because the network size is large (\(n \geq 100\)) and the Boolean functions are not simple.
+
+%Third, \texttt{PyBoolNet} could handle every model of network size less than or equal to 300.
+%It only failed in one model of network size 350 but 29 models of network size 400.
+%The average running time vastly increases as the network size increases.
+%As compared to the four methods of our approach, the performance of \texttt{PyBoolNet} is comparable for the 100-node and 200-node models.
+%However, from \(n = 250\), the performance difference is exhibited more clearly.
+
+Third, \dots
 

### 2023-03-07 trappist/cp.py
Not sure asyncio.run() is enough

@@ -16,6 +16,7 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 
+from asyncio import run
 from dataclasses import asdict
 from datetime import timedelta
 from time import perf_counter
@@ -82,7 +83,7 @@ def solve_cp(
     ):
         start = perf_counter()
         # result = inst.solve(timeout=remaining, processes=nprocs)
-        result = inst.solve_async(timeout=remaining, processes=nprocs)
+        result = run(inst.solve_async(timeout=remaining, processes=nprocs))
         end = perf_counter()
         if remaining is not None:
             remaining -= timedelta(seconds=(end - start))

### 2023-03-07 trappist/ilp.py

@@ -16,6 +16,7 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 
+from asyncio import run
 from dataclasses import asdict
 from datetime import timedelta
 from time import perf_counter
@@ -73,7 +74,7 @@ def solve_ilp(
         max_output == 0 or nsol < max_output
     ):
         start = perf_counter()
-        result = inst.solve(timeout=remaining, processes=nprocs)
+        result = run(inst.solve_async(timeout=remaining, processes=nprocs))
         end = perf_counter()
         if remaining is not None:
             remaining -= timedelta(seconds=(end - start))

### 2023-03-07 trappist/cp.py
can run on notebook

@@ -16,7 +16,7 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 
-from asyncio import run
+from asyncio import run, get_running_loop
 from dataclasses import asdict
 from datetime import timedelta
 from time import perf_counter
@@ -26,6 +26,7 @@ from minizinc import Instance, Model, Result, Solver, Status
 
 import networkx as nx  # TODO maybe replace with lists/dicts
 
+from concurrent.futures import ThreadPoolExecutor
 
 def create_cp(petri_net: nx.DiGraph) -> Model:
     """Create the CP program for the max conflict-free siphons of petri net."""
@@ -83,7 +84,15 @@ def solve_cp(
     ):
         start = perf_counter()
         # result = inst.solve(timeout=remaining, processes=nprocs)
-        result = run(inst.solve_async(timeout=remaining, processes=nprocs))
+        #result = run(inst.solve_async(timeout=remaining, processes=nprocs))
+
+        try:
+            get_running_loop()
+            with ThreadPoolExecutor(1) as pool:
+                result = pool.submit(lambda: run(inst.solve_async(timeout=remaining, processes=nprocs))).result()
+        except RuntimeError:
+            result = run(inst.solve_async(timeout=remaining, processes=nprocs))
+
         end = perf_counter()
         if remaining is not None:
             remaining -= timedelta(seconds=(end - start))

### 2023-03-07 trappist/ilp.py

@@ -16,7 +16,7 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 
-from asyncio import run
+from asyncio import run, get_running_loop
 from dataclasses import asdict
 from datetime import timedelta
 from time import perf_counter
@@ -28,6 +28,7 @@ import networkx as nx  # TODO maybe replace with lists/dicts
 
 from .cp import cp_to_bool, zincify
 
+from concurrent.futures import ThreadPoolExecutor
 
 def create_ilp(petri_net: nx.DiGraph) -> Model:
     """Create the ILP program for the max conflict-free siphons of petri net."""
@@ -74,7 +75,14 @@ def solve_ilp(
         max_output == 0 or nsol < max_output
     ):
         start = perf_counter()
-        result = run(inst.solve_async(timeout=remaining, processes=nprocs))
+        #result = run(inst.solve_async(timeout=remaining, processes=nprocs))
+        try:
+            get_running_loop()
+            with ThreadPoolExecutor(1) as pool:
+                result = pool.submit(lambda: run(inst.solve_async(timeout=remaining, processes=nprocs))).result()
+        except RuntimeError:
+            result = run(inst.solve_async(timeout=remaining, processes=nprocs))
+
         end = perf_counter()
         if remaining is not None:
             remaining -= timedelta(seconds=(end - start))

### 2023-03-08 trappist/ilp.py
ILP method works now

@@ -52,7 +52,7 @@ def create_ilp(petri_net: nx.DiGraph) -> Model:
                 if zsucc not in or_preds:  # optimize obvious tautologies
                     or_string = " + ".join(or_preds)
                     model.add_string(f"constraint {zsucc} <= {or_string};\n")
-    model.add_string(f"solve maximize {' + '.join(variables)};\n")
+    model.add_string(f"solve maximize ({' + '.join(variables)});\n")
     return model
 
 
@@ -76,6 +76,7 @@ def solve_ilp(
     ):
         start = perf_counter()
         #result = run(inst.solve_async(timeout=remaining, processes=nprocs))
+        
         try:
             get_running_loop()
             with ThreadPoolExecutor(1) as pool:
@@ -96,7 +97,8 @@ def solve_ilp(
             # non subset constrait for maximality
             non_sub = []
             for k, v in d.items():
-                if not v:
+                #if not v:
+                if not v and k != "objective":
                     non_sub.append(k)
             or_string = " + ".join(non_sub)
             model.add_string(f"constraint 1 <= {or_string};")

### 2023-03-09 manuscripts/TCS2022/tcs2022.tex
Timeout problem of CP and ILP for N-K models

@@
     \midrule
-    \rownb & arellano\_rootstem & 9 & 4 & \textbf{0.13} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & calzone\_cellfate & 28 & 27 & \textbf{0.12} & 0.02 & 0.01 & 0.01 & - & - & 0.01\\
-    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & \textbf{0.11} & 0.03 & 0.01 & 0.01 & - & - & 0.01\\
-    \rownb & davidich\_yeast & 10 & 12 & \textbf{0.11} & 0.02 & 0.01 & 0.01 & - & - & 0.01\\
-    \rownb & dinwoodie\_life & 15 & 7 & \textbf{0.11} & 0.01 & 0.00 & 0.01 & - & - & 0.01\\
-    \rownb & dinwoodie\_stomatal & 13 & 1 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & faure\_cellcycle & 10 & 2 & \textbf{0.11} & 0.02 & 0.01 & 0.01 & - & - & 0.01\\
-    \rownb & grieco\_mapk & 53 & 18 & \textbf{0.19} & 0.03 & 0.02 & 0.03 & - & - & 0.02\\
-    \rownb & irons\_yeast & 18 & 1 & \textbf{0.12} & 0.03 & 0.01 & 0.01 & - & - & 0.02\\
-    \rownb & jaoude\_thdiff & 103 & $1000^+$ & N/A & \textbf{0.85} & \textbf{0.45} & \textbf{0.56} & - & - & 0.09\\
-    \rownb & klamt\_tcr & 40 & 8 & \textbf{0.11} & 0.01 & 0.01 & 0.01 & - & - & 0.02\\
-    \rownb & krumsiek\_myeloid & 11 & 6 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & multivalued & 13 & 4 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & n12c5 & 11 & 5 & \textbf{0.11} & \textbf{17.83} & 0.01 & 0.01 & - & - & 0.01\\
-    \rownb & n3s1c1a & 2 & 2 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & n3s1c1b & 2 & 2 & \textbf{0.09} & 0.02 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & n5s3 & 4 & 3 & \textbf{0.10} & 0.02 & \textbf{NM} & 0.00 & - & - & 0.01\\
-    \rownb & n6s1c2 & 5 & 3 & \textbf{0.10} & 0.02 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & n7s3 & 6 & 3 & \textbf{0.11} & 0.02 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & raf & 3 & 2 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & randomnet\_n15k3 & 15 & 3 & \textbf{0.10} & 0.02 & \textbf{NM} & 0.01 & - & - & 0.01\\
-    \rownb & randomnet\_n7k3 & 7 & 10 & \textbf{0.10} & 0.01 & \textbf{NM} & 0.00 & - & - & 0.01\\
-    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{0.15} & \textbf{0.94} & 0.02 & 0.02 & - & - & 0.02\\
-    \rownb & saadatpour\_guardcell & 13 & 1 & \textbf{0.10} & 0.06 & 0.00 & 0.00 & - & - & 0.02\\
-    \rownb & selvaggio\_emt & 56 & $1000^+$ & N/A & \textbf{0.48} & \textbf{0.28} & \textbf{0.28} & - & - & 0.09\\
-    \rownb & tournier\_apoptosis & 12 & 3 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & xiao\_wnt5a & 7 & 4 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & - & - & 0.01\\
-    \rownb & zhang\_tlgl & 60 & 156 & \textbf{0.60} & 0.09 & 0.09 & 0.07 & - & - & 0.04\\
-    \rownb & zhang\_tlgl\_v2 & 60 & 258 & \textbf{0.64} & 0.04 & 0.08 & 0.11 & - & - & 0.04\\
+    \rownb & arellano\_rootstem & 9 & 4 & \textbf{0.13} & 0.01 & 0.00 & 0.00 & \textbf{0.97} & \textbf{0.96} & 0.01\\
+    \rownb & calzone\_cellfate & 28 & 27 & \textbf{0.12} & 0.02 & 0.01 & 0.01 & \textbf{5.59} & \textbf{6.03} & 0.01\\
+    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & \textbf{0.11} & 0.03 & 0.01 & 0.01 & \textbf{6.56} & \textbf{6.99} & 0.01\\
+    \rownb & davidich\_yeast & 10 & 12 & \textbf{0.11} & 0.02 & 0.01 & 0.01 & \textbf{2.56} & \textbf{2.21} & 0.01\\
+    \rownb & dinwoodie\_life & 15 & 7 & \textbf{0.11} & 0.01 & 0.00 & 0.01 & \textbf{1.68} & \textbf{1.39} & 0.01\\
+    \rownb & dinwoodie\_stomatal & 13 & 1 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & \textbf{0.39} & \textbf{0.29} & 0.01\\
+    \rownb & faure\_cellcycle & 10 & 2 & \textbf{0.11} & 0.02 & 0.01 & 0.01 & \textbf{0.58} & \textbf{0.46} & 0.01\\
+    \rownb & grieco\_mapk & 53 & 18 & \textbf{0.19} & 0.03 & 0.02 & 0.03 & \textbf{3.93} & \textbf{10.46} & 0.02\\
+    \rownb & irons\_yeast & 18 & 1 & \textbf{0.12} & 0.03 & 0.01 & 0.01 & \textbf{0.37} & \textbf{0.39} & 0.02\\
+    \rownb & jaoude\_thdiff & 103 & $1000^+$ & N/A & \textbf{0.85} & \textbf{0.45} & \textbf{0.56} & \textbf{DNF} & \textbf{DNF} & 0.09\\
+    \rownb & klamt\_tcr & 40 & 8 & \textbf{0.11} & 0.01 & 0.01 & 0.01 & \textbf{1.98} & \textbf{1.22} & 0.02\\
+    \rownb & krumsiek\_myeloid & 11 & 6 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & \textbf{1.48} & \textbf{1.26} & 0.01\\
+    \rownb & multivalued & 13 & 4 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & \textbf{0.93} & \textbf{0.86} & 0.01\\
+    \rownb & n12c5 & 11 & 5 & \textbf{0.11} & \textbf{17.83} & 0.01 & 0.01 & \textbf{1.21} & \textbf{1.10} & 0.01\\
+    \rownb & n3s1c1a & 2 & 2 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & \textbf{0.63} & \textbf{0.49} & 0.01\\
+    \rownb & n3s1c1b & 2 & 2 & \textbf{0.09} & 0.02 & 0.00 & 0.00 & \textbf{0.56} & \textbf{0.49} & 0.01\\
+    \rownb & n5s3 & 4 & 3 & \textbf{0.10} & 0.02 & \textbf{NM} & 0.00 & \textbf{0.74} & \textbf{0.69} & 0.01\\
+    \rownb & n6s1c2 & 5 & 3 & \textbf{0.10} & 0.02 & 0.00 & 0.00 & \textbf{0.91} & \textbf{0.59} & 0.01\\
+    \rownb & n7s3 & 6 & 3 & \textbf{0.11} & 0.02 & 0.00 & 0.00 & \textbf{0.79} & \textbf{0.68} & 0.01\\
+    \rownb & raf & 3 & 2 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & \textbf{0.55} & \textbf{0.39} & 0.01\\
+    \rownb & randomnet\_n15k3 & 15 & 3 & \textbf{0.10} & 0.02 & \textbf{NM} & 0.01 & \textbf{0.77} & \textbf{0.67} & 0.01\\
+    \rownb & randomnet\_n7k3 & 7 & 10 & \textbf{0.10} & 0.01 & \textbf{NM} & 0.00 & \textbf{2.07} & \textbf{1.46} & 0.01\\
+    \rownb & remy\_tumorigenesis & 34 & 25 & \textbf{0.15} & \textbf{0.94} & 0.02 & 0.02 & \textbf{5.98} & \textbf{7.98} & 0.02\\
+    \rownb & saadatpour\_guardcell & 13 & 1 & \textbf{0.10} & 0.06 & 0.00 & 0.00 & \textbf{0.53} & \textbf{0.45} & 0.02\\
+    \rownb & selvaggio\_emt & 56 & $1000^+$ & N/A & \textbf{0.48} & \textbf{0.28} & \textbf{0.28} & \textbf{DNF} & \textbf{DNF} & 0.09\\
+    \rownb & tournier\_apoptosis & 12 & 3 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & \textbf{0.74} & \textbf{0.75} & 0.01\\
+    \rownb & xiao\_wnt5a & 7 & 4 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & \textbf{1.00} & \textbf{0.89} & 0.01\\
+    \rownb & zhang\_tlgl & 60 & 156 & \textbf{0.60} & 0.09 & 0.09 & 0.07 & \textbf{37.26} & \textbf{DNF} & 0.04\\
+    \rownb & zhang\_tlgl\_v2 & 60 & 258 & \textbf{0.64} & 0.04 & 0.08 & 0.11 & \textbf{69.95} & \textbf{DNF} & 0.04\\
     \bottomrule
@@
 			xmin=-1, xmax=180,
-			log basis x={10},
+			%xmode=log,
+			%log basis x={10},
 			%scale only axis,
-			ymin=100, ymax=134,
+			ymin=0, ymax=134,
 			%ymode=log,
-			%log basis y={10},
+			%log basis y={2},
 			xtick={0.5, 1, 2, 5, 10, 20, 50, 100, 180},
 			xticklabels = {0.5, 1, 2, 5, 10, 20, 50, 100, 180},
-			ytick={100, 110, 120, 130},
-			y=0.5cm/4,
+			ytick={0,20,40,60,80,100,120},
+			y=0.5cm/10,
 			%legend style={font=\fontsize{7}{8}\selectfont},
@@
 
-%		\addplot[
-%			smooth,
-%			thick,
-%			%dashed,
-%			color=green,
-%			mark=star,
-%		]
-%		coordinates {
-%			(0.5, 143)(1, 176)(2, 195)(5, 222)(10, 249)(20, 264)(50, 287)(100, 308)(180, 322)
-%		}; \addlegendentry{\texttt{CP}}
+		\addplot[
+			smooth,
+			thick,
+			%dashed,
+			color=green,
+			mark=star,
+		]
+		coordinates {
+			(0.5, 4)(1, 33)(2, 59)(5, 87)(10, 109)(20, 118)(50, 122)(100, 130)(180, 133)
+		}; \addlegendentry{\texttt{CP}}
 
@@
 %		coordinates {
-		%			(0.5, 143)(1, 176)(2, 195)(5, 222)(10, 249)(20, 264)(50, 287)(100, 308)(180, 322)
+%			(0.5, 4)(1, 33)(2, 59)(5, 87)(10, 109)(20, 118)(50, 122)(100, 130)(180, 133)
 %		}; \addlegendentry{\texttt{CP}}
@@
     \midrule % n < 100
-    \rownb & metastatic~\cite{DesignPrinciplesGeneNetworks} & 10 & 4 & \textbf{0.10} & 0.04 & NM & 0.01 & - & - & 0.02 \\
-    \rownb & Arabidopsis\_thaliana~\cite{DesignPrinciplesGeneNetworks} & 15 & 8 & \textbf{0.10} & 0.06 & NM & 0.01 & - & - & 0.02 \\
-    \rownb & p53\_high\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.38 & \textbf{1.76} & NM & 0.08 & - & - & 0.14 \\
-    \rownb & p53\_low\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.41 & \textbf{1.76} & NM & 0.07 & - & - & 0.14 \\
-    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & 32 & \textbf{DNF} & \textbf{DNF} & NM & 0.03 & - & - & 0.19 \\
-    \rownb & DNA\_damage~\cite{DesignPrinciplesGeneNetworks} & 26 & 16 & \textbf{0.24} & \textbf{0.33} & NM & 0.02 & - & - & 0.05 \\
-    \rownb & Rho-GTPases~\cite{DesignPrinciplesGeneNetworks} & 33 & 2 & 0.17 & 0.57 & \textbf{40.39} & 0.07 & - & - & 0.11 \\
-    \rownb & Pluripotency~\cite{YachieKinoshita2018} & 36 & 440 & \textbf{DNF} & \textbf{DNF} & NM & 0.16 & - & - & 0.28 \\
-    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & 276 & 0.37 & 0.43 & NM & 0.07 & - & - & 0.06 \\
-    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $1000^+$ & N/A & 0.11 & 0.36 & 0.17 & - & - & 0.06 \\
-    \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & 128 & 0.33 & 0.05 & 0.07 & 0.06 & - & - & 0.05 \\
-    \rownb & Cacace\_TdevModel~\cite{Cacace2020} & 61 & 28 & \textbf{1.29} & \textbf{5.67} & NM & 0.06 & - & - & 0.08 \\
-    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $1000^+$ & N/A & \textbf{DNF} & 0.50 & 0.34 & - & - & 0.33 \\
-    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & \textbf{39.22} & \textbf{1.01} & 0.20 & 0.12 & - & - & 0.05 \\
-    \rownb & Bcell~\cite{Dutta2019} & 73 & 72 & 0.23 & 0.04 & 0.08 & 0.06 & - & - & 0.05 \\
-    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $1000^+$ & N/A & 0.09 & 0.55 & 0.37 & - & - & 0.15 \\
-    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $1000^+$ & N/A & \textbf{107.57} & 0.76 & 0.56 & - & - & 0.16 \\
+    \rownb & metastatic~\cite{DesignPrinciplesGeneNetworks} & 10 & 4 & \textbf{0.10} & 0.04 & NM & 0.01 & \textbf{1.15} & \textbf{0.89} & 0.02 \\
+    \rownb & Arabidopsis\_thaliana~\cite{DesignPrinciplesGeneNetworks} & 15 & 8 & \textbf{0.10} & 0.06 & NM & 0.01 & \textbf{2.06} & \textbf{1.83} & 0.02 \\
+    \rownb & p53\_high\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.38 & \textbf{1.76} & NM & 0.08 & 0.53 & 0.43 & 0.14 \\
+    \rownb & p53\_low\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.41 & \textbf{1.76} & NM & 0.07 & 0.58 & 0.48 & 0.14 \\
+    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & 32 & \textbf{DNF} & \textbf{DNF} & NM & 0.03 & \textbf{8.41} & \textbf{12.38} & 0.19 \\
+    \rownb & DNA\_damage~\cite{DesignPrinciplesGeneNetworks} & 26 & 16 & \textbf{0.24} & \textbf{0.33} & NM & 0.02 & \textbf{3.91} & \textbf{5.33} & 0.05 \\
+    \rownb & Rho-GTPases~\cite{DesignPrinciplesGeneNetworks} & 33 & 2 & 0.17 & 0.57 & \textbf{40.39} & 0.07 & \textbf{0.74} & 0.56 & 0.11 \\
+    \rownb & Pluripotency~\cite{YachieKinoshita2018} & 36 & 440 & \textbf{DNF} & \textbf{DNF} & NM & 0.16 & \textbf{138.92} & \textbf{DNF} & 0.28 \\
+    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & 276 & 0.37 & 0.43 & NM & 0.07 & \textbf{72.40} & \textbf{DNF} & 0.06 \\
+    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $1000^+$ & N/A & 0.11 & 0.36 & 0.17 & \textbf{DNF} & \textbf{DNF} & 0.06 \\
+    \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & 128 & 0.33 & 0.05 & 0.07 & 0.06 & \textbf{32.66} & \textbf{126.22} & 0.05 \\
+    \rownb & Cacace\_TdevModel~\cite{Cacace2020} & 61 & 28 & \textbf{1.29} & \textbf{5.67} & NM & 0.06 & \textbf{7.51} & \textbf{23.15} & 0.08 \\
+    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $1000^+$ & N/A & \textbf{DNF} & 0.50 & 0.34 & \textbf{DNF} & \textbf{DNF} & 0.33 \\
+    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & \textbf{39.22} & \textbf{1.01} & 0.20 & 0.12 & \textbf{75.81} & \textbf{DNF} & 0.05 \\
+    \rownb & Bcell~\cite{Dutta2019} & 73 & 72 & 0.23 & 0.04 & 0.08 & 0.06 & \textbf{18.95} & \textbf{81.85} & 0.05 \\
+    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $1000^+$ & N/A & 0.09 & 0.55 & 0.37 & \textbf{DNF} & \textbf{DNF} & 0.15 \\
+    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $1000^+$ & N/A & \textbf{107.57} & 0.76 & 0.56 & \textbf{DNF} & \textbf{DNF} & 0.16 \\
 
     \midrule % 100 <= n < 200
-     \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & 78 & \textbf{56.81} & \textbf{4.25} & 0.23 & 0.17 & - & - & 0.19 \\
-    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & 82 & \textbf{DNF} & \textbf{14.01} & 0.27 & 0.20 & - & - & 0.25 \\
-    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $1000^+$ & N/A & 0.54 & 1.09 & 0.84 & - & - & 0.27 \\
-    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $1000^+$ & N/A & 0.16 & 1.07 & 1.06 & - & - & 0.16 \\
-    \rownb & angiofull~\cite{Weinstein2017} & 142 & $1000^+$ & N/A & 0.17 & 1.06 & 0.88 & - & - & 0.23 \\
-    \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & 492 & \textbf{DNF} & \textbf{11.28} & 0.78 & 0.69 & - & - & 0.35 \\
-    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $1000^+$ & N/A & \textbf{DNF} & 1.43 & 1.55 & - & - & 0.19 \\
-    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & $1000^+$ & N/A & \textbf{13.58} & 1.76 & 1.51 & - & - & 0.27 \\
+     \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & 78 & \textbf{56.81} & \textbf{4.25} & 0.23 & 0.17 & \textbf{25.20} & \textbf{DNF} & 0.19 \\
+    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & 82 & \textbf{DNF} & \textbf{14.01} & 0.27 & 0.20 & \textbf{27.55} & \textbf{DNF} & 0.25 \\
+    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $1000^+$ & N/A & 0.54 & 1.09 & 0.84 & \textbf{DNF} & \textbf{DNF} & 0.27 \\
+    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $1000^+$ & N/A & 0.16 & 1.07 & 1.06 & \textbf{DNF} & \textbf{DNF} & 0.16 \\
+    \rownb & angiofull~\cite{Weinstein2017} & 142 & $1000^+$ & N/A & 0.17 & 1.06 & 0.88 & \textbf{DNF} & \textbf{DNF} & 0.23 \\
+    \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & 492 & \textbf{DNF} & \textbf{11.28} & 0.78 & 0.69 & \textbf{DNF} & \textbf{DNF} & 0.35 \\
+    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $1000^+$ & N/A & \textbf{DNF} & 1.43 & 1.55 & \textbf{DNF} & \textbf{DNF} & 0.19 \\
+    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & $1000^+$ & N/A & \textbf{13.58} & 1.76 & 1.51 & \textbf{DNF} & \textbf{DNF} & 0.27 \\
 
     \midrule % 200 <= n
-    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $1000^+$ & N/A & 1.13 & 1.47 & 1.43 & - & - & 0.31 \\
-    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & $1000^+$ & N/A & \textbf{DNF} & 1.52 & 2.26 & - & - & 0.35 \\
-    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $1000^+$ & N/A & \textbf{4.99} & NM & 1.96 & - & - & 0.28 \\
-    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & $1000^+$ & N/A & 0.42 & 2.36 & \textbf{4.91} & - & - & 0.44 \\
-    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & $1000^+$ & N/A & \textbf{DNF} & 2.56 & \textbf{5.62} & - & - & 0.46 \\
-    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & $1000^+$ & N/A & 0.36 & 2.99 & \textbf{4.81} & - & - & 0.37 \\
-    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & $1000^+$ & N/A & \textbf{DNF} & NM & \textbf{18.21} & - & - & 0.79 \\
+    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $1000^+$ & N/A & 1.13 & 1.47 & 1.43 & \textbf{DNF} & \textbf{DNF} & 0.31 \\
+    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & $1000^+$ & N/A & \textbf{DNF} & 1.52 & 2.26 & \textbf{DNF} & \textbf{DNF} & 0.35 \\
+    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $1000^+$ & N/A & \textbf{4.99} & NM & 1.96 & \textbf{DNF} & \textbf{DNF} & 0.28 \\
+    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & $1000^+$ & N/A & 0.42 & 2.36 & \textbf{4.91} & \textbf{DNF} & \textbf{DNF} & 0.44 \\
+    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & $1000^+$ & N/A & \textbf{DNF} & 2.56 & \textbf{5.62} & \textbf{DNF} & \textbf{DNF} & 0.46 \\
+    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & $1000^+$ & N/A & 0.36 & 2.99 & \textbf{4.81} & \textbf{DNF} & \textbf{DNF} & 0.37 \\
+    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & $1000^+$ & N/A & \textbf{DNF} & NM & \textbf{18.21} & \textbf{DNF} & \textbf{DNF} & 0.79 \\
 

### 2023-03-09 trappist/bnet.py

@@ -54,7 +54,7 @@ def read_bnet(fileobj: IO, method: str) -> nx.DiGraph:
     net = nx.DiGraph()
 
     # big bnets…
-    setrecursionlimit(2048)
+    setrecursionlimit(204800)
     for line in fileobj.readlines():
         if line.startswith("#") or line.startswith("targets, factors"):
             continue

### 2023-03-09 trappist/cp.py
use compute_trap_spaces(time_limit)

@@ -16,7 +16,8 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 
-from asyncio import run, get_running_loop
+from asyncio import get_running_loop, run
+from concurrent.futures import ThreadPoolExecutor
 from dataclasses import asdict
 from datetime import timedelta
 from time import perf_counter
@@ -26,7 +27,6 @@ from minizinc import Instance, Model, Result, Solver, Status
 
 import networkx as nx  # TODO maybe replace with lists/dicts
 
-from concurrent.futures import ThreadPoolExecutor
 
 def create_cp(petri_net: nx.DiGraph) -> Model:
     """Create the CP program for the max conflict-free siphons of petri net."""
@@ -83,15 +83,14 @@ def solve_cp(
         max_output == 0 or nsol < max_output
     ):
         start = perf_counter()
-        # result = inst.solve(timeout=remaining, processes=nprocs)
-        #result = run(inst.solve_async(timeout=remaining, processes=nprocs))
 
         try:
             get_running_loop()
             with ThreadPoolExecutor(1) as pool:
-                result = pool.submit(lambda: run(inst.solve_async(timeout=remaining, processes=nprocs))).result()
+                coroutine = inst.solve_async(timeout=remaining, processes=nprocs)
+                result = pool.submit(lambda: run(coroutine)).result()  # noqa: B023
         except RuntimeError:
-            result = run(inst.solve_async(timeout=remaining, processes=nprocs))
+            result = inst.solve(timeout=remaining, processes=nprocs)
 
         end = perf_counter()
         if remaining is not None:

### 2023-03-09 trappist/ilp.py

@@ -16,7 +16,8 @@ You should have received a copy of the GNU General Public License
 along with this program.  If not, see <https://www.gnu.org/licenses/>.
 """
 
-from asyncio import run, get_running_loop
+from asyncio import get_running_loop, run
+from concurrent.futures import ThreadPoolExecutor
 from dataclasses import asdict
 from datetime import timedelta
 from time import perf_counter
@@ -28,7 +29,6 @@ import networkx as nx  # TODO maybe replace with lists/dicts
 
 from .cp import cp_to_bool, zincify
 
-from concurrent.futures import ThreadPoolExecutor
 
 def create_ilp(petri_net: nx.DiGraph) -> Model:
     """Create the ILP program for the max conflict-free siphons of petri net."""
@@ -75,14 +75,14 @@ def solve_ilp(
         max_output == 0 or nsol < max_output
     ):
         start = perf_counter()
-        #result = run(inst.solve_async(timeout=remaining, processes=nprocs))
-        
+
         try:
             get_running_loop()
             with ThreadPoolExecutor(1) as pool:
-                result = pool.submit(lambda: run(inst.solve_async(timeout=remaining, processes=nprocs))).result()
+                coroutine = inst.solve_async(timeout=remaining, processes=nprocs)
+                result = pool.submit(lambda: run(coroutine)).result()  # noqa: B023
         except RuntimeError:
-            result = run(inst.solve_async(timeout=remaining, processes=nprocs))
+            result = inst.solve(timeout=remaining, processes=nprocs)
 
         end = perf_counter()
         if remaining is not None:
@@ -97,7 +97,6 @@ def solve_ilp(
             # non subset constrait for maximality
             non_sub = []
             for k, v in d.items():
-                #if not v:
                 if not v and k != "objective":
                     non_sub.append(k)
             or_string = " + ".join(non_sub)

### 2023-03-10 manuscripts/TCS2022/tcs2022.tex
Finished experimental analysis

@@
 Fourth, we discuss in detail how to compute several special types of trap spaces in a Boolean network.
-Besides minimal trap spaces, these special types also play crucial roles in analyzing Boolean networks~\cite{Rozum2021}.
+Besides minimal trap spaces, these special types also play crucial roles in analyzing and controlling Boolean networks~\cite{Rozum2021}.
 %Fifth, we present the idea for using our Petri net approach to handle the problem of inconsistent and incomplete data in modeling biological systems.
 Fifth, regarding the implementation, we have developed a new converter that directly reads a \texttt{.bnet} file and builds the Petri net encoding, instead of using the \texttt{PNML} conversion of \texttt{bioLQM}~\cite{DBLP:conf/cmsb/TrinhBHS22}.
-Finally, we conduct a more comprehensive benchmark on more real-world models from various sources and randomly generated models to evaluate all the proposed methods (the benchmark conducted in~\cite{DBLP:conf/cmsb/TrinhBHS22} considers only dozens of representative real-world models).
+Finally, we conduct a more extensive benchmark on more real-world models from various sources and randomly generated models to evaluate all the proposed methods (the benchmark conducted in~\cite{DBLP:conf/cmsb/TrinhBHS22} considers only dozens of representative real-world models), with more comprehensive insights are obtained.
 
@@
 To evaluate the performance of the newly proposed methods (implemented as a Python package named \texttt{Trappist}) and the state-of-the-art methods (\texttt{bioLQM}\footnote{\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}}, \texttt{PyBoolNet}~\cite{klarner2015computing,klarner2017pyboolnet}, and \texttt{mpbn}~\cite{Paulev2020}), we compared them on both \texttt{PyBoolNet}'s own model repository and many real-world models from various sources in the literature.
+To our knowledge, these models are a highly representative sample of Boolean models currently available in the literature.
 It is worth noting that \texttt{mpbn}~\cite{Paulev2020} only handles locally-monotonic models, whereas the other methods can handle general models.
-To obtain a more comprehensive comparison, we also used random models generated by a third-party software (i.e., \texttt{BoolNet R} package~\cite{mussel2010boolnet}).
+To obtain a more comprehensive comparison, we also used random models generated by a third-party software \texttt{BoolNet R} package~\cite{mussel2010boolnet}.
 As explained in Section~\ref{sec:case_study}, in our benchmarks, we only searched for the first 1000 minimal trap spaces for each model.
@@
   \label{tab:pyboolnet_repo}
-  \scalebox{0.9}{
+  \scalebox{0.85}{
   \begin{tabular}{rlrrrrrrrrr}
@@
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 minimal trap spaces.
+``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces) within the time limit of three minutes.
 In the case of \texttt{bioLQM}, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of \texttt{bioLQM} because it always requires to compute all minimal trap spaces.
@@
 ``NM'' indicates a non-locally-monotonic model.
-There are four variants of \texttt{Trappist}: \texttt{SAT} (i.e., the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{CP} (i.e., the CP-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{ILP} (i.e., the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and \texttt{ASP} (i.e., the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
+There are four variants of \texttt{Trappist}: \texttt{SAT} (i.e., \texttt{Trappist-MaxSAT}, the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{CP} (i.e., \texttt{Trappist-CP}, the CP-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{ILP} (i.e., \texttt{Trappist-ILP}, the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and \texttt{ASP} (i.e., \texttt{Trappist-ASP}, the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
+
+We first analyze the results of the four variants of \texttt{Trappist}.
+We can see that \texttt{Trappist-MaxSAT} and \texttt{Trappist-ASP} are comparable in most models, but \texttt{Trappist-ASP} is much faster for the jaoude\_thdiff and selvaggio\_emt models where the number of minimal trap spaces is greater than 1000.
+The latter can be explained by the fact that \texttt{Trappist-MaxSAT} follows an iterative approach, i.e., it restarts the search each time a solution is found (see Subsection~\ref{subsec:computation_csp}).
+The latter can be explained by the fact that \texttt{Trappist-MaxSAT} follows an iterative approach, i.e., it restarts the search each time a solution is found (see Subsection~\ref{subsec:computation_csp}).
+This iterative approach may be less efficient than the way ASP solvers use to enumerate multiple solutions (answer sets), which is an advantage of ASP solvers~\cite{DBLP:journals/aicom/GebserKKOSS11}.
+Hence, when the number of solutions increases, the inferiority of \texttt{Trappist-MaxSAT} compared to \texttt{Trappist-ASP} will be exhibited more clearly.
+The two remaining variants, \texttt{Trappist-CP} and \texttt{Trappist-ILP}, are much less efficient than \texttt{Trappist-MaxSAT} and \texttt{Trappist-ASP} in every model, even are \(2000 \times\) slower in some models.
+The first reason for their bad performance is that they are also iterative methods like \texttt{Trappist-MaxSAT}, thus they are not efficient for "enumeration" problems.
+Upon closer inspection, for the Boolean CSP characterizing conflict-free siphons, CP seems to be something that is a "less-efficient-SAT".
+For ILP, it may be even worse, since the problem is purely Boolean (no
+real or integer numbers whatsoever).
+This is confirmed by the observation that for some quite large models (e.g., the grieco\_mapk, zhang\_tlgl, and zhang\_tlgl\_v2 models), \texttt{Trappist-ILP} is much slower than \texttt{Trappist-CP}.
+Note that the inferiority of ILP compared to ASP with respect to the trap space enumeration has been reported in~\cite{klarner2015computing}.
+Hereafter, we shall compare the best variant of \texttt{Trappist} (i.e., \texttt{Trappist-ASP}) with other methods.
 
 As shown in Table~\ref{tab:pyboolnet_repo}, for most of the models of the \texttt{PyBoolNet} repository, the results are comparable with all minimal trap spaces found very fast.
-For 3 of the 29 models, \texttt{mpbn} did not give any answer because these models are locally-monotonic.
-%Note that on some very small models, \texttt{Trappist} is sometimes slower than \texttt{PyBoolNet} and/or \texttt{mpbn}, but still significantly under one second.
-%On the contrary, on every model that was a bit challenging for \texttt{PyBoolNet} or \texttt{mpbn}, the new method is far more efficient with speedups between one and two orders of magnitude.
+However upon closer inspection, we can see some notable differences.
+First, \texttt{Trappist-ASP} is far more efficient than \texttt{bioLQM} in every model with speedups between \(5\times\) and \(16 \times\).
+Second, for small models, \texttt{PyBoolNet} and \texttt{mpbn} are comparable to \texttt{Trappist-ASP}.
+However, on every model that was a bit challenging for \texttt{PyBoolNet} or \texttt{mpbn}, \texttt{Trappist-ASP} is far more efficient with speedups between \(3\times\) and \(5 \times\) for the case of \texttt{mpbn}, and between \(5\times\) and \(1783 \times\) for the case of \texttt{PyBoolNet}.
+In particular, the second best variant of \texttt{Trappist} (i.e., \texttt{Trappist-MaxSAT}) is even far more efficient than \texttt{bioLQM} and \texttt{PyBoolNet}, is comparable to \texttt{mpbn} on every model.
+It is worth noting that for 3 of the 29 models, \texttt{mpbn} did not give any answer because these models are locally-monotonic but all the other methods did, which confirms the limit of \texttt{mpbn} on the applicable class of models.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 	\centering
@@
 
-%		\addplot[
-%			smooth,
-%			thick,
-%			dashed,
-%			color=magenta,
-%			mark=star,
-%		]
-%		coordinates {
-%			(0.5, 144)(1, 177)(2, 196)(5, 223)(10, 250)(20, 265)(50, 288)(100, 309)(180, 323)
-%		}; \addlegendentry{\texttt{ILP}}
+		\addplot[
+			smooth,
+			thick,
+			dashed,
+			color=magenta,
+			mark=star,
+		]
+		coordinates {
+			(0.5, 9)(1, 34)(2, 60)(5, 84)(10, 97)(20, 109)(50, 113)(100, 121)(180, 124)
+		}; \addlegendentry{\texttt{ILP}}
 
@@
 		%		coordinates {
-		%			(0.5, 144)(1, 177)(2, 196)(5, 223)(10, 250)(20, 265)(50, 288)(100, 309)(180, 323)
+		%			(0.5, 9)(1, 34)(2, 60)(5, 84)(10, 97)(20, 109)(50, 113)(100, 121)(180, 124)
 		%		}; \addlegendentry{\texttt{ILP}}
@@
 	\end{tikzpicture}
-	\caption{Cumulative numbers of the \texttt{BBM} models that have less than 1000 minimal trap spaces (above) and \texttt{BBM} models (below) solved by the competing methods with respect to enumerating the first 1000 minimal trap spaces.}
+	\caption{Cumulative numbers of the \texttt{BBM} models that have less than 1000 minimal trap spaces (above) and \texttt{BBM} models (below) solved by the compared methods with respect to enumerating the first 1000 minimal trap spaces.}
 	\label{fig:BBM_cumul}
@@
 
-Figure~\ref{fig:BBM_cumul} (above) shows cumulative numbers of the \texttt{BBM} models that have less than 1000 minimal trap spaces solved by the competing methods with respect to enumerating the first 1000 minimal trap spaces.
+Figure~\ref{fig:BBM_cumul} (above) shows cumulative numbers of the \texttt{BBM} models that have less than 1000 minimal trap spaces solved by the compared methods with respect to enumerating the first 1000 minimal trap spaces.
 The number of such models is 134 (per all 211 models), and 15 of them are non-locally-monotonic.
-We can first see that \dots
-
-Figure~\ref{fig:BBM_cumul} (below) shows cumulative numbers of the \texttt{BBM} models solved by the competing methods (except \texttt{bioLQM}) with respect to enumerating the first 1000 minimal trap spaces.
-\dots
-
-%Table~\ref{tab:bbm_repo} shows the experimental results on the 211 real-world models from the \texttt{BBM} repository.
-%Column 2 expresses the numbers of failures (i.e., did not finish the computation within a time limit of three minutes) of each method.
-%For the case of \texttt{bioLQM}, we only considered the models that have at most 1000 minimal trap spaces.
-%The number of such models is 134 (per all 211 models) and is denoted inside the parentheses.
-%For the case of \texttt{mpbn}, we only considered the models that are locally-monotonic.
-%The number of such models is 187 (per all 211 models) and is denoted inside the parentheses.
-%Columns 3-5 express the average running time (in seconds) of each method for the models having at most 1000 minimal trap spaces, the locally-monotonic models, and all the models, respectively.
-%Note that when computing the average running time, if the running time exceeds 180s, it is considered as 180s.
-%From the results shown in Table~\ref{tab:bbm_repo}, we reported several observations as follows.
+This model set allows us to fairly consider \texttt{bioLQM} for comparison, since \texttt{bioLQM} always requires to compute all minimal trap spaces.
+We can first see that \texttt{Trappist-ASP} and \texttt{Trappist-MaxSAT} are still the two best methods as they can handle every model within 1s as well as they always can handle more models than all the remaining methods on every time limit.
+Second, \texttt{Trappist-CP} is better than \texttt{Trappist-ILP}, which is consistent with their comparison shown in the previous subsection.
+Third, one notable remark is that for the time limit of 100s or 180s, \texttt{Trappist-CP} can handle more models than all \texttt{bioLQM}, \texttt{PyBoolNet}, and \texttt{mpbn}.
+This remark shows that even with a not best implementation, our alternative approach is still better than the state-of-the-art methods on a certain set of real-world models.
+This is supported by the fact that our alternative approach avoids the need for computing prime implicants (as opposed to \texttt{PyBoolNet}) and can handle non-locally-monotonic Boolean networks (as opposed to \texttt{mpbn}).
+
+Figure~\ref{fig:BBM_cumul} (below) shows cumulative numbers of the \texttt{BBM} models solved by the compared methods (except \texttt{bioLQM}, \texttt{Trappist-CP}, and \texttt{Trappist-ILP}) with respect to enumerating the first 1000 minimal trap spaces.
+We omit the results of \texttt{Trappist-CP} and \texttt{Trappist-ILP} because they can handle no model with more than 1000 minimal trap spaces.
+Again, we can see that \texttt{Trappist-ASP} and \texttt{Trappist-MaxSAT} are the two best methods as they can handle every but one model within 5s as well as they always can handle much more models than both \texttt{PyBoolNet} and \texttt{mpbn} on every time limit.
+Note that with the time limit of 0.5s, \texttt{Trappist-ASP} can handle more 14 models than \texttt{Trappist-MaxSAT}, which is opposed to the case of models with less than 1000 minimal trap spaces (see Figure~\ref{fig:BBM_cumul} (above)).
+This observation confirms the disadvantage of \texttt{Trappist-MaxSAT} compared to \texttt{Trappist-ASP} for the case of many minimal trap spaces.
 
@@
 Table~\ref{tab:result_real} shows the obtained experimental results.
-``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces) within the time limit of three minutes.
 A number in bold indicates a ratio greater than or equal to 10 compared to the best result.
@@
   \label{tab:result_real}
-  \scalebox{0.82}{
+  \scalebox{0.76}{
   \begin{tabular}{rlrrrrrrrrr}
@@
 
-The first observation is that for 11 of the 32 models (more than 34\%), \texttt{mpbn} did not give any answer because these models as not locally-monotonic.
-For 22 of the 32 models where \texttt{mpbn} returned the answers, \texttt{mpbn} and \texttt{Trappist} are comparable in computation time, though surprisingly \texttt{mpbn} appears quite slower on average than the best variant of \texttt{Trappist}.
-In particular, for the Rho-GTPases model, \texttt{mpbn} is \(577\times\) slower than the best variant of \texttt{Trappist}.
-However, since \texttt{mbpn} can handle only locally-monotonic models and \texttt{Trappist} can handle general models, it is difficult to further compare between them.
-Hence, we focus on only comparisons among \texttt{bioLQM}, \texttt{PyBoolNet}, and \texttt{Trappist} in the following observations.
+First, we obtained some observations on the four variants of \texttt{Trappist} consistent with the observations obtained in the previous subsections.
+More specifically, \texttt{Trappist-ASP} is still the best variant with the running time is always less than one second for every model, and following by \texttt{Trappist-MaxSAT}.
+In particular, the difference in running time between \texttt{Trappist-ASP} and \texttt{Trappist-MaxSAT} is bigger for larger models or models with more than 1000 minimal trap spaces.
+\texttt{Trappist-CP} and \texttt{Trappist-ILP} still have bad performance, with \texttt{Trappist-CP} is a better than \texttt{Trappist-ILP}.
+They still can handle no model with more than 1000 minimal trap spaces.
+However, \texttt{Trappist-CP} or \texttt{Trappist-ILP} can handle the FT-GRN and Pluripotency models, whereas all \texttt{bioLQM}, \texttt{PyBoolNet}, and \texttt{mpbn} cannot.
+
+Second, \texttt{Trappist-ASP} (even \texttt{Trappist-MaxSAT}) is far more efficient than both \texttt{bioLQM} and \texttt{PyBoolNet} on every model where the comparison is possible.
+For most models, the speedups of \texttt{Trappist-ASP} compared to \texttt{bioLQM} and \texttt{PyBoolNet} are between one and three orders of magnitude.
+This again confirms the superiority of \texttt{Trappist-ASP} compared to the other methods that can handle general Boolean networks.
+
+Third, for 11 of the 32 models (more than 34\%), \texttt{mpbn} did not give any answer because these models as non-locally-monotonic.
+For 21 of the 32 models where \texttt{mpbn} returned the answers, \texttt{mpbn} and \texttt{Trappist-ASP} are comparable in computation time, with \texttt{mpbn} appears quite slower on average.
+In particular, for the Rho-GTPases model, \texttt{mpbn} is \(577\times\) slower than \texttt{Trappist-ASP}.
+This observation along with the comparisons between \texttt{mpbn} and \texttt{Trappist-ASP} in the previous subsections are quite surprising because the ASP encoding of \texttt{mpbn} only requires the DNF for the activation part of a Boolean function, whereas that of \texttt{Trappist-ASP} requires both the activation and inhibition parts (see Subsection~\ref{subsec:computation_asp}).
+However, the reason maybe lies on the differences in the ASP encoding characteristics of the two methods and the fact that \texttt{mpbn} needs to spend time for checking the locally-monotonicity of each Boolean function in a Boolean network.
+It is possible that \texttt{mpbn} may outperform \texttt{Trappist} for a certain set of models, but not for the set of real-world models considered in this article.
+
+Fourth, regarding the comparison of the ASP-based methods (i.e., \texttt{PyBoolNet}, \texttt{mpbn}, and \texttt{Trappist-ASP}), we note that for all the models where \texttt{PyBoolNet} did not finish before the time limit, the timeout occurred during the computation of the prime-implicants.
+Hence, not even a single minimal trap space was output by that method.
+For all the remaining models, once \texttt{PyBoolNet} went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
+Hence, with the experimental results shown in this subsection as well as the two previous subsections, the practical differences between the ASP encoding of \texttt{Trappist-ASP} and that of \texttt{PyBoolNet} are not distinctly exposed.
+The fact that our new ASP encoding is guaranteed to be linear in the number of nodes of the original model (see Subsection~\ref{subsec:computation_asp}) does not seem to be crucial here, however a much deeper analysis of those cases shall be shown in the next subsection.
 
@@
 In addition, each node has only three input nodes, i.e., the number of prime-implicants of the associated Boolean function is small.
-Hence, \texttt{PyBoolNet} always passed the phase of computing prime-implicants in every model even within 1s, which enables us to compare the ASP encoding of \texttt{PyBoolNet} and that of \texttt{Trappist}.
-
-\begin{figure}[ht]
+Hence, \texttt{PyBoolNet} always passed the phase of computing prime-implicants in every model even within one second, which enables us to compare the ASP encoding of \texttt{PyBoolNet} and that of \texttt{Trappist-ASP}.
+
+\begin{figure}[!ht]
 	\centering
@@
 	coordinates {
-		(0.5, 143)(1, 176)(2, 195)(5, 222)(10, 249)(20, 264)(50, 287)(100, 308)(180, 322)
+		(0.5, 13)(1, 81)(2, 127)(5, 155)(10, 165)(20, 178)(50, 202)(100, 214)(180, 226)
 	}; \addlegendentry{\texttt{CP}}
@@
 	coordinates {
-		(0.5, 144)(1, 177)(2, 196)(5, 223)(10, 250)(20, 265)(50, 288)(100, 309)(180, 323)
+		(0.5, 0)(1, 0)(2, 0)(5, 1)(10, 1)(20, 1)(50, 9)(100, 25)(180, 39)
 	}; \addlegendentry{\texttt{ILP}}
@@
 	\end{tikzpicture}
-	\caption{Cumulative numbers of random models solved by the competing methods with respect to enumerating all the minimal trap spaces.}
+	\caption{Cumulative numbers of random models solved by the compared methods with respect to enumerating all the minimal trap spaces.}
 	\label{fig:random_cumul_first}
@@
 
-Figure~\ref{fig:random_cumul_first} shows cumulative numbers of random models solved by the competing methods with respect to enumerating all the minimal trap spaces.
-The number of succeeded models within three minutes for each method is as follows: \texttt{bioLQM} (0), \texttt{PyBoolNet} (320), \texttt{mpbn} (0), \texttt{Trappist-maxSAT} (338), \texttt{Trappist-CP} (0), \texttt{Trappist-ILP} (0), \texttt{Trappist-ASP} (349).
-We can see that \texttt{Trappist-ASP} is the method that could handle the most number of models.
-Note that none of the other methods could handle the model failed by \texttt{Trappist-ASP}.
-Upon closer inspection, we obtained several observations consistent with those obtained for real-world models as follows.
+Figure~\ref{fig:random_cumul_first} shows cumulative numbers of random models solved by the compared methods with respect to enumerating all the minimal trap spaces.
+The number of succeeded models within three minutes for each method is: \texttt{bioLQM} (0), \texttt{PyBoolNet} (320), \texttt{mpbn} (0), \texttt{Trappist-maxSAT} (338), \texttt{Trappist-CP} (226), \texttt{Trappist-ILP} (39), \texttt{Trappist-ASP} (349).
+We can see that \texttt{Trappist-ASP} is the only method that can handle every but one model.
+Note that none of the other methods can handle the only model failed by \texttt{Trappist-ASP}.
+We also obtained some observations consistent with those obtained for real-world models.
+More specifically, \texttt{Trappist-MaxSAT} is still the second best method and \texttt{Trappist-CP} is better than \texttt{Trappist-ILP}.
+Upon closer inspection, we obtained several notable observations as follows.
 
@@
 
-Second, surprisingly \texttt{bioLQM} could not handle any model.
+Second, surprisingly \texttt{bioLQM} cannot handle any model.
 One of the reason may be that the BDD characterizing all trap spaces is too large, and its computation is slow.
@@
 
-%Third, \texttt{PyBoolNet} could handle every model of network size less than or equal to 300.
-%It only failed in one model of network size 350 but 29 models of network size 400.
-%The average running time vastly increases as the network size increases.
-%As compared to the four methods of our approach, the performance of \texttt{PyBoolNet} is comparable for the 100-node and 200-node models.
-%However, from \(n = 250\), the performance difference is exhibited more clearly.
-
-Third, \dots
-
-Finally, \dots
+Third, for every time limit, \texttt{Trappist-ASP} can always handle much models than \texttt{PyBoolNet}, ranging from 29 to 65 more models.
+Since the time for the phase of computing prime-implicants of \texttt{PyBoolNet} is negligible in every model, most of the running time of \texttt{PyBoolNet} was spent for its ASP solving phase.
+Hence, we can easily see that the ASP encoding of \texttt{Trappist-ASP} is much better than that of \texttt{PyBoolNet}. 
+This observation is consistent with the theoretical comparison in the ASP encoding between \texttt{Trappist-ASP} and \texttt{PyBoolNet} mentioned in Subsection~\ref{subsec:computation_asp}.
+
+%More models handled by Trappist-ASP: 38, 37, 51, 54, 57, 65, 54, 41, 29
+
+\subsection{Experimental summary}
+\label{subsec:Experimental_summary}
+
+We have tested our alternative approach on many Boolean network models of various sizes and types (e.g., real-world models, randomly generated models).
+This indicates the high coverage and comprehensiveness of the experiments.
+
+Among the four variants of the alternative approach, \texttt{Trappist-ASP} is the best method as it vastly outperforms all the other variants.
+The second best one is \texttt{Trappist-MaxSAT}.
+The two remaining variants (i.e., \texttt{Trappist-CP} and \texttt{Trappist-ILP}) give bad performance for most models.
+However, for some certain cases, they are even better than all the state-of-the-art methods (i.e., \texttt{bioLQM}, \texttt{PyBoolNet}, and \texttt{mpbn}).
+This is evidence for the advantages of the alternative approach compared to the state-of-the-art ones.
+
+Regarding general Boolean networks, \texttt{Trappist-ASP} (even \texttt{Trappist-MaxSAT}) is far more efficient than both \texttt{bioLQM} and \texttt{PyBoolNet}.
+The speedups of \texttt{Trappist-ASP} or \texttt{Trappist-MaxSAT} are large, even between one and three orders of magnitude for most models.
+In addition, the experimental results also confirm that the ASP encoding of \texttt{Trappist-ASP} is much better than that of \texttt{PyBoolNet}.
+
+Regarding locally-monotonic Boolean networks, the performance of \texttt{mpbn} is comparable to that of \texttt{Trappist-ASP} or \texttt{Trappist-MaxSAT}.
+However, \texttt{mpbn} is quite slower than \texttt{Trappist-ASP} in average.
+This shows the practical advantage of \texttt{Trappist-ASP} compared to \texttt{mpbn}, though its ASP encoding may be more complex than that of \texttt{mpbn} in theory.
 
@@
 The proposed methods have been evaluated on many real-world models from the literature as well as randomly generated models.
-The experimental results show that the new approach vastly outperforms all the state-of-the-art methods in terms of general Boolean networks and is comparable to the \texttt{mpbn} method even better in average in terms of locally-monotonic Boolean networks.
+The experimental results show that the new approach vastly outperforms all the state-of-the-art methods in terms of general Boolean networks and is comparable to the \texttt{mpbn} method even much better in average in terms of locally-monotonic Boolean networks.
 We believe that this opens up the way to a much better analysis of large Boolean networks, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
@@
 Exploring those connections might lead both to interesting topics of research for Petri nets, like a notion of trap-spaces, and for Boolean networks.
-We also believe that the connection between trap spaces of Boolean networks and siphons of Petri nets can be a very useful tool for exploring and proving more new properties of trap spaces in Boolean networks, as we have used it to successfully prove the separation of minimal trap spaces.
-Diving into this direction is one of our future work.
+We also believe that the connection between trap spaces of Boolean networks and siphons of Petri nets can be a very useful tool for exploring and proving more new properties of trap spaces in Boolean networks, as we have used it to successfully prove the independence of trap spaces to the update scheme and the separation of minimal trap spaces.
+Diving into this direction is promising and one of our future work.
 

### 2023-03-10 bench.sh

@@ -1,4 +1,4 @@
 #!/usr/bin/env bash
 
 docker build -t trappist $(dirname -- ${BASH_SOURCE[0]})
-docker run -v $(pwd):/notebook/work trappist jupyter nbconvert --to notebook --output work/Benchmark.ipynb --execute --allow-errors Benchmark.ipynb
+docker run -v $(pwd):/notebook/work trappist jupyter nbconvert --to notebook --output work/Benchmark-TCS-Random.ipynb --execute --allow-errors Benchmark-TCS-Random.ipynb

### 2023-03-11 manuscripts/TCS2022/tcs2022.tex
Justification for considering special types of trap spaces

@@
 
-In the field of systems biology, biologists may want to compute more special types of trap spaces beyond minimal trap spaces~\cite{klarner2017pyboolnet}, which also play crucial roles in analysis of Boolean networks.
+In the field of systems biology, biologists may want to compute more special types of trap spaces beyond minimal trap spaces~\cite{klarner2017pyboolnet}, which also play crucial roles in analysis and control of Boolean networks~\cite{cifuentes2020control,Rozum2021}.
 We shall show that our proposed methods can be easily adjusted to compute popular types of trap spaces.
-We illustrate the adjustments via the ASP-based method (see Subsection~\ref{subsec:computation_asp}), but these adjustments are completely applicable for other approaches such as MaxSAT, CP, and ILP\@.
+We illustrate the adjustments via the ASP-based method (see Subsection~\ref{subsec:computation_asp}) because ASP is declarative by nature, but these adjustments are completely applicable for other approaches such as MaxSAT, CP, and ILP\@.
 
@@
 Second, we consider fixed points in Boolean networks.
+To date, the analysis of the fixed points of a Boolean network remains
+a very useful tool in understanding the behavior of complex biological models not only due to the fact that in some cases the full computation of complex attractors remains intractable, but also because for many biological systems, the expected long-term behavior is not cyclic~\cite{corral2021interplay}.
+Furthermore, the fixed point computation is also the crucial starting point for several state-of-the-art methods for computing complex attractors of Boolean networks~\cite{DBLP:conf/bcb/TrinhHB22}.
 Let \(s\) be a fixed point of a Boolean network \(\mathcal{N}\).
@@
 Third, we consider the trap spaces intersecting a given subspace \(m^{*}\) of a Boolean network.
+Such trap spaces are used in the trap space-based control method~\cite{cifuentes2020control}.
 A trap space \(m\) intersects \(m^{*}\) if and only if \(S_{\mathcal{N}}[m] \cap S_{\mathcal{N}}[m^{*}] \neq \emptyset\).
@@
 Finally, we consider the trap spaces that are inside a given subspace \(m^{*}\) of a Boolean network.
+Such trap spaces are used in the iterative procedure of building the succession diagram of a Boolean network~\cite{Rozum2021}, which is hierarchical.
 We first adjust \(\mathcal{L}\) to characterize all such trap spaces.

### 2023-03-15 manuscripts/TCS2022/tcs2022.tex
Pages 1--9, nothing crucial

@@
 \newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
+
+\usepackage{amsmath}
+\newcommand{\IN}{\mathit{IN}}
+\DeclareMathOperator{\pred}{pred}
+\let\succ\relax
+\DeclareMathOperator{\succ}{succ}
 
@@
 %% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
+%% or \MSC[2008] code \sep code 2000 is the default
 
@@
 By this approach, it requires the computation of all solutions, whereas the ASP-based methods~\cite{klarner2015computing,Paulev2020} can start enumerating them as they are found.
-Moreover, the main issue with the BDD-based method is that the number of generic trap spaces of a Boolean network may be extremely larger than the number of minimal trap spaces of this Boolean network.
+Moreover, the main issue with the BDD-based method is that the number of generic trap spaces of a Boolean network may be extremely larger than its number of minimal trap spaces.
 This issue limits the efficiency of the BDD-based method.
@@
 It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
-We then demonstrate its efficiency and compare it to the state-of-the-art methods for computing minimal trap spaces in Boolean networks on many real-world models from various sources in the literature and randomly generated models.
+We then demonstrate its efficiency and compare it to the state-of-the-art methods for computing minimal trap spaces of Boolean networks on many real-world models from various sources in the literature and on randomly generated models.
 
@@
     \item \(F = \{f_1, \dots, f_n\}\) is the set of update functions.
-    Each function \(f_i\) is associated with node \(v_i\) and satisfies \(f_i \colon \mathbb{B}^{\vert IN(v_i)\vert} \mapsto \mathbb{B}\) where \(\mathbb{B} = \{0, 1\}\) and \(IN(v_i)\) denotes the set of input nodes of \(v_i\).
+    Each function \(f_i\) is associated with node \(v_i\) and satisfies \(f_i \colon \mathbb{B}^{\vert \IN(v_i)\vert} \mapsto \mathbb{B}\) where \(\mathbb{B} = \{0, 1\}\) and \(\IN(v_i)\) denotes the set of input nodes of \(v_i\).
     Note that a node \(v_i \in V\) is called a \emph{source} node if and only if \(f_{i} = v_i\).
@@
 We denote the set of all possible states of a Boolean network \(\mathcal{N}\) by \(\mathcal{S}_{\mathcal{N}} = \mathbb{B}^n\).
-At each time step \(t\), node \(v_i\) can update its state by
-\[v_i(t + 1) = f_i(v(t))\]
-where \(v(t)\) is the state of \(\mathcal{N}\) at time \(t\) and \(v_i(t + 1)\) is the state of node \(v_i\) at time \(t + 1\).
-Note that for simplicity, we write \(f_i(v(t))\) even \(IN(v_i) \subset V\) (i.e., \(IN(v_i)\) does not contain some nodes of \(V\)).
-An update scheme of a Boolean network specifies the way that the nodes update their states through time evolution~\cite{thomas1991regulatory}.
+At each time step \(t\), node \(v_i\) can, depending on the update scheme, update its state by
+\[v_i(t + 1) =
+  \begin{cases}
+    &f_i(v(t))\\
+    \text{or} &v(t)
+  \end{cases}
+\]
+where \(v(t)\) is the state of \(\mathcal{N}\) at time \(t\). % and \(v_i(t + 1)\) is the state of node \(v_i\) at time \(t + 1\).
+Note that for simplicity, we write \(f_i(v(t))\) even if \(\IN(v_i) \subsetneq V\) (i.e., \(\IN(v_i)\) does not contain some nodes of \(V\)).
+An update scheme of a Boolean network specifies which nodes update their states, as defined above, through time evolution~\cite{thomas1991regulatory}.
 Following the update scheme, the Boolean network transits from a state to another state (possibly identical).
@@
 Then the dynamics of \(\mathcal{N}\) is captured by the directed graph \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\) called the State Transition Graph (STG).
-There are two main types of update schemes~\cite{thomas1991regulatory}: synchronous, where all the nodes are update simultaneously, and fully asynchronous, where only one node is nondeterministically selected to be updated.
+There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: synchronous, where all the nodes are update simultaneously, and fully asynchronous, where only one node is selected to be updated.
 
@@
 
-An non-empty set \(T \subseteq \mathcal{S}_{\mathcal{N}}\) is a trap set with respect to \(\rightarrow\) if for every \(x \in T\) and \(y \in S\) with \(x \rightarrow y\) it holds that \(y \in T\)~\cite{klarner2015computing}.
+A non-empty set \(T \subseteq \mathcal{S}_{\mathcal{N}}\) is a trap set with respect to \(\rightarrow\) if for every \(x \in T\) and \(y \in S\) with \(x \rightarrow y\) it holds that \(y \in T\)~\cite{klarner2015computing}.
 An attractor of \(\mathcal{N}\) with respect to \(\rightarrow\) can be defined as an inclusion-wise minimal trap set of \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\).
 An attractor can be also seen as a terminal strongly connected component of \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\)~\cite{chatain2014characterization}.
-An attractor of size 1 is called a fixed point, otherwise a cyclic attractor~\cite{klarner2015computing}.
+An attractor of size 1 is called a fixed point, otherwise it is called a cyclic or complex attractor~\cite{klarner2015computing}.
 
@@
 \mathcal{S}_{\mathcal{N}}[m] := \{s \in \mathcal{S}_{\mathcal{N}} \mid \forall v \in D_m \colon s(v) = m(v)\}.
-\] For example, \(m = \star\star1\) (for simplicity, we write subspaces likes states) means that \(D_m = \{v_3\}, m(v_3) = 1\), and it is equivalent to the set of states \(\{001, 011, 101, 111\}\).
+\] For example, \(m = \star\star1\) (for simplicity, we shall write subspaces likes states as a sequence of values) means that \(D_m = \{v_3\}, m(v_3) = 1\), and it is equivalent to the set of states \(\{001, 011, 101, 111\}\).
 We denote \(\mathcal{S}_{\mathcal{N}}^{\star} = (\mathbb{B} \cup \{\star\})^n\) the set of all possible subspaces of \(\mathcal{N}\).
@@
 For example, let us consider the Boolean network shown in Example~\ref{example:BN}.
-Figure~\ref{fig:stg} shows the dynamics of this model under the fully asynchronous update (i.e., only one node is nondeterministically selected in order to be updated at each time step).
-The model has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
-Since \(m_1 < m_2\), \(m_1\) is a minimal trap space of the Boolean network.
+Figure~\ref{fig:stg} shows the dynamics of this model under the fully asynchronous update (i.e., only one node is updated at each time step).
+The model has all trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
+Since \(m_1 < m_2\), \(m_1\) is the only minimal trap space of the Boolean network.
 
@@
 Marking \(m\) can be seen as a subset of \(P\) that contains all marked places by \(m\).
-We shall write \(pred(x)\) (resp.\ \(succ(x)\)) to represent the set of vertices that have a (non-zero weighted) arc leading to (resp.\ coming from) \(x\).
+We shall write \(\pred(x)\) (resp.\ \(\succ(x)\)) to represent the set of vertices that have a (non-zero weighted) arc leading to (resp.\ coming from) \(x\).
 In this work, we consider a class of Petri nets called 1-safe Petri nets where every place has at most 1 token and all arcs are of weight 1.
 In this case, weights are implicitly omitted in the arcs of a Petri net.
-Then, a transition \(t \in T\) is \emph{enabled} at a marking \(m\) if and only if \(pred(t) \subseteq m\).
+Then, a transition \(t \in T\) is \emph{enabled} at a marking \(m\) if and only if \(\pred(t) \subseteq m\).
 A marking \(m\) is called a deadlock if there are no enabled transitions at \(m\).
-The firing of \(t\) leads to a new marking \(m'\) specified by \(m' = (m \backslash pred(t)) \cup succ(t)\).
+The firing of \(t\) leads to a new marking \(m'\) specified by \(m' = (m \backslash \pred(t)) \cup \succ(t)\).
 Note that when multiple transitions are enabled, we need to embed one firing scheme (similar to the update scheme of a Boolean network) to the Petri net.
@@
 
-\noindent{\small \verb|java -jar GINsim.jar -lqm <input.{sbml,bnet,zginml,...}> <output.pnml>|}
+{\small \verb|java -jar GINsim.jar -lqm <input.{sbml,bnet,...}> <output.pnml>|}
 
@@
 This does not change the worst-case-complexity (obtaining a single DNF being exponential) but might matter a lot in practice.
-As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM} and directly in our tool using the \texttt{pyeda}\footnote{\url{https://pyeda.readthedocs.io/en/latest/}} library, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.
+As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM} or directly in our tool using the \texttt{pyeda}\footnote{\url{https://pyeda.readthedocs.io/en/latest/}} library, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.
 
@@
   A \emph{siphon} of a Petri net \((P, T, W)\) is a set of places \(S\) such that:
-  \[\forall t\in T, S\cap succ(t)\not =\emptyset\Rightarrow S\cap pred(t)\not =\emptyset.\]
+  \[\forall t\in T, S\cap \succ(t)\not =\emptyset\Rightarrow S\cap \pred(t)\not =\emptyset.\]
   Note that \(\emptyset\) is trivially a siphon.
@@
 
-Let \(pred(S) := \bigcup_{s \in S}pred(s)\) and \(succ(S) := \bigcup_{s \in S}succ(s)\).
-If \(S = \emptyset\), then conventionally \(pred(S) = succ(S) = \emptyset\).
+Let \(\pred(S) := \bigcup_{s \in S}\pred(s)\) and \(\succ(S) := \bigcup_{s \in S}\succ(s)\).
+If \(S = \emptyset\), then conventionally \(\pred(S) = \succ(S) = \emptyset\).
 We have an important property on siphons~\cite{DBLP:journals/isci/LiuB16} as follows.
@@
   Let \(S\) be a siphon of a Petri net \((P, T, W)\).
-  Then \(pred(S) \subseteq succ(S)\).
+  Then \(\pred(S) \subseteq \succ(S)\).
 
@@
   Assume that \(S\) is not a siphon of \(\mathcal{P}\).
-  Then, there is a transition \(t \in T\) such that \(S \cap succ(t) \neq \emptyset\) but \(S \cap pred(t) = \emptyset\).
-  This implies that there is a place \(p \in S\) such that \(p \in succ(t)\) but \(p \not \in pred(t)\).
+  Then, there is a transition \(t \in T\) such that \(S \cap \succ(t) \neq \emptyset\) but \(S \cap \pred(t) = \emptyset\).
+  This implies that there is a place \(p \in S\) such that \(p \in \succ(t)\) but \(p \not \in \pred(t)\).
   Let \(v\) be the corresponding node in \(\mathcal{N}\) of \(p\).
@@
   %In addition, there is also no arc or a bidirectional arc between \(t\) and another place rather than \(p_v\) and \(\overline{p}_v\).
-  We follow the following procedure to find a state \(s \in \mathcal{S}_{\mathcal{N}}[m]\) such that \(m_s(p') = 1, \forall p' \in pred(t)\) where \(m_s\) is the corresponding marking in \(\mathcal{P}\) of \(s\).
-  For every place \(p' \in pred(t)\), let \(p''\) be the complementary place of \(p'\) and \(v'\) be the corresponding node in \(\mathcal{N}\) of \(p'\) and \(p''\).
+  We follow the following procedure to find a state \(s \in \mathcal{S}_{\mathcal{N}}[m]\) such that \(m_s(p') = 1, \forall p' \in \pred(t)\) where \(m_s\) is the corresponding marking in \(\mathcal{P}\) of \(s\).
+  For every place \(p' \in \pred(t)\), let \(p''\) be the complementary place of \(p'\) and \(v'\) be the corresponding node in \(\mathcal{N}\) of \(p'\) and \(p''\).
   If \(p'' \not \in S\), then \(v' \not \in D_m\) and we can always set a Boolean value to \(s(v')\) such that \(s \in \mathcal{S}_{\mathcal{N}}[m]\) and \(m_s(p') = 1\).
@@
 Instead, we can first compute the mirror \(S_m\) of \(m\) in the Petri net encoding.
-Then, by Proposition~\ref{prop:siphon_set} and Theorem~\ref{theo:ts_2_sp}, we can check if \(pred(S_m) \subseteq succ(S_m)\).
+Then, by Proposition~\ref{prop:siphon_set} and Theorem~\ref{theo:ts_2_sp}, we can check if \(\pred(S_m) \subseteq \succ(S_m)\).
 Note that the Petri net construction is less computationally demanding than the prime-implicant computation because it only requires computing generic (not prime) implicants of the Boolean functions~\cite{chatain2014characterization}.
@@
 Suppose that \(S\) is a generic siphon of \(\mathcal{P}\).
-If a place \(p\) should belong to \(S\), then by Proposition~\ref{prop:siphon_set} all the transitions in \(pred(p)\) must belong to \(succ(S)\).
-A transition \(t\) belongs to \(succ(S)\) if and only if there is at least one place \(p'\) in \(S\) such that \(p' \in pred(t)\).
-Hence, for each transition \(t \in pred(p)\), we can state that
+If a place \(p\) should belong to \(S\), then by Proposition~\ref{prop:siphon_set} all the transitions in \(\pred(p)\) must belong to \(\succ(S)\).
+A transition \(t\) belongs to \(\succ(S)\) if and only if there is at least one place \(p'\) in \(S\) such that \(p' \in \pred(t)\).
+Hence, for each transition \(t \in \pred(p)\), we can state that
 \begin{equation}
 \label{eq:siphon}
-p \in S \Rightarrow \bigvee_{p' \in pred(t)}p' \in S.
+p \in S \Rightarrow \bigvee_{p' \in \pred(t)}p' \in S.
 \end{equation}
-The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\) fully characterizes all generic siphons of a Petri net and has been used with SAT solvers in~\cite{oanea2010new,nabli2016enumerating}.
+The system of all the rules of the above form with respect to all pairs \((p, t)\) where \(p \in P, t \in T, t \in \pred(p)\) fully characterizes all generic siphons of a Petri net and has been used with SAT solvers in~\cite{oanea2010new,nabli2016enumerating}.
 To make \(S\) to be a conflict-free siphon, we need to add to the system the rule
@@
     \item \(C = \{\neg p_v \vee \neg \overline{p}_v = 1 \mid \forall v \in V\} \bigwedge
-\{(p = 1 \rightarrow \bigvee_{p' \in pred(t)}p' = 1) \mid p \in P, t \in pred(p)\}\).
+\{(p = 1 \rightarrow \bigvee_{p' \in \pred(t)}p' = 1) \mid p \in P, t \in \pred(p)\}\).
   \end{itemize}
@@
 \[
-  (\neg p \vee \bigvee_{p' \in pred(t)}p'), \forall p \in P, \forall t \in pred(p).
+  (\neg p \vee \bigvee_{p' \in \pred(t)}p'), \forall p \in P, \forall t \in \pred(p).
 \]
@@
 The set of all atoms in \(\mathcal{L}\) is given as \(\mathcal{A} = \bigcup_{v \in V}\{\verb|p-v|, \verb|n-v|\}\).
-For each pair \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\), we translate the rule (\ref{eq:siphon}) into the ASP rule
+For each pair \((p, t)\) where \(p \in P, t \in T, t \in \pred(p)\), we translate the rule (\ref{eq:siphon}) into the ASP rule
 \[
@@
 \]
-where \(\verb|a| \in \mathcal{A}\) is the atom representing place \(p\) and \(\{\verb|a_1|, \dots,\verb|a_k|\} \subseteq \mathcal{A}\) is the set of atoms representing places in \(pred(t)\).
+where \(\verb|a| \in \mathcal{A}\) is the atom representing place \(p\) and \(\{\verb|a_1|, \dots,\verb|a_k|\} \subseteq \mathcal{A}\) is the set of atoms representing places in \(\pred(t)\).
 The rule (\ref{eq:conflict}) is translated into the ASP rule
@@
 The set of all binary variables in \(\mathcal{I}\) is \(\bigcup_{v \in V}\{\verb|p-v|, \verb|n-v|\}\).
-For each pair \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\), we translate the rule (\ref{eq:siphon}) into the ILP inequality
+For each pair \((p, t)\) where \(p \in P, t \in T, t \in \pred(p)\), we translate the rule (\ref{eq:siphon}) into the ILP inequality
 \[
@@
 \]
-where \verb|a| is the binary variable representing place \(p\) and \{\verb|a_1|, \dots,\verb|a_k|\} is the set of binary variable representing places in \(pred(t)\).
+where \verb|a| is the binary variable representing place \(p\) and \{\verb|a_1|, \dots,\verb|a_k|\} is the set of binary variable representing places in \(\pred(t)\).
 The rule (\ref{eq:conflict}) is translated into the ILP inequality

### 2023-03-15 manuscripts/TCS2022/tcs2022.tex
Section 3

@@
   This implies that there is a place \(p \in S\) such that \(p \in \succ(t)\) but \(p \not \in \pred(t)\).
-  Let \(v\) be the corresponding node in \(\mathcal{N}\) of \(p\).
+  Let \(v\) be the node in \(\mathcal{N}\) corresponding to \(p\).
   By the characteristics of the encoding~\cite{chaouiya2004qualitative}, there is a directional arc from \(t\) to \(p\) and a directional arc from the complementary place of \(p\) to \(t\).
@@
 
-From the proof of Theorem~\ref{theo:ts_2_sp}, we can see that this theorem still holds for any update scheme of the Boolean network.
+From the proof of Theorem~\ref{theo:ts_2_sp}, we can see that the theorem holds for any update scheme associated to the Boolean network.
 Since the Petri net encoding of a Boolean network is independent of its update scheme and siphons are a static property of a Petri net, we can imply that trap spaces of a Boolean network are independent of its update scheme.
 Note that the original proof for this property of trap spaces (see Theorem 1 of~\cite{klarner2015computing}) only considers the two popular update schemes (i.e., synchronous and fully asynchronous).
-This exhibits the very first theoretical application of the connection between trap spaces of Boolean networks and siphons of Petri nets.
+Theorem~\ref{theo:ts_2_sp} exhibits the very first theoretical application of the connection between trap spaces of Boolean networks and siphons of Petri nets.
 
@@
 We use it to prove a property of minimal trap spaces, which has surprisingly not been formally proved.
+\sylvain{I'm not sure how much we should insist on the ``surprise''.
+The proof directly on trap-spaces is quite obvious and follows the same structure: the intersection would be a trap-space, and smaller than the original ones, hence a contradiction\ldots}
 Specifically, all minimal trap spaces of a Boolean network are mutually disjoint.
@@
 
-A naturally computational application of Theorem~\ref{theo:ts_2_sp} is that we can efficiently decide whether a subspace \(m\) is a trap space.
+A natural computational application of Theorem~\ref{theo:ts_2_sp} is that we can efficiently decide whether a subspace \(m\) is a trap space.
 In \texttt{PyBoolNet}~\cite{klarner2017pyboolnet}, this is checked by using the percolation on the prime-implicants of the Boolean functions.
@@
 See Subsection~\ref{subsec:computation_special_ts} for more discussions about many special types of trap spaces.
-It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap spaces that might be useful for the analysis of Petri nets, but this is out of the scope of this present article.
-
-It is noted that there are no existing methods specifically designed for computing maximal conflict-free siphons (even maximal siphons) of a Petri net.
+It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap spaces that might be useful for the analysis of Petri nets, but this is out of the scope of the present article.
+
+Note that there are no existing methods specifically designed for computing maximal conflict-free siphons (even maximal siphons) of a Petri net.
 The reason might be that researchers mainly focus on minimal generic siphons~\cite{DBLP:journals/isci/LiuB16} in the field of Petri nets.

### 2023-03-15 manuscripts/TCS2022/tcs2022.tex
Section 4

@@
 In the field of systems biology, biologists may want to compute more special types of trap spaces beyond minimal trap spaces~\cite{klarner2017pyboolnet}, which also play crucial roles in analysis and control of Boolean networks~\cite{cifuentes2020control,Rozum2021}.
-We shall show that our proposed methods can be easily adjusted to compute popular types of trap spaces.
+We shall show that our proposed methods can be easily adjusted to compute such popular types of trap spaces.
 We illustrate the adjustments via the ASP-based method (see Subsection~\ref{subsec:computation_asp}) because ASP is declarative by nature, but these adjustments are completely applicable for other approaches such as MaxSAT, CP, and ILP\@.
 
-First, the work by~\cite{Rozum2021} uses the concept of stable motifs to build the succession diagram of a Boolean network, a summary of the decisions in the network dynamics that lead to successively more restrictive nested stable motifs.
+First, the work by~\cite{Rozum2021} uses the concept of \emph{stable motifs} to build the succession diagram of a Boolean network, a summary of the decisions in the network dynamics that lead to successively more restrictive nested stable motifs.
 The succession diagram is useful for control and decision making on this Boolean network.
@@
 
-Second, we consider fixed points in Boolean networks.
+Second, we consider \emph{fixed points} in Boolean networks.
 To date, the analysis of the fixed points of a Boolean network remains
@@
 
-Third, we consider the trap spaces intersecting a given subspace \(m^{*}\) of a Boolean network.
+Third, we consider the trap spaces \emph{intersecting} a given subspace \(m^{*}\) of a Boolean network.
 Such trap spaces are used in the trap space-based control method~\cite{cifuentes2020control}.
@@
 
-Finally, we consider the trap spaces that are inside a given subspace \(m^{*}\) of a Boolean network.
+Finally, we consider the trap spaces that are \emph{inside} a given subspace \(m^{*}\) of a Boolean network.
 Such trap spaces are used in the iterative procedure of building the succession diagram of a Boolean network~\cite{Rozum2021}, which is hierarchical.

### 2023-03-16 manuscripts/TCS2022/tcs2022.tex
Done for this pass…

@@
 
+\sylvain{I think that the result above (0.19s) is only about the F5 model, but do we also have the time for the F4 model, if so, it should be nice to include it too.}
+
 \section{Evaluation}%
@@
 
-To evaluate the performance of the newly proposed methods (implemented as a Python package named \texttt{Trappist}) and the state-of-the-art methods (\texttt{bioLQM}\footnote{\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}}, \texttt{PyBoolNet}~\cite{klarner2015computing,klarner2017pyboolnet}, and \texttt{mpbn}~\cite{Paulev2020}), we compared them on both \texttt{PyBoolNet}'s own model repository and many real-world models from various sources in the literature.
-To our knowledge, these models are a highly representative sample of Boolean models currently available in the literature.
+To evaluate the performance of the newly proposed methods (implemented as a Python package named \texttt{Trappist} and available on the Python package index\footnote{\url{https://pypi.org/project/trappist/}}) and the state-of-the-art methods (\texttt{bioLQM}\footnote{\url{http://colomoto.org/biolqm/doc/tools-trapspace.html}}, \texttt{PyBoolNet}~\cite{klarner2015computing,klarner2017pyboolnet}, and \texttt{mpbn}~\cite{Paulev2020}), we compared them on both \texttt{PyBoolNet}'s own model repository and many real-world models from various sources in the literature.
+To our knowledge, these models are a highly representative sample of Boolean models currently available.
 It is worth noting that \texttt{mpbn}~\cite{Paulev2020} only handles locally-monotonic models, whereas the other methods can handle general models.
@@
 
-All the models and a Jupyter notebook realizing the benchmarks can be found at~\url{https://github.com/soli/trap-spaces-as-siphons}.
+All the models and some Jupyter notebooks realizing the benchmarks (and named \texttt{TCS-Benchmark-<...>.ipynb}) can be found at~\url{https://github.com/soli/trap-spaces-as-siphons/}.
 These can be run on a Docker image in the cloud by clicking the ``Binder'' button.
@@
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 minimal trap spaces.
-``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces) within the time limit of three minutes.
+``DNF'' means that the method did not finish the computation within the time limit of three minutes.
 In the case of \texttt{bioLQM}, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of \texttt{bioLQM} because it always requires to compute all minimal trap spaces.
@@
 ``NM'' indicates a non-locally-monotonic model.
+\sylvain{should we also highlight in some way the (tied-)best result?}
 There are four variants of \texttt{Trappist}: \texttt{SAT} (i.e., \texttt{Trappist-MaxSAT}, the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{CP} (i.e., \texttt{Trappist-CP}, the CP-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{ILP} (i.e., \texttt{Trappist-ILP}, the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and \texttt{ASP} (i.e., \texttt{Trappist-ASP}, the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
@@
 We can see that \texttt{Trappist-MaxSAT} and \texttt{Trappist-ASP} are comparable in most models, but \texttt{Trappist-ASP} is much faster for the jaoude\_thdiff and selvaggio\_emt models where the number of minimal trap spaces is greater than 1000.
-The latter can be explained by the fact that \texttt{Trappist-MaxSAT} follows an iterative approach, i.e., it restarts the search each time a solution is found (see Subsection~\ref{subsec:computation_csp}).
-The latter can be explained by the fact that \texttt{Trappist-MaxSAT} follows an iterative approach, i.e., it restarts the search each time a solution is found (see Subsection~\ref{subsec:computation_csp}).
+The latter can be explained by the fact that \texttt{Trappist-MaxSAT} follows an iterative approach, i.e., it restarts the search with a new constraint each time a solution is found (see Subsection~\ref{subsec:computation_csp}).
 This iterative approach may be less efficient than the way ASP solvers use to enumerate multiple solutions (answer sets), which is an advantage of ASP solvers~\cite{DBLP:journals/aicom/GebserKKOSS11}.
 Hence, when the number of solutions increases, the inferiority of \texttt{Trappist-MaxSAT} compared to \texttt{Trappist-ASP} will be exhibited more clearly.
-The two remaining variants, \texttt{Trappist-CP} and \texttt{Trappist-ILP}, are much less efficient than \texttt{Trappist-MaxSAT} and \texttt{Trappist-ASP} in every model, even are \(2000 \times\) slower in some models.
+The two remaining variants, \texttt{Trappist-CP} and \texttt{Trappist-ILP}, are much less efficient than \texttt{Trappist-MaxSAT} and \texttt{Trappist-ASP} in every model, even are more than three orders of magnitude slower in some models.
 The first reason for their bad performance is that they are also iterative methods like \texttt{Trappist-MaxSAT}, thus they are not efficient for "enumeration" problems.
-Upon closer inspection, for the Boolean CSP characterizing conflict-free siphons, CP seems to be something that is a "less-efficient-SAT".
+Upon closer inspection, for the Boolean CSP characterizing conflict-free siphons, CP seems to be something that is a "less-efficient-SAT", handling mostly Boolean constraints and making little use of the global constraints only added for the iterative part.
 For ILP, it may be even worse, since the problem is purely Boolean (no
@@
 However, on every model that was a bit challenging for \texttt{PyBoolNet} or \texttt{mpbn}, \texttt{Trappist-ASP} is far more efficient with speedups between \(3\times\) and \(5 \times\) for the case of \texttt{mpbn}, and between \(5\times\) and \(1783 \times\) for the case of \texttt{PyBoolNet}.
-In particular, the second best variant of \texttt{Trappist} (i.e., \texttt{Trappist-MaxSAT}) is even far more efficient than \texttt{bioLQM} and \texttt{PyBoolNet}, is comparable to \texttt{mpbn} on every model.
+In particular, the second best variant of \texttt{Trappist} (i.e., \texttt{Trappist-MaxSAT}) is even far more efficient than \texttt{bioLQM} and \texttt{PyBoolNet}, and is comparable to \texttt{mpbn} on every model.
 It is worth noting that for 3 of the 29 models, \texttt{mpbn} did not give any answer because these models are locally-monotonic but all the other methods did, which confirms the limit of \texttt{mpbn} on the applicable class of models.
@@
 	\end{tikzpicture}
-	\caption{Cumulative numbers of the \texttt{BBM} models that have less than 1000 minimal trap spaces (above) and \texttt{BBM} models (below) solved by the compared methods with respect to enumerating the first 1000 minimal trap spaces.}
+	\caption{Cumulative numbers of the \texttt{BBM} models that have less than 1000 minimal trap spaces (upper panel) and \texttt{BBM} models solved by the compared methods with respect to enumerating the first 1000 minimal trap spaces (lower panel).}
 	\label{fig:BBM_cumul}
@@
 
-Figure~\ref{fig:BBM_cumul} (above) shows cumulative numbers of the \texttt{BBM} models that have less than 1000 minimal trap spaces solved by the compared methods with respect to enumerating the first 1000 minimal trap spaces.
+Figure~\ref{fig:BBM_cumul} (upper panel) shows cumulative numbers of the \texttt{BBM} models that have less than 1000 minimal trap spaces solved by the compared methods with respect to enumerating the first 1000 minimal trap spaces.
 The number of such models is 134 (per all 211 models), and 15 of them are non-locally-monotonic.
@@
 
-Figure~\ref{fig:BBM_cumul} (below) shows cumulative numbers of the \texttt{BBM} models solved by the compared methods (except \texttt{bioLQM}, \texttt{Trappist-CP}, and \texttt{Trappist-ILP}) with respect to enumerating the first 1000 minimal trap spaces.
+Figure~\ref{fig:BBM_cumul} (lower panel) shows cumulative numbers of the \texttt{BBM} models solved by the compared methods (except \texttt{bioLQM}, \texttt{Trappist-CP}, and \texttt{Trappist-ILP}) with respect to enumerating the first 1000 minimal trap spaces.
 We omit the results of \texttt{Trappist-CP} and \texttt{Trappist-ILP} because they can handle no model with more than 1000 minimal trap spaces.
-Again, we can see that \texttt{Trappist-ASP} and \texttt{Trappist-MaxSAT} are the two best methods as they can handle every but one model within 5s as well as they always can handle much more models than both \texttt{PyBoolNet} and \texttt{mpbn} on every time limit.
-Note that with the time limit of 0.5s, \texttt{Trappist-ASP} can handle more 14 models than \texttt{Trappist-MaxSAT}, which is opposed to the case of models with less than 1000 minimal trap spaces (see Figure~\ref{fig:BBM_cumul} (above)).
+Again, we can see that \texttt{Trappist-ASP} and \texttt{Trappist-MaxSAT} are the two best methods as they can handle every but one model within 5s.
+They also always handle many more models than both \texttt{PyBoolNet} and \texttt{mpbn} on every time limit.
+Note that with the time limit of 0.5s, \texttt{Trappist-ASP} can handle 14 more models than \texttt{Trappist-MaxSAT}, which is opposed to the case of models with less than 1000 minimal trap spaces (see Figure~\ref{fig:BBM_cumul} (upper panel)).
 This observation confirms the disadvantage of \texttt{Trappist-MaxSAT} compared to \texttt{Trappist-ASP} for the case of many minimal trap spaces.
@@
 First, we obtained some observations on the four variants of \texttt{Trappist} consistent with the observations obtained in the previous subsections.
-More specifically, \texttt{Trappist-ASP} is still the best variant with the running time is always less than one second for every model, and following by \texttt{Trappist-MaxSAT}.
+More specifically, \texttt{Trappist-ASP} is still the best variant with a running time below one second for every model, and followed by \texttt{Trappist-MaxSAT}.
 In particular, the difference in running time between \texttt{Trappist-ASP} and \texttt{Trappist-MaxSAT} is bigger for larger models or models with more than 1000 minimal trap spaces.
-\texttt{Trappist-CP} and \texttt{Trappist-ILP} still have bad performance, with \texttt{Trappist-CP} is a better than \texttt{Trappist-ILP}.
+\texttt{Trappist-CP} and \texttt{Trappist-ILP} still have a much worse performance, with \texttt{Trappist-CP} better than \texttt{Trappist-ILP}.
 They still can handle no model with more than 1000 minimal trap spaces.
@@
 Third, for 11 of the 32 models (more than 34\%), \texttt{mpbn} did not give any answer because these models as non-locally-monotonic.
-For 21 of the 32 models where \texttt{mpbn} returned the answers, \texttt{mpbn} and \texttt{Trappist-ASP} are comparable in computation time, with \texttt{mpbn} appears quite slower on average.
+For 21 of the 32 models where \texttt{mpbn} returned the answers, \texttt{mpbn} and \texttt{Trappist-ASP} are roughly comparable in computation time, but \texttt{mpbn} appears quite slower on average.
 In particular, for the Rho-GTPases model, \texttt{mpbn} is \(577\times\) slower than \texttt{Trappist-ASP}.
 This observation along with the comparisons between \texttt{mpbn} and \texttt{Trappist-ASP} in the previous subsections are quite surprising because the ASP encoding of \texttt{mpbn} only requires the DNF for the activation part of a Boolean function, whereas that of \texttt{Trappist-ASP} requires both the activation and inhibition parts (see Subsection~\ref{subsec:computation_asp}).
-However, the reason maybe lies on the differences in the ASP encoding characteristics of the two methods and the fact that \texttt{mpbn} needs to spend time for checking the locally-monotonicity of each Boolean function in a Boolean network.
-It is possible that \texttt{mpbn} may outperform \texttt{Trappist} for a certain set of models, but not for the set of real-world models considered in this article.
+However, the reason may lie on the differences in the ASP encoding characteristics of the two methods and the fact that \texttt{mpbn} needs to spend time checking the locally-monotonicity of each Boolean function in a Boolean network.
+We expect that \texttt{mpbn} may outperform \texttt{Trappist} for a certain set of models, but not for the set of real-world models considered in this article.
 
@@
 
-We randomly generated a set of N-K models~\cite{glass1973logical} with network size \(n\) in the set \{100, 150, 200, 250, 300, 350, 400\} and \(K = 3\) (i.e., each node has exactly three input nodes).
+We randomly generated a set of N-K models~\cite{glass1973logical} with network size \(n\) in the set \{100, 150, 200, 250, 300, 350, 400\} and in-degree \(K = 3\) (i.e., each node has exactly three input nodes).
 We chose N-K models because they are a useful tool for studying the dynamics of Boolean networks~\cite{glass1973logical,klarner2015computing,Rozum2021}.
@@
 The number of succeeded models within three minutes for each method is: \texttt{bioLQM} (0), \texttt{PyBoolNet} (320), \texttt{mpbn} (0), \texttt{Trappist-maxSAT} (338), \texttt{Trappist-CP} (226), \texttt{Trappist-ILP} (39), \texttt{Trappist-ASP} (349).
-We can see that \texttt{Trappist-ASP} is the only method that can handle every but one model.
-Note that none of the other methods can handle the only model failed by \texttt{Trappist-ASP}.
+We can see that \texttt{Trappist-ASP} is the only method that can handle every model, but one.
+Note that none of the other methods can handle that only model failed by \texttt{Trappist-ASP}.
 We also obtained some observations consistent with those obtained for real-world models.
@@
 
-First, \texttt{mpbn} did not be able to handle any model because all the models are non-locally-monotonic.
+First, \texttt{mpbn} was not able to handle any model because all the models are non-locally-monotonic.
 Recall that a Boolean network is non-locally-monotonic if only one of its Boolean functions is non-locally-monotonic.
-Hence, it is apparent that all the randomly generated models are non-locally-monotonic because of the number of nodes is large (\(n \geq 100\)).
-This observation confirms the limit on the applicable model class of \texttt{mpbn}.
+Hence, it is apparent that all this type of randomly generated models are non-locally-monotonic because of the number of nodes is large (\(n \geq 100\)).
+This observation confirms a limit on the applicable model class of \texttt{mpbn}.
 
@@
 
-Third, for every time limit, \texttt{Trappist-ASP} can always handle much models than \texttt{PyBoolNet}, ranging from 29 to 65 more models.
+Third, for every time limit, \texttt{Trappist-ASP} can always handle many more models than \texttt{PyBoolNet}, ranging from 29 to 65 more models.
 Since the time for the phase of computing prime-implicants of \texttt{PyBoolNet} is negligible in every model, most of the running time of \texttt{PyBoolNet} was spent for its ASP solving phase.
@@
 
-We have tested our alternative approach on many Boolean network models of various sizes and types (e.g., real-world models, randomly generated models).
+We have tested our alternative approach on many Boolean network models of various sizes and types (e.g., real-world models, randomly generated models) on existing and newly created benchmarks.
 This indicates the high coverage and comprehensiveness of the experiments.
@@
 The two remaining variants (i.e., \texttt{Trappist-CP} and \texttt{Trappist-ILP}) give bad performance for most models.
-However, for some certain cases, they are even better than all the state-of-the-art methods (i.e., \texttt{bioLQM}, \texttt{PyBoolNet}, and \texttt{mpbn}).
-This is evidence for the advantages of the alternative approach compared to the state-of-the-art ones.
+However, for certain cases, they are still better than all state-of-the-art methods (i.e., \texttt{bioLQM}, \texttt{PyBoolNet}, and \texttt{mpbn}).
+This is evidence for the advantages of an alternative approach compared to what preexisted.
 
@@
 The speedups of \texttt{Trappist-ASP} or \texttt{Trappist-MaxSAT} are large, even between one and three orders of magnitude for most models.
-In addition, the experimental results also confirm that the ASP encoding of \texttt{Trappist-ASP} is much better than that of \texttt{PyBoolNet}.
-
-Regarding locally-monotonic Boolean networks, the performance of \texttt{mpbn} is comparable to that of \texttt{Trappist-ASP} or \texttt{Trappist-MaxSAT}.
-However, \texttt{mpbn} is quite slower than \texttt{Trappist-ASP} in average.
+In addition, the experimental results also confirm that the ASP encoding of \texttt{Trappist-ASP} is much more efficient than that of \texttt{PyBoolNet}.
+
+Regarding locally-monotonic Boolean networks, the performance of \texttt{mpbn} is roughly comparable to that of \texttt{Trappist-ASP} or \texttt{Trappist-MaxSAT}.
+However, \texttt{mpbn} is quite slower than \texttt{Trappist-ASP} on average.
 This shows the practical advantage of \texttt{Trappist-ASP} compared to \texttt{mpbn}, though its ASP encoding may be more complex than that of \texttt{mpbn} in theory.
@@
 The proposed methods have been evaluated on many real-world models from the literature as well as randomly generated models.
-The experimental results show that the new approach vastly outperforms all the state-of-the-art methods in terms of general Boolean networks and is comparable to the \texttt{mpbn} method even much better in average in terms of locally-monotonic Boolean networks.
+The experimental results show that the new approach vastly outperforms all the state-of-the-art methods in terms of general Boolean networks and is comparable to the \texttt{mpbn} method even much better on average in terms of locally-monotonic Boolean networks.
 We believe that this opens up the way to a much better analysis of large Boolean networks, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
@@
 
+\sylvain{Do you want to say a word about specific types of trap-spaces? This currently does not really appear in the conclusion.}
+
 %% The Appendices part is started with the command \appendix;

### 2023-03-17 manuscripts/TCS2022/tcs2022.tex
response

@@
 Then the dynamics of \(\mathcal{N}\) is captured by the directed graph \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\) called the State Transition Graph (STG).
-There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: synchronous, where all the nodes are update simultaneously, and fully asynchronous, where only one node is selected to be updated.
+There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: synchronous, where all the nodes are update simultaneously, and fully asynchronous, where only one node is selected in random to be updated.
 
@@
 The proof directly on trap-spaces is quite obvious and follows the same structure: the intersection would be a trap-space, and smaller than the original ones, hence a contradiction\ldots}
+\vangiang{I agree with you. I have added an argument for this.}
 Specifically, all minimal trap spaces of a Boolean network are mutually disjoint.
 This property is important because we can use it to approximate the set of attractors of the Boolean network under any update scheme~\cite{klarner2015computing} or to compute exactly the set of complex attractors of the Boolean network under the fully asynchronous update scheme~\cite{DBLP:conf/bcb/TrinhHB22}.
+Note that it would be not difficult to obtain a direct proof on trap spaces for this property, which follows the same structure as the proof on siphons.
+However, we emphasize here the potential of using the connection between Boolean networks and Petri nets to explore and prove properties of trap spaces in Boolean networks.
 
@@
 
-Using the ASP-based method presented in Section~\ref{subsec:computation_asp}, it is possible to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) within 0.19s, which is much quicker than \texttt{mpbn}.
-Unfortunately since this was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.
+For the ``RA-apoptosis'' model, using the ASP-based method presented in Section~\ref{subsec:computation_asp}, it is possible to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) within 0.19s, which is much quicker than \texttt{mpbn}.
+The needed time for the ``Alzheimer'' model is 0.79s.
+Unfortunately since this method was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.
 
 \sylvain{I think that the result above (0.19s) is only about the F5 model, but do we also have the time for the F4 model, if so, it should be nice to include it too.}
+\vangiang{I agree. Simply included it.}
 
@@
 \sylvain{should we also highlight in some way the (tied-)best result?}
+\vangiang{I think we should not because such highlights seem not much support the experimental analysis, but may make the table a bit awkward.}
 There are four variants of \texttt{Trappist}: \texttt{SAT} (i.e., \texttt{Trappist-MaxSAT}, the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{CP} (i.e., \texttt{Trappist-CP}, the CP-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{ILP} (i.e., \texttt{Trappist-ILP}, the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and \texttt{ASP} (i.e., \texttt{Trappist-ASP}, the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
@@
 In this article we have explored and proved for the first time the equivalence between (minimal) trap spaces of a general Boolean network and (maximal) conflict-free siphons of its Petri net encoding.
-We have shown several important applications of this finding to studying properties of trap spaces in Boolean networks.
+We have shown several useful applications of this finding to studying properties of trap spaces in Boolean networks.
 As an important practical application of the equivalence, we have proposed a new approach for the computation of minimal trap spaces in Boolean networks, based on the enumeration of maximal conflict-free siphons of Petri nets.
 We have also proposed the four possible methods using MaxSAT, CP, ILP, and ASP for implementing the new approach.
-The proposed methods have been evaluated on many real-world models from the literature as well as randomly generated models.
+In particular, we have shown how to adjust our approach to compute several specific types of trap spaces (e.g., maximal trap spaces, fixed points), which besides minimal trap spaces also play crucial rules in the analysis and control of Boolean networks.
+The proposed methods for the minimal trap space computation have been evaluated on many real-world models from the literature as well as randomly generated models.
 The experimental results show that the new approach vastly outperforms all the state-of-the-art methods in terms of general Boolean networks and is comparable to the \texttt{mpbn} method even much better on average in terms of locally-monotonic Boolean networks.
@@
 \sylvain{Do you want to say a word about specific types of trap-spaces? This currently does not really appear in the conclusion.}
+\vangiang{Thank you for your reminder. I have mentioned the computation of specific types of trap spaces in the conclusion.}
 

### 2023-03-20 manuscripts/TCS2022/tcs2022.tex
Mostly agreed

@@
 Then the dynamics of \(\mathcal{N}\) is captured by the directed graph \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\) called the State Transition Graph (STG).
-There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: synchronous, where all the nodes are update simultaneously, and fully asynchronous, where only one node is selected in random to be updated.
+There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: synchronous, where all the nodes are update simultaneously, and fully asynchronous, where only one node is selected non-deterministically to be updated.
+\sylvain{I removed the ``random'' part since actually, how the only node is selected is not strictly part of the asynchronous update\ldots Updating always the nodes in the order in which they appear in the model for instance, though pretty bad, would be an admissible async update run.
+Usually for such things where you could even imagine an ``opponent'' chosing which node to update, you can say it's ``non-deterministic'' but not necessarily ``random''.}
 
@@
 We use it to prove a property of minimal trap spaces, which has surprisingly not been formally proved.
-\sylvain{I'm not sure how much we should insist on the ``surprise''.
-The proof directly on trap-spaces is quite obvious and follows the same structure: the intersection would be a trap-space, and smaller than the original ones, hence a contradiction\ldots}
-\vangiang{I agree with you. I have added an argument for this.}
 Specifically, all minimal trap spaces of a Boolean network are mutually disjoint.
@@
 Unfortunately since this method was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.
-
-\sylvain{I think that the result above (0.19s) is only about the F5 model, but do we also have the time for the F4 model, if so, it should be nice to include it too.}
-\vangiang{I agree. Simply included it.}
 
@@
 ``NM'' indicates a non-locally-monotonic model.
-\sylvain{should we also highlight in some way the (tied-)best result?}
-\vangiang{I think we should not because such highlights seem not much support the experimental analysis, but may make the table a bit awkward.}
 There are four variants of \texttt{Trappist}: \texttt{SAT} (i.e., \texttt{Trappist-MaxSAT}, the MaxSAT-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{CP} (i.e., \texttt{Trappist-CP}, the CP-based method shown in Subsection~\ref{subsec:computation_csp}), \texttt{ILP} (i.e., \texttt{Trappist-ILP}, the ILP-based method shown in Subsection~\ref{subsec:computation_ilp}), and \texttt{ASP} (i.e., \texttt{Trappist-ASP}, the ASP-based method shown in Subsection~\ref{subsec:computation_asp}).
@@
 We have also proposed the four possible methods using MaxSAT, CP, ILP, and ASP for implementing the new approach.
-In particular, we have shown how to adjust our approach to compute several specific types of trap spaces (e.g., maximal trap spaces, fixed points), which besides minimal trap spaces also play crucial rules in the analysis and control of Boolean networks.
+In particular, we have shown how to adjust our approach to compute several specific types of trap spaces (e.g., maximal trap spaces, fixed points), which besides minimal trap spaces also play crucial roles in the analysis and control of Boolean networks.
 The proposed methods for the minimal trap space computation have been evaluated on many real-world models from the literature as well as randomly generated models.
@@
 
-\sylvain{Do you want to say a word about specific types of trap-spaces? This currently does not really appear in the conclusion.}
-\vangiang{Thank you for your reminder. I have mentioned the computation of specific types of trap spaces in the conclusion.}
-
 %% The Appendices part is started with the command \appendix;

### 2023-03-20 manuscripts/TCS2022/tcs2022.tex
First complete draft

@@
 There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: synchronous, where all the nodes are update simultaneously, and fully asynchronous, where only one node is selected non-deterministically to be updated.
-\sylvain{I removed the ``random'' part since actually, how the only node is selected is not strictly part of the asynchronous update\ldots Updating always the nodes in the order in which they appear in the model for instance, though pretty bad, would be an admissible async update run.
-Usually for such things where you could even imagine an ``opponent'' chosing which node to update, you can say it's ``non-deterministic'' but not necessarily ``random''.}
 

### 2023-04-03 manuscripts/TCS2022/tcs2022.tex
final check

@@
 From the observation that the transcriptional regulation behaved in a sigmoid step-like way, came the original idea to represent models of gene regulation as discrete event systems.
-Those Gene Regulation Networks (GRN) use thresholds or equivalently logical functions to represent the different regulations~\cite{glass1973logical,thomas1973boolean,thomas1990biological,thomas1991regulatory}.
-
-Boolean net modeling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}, even for modelling post-transcriptional mechanisms.
+Those gene regulation networks use thresholds or equivalently logical functions to represent the different regulations~\cite{glass1973logical,thomas1973boolean,thomas1990biological,thomas1991regulatory}.
+
+Boolean network modeling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}, even for modelling post-transcriptional mechanisms.
 This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas~\cite{aghamiri2020automated}.
@@
 
-It is worth noting that the recent method presented in~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces avoids the prime-implicants computation by relying on the \emph{most-permissive} semantics of Boolean networks.
+It is worth noting that the recent method presented in~\cite{DBLP:conf/ictai/ChevalierFPZ19} for computing minimal trap spaces avoids the prime-implicant computation by relying on the \emph{most-permissive} semantics of Boolean networks.
 This method has been implemented in the tool \texttt{mpbn}\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
@@
 In addition, the \texttt{bioLQM} platform also provides another method using Binary Decision Diagrams (BDDs) in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html}.
-This method avoids the prime-implicants computation as it characterizes the set of generic trap spaces of a Boolean network by a BDD, then filters this set to get the set of all minimal trap spaces.
-By this approach, it requires the computation of all solutions, whereas the ASP-based methods~\cite{klarner2015computing,Paulev2020} can start enumerating them as they are found.
+This method avoids the prime-implicant computation as it characterizes the set of generic trap spaces of a Boolean network by a BDD, then filters this set to get the set of all minimal trap spaces.
+By this approach, it requires the computation of all solutions, whereas the methods~\cite{klarner2015computing,Paulev2020} based on Answer Set Programming (ASP) can start enumerating them as they are found.
 Moreover, the main issue with the BDD-based method is that the number of generic trap spaces of a Boolean network may be extremely larger than its number of minimal trap spaces.
@@
     &f_i(v(t))\\
-    \text{or} &v(t)
+    \text{or} &v_i(t)
   \end{cases}
 \]
-where \(v(t)\) is the state of \(\mathcal{N}\) at time \(t\). % and \(v_i(t + 1)\) is the state of node \(v_i\) at time \(t + 1\).
+where \(v(t)\) (resp. \(v_i(t)\)) is the state of \(\mathcal{N}\) (resp. the state of node \(v_i\)) at time \(t\). % and \(v_i(t + 1)\) is the state of node \(v_i\) at time \(t + 1\).
 Note that for simplicity, we write \(f_i(v(t))\) even if \(\IN(v_i) \subsetneq V\) (i.e., \(\IN(v_i)\) does not contain some nodes of \(V\)).
@@
 Then the dynamics of \(\mathcal{N}\) is captured by the directed graph \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\) called the State Transition Graph (STG).
-There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: synchronous, where all the nodes are update simultaneously, and fully asynchronous, where only one node is selected non-deterministically to be updated.
+There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: \emph{synchronous}, where all the nodes are update simultaneously, and \emph{fully asynchronous}, where only one node is selected non-deterministically to be updated.
 
@@
 A subspace \(m\) of a Boolean network \(\mathcal{N} = (V, F)\) is a mapping \(m \colon V \mapsto \mathbb{B} \cup \{\star\}\).
-\(m(v_i) \in \mathbb{B}\) means that the value of \(v_i\) is fixed in \(m\) and \(v_i\) is called a fixed variable.
-\(m(v_i) \in \star\) means that the value of \(v_i\) is free in \(m\) and \(v_i\) is called a free variable.
+\(m(v_i) \in \mathbb{B}\) means that the value of \(v_i\) is fixed in \(m\) and \(v_i\) is called a \emph{fixed} variable.
+\(m(v_i) \in \star\) means that the value of \(v_i\) is free in \(m\) and \(v_i\) is called a \emph{free} variable.
 We denote \(D_m\) the set of all fixed variables of \(m\).
@@
 Figure~\ref{fig:stg} shows the dynamics of this model under the fully asynchronous update (i.e., only one node is updated at each time step).
-The model has all trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
+The model has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
 Since \(m_1 < m_2\), \(m_1\) is the only minimal trap space of the Boolean network.
@@
 This does not change the worst-case-complexity (obtaining a single DNF being exponential) but might matter a lot in practice.
-As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM} or directly in our tool using the \texttt{pyeda}\footnote{\url{https://pyeda.readthedocs.io/en/latest/}} library, and the one based on the most-permissive semantics compare in the Section~\ref{sec:eval} on evaluation.
+As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM} or directly in our tool using the \texttt{pyeda}\footnote{\url{https://pyeda.readthedocs.io/en/latest/}} library, and the one based on the most-permissive semantics compare in Section~\ref{sec:eval} on evaluation.
 
@@
 Siphons are a static and classical property of Petri nets~\cite{peterson1981petri}.
-Note however that the use of siphons for the analysis of biological models, though it is not new, has been mostly relevant to the ODE-based continuous semantics of Chemical Reaction Networks~\cite{angeli2007petri,angeli2011persistence,degrand2020graphical}.
+Note however that the use of siphons for the analysis of biological models, though it is not new, has been mostly relevant to the ODE-based continuous semantics of chemical reaction networks~\cite{angeli2007petri,angeli2011persistence,degrand2020graphical}.
 We recall here the basic definition establishing that to produce something in a siphon you must consume something from the siphon.
@@
 
-\section{Minimal trap spaces as maximal conflict-free siphons}%
+\section{Trap spaces as conflict-free siphons}%
 \label{sec:Main_finding}
@@
 
-Note that there are no existing methods specifically designed for computing maximal conflict-free siphons (even maximal siphons) of a Petri net.
+Note that there are no existing methods specifically designed for computing maximal conflict-free siphons (even maximal generic siphons) of a Petri net.
 The reason might be that researchers mainly focus on minimal generic siphons~\cite{DBLP:journals/isci/LiuB16} in the field of Petri nets.
@@
 We set a soft clause for each variable of the CSP and then use a ``minimal correction subset'' blocking strategy, which will ensure set-inclusion maximality of the solutions.
-This is what is implemented in \texttt{Trappist} using the \texttt{RC2} MaxSAT solver~\cite{DBLP:journals/jsat/IgnatievMM19} available through the \texttt{python-sat} package\footnote{\url{https://pysathq.github.io/docs/html/api/examples/rc2.html}}.
+We implement this approach by using the \texttt{RC2} MaxSAT solver~\cite{DBLP:journals/jsat/IgnatievMM19} available through the \texttt{python-sat} package\footnote{\url{https://pysathq.github.io/docs/html/api/examples/rc2.html}}.
 
@@
 \]
-where \verb|a| is the binary variable representing place \(p\) and \{\verb|a_1|, \dots,\verb|a_k|\} is the set of binary variable representing places in \(\pred(t)\).
+where \verb|a| is the binary variable representing place \(p\) and \{\verb|a_1|, \dots,\verb|a_k|\} is the set of binary variables representing places in \(\pred(t)\).
 The rule (\ref{eq:conflict}) is translated into the ILP inequality
@@
 
-Since we used the MiniZinc framework to interface with the CP solver, it was simple to make the slight modifications described above and use that same interface to call the Coin-OR CBC solver\footnote{\url{https://github.com/coin-or/Cbc}}~\cite{john_forrest_2022_6522795}.
+Since we used the MiniZinc framework to interface with the CP solver, it was simple to make the slight modifications described above and to use that same interface to call the Coin-OR CBC solver\footnote{\url{https://github.com/coin-or/Cbc}}~\cite{john_forrest_2022_6522795}.
 
@@
 One of the main maps studied this way represents knowledge about the Rheumatoïd Arthritis~\cite{singh2018computational}, and was the main motivation for the development of a tool to automatically transform it into an executable Boolean network~\cite{aghamiri2020automated}.
-In the supplementary material of the paper, an excerpt of the map, focused around the apoptosis (cell death) module is transformed into a model of \emph{reasonable} size, namely 180 Boolean variables (model \verb|F5_RA_apoptosis_executable_module.sbml| of supplementary material S3, and model ``RA-apoptosis'' of Section~\ref{sec:eval}).
+In the supplementary material of the paper, an excerpt of the map, focused around the apoptosis (cell death) module is transformed into a model of \emph{reasonable} size, namely 180 Boolean variables (model \verb|F5_RA_apoptosis_executable_module.sbml| of supplementary material S3, and model ``RA\_apoptosis'' of Subsection~\ref{subsec:Selected_models}).
 The study of such model, though, is a big hurdle.
@@
 However, using a state-of-the-art tool like \texttt{PyBoolNet}~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
-\texttt{mpbn}~\cite{Paulev2020} can return the first 1000 solution within 1.43s, but indeed, it limits the modeling range of the modelers as it does not permit using non-locally-monotonic Boolean functions.
+\texttt{mpbn}~\cite{Paulev2020} can return the first 1000 solutions within 1.43s, but indeed, it limits the modeling range of the modelers as it does not permit using non-locally-monotonic Boolean functions.
 This is also true for the Alzheimer model also mentioned in that same article and originally from~\cite{ogishima2016alzpathway} (\verb|F4| file in the original supplementary material, and ``Alzheimer'' in Table~\ref{tab:result_real}), where \texttt{PyBoolNet} also fails at the prime-implicant computation and \texttt{mpbn} does not give any answer because this model is actually non-locally-monotonic.
@@
 
-For the ``RA-apoptosis'' model, using the ASP-based method presented in Section~\ref{subsec:computation_asp}, it is possible to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) within 0.19s, which is much quicker than \texttt{mpbn}.
+For the ``RA\_apoptosis'' model, using the ASP-based method presented in Subsection~\ref{subsec:computation_asp}, it is possible to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) within 0.19s, which is much quicker than \texttt{mpbn}.
 The needed time for the ``Alzheimer'' model is 0.79s.
@@
 \subsection{Selected models}
+\label{subsec:Selected_models}
 
@@
 
-Third, for 11 of the 32 models (more than 34\%), \texttt{mpbn} did not give any answer because these models as non-locally-monotonic.
+Third, for 11 of the 32 models (more than 34\%), \texttt{mpbn} did not give any answer because these models are non-locally-monotonic.
 For 21 of the 32 models where \texttt{mpbn} returned the answers, \texttt{mpbn} and \texttt{Trappist-ASP} are roughly comparable in computation time, but \texttt{mpbn} appears quite slower on average.
@@
 This observation along with the comparisons between \texttt{mpbn} and \texttt{Trappist-ASP} in the previous subsections are quite surprising because the ASP encoding of \texttt{mpbn} only requires the DNF for the activation part of a Boolean function, whereas that of \texttt{Trappist-ASP} requires both the activation and inhibition parts (see Subsection~\ref{subsec:computation_asp}).
-However, the reason may lie on the differences in the ASP encoding characteristics of the two methods and the fact that \texttt{mpbn} needs to spend time checking the locally-monotonicity of each Boolean function in a Boolean network.
+However, the reason may lie on the differences in the ASP encoding characteristics of the two methods and the fact that \texttt{mpbn} needs to spend time checking the local-monotonicity of each Boolean function in a Boolean network.
 We expect that \texttt{mpbn} may outperform \texttt{Trappist} for a certain set of models, but not for the set of real-world models considered in this article.
@@
 Hence, we searched for all solutions in each model, which makes the comparison to \texttt{bioLQM} more comprehensive.
-In addition, each node has only three input nodes, i.e., the number of prime-implicants of the associated Boolean function is small.
+In addition, each node has only three input nodes, leading to the number of prime-implicants of the associated Boolean function is small.
 Hence, \texttt{PyBoolNet} always passed the phase of computing prime-implicants in every model even within one second, which enables us to compare the ASP encoding of \texttt{PyBoolNet} and that of \texttt{Trappist-ASP}.
@@
 
-Finally, we think that the links between Petri nets and Boolean networks that we stumbled upon in this method might have deeper roots.
+Finally, we think that the links between Petri nets and Boolean networks that we stumbled upon in this article might have deeper roots.
 Exploring those connections might lead both to interesting topics of research for Petri nets, like a notion of trap-spaces, and for Boolean networks.

### 2023-04-04 manuscripts/TCS2022/tcs2023.tex
rename to 2023

--- 2023-04-03 manuscripts/TCS2022/tcs2022.tex
@@
 First, more formal definitions are given and the existing proofs are made more detailed.
-In particular, an updated proof provides another way to prove the independence of trap spaces of a Boolean network on its update scheme, which was originally proved in~\cite{klarner2015computing}.
+In particular, an updated proof provides another way to prove the independence of trap spaces of a Boolean network with respect to its update scheme, which was originally proved in~\cite{klarner2015computing}.
 Second, we showcase a theoretical application of the connection between trap spaces in Boolean networks and conflict-free siphons in Petri nets.
@@
 Fifth, regarding the implementation, we have developed a new converter that directly reads a \texttt{.bnet} file and builds the Petri net encoding, instead of using the \texttt{PNML} conversion of \texttt{bioLQM}~\cite{DBLP:conf/cmsb/TrinhBHS22}.
-Finally, we conduct a more extensive benchmark on more real-world models from various sources and randomly generated models to evaluate all the proposed methods (the benchmark conducted in~\cite{DBLP:conf/cmsb/TrinhBHS22} considers only dozens of representative real-world models), with more comprehensive insights are obtained.
+Finally, we conduct a more extensive benchmark on more real-world models from various sources and randomly generated models to evaluate all the proposed methods (the benchmark conducted in~\cite{DBLP:conf/cmsb/TrinhBHS22} considers only a few dozens of representative real-world models), therefore obtaining more comprehensive insights.
 
@@
 Note that the Petri net construction is less computationally demanding than the prime-implicant computation because it only requires computing generic (not prime) implicants of the Boolean functions~\cite{chatain2014characterization}.
-In addition, the time complexity of the above checking method is quadratic in the number of transitions of the Petri net in worst cases.
+In addition, the worst case time complexity of the above checking method is quadratic in the number of transitions of the Petri net.
 
@@
 As an important practical application of the equivalence, we have proposed a new approach for the computation of minimal trap spaces in Boolean networks, based on the enumeration of maximal conflict-free siphons of Petri nets.
-We have also proposed the four possible methods using MaxSAT, CP, ILP, and ASP for implementing the new approach.
+We have also proposed four possible methods using MaxSAT, CP, ILP, and ASP for implementing the new approach.
 In particular, we have shown how to adjust our approach to compute several specific types of trap spaces (e.g., maximal trap spaces, fixed points), which besides minimal trap spaces also play crucial roles in the analysis and control of Boolean networks.

### 2023-04-11 manuscripts/TCS2023/tcs2023.tex
Intro

--- 2023-04-04 manuscripts/TCS2022/tcs2023.tex
@@
 
-Boolean network modeling has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}, even for modelling post-transcriptional mechanisms.
-This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas~\cite{aghamiri2020automated}.
-Besides simulation, the analysis of such models is mostly based on attractor computation, since those correspond roughly to observable biological \emph{phenotypes}.
-The recent use of trap spaces~\cite{klarner2015computing} made a real breakthrough in that field allowing to consider medium-sized models that used to be out of reach.
-However, with the continuing increase in model size and complexity of Boolean update functions, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits.
+Boolean modeling made available some powerful analyses and corresponding insight for gene regulation models.
+Then, over the years, its use increased even for modelling post-transcriptional mechanisms, supported by the many cases where precise biological data was not sufficiently available to build a detailed quantitative model~\cite{wang2012boolean}.
+This lack of data is more frequent for large and very large models, which led to a steady increase in the size of logical models \emph{à la} Thomas~\cite{aghamiri2020automated}.
+The main analysis tool for such models is the computation of its fixed and periodic attractors, since those correspond roughly to observable biological \emph{phenotypes}.
+The recent use of trap spaces~\cite{klarner2015computing} made a real breakthrough in that field allowing to consider medium-sized models that used to be out of reach and for which only simulation was available.
+However, with the most recent models both being quite large and using rather complex update functions, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits.
 More specifically, the number of prime implicants of a Boolean function is in general exponential in the number of input nodes of this function~\cite{klarner2015computing}.
@@
 \bibliographystyle{elsarticle-num-names}
-\bibliography{tcs2022.bib}
+\bibliography{tcs2023.bib}
 

### 2023-04-11 manuscripts/TCS2023/tcs2023.tex
Some small changes

--- 2023-04-04 manuscripts/TCS2022/tcs2023.tex
@@
 \newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
+
+\newcommand{\revBelaid}[1]{\textcolor{red}{#1}}
 
@@
 We give a Boolean network \(\mathcal{N} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
-Herein, \(\land\), \(\lor\), and \(\neg\) denote the conjunction, disjunction, and negation logical operators, respectively.\label{example:BN}
+Herein, \(\land\), \(\lor\), and \(\neg\) denote the \revBelaid{logical conjunction, disjunction, and negation operators}, respectively.\label{example:BN}
 \end{example}
@@
 This does not change the worst-case-complexity (obtaining a single DNF being exponential) but might matter a lot in practice.
-As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM} or directly in our tool using the \texttt{pyeda}\footnote{\url{https://pyeda.readthedocs.io/en/latest/}} library, and the one based on the most-permissive semantics compare in Section~\ref{sec:eval} on evaluation.
+As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM} or directly in our tool using the \texttt{pyeda}\footnote{\url{https://pyeda.readthedocs.io/en/latest/}} library, and the one based on the most-permissive semantics \revBelaid{compare with each other in Section~\ref{sec:eval}}.
 
@@
 Intuitively, a siphon is a set of places that once unmarked remains so.
-If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the unmarked places remain unmarked).
+If it is conflict-free then its dual corresponds to a \revBelaid{subspace} of the model such that whatever update, the fixed values remain so (since the unmarked places remain unmarked).
 This is precisely the definition of a trap space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap space.

### 2023-04-12 manuscripts/TCS2023/tcs2023.tex
Remove highlights
Final changes

--- 2023-04-11 manuscripts/TCS2023/tcs2023.tex
@@
 We give a Boolean network \(\mathcal{N} = (V, F)\), where \(V = (x_1, x_2)\) and \(F = (f_1, f_2)\) with \(f_1 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2), f_2 = (x_1 \land x_2) \lor (\neg x_1 \land \neg x_2)\).
-Herein, \(\land\), \(\lor\), and \(\neg\) denote the conjunction, disjunction, and negation logical operators, respectively.\label{example:BN}
+Herein, \(\land\), \(\lor\), and \(\neg\) denote the logical conjunction, disjunction, and negation operators, respectively.\label{example:BN}
 \end{example}
@@
 This does not change the worst-case-complexity (obtaining a single DNF being exponential) but might matter a lot in practice.
-As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM} or directly in our tool using the \texttt{pyeda}\footnote{\url{https://pyeda.readthedocs.io/en/latest/}} library, and the one based on the most-permissive semantics compare in Section~\ref{sec:eval} on evaluation.
+As such, we will explore how this transformation, here using BDDs in \texttt{bioLQM} or directly in our tool using the \texttt{pyeda}\footnote{\url{https://pyeda.readthedocs.io/en/latest/}} library, and the one based on the most-permissive semantics compare with each other in Section~\ref{sec:eval}.
 
@@
 \label{prop:siphon_set}
-  Let \(S\) be a siphon of a Petri net \((P, T, W)\).
-  Then \(\pred(S) \subseteq \succ(S)\).
+
+  A set \(S\) of places is a siphon of a Petri net \((P, T, W)\) if and only if \(\pred(S) \subseteq \succ(S)\).
 
@@
 Intuitively, a siphon is a set of places that once unmarked remains so.
-If it is conflict-free then its dual corresponds to a partial-state of the model such that whatever update, the fixed values remain so (since the unmarked places remain unmarked).
+If it is conflict-free then its dual corresponds to a subspace of the model such that whatever update, the fixed values remain so (since the unmarked places remain unmarked).
 This is precisely the definition of a trap space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap space.
@@
   A \emph{mirror} of \(m\) is a set of places \(S\) in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{N}\) such that:
-  \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S\] and \[\forall v \in V \setminus D_m, p_v \not \in S, \overline{p}_v \not \in S.\]
+  \[\forall v \in D_m \left[m(v) = 0 \Leftrightarrow p_v \in S \land m(v) = 1 \Leftrightarrow \overline{p}_v \in S\right]\] and \[\forall v \in V \setminus D_m \left[p_v \not \in S \land \overline{p}_v \not \in S\right].\]
 

### 2023-04-13 manuscripts/TCS2023/tcs2023.tex
Fixed some overfull hboxes

@@
 This method has been implemented in the tool \texttt{mpbn}\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
-However, this method is only applicable for \emph{locally-monotonic} Boolean networks, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean networks (i.e., including both locally-monotonic and non-locally-monotonic ones).
+However, this method is only applicable for \emph{locally-monotonic} Boolean networks, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean networks (i.e., including both locally-mono\-tonic and non-locally-monotonic ones).
 In addition, the \texttt{bioLQM} platform also provides another method using Binary Decision Diagrams (BDDs) in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html}.
@@
 
-This method was implemented using the state-of-the-art CP solver Chuffed\footnote{\url{https://github.com/chuffed/chuffed}}~\cite{demirovic2018solution} via its MiniZinc~\cite{nethercote2007minizinc} interface.
+This method was implemented using the state-of-the-art CP solver Chuffed\footnote{\url{https://github.com/chuffed/chuffed}} \cite{demirovic2018solution} via its MiniZinc~\cite{nethercote2007minizinc} interface.
 Because it is a high-level interface, the backtrack-and-replay method of~\cite{nabli2016enumerating} was not used but rather the alternative implementation with two global constraints for lexicographic ordering (ensuring enumeration of solutions) and iterated non-subset of each already found solution (for maximality).
@@
 
-Fourth, regarding the comparison of the ASP-based methods (i.e., \texttt{PyBoolNet}, \texttt{mpbn}, and \texttt{Trappist-ASP}), we note that for all the models where \texttt{PyBoolNet} did not finish before the time limit, the timeout occurred during the computation of the prime-implicants.
+Fourth, regarding the comparison of the ASP-based methods (i.e.,\linebreak \texttt{PyBoolNet}, \texttt{mpbn}, and \texttt{Trappist-ASP}), we note that for all the models where \texttt{PyBoolNet} did not finish before the time limit, the timeout occurred during the computation of the prime-implicants.
 Hence, not even a single minimal trap space was output by that method.
@@
 
-Regarding general Boolean networks, \texttt{Trappist-ASP} (even \texttt{Trappist-MaxSAT}) is far more efficient than both \texttt{bioLQM} and \texttt{PyBoolNet}.
+Regarding general Boolean networks, \texttt{Trappist-ASP} (even \texttt{Trappist\-MaxSAT}) is far more efficient than both \texttt{bioLQM} and \texttt{PyBoolNet}.
 The speedups of \texttt{Trappist-ASP} or \texttt{Trappist-MaxSAT} are large, even between one and three orders of magnitude for most models.

### 2023-06-06 manuscripts/TCS2023/response.txt
Cleanup before addressing reviewer comments

We would first like to thank the reviewers for their insightful comments.

* Reviewer 1

"In my opinion, this submission is a solid paper, quite well written, and should
become a reference for anyone doing dynamics analysis of Boolean models. Below I
list several minor remarks, fixes, and suggestions for improvement. I believe
this work can be published after these minor corrections are implemented."

We fixed all the minor remarks highlighted by the reviewer, as can be seen in
red in the uploaded version.  Indeed it improved substantially the readability
of the article.

### 2023-06-06 manuscripts/TCS2023/tcs2023.tex
Cleanup before addressing reviewer comments

@@
 
-%% Use the option review to obtain double line spacing
-%% \documentclass[preprint,review,12pt]{elsarticle}
-
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn]{elsarticle}
-%% \documentclass[final,3p,times]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn]{elsarticle}
-%% \documentclass[final,5p,times]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn]{elsarticle}
-
-%% The amssymb package provides various useful mathematical symbols
 \usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
 \usepackage{amsthm}
@@
 
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
 \usepackage{lineno}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Drawing
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \usepackage{tikz}
@@
 
-% drawing automa
 \usetikzlibrary{positioning,automata}
 
-% for captions in figures
 \usepackage[labelformat=simple]{subcaption}
@@
 \usepackage{booktabs}
-
-% table row numbers
 \newcounter{rownumber}
@@
 
-\newcommand{\vangiang}[1]{\textcolor{magenta}{#1}}
-\newcommand{\belaid}[1]{\textcolor{blue}{#1}}
-\newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
+% \newcommand{\vangiang}[1]{\textcolor{magenta}{#1}}
+% \newcommand{\belaid}[1]{\textcolor{blue}{#1}}
+% \newcommand{\sylvain}[1]{\textcolor{teal}{#1}}
+\newcommand{\change}[1]{\textcolor{red}{#1}}
 
@@
 
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%% \fntext[label3]{}
-
 \title{Trap spaces of Boolean networks are conflict-free siphons of their Petri net encoding}
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%%
-%% \affiliation[label2]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
 
@@
 
-\affiliation[1]{organization={LIS, Aix-Marseille University},%Department and Organization
-            %addressline={},
+\affiliation[1]{organization={LIS, Aix-Marseille University},
             city={Marseille},
-            %postcode={},
-            %state={},
             country={France}}
@@
 
-\affiliation[2]{organization={Lifeware team, Inria Saclay center},%Department and Organization
-            %addressline={},
+\affiliation[2]{organization={Lifeware team, Inria Saclay center},
             city={Palaiseau},
-            %postcode={},
-            %state={},
             country={France}}
@@
 \begin{abstract}
-%% Text of abstract
 
 Boolean network modeling of gene regulation but also of post-trans\-criptomic systems has proven over the years that it can bring powerful analyses and corresponding insight to the many cases where precise biological data is not sufficiently available to build a detailed quantitative model.
-%This is even more true for very large models where such data is frequently missing and led to a constant increase in size of logical models \emph{à la} Thomas.
 Besides simulation, the analysis of such models is mostly based on attractor computation, since those correspond roughly to observable biological \emph{phenotypes}.
@@
 \end{abstract}
-
-%%Graphical abstract
-% \begin{graphicalabstract}
-% %\includegraphics{grabs}
-% \end{graphicalabstract}
 
@@
 \begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-
-%% PACS codes here, in the form: \PACS code \sep code
-
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code 2000 is the default
-
 Logical model \sep Boolean network \sep Trap space \sep Attractor computation \sep Petri net \sep Siphon \sep Systems biology
@@
 
-%% main text
 \section{Introduction}
@@
 
-%After recalling several technical preliminaries (Section~\ref{sec:Preliminaries}), we expose the concrete need for such a method and detail its implementation using Answer Set Programming (Section~\ref{sec:Computation}).
-
-%All models used for evaluation and the implementation of the presented method are available at \url{https://github.com/soli/trap-spaces-as-siphons} and executable as a CoLoMoTo docker image.
-
-% Mention the extensions for the journal version
 Herein we revise and extend our previous work in~\cite{DBLP:conf/cmsb/TrinhBHS22} as follows.
@@
 Besides minimal trap spaces, these special types also play crucial roles in analyzing and controlling Boolean networks~\cite{Rozum2021}.
-%Fifth, we present the idea for using our Petri net approach to handle the problem of inconsistent and incomplete data in modeling biological systems.
 Fifth, regarding the implementation, we have developed a new converter that directly reads a \texttt{.bnet} file and builds the Petri net encoding, instead of using the \texttt{PNML} conversion of \texttt{bioLQM}~\cite{DBLP:conf/cmsb/TrinhBHS22}.
@@
 
-%%% Paper structure %%%
 The rest of this paper is organized as follows: Section~\ref{sec:Preliminaries} recalls the basic concepts including Boolean networks, attractors, trap spaces, Petri nets, and siphons.
@@
 We shall briefly recall here some preliminaries on Boolean networks related to trap spaces and Petri nets.
-%In the case of multi-level logical models, an encoding into a Boolean network is always possible~\cite{Didier2011}.
-%\vangiang{Remove this statement because there is not sure if the encoded Boolean network preserves the trap spaces of the original multi-level logical model.}
 
@@
 \]
-where \(v(t)\) (resp. \(v_i(t)\)) is the state of \(\mathcal{N}\) (resp. the state of node \(v_i\)) at time \(t\). % and \(v_i(t + 1)\) is the state of node \(v_i\) at time \(t + 1\).
+where \(v(t)\) (resp. \(v_i(t)\)) is the state of \(\mathcal{N}\) (resp. the state of node \(v_i\)) at time \(t\).
 Note that for simplicity, we write \(f_i(v(t))\) even if \(\IN(v_i) \subsetneq V\) (i.e., \(\IN(v_i)\) does not contain some nodes of \(V\)).
@@
   Without loss of generality, we assume that \(p = p_v\), then there is a directional arc from \(t\) to \(p_v\) and a directional arc from \(\overline{p}_v\) to \(t\).
-  %In addition, there is also no arc or a bidirectional arc between \(t\) and another place rather than \(p_v\) and \(\overline{p}_v\).
   We follow the following procedure to find a state \(s \in \mathcal{S}_{\mathcal{N}}[m]\) such that \(m_s(p') = 1, \forall p' \in \pred(t)\) where \(m_s\) is the corresponding marking in \(\mathcal{P}\) of \(s\).
@@
 In the case that \(m^{*} = \varepsilon\), we have all maximal trap spaces of the original Boolean network.
-
-%\section{Inconsistent and incomplete data}%
-%\label{sec:incomplete-data}
-%
-%So far we have assumed that we are always able to derive complete
-%and consistent truth tables (i.e., Boolean functions) that correctly capture the behavior of each node in a Boolean network.
-%However, in practice it is rarely the
-%case that a Boolean network of a biological system is fully understood and indeed, this is one important reason for modeling such networks.
-%The data provided may be incomplete in the sense that information is missing about what happens in certain states, or it may be inconsistent in that we have conflicting information.
-%The result is that the behavior of some nodes under certain conditions may be non-deterministic~\cite{Steggles2006}.
-%
-%Non-deterministic behavior is problematic for Boolean networks with the synchronous update scheme, which are unable to represent the possibility of more than one next state.
-%It can be addressed by using Boolean networks with asynchronous update schemes because of their non-determinism.
-%However, multiple update functions may be considered for each node, which is problematic for existing analysis methods in Boolean networks.
-%For example, if the truth table of a node have \(k\) unknown-output rows, then this node may have \(2^k\) possible Boolean functions and the analysis methods such as \texttt{PyBoolNet} and \texttt{mpbn} need to consider \(2^k\) possible Boolean networks due to they require complete Boolean functions for their underlying processing.
-%Moreover, one interesting question maybe raised from systems biologists is if there is a common behavioral property (e.g., minimal trap space, fixed point, attractor) among all possible combinations.
-%To answer this question, the existing analysis methods maybe need to consider \(2^k\) possible Boolean networks as well.
-%
-%We shall show how our Petri net approach presented the previous section can handle the problem of inconsistent and incomplete data.
-%The idea is to build the Petri net encoding of the Boolean network from its partial truth tables.
-%More specifically, for each unknown-output row in the truth table of a node, we create two conflicting transitions representing the two possible outputs (i.e., 0 or 1).
-%For the remaining rows, we create transitions as usual following the Petri net encoding~\cite{chatain2014characterization}.
-%Because of the non–deterministic choice mechanisms of Petri nets, the new Petri net is still equivalent in dynamics to the incomplete Boolean network.
-%One notable advantage of the connection between trap spaces of a Boolean network and conflict-free siphons of its Petri net encoding is that it still holds if the Petri net encoding changes (e.g., more or less transitions) but still maintain the equivalence in dynamics to the Boolean network.
-%Hence, the minimal trap spaces of the incomplete Boolean network can be computed by computing maximal conflict-free siphons of the resulting Petri net.
-%As a consequence, we can easily answer the question on common behavioral property.
-%
-%\begin{example}
-%Consider an incomplete Boolean network \(\mathcal{N} = (V, F)\) (mentioned in~\cite{Steggles2006}), where \(V = (g_1, g_2, g_3)\) and \(F = (f_1, f_2, f_3)\) with \(f_1 = g_2, f_3 = \neg g_1\).
-Herein, \(f_2\) is an incomplete function and the partial truth table is as follows.
-%\begin{center}
-%	\begin{tabular}{rr|r}
-%		\toprule
-%		$g_1$ & $g_3$ & $g_2$ \\ \midrule
-%		0 & 0 & 0 \\
-%		0 & 1 & 0 \\
-%		1 & 0 & ? \\
-%		1 & 1 & 1 \\
-%		\bottomrule
-%	\end{tabular}
-%\end{center}
-%We have two possible update functions for \(g_2\): \(f_2 = g_1 \land g_3\) for the case that \(? = 0\) and \(f_2 = g_1\) for the case that \(? = 1\).
-%Accordingly, we have two possible Boolean networks.
-%The first one has one minimal trap space: 001.
-%The second one has two minimal trap spaces: 001 and 110.
-%Both have a common minimal trap space: 001.
-%The asynchronous dynamics of the incomplete Boolean network is shown in Figure~\ref{fig:incomplete-dynamics-new-PN}a.
-%The resulting Petri net following our idea is given in Figure~\ref{fig:incomplete-dynamics-new-PN}b.
-%Herein, \(t^3_{g_2}\) and \(t^4_{g_2}\) are the two conflicting transitions representing the unknown-output row.
-%This Petri net has only one maximal conflict-free siphon \(\{p_{g_1}, p_{g_2}, \overline{p}_{g_3}\}\) corresponding to the common minimal trap space 001.
-%
-%\label{example:incomplete-BN}%
-%\end{example}
-%
-%\begin{figure}[!htb]
-%	\centering
-%	\begin{subfigure}[b]{0.3\textwidth}
-%		\centering
-%		\begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]
-%		\node[] (0) [] {000};
-%		\node[] (1) [left=of 0] {001};
-%		\node[] (4) [above=of 0] {100};
-%		\node[] (6) [above=of 1] {110};
-%		\node[] (5) [below=of 1] {101};
-%		\node[] (2) [left=of 6] {010};
-%		\node[] (3) [left=of 1] {011};
-%		\node[] (7) [below=of 3] {111};
-%
-%		\draw[->] (0) edge [] (1);
-%		\draw[->] (4) edge [] (0);
-%		\draw[->] (4) edge [bend left] (6);
-%		\draw[->] (6) edge [bend left] (4);
-%
-%		\draw[->] (5) edge [] (1);
-%		\draw[->] (5) edge [] (7);
-%
-%		\draw[->] (2) edge [] (6);
-%		\draw[->] (2) edge [] (0);
-%		\draw[->] (2) edge [] (3);
-%
-%		\draw[->] (3) edge [] (1);
-%		\draw[->] (3) edge [] (7);
-%
-%		\draw[->] (7) edge [] (6);
-%
-%		\end{tikzpicture}
-%		\caption{}%
-%		%\label{fig:stg}
-%	\end{subfigure}\hfill%
-%	\begin{subfigure}[b]{0.6\textwidth}
-%		\centering
-%		\begin{tikzpicture}[node distance=1cm and 1cm, every node/.style={scale=1.0}]\footnotesize
-%		\node[circle,draw,label=left:$p_{g_1}$] (x1) {};
-%		\node[circle,draw,label=left:$\overline{p}_{g_1}$] (nx1) [below=of x1, yshift=-0.5cm] {};
-%
-%		\node[circle,draw,label=right:$p_{g_2}$] (x2) [right=of x1, xshift=2.5cm] {};
-%		\node[circle,draw,label=right:$\overline{p}_{g_2}$] (nx2) [below=of x2, yshift=-0.5cm] {};
-%
-%		\node[circle,draw,label=left:$p_{g_3}$] (x3) [below=of nx1, yshift=-0.5cm] {};
-%		\node[circle,draw,label=right:$\overline{p}_{g_3}$] (nx3) [below=of nx2, yshift=-0.5cm] {};
-%
-%		\node[rectangle,draw,label=above:$t^1_{g_2}$] (t1x2) [right=of x1] {};
-%		\node[rectangle,draw,label=above:$t^2_{g_2}$] (t2x2) [right=of nx1] {};
-%
-%		\node[rectangle,draw,label=left:$t^3_{g_2}$] (t3x2) [left=of nx2] {};
-%		\node[rectangle,draw,label=left:$t^4_{g_2}$] (t4x2) [left=of nx3] {};
-%
-%		\draw[->] (x2) -- (t1x2);
-%		\draw[->] (t1x2) -- (nx2);
-%		\draw[<->] (t1x2) -- (nx1);
-%
-%		\draw[->] (nx2) edge [bend left] (t2x2);
-%		\draw[->] (t2x2) -- (x2);
-%		\draw[<->] (t2x2) -- (x1);
-%		\draw[<->] (t2x2) -- (x3);
-%
-%		\draw[->] (x2) -- (t3x2);
-%		\draw[->] (t3x2) -- (nx2);
-%		\draw[<->] (t3x2) -- (x1);
-%		\draw[<->] (t3x2) -- (nx3);
-%
-%		\draw[->] (nx2) -- (t4x2);
-%		\draw[->] (t4x2) -- (x2);
-%		\draw[<->] (t4x2) edge [bend left=15] (x1);
-%		\draw[<->] (t4x2) -- (nx3);
-%
-%		\end{tikzpicture}
-%		\caption{}%
-%		%\label{fig:PN}
-%	\end{subfigure}%
-%	\caption{Asynchronous dynamics (self arcs are omitted for simplicity) and new Petri net encoding (the transitions corresponding to \(g_1\) and \(g_3\) are omitted for simplicity) of the incomplete Boolean network of Example~\ref{example:incomplete-BN}.}%
-%	\label{fig:incomplete-dynamics-new-PN}
-%\end{figure}
-%
-%In summary, we can still perform meaningful analysis on the resulting Petri net.
-%Note that the new Petri net may have more transitions, however the number of additional transitions is only \(O(2k)\) where \(k\) is the number of unknown-output rows in the truth tables.
-%This is actually a significant advantage, since we do not need to consider \(2^k\) possible Boolean networks.
-%It is worth noting that the approach proposed in~\cite{Steggles2006} is only shown to deal with the case that \(k = 1\), thus it is not clear how this approach can handle the case that \(k > 1\).
-%Moreover, in the context of Boolean models construction of biological systems from prior knowledge and experimental data, the work by~\cite{chevalier2020synthesis} also deals with some kind of uncertainty.
-%It uses ASP to represent all possible Boolean networks possessing the same given property, where each solution of this ASP is equivalent to a distinct Boolean network.
-%Then it samples through that solution space with heuristics to select \emph{diverse} Boolean networks, and proceed further simulations or analysis on every individual Boolean network.
-%Compared to our Petri net approach presented above, it still requires to consider multiple Boolean networks as well as it is only applicable for locally-monotonic Boolean networks~\cite{chevalier2020synthesis}.
-%
-%To this end, as more data becomes available for the underlying biological system, the Petri net model can be refined to reduce the
-%amount of non-determinism (i.e., the number of conflicting transitions) it contains.
-%Specifically, if we know the output for an unknown-output row of a node, we can remove one of the two conflicting transitions corresponding to this unknown-output row.
-%Hence, Petri nets provide an interesting means of documenting the development of knowledge for a biological system, and our Petri net-based approach can be more useful in the model refinement loop.
 
@@
 
-%\setlength{\tabcolsep}{4pt}
-%\begin{table}[!htb]
-%  \caption{Results on the real-world models from the BBM repository.}
-%  \centering%
-%  \label{tab:bbm_repo}
-%  \begin{tabular}{lrrrr}
-%    \toprule
-%    Method & \# failures & avg-lqm (s) & avg-mono (s) & avg-all (s) \\ \midrule
-%    \texttt{bioLQM} & 9 (134) & 12.87 & N/A & N/A \\
-%    \texttt{PyBoolNet} & 12 & 8.87 & 11.00 & 13.59 \\
-%    \texttt{mpbn} & 2 (187) & N/A & 2.31 & N/A \\
-%    \texttt{Trappist}-MaxSAT & 1 & 0.03 & 1.09 & 1.01 \\
-%    \texttt{Trappist}-CP & - & - & - & - \\
-%    \texttt{Trappist}-ILP & - & - & - & - \\
-%    \texttt{Trappist}-ASP & 1 & 0.05 & 1.02 & 0.93 \\
-%    \bottomrule
-%  \end{tabular}
-%\end{table}
-%\setcounter{rownumber}{0}
-
 \begin{figure}[!ht]
@@
 
-%The second observation is that the proposed method vastly outperforms \texttt{PyBoolNet} in computational time, on each and every model, and sometimes with orders of magnitude of difference (e.g., for most models in the 100--1000 nodes size range).
-%Note that for all the cases where \texttt{PyBoolNet} did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-implicants.
-%Hence, not even a single minimal trap space was output by that method.
-%The computational advantage is therefore immediately a practical advantage since on the one hand the state-of-the-art method did not allow any analysis whatsoever of the models, and on the other hand the proposed method could provide, very often under one second, the first thousand minimal trap spaces.
-%For modellers having a critical look at a model and in a \emph{model, invalidate, refine} loop this means a huge difference in the models that are amenable to study.
-%
-%Note that even with a very restricted time-limit of two minutes, it was possible with the proposed technique to find \emph{all} minimal trap spaces of small models (roughly under 130 nodes, i.e., considered as quite big up to now).
-%Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
-%Since the ASP code is declarative by nature, it is also possible to add to it supplementary constraints coming from the modeler in case one is looking for specific attractors.
-%Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration.
-%
-%The third observation is that for all the models where \texttt{PyBoolNet} finished before the timeout, once \texttt{PyBoolNet} went through the prime-implicant phase, its ASP solving phase quickly returned the first 1000 minimal trap spaces, all under one second.
-%For these models, the ASP solving phase of the proposed method also took very short time, all under one second.
-%Hence, with the experimental results shown in this paper, the practical differences between our ASP encoding and that of \texttt{PyBoolNet} are not distinctly exposed.
-%The fact that our new ASP encoding is guaranteed to be linear in the number of nodes of the original model does not seem to be crucial here, however a much deeper analysis of those cases remains to be done.
-%
-%Note that though enumerating the extremal siphons of a Petri net is exponential (see~\cite{nabli2016enumerating} for instance) this is apparently not the bottleneck of the proposed method, showing once again that networks obtained from biochemical models do have a specific structure.
-
 \subsection{Randomly generated models}
@@
 
-%\begin{table}[!htb]
-%  \caption{Results on N-K models.}
-%  \centering%
-%  \label{tab:result_N_K}
-%  \begin{tabular}{rrrrrrrr}
-%    \toprule
-%    & & & & \multicolumn{4}{c}{\texttt{Trappist}}\\
-%    \cmidrule(rr){5-8}
-%    \(n\) & \texttt{LQM} & \texttt{mpbn} & \texttt{PBN} & \texttt{SAT} & \texttt{CP} & \texttt{ILP} & \texttt{ASP} \\
-%    \midrule
-%    100 & 50 ($>$ 180) & 50 (N/A) & 0 (0.07) & 0 (0.05) & - () & - () & 0 (0.09) \\
-%    150 & 50 ($>$ 180) & 50 (N/A) & 0 (0.14) & 0 (0.10) & - () & - () & 0 (0.14) \\
-%    200 & 50 ($>$ 180) & 50 (N/A) & 0 (0.43) & 0 (0.25) & - () & - () & 0 (0.24) \\
-%    250 & 50 ($>$ 180) & 50 (N/A) & 0 (1.92) & 0 (1.04) & - () & - () & 0 (0.56) \\
-%    300 & 50 ($>$ 180) & 50 (N/A) & 0 (9.68) & 0 (4.46) & - () & - () & 0 (1.83) \\
-%    350 & 50 ($>$ 180) & 50 (N/A) & 1 (46.54) & 0 (20.09) & - () & - () & 0 (6.10) \\
-%    400 & 50 ($>$ 180) & 50 (N/A) & 29 (144.09) & 12 (90.36) & - () & - () & 1 (33.01) \\
-%    \bottomrule
-%  \end{tabular}
-%\end{table}
-
-%Table~\ref{tab:result_N_K} shows the experimental results on N-K models.
-%Column \(n\) denotes the network size.
-%Columns \texttt{LQM} and \texttt{PBN} show the results of \texttt{bioLQM} and \texttt{PyBoolNet}, respectively.
-%For each method, the number outside the parentheses indicates the number of failures, whereas the number inside the parentheses indicates the average running time (in seconds).
-%Note that when computing the average running time, if the running time exceeds 180s, it is considered as 180s.
-%From these results, we obtained several observations consistent with those obtained for real-world models.
-
 Figure~\ref{fig:random_cumul_first} shows cumulative numbers of random models solved by the compared methods with respect to enumerating all the minimal trap spaces.
@@
 
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
-%% \appendix
-
-%% \section{}
-%% \label{}
-
-%% For citations use:
-%%       \citet{<label>} ==> Jones et al. [21]
-%%       \citep{<label>} ==> [21]
-%%
-
-%% If you have bibdatabase file and want bibtex to generate the
-%% bibitems, please use
-%%
 \bibliographystyle{elsarticle-num-names}
@@
 
-%% else use the following coding to input the bibitems directly in the
-%% TeX file.
-
-% \begin{thebibliography}{00}
-
-% %% \bibitem[Author(year)]{label}
-% %% Text of bibliographic item
-
-% \bibitem[ ()]{}
-
-% \end{thebibliography}
 \end{document}
@@
 \endinput
-%%
-%% End of file `elsarticle-template-num-names.tex'.

### 2023-06-06 manuscripts/TCS2023/tcs2023.tex
Remarks of Reviewer 1, up to proof of Thm. 3.1

@@
 The recent use of trap spaces made a real breakthrough in that field allowing to consider medium-sized models that used to be out of reach.
-However, with the continuing increase in model size and complexity of Boolean update functions, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits due to the difficulty of the prime-implicant computation.
+However, with the continuing increase in model size and complexity of Boolean update functions, the state-of-the-art computation of minimal trap spaces based on \emph{\change{prime implicants}} shows its limits due to the difficulty of the prime-implicant computation.
 
@@
 Besides important theoretical applications in studying properties of trap spaces, the connection enables us to propose an alternative approach to compute minimal trap spaces, and hence complex attractors, of a general Boolean network.
-It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
+It replaces the need for \change{prime implicants} by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
 We then demonstrate its efficiency and compare it to the state-of-the-art methods on a large collection of real-world and randomly generated models.
@@
 The recent use of trap spaces~\cite{klarner2015computing} made a real breakthrough in that field allowing to consider medium-sized models that used to be out of reach and for which only simulation was available.
-However, with the most recent models both being quite large and using rather complex update functions, the state-of-the-art computation of minimal trap spaces based on \emph{prime-implicants} shows its limits.
+However, with the most recent models both being quite large and using rather complex update functions, the state-of-the-art computation of minimal trap spaces based on \emph{\change{prime implicants}} shows its limits.
 More specifically, the number of prime implicants of a Boolean function is in general exponential in the number of input nodes of this function~\cite{klarner2015computing}.
@@
 This method has been implemented in the tool \texttt{mpbn}\footnote{\url{https://github.com/bnediction/mpbn}} demonstrated in~\cite{Paulev2020} for handling medium-sized models from the literature and very large synthetic models (up to 100,000 nodes).
-However, this method is only applicable for \emph{locally-monotonic} Boolean networks, whereas the prime-implicants based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean networks (i.e., including both locally-mono\-tonic and non-locally-monotonic ones).
+However, this method is only applicable for \emph{locally-monotonic} Boolean networks, whereas the \change{prime implicants} based method~\cite{klarner2015computing} is applicable for \emph{general} Boolean networks (i.e., including both locally-mono\-tonic and non-locally-monotonic ones).
 In addition, the \texttt{bioLQM} platform also provides another method using Binary Decision Diagrams (BDDs) in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html}.
@@
 Specifically, based on the connection, we propose an alternative approach to compute minimal trap spaces, and hence complex attractors, of a general Boolean network.
-It replaces the need for prime-implicants by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
+It replaces the need for \change{prime implicants} by a completely different technique, namely the enumeration of maximal siphons in the Petri net encoding of the original model.
 We then demonstrate its efficiency and compare it to the state-of-the-art methods for computing minimal trap spaces of Boolean networks on many real-world models from various sources in the literature and on randomly generated models.
@@
 Then the dynamics of \(\mathcal{N}\) is captured by the directed graph \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\) called the State Transition Graph (STG).
-There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: \emph{synchronous}, where all the nodes are update simultaneously, and \emph{fully asynchronous}, where only one node is selected non-deterministically to be updated.
+There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: \emph{synchronous}, where all the nodes are \change{updated} simultaneously, and \emph{fully asynchronous}, where only one node is selected non-deterministically to be updated.
 
@@
 We denote \(\mathcal{S}_{\mathcal{N}}^{\star} = (\mathbb{B} \cup \{\star\})^n\) the set of all possible subspaces of \(\mathcal{N}\).
-Note that \(\left\vert\mathcal{S}_{\mathcal{N}}^{\star}\right\vert = 3^n\) and \(S_{\mathcal{N}} \subset \mathcal{S}_{\mathcal{N}}^{\star}\)~\cite{klarner2015computing}.
+Note that \(\left\vert\mathcal{S}_{\mathcal{N}}^{\star}\right\vert = 3^n\) and \change{\(\mathcal{S}_{\mathcal{N}} \in \mathcal{S}_{\mathcal{N}}^{\star}\)}~\cite{klarner2015computing}.
 
@@
 For example, let us consider the Boolean network shown in Example~\ref{example:BN}.
-Figure~\ref{fig:stg} shows the dynamics of this model under the fully asynchronous update (i.e., only one node is updated at each time step).
+Figure~\ref{fig:stg} shows the dynamics of this model under the fully asynchronous update \change{scheme} (i.e., only one node is updated at each time step).
 The model has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
@@
 \end{tikzpicture}
-\caption{State transition graph, under the fully asynchronous update.}%
+\caption{State transition graph, under the fully asynchronous update \change{scheme}.}%
 \label{fig:stg}
@@
 In this work, we consider a class of Petri nets called 1-safe Petri nets where every place has at most 1 token and all arcs are of weight 1.
+\change{Note that in such nets we have \(m : P \mapsto \{0,1\}\), we might therefore represent a marking by the equivalent set of places containing a token and will use this notation for simplicity.}
 In this case, weights are implicitly omitted in the arcs of a Petri net.
@@
 Each conjunct of the logical function that activates the \emph{gene} will lead to a transition \(t\), consuming the inactive place (i.e., a directional arc from \(\overline{p}_v\) to \(t\)), producing the active place (i.e., a directional arc from \(t\) to \(p_v\)), and with all other literals both consumed and produced (i.e., a bidirectional arc).
-And conversely for the inactivation.
+\change{Conversely a transition is added from the active place to the inactive place for each conjunct of the negation of that function.}
 Let \(s\) be a state of the Boolean network and \(m_s\) be its corresponding marking in the encoded Petri net.
-It holds that \(\forall v \in V\), \(s(v) = 0\) if and only if \(m_s(\overline{p}_v) = 1\) and \(s(v) = 1\) if and only if \(m_s(p_v) = 1\).
+It holds that \(\forall v \in V\), \(s(v) = 0\) if and only if \(m_s(\overline{p}_v) = 1\) \change{and \(m_s(p_v) = 0\)} and \(s(v) = 1\) if and only if \(m_s(p_v) = 1\) \change{and \(m_s(\overline{p}_v) = 0\)}.
 Note also that at any marking \(m\) of the Petri net encoding a Boolean network, it always holds that \(m(p_v) + m(\overline{p}_v) = 1\).
@@
 In~\cite{chaouiya2004qualitative} this took the form of the whole truth table of the Boolean functions, but as shown in Appendix 1 of~\cite{chatain2014characterization} computing Disjunctive Normal Forms (DNF) of each Boolean function is enough.
-Though this might appear quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
+Though this might appear quite computationally intensive it is important to remark first that contrary to the \change{prime implicants} case, there is no need to find \emph{minimal} DNFs.
 One way to look at this is to consider that this amounts to a similar approach as that used in~\cite{DBLP:conf/ictai/ChevalierFPZ19} but with the encoding of both activation and inhibition functions as DNFs in order to take into account possible non-local-monotonicity.
@@
 
-First, we add a definition related to any set of places of a Petri net encoding a Boolean network, and notably a siphon of such a net.
+\change{First let us associate subspaces and sets of places in the Petri net encoding.}
+\begin{definition}\label{def:mirror}
+
+  Let \(m\) be a subspace of Boolean network \(\mathcal{N} = (V, F)\).
+  A \emph{mirror} of \(m\) is a set of places \(S\) in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{N}\) such that:
+  \[\forall v \in D_m \left[m(v) = 0 \Leftrightarrow p_v \in S \land m(v) = 1 \Leftrightarrow \overline{p}_v \in S\right]\] and \[\forall v \in V \setminus D_m \left[p_v \not \in S \land \overline{p}_v \not \in S\right].\]
+
+\end{definition}
+
+\change{Now}, we add a definition related to any set of places of a Petri net encoding a Boolean network, and notably a siphon of such a net.
 
@@
 Intuitively, a siphon is a set of places that once unmarked remains so.
-If it is conflict-free then its dual corresponds to a subspace of the model such that whatever update, the fixed values remain so (since the unmarked places remain unmarked).
-This is precisely the definition of a trap space and maximality of the siphon is equivalent to as many fixed values as possible, hence minimality of the trap space.
+If it is conflict-free \change{it is possible to associate a subspace to it, more precisely it is the \emph{mirror} of a subspace.
+Since it is a siphon, the fixed values will remain so whatever update happens, as the unmarked places remain unmarked.
+The subspace corresponding to that conflict-free siphon is therefore a trap space, and the maximality of the siphon is equivalent to the minimality of the trap space (as many fixed values as possible).}
 For example, the Boolean network given in Example~\ref{example:BN} has two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
@@
 
-\begin{definition}
-
-  Let \(m\) be a subspace of Boolean network \(\mathcal{N} = (V, F)\).
-  A \emph{mirror} of \(m\) is a set of places \(S\) in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{N}\) such that:
-  \[\forall v \in D_m \left[m(v) = 0 \Leftrightarrow p_v \in S \land m(v) = 1 \Leftrightarrow \overline{p}_v \in S\right]\] and \[\forall v \in V \setminus D_m \left[p_v \not \in S \land \overline{p}_v \not \in S\right].\]
-
-\end{definition}
-
 \begin{theorem}%
@@
 
-  First, we show that if \(m\) is a trap space of \(\mathcal{N}\), then \(S\) is a conflict-free siphon of \(\mathcal{P}\) (*).
+  \change{\textit{First, we show that if \(m\) is a trap space of \(\mathcal{N}\), then \(S\) is a conflict-free siphon of \(\mathcal{P}\) (*).}}
+
   If \(D_m = \emptyset\), then \(S = \emptyset\) is trivially a conflict-free siphon of \(\mathcal{P}\).
@@
   Without loss of generality, we assume that \(p = p_v\), then there is a directional arc from \(t\) to \(p_v\) and a directional arc from \(\overline{p}_v\) to \(t\).
+
   We follow the following procedure to find a state \(s \in \mathcal{S}_{\mathcal{N}}[m]\) such that \(m_s(p') = 1, \forall p' \in \pred(t)\) where \(m_s\) is the corresponding marking in \(\mathcal{P}\) of \(s\).
   For every place \(p' \in \pred(t)\), let \(p''\) be the complementary place of \(p'\) and \(v'\) be the corresponding node in \(\mathcal{N}\) of \(p'\) and \(p''\).
-  If \(p'' \not \in S\), then \(v' \not \in D_m\) and we can always set a Boolean value to \(s(v')\) such that \(s \in \mathcal{S}_{\mathcal{N}}[m]\) and \(m_s(p') = 1\).
+
+  If \(p'' \not \in S\), then \(v' \not \in D_m\) and we can always set \change{the} Boolean value to \(s(v')\) such that \(s \in \mathcal{S}_{\mathcal{N}}[m]\) and \(m_s(p') = 1\).
+
   If \(p'' \in S\), then \(v' \in D_m\) and we set \(s(v') = m(v')\).
-  In this case, if \(p' = p_{v'}\) then \(s(v') = m(v') = 1\) leading to \(m_s(p') = 1\), if \(p' = \overline{p}_{v'}\) then \(s(v') = m(v') = 0\) leading to \(m_s(p') = 1\).
-  For the remaining nodes of \(\mathcal{N}\), we can always set Boolean values to these nodes to preserve that \(s \in \mathcal{S}_{\mathcal{N}}[m]\).
-  We also have \(m_s(p_v) = 0\) by the characteristics of the encoding~\cite{chaouiya2004qualitative}.
+  In this case, if \(p' = p_{v'}\) then \(s(v') = m(v') = 1\) leading to \(m_s(p') = 1\), if \(p' = \overline{p}_{v'}\) then \(s(v') = m(v') = 0\) leading to \(m_s(p') = \change{0}\).
+
+  For the remaining nodes of \(\mathcal{N}\), we can always set Boolean values to these nodes to preserve that \(s \in \mathcal{S}_{\mathcal{N}}[m]\) \change{by applying the same procedure}.
+  We also have \(m_s(p_v) = 0\) by the characteristics of the encoding~\cite{chaouiya2004qualitative} \change{(and Definition \ref{def:mirror})}.
   Now, \(t\) is enabled at marking \(m_s\).
@@
   This implies that \(s' \not \in \mathcal{S}_{\mathcal{N}}[m]\).
+
   For any firing scheme of \(\mathcal{P}\), the firing of \(t\) always happens.
@@
 
-  Second, we show that if \(S\) is a conflict-free siphon of \(\mathcal{P}\), then \(m\) is a trap space of \(\mathcal{N}\) (**).
+  \change{\textit{Second, we show that if \(S\) is a conflict-free siphon of \(\mathcal{P}\), then \(m\) is a trap space of \(\mathcal{N}\) (**).}}
+
   By the definition of a mirror, \(m\) is a subspace of \(\mathcal{N}\).
@@
   If \(p = p_{v}\), then \(m_s(p_{v}) = 1\) leading to \(m(v) = s(v) = 1\) by the characteristics of the encoding~\cite{chaouiya2004qualitative}.
-  By the definition of a mirror, \(m(v) = 0\) because \(p_{v} \in S\), which is a contradiction.
+  By the definition of a mirror, \(m(v) = 0\) because \(p_{v} \in S\), \change{meaning that \(m_s(p_{v}) = 0\)}, which is a contradiction.
+
   It is symmetric for the case that \(p = \overline{p}_{v}\).
@@
 A natural computational application of Theorem~\ref{theo:ts_2_sp} is that we can efficiently decide whether a subspace \(m\) is a trap space.
-In \texttt{PyBoolNet}~\cite{klarner2017pyboolnet}, this is checked by using the percolation on the prime-implicants of the Boolean functions.
-As we have mentioned at the beginning of this article, the computation of prime-implicants is a demanding task for complex Boolean networks, even is sometimes intractable.
+In \texttt{PyBoolNet}~\cite{klarner2017pyboolnet}, this is checked by using the percolation on the \change{prime implicants} of the Boolean functions.
+As we have mentioned at the beginning of this article, the computation of \change{prime implicants} is a demanding task for complex Boolean networks, even is sometimes intractable.
 Hence, the checking method in~\cite{klarner2017pyboolnet} shows its limitations.
@@
 for each \(v \in V\).
-Note that the number of atoms of \(\mathcal{L}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean network and that number might be exponential in \(n\).
+Note that the number of atoms of \(\mathcal{L}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of \change{prime implicants} of the Boolean network and that number might be exponential in \(n\).
 In~\cite{DBLP:conf/ictai/ChevalierFPZ19}, there is an ASP characterization of trap spaces that does not rely on minimal DNFs either and thus seems very similar to our ASP encoding.
@@
 We used a set of real-world Boolean networks lying in various scales collected from numerous bibliographic sources in the literature.
-Most of these models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants), and have never been fully analyzed.
+Most of these models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of \change{prime implicants}), and have never been fully analyzed.
 Note that these models are not included in the \texttt{PyBoolNet} and \texttt{BBM} repositories.
@@
 
-Fourth, regarding the comparison of the ASP-based methods (i.e.,\linebreak \texttt{PyBoolNet}, \texttt{mpbn}, and \texttt{Trappist-ASP}), we note that for all the models where \texttt{PyBoolNet} did not finish before the time limit, the timeout occurred during the computation of the prime-implicants.
+Fourth, regarding the comparison of the ASP-based methods (i.e.,\linebreak \texttt{PyBoolNet}, \texttt{mpbn}, and \texttt{Trappist-ASP}), we note that for all the models where \texttt{PyBoolNet} did not finish before the time limit, the timeout occurred during the computation of the \change{prime implicants}.
 Hence, not even a single minimal trap space was output by that method.
@@
 Hence, we searched for all solutions in each model, which makes the comparison to \texttt{bioLQM} more comprehensive.
-In addition, each node has only three input nodes, leading to the number of prime-implicants of the associated Boolean function is small.
-Hence, \texttt{PyBoolNet} always passed the phase of computing prime-implicants in every model even within one second, which enables us to compare the ASP encoding of \texttt{PyBoolNet} and that of \texttt{Trappist-ASP}.
+In addition, each node has only three input nodes, leading to the number of \change{prime implicants} of the associated Boolean function is small.
+Hence, \texttt{PyBoolNet} always passed the phase of computing \change{prime implicants} in every model even within one second, which enables us to compare the ASP encoding of \texttt{PyBoolNet} and that of \texttt{Trappist-ASP}.
 
@@
 Third, for every time limit, \texttt{Trappist-ASP} can always handle many more models than \texttt{PyBoolNet}, ranging from 29 to 65 more models.
-Since the time for the phase of computing prime-implicants of \texttt{PyBoolNet} is negligible in every model, most of the running time of \texttt{PyBoolNet} was spent for its ASP solving phase.
+Since the time for the phase of computing \change{prime implicants} of \texttt{PyBoolNet} is negligible in every model, most of the running time of \texttt{PyBoolNet} was spent for its ASP solving phase.
 Hence, we can easily see that the ASP encoding of \texttt{Trappist-ASP} is much better than that of \texttt{PyBoolNet}.

### 2023-06-06 manuscripts/TCS2023/tcs2023.tex
Done with reviewer 1's comments

@@
 \newtheorem{proposition}{Proposition}[section]
+\newtheorem{corollary}{Corollary}[section]
 
@@
 
-From the proof of Theorem~\ref{theo:ts_2_sp}, we can see that the theorem holds for any update scheme associated to the Boolean network.
-Since the Petri net encoding of a Boolean network is independent of its update scheme and siphons are a static property of a Petri net, we can imply that trap spaces of a Boolean network are independent of its update scheme.
+\change{Note that this proof gives us as corollary a well-known result on trap spaces.}
+
+\begin{corollary}
+  \change{Trap spaces of a Boolean network are independent of the update scheme.}
+\end{corollary}
+
+\begin{proof}
+\change{From the proof of Theorem~\ref{theo:ts_2_sp}, we can see that the theorem holds for any update scheme associated to the Boolean network.
+Since the Petri net encoding of a Boolean network is independent of its update scheme and siphons are a static property of a Petri net, we get that trap spaces of a Boolean network are independent of its update scheme.}
+\end{proof}
+
 Note that the original proof for this property of trap spaces (see Theorem 1 of~\cite{klarner2015computing}) only considers the two popular update schemes (i.e., synchronous and fully asynchronous).
@@
 See Subsection~\ref{subsec:computation_special_ts} for more discussions about many special types of trap spaces.
-It might actually be possible to generalize our result to any 1-safe place-complementary Petri net to define a notion of trap spaces that might be useful for the analysis of Petri nets, but this is out of the scope of the present article.
+It might actually be possible to generalize our result to any 1-safe place-complementary \change{(i.e., places are defined by pairs such that the markings are complementary)} Petri net to define a notion of trap spaces that might be useful for the analysis of Petri nets, but this is out of the scope of the present article.
 
@@
 
-The following Boolean Constraint Satisfaction Problem (CSP) directly derives from the above characterization:
+\change{A Constraint Satisfaction Problem (CSP) is defined by a triple giving its variables, their domains, and the constraints on those variables.}
+The following Boolean CSP directly derives from the above characterization:
 \begin{definition}
@@
 From the encoded CSP, the set of all \emph{minimal} siphons of the Petri net can be enumerated in the set inclusion order.
-For enumerating siphons in the set inclusion order, the proposed method by~\cite{nabli2016enumerating} uses the technique that labels directly the Boolean variables with increasing value selection (i.e., to test first the absence, then the presence of a place in the candidate solution).
+For enumerating siphons in the set inclusion order, the \change{method proposed in}~\cite{nabli2016enumerating} uses the technique that labels directly the Boolean variables with increasing value selection (i.e., to test first the absence, then the presence of a place in the candidate solution).
 The method has two implementations, one uses an iterated SAT procedure and the other uses Constraint Programming (CP) with backtracking.
@@
 
-However, using a state-of-the-art tool like \texttt{PyBoolNet}~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
+However, using a state-of-the-art tool like \texttt{PyBoolNet}~\cite{klarner2015computing} on that model \change{unfortunately} fails at the phase of prime-implicant generation.
 \texttt{mpbn}~\cite{Paulev2020} can return the first 1000 solutions within 1.43s, but indeed, it limits the modeling range of the modelers as it does not permit using non-locally-monotonic Boolean functions.
@@
 The current practice usually revolves then around fixing some source nodes to plausible values and reducing the model accordingly.
-While this approach makes sense, it relies on potentially arbitrary decisions, and \emph{hides away} critical modelling choices that were actually not part of the original Boolean network or even of the starting map.
-
-For the ``RA\_apoptosis'' model, using the ASP-based method presented in Subsection~\ref{subsec:computation_asp}, it is possible to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) within 0.19s, which is much quicker than \texttt{mpbn}.
+While this approach makes sense, it relies on potentially arbitrary decisions, and \emph{hides away} critical modelling choices that were \change{clearly} not part of the original Boolean network or even of the starting map.
+
+For the ``RA\_apoptosis'' model, using the ASP-based method presented in Subsection~\ref{subsec:computation_asp}, it is \change{now} possible to obtain the first 1000 minimal trap spaces (including ones that contain more than one state) within 0.19s, which is much quicker than \texttt{mpbn}.
 The needed time for the ``Alzheimer'' model is 0.79s.
-Unfortunately since this method was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.
+% Unfortunately since this method was not available at the time, the analysis of the model remained very high-level and qualitative, instead of being able to use the rich information of computed minimal trap spaces.
 
@@
     \rownb & irons\_yeast & 18 & 1 & \textbf{0.12} & 0.03 & 0.01 & 0.01 & \textbf{0.37} & \textbf{0.39} & 0.02\\
-    \rownb & jaoude\_thdiff & 103 & $1000^+$ & N/A & \textbf{0.85} & \textbf{0.45} & \textbf{0.56} & \textbf{DNF} & \textbf{DNF} & 0.09\\
+    \rownb & jaoude\_thdiff & 103 & $1000^+$ & N/A & \textbf{0.85} & \textbf{0.45} & \textbf{0.56} & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.09\\
     \rownb & klamt\_tcr & 40 & 8 & \textbf{0.11} & 0.01 & 0.01 & 0.01 & \textbf{1.98} & \textbf{1.22} & 0.02\\
@@
     \rownb & saadatpour\_guardcell & 13 & 1 & \textbf{0.10} & 0.06 & 0.00 & 0.00 & \textbf{0.53} & \textbf{0.45} & 0.02\\
-    \rownb & selvaggio\_emt & 56 & $1000^+$ & N/A & \textbf{0.48} & \textbf{0.28} & \textbf{0.28} & \textbf{DNF} & \textbf{DNF} & 0.09\\
+    \rownb & selvaggio\_emt & 56 & $1000^+$ & N/A & \textbf{0.48} & \textbf{0.28} & \textbf{0.28} & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.09\\
     \rownb & tournier\_apoptosis & 12 & 3 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & \textbf{0.74} & \textbf{0.75} & 0.01\\
     \rownb & xiao\_wnt5a & 7 & 4 & \textbf{0.10} & 0.01 & 0.00 & 0.00 & \textbf{1.00} & \textbf{0.89} & 0.01\\
-    \rownb & zhang\_tlgl & 60 & 156 & \textbf{0.60} & 0.09 & 0.09 & 0.07 & \textbf{37.26} & \textbf{DNF} & 0.04\\
-    \rownb & zhang\_tlgl\_v2 & 60 & 258 & \textbf{0.64} & 0.04 & 0.08 & 0.11 & \textbf{69.95} & \textbf{DNF} & 0.04\\
+    \rownb & zhang\_tlgl & 60 & 156 & \textbf{0.60} & 0.09 & 0.09 & 0.07 & \textbf{37.26} & \textbf{\change{NF}} & 0.04\\
+    \rownb & zhang\_tlgl\_v2 & 60 & 258 & \textbf{0.64} & 0.04 & 0.08 & 0.11 & \textbf{69.95} & \textbf{\change{NF}} & 0.04\\
     \bottomrule
@@
 Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 minimal trap spaces.
-``DNF'' means that the method did not finish the computation within the time limit of three minutes.
-In the case of \texttt{bioLQM}, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not recorded the running time of \texttt{bioLQM} because it always requires to compute all minimal trap spaces.
+``\change{NF}'' means that the method did not finish the computation within the time limit of three minutes.
+In the case of \texttt{bioLQM}, ``N/A'' means that the number of all minimal trap spaces of the model is larger than 1000 and we did not record the running time of \texttt{bioLQM} because it always requires to compute all minimal trap spaces.
 A number in bold indicates a ratio greater than three compared to the best result.
@@
 In particular, the second best variant of \texttt{Trappist} (i.e., \texttt{Trappist-MaxSAT}) is even far more efficient than \texttt{bioLQM} and \texttt{PyBoolNet}, and is comparable to \texttt{mpbn} on every model.
-It is worth noting that for 3 of the 29 models, \texttt{mpbn} did not give any answer because these models are locally-monotonic but all the other methods did, which confirms the limit of \texttt{mpbn} on the applicable class of models.
+It is worth noting that for 3 of the 29 models, \texttt{mpbn} did not give any answer because these models are \change{non-}locally-monotonic but all the other methods did, which confirms the limit of \texttt{mpbn} on the applicable class of models.
 
@@
 
-Currently, a research group has made a great effort for building a collection (called \texttt{BBM}) of real-world Boolean models from various sources used in systems biology.
+\change{The research group behind the \texttt{BBM} repository has recently undertaken considerable effort for building a collection of real-world Boolean models from various sources used in systems biology.}
 It aims to be a comprehensive collection suitable for benchmarking and testing new tools and methods.
@@
 This model set allows us to fairly consider \texttt{bioLQM} for comparison, since \texttt{bioLQM} always requires to compute all minimal trap spaces.
-We can first see that \texttt{Trappist-ASP} and \texttt{Trappist-MaxSAT} are still the two best methods as they can handle every model within 1s as well as they always can handle more models than all the remaining methods on every time limit.
+We can first see that \texttt{Trappist-ASP} and \texttt{Trappist-MaxSAT} are still the two best methods as they can handle every model within 1s \change{and} always can handle more models than all the remaining methods on every time limit.
 Second, \texttt{Trappist-CP} is better than \texttt{Trappist-ILP}, which is consistent with their comparison shown in the previous subsection.
 Third, one notable remark is that for the time limit of 100s or 180s, \texttt{Trappist-CP} can handle more models than all \texttt{bioLQM}, \texttt{PyBoolNet}, and \texttt{mpbn}.
-This remark shows that even with a not best implementation, our alternative approach is still better than the state-of-the-art methods on a certain set of real-world models.
+This remark shows that even \change{without focusing on the optimization of our implementation}, our alternative approach is still better than the state-of-the-art methods on a certain set of real-world models.
 This is supported by the fact that our alternative approach avoids the need for computing prime implicants (as opposed to \texttt{PyBoolNet}) and can handle non-locally-monotonic Boolean networks (as opposed to \texttt{mpbn}).
@@
 \begin{table}[!htb]
-  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (\texttt{LQM}), \texttt{PyBoolNet} (\texttt{PBN}), \texttt{mpbn} and the four variants of \texttt{Trappist} on selected models from the literature.}
+  \caption{Timing comparisons (in seconds) between \texttt{bioLQM} (\texttt{LQM}), \texttt{PyBoolNet} (\texttt{PBN}), \texttt{mpbn} and the four variants of \texttt{Trappist} on selected models from the literature.
+  \change{The models are sorted by size with a horizontal rule inserted to split at 100 and 200 nodes, as in~\cite{DBLP:conf/cmsb/TrinhBHS22}}}
 % in bold if greater than or equal to 10
@@
     \rownb & p53\_low\_dna~\cite{DesignPrinciplesGeneNetworks} & 16 & 1 & 0.41 & \textbf{1.76} & NM & 0.07 & 0.58 & 0.48 & 0.14 \\
-    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & 32 & \textbf{DNF} & \textbf{DNF} & NM & 0.03 & \textbf{8.41} & \textbf{12.38} & 0.19 \\
+    \rownb & FT-GRN~\cite{ChvezHernndez2022} & 23 & 32 & \textbf{\change{NF}} & \textbf{\change{NF}} & NM & 0.03 & \textbf{8.41} & \textbf{12.38} & 0.19 \\
     \rownb & DNA\_damage~\cite{DesignPrinciplesGeneNetworks} & 26 & 16 & \textbf{0.24} & \textbf{0.33} & NM & 0.02 & \textbf{3.91} & \textbf{5.33} & 0.05 \\
     \rownb & Rho-GTPases~\cite{DesignPrinciplesGeneNetworks} & 33 & 2 & 0.17 & 0.57 & \textbf{40.39} & 0.07 & \textbf{0.74} & 0.56 & 0.11 \\
-    \rownb & Pluripotency~\cite{YachieKinoshita2018} & 36 & 440 & \textbf{DNF} & \textbf{DNF} & NM & 0.16 & \textbf{138.92} & \textbf{DNF} & 0.28 \\
-    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & 276 & 0.37 & 0.43 & NM & 0.07 & \textbf{72.40} & \textbf{DNF} & 0.06 \\
-    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $1000^+$ & N/A & 0.11 & 0.36 & 0.17 & \textbf{DNF} & \textbf{DNF} & 0.06 \\
+    \rownb & Pluripotency~\cite{YachieKinoshita2018} & 36 & 440 & \textbf{\change{NF}} & \textbf{\change{NF}} & NM & 0.16 & \textbf{138.92} & \textbf{\change{NF}} & 0.28 \\
+    \rownb & Pluripotent~\cite{DesignPrinciplesGeneNetworks} & 36 & 276 & 0.37 & 0.43 & NM & 0.07 & \textbf{72.40} & \textbf{\change{NF}} & 0.06 \\
+    \rownb & Pancreatic\_Cancer~\cite{DesignPrinciplesGeneNetworks} & 43 & $1000^+$ & N/A & 0.11 & 0.36 & 0.17 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.06 \\
     \rownb & Drosophila~\cite{RodrguezVega2014} & 52 & 128 & 0.33 & 0.05 & 0.07 & 0.06 & \textbf{32.66} & \textbf{126.22} & 0.05 \\
     \rownb & Cacace\_TdevModel~\cite{Cacace2020} & 61 & 28 & \textbf{1.29} & \textbf{5.67} & NM & 0.06 & \textbf{7.51} & \textbf{23.15} & 0.08 \\
-    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $1000^+$ & N/A & \textbf{DNF} & 0.50 & 0.34 & \textbf{DNF} & \textbf{DNF} & 0.33 \\
-    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & \textbf{39.22} & \textbf{1.01} & 0.20 & 0.12 & \textbf{75.81} & \textbf{DNF} & 0.05 \\
+    \rownb & hedgehog~\cite{DesignPrinciplesGeneNetworks} & 65 & $1000^+$ & N/A & \textbf{\change{NF}} & 0.50 & 0.34 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.33 \\
+    \rownb & EMT~\cite{Rozum2021} & 69 & 268 & \textbf{39.22} & \textbf{1.01} & 0.20 & 0.12 & \textbf{75.81} & \textbf{\change{NF}} & 0.05 \\
     \rownb & Bcell~\cite{Dutta2019} & 73 & 72 & 0.23 & 0.04 & 0.08 & 0.06 & \textbf{18.95} & \textbf{81.85} & 0.05 \\
-    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $1000^+$ & N/A & 0.09 & 0.55 & 0.37 & \textbf{DNF} & \textbf{DNF} & 0.15 \\
-    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $1000^+$ & N/A & \textbf{107.57} & 0.76 & 0.56 & \textbf{DNF} & \textbf{DNF} & 0.16 \\
+    \rownb & mast\_cell~\cite{aghamiri2020automated} & 73 & $1000^+$ & N/A & 0.09 & 0.55 & 0.37 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.15 \\
+    \rownb & Corral\_ThIL17diff~\cite{corral2021interplay} & 92 & $1000^+$ & N/A & \textbf{107.57} & 0.76 & 0.56 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.16 \\
 
     \midrule % 100 <= n < 200
-     \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & 78 & \textbf{56.81} & \textbf{4.25} & 0.23 & 0.17 & \textbf{25.20} & \textbf{DNF} & 0.19 \\
-    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & 82 & \textbf{DNF} & \textbf{14.01} & 0.27 & 0.20 & \textbf{27.55} & \textbf{DNF} & 0.25 \\
-    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $1000^+$ & N/A & 0.54 & 1.09 & 0.84 & \textbf{DNF} & \textbf{DNF} & 0.27 \\
-    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $1000^+$ & N/A & 0.16 & 1.07 & 1.06 & \textbf{DNF} & \textbf{DNF} & 0.16 \\
-    \rownb & angiofull~\cite{Weinstein2017} & 142 & $1000^+$ & N/A & 0.17 & 1.06 & 0.88 & \textbf{DNF} & \textbf{DNF} & 0.23 \\
-    \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & 492 & \textbf{DNF} & \textbf{11.28} & 0.78 & 0.69 & \textbf{DNF} & \textbf{DNF} & 0.35 \\
-    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $1000^+$ & N/A & \textbf{DNF} & 1.43 & 1.55 & \textbf{DNF} & \textbf{DNF} & 0.19 \\
-    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & $1000^+$ & N/A & \textbf{13.58} & 1.76 & 1.51 & \textbf{DNF} & \textbf{DNF} & 0.27 \\
+     \rownb & Adhesion\_CIP~\cite{guberman2020boolean} & 121 & 78 & \textbf{56.81} & \textbf{4.25} & 0.23 & 0.17 & \textbf{25.20} & \textbf{\change{NF}} & 0.19 \\
+    \rownb & EMT\_Mech~\cite{Sullivan2022} & 136 & 82 & \textbf{\change{NF}} & \textbf{14.01} & 0.27 & 0.20 & \textbf{27.55} & \textbf{\change{NF}} & 0.25 \\
+    \rownb & macrophage~\cite{DesignPrinciplesGeneNetworks} & 136 & $1000^+$ & N/A & 0.54 & 1.09 & 0.84 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.27 \\
+    \rownb & angiogenesis~\cite{DesignPrinciplesGeneNetworks} & 141 & $1000^+$ & N/A & 0.16 & 1.07 & 1.06 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.16 \\
+    \rownb & angiofull~\cite{Weinstein2017} & 142 & $1000^+$ & N/A & 0.17 & 1.06 & 0.88 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.23 \\
+    \rownb & EMT\_Mech\_TGFbeta~\cite{Sullivan2022} & 150 & 492 & \textbf{\change{NF}} & \textbf{11.28} & 0.78 & 0.69 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.35 \\
+    \rownb & RA\_apoptosis~\cite{aghamiri2020automated} & 180 & $1000^+$ & N/A & \textbf{\change{NF}} & 1.43 & 1.55 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.19 \\
+    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & $1000^+$ & N/A & \textbf{13.58} & 1.76 & 1.51 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.27 \\
 
     \midrule % 200 <= n
-    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $1000^+$ & N/A & 1.13 & 1.47 & 1.43 & \textbf{DNF} & \textbf{DNF} & 0.31 \\
-    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & $1000^+$ & N/A & \textbf{DNF} & 1.52 & 2.26 & \textbf{DNF} & \textbf{DNF} & 0.35 \\
-    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $1000^+$ & N/A & \textbf{4.99} & NM & 1.96 & \textbf{DNF} & \textbf{DNF} & 0.28 \\
-    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & $1000^+$ & N/A & 0.42 & 2.36 & \textbf{4.91} & \textbf{DNF} & \textbf{DNF} & 0.44 \\
-    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & $1000^+$ & N/A & \textbf{DNF} & 2.56 & \textbf{5.62} & \textbf{DNF} & \textbf{DNF} & 0.46 \\
-    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & $1000^+$ & N/A & 0.36 & 2.99 & \textbf{4.81} & \textbf{DNF} & \textbf{DNF} & 0.37 \\
-    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & $1000^+$ & N/A & \textbf{DNF} & NM & \textbf{18.21} & \textbf{DNF} & \textbf{DNF} & 0.79 \\
+    \rownb & Snf1-pathway~\cite{Lubitz2015} & 202 & $1000^+$ & N/A & 1.13 & 1.47 & 1.43 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.31 \\
+    \rownb & T-cell-co-receptor~\cite{DesignPrinciplesGeneNetworks} & 206 & $1000^+$ & N/A & \textbf{\change{NF}} & 1.52 & 2.26 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.35 \\
+    \rownb & TcellCheckPoint~\cite{hernandez2020computational} & 218 & $1000^+$ & N/A & \textbf{4.99} & NM & 1.96 & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.28 \\
+    \rownb & Mycobacterium~\cite{DesignPrinciplesGeneNetworks} & 317 & $1000^+$ & N/A & 0.42 & 2.36 & \textbf{4.91} & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.44 \\
+    \rownb & Leishmania~\cite{DesignPrinciplesGeneNetworks} & 342 & $1000^+$ & N/A & \textbf{\change{NF}} & 2.56 & \textbf{5.62} & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.46 \\
+    \rownb & Cholocystokinin~\cite{aghamiri2020automated} & 383 & $1000^+$ & N/A & 0.36 & 2.99 & \textbf{4.81} & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.37 \\
+    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & $1000^+$ & N/A & \textbf{\change{NF}} & NM & \textbf{18.21} & \textbf{\change{NF}} & \textbf{\change{NF}} & 0.79 \\
 
@@
 Hence, we searched for all solutions in each model, which makes the comparison to \texttt{bioLQM} more comprehensive.
-In addition, each node has only three input nodes, leading to the number of \change{prime implicants} of the associated Boolean function is small.
+In addition, each node has only three input nodes, \change{leading to a small number of prime implicants of
+the associated Boolean function}.
 Hence, \texttt{PyBoolNet} always passed the phase of computing \change{prime implicants} in every model even within one second, which enables us to compare the ASP encoding of \texttt{PyBoolNet} and that of \texttt{Trappist-ASP}.
@@
 Recall that a Boolean network is non-locally-monotonic if only one of its Boolean functions is non-locally-monotonic.
-Hence, it is apparent that all this type of randomly generated models are non-locally-monotonic because of the number of nodes is large (\(n \geq 100\)).
+Hence, it is apparent that all \change{these types} of randomly generated models are non-locally-monotonic because of the number of nodes is large (\(n \geq 100\)).
 This observation confirms a limit on the applicable model class of \texttt{mpbn}.

### 2023-06-06 manuscripts/TCS2023/response.txt
Done with reviewer 1's comments
Some answers for Reviewer 2

@@
 
-We fixed all the minor remarks highlighted by the reviewer, as can be seen in
-red in the uploaded version.  Indeed it improved substantially the readability
-of the article.
+We thank the reviewer for their kind words and for a very thorough reading of
+our article.  We fixed all the minor remarks highlighted by the reviewer, as can
+be seen in red in the uploaded version.  We believe it improved substantially
+the readability of the article.
+
+* Reviewer 2
+
+"1. The authors employ an existing method of translating BNs to PNs. Anyway, they introduce a novel notion of a
+specific class of siphons. It would be great to see more discussion on that general concept. Especially, I am not
+familiar with the complexity of computing general siphons in PNs. It would be good to recall complexity issues
+around that and to discuss (at least informally) the complexity of the problem of computing all conflict-free siphons
+of the specific class of safe PNs."
+
+TODO
+
+"2. It is claimed that the PNs-based encoding (and also the results) are independent of the semantics of the update
+scheme. It is not clear how this encoding could work with most-permissive semantics. I can imagine synchronous or
+generally synchronous schemes to be embedded (still raising some questions on the complexity issues and
+adaptability of the three methods presented in trappist)."
+
+As explained in the article, and now extracted as a Corollary, the trap-spaces
+of a Boolean network are indeed independent of any update scheme.  The
+most-permissive semantics is not an update scheme (since it goes outside of the
+Boolean state space) but an approximation technique for reachability and
+trap-spaces.  As shown in the comparison with the mpbn tool that is based on
+that semantics, the proposed method still outperforms any existing technique for
+computing trap spaces, while also being able to handle non-locally-monotonic
+networks.
+
+"3. The paper is quite skeptic on the topic of using BDDs for identifying trap spaces. Anyway, BDDs have recently
+proven to be useful for computing attractors (e.g., Benes et al. CAV 2021). I am aware the trap spaces are much more
+complicated structures than (general) SCCs. Anyway, it might be good to see some deeper discussion that gives the
+reader more information on why BDD would not work. Or the claim could be rather made weaker."
+
+Since we do use BDDs in our implementation, in our case to obtain the Petri net
+encoding, we are fully aware of their great characteristics.  As explained in
+the introduction, what makes the current BDD-based approach, as implemented in
+bioLQM, not efficient is that it tries to represent *all* trap spaces of a model
+as a BDD and then to filter them to get minimal trap spaces. This is where the
+inefficiency lies, and not in the specific use of BDDs per se.
+
+We have added some points to highlight that we are not at any point claiming
+that BDDs in general are not suited to this kind of task.
+
+"4. At last of but not least, an investigative (may be naive) question comes to my mind: Are there (at the PN level) any
+static analysis approaches that might help even (or improve) in identifying conflict-free siphons based on the
+network structure? That would made the idea of the paper even more promising."
+
+We are not aware of any such technique, but have now highlighted (at the end of
+Section 3 and in conclusion) that the link between Petri nets and Trap spaces
+could bring novelties both ways.
+
+"The critical problem I see in this submission is that several parts of the related work are missing (especially the
+coverage of TCS background is quite low).
+First, I think there must be definitely more background on the related problems in the domain of PNs. There exist
+several approaches to compute siphons. I am not from the PN domain but by some fast googling I see many papers
+on computing siphons. It would be good to include some of them in the background and discuss if they might be also
+considered for computing conflict-free siphons. I can believe the simple argument is that existing approaches are
+general and hence cannot be that efficient as the approach presented in the paper. Anyway, that should be
+discussed explicitly to add some more rigour."
+
+As explained at the very end of Section 3, there is, to our knowledge, no method
+to compute maximal generic siphons, and even less maximal conflict-free
+siphons. We refer to a recent survey [34] but as it shows, the focus is on the
+minimal siphons, and of course generic ones.  We have added a sentence to
+explain even more why these methods are not adapted to our problem.
+
+"Second, in my opinion some work on attractor identification might be seemed as strongly related since there are
+several useful concepts that have strong relation to trap spaces. In particular, there is the quite recent phd work of
+Fontanels et al. elaborating on the concept of phenotype control where I think that phenotypes can somehow gain
+from computing the trap spaces. Also the network reductions as recently introduced by Tonello et al. (arxiv 2023)
+seem to be promising for shifting the scalability further. This comes to question especially on lines 692-698."
+
+TODO
+
+We would like to thank the reviewer for their useful comments. We also addressed all the minor issues they pointed out.
+
+Minor issues
+==========
+page 5. The general definition of BNs semantics at the line 35 is quite non-standard. I understand the attempt to
+capture a sort of subclass of update modes abstractly but I think it still would make sense to be more explicit and
+give some concrete instances (that can be discussed throughout the paper). In the current version I have the feeling
+you focus primarily on asynchronous update mode.
+
+Do we have a reference for this?
+
+Figure 1: add the topology of the BN
+
+TODO
+
+line 172: there is the abuse of notion $m$ (already used for "markings" of BNs)
+
+TODO
+m = subspaces = markings is indeed not nice

### 2023-06-06 manuscripts/TCS2023/tcs2023.tex
Some answers for Reviewer 2

@@
 By this approach, it requires the computation of all solutions, whereas the methods~\cite{klarner2015computing,Paulev2020} based on Answer Set Programming (ASP) can start enumerating them as they are found.
-Moreover, the main issue with the BDD-based method is that the number of generic trap spaces of a Boolean network may be extremely larger than its number of minimal trap spaces.
-This issue limits the efficiency of the BDD-based method.
+Moreover, the main issue with \change{this} BDD-based method is that the number of generic trap spaces of a Boolean network may be extremely larger than its number of minimal trap spaces.
+This issue limits the efficiency of the \change{current} BDD-based method.
 The study~\cite{DBLP:journals/tcs/NoualRS13} highlights the need for non-locally-monotonic Boolean networks in both biological and theoretical aspects.
@@
 
-A non-empty set \(T \subseteq \mathcal{S}_{\mathcal{N}}\) is a trap set with respect to \(\rightarrow\) if for every \(x \in T\) and \(y \in S\) with \(x \rightarrow y\) it holds that \(y \in T\)~\cite{klarner2015computing}.
+A non-empty set \(T \subseteq \mathcal{S}_{\mathcal{N}}\) is a \change{\emph{trap set}} with respect to \(\rightarrow\) if for every \(x \in T\) and \change{\(y \in \mathcal{S}_{\mathcal{N}}\)} with \(x \rightarrow y\) it holds that \(y \in T\)~\cite{klarner2015computing}.
 An attractor of \(\mathcal{N}\) with respect to \(\rightarrow\) can be defined as an inclusion-wise minimal trap set of \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\).
@@
 A \emph{trap space} is defined as a subspace that is also a trap set.
-It is noted that trap spaces of a Boolean network are independent of the update scheme of this model~\cite{klarner2015computing}.
+It is noted that trap spaces of a Boolean network are independent of the update scheme of this model~\cite{klarner2015computing}, \change{we provide in Corollary~\ref{cor:indep} another proof of this}.
 Then, we define a partial order \(<\) on \(\mathcal{S}_{\mathcal{N}}^{\star}\) as: \(m < m'\) if and only if \(\mathcal{S}_{\mathcal{N}}[m] \subseteq \mathcal{S}_{\mathcal{N}}[m']\) and \(\mathcal{S}_{\mathcal{N}}[m] \neq \mathcal{S}_{\mathcal{N}}[m']\).
@@
 
-\begin{corollary}
+\begin{corollary}\label{cor:indep}
   \change{Trap spaces of a Boolean network are independent of the update scheme.}
@@
 It might actually be possible to generalize our result to any 1-safe place-complementary \change{(i.e., places are defined by pairs such that the markings are complementary)} Petri net to define a notion of trap spaces that might be useful for the analysis of Petri nets, but this is out of the scope of the present article.
+\change{Note also that conversely, investigating static analyses on such 1-safe place-complementary nets might allow for a more efficient computation of their siphons and hence of trap spaces.}
 
@@
 The reason might be that researchers mainly focus on minimal generic siphons~\cite{DBLP:journals/isci/LiuB16} in the field of Petri nets.
+\change{While adapting those methods to obtain minimal conflict-free siphons would sometimes be possible, the switch from minimality to maximality is quite a leap.}
 Hence, we here propose several methods for computing maximal conflict-free siphons of a Petri net.
@@
 Second, surprisingly \texttt{bioLQM} cannot handle any model.
-One of the reason may be that the BDD characterizing all trap spaces is too large, and its computation is slow.
+One of the reason may be that the BDD characterizing all \change{generic} trap spaces is too large, and its computation is slow.
 In addition, having too many generic trap spaces before the filtering process may be also a reason.
@@
 They could replace our proposed methods if they give significantly better performance.
+\change{Making use of the specific structure (1-safe, place-complementary) might also reveal new techniques to be considered.}
 However, the current methods appear to already perform very well even on the biggest models we have considered.

### 2023-06-10 manuscripts/TCS2023/tcs2023.tex
Update scheme

@@
 
-A state \(v \in \mathbb{B}^{n}\) is as a mapping \(v \colon V \mapsto \mathbb{B}\) that assigns either 0 (inactive) or 1 (active) to each node.
+\change{A state \(s \in \mathbb{B}^{n}\) is as a mapping \(s \colon V \mapsto \mathbb{B}\) that assigns either 0 (inactive) or 1 (active) to each node.}
 We denote the set of all possible states of a Boolean network \(\mathcal{N}\) by \(\mathcal{S}_{\mathcal{N}} = \mathbb{B}^n\).
-At each time step \(t\), node \(v_i\) can, depending on the update scheme, update its state by
-\[v_i(t + 1) =
-  \begin{cases}
-    &f_i(v(t))\\
-    \text{or} &v_i(t)
-  \end{cases}
-\]
-where \(v(t)\) (resp. \(v_i(t)\)) is the state of \(\mathcal{N}\) (resp. the state of node \(v_i\)) at time \(t\).
-Note that for simplicity, we write \(f_i(v(t))\) even if \(\IN(v_i) \subsetneq V\) (i.e., \(\IN(v_i)\) does not contain some nodes of \(V\)).
-An update scheme of a Boolean network specifies which nodes update their states, as defined above, through time evolution~\cite{thomas1991regulatory}.
+\change{At each time step \(t\), node \(v_i\) can update its state by
+\[s'(v_i) = f_i(s)
+\]
+where \(s\) (resp. \(s'\)) is the state of \(\mathcal{N}\) at time \(t\) (resp. \(t + 1\)).
+Note that for simplicity, we write \(f_i(s)\) even if \(\IN(v_i) \subsetneq V\) (i.e., \(\IN(v_i)\) does not contain some nodes of \(V\)).
+An update scheme of a Boolean network specifies the way that the nodes update their states through time evolution~\cite{gershenson2004introduction}.
+There are many different update schemes, but the two main types~\cite{gershenson2004introduction} are: \emph{synchronous}, where all the nodes are \change{updated} simultaneously, and \emph{fully asynchronous}, where only one node is selected non-deterministically to be updated.}
 Following the update scheme, the Boolean network transits from a state to another state (possibly identical).
 This transition is called the \emph{state transition} and denoted by \(\rightarrow \subseteq \mathcal{S}_{\mathcal{N}} \times \mathcal{S}_{\mathcal{N}}\).
+\change{For example, under the synchronous update scheme, we have \(x \rightarrow y\) if and only if \(y(v_i) = f_i(x), \forall v_i \in V\), whereas under the fully asynchronous update scheme, we have \(x \rightarrow y\) if and only if there is a node \(v_i \in V\) such that \(y(v_i) = f_i(x)\) and \(y(v_j) = x(v_j), \forall v_j \in V, j \neq i\).}
 Then the dynamics of \(\mathcal{N}\) is captured by the directed graph \((\mathcal{S}_{\mathcal{N}}, \rightarrow)\) called the State Transition Graph (STG).
-There are many different update schemes, but the two main types~\cite{thomas1991regulatory} are: \emph{synchronous}, where all the nodes are \change{updated} simultaneously, and \emph{fully asynchronous}, where only one node is selected non-deterministically to be updated.
 

### 2023-06-11 manuscripts/TCS2023/tcs2023.tex
Attractor identification, phenotype control, reduction

@@
 Specifically, all minimal trap spaces of a Boolean network are mutually disjoint.
-This property is important because we can use it to approximate the set of attractors of the Boolean network under any update scheme~\cite{klarner2015computing} or to compute exactly the set of complex attractors of the Boolean network under the fully asynchronous update scheme~\cite{DBLP:conf/bcb/TrinhHB22}.
+\change{This property is important because it can benefit attractor identification of Boolean networks.
+Specifically, in~\cite{Klarner2015att}, the authors use random walks inside each minimal trap space to obtain approximations for attractors of a Boolean network under the fully asynchronous update scheme, then they use CTL model checking to verify the quality of the approximations.
+In~\cite{DBLP:conf/bcb/TrinhHB22}, the authors use the set of minimal trap spaces as a seed to speedup their previous attractor identification method that relies on feedback vertex sets and reachablity analysis.
+The soundness of the two above approaches comes from the separation of minimal trap spaces.
+}
 Note that it would be not difficult to obtain a direct proof on trap spaces for this property, which follows the same structure as the proof on siphons.
@@
 
-First, the work by~\cite{Rozum2021} uses the concept of \emph{stable motifs} to build the succession diagram of a Boolean network, a summary of the decisions in the network dynamics that lead to successively more restrictive nested stable motifs.
+\change{First, the work presented in~\cite{Rozum2021} uses the concept of \emph{stable motifs} to build the succession diagram of a Boolean network, a summary of the decisions in the network dynamics that lead to successively more restrictive nested stable motifs.
 The succession diagram is useful for control and decision making on this Boolean network.
 In particular, the proposed control methods are independent to the update scheme.
-It has been shown that a stable motif of a Boolean network is equivalent to a maximal trap space of this Boolean network~\cite{Rozum2021}.
+Note that, in~\cite{Rozum2021}, the succession diagram is also used to identify all attractors of a Boolean network under the fully asynchronous update scheme.
+It has been shown that a stable motif of a Boolean network is equivalent to a \emph{maximal trap space} of this Boolean network~\cite{Rozum2021}.
+Indeed, the computation of stable motif is a bottleneck of the methods proposed in~\cite{Rozum2021}.}
 Hence, it is necessary to develop an efficient method for computing maximal trap spaces of a Boolean network.
@@
 Third, we consider the trap spaces \emph{intersecting} a given subspace \(m^{*}\) of a Boolean network.
-Such trap spaces are used in the trap space-based control method~\cite{cifuentes2020control}.
-A trap space \(m\) intersects \(m^{*}\) if and only if \(S_{\mathcal{N}}[m] \cap S_{\mathcal{N}}[m^{*}] \neq \emptyset\).
+\change{Such trap spaces (along with minimal trap space) are used in the phenotype control method~\cite{cifuentes2020control}.
+This method uses the prime implicant-based method to compute trap spaces, which has been shown inefficient.
+Hence, having a more efficient method for computing such trap spaces can push the barrier previously existing in this control method.
+}A trap space \(m\) intersects \(m^{*}\) if and only if \(S_{\mathcal{N}}[m] \cap S_{\mathcal{N}}[m^{*}] \neq \emptyset\).
 It follows that for every \(v\), if \(m^{*}(v) = 0\) then \(m(v) = 0\) or \(m(v) = \star\), if \(m^{*}(v) = 1\) then \(m(v) = 1\) or \(m(v) = \star\).
@@
 
+\change{Another direction to speed up our approach in the side of Boolean networks is to apply reduction techniques to the original Boolean network.
+Many reduction techniques on Boolean networks~\cite{DBLP:journals/tcs/NaldiRTC11,schwieger2023reduction} have been proposed and some of them fully preserve attractors of a Boolean network under the fully asynchronous Boolean networks.
+In particular, a reduction technique on elimination of negatively auto-regulated nodes with respect to asynchronous attractors has recently been proposed~\cite{schwieger2023reduction}.
+However, there are two major issues needed to be considered.
+First, the question of whether these reduction techniques fully preserve minimal trap spaces of a Boolean network is still open.
+Second, although these reduction techniques can reduce the number of nodes, they can also increase the complexity of Boolean update functions~\cite{DBLP:journals/tcs/NaldiRTC11}, which is also an importan factor.
+It raises the question of whether they really simplify the computational burden of trap space computation.
+We will deeply investigate the two issues.
+Furthermore, we believe that the connection between trap spaces and siphons can be a very useful tool for addressing the first issue.
+}
+
 It is worth noting that there may be possibly other methods for computing minimal/maximal conflict-free siphons in Petri nets, like the methods for generic siphon computation in the field of Petri nets (see~\cite{DBLP:journals/isci/LiuB16} for a survey about these methods).

### 2023-06-11 manuscripts/TCS2023/tcs2023.tex
typos

@@
 It has been shown that a stable motif of a Boolean network is equivalent to a \emph{maximal trap space} of this Boolean network~\cite{Rozum2021}.
-Indeed, the computation of stable motif is a bottleneck of the methods proposed in~\cite{Rozum2021}.}
+Indeed, the computation of stable motifs is a bottleneck of the methods proposed in~\cite{Rozum2021}.}
 Hence, it is necessary to develop an efficient method for computing maximal trap spaces of a Boolean network.
@@
 Third, we consider the trap spaces \emph{intersecting} a given subspace \(m^{*}\) of a Boolean network.
-\change{Such trap spaces (along with minimal trap space) are used in the phenotype control method~\cite{cifuentes2020control}.
-This method uses the prime implicant-based method to compute trap spaces, which has been shown inefficient.
+\change{Such trap spaces (along with minimal trap spaces) are used in the phenotype control method~\cite{cifuentes2020control}.
+This method uses the prime implicant-based method~\cite{klarner2015computing,klarner2017pyboolnet} to compute trap spaces, which has been shown inefficient.
 Hence, having a more efficient method for computing such trap spaces can push the barrier previously existing in this control method.
@@
 \change{Another direction to speed up our approach in the side of Boolean networks is to apply reduction techniques to the original Boolean network.
-Many reduction techniques on Boolean networks~\cite{DBLP:journals/tcs/NaldiRTC11,schwieger2023reduction} have been proposed and some of them fully preserve attractors of a Boolean network under the fully asynchronous Boolean networks.
+Many reduction techniques on Boolean networks~\cite{DBLP:journals/tcs/NaldiRTC11,schwieger2023reduction} have been proposed and some of them fully preserve attractors of a Boolean network under the fully asynchronous update scheme.
 In particular, a reduction technique on elimination of negatively auto-regulated nodes with respect to asynchronous attractors has recently been proposed~\cite{schwieger2023reduction}.
@@
 First, the question of whether these reduction techniques fully preserve minimal trap spaces of a Boolean network is still open.
-Second, although these reduction techniques can reduce the number of nodes, they can also increase the complexity of Boolean update functions~\cite{DBLP:journals/tcs/NaldiRTC11}, which is also an importan factor.
+Second, although these reduction techniques can reduce the number of nodes, they can also increase the complexity of Boolean update functions~\cite{DBLP:journals/tcs/NaldiRTC11}, which is also an importan factor for the efficiency of computation methods.
 It raises the question of whether they really simplify the computational burden of trap space computation.

### 2023-06-11 manuscripts/TCS2023/tcs2023.tex
Notation of marking

@@
 \end{definition}
-A \emph{marking} for a Petri net is a mapping \(m : P \mapsto \mathbb{N}\) that assigns a number of tokens to each place.
-A place \(p\) is marked by a marking \(m\) if and only if \(m(p) > 0\).
-Marking \(m\) can be seen as a subset of \(P\) that contains all marked places by \(m\).
+A \emph{marking} for a Petri net is a mapping \change{\(M : P \mapsto \mathbb{N}\)} that assigns a number of tokens to each place.
+A place \(p\) is \change{\emph{marked}} by a marking \change{\(M\)} if and only if \change{\(M(p) > 0\)}.
 We shall write \(\pred(x)\) (resp.\ \(\succ(x)\)) to represent the set of vertices that have a (non-zero weighted) arc leading to (resp.\ coming from) \(x\).
 In this work, we consider a class of Petri nets called 1-safe Petri nets where every place has at most 1 token and all arcs are of weight 1.
-\change{Note that in such nets we have \(m : P \mapsto \{0,1\}\), we might therefore represent a marking by the equivalent set of places containing a token and will use this notation for simplicity.}
+\change{Note that in such nets we have \change{\(M : P \mapsto \{0,1\}\)}, we might therefore represent a marking by the equivalent set of places containing a token and will use this notation for simplicity.}
 In this case, weights are implicitly omitted in the arcs of a Petri net.
-Then, a transition \(t \in T\) is \emph{enabled} at a marking \(m\) if and only if \(\pred(t) \subseteq m\).
-A marking \(m\) is called a deadlock if there are no enabled transitions at \(m\).
-The firing of \(t\) leads to a new marking \(m'\) specified by \(m' = (m \backslash \pred(t)) \cup \succ(t)\).
+Then, a transition \(t \in T\) is \emph{enabled} at a marking \change{\(M\)} if and only if \change{\(\pred(t) \subseteq M\)}.
+A marking \change{\(M\)} is called a deadlock if there are no enabled transitions at \change{\(M\)}.
+The firing of \(t\) leads to a new marking \change{\(M'\)} specified by \change{\(M' = (M \backslash \pred(t)) \cup \succ(t)\)}.
 Note that when multiple transitions are enabled, we need to embed one firing scheme (similar to the update scheme of a Boolean network) to the Petri net.
@@
 \change{Conversely a transition is added from the active place to the inactive place for each conjunct of the negation of that function.}
-Let \(s\) be a state of the Boolean network and \(m_s\) be its corresponding marking in the encoded Petri net.
-It holds that \(\forall v \in V\), \(s(v) = 0\) if and only if \(m_s(\overline{p}_v) = 1\) \change{and \(m_s(p_v) = 0\)} and \(s(v) = 1\) if and only if \(m_s(p_v) = 1\) \change{and \(m_s(\overline{p}_v) = 0\)}.
-Note also that at any marking \(m\) of the Petri net encoding a Boolean network, it always holds that \(m(p_v) + m(\overline{p}_v) = 1\).
+Let \(s\) be a state of the Boolean network and \change{\(M_s\)} be its corresponding marking in the encoded Petri net.
+\change{It holds that \(\forall v \in V\), \(s(v) = 0\) if and only if \(M_s(\overline{p}_v) = 1\) \change{and \(M_s(p_v) = 0\)} and \(s(v) = 1\) if and only if \(M_s(p_v) = 1\) \change{and \(M_s(\overline{p}_v) = 0\)}.
+Note also that at any marking \(M\) of the Petri net encoding a Boolean network, it always holds that \(M(p_v) + M(\overline{p}_v) = 1\).}
 
@@
 For each node \(v\) of \(\mathcal{N}\), only transitions corresponding to \(v\) can change the current marking of \(p_v\) or \(\overline{p}_v\).
-In addition, at any marking at most one of such transitions is enabled because \(m(p_v) + m(\overline{p}_v) = 1\) holds.
+\change{In addition, at any marking at most one of such transitions is enabled because \(M(p_v) + M(\overline{p}_v) = 1\) holds.}
 Hence, for any update scheme in \(\mathcal{N}\), we have a corresponding firing scheme in \(\mathcal{P}\), which preserves the equivalence between the dynamics of \(\mathcal{N}\) and \(\mathcal{P}\)~\cite{DBLP:journals/nc/ChatainHKPT20}.
@@
 
-  We follow the following procedure to find a state \(s \in \mathcal{S}_{\mathcal{N}}[m]\) such that \(m_s(p') = 1, \forall p' \in \pred(t)\) where \(m_s\) is the corresponding marking in \(\mathcal{P}\) of \(s\).
+  We follow the following procedure to find a state \(s \in \mathcal{S}_{\mathcal{N}}[m]\) such that \change{\(M_s(p') = 1, \forall p' \in \pred(t)\)} where \change{\(M_s\)} is the corresponding marking in \(\mathcal{P}\) of \(s\).
   For every place \(p' \in \pred(t)\), let \(p''\) be the complementary place of \(p'\) and \(v'\) be the corresponding node in \(\mathcal{N}\) of \(p'\) and \(p''\).
 
-  If \(p'' \not \in S\), then \(v' \not \in D_m\) and we can always set \change{the} Boolean value to \(s(v')\) such that \(s \in \mathcal{S}_{\mathcal{N}}[m]\) and \(m_s(p') = 1\).
+  If \(p'' \not \in S\), then \(v' \not \in D_m\) and we can always set \change{the} Boolean value to \(s(v')\) such that \(s \in \mathcal{S}_{\mathcal{N}}[m]\) and \change{\(M_s(p') = 1\)}.
 
   If \(p'' \in S\), then \(v' \in D_m\) and we set \(s(v') = m(v')\).
-  In this case, if \(p' = p_{v'}\) then \(s(v') = m(v') = 1\) leading to \(m_s(p') = 1\), if \(p' = \overline{p}_{v'}\) then \(s(v') = m(v') = 0\) leading to \(m_s(p') = \change{0}\).
+  In this case, if \(p' = p_{v'}\) then \(s(v') = m(v') = 1\) leading to \change{\(M_s(p') = 1\)}, if \(p' = \overline{p}_{v'}\) then \(s(v') = m(v') = 0\) leading to \change{\(M_s(p') = \change{0}\)}.
 
   For the remaining nodes of \(\mathcal{N}\), we can always set Boolean values to these nodes to preserve that \(s \in \mathcal{S}_{\mathcal{N}}[m]\) \change{by applying the same procedure}.
-  We also have \(m_s(p_v) = 0\) by the characteristics of the encoding~\cite{chaouiya2004qualitative} \change{(and Definition \ref{def:mirror})}.
-  Now, \(t\) is enabled at marking \(m_s\).
-  Its firing leads to a new marking \(m'_s\) such that \(m'_s(p_v) = 1\) and \(m'_s(\overline{p}_v) = 0\).
-  Let \(s'\) be the corresponding state in \(\mathcal{N}\) of \(m'_s\).
-  We have \(s'(v) = 1\) because \(m'_s(p_v) = 1\) and \(m(v) = 0\) because \(p_v \in S\).
+  We also have \change{\(M_s(p_v) = 0\)} by the characteristics of the encoding~\cite{chaouiya2004qualitative} \change{(and Definition \ref{def:mirror})}.
+  Now, \(t\) is enabled at marking \change{\(M_s\)}.
+  Its firing leads to a new marking \change{\(M'_s\) such that \(M'_s(p_v) = 1\) and \(M'_s(\overline{p}_v) = 0\).}
+  Let \(s'\) be the corresponding state in \(\mathcal{N}\) of \change{\(M'_s\)}.
+  We have \(s'(v) = 1\) because \change{\(M'_s(p_v) = 1\) and \(m(v) = 0\) because \(p_v \in S\).}
   This implies that \(s' \not \in \mathcal{S}_{\mathcal{N}}[m]\).
@@
   By the definition of a mirror, \(m\) is a subspace of \(\mathcal{N}\).
-  Let \(s\) be an arbitrary state in \(\mathcal{S}_{\mathcal{N}}[m]\) and \(m_s\) be its corresponding marking in \(\mathcal{P}\).
-  Assume that there is a place \(p \in S\) such that \(m_s(p) = 1\).
+  Let \(s\) be an arbitrary state in \(\mathcal{S}_{\mathcal{N}}[m]\) and \change{\(M_s\)} be its corresponding marking in \(\mathcal{P}\).
+  Assume that there is a place \(p \in S\) such that \change{\(M_s(p) = 1\)}.
   Let \(v\) be the corresponding node in \(\mathcal{N}\) of \(p\).
   Since \(p \in S\), \(v \in D_m\) and \(m(v) = s(v)\).
-  If \(p = p_{v}\), then \(m_s(p_{v}) = 1\) leading to \(m(v) = s(v) = 1\) by the characteristics of the encoding~\cite{chaouiya2004qualitative}.
-  By the definition of a mirror, \(m(v) = 0\) because \(p_{v} \in S\), \change{meaning that \(m_s(p_{v}) = 0\)}, which is a contradiction.
+  If \(p = p_{v}\), then \change{\(M_s(p_{v}) = 1\)} leading to \(m(v) = s(v) = 1\) by the characteristics of the encoding~\cite{chaouiya2004qualitative}.
+  By the definition of a mirror, \(m(v) = 0\) because \(p_{v} \in S\), \change{meaning that \(M_s(p_{v}) = 0\)}, which is a contradiction.
 
   It is symmetric for the case that \(p = \overline{p}_{v}\).
-  Hence, \(m_s(p) = 0, \forall p \in S\).
-  In any marking \(m'_s\) reachable from \(m_s\) regardless of the firing scheme of \(\mathcal{P}\), we have \(m'_s(p) = 0, \forall p \in S\) by the dynamical property on markings of a siphon~\cite{DBLP:journals/isci/LiuB16}.
-  Let \(s'\) be the corresponding state in \(\mathcal{N}\) of \(m'_s\).
+  Hence, \change{\(M_s(p) = 0, \forall p \in S\)}.
+  In any marking \change{\(M'_s\) reachable from \(M_s\) regardless of the firing scheme of \(\mathcal{P}\), we have \(M'_s(p) = 0, \forall p \in S\) by the dynamical property on markings of a siphon~\cite{DBLP:journals/isci/LiuB16}.}
+  Let \(s'\) be the corresponding state in \(\mathcal{N}\) of \change{\(M'_s\)}.
   For every node \(v \in D_m\), we have all two cases as follows.
-  Case 1: \(p_v \in S\), then \(m'_s(p_v) = 0\), thus \(s'(v) = 0 = m(v)\).
-  Case 2: \(\overline{p}_v \in S\), then \(m'_s(\overline{p}_v) = 0\), thus \(s'(v) = 1 = m(v)\).
+  Case 1: \(p_v \in S\), then \change{\(M'_s(p_v) = 0\)}, thus \(s'(v) = 0 = m(v)\).
+  Case 2: \(\overline{p}_v \in S\), then \change{\(M'_s(\overline{p}_v) = 0\)}, thus \(s'(v) = 1 = m(v)\).
   Hence, \(s'(v) = m(v)\) for every \(v \in D_m\).

### 2023-06-11 manuscripts/TCS2023/tcs2023.tex
Structure-based approaches

@@
 Although these approaches do not directly support the minimal/maximal conflict-free siphon computation now, we plan to investigate them in the future.
-They could replace our proposed methods if they give significantly better performance.
-\change{Making use of the specific structure (1-safe, place-complementary) might also reveal new techniques to be considered.}
+\change{
+In particular, several approaches based on the network structure at the Petri net level (e.g., the decomposition approaches~\cite{DBLP:journals/tsmc/CordoneFP05,DBLP:journals/ieeejas/YouKW22} for identifying minimal generic siphons) can be adapted to help the identification of minimal conflict-free siphons.
+Making use of the specific structure (1-safe, place-complementary) might also reveal new techniques to be considered.
+It is potentially possible because in the field of Petri nets, most of the methods for identifying minimal generic siphons focus on various net classes with special structures~\cite{DBLP:journals/isci/LiuB16}.
+The above potential approaches could replace our proposed methods if they give significantly better performance.}
 However, the current methods appear to already perform very well even on the biggest models we have considered.

### 2023-06-12 manuscripts/TCS2023/tcs2023.tex
Complexity

@@
 
+\change{First, we discuss the complexity of siphon computation in Petri nets.
+Siphons are a prominent concept in the field of Petri nets, but unfortunately there are very few studies focusing on the complexity aspect.
+In this field, researchers mainly focus on practical methods for computing minimal generic siphons (also many related types) in general or special Petri nets and the applications of such types to the control of real-world systems modeled by Petri nets~\cite{DBLP:journals/isci/LiuB16}.
+The problem of finding a minimal siphon of a 1-safe Petri net is solvable in polynomial time~\cite{tanimoto1996finding}.
+Clearly, the problem of finding a siphon of a 1-safe Petri net is also solvable in polynomial time.
+However, the problem of computing all (minimal) siphons is not easier than the problem of computing a (minimal) siphon but its complexity still not clear.
+Note that the number of siphons (even minimal siphons) can be exponential in the number of places of the Petri net~\cite{DBLP:journals/isci/LiuB16}.
+Moreover, there is no complexity result for the case of maximal siphons.
+Regarding the conflict-free siphons, we believe that the polynomial algorithm for computing a minimal generic siphon presented in~\cite{tanimoto1996finding} can be adapted to find a (minimal) conflict-free siphon.
+This is not in contrast to the NP-hardness of some problems on trap spaces in Boolean networks~\cite{moon2023computational} because in general the number of transitions of the Petri net encoding of a Boolean network can be exponential in the number of nodes of this Boolean network.
+However, again the complexity of the problem of computing all (minimal/maximal) conflict-free siphons is still open.}
+
 \subsection{Characterization}%
@@
 
-First, we show the characterization of all conflict-free siphons of the encoded Petri net \(\mathcal{P} = (P, T, W)\).
+\change{We here show} the characterization of all conflict-free siphons of the encoded Petri net \(\mathcal{P} = (P, T, W)\).
 Suppose that \(S\) is a generic siphon of \(\mathcal{P}\).
@@
 First, the question of whether these reduction techniques fully preserve minimal trap spaces of a Boolean network is still open.
-Second, although these reduction techniques can reduce the number of nodes, they can also increase the complexity of Boolean update functions~\cite{DBLP:journals/tcs/NaldiRTC11}, which is also an importan factor for the efficiency of computation methods.
+Second, although these reduction techniques can reduce the number of nodes, they can also increase the complexity of Boolean update functions~\cite{DBLP:journals/tcs/NaldiRTC11}, which is also an importan factor for the performance of computation methods.
 It raises the question of whether they really simplify the computational burden of trap space computation.

### 2023-06-13 manuscripts/TCS2023/tcs2023.tex
Topology
Finished revising

@@
 For example, let us consider the Boolean network shown in Example~\ref{example:BN}.
-Figure~\ref{fig:stg} shows the dynamics of this model under the fully asynchronous update \change{scheme} (i.e., only one node is updated at each time step).
+\change{Figure~\ref{fig:stg}} shows the dynamics of this model under the fully asynchronous update \change{scheme} (i.e., only one node is updated at each time step).
 The model has all two trap spaces, \(m_1 = 11\) and \(m_2 = \star\star\).
@@
 \centering
-\begin{subfigure}[b]{0.3\textwidth}
+\begin{subfigure}[b]{0.25\textwidth}
+  \centering
+  \begin{tikzpicture}
+    \node (1) {\(x_{1}\)};
+    \node (2) [right=of 1] {\(x_{2}\)};
+
+    \draw[->] (1) edge [bend left=45] (2);
+    \draw[->] (2) edge [bend left=45] (1);
+    \draw[->] (1) to [loop left] (1);
+    \draw[->] (2) to [loop right] (2);
+  \end{tikzpicture}
+  \caption{\change{Influence graph of the model. Each arc is actually both positive and negative.}}
+\end{subfigure}\hfill%
+\begin{subfigure}[b]{0.25\textwidth}
 \centering
@@
 \end{subfigure}\hfill%
-\begin{subfigure}[b]{0.6\textwidth}
+\begin{subfigure}[b]{0.4\textwidth}
 \centering
@@
 \label{fig:PN}
-\end{subfigure}%
-\caption{Dynamics and encoding of the Boolean network of Example~\ref{example:BN}.}%
+\end{subfigure}\hfill%
+\caption{\change{Influence graph, dynamics, and Petri net encoding of the Boolean network of Example~\ref{example:BN}}.}%
 \label{fig:stg_and_PN}

### 2023-06-16 manuscripts/TCS2023/tcs2023.tex
Reference for BBM repository

@@
 
-\change{The research group behind the \texttt{BBM} repository has recently undertaken considerable effort for building a collection of real-world Boolean models from various sources used in systems biology.}
+\change{The research group behind the \texttt{BBM} repository~\cite{pastva2023repository} has recently undertaken considerable effort for building a collection of real-world Boolean models from various sources used in systems biology.}
 It aims to be a comprehensive collection suitable for benchmarking and testing new tools and methods.

### 2023-07-11 manuscripts/TCS2023/tcs2023.tex
typo

@@
 First, the question of whether these reduction techniques fully preserve minimal trap spaces of a Boolean network is still open.
-Second, although these reduction techniques can reduce the number of nodes, they can also increase the complexity of Boolean update functions~\cite{DBLP:journals/tcs/NaldiRTC11}, which is also an importan factor for the performance of computation methods.
+Second, although these reduction techniques can reduce the number of nodes, they can also increase the complexity of Boolean update functions~\cite{DBLP:journals/tcs/NaldiRTC11}, which is also an important factor for the performance of computation methods.
 It raises the question of whether they really simplify the computational burden of trap space computation.
